### 2016-06-13 paper/structureScene.tex
initial commit

\documentclass[journal]{IEEEtran}
\usepackage{cite}
% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
%
\usepackage[cmex10]{amsmath}

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
\usepackage{fixltx2e}

\usepackage{stfloats}
% LaTeX2e). It also provides a command:
\fnbelowfloat

\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\usepackage{url}
\usepackage{color}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{color}
\newcommand{\gl}[1]{\textcolor{red}{Gregoire : #1}}
\newcommand{\mr}[1]{\textcolor{magenta}{Mathias : #1}}
\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu : #1}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
%\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
\title{Unsupervised Acoustic Scene Micro-Segmentation for High Level (Semantic ?) Attributes Prediction}
%
%

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This paper introduces a 
\end{abstract}

% Note that keywords are not normally used for peerreview paper.
\begin{IEEEkeywords}
Acoustic scene classification, segmentation.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}

\IEEEPARstart{O}{ver} the past decades, the amount of audio data recorded from our sonic environment has considerably grown. 1. Recent research areas such as eco-acoustics \cite{ECOACOUSTICS2014, krause} start to consider massive deployment of acoustic sensors around the world in order to measure potential animal biodiversity modification over large temporal scales due to human activity or climate change \cite{NessSST13, stowell13a, stowell13b}.

2. Recent regulations -> pleasantness

We believe that those matters have strong societal impacts, but those fields of research are still in infancy and consequently very few well built dataset are available for evaluation purposes.

A relatively more mature field of investigation is the classification

despite its lower value, it has the advantage of being rooted by numerous works in cognitive psychology and been considered in the data processing field for a while, with availability of some well designed evaluation datasets.

In this paper, we investigate the potential of a new unsupervised segmentation paradigm that allows us to

This data has to be processed

Environment sound

describe in terms of high level attributes

be they cognitive ones such as pleasantness, or descriptive ones such as the type of environment where the

%As part of the aforementioned research areas and applications, the emerging field of \emph{Acoustic Scene Analysis} (also called \emph{Sound Scene Analysis}) \cite{Stowell15} aims to develop approaches and systems for the automatic analysis of environmental sounds and soundscapes (originating both from urban or nature environments). While research methodologies in related fields such as Automatic Speech Recognition (ASR) \cite{Rabiner93} and Music Information Retrieval (MIR) \cite{Muller07} are now well established, research addressing Acoustic Scene Analysis remains relatively young.

Whilst the range of applications are very large, so is the need

holistic / event based

skeleton of events on a bed of texture.

debate about the kind of process underlying perception

strong indication that the detection of specific events, named markers suffice to trigger class prediction

From a numerical data processing overview, the holistic scheme as a simplicity, but clearly face poor performance on realistic conditions \cite{jasa-el}.

If available, the complete description of the scene in terms of event occurences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas.

\section{Overview}

\section{Micro Structure}

\subsection{In Music}

\subsection{In Environmental Scenes}

\section{Pleasantness Prediction}

\section{Environnment Prediction}

\section{Discussion}

\section{Conclusion}

\bibliographystyle{unsrt}
\bibliography{biblio}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Mathieu Lagrange}
%Biography text here.
%\end{IEEEbiography}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}{John Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

%\begin{IEEEbiographynophoto}{Jane Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2016-06-21 paper/structureScene.tex
first structure
populating bib

@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Unsupervised Acoustic Scene Micro-Segmentation for High Level (Semantic ?) Attributes Prediction}
-%
-%
+\title{Acoustic Scene Classification: \\ Supervised and Unsupervised Approaches using the Scattering Transform}
+
+\author{Vincent Lostanlen, Gr\'egoire Lafay, Joakim Anden and Mathieu Lagrange}
 
@@
 \IEEEPARstart{O}{ver} the past decades, the amount of audio data recorded from our sonic environment has considerably grown.
-1.
 Recent research areas such as eco-acoustics \cite{ECOACOUSTICS2014, krause} start to consider massive deployment of acoustic sensors around the world in order to measure potential animal biodiversity modification over large temporal scales due to human activity or climate change \cite{NessSST13, stowell13a, stowell13b}.
 
-2. Recent regulations -> pleasantness
+Also, some recent works demonstrate the interest of such approaches for the characterization of human acoustic pleasantness in urban areas \cite{lafay, lavandier} and the prediction of annoyance due to the traffic \cite{gloaguen}.
+We believe that those case study are of interest for the signal processing community as they have strong societal impacts and raises interesting research avenues.
+Unfortunately, those fields of research are still in infancy and consequently very few well built datasets are available for evaluation purposes.
 
-We believe that those matters have strong societal impacts, but those fields of research are still in infancy and consequently very few well built dataset are available for evaluation purposes.
-
-A relatively more mature field of investigation is the classification
-
-despite its lower value, it has the advantage of being rooted by numerous works in cognitive psychology and been considered in the data processing field for a while, with availability of some well designed evaluation datasets.
+A closely related field of investigation that is relatively more mature is the classification of acoustic scenes whose aim is to predict labels by processing audio data.
+Despite its lower value, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois} and been considered in the data processing field for a while, with availability of some well designed evaluation datasets.
 
@@
 
-From a numerical data processing overview, the holistic scheme as a simplicity, but clearly face poor performance on realistic conditions \cite{jasa-el}.
+From a numerical data processing overview, the holistic scheme as a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
 
@@
 
-\section{Overview}
+\section{Previous work}
 
-\section{Micro Structure}
+GL 1
 
-\subsection{In Music}
+ML 2
 
-\subsection{In Environmental Scenes}
+\subsection{Goal}
 
-\section{Pleasantness Prediction}
+\subsection{Methods}
 
-\section{Environnment Prediction}
+\subsection{Datasets}
+
+defreville \cite{aucouturier2007bag}
+
+\cite{lagrange:hal-01082501}
+
+dcase2013 \cite{7100934} \cite{giannoulis2013database}
+
+For each scene type, three different recordists (DG, DS,
+EB) visited a wide variety of locations in Greater London over
+a period of months (Summer and Autumn 2012), and in each
+scene recorded a few minutes of audio. We ensured that no
+systematic variations in the recordings covaried with scene
+type: all recordings were made in moderate weather condi-
+tions, and varying times of day and week, and each recordist
+recorded each scene type.
+
+Rouen \cite{rakotomamonjy2015histogram}
+
+\url{https://sites.google.com/site/alainrakotomamonjy/home/audio-scene}
+
+dCase2016 \cite{Mesaros2016_EUSIPCO}
+
+advantages of dcase2013:
+
+controled intra class diversity
+
+despite its size, it is still challenging as state-of-the-art systems achieves 76 \% \cite{6701890} (winner of the 2013 challenge RQA+SVM).
+For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively.
+On the latter, the state-of-the-art is currently 95 \% \cite{bisot2016acoustic}.
+
+\section{Scattering}
+
+VL
+
+\section{Unsupervised approach}
+
+GL
+
+\subsection{Results}
+
+GL
+
+\section{Supervised approach}
+
+VL
+
+\subsection{Results}
+
+VL
 
@@
 
+ML
+
 \section{Conclusion}
+
+ML
 

### 2016-06-21 paper/structureScene.tex
missing \fi to compile
write first paragraph of section 3

@@
   \usepackage[pdftex]{graphicx}
+\fi
 %
@@
 
-\section{Scattering}
+The scattering transform is a multiscale representation of audio signals which cascades an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
+The modulation filterbank may either be performed solely on the time variable or on both time and frequency.
+In this section, we present the design of both the temporal scattering transform and the time-frequency scattering transform.
 

### 2016-06-21 paper/structureScene.tex
write on the difference between scalogram and CQT

@@
 VL
+The wavelet transform $\boldsymbol{}$ of an audio signal
+$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets:
+$\boldsymbol{y_1}(t, \gamma_1) =
+(\boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}})(t)$.
+Applying pointwise complex modulus to $\boldsymbol{y_1}$ yields
+the wavelet scalogram
+$\boldsymbol{x_1}(t, \gamma_1) = \vert \boldsymbol{y_1}(t, \gamma_1)\vert$.
+The wavelet scalogram bears resemblance with the constant-Q transform (CQT),
+which is derived from the short-time Fourier transform by averaging the frequency
+axis into constant-Q subbands of center frequencies $2^{-\gamma/Q}$.
+Like the CQT, the wavelet scalogram is
+a time-frequency representation indexed by time $t$ and log-frequency $\gamma_1$.
+However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
+theoretical limit of optimal time-frequency localization across the whole
+frequency range, whereas the CQT has a fixed temporal resolution.
+Therefore, the wavelet scalogram has a better temporal localization at high
+frequencies, which allows to observe steady amplitude modulations.
 

### 2016-06-21 paper/structureScene.tex
call section 3 "scattering transforms"

@@
 
+\section{Scattering transforms}
 The scattering transform is a multiscale representation of audio signals which cascades an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
-The modulation filterbank may either be performed solely on the time variable or on both time and frequency.
-In this section, we present the design of both the temporal scattering transform and the time-frequency scattering transform.
+Depending upon the chosen architecture, the modulation filterbank may either be performed solely over the time variable or on both time and frequency.
+In this section, we present both the temporal scattering transform and the time-frequency scattering transform.
 

### 2016-06-21 paper/structureScene.tex
write on invariance to translation

@@
 \section{Scattering transforms}
-The scattering transform is a multiscale representation of audio signals which cascades an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
+Scattering transforms are a multiscale representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
 Depending upon the chosen architecture, the modulation filterbank may either be performed solely over the time variable or on both time and frequency.
 In this section, we present both the temporal scattering transform and the time-frequency scattering transform.
+
+The notion of invariance to translation plays an essential role in auditory scene classification.
+Indeed, the starting time of the recordings are chosen arbitrarily, and thus do not convey any information about the class.
+To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space prior to training a classifier.
 

### 2016-06-21 paper/structureScene.tex
write on S_1 x

@@
 Indeed, the starting time of the recordings are chosen arbitrarily, and thus do not convey any information about the class.
-To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space prior to training a classifier.
+To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space prior to training the classifier.
+From any set of descriptors, \eg a time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$,
+a translation-invariant representation up to time skills of duration $T$ can be simply obtained
+by concolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
+set to $1/T$:
+\begin{equation}
+\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \phi) (t).
+\end{equation}
 

### 2016-06-21 paper/structureScene.tex
usepackage xspace to define \eg
subsection titles
more on section III.A

@@
 \usepackage{color}
+\usepackage{xspace}
+\newcommand*{\eg}{e.g.\@\xspace}
 
@@
 
+\subsection{Invariance and stability in acoustic scene classification}
 The notion of invariance to translation plays an essential role in auditory scene classification.
@@
 To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space prior to training the classifier.
-From any set of descriptors, \eg a time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$,
+From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ of  a signal $\boldsymbol{x}(t)$ --
 a translation-invariant representation up to time skills of duration $T$ can be simply obtained
-by concolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
+by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
 set to $1/T$:
@@
 \end{equation}
+As a downside, the transient information in $\boldsymbol{x_1}$ at finer scales than $T$ are lost by this low-pass filtering, hence a lack of discriminability in feature space.
+To address this issue, the temporal scattering transform is to recover this information by convolving $x_1$ with wavelets whose center frequencies are above $1/T$, and subsequently applying complex modulus.
+\subsection{Temporal scattering}
 
@@
 
+\subsection{Time-frequency scattering}
 \section{Unsupervised approach}

### 2016-06-21 paper/structureScene.tex
rephrase scalogram vs CQT

@@
 
-VL
 The wavelet transform $\boldsymbol{}$ of an audio signal
@@
 which is derived from the short-time Fourier transform by averaging the frequency
-axis into constant-Q subbands of center frequencies $2^{-\gamma/Q}$.
-Like the CQT, the wavelet scalogram is
-a time-frequency representation indexed by time $t$ and log-frequency $\gamma_1$.
+axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}$.
+Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
 However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
@@
 \subsection{Time-frequency scattering}
+
 \section{Unsupervised approach}

### 2016-06-21 paper/structureScene.tex
define wavelet filter bank

@@
 \subsection{Temporal scattering}
+Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
+dimensionless frequency $1$ and bandwidth $1/Q$.
+A filter bank of wavelets is built by dilating $\boldsymbol{\psi}$
+according to a geometric sequence of scales $2^{-\gamma_1/Q}$,
+where the variable $\gamma$ takes integer values.
+\begin{equation}
+\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q} \boldsymbol{\psi}(-2^{\gamma_1/Q} t)
+\end{equation}
+We denote by $\boldsymbol{\psi_{\gamma_1}}(t)$ the resulting wavelets.
 

### 2016-06-21 paper/structureScene.tex
mention fourth-order Gammatone
Anden -> Andén + Oxford comma in author list
write on stability

@@
 
-\author{Vincent Lostanlen, Gr\'egoire Lafay, Joakim Anden and Mathieu Lagrange}
+\author{Vincent Lostanlen, Gr\'egoire Lafay, Joakim And\'en, and Mathieu Lagrange}
 
@@
 As a downside, the transient information in $\boldsymbol{x_1}$ at finer scales than $T$ are lost by this low-pass filtering, hence a lack of discriminability in feature space.
-To address this issue, the temporal scattering transform is to recover this information by convolving $x_1$ with wavelets whose center frequencies are above $1/T$, and subsequently applying complex modulus.
+To address this issue, the temporal scattering transform recovers this information by convolving $x_1$ with wavelets whose center frequencies are above $1/T$ and subsequently applying complex modulus.
+
+By resorting to wavelet modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to small time warps,
+in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2010}.
+Besides invariance to translation, this stability property is crucial in signal classification, as it entails a guarantee of robustness.
+
 \subsection{Temporal scattering}
@@
 The wavelet scalogram bears resemblance with the constant-Q transform (CQT),
-which is derived from the short-time Fourier transform by averaging the frequency
+which is derived from the short-time Fourier transform (STFT) by averaging the frequency
 axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}$.
@@
 theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the CQT has a fixed temporal resolution.
+frequency range, whereas the CQT has a fixed temporal resolution, constrained by the support of the STFT analyzing window.
 Therefore, the wavelet scalogram has a better temporal localization at high
 frequencies, which allows to observe steady amplitude modulations.
+
+In all subsequent experiments,
+$\boldsymbol{\psi}$ is designed as a fourth-order Gammatone
+wavelet of quality factor $Q = 8$.
 

### 2016-06-21 paper/structureScene.tex
define \ie
minor changes at the beginning of Section III

@@
 \newcommand*{\eg}{e.g.\@\xspace}
+\newcommand*{\ie}{i.e.\@\xspace}
 
@@
 \section{Scattering transforms}
-Scattering transforms are a multiscale representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
+Scattering transforms are translation-invariant representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
 Depending upon the chosen architecture, the modulation filterbank may either be performed solely over the time variable or on both time and frequency.
@@
 \begin{equation}
-\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \phi) (t).
+\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \boldsymbol{\phi}) (t).
 \end{equation}

### 2016-06-21 paper/structureScene.tex
more on stability

@@
 in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2010}.
-Besides invariance to translation, this stability property is crucial in signal classification, as it entails a guarantee of robustness.
+Besides invariance to translation, this stability property is crucial in signal classification, as it entails a guarantee of robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.
 

### 2016-06-21 paper/structureScene.tex
write on the number of layers and the required invariance

@@
 
+The scattering transform is defined mathematically as an infinite cascade of nonlinear layers, each consisting of a wavelet transform modulus operator.
+However, to achieve invariance to translation up to $T = 186\,\mathrm{ms}$, \ie of the order of the minimal duration between non-overlapping acoustic events, however, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ suffice.
+The next section describes the operations involved in temporal scattering.
+
 \subsection{Temporal scattering}

### 2016-06-21 paper/structureScene.tex
rephrase end of III.A

--- 2016-06-21 paper/structureScene.tex
@@
 
+The scattering transform is defined mathematically as an infinite cascade of nonlinear layers, each consisting of a wavelet transform modulus operator.
+In practice, to achieve invariance to translation up to $T = 186\,\mathrm{ms}$, \ie of the order of the minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ suffice.
+The next section describes the operations involved in temporal scattering.
+
 \subsection{Temporal scattering}

### 2016-06-21 paper/structureScene.tex
more on wavelets

@@
 Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
-dimensionless frequency $1$ and bandwidth $1/Q$.
-A filter bank of wavelets is built by dilating $\boldsymbol{\psi}$
-according to a geometric sequence of scales $2^{-\gamma_1/Q}$,
-where the variable $\gamma$ takes integer values.
+dimensionless frequency $1$ and bandwidth $1/Q_1$.
+A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
+according to a geometric sequence of temporal scales $2^{-\gamma_1/Q_1}$.
+The variable $\gamma_1$, akin to a log-frequency, takes integer values between $0$ and $(J_1 \times Q_1 - 1)$.
+In the sequel, the number of filters per octave $Q_1$ is set to $8$, and
+the number of octaves $J_1$ is set to $10$, \ie the full audible range.
+We denote by
 \begin{equation}
-\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q} \boldsymbol{\psi}(-2^{\gamma_1/Q} t)
+\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(-2^{\gamma_1/Q_1} t)
 \end{equation}
-We denote by $\boldsymbol{\psi_{\gamma_1}}(t)$ the resulting wavelets.
+the wavelets resulting from the dilation of $\boldsymbol{\psi}$(t).
+For some $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
+has a center frequency of $2^{-\gamma_1/Q_1}$, a bandwidth of $2^{-\gamma_1/Q_1}/Q_1$, and a quality factor of $Q_1$.
 

### 2016-06-21 paper/structureScene.tex
display definition of x_1

@@
 the wavelet scalogram
-$\boldsymbol{x_1}(t, \gamma_1) = \vert \boldsymbol{y_1}(t, \gamma_1)\vert$.
+\begin{equation}
+\boldsymbol{x_1}(t, \gamma_1)
+= \vert \boldsymbol{y_1}(t, \gamma_1)\vert
+= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t).
+\end{equation}
 The wavelet scalogram bears resemblance with the constant-Q transform (CQT),

### 2016-06-22 paper/structureScene.tex
minor improvements to section 3

@@
 In the sequel, the number of filters per octave $Q_1$ is set to $8$, and
-the number of octaves $J_1$ is set to $10$, \ie the full audible range.
+the number of octaves $J_1$ is set to $10$, \ie the maximum range of human hearing.
 We denote by
@@
 \end{equation}
-The wavelet scalogram bears resemblance with the constant-Q transform (CQT),
+The wavelet scalogram bears a resemblance to the constant-Q transform (CQT),
 which is derived from the short-time Fourier transform (STFT) by averaging the frequency
@@
 
-In all subsequent experiments,
-$\boldsymbol{\psi}$ is designed as a fourth-order Gammatone
-wavelet of quality factor $Q = 8$.
+In all subsequent experiments, wavelets
+$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as a fourth-order Gammatone
+wavelet with one vanishing moment \cite{Venkitaraman2014}.
 
@@
 \section{Supervised approach}
-
-VL
 

### 2016-06-22 paper/structureScene.tex
bis

@@
 A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
-according to a geometric sequence of temporal scales $2^{-\gamma_1/Q_1}$.
+according to a geometric sequence of temporal scales $2^{\gamma_1/Q_1}$.
 The variable $\gamma_1$, akin to a log-frequency, takes integer values between $0$ and $(J_1 \times Q_1 - 1)$.
-In the sequel, the number of filters per octave $Q_1$ is set to $8$, and
-the number of octaves $J_1$ is set to $10$, \ie the maximum range of human hearing.
+In the sequel, we set the number of wavelets per octave $Q_1$ to $8$, and
+the number of octaves $J_1$ to $10$, \ie the maximum range of human hearing.
 We denote by
 \begin{equation}
-\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(-2^{\gamma_1/Q_1} t)
+\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(2^{-\gamma_1/Q_1} t)
 \end{equation}
@@
 \end{equation}
-The wavelet scalogram bears a resemblance to the constant-Q transform (CQT),
-which is derived from the short-time Fourier transform (STFT) by averaging the frequency
+The wavelet scalogram bears resemblance to the constant-Q transform (CQT),
+which is derived from the short-term Fourier transform (STFT) by averaging the frequency
 axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}$.

### 2016-06-22 paper/structureScene.tex
specify a center frequency in Hertz for psi

@@
 Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
-dimensionless frequency $1$ and bandwidth $1/Q_1$.
+center frequency $\xi$ and bandwidth $\xi/Q_1$.
 A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
-according to a geometric sequence of temporal scales $2^{\gamma_1/Q_1}$.
+according to a geometric sequence of scales $2^{\gamma_1/Q_1}$.
 The variable $\gamma_1$, akin to a log-frequency, takes integer values between $0$ and $(J_1 \times Q_1 - 1)$.
-In the sequel, we set the number of wavelets per octave $Q_1$ to $8$, and
-the number of octaves $J_1$ to $10$, \ie the maximum range of human hearing.
+In the sequel, we set $\xi$ to 20 kHz, \ie close to the Nyquist frequency of audio recordings ; the number of octaves $J_1$ to $10$, \ie the maximum range of human hearing ; and the number of wavelets per octave $Q_1$ to $8$.
 We denote by
@@
 For some $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
-has a center frequency of $2^{-\gamma_1/Q_1}$, a bandwidth of $2^{-\gamma_1/Q_1}/Q_1$, and a quality factor of $Q_1$.
+has a center frequency of $2^{-\gamma_1/Q_1}\xi$, a bandwidth of $2^{-\gamma_1/Q_1}\xi/Q_1$, and a quality factor of $Q_1$.
 
@@
 which is derived from the short-term Fourier transform (STFT) by averaging the frequency
-axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}$.
+axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}\xi$.
 Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.

### 2016-06-22 paper/structureScene.tex
minimal amplitude modulation scale ~ 1ms

@@
 Therefore, the wavelet scalogram has a better temporal localization at high
-frequencies, which allows to observe steady amplitude modulations.
+frequencies than the CQT, at the expense of a greater computational cost.
+This property allows to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi$ for $\gamma = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi$.
 

### 2016-06-22 paper/structureScene.tex
start writing on amplitude modulations
\xi -> \xi_1

@@
 Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
-center frequency $\xi$ and bandwidth $\xi/Q_1$.
+center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$.
 A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
@@
 The variable $\gamma_1$, akin to a log-frequency, takes integer values between $0$ and $(J_1 \times Q_1 - 1)$.
-In the sequel, we set $\xi$ to 20 kHz, \ie close to the Nyquist frequency of audio recordings ; the number of octaves $J_1$ to $10$, \ie the maximum range of human hearing ; and the number of wavelets per octave $Q_1$ to $8$.
+In the sequel, we set $\xi_1$ to 20 kHz, \ie close to the Nyquist frequency of audio recordings ; the number of octaves $J_1$ to $10$, \ie the maximum range of human hearing ; and the number of wavelets per octave $Q_1$ to $8$.
 We denote by
@@
 For some $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
-has a center frequency of $2^{-\gamma_1/Q_1}\xi$, a bandwidth of $2^{-\gamma_1/Q_1}\xi/Q_1$, and a quality factor of $Q_1$.
+has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$, a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$, and a quality factor of $Q_1$.
 
@@
 which is derived from the short-term Fourier transform (STFT) by averaging the frequency
-axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}\xi$.
+axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}\xi_1$.
 Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
@@
 frequencies than the CQT, at the expense of a greater computational cost.
-This property allows to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi$ for $\gamma = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi$.
+This property allows to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+
+Among auditory scenes, amplitude modulations may be caused by a broad variety of mechanical interactions, including collision, friction, and turbulent flow.
+They also account for some higher-level cognitive attributes of sound, \eg prosody in speech or rhythm in music.
 

### 2016-06-22 paper/structureScene.tex
title of section III = "Scattering representations"

@@
 
-\section{Scattering transforms}
+\section{Scattering representations}
 Scattering transforms are translation-invariant representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.

### 2016-06-22 paper/structureScene.tex
definition of y_1 is unnecessary.

@@
 As a downside, the transient information in $\boldsymbol{x_1}$ at finer scales than $T$ are lost by this low-pass filtering, hence a lack of discriminability in feature space.
-To address this issue, the temporal scattering transform recovers this information by convolving $x_1$ with wavelets whose center frequencies are above $1/T$ and subsequently applying complex modulus.
+To address this issue, the temporal scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$, and subsequently applying complex modulus.
 
@@
 The wavelet transform $\boldsymbol{}$ of an audio signal
-$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets:
-$\boldsymbol{y_1}(t, \gamma_1) =
-(\boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}})(t)$.
+$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets.
 Applying pointwise complex modulus to $\boldsymbol{y_1}$ yields
@@
 \boldsymbol{x_1}(t, \gamma_1)
-= \vert \boldsymbol{y_1}(t, \gamma_1)\vert
 = \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t).

### 2016-06-22 paper/structureScene.tex
define x_2

@@
 They also account for some higher-level cognitive attributes of sound, \eg prosody in speech or rhythm in music.
+Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a translation-invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator:
+\begin{equation}
+\boldsymbol{x_2}(t,\gamma_1,\gamma_2) =
+\vert \boldsymbol{x_1} \ast \boldsymbol{\psi_{\gamma_2}} \vert(t,\gamma_1),
+\end{equation}
+where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order log-frequency index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
 

### 2016-06-22 paper/structureScene.tex
define S_2 x

--- 2016-06-22 paper/structureScene.tex
@@
 They also account for some higher-level cognitive attributes of sound, \eg prosody in speech or rhythm in music.
-
+Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a translation-invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
+The amplitude modulation spectrum resulting from this operator is
+\begin{equation}
+\boldsymbol{x_2}(t,\gamma_1,\gamma_2) =
+\vert \boldsymbol{x_1} \ast \boldsymbol{\psi_{\gamma_2}} \vert(t,\gamma_1),
+\end{equation}
+where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order log-frequency index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
+In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
+\begin{equation}
+\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
+(\boldsymbol{x_2} \ast \phi)(t,\gamma_1,\gamma_2).
+\end{equation}
 In all subsequent experiments, wavelets

### 2016-06-22 paper/structureScene.tex
define Sx

@@
 where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order log-frequency index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
-In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
+In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$. Finally, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation:
 \begin{equation}
@@
 \end{equation}
-In all subsequent experiments, wavelets
+The scattering transform consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand index for $\gamma_1$ or $(\gamma_1,\gamma_2)$.
+
+In all experiments, wavelets
 $\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as a fourth-order Gammatone

### 2016-06-22 figs/wavelets/gammatone.m
write gammatone.m

@@ -1,0 +1,35 @@
+nfo = 8;
+
+opts{1}.time.nFilters_per_octave = nfo;
+opts{1}.time.size = 2^19;
+opts{1}.time.T = 2^14;
+opts{1}.time.is_chunked = false;
+opts{1}.time.gamma_bounds = [0 nfo*11]; % Restrict to top 11 acoustic octaves
+opts{1}.time.wavelet_handle = @gammatone_1d;
+opts{2}.banks.time.nFilters_per_octave = 1;
+opts{2}.banks.time.wavelet_handle = @gammatone_1d;
+opts{2}.banks.time.T = 2^17;
+
+archs = sc_setup(opts);
+
+psi1_fts = display_bank(archs{1}.banks{1});
+psi2_fts = display_bank(archs{2}.banks{1});
+
+%%
+psi1_ft = psi1_fts(:, 40);
+psi1_ift = ifftshift(ifft(psi1_ft));
+psi1_ift = psi1_ift((end/2) + (-300:1000));
+psi1 = [real(psi1_ift),imag(psi1_ift),abs(psi1_ift)];
+
+psi2_ft = psi2_fts(:, 11);
+psi2_ift = ifftshift(ifft(psi2_ft));
+psi2_ift = psi2_ift((end/2) + (-3000:10000));
+psi2 = [real(psi2_ift),imag(psi2_ift),abs(psi2_ift)];
+
+figure(1);
+plot(psi1)
+axis off
+export_fig -transparent gammatone_Q8.png
+plot(psi2)
+axis off
+export_fig -transparent gammatone_Q1.png

### 2016-06-22 paper/structureScene.tex
definition of x_2
define wavelet on the log-frequency axis
place fig 1

@@
 \usepackage{color}
+\usepackage{graphicx}
 \usepackage{xspace}
@@
 
+\begin{figure}
+\begin{center}
+\setlength{\unitlength}{1cm}
+\begin{picture}(5,2)
+ \put(-1.5,0.0){\includegraphics[height=2cm,width=3.5cm]{gammatone_Q8.png}}
+ \put(2.5,0.0){\includegraphics[height=2cm,width=3.5cm]{gammatone_Q1.png}}
+\end{picture}
+\caption{
+Left: Gammatone wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$ with a quality factor of $Q_1=8$.
+Right: Gammatone wavelet $\boldsymbol{\psi_{\gamma_2}}(t)$ with a quality factor of $Q_2=1$.}
+\end{center}
+\end{figure}
+
 \subsection{Time-frequency scattering}
+
+\begin{equation}
+\boldsymbol{\psi_{\gamma_{\mathrm{fr}}, \theta_{\mathrm{fr}}}}(\gamma_1)
+= \theta_{\mathrm{fr}} 2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}
+\boldsymbol{\psi_{\theta_{\mathrm{fr}}}}( \theta_{\mathrm{fr}}  2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}\gamma_1)
+\end{equation}
+
+\begin{equation}
+\boldsymbol{x_2}(t,\gamma_1,\gamma_2,\gamma_{\mathrm{fr}},\theta_{\mathrm{fr}}) =
+\vert \boldsymbol{x_1}
+\overset{t}{\ast} \boldsymbol{\psi_{\gamma_2}}
+\overset{\gamma_1}{\ast} \boldsymbol{\psi_{\gamma_{\mathrm{fr}}, \theta_{\mathrm{fr}}}}
+\vert (t,\gamma_1)
+\end{equation}
+Again, we use the generic literal $\gamma$ as a shorthand for $\gamma_1$ or $(\gamma_1, \gamma_2, \gamma_{\mathrm{fr}}, \theta_{\mathrm{fr}})$.
 

### 2016-06-22 paper/structureScene.tex
comment on the asymmetry of Gammatones, ref Fastl

@@
 where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order log-frequency index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
-In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$. Finally, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation:
+In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
+Finally, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation, yielding
 \begin{equation}
@@
 \end{equation}
-The scattering transform consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand index for $\gamma_1$ or $(\gamma_1,\gamma_2)$.
-
-In all experiments, wavelets
-$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as a fourth-order Gammatone
-wavelet with one vanishing moment \cite{Venkitaraman2014}.
+The scattering transform consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand index for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
 
@@
 \caption{
+\label{fig:gammatones}
 Left: Gammatone wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$ with a quality factor of $Q_1=8$.
@@
 \end{figure}
+
+Wavelets
+$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
+wavelets with one vanishing moment \cite{Venkitaraman2014}, as shown in Figure \ref{fig:gammatones}.
+In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the widespread Morlet wavelets.
+Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \cite{Fastl2007}.
+Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay, hence an
 

### 2016-06-22 paper/structureScene.tex
cite Andén MLSP 2015

@@
 Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \cite{Fastl2007}.
-Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay, hence an
+Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
 
 \subsection{Time-frequency scattering}
+\cite{Anden2015}
 

### 2016-06-22 paper/structureScene.tex
cite Mallat 2012

@@
 \usepackage{color}
-\newcommand{\gl}[1]{\textcolor{red}{Gregoire : #1}}
-\newcommand{\mr}[1]{\textcolor{magenta}{Mathias : #1}}
+\newcommand{\vl}[1]{\textcolor{orange}{Vincent : #1}}
+\newcommand{\gl}[1]{\textcolor{red}{Gr\'egoire : #1}}
+\newcommand{\ja}[1]{\textcolor{magenta}{Joakim : #1}}
 \newcommand{\ml}[1]{\textcolor{blue}{ Mathieu : #1}}
@@
 By resorting to wavelet modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to small time warps,
-in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2010}.
+in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}.
 Besides invariance to translation, this stability property is crucial in signal classification, as it entails a guarantee of robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.

### 2016-06-23 experiments/supervised/dcase_scattering.m
no longer store scattergrams

@@ -49,29 +49,22 @@ end
 %%
 nFiles = length(names);
 X_features = cell(1, nFiles);
-X_scattergrams = cell(1,nFiles);
-parfor file_index = 1:nFiles
+for file_index = 1:nFiles
     name = names{file_index};
     path = fullfile(dataset_path, name);
     stereo_waveform = audioread(path);
     multichannel_waveform = stereo_waveform * mixing_matrix;
     azimuth_features = cell(1, nAzimuths);
-    azimuth_scattergrams = cell(1, nAzimuths);
     for azimuth_index = 1:nAzimuths
-        [features, scattergram] = multichannel_scattering( ...
-            multichannel_waveform(:, azimuth_index), archs);
+        features = ...
+            multichannel_scattering(multichannel_waveform(:, azimuth_index), archs);
         azimuth_features{azimuth_index} = features;
-        azimuth_scattergrams{azimuth_index} = scattergram;
         disp([name, ...
             ', azimuth ', num2str(azimuth_index), ...
             ' finished on worker ', num2str(labindex()), ...
             ' at ', datestr(now(), 'HH:MM:SS')]);
     end
     X_features{file_index} = cat(3, azimuth_features);
-    X_scattergrams{file_index} = cat(4, azimuth_scattergrams);
 end
 X_features = cat(4, X_features);
-X_scattergrams = cat(5,  X_scattergrams);
-save(fullfile('memoized_features', out_file_name), ...
-    'X_features', 'X_scattergrams');
 end

### 2016-06-23 experiments/supervised/supervised_script.m
unnecessary line in script

@@ -33,7 +33,6 @@ opts{1}.time.T = 2^14;
 opts{1}.time.is_chunked = false;
 opts{1}.time.gamma_bounds = [0 nfo*11]; % Restrict to top 11 acoustic octaves
 opts{1}.time.wavelet_handle = @gammatone_1d;
-opts{1}.time.S_log2_oversampling = 0;
 opts{2}.banks.time.nFilters_per_octave = 1;
 opts{2}.banks.time.wavelet_handle = @gammatone_1d;
 opts{2}.banks.time.sibling_mask_factor = 2^6;

### 2016-06-23 experiments/supervised/dcase_scattering.m
save in HDF5 format

@@ -67,4 +67,5 @@ for file_index = 1:nFiles
     X_features{file_index} = cat(3, azimuth_features);
 end
 X_features = cat(4, X_features);
+save(fullfile('memoized_features', out_file_name), 'X_features', '-v7.3');
 end

### 2016-06-23 experiments/supervised/multichannel_scattering.m
update multichannel scattering
* parfor is now along chunks
* works for joint scattering

@@ -1,4 +1,4 @@
-function [features, scattergram] = multichannel_scattering(waveform, archs)
+function features = multichannel_scattering(waveform, archs)
 %% Truncation
 original_length = 30 * 44100;
 truncated_length = 5 * 2^18;
@@ -23,54 +23,14 @@ nChunks = 4;
 chunks = reshape(chunks, size(chunks, 1), nChunks * nAzimuths);
 
 %%
-nLayers = length(archs);
-S = cell(1, nLayers);
-U = cell(1, nLayers);
-Y = cell(1, nLayers);
-
-U{1+0} = initialize_U(chunks, archs{1}.banks{1});
-
-%% Propagation cascade
-for layer = 1:nLayers
-    arch = archs{layer};
-    previous_layer = layer - 1;
-    % Scatter iteratively layer U to get sub-layers Y
-    if isfield(arch, 'banks')
-        Y{layer} = U_to_Y(U{1+previous_layer}, arch.banks);
-    else
-        Y{layer} = U(1+previous_layer);
-    end
-    
-    % Apply nonlinearity to last sub-layer Y to get layer U
-    if isfield(arch, 'nonlinearity') 
-        U{1+layer} = Y_to_U(Y{layer}{end}, arch.nonlinearity);
-    end
-    
-    % Blur/pool first layer Y to get layer S
-    if isfield(arch, 'invariants')
-        S{1+previous_layer} = Y_to_S(Y{layer}, arch);
-    end
+S_chunks = cell(1, nChunks);
+parfor chunk_index = 1:nChunks
+    S_chunks{chunk_index} = sc_propagate(chunks(:, chunk_index), archs);
 end
 
-%%
-S1 = S{1+1}.data((1+end/4):(3*end/4), :, :);
-S1 = reshape(S1, size(S1, 1) * nChunks, nAzimuths, size(S1, 3));
-S1 = permute(S1, [3, 1, 4, 2]);
-
-%%
-[nLambda1s, nFrames, ~] = size(S1);
-if iscell(S{1+2})
-else
-    nLambda2s = length(S{1+2}.data);
-    scattergram = cat(3, S1, zeros(nLambda1s, nFrames, nLambda2s, nAzimuths));
-    features = S1(:, :, 1, ceil(end/2));
-    for lambda2_index = 1:nLambda2s
-        band = S{1+2}.data{lambda2_index}((1+end/4):(3*end/4), :, :);
[…]

### 2016-06-23 paper/structureScene.tex
transitions
some elements on time-frequency scattering

@@
 
+\usepackage{amssymb}
+\usepackage{amsmath}
 \usepackage{pgf}
@@
 In practice, to achieve invariance to translation up to $T = 186\,\mathrm{ms}$, \ie of the order of the minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ suffice.
-The next section describes the operations involved in temporal scattering.
+The next subsection describes the operations involved in temporal scattering.
 
@@
 
+The next subsection extends the second-order scattering operator to joint wavelet transforms over the time and log-frequency variables.
+
 \subsection{Time-frequency scattering}
+
 \cite{Anden2015}
+%The temporal scale $2^{\gamma_2}/\xi_2$ takes $5$ values between $23\,\mathrm{ms}$ and $370\,\mathrm{ms}$.
+%The log-frequential scale $\beta$ takes $5$ values between $1/4^{\mathrm{th}}$ of an octave and $4$ octaves, as well as their $5$ "mirror" frequencies,
+%according to a geometric sequence.
+%In addition, the edge case $\gamma_{\mathrm{fr}} = \infty$ corresponds
+%to a log-frequential low-pass filter along $4$ octaves
+%according to all $5$ temporal scales $2^{\gamma_2}/\xi_2$.
 
@@
 \section{Supervised approach}
+
+\subsection{}
+
+\begin{equation}
+\widetilde{\mathbf{S}}\boldsymbol{X}(\gamma, n) =
+\dfrac{\log \mathbf{S}\boldsymbol{X}(\gamma, n) -
+\mathbb{E}[\log \mathbf{S}\boldsymbol{X}(\gamma)]}{\sigma(\log \mathbf{S}\boldsymbol{X}(\gamma))}
+\end{equation}
+where $\mathbb{E}[\log \mathbf{S}\boldsymbol{X}(\gamma)] = \frac{1}{N} \sum_{n=1}^{N} \log \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the expected value of $\log \mathbf{S}\boldsymbol{X}(\gamma,n)$ and
+\begin{equation}
+\sigma(\log \mathbf{S}\boldsymbol{X} (\gamma)) =
+\frac{1}{N-1} \sum_{n=1}^{N}
+( \log \mathbf{S}\boldsymbol{X}(\gamma,n) - \mathbb{E}[\log \mathbf{S}\boldsymbol{X}] )
+\end{equation}
+ is the sample standard deviation.
 

### 2016-06-23 paper/structureScene.tex
one sentence on invertibility and energy conservation
rephrase a sentence about the CQT

@@
 theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the CQT has a fixed temporal resolution, constrained by the support of the STFT analyzing window.
+frequency range, whereas the temporal resolution of the CQT is fixed by the support of the STFT analyzing window.
 Therefore, the wavelet scalogram has a better temporal localization at high
@@
 \cite{Anden2015}
-%The temporal scale $2^{\gamma_2}/\xi_2$ takes $5$ values between $23\,\mathrm{ms}$ and $370\,\mathrm{ms}$.
-%The log-frequential scale $\beta$ takes $5$ values between $1/4^{\mathrm{th}}$ of an octave and $4$ octaves, as well as their $5$ "mirror" frequencies,
-%according to a geometric sequence.
-%In addition, the edge case $\gamma_{\mathrm{fr}} = \infty$ corresponds
-%to a log-frequential low-pass filter along $4$ octaves
-%according to all $5$ temporal scales $2^{\gamma_2}/\xi_2$.
-
-\begin{equation}
-\boldsymbol{\psi_{\gamma_{\mathrm{fr}}, \theta_{\mathrm{fr}}}}(\gamma_1)
-= \theta_{\mathrm{fr}} 2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}
-\boldsymbol{\psi_{\theta_{\mathrm{fr}}}}( \theta_{\mathrm{fr}}  2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}\gamma_1)
-\end{equation}
+Since it operates over a complex-valued input, the filterbank over log-frequencies $\gamma_1$ must cover the whole Fourier spectrum, including negative frequencies and the null frequency, to ensure invertibility and energy conservation.
 

### 2016-06-23 paper/structureScene.tex
my comments in orange color

@@
 Left: Gammatone wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$ with a quality factor of $Q_1=8$.
-Right: Gammatone wavelet $\boldsymbol{\psi_{\gamma_2}}(t)$ with a quality factor of $Q_2=1$.}
+Right: Gammatone wavelet $\boldsymbol{\psi_{\gamma_2}}(t)$ with a quality factor of $Q_2=1$.
+\vl{we should discuss color schemes and axis labels}}
 \end{center}
@@
 
+\vl{missing sentence here ?}
 The next subsection extends the second-order scattering operator to joint wavelet transforms over the time and log-frequency variables.

### 2016-06-24 paper/structureScene.tex
formula for the center frequencies
try reformulating the title
Scattering Auditory Scenes for Classification: Supervised and
Unsupervised Approaches

@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Acoustic Scene Classification: \\ Supervised and Unsupervised Approaches using the Scattering Transform}
+\title{Scattering Auditory Scenes for Classification: Supervised and Unsupervised Approaches}
 
@@
 Since it operates over a complex-valued input, the filterbank over log-frequencies $\gamma_1$ must cover the whole Fourier spectrum, including negative frequencies and the null frequency, to ensure invertibility and energy conservation.
+Therefore, wavelets on $\gamma_1$ have center frequencies of the form
+$\theta_{\mathrm{fr}}  2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}$, where $\gamma_{\mathrm{fr}}$ takes integer values betwen $0$ and $(J_\mathrm{fr} \times Q_\mathrm{fr} - 1)$ and $\theta_{\mathrm{fr}} = \pm 1$.
 
@@
 
-\section{Supervised approach}
+\section{Supervised approach: support vector machines}
+Support vector machines are large-margin binary discriminators with tunable regularization of misclassified examples.
 

### 2016-06-26 paper/structureScene.tex
write intro of section IV

@@
 
+\section{Feature design}
+Prior to supervised or unsupervised classification, is it beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power. 
+This section reviews binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.
+The former is novel and specific to stereophonic recordings, whereas the next ones would also apply to many other problems.
+
 \section{Unsupervised approach}

### 2016-06-26 paper/structureScene.tex
start writing section IV

@@
 
+\subsection{Binaural data augmentation}
+The recording condition of most acoustic scene datasets follow a binaural protocol, \ie a pair of microphones respectively located over the ears of the recordist.
+This protocol provides a realistic description of the spatial auditory environment, as it reproduces the natural listening conditions of humans.
+% In particular, the inter-intensity difference (IID) between left and right channel conveys the approximate azimuth of each sound source that make up the scene with respect to the listener.
+Yet, since the location of the recordist may vary across instances of the same class, the feature space for classification should be invariant to the location of sources.
+In order to achieve invariance to location, averaging the left and right channels into a monophonic signal is by far the most widespread technique.
+
+\begin{equation}
+\boldsymbol{x}(t) =
+\alpha \boldsymbol{\times x_\mathsf{L}}\!(t) +
+(1-\alpha) \boldsymbol{\times x_\mathsf{R}}\!(t)
+\end{equation}
+In practice, $\alpha$ takes values in ${0\%, 25\%, 50\%, 75\%, 100\%}$.
+
+\subsection{Feature selection}
+
+\subsection{Logarithmic compression}
+% linearizes the dependency with respect to the distance to the source
+
+\subsection{Standardization}
+
+\subsection{Early vs. late integration}
+Fine-grained classification of acoustic scenes is only made
+possible by integrating signal information over a long temporal context.
+Whereas a few seconds are often sufficient to recognize a speaker,
+a musical instrument, or a genre, it may require up to 30 seconds
+to disambiguate closely related acoustic scenes.
+
 \section{Unsupervised approach}

### 2016-06-26 paper/structureScene.tex
write more on the problem of converting to mono

@@
 This protocol provides a realistic description of the spatial auditory environment, as it reproduces the natural listening conditions of humans.
-% In particular, the inter-intensity difference (IID) between left and right channel conveys the approximate azimuth of each sound source that make up the scene with respect to the listener.
-Yet, since the location of the recordist may vary across instances of the same class, the feature space for classification should be invariant to the location of sources.
-In order to achieve invariance to location, averaging the left and right channels into a monophonic signal is by far the most widespread technique.
+In particular, the interaural level difference (ILD) between left and right channel conveys the approximate azimuth of each sound source that make up the scene with respect to the listener.
+\vl{ref needed}
+Yet, since the location of the recordist varies across instances of the same class, the feature space for classification should be invariant to these azimuths.
 
+In order to achieve invariance to ILD, averaging the left and right channels into a monophonic signal is by far the most widespread approach.
+An important drawback is that it biases the measurement towards the center of the scene, as lateral sources get damped.
+%Moreover, sources 
 \begin{equation}

### 2016-06-26 paper/structureScene.tex
write more on section IV

@@
 In order to achieve invariance to ILD, averaging the left and right channels into a monophonic signal is by far the most widespread approach.
-An important drawback is that it biases the measurement towards the center of the scene, as lateral sources get damped.
-%Moreover, sources 
+However, it favors the center of the scene while attenuating the presence of lateral sources.
+Moreover, sources that are on distinct azimuths get mixed, which may cause bias at training time.
+Instead, so as to take advantage of stereophonic data, we perform multiiple combinations of the left channel $\boldsymbol{x_{\mathsf{L}}}(t)$ and right channel $\boldsymbol{x_{\mathsf{R}}}(t)$ into monophonic signals $\boldsymbol{x_{\alpha}}(t)$ according to the equation
 \begin{equation}
-\boldsymbol{x}(t) =
-\alpha \boldsymbol{\times x_\mathsf{L}}\!(t) +
-(1-\alpha) \boldsymbol{\times x_\mathsf{R}}\!(t)
+\boldsymbol{x_{\alpha}}(t) =
+\dfrac{1+\alpha}{2} \boldsymbol{x_\mathsf{L}}(t) +
+\dfrac{1-\alpha}{2} \boldsymbol{x_\mathsf{R}}(t),
 \end{equation}
-In practice, $\alpha$ takes values in ${0\%, 25\%, 50\%, 75\%, 100\%}$.
+where $\alpha$ is a panoramic mixing weight, taking values between $-1$ and $+1$, which could be deterministic or random.
+In following experiments, we deterministically set $\alpha$ to the values $-1$, $\frac{1}{2}$, $0$, $+\frac{1}{2}$, and $+1$.
+This is a form of data augmentation, because $5$ monophonic signals are derived from each binaural recording in the training set.
 
 \subsection{Feature selection}
+The starting scale $\gamma_{1,\min}$ corresponding to some upper bound $\xi_{1,\max}$ on the frequency axis is given by
+\begin{equation}
+\gamma_{1,\min} =
+\left\lfloor Q_1 \log_2 \frac{\xi_1}{\xi_{1,\max}} \right\rceil
+\end{equation}
+where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.
 
 \subsection{Logarithmic compression}
+Many pattern recognition algorithms, including nearest neighbour classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
+
+%Logarithmic compression is corroborated by psychoacoustics
 % linearizes the dependency with respect to the distance to the source
 
+And\'en \cite{Anden2014}
+\begin{equation}
+\log \dfrac{\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)}{\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)} =
+\log \mathbf{S_2}\boldsymbol{x}(t, \gamma_1, \gamma_2) -
+\log \mathbf{S_1}\boldsymbol{x}(t, \gamma_1)
+\end{equation}
+
 \subsection{Standardization}
+
+\begin{equation}
+\widetilde{\mathbf{S}}\boldsymbol{X}(\gamma, n) =
+\dfrac{ \mathbf{S}\boldsymbol{X}(\gamma, n) -
+\mathbb{E}[ \mathbf{S}\boldsymbol{X}(\gamma)]}{\sigma( \mathbf{S}\boldsymbol{X}(\gamma))}
+\end{equation}
+where $\mathbb{E}[ \mathbf{S}\boldsymbol{X}(\gamma)] = \frac{1}{N} \sum_{n=1}^{N} \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the expected value of $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and
+\begin{equation}
+\sigma(\mathbf{S}\boldsymbol{X} (\gamma)) =
+\sqrt{\frac{1}{N-1} \sum_{n=1}^{N}
+\left( \mathbf{S}\boldsymbol{X}(\gamma,n) - \mathbb{E}[\mathbf{S}\boldsymbol{X}] \right)^2}
+\end{equation}
+ is the sample standard deviation.
 
@@
 possible by integrating signal information over a long temporal context.
-Whereas a few seconds are often sufficient to recognize a speaker,
-a musical instrument, or a genre, it may require up to 30 seconds
-to disambiguate closely related acoustic scenes.
+
+\cite{Joder2009}
+
+The next section discusses more advanced methods for the temporal integration of scattering coefficients in an unsupervised setting.
 
@@
 Support vector machines are large-margin binary discriminators with tunable regularization of misclassified examples.
-
-\subsection{}
-
-\begin{equation}
-\widetilde{\mathbf{S}}\boldsymbol{X}(\gamma, n) =
-\dfrac{\log \mathbf{S}\boldsymbol{X}(\gamma, n) -
-\mathbb{E}[\log \mathbf{S}\boldsymbol{X}(\gamma)]}{\sigma(\log \mathbf{S}\boldsymbol{X}(\gamma))}
-\end{equation}
-where $\mathbb{E}[\log \mathbf{S}\boldsymbol{X}(\gamma)] = \frac{1}{N} \sum_{n=1}^{N} \log \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the expected value of $\log \mathbf{S}\boldsymbol{X}(\gamma,n)$ and
-\begin{equation}
-\sigma(\log \mathbf{S}\boldsymbol{X} (\gamma)) =
-\frac{1}{N-1} \sum_{n=1}^{N}
-( \log \mathbf{S}\boldsymbol{X}(\gamma,n) - \mathbb{E}[\log \mathbf{S}\boldsymbol{X}] )
-\end{equation}
- is the sample standard deviation.
 

### 2016-06-27 paper/structureScene.tex
add a lot of references regarding acoustic scenes classification. Add short descriptions of the dcase 2013 participant. add refList.txt (old references). Clean biblio.bib. Remove .bbl and .blg

@@
 
-Also, some recent works demonstrate the interest of such approaches for the characterization of human acoustic pleasantness in urban areas \cite{lafay, lavandier} and the prediction of annoyance due to the traffic \cite{gloaguen}.
+Also, some recent works demonstrate the interest of such approaches for the characterization of human acoustic pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} \gl{What was the lavandier article ?} and the prediction of annoyance due to the traffic \cite{gloaguen}.
 We believe that those case study are of interest for the signal processing community as they have strong societal impacts and raises interesting research avenues.
@@
 A closely related field of investigation that is relatively more mature is the classification of acoustic scenes whose aim is to predict labels by processing audio data.
-Despite its lower value, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois} and been considered in the data processing field for a while, with availability of some well designed evaluation datasets.
+Despite its lower value, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and been considered in the data processing field for a while, with availability of some well designed evaluation datasets.
 
@@
 
-GL 1
+\gl{bag of frame approach:} \cite{aucouturier2007bag}: bag of frame approach. \cite{lagrange:hal-01082501}: why the bag of frame approach does no work. \\
 
-ML 2
+\gl{DCASE challenge:} \cite{chum2013ieee}: CHR system, using HMM and frame based classification.
+\cite{elizalde2013vector} ELF system, I vector approach using probabilistic LDA as classifier.
+\cite{geiger2013large} GSR system, basically mel/MFCC using SVM as classifier.
+\cite{krijnders2013tone} KH system, tone fit features computed from cochleogram (but similar results are obtained with MFCCs) using SVM as classifier.
+\cite{li2013auditory} LTT system, Tree bagger classifier used on segments (MFCCs) prior to a majority vote.
+\cite{lee2013acoustic} NHL system, unsupervised feature learning approach on mel frequency spectrogram with selective Max-pooling.
+Supervised training is performed using SVM.
+\cite{nogueira2013sound} NR1 system, MFCC and spatial features (inter-aural differences) with SVM as classifier.
+\cite{olivetti2013wonder} OE system, normalized compression distance (I honestly did not understand a thing about it, but this the worst method of the challenge anyway) using  Random Forest  classifier.
+\cite{patil2013multiresolution} PE system, 2D Gabor filters applied on 1sec segments with SVM as classifier.
+\cite{rakotomamonjy2015histogram} RG system, histogram of gradient representation, computed from a constant Q transform with SVM as classifier.
+\cite{roma2013recurrence} RNH system, RQA stacked with MFCC, SVM as classifier.
+\\
 
-\subsection{Goal}
+\gl{Other paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, barchiesi2015acoustic, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory, bisot2016acoustic}
 
-\subsection{Methods}
+\subsection{Describing a scene (Feature extraction)}
+
+\subsection{Classifying a scene}
 
@@
 
-dcase2013 \cite{7100934} \cite{giannoulis2013database}
+DCASE 2013 challenge reference papers. \cite{giannoulis2013database, 7100934} \\
 

### 2016-06-27 paper/structureScene.tex
write more of section IV

--- 2016-06-26 paper/structureScene.tex
@@
 
-despite its size, it is still challenging as state-of-the-art systems achieves 76 \% \cite{6701890} (winner of the 2013 challenge RQA+SVM).
+despite its size, it is still challenging as state-of-the-art systems achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
 For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively.
@@
 where $\alpha$ is a panoramic mixing weight, taking values between $-1$ and $+1$, which could be deterministic or random.
+
 In following experiments, we deterministically set $\alpha$ to the values $-1$, $\frac{1}{2}$, $0$, $+\frac{1}{2}$, and $+1$.
 This is a form of data augmentation, because $5$ monophonic signals are derived from each binaural recording in the training set.
+At test time, only the center mix ($\alpha = 0$) is used to perform a prediction.
 
 \subsection{Feature selection}
+In a scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$, some coefficients $\gamma$ may consistently bear negligible information for the task of interest.
+This has lead \cite{roma2013}
+
 The starting scale $\gamma_{1,\min}$ corresponding to some upper bound $\xi_{1,\max}$ on the frequency axis is given by
@@
 \subsection{Logarithmic compression}
-Many pattern recognition algorithms, including nearest neighbour classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
+Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
+Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design, and their distribution is typically skewed towards the right.
 
-%Logarithmic compression is corroborated by psychoacoustics
-% linearizes the dependency with respect to the distance to the source
+This issue can be addressed by applying to all coefficients a pointwise concave transformation, \eg logarithmic, hence reducing skewness.
+In particular, logarithmic compression is ubiquitous in audio signal processing, as it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
+In the context of auditory scenes, we must also stress that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone.
+Logarithmic compression can linearize this dependency, which arguably facilitates the construction of an invariant to this distance at the classifier stage.
 
@@
 \end{equation}
+
+\subsection{Early vs. late integration}
+Fine-grained classification of acoustic scenes is only made
+possible by integrating signal information over a long temporal context.
+A straightforward approach, known as early integration, consists in summarizing the multivariate time series of scattering coefficients
+
+Early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff. We refer to \cite{Joder2009} for a comprehensive review of this problematic.
 
@@
 
-\subsection{Early vs. late integration}
-Fine-grained classification of acoustic scenes is only made
-possible by integrating signal information over a long temporal context.
-
-\cite{Joder2009}
-
-The next section discusses more advanced methods for the temporal integration of scattering coefficients in an unsupervised setting.
+The next section discusses adaptive methods for the temporal integration of scattering coefficients in an unsupervised setting.
 

### 2016-06-27 paper/structureScene.tex
finish writing subsection IV. C

@@
 Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design, and their distribution is typically skewed towards the right.
+This issue can however be addressed by applying to all coefficients a pointwise concave transformation, \eg logarithmic, hence reducing skewness.
 
-This issue can be addressed by applying to all coefficients a pointwise concave transformation, \eg logarithmic, hence reducing skewness.
 In particular, logarithmic compression is ubiquitous in audio signal processing, as it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
-In the context of auditory scenes, we must also stress that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone.
-Logarithmic compression can linearize this dependency, which arguably facilitates the construction of an invariant to this distance at the classifier stage.
+In the context of auditory scenes, we must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone, which is a spurious factor a variability to the task of auditory scene classification.
+Logarithmic compression can linearize this dependency, which arguably facilitates the construction of a powerful invariant at the classifier stage.
 
-And\'en \cite{Anden2014}
+Given a task of musical genre recognition, \cite{Anden2014} has advocated for the renormalization of second-order scattering coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ by the corresponding first-order scattering coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, as it provably decorrelates the amplitudes of their activations.
+Interestingly, taking the logarithm of renormalized coefficients would yield
 \begin{equation}
@@
 \log \mathbf{S_2}\boldsymbol{x}(t, \gamma_1, \gamma_2) -
-\log \mathbf{S_1}\boldsymbol{x}(t, \gamma_1)
+\log \mathbf{S_1}\boldsymbol{x}(t, \gamma_1),
 \end{equation}
+\ie a linear combination of the logarithms of first- and second-order coefficients.
+Therefore, the theoretical insight brought by \cite{Anden2014} in favor of renormalized scattering also applies to log-scattering up to a linear transformation in feature space, to which affine classifiers are not sensitive.
 

### 2016-06-27 paper/structureScene.tex
write on temporal integration

@@
 
-\section{Scattering representations}
+\section{Scattering representations \label{sec:scattering}}
 Scattering transforms are translation-invariant representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
@@
 
-\subsection{Early vs. late integration}
-Fine-grained classification of acoustic scenes is only made
+\subsection{Early vs. late temporal integration}
+Owing to the scarcity of salient events in many natural soundscapes, fine-grained classification of auditory scenes is only made
 possible by integrating signal information over a long temporal context.
-A straightforward approach, known as early integration, consists in summarizing the multivariate time series of scattering coefficients
+Indeed, whereas a few seconds are often sufficient to recognize a speaker,
+a musical instrument, or a genre, it may require up to 30 seconds
+to disambiguate auditory scenes which share part of their semantic content, \eg a subway train from a subway station or a quiet street from a park.
+
+A straightforward approach, known as early integration, consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene, by retaining their average values only.
+Going back to the definition of the scattering transform given in section \ref{sec:scattering}, this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
 

### 2016-06-27 paper/structureScene.tex
typo

@@
 frequencies than the CQT, at the expense of a greater computational cost.
-This property allows to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+This property allows to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
 

### 2016-06-27 paper/structureScene.tex
write more on section IV B

@@
 \subsection{Feature selection}
-In a scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$, some coefficients $\gamma$ may consistently bear negligible information for the task of interest.
-This has lead \cite{roma2013}
+In a multivariate time series of descriptors $\mathbf{S}\boldsymbol{x}(t,\gamma)$, some coefficients $\gamma$ may consistently bear negligible information for the task of interest.
+This fact may cause the classifier to overfit the training set, especially when the number of features is large and the number of training instances is small, as is the case here.
+
+%Many approaches coexist for feature selection, including matrix factorization, orthogonal matching pursuit.
+This is inspired by the state-of-the-art results of \cite{roma2013}, which rely on cepstral descriptors computed up to a cutoff frequency as small as 900 Hz.
 

### 2016-06-27 paper/structureScene.tex
prior to -> before

@@
 \cite{krijnders2013tone} KH system, tone fit features computed from cochleogram (but similar results are obtained with MFCCs) using SVM as classifier.
-\cite{li2013auditory} LTT system, Tree bagger classifier used on segments (MFCCs) prior to a majority vote.
+\cite{li2013auditory} LTT system, Tree bagger classifier used on segments (MFCCs) before a majority vote.
 \cite{lee2013acoustic} NHL system, unsupervised feature learning approach on mel frequency spectrogram with selective Max-pooling.
@@
 Indeed, the starting time of the recordings are chosen arbitrarily, and thus do not convey any information about the class.
-To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space prior to training the classifier.
+To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space before training the classifier.
 From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ of  a signal $\boldsymbol{x}(t)$ --
@@
 
+\subsection{Time-frequency scattering}
+The most salient events among auditory scenes draw highly structured patterns in the time-frequency plane.
 \vl{missing sentence here ?}
-The next subsection extends the second-order scattering operator to joint wavelet transforms over the time and log-frequency variables.
+Despite its invariance and stability properties, the temporal scattering transform fails to characterize these patterns, because it extracts the amplitude modulations of each wavelet subband separately.
 
-\subsection{Time-frequency scattering}
+A recent improvement to the temporal scattering transform has been brought by \cite{Anden2015},
 
-\cite{Anden2015}
 Since it operates over a complex-valued input, the filterbank over log-frequencies $\gamma_1$ must cover the whole Fourier spectrum, including negative frequencies and the null frequency, to ensure invertibility and energy conservation.
@@
 $\theta_{\mathrm{fr}}  2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}$, where $\gamma_{\mathrm{fr}}$ takes integer values betwen $0$ and $(J_\mathrm{fr} \times Q_\mathrm{fr} - 1)$ and $\theta_{\mathrm{fr}} = \pm 1$.
-
 \begin{equation}
@@
 \section{Feature design}
-Prior to supervised or unsupervised classification, is it beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power. 
+Before supervised or unsupervised classification, is it beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power. 
 This section reviews binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.

### 2016-06-27 paper/structureScene.tex
update section III

@@
 
-A recent improvement to the temporal scattering transform has been brought by \cite{Anden2015},
+This issue was remedied by time-frequency scattering \cite{Anden2015}, which performs a wavelet decomposition over the log-frequency variable $\gamma_1$ after the second-order wavelet decomposition over time.
+While temporal scattering merely extracts the average amplitude modulation spectrum, time-frequency scattering computes spectral modulations at various temporal rates and spectral scales.
+% link with GBFB and STRF which have proven useful for acoustic scene classification.
 
@@
 This section reviews binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.
-The former is novel and specific to stereophonic recordings, whereas the next ones would also apply to many other problems.
+The former is novel and specific to stereophonic recordings, whereas the next ones would also apply to many other kinds of data.
 

### 2016-06-27 figs/pipeline/pipeline.m
load memoized features and standardization

@@ -64,8 +64,21 @@ truncated_scalogram = log_scalogram(starting_gamma:end, :);
 figure(12);
 imagesc(max(truncated_scalogram, 0));
 
+%% Load memoized features
+memoized_path = ...
+    'experiments/supervised/memoized_features/dcase2013_timeQ8_train.mat';
+training_data = load(memoized_path);
+X_train = training_data.dcase2013_timeQ8_train.X_train;
+
+%% Compute mean and standard deviation
+X_sizes = size(X_train);
+X_train = reshape(X_train, X_sizes(1), prod(X_sizes(2:end)));
+X_mean = mean(X_train, 2);
+X_std = std(X_train, [], 2);
+gamma_range = starting_gamma + (1:size(truncated_scalogram, 1)) - 1;
+X_mean = X_mean(gamma_range);
+X_std = X_std(gamma_range);
+
 %% Standardization
-scalogram_mean = mean(truncated_scalogram, 2);
-scalogram_std = std(truncated_scalogram, [], 2);
-centered_scalogram = bsxfun(@minus, truncated_scalogram, scalogram_mean);
-imagesc(centered_scalogram)
+centered_scalogram = bsxfun(@minus, truncated_scalogram, X_mean);
+standardized_scalogram = bsxfun(@rdivide, centered_scalogram, X_std);

### 2016-06-27 paper/structureScene.tex
write more IV. D

@@
 a musical instrument, or a genre, it may require up to 30 seconds
-to disambiguate auditory scenes which share part of their semantic content, \eg a subway train from a subway station or a quiet street from a park.
+to disambiguate auditory scenes which share part of their semantic content, \eg a train from a subway station or a quiet street from a park.
+Depending on whether aggregation is performed in feature space or in decision space, the corresponding method is referred to as early or late integration.
 
-A straightforward approach, known as early integration, consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene, by retaining their average values only.
+A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene, by retaining their average values only.
 Going back to the definition of the scattering transform given in section \ref{sec:scattering}, this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
+With a slight abuse of notation, we denote by
+\begin{equation}
+\mathbf{S}\boldsymbol{x}(\gamma) =
+\int_{-\infty}^{+\infty} \mathbf{S}\boldsymbol{x}(t,\gamma)\;\mathrm{d}t
+\end{equation}
+the summarized features.
+
+In turn, late integration
 

### 2016-06-27 paper/structureScene.tex
write write write

@@
 
-In turn, late integration
+In turn, a late integration scheme relies on probabilistic estimates over short-term windows of length $T$, which are subsequently aggregated to produce a final decision
+\begin{equation}
+\hat{y} = \arg \max_{y} \rho\Big(\big\{ \mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right] \big\}_{t} \Big),
+\end{equation}
+where $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.
 
-Early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff. We refer to \cite{Joder2009} for a comprehensive review of this problematic.
+The principal drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene -- if we don't account for data augmentation.
+In the context of the DCASE 2013 dataset, this corresponds to merely $10$ training instances per class, and $100$ instances overall, hence a risk of overfitting.
+On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene, resulting in $12800$ instances overall.
+However, many of these instances may be silent or lack any salient properties of the category they are assigned to, hence an increase in bias and a risk of underfitting.
+In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff.
+We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
 
@@
 Support vector machines are large-margin binary discriminators with tunable regularization of misclassified examples.
+% The training set is linearly separable in feature space.
 

### 2016-06-28 paper/structureScene.tex
cite Wang, Blauert, revise section III

@@
 From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ of  a signal $\boldsymbol{x}(t)$ --
-a translation-invariant representation up to time skills of duration $T$ can be simply obtained
+a translation-invariant representation up to time shifts of duration $T$ can be simply obtained
 by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
@@
 
-By resorting to wavelet modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to small time warps,
+By resorting to wavelet transform modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to small time warps,
 in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}.
@@
 In practice, to achieve invariance to translation up to $T = 186\,\mathrm{ms}$, \ie of the order of the minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ suffice.
+
 The next subsection describes the operations involved in temporal scattering.
@@
 the wavelets resulting from the dilation of $\boldsymbol{\psi}$(t).
-For some $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
+For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
 has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$, a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$, and a quality factor of $Q_1$.
@@
 \end{equation}
-The scattering transform consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand index for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
+The scattering transform consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
 
@@
 \subsection{Binaural data augmentation}
-The recording condition of most acoustic scene datasets follow a binaural protocol, \ie a pair of microphones respectively located over the ears of the recordist.
+Most acoustic scene datasets are recorded according to a binaural protocol, \ie with a pair of in-ear microphones \cite{Wang2006}.
 This protocol provides a realistic description of the spatial auditory environment, as it reproduces the natural listening conditions of humans.
-In particular, the interaural level difference (ILD) between left and right channel conveys the approximate azimuth of each sound source that make up the scene with respect to the listener.
-\vl{ref needed}
+In particular, the interaural level difference (ILD) between left and right channel conveys the approximate azimuth of each sound source that make up the scene with respect to the listener \cite{Blauert2003}.
 Yet, since the location of the recordist varies across instances of the same class, the feature space for classification should be invariant to these azimuths.

### 2016-06-28 paper/structureScene.tex
revise section IV

@@
 In order to achieve invariance to ILD, averaging the left and right channels into a monophonic signal is by far the most widespread approach.
-However, it favors the center of the scene while attenuating the presence of lateral sources.
+However, it favors the center of the scene while attenuating lateral cues.
 Moreover, sources that are on distinct azimuths get mixed, which may cause bias at training time.
-Instead, so as to take advantage of stereophonic data, we perform multiiple combinations of the left channel $\boldsymbol{x_{\mathsf{L}}}(t)$ and right channel $\boldsymbol{x_{\mathsf{R}}}(t)$ into monophonic signals $\boldsymbol{x_{\alpha}}(t)$ according to the equation
+
+Instead, so as to leverage stereophonic data, we perform multiple combinations of the left channel $\boldsymbol{x_{\mathsf{L}}}(t)$ and right channel $\boldsymbol{x_{\mathsf{R}}}(t)$ into monophonic signals $\boldsymbol{x_{\alpha}}(t)$ according to the equation
 \begin{equation}
@@
 This is a form of data augmentation, because $5$ monophonic signals are derived from each binaural recording in the training set.
-At test time, only the center mix ($\alpha = 0$) is used to perform a prediction.
+At test time, only the center mix ($\alpha = 0$) is used to perform the class.
 
@@
 
-%Many approaches coexist for feature selection, including matrix factorization, orthogonal matching pursuit.
 This is inspired by the state-of-the-art results of \cite{roma2013}, which rely on cepstral descriptors computed up to a cutoff frequency as small as 900 Hz.

### 2016-06-28 paper/structureScene.tex
finish writing part IV. E

@@
 
-In turn, a late integration scheme relies on probabilistic estimates over short-term windows of length $T$, which are subsequently aggregated to produce a final decision
+Conversely, a late integration scheme relies on probabilistic estimates over short-term windows of length $T$, which are subsequently aggregated to produce a final decision
 \begin{equation}
@@
 \end{equation}
-where $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.
+where $\hat{y}$ is the estimated class label and $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.
 
 The principal drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene -- if we don't account for data augmentation.
-In the context of the DCASE 2013 dataset, this corresponds to merely $10$ training instances per class, and $100$ instances overall, hence a risk of overfitting.
+In the context of the DCASE 2013 dataset, this corresponds to merely $10$ training instances per class, and $100$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
 On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene, resulting in $12800$ instances overall.
@@
 \subsection{Standardization}
-
+Let $\mathbf{S}\boldsymbol{X}(\gamma,n)$ be a dataset, where the indices $\gamma$ and $n$ respectively denotes features and examples.
+It is commonly acknowledged that support vector machines should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges.
+This operation may also help making the dataset linearly separable in feature space.
+\vl{Check this claim in DCASE 2013 time scattering features.}
+To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we apply the affine transformation
 \begin{equation}
@@
 \dfrac{ \mathbf{S}\boldsymbol{X}(\gamma, n) -
-\mathbb{E}[ \mathbf{S}\boldsymbol{X}(\gamma)]}{\sigma( \mathbf{S}\boldsymbol{X}(\gamma))}
+\mathbb{E}[ \mathbf{S}\boldsymbol{X}](\gamma)}{\sigma[ \mathbf{S}\boldsymbol{X}](\gamma)}
 \end{equation}
-where $\mathbb{E}[ \mathbf{S}\boldsymbol{X}(\gamma)] = \frac{1}{N} \sum_{n=1}^{N} \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the expected value of $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and
+where $\mathbb{E}[ \mathbf{S}\boldsymbol{X}](\gamma) = \frac{1}{N} \sum_{n=1}^{N} \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the expected value of $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and
 \begin{equation}
-\sigma(\mathbf{S}\boldsymbol{X} (\gamma)) =
+\sigma[\mathbf{S}\boldsymbol{X}] (\gamma) =
 \sqrt{\frac{1}{N-1} \sum_{n=1}^{N}
@@
  is the sample standard deviation.
+
+ The vectors $\mathbb{E}[\mathbf{S}\boldsymbol{X}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{X}](\gamma)$ are estimated from the training set only, and the same affine transformation is subsequently applied to all samples in the training set and the test set.
 
@@
 Support vector machines are large-margin binary discriminators with tunable regularization of misclassified examples.
-% The training set is linearly separable in feature space.
 

### 2016-06-29 experiments/supervised/get_timefrequency_frequencies.m
write function get_timefrequency_frequencies

@@ -1,0 +1,114 @@
+function frequencies = get_timefrequency_frequencies(nfo)
+
+% Prepare scattering "architectures", i.e. filter banks and nonlinearities
+opts{1}.time.nFilters_per_octave = nfo;
+opts{1}.time.size = 2^15;
+opts{1}.time.T = 2^14;
+opts{1}.time.is_chunked = false;
+opts{1}.time.gamma_bounds = [0 nfo*11]; % Restrict to top 11 acoustic octaves
+opts{1}.time.wavelet_handle = @gammatone_1d;
+opts{1}.time.S_log2_oversampling = 0;
+opts{2}.banks.time.nFilters_per_octave = 1;
+opts{2}.banks.time.wavelet_handle = @gammatone_1d;
+opts{2}.banks.time.sibling_mask_factor = 2^3;
+opts{2}.banks.time.T = 2^17;
+opts{2}.banks.gamma.nFilters_per_octave = 1;
+opts{2}.banks.gamma.T = 2^nextpow2(nfo * 4);
+opts{2}.invariants.time.size = 2^15;
+opts{2}.invariants.time.T = 2^14;
+opts{2}.invariants.time.subscripts = 1;
+opts{3}.invariants.time.size = 2^15;
+opts{3}.invariants.time.T = 2^14;
+opts{3}.invariants.time.subscripts = 1;
+archs = sc_setup(opts);
+
+x = randn(2^15, 1);
+S = sc_propagate(x, archs);
+
+%%
+S1_refs = generate_refs(S{1+1}.data, 1, S{1+1}.ranges{1+0});
+for S1_ref_index = 1:size(S1_refs, 2)
+    S{1+1}.data = subsasgn(S{1+1}.data, S1_refs(:, S1_ref_index), S1_ref_index);
+end
+S2psi_refs = generate_refs(S{1+2}{1,1}.data, 1, S{1+2}{1,1}.ranges{1+0});
+for S2psiref_index = 1:size(S2psi_refs, 2)
+    S{1+2}{1,1}.data = subsasgn( ...
+        S{1+2}{1,1}.data, S2psi_refs(:, S2psiref_index), ...
+        S2psiref_index);
+end
+S2phi_refs = generate_refs(S{1+2}{1,2}.data, 1, S{1+2}{1,2}.ranges{1+0});
+for S2phi_ref_index = 1:size(S2phi_refs, 2)
+    S{1+2}{1,2}.data = subsasgn( ...
[…]

### 2016-07-01 experiments/sceneSvm/scsv1learn.m
update config user name; create scsvReport: add tex report repository;

@@ -11,7 +11,7 @@ function [config, store, obs] = scsv1learn(config, setting, data)
 % Date: 29-Jun-2016
 
 % Set behavior for debug mode
-if nargin==0, sceneSvm('do', 1, 'mask', {},'parallel',4); return; else store=[]; obs=[]; end
+if nargin==0, sceneSvm('do', 1, 'mask', {}); return; else store=[]; obs=[]; end
 
 expRandomSeed();
 
@@ -43,7 +43,7 @@ end
 acc = [];
 for k=1:5
     
-    trainSetting = ' -q ';
+    trainSetting = ' -q';
     
     switch setting.kernel
         case 'none'

### 2016-07-01 experiments/sceneSvm/scsvReport.m

@@ -1,11 +1,37 @@
-function config = scsvReport(config)                      
+function config = scsvReport(config)
 % scsvReport REPORTING of the expLanes experiment sceneSvm
-%    config = scsvInitReport(config)                      
-%       config : expLanes configuration state             
-                                                          
-% Copyright: Mathieu Lagrange                             
-% Date: 29-Jun-2016                                       
-                                                          
-if nargin==0, sceneSvm('report', 'r'); return; end        
-                                                          
-config = expExpose(config, 't', 'mask', {});                          
+%    config = scsvInitReport(config)
+%       config : expLanes configuration state
+
+% Copyright: Mathieu Lagrange
+% Date: 29-Jun-2016
+
+if nargin==0, sceneSvm('report', 'rcv'); return; end
+
+%% label
+
+setVal={'train','test'};
+
+%% get best setting
+
+set=1;
+config = expExpose(config, 't','fontSize','scriptsize','step', 1, 'mask', {0 0 0 set 0 0},'obs',1,'precision', 2);
+
+cut=1;
+select=2;
+gcaStyle='set(gca,''xtick'',1:6,''xticklabel'',[])';
+
+config = expExpose(config, 'p','fontSize','scriptsize','step', 1, 'mask', {0 0 cut set 0 select},'obs',1,'precision', 2, ...
+    'plotAxisProperties',{'fontsize',16},...
+    'plotCommand',{'legend(''mfcc, objBased, closest'',''mfcc, objBased, EMD'',''Scat, objBased, closest'',''Scat, objBased, EMD'',''mfcc, baseline'',''Scat, baseline'')', ...
+    'ylabel(''accuracy'')','ylim([.3 1])',gcaStyle,['title(''MFCC vs. Scattering using SVM on ' setVal{set} ''')']});
+
+
+%% run test set with best setting
+set=2;
+
+config = expExpose(config, 'p','fontSize','scriptsize','step', 1, 'mask', {0 0 cut set 0 select},'obs',1,'precision', 2, ...
+    'plotAxisProperties',{'fontsize',16},...
+    'plotCommand',{'legend(''mfcc, objBased, closest'',''mfcc, objBased, EMD'',''Scat, objBased, closest'',''Scat, objBased, EMD'',''mfcc, baseline'',''Scat, baseline'')', ...
+    'ylabel(''accuracy'')','ylim([.3 1])',gcaStyle,['title(''MFCC vs. Scattering using SVM on ' setVal{set} ''')']});
+

### 2016-07-05 paper/structureScene.tex
update paper
work on intro and section structure

@@
 
-Also, some recent works demonstrate the interest of such approaches for the characterization of human acoustic pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} \gl{What was the lavandier article ?} and the prediction of annoyance due to the traffic \cite{gloaguen}.
+Also, some recent works demonstrate the interest of such approaches for the characterization of human acoustic pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} and the prediction of annoyance due to the traffic \cite{gloaguen}.
 We believe that those case study are of interest for the signal processing community as they have strong societal impacts and raises interesting research avenues.
-Unfortunately, those fields of research are still in infancy and consequently very few well built datasets are available for evaluation purposes.
 
+Unfortunately, those fields of research are still in infancy and consequently very few well built datasets are readily available for evaluation purposes for the signal processing community.
 A closely related field of investigation that is relatively more mature is the classification of acoustic scenes whose aim is to predict labels by processing audio data.
-Despite its lower value, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and been considered in the data processing field for a while, with availability of some well designed evaluation datasets.
+Despite being narrower in terms of range of applications, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and been considered in the data processing field for a while, with availability of some well designed evaluation datasets.
+
+Despite the differences, the following question is valid both for the above cited future application and the acoustic scene classification (ASC).
+How can we efficiently and effectively process large amount of audio data using representations that are both compact, generic in their design and with a large application range ?
+
+The research described in this paper tackles that question by demonstrating the usefulness of two main contributions: 1) the use of the scattering transform allows us to extract discriminative features with almost no parameter optimization and 2) object-based modeling of the acoustic scene based on individual unsupervised clustering of each acoustic scene  allows us to effectively focus the description on areas of interest.
+Those contributions are extensively described, respectively in Section \ref{sec:scattering} and \ref{sec:object} and evaluated both in an unsupervised setting, \textit{i.e.} acoustic scene similarity (ASS) and in a supervised setting, \textit{i.e.}  acoustic scene classification (ASC), respectively in Section \ref{sec:similarity} and Section \ref{sec:classification}.
+
+paper
 
@@
 where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.
+% Advantage over PCA: we keep nonnegativity and the features remain uninterpretable.
 
@@
 
-\section{Unsupervised approach}
+\section{Object-based modeling of the scene}
+
+\subsection{Protocol}
+
+\subsection{Results}
+
+\section{Acoustic Scene Similarity}
 
@@
 
-\subsection{Results}
-
-GL
-
-\section{Supervised approach: support vector machines}
+\section{Acoustic Scene Classification}
 Support vector machines are large-margin binary discriminators with tunable regularization of misclassified examples.
 
+\subsection{Protocol}
+
 \subsection{Results}
-
-VL
 

### 2016-07-07 paper/structureScene.tex
raw edit of beginning of section 2

@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Scattering Auditory Scenes for Classification: Supervised and Unsupervised Approaches}
+\title{Object-based Auditory Scenes Similarity and Classification Using Scattering Features}
 
@@
 \begin{abstract}
-This paper introduces a 
+This paper introduces a generic hierarchical modeling of acoustic scenes using time/frequency scattering features for time spans below the second and object-based segmentation for larger time spans that effectiely model the scene with almost no requirement for optimization.
+Application of the proposed approach to the acoustic scene similarity and acoustic scene classification tasks demonstrates its potential.
+
+More precisely, the proposed approach achieves 
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Acoustic scene classification, segmentation.
+Acoustic scene classification, acoustic scene segmentation.
 \end{IEEEkeywords}
@@
 
-Despite the differences, the following question is valid both for the above cited future application and the acoustic scene classification (ASC).
+Despite the differences, the following question is valid both for the above cited future application and the acoustic scene classification (ASC) task.
 How can we efficiently and effectively process large amount of audio data using representations that are both compact, generic in their design and with a large application range ?
@@
 The research described in this paper tackles that question by demonstrating the usefulness of two main contributions: 1) the use of the scattering transform allows us to extract discriminative features with almost no parameter optimization and 2) object-based modeling of the acoustic scene based on individual unsupervised clustering of each acoustic scene  allows us to effectively focus the description on areas of interest.
-Those contributions are extensively described, respectively in Section \ref{sec:scattering} and \ref{sec:object} and evaluated both in an unsupervised setting, \textit{i.e.} acoustic scene similarity (ASS) and in a supervised setting, \textit{i.e.}  acoustic scene classification (ASC), respectively in Section \ref{sec:similarity} and Section \ref{sec:classification}.
+After studying the state of the art in the field of acoustic scene modeling in Section \ref{sec:soa}, those contributions are extensively described, respectively in Section \ref{sec:scattering} and \ref{sec:object} and evaluated both in an unsupervised setting, \textit{i.e.} acoustic scene similarity (ASS) and in a supervised setting, \textit{i.e.}  acoustic scene classification (ASC), respectively in Section \ref{sec:similarity} and Section \ref{sec:classification}.
 
-paper
+\section{Previous work} \label{sec:soa}
 
-In this paper, we investigate the potential of a new unsupervised segmentation paradigm that allows us to
+As stated in the introduction, the field of eco-acoustics is relatively young, but significant works are worth mentioning.
+One avenue of research issued from bio acoustic studies, consider the audio stream as a set of events produced animals, and is processed to retrieve the precise species that produced the sound and potentially the type of vocalization in order to predict some behavior of the animal under study \cite{dan}.
+Another, more recent, avenue of research, process the audio stream in an holistic manner in order to infer large time-scale properties of the bio-acoustic scene for example to estimate which part of the stream need to be analyzed in details, \textit{i.e.} by a human \cite{australians}.
 
-This data has to be processed
+The same dichotomy applies for the field of audio scene analysis, where the interest can be whether 1) to isolate and identify precise events, \textit{i.e.} acoustic event detection (AED) or 2) to predict global attributes assigned to large time spans of audio, \textit{i.e.} ASC.
+For the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag} where the audio is modelled as high level statistics comptued from a set of local features.
+Even though this approach as the advantage of high genericity and follows some evidence found by psycho perception research \cite{macdermott}, for the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be equivalent to  \cite{lagrange:hal-01082501}.
 
-Environment sound
+citer papier event based de l'intro de jj
 
-describe in terms of high level attributes
+ there is evidence that modelling the scene as a set of events allows us to predict high level semantic information such as pleasantness
 
-be they cognitive ones such as pleasantness, or descriptive ones such as the type of environment where the
-
-%As part of the aforementioned research areas and applications, the emerging field of \emph{Acoustic Scene Analysis} (also called \emph{Sound Scene Analysis}) \cite{Stowell15} aims to develop approaches and systems for the automatic analysis of environmental sounds and soundscapes (originating both from urban or nature environments).
-While research methodologies in related fields such as Automatic Speech Recognition (ASR) \cite{Rabiner93} and Music Information Retrieval (MIR) \cite{Muller07} are now well established, research addressing Acoustic Scene Analysis remains relatively young.
-
-Whilst the range of applications are very large, so is the need
-
-holistic / event based
-
-skeleton of events on a bed of texture.
-
-debate about the kind of process underlying perception
-
-strong indication that the detection of specific events, named markers suffice to trigger class prediction
-
-From a numerical data processing overview, the holistic scheme as a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
-
-If available, the complete description of the scene in terms of event occurences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas.
-
-\section{Previous work}
-
-\gl{bag of frame approach:} \cite{aucouturier2007bag}: bag of frame approach. \cite{lagrange:hal-01082501}: why the bag of frame approach does no work. \\
+Eventhough the precise list of events that occur in a given time is unachievable to get even for trained human, research shows that using only a few events, so called markers \cite{} is enough to reliably predict high level attributes.
+For example, the presence of a few bird calls in a smooth traffic hubhub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
 
@@
 
+describe in terms of high level attributes
+
+be they cognitive ones such as pleasantness, or descriptive ones such as the type of environment where the
+
+%As part of the aforementioned research areas and applications, the emerging field of \emph{Acoustic Scene Analysis} (also called \emph{Sound Scene Analysis}) \cite{Stowell15} aims to develop approaches and systems for the automatic analysis of environmental sounds and soundscapes (originating both from urban or nature environments).
+While research methodologies in related fields such as Automatic Speech Recognition (ASR) \cite{Rabiner93} and Music Information Retrieval (MIR) \cite{Muller07} are now well established, research addressing Acoustic Scene Analysis remains relatively young.
+
+Whilst the range of applications are very large, so is the need
+
+holistic / event based
+
+skeleton of events on a bed of texture.
+
+debate about the kind of process underlying perception
+
+strong indication that the detection of specific events, named markers suffice to trigger class prediction
+
+From a numerical data processing overview, the holistic scheme as a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
+
+If available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas.
+
+\section{Acoustic Scene Similarity}
+
 \subsection{Protocol}
@@
 \subsection{Results}
-
-\section{Acoustic Scene Similarity}
-
-GL
 

### 2016-07-07 paper/structureScene.tex
page budget
typo

@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Object-based Auditory Scenes Similarity and Classification Using Scattering Features}
+\title{Object-based Auditory Scenes Similarity Retrieval and Classification With Scattering Transforms}
 
@@
 This is a form of data augmentation, because $5$ monophonic signals are derived from each binaural recording in the training set.
-At test time, only the center mix ($\alpha = 0$) is used to perform the class.
+At test time, only the center mix ($\alpha = 0$) is used to predict the class.
 
@@
 
+\ml{GL 2 colonnes}
+
 describe in terms of high level attributes
@@
 
+\ml{GL 1 colonne}
+
 \subsection{Protocol}
@@
 
+\ml{VL 2 colonnes}
+
 \subsection{Protocol}
@@
 
-\section{Discussion}
-
-ML
-
 \section{Conclusion}
 
-ML
+\ml{ML 1/2 colonne}
 

### 2016-07-08 paper/structureScene.tex
proof read of scattering sections

@@
 Scattering transforms are translation-invariant representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
-Depending upon the chosen architecture, the modulation filterbank may either be performed solely over the time variable or on both time and frequency.
+Depending upon the chosen architecture, the modulation filterbank may either be performed solely over the time variable or on both time and frequency variables.
 In this section, we present both the temporal scattering transform and the time-frequency scattering transform.
@@
 frequencies than the CQT, at the expense of a greater computational cost.
-This property allows to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+This property allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
 
@@
 They also account for some higher-level cognitive attributes of sound, \eg prosody in speech or rhythm in music.
+\ml{je ne parlerai pas de cognition a ce stade, je dirais plutot que les modulations s'expriment aussi bien au niveau physique à l'échelle fine (interactions mecaniques) qu'a une echelle temporelle plus large (prososdie, rhythme).
+Apres, si tu as des citations, tu peux dire que ce sont des indices importants pour la perception, ce qui est le cas}
 Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a translation-invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
@@
 In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Finally, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation, yielding
+Finally, \ml{finnaly a un sens particulier en Anglais, il implique une sorte de soulagement.
+Dans des documents techniques on met plutot lastly} the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation, yielding
 \begin{equation}
@@
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
+\ml{tu peux citer le travail publié a Nature de lewicki, qui montrent que ces formes se retrouvent par ICA de sons naturels http://www.cs.cmu.edu/~learning/talks-2002/02-10-21.lewicki.pdf}
 
@@
 The most salient events among auditory scenes draw highly structured patterns in the time-frequency plane.
-\vl{missing sentence here ?}
+\vl{missing sentence here ? oui, en lien avec les strf, tu peux donner des exmeples d'orientation de modulation frequentielle}
 Despite its invariance and stability properties, the temporal scattering transform fails to characterize these patterns, because it extracts the amplitude modulations of each wavelet subband separately.
@@
 Therefore, wavelets on $\gamma_1$ have center frequencies of the form
-$\theta_{\mathrm{fr}}  2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}$, where $\gamma_{\mathrm{fr}}$ takes integer values betwen $0$ and $(J_\mathrm{fr} \times Q_\mathrm{fr} - 1)$ and $\theta_{\mathrm{fr}} = \pm 1$.
+$\theta_{\mathrm{fr}}  2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}$, where $\gamma_{\mathrm{fr}}$ takes integer values between $0$ and $(J_\mathrm{fr} \times Q_\mathrm{fr} - 1)$ and $\theta_{\mathrm{fr}} = \pm 1$.
 \begin{equation}
@@
 \section{Feature design}
-Before supervised or unsupervised classification, is it beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power. 
-This section reviews binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.
-The former is novel and specific to stereophonic recordings, whereas the next ones would also apply to many other kinds of data.
+Before supervised or unsupervised classification, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power. 
+This section reviews \ml{several means of achieving those goals:} binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.
+\ml{The former has, to the best of our knowledge, not being considered for ASC and specific to multichannel recordings}, whereas the next ones would also apply to many other kinds of data.
 
@@
 
-In order to achieve invariance to ILD, averaging the left and right channels into a monophonic signal is by far the most widespread approach.
-However, it favors the center of the scene while attenuating lateral cues.
-Moreover, sources that are on distinct azimuths get mixed, which may cause bias at training time.
+In order to achieve invariance to ILD, averaging the left and right channels into a monophonic signal is by far the most widespread approach \ml{for classification purposes}.
+However, it favors the center of the scene while attenuating lateral cues and sources that are on distinct azimuths get mixed, which may cause bias at training time.
 
@@
 where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.
-% Advantage over PCA: we keep nonnegativity and the features remain uninterpretable.
+% Advantage over PCA: we keep nonnegativity and the features remain interpretable.
 
@@
 Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
-Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design, and their distribution is typically skewed towards the right.
+Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design, and their distribution is typically skewed towards the right \ml{c'est quoi la droite, sans figure c'est pas facile de voir de quoi tu parles}.
 This issue can however be addressed by applying to all coefficients a pointwise concave transformation, \eg logarithmic, hence reducing skewness.
@@
 \subsection{Early vs. late temporal integration}
-Owing to the scarcity of salient events in many natural soundscapes, fine-grained classification of auditory scenes is only made
+Owing to the scarcity of salient events in many natural \ml{sound scenes soundscapes est un modele de perception de la scene}, fine-grained classification of auditory scenes is only made
 possible by integrating signal information over a long temporal context.
@@
 
-The principal drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene -- if we don't account for data augmentation.
+The principal drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene -- if data augmentation is not considered.
 In the context of the DCASE 2013 dataset, this corresponds to merely $10$ training instances per class, and $100$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
@@
 We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
+
+\ml{Superbe section !!}
 

### 2016-07-08 paper/structureScene.tex
\vl{Update number of instances with new train / test split}

@@
 On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene, resulting in $12800$ instances overall.
+\vl{Update number of instances with new train / test split}
 However, many of these instances may be silent or lack any salient properties of the category they are assigned to, hence an increase in bias and a risk of underfitting.

### 2016-07-08 paper/structureScene.tex
update section IV intro

@@
 Before supervised or unsupervised classification, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power. 
-This section reviews \ml{several means of achieving those goals:} binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.
-\ml{The former has, to the best of our knowledge, not being considered for ASC and specific to multichannel recordings}, whereas the next ones would also apply to many other kinds of data.
+In this section, we review several ways of achieving these properties, namely binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.
+The former has, to the best of our knowledge, not being considered for ASC and is specific to multichannel recordings, whereas the next ones would also apply to many other kinds of data.
 
@@
 We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
-
-\ml{Superbe section !!}
 

### 2016-07-08 paper/structureScene.tex
update IV.A. Request for comments

@@
 In particular, the interaural level difference (ILD) between left and right channel conveys the approximate azimuth of each sound source that make up the scene with respect to the listener \cite{Blauert2003}.
-Yet, since the location of the recordist varies across instances of the same class, the feature space for classification should be invariant to these azimuths.
+Yet, the location of the recordist may vary across instances of the same class.
+Therefore, any global rotations of all azimuths, corresponding to a rotation of the head of the recordist, should be discarded as a spurious source of variability in the observed data.
 
-In order to achieve invariance to ILD, averaging the left and right channels into a monophonic signal is by far the most widespread approach \ml{for classification purposes}.
-However, it favors the center of the scene while attenuating lateral cues and sources that are on distinct azimuths get mixed, which may cause bias at training time.
+For classification purposes, averaging the left and right channels into a monophonic signal is by far the most widespread approach, because the resulting monophonic signal is approximately invariant to azimuth.
+However, it favors the center of the scene while attenuating lateral cues.
+Moreover, sources that are on distinct azimuths get mixed, which may cause bias at training time.
+\vl{qu'en penses-tu ML ?}
 

### 2016-07-08 paper/structureScene.tex
write more on IV.B

@@
 This fact may cause the classifier to overfit the training set, especially when the number of features is large and the number of training instances is small, as is the case here.
+Selecting features appropriately before the training stage may circumvent this problem.
+However, the parameters of the chosen feature selection algorithm are themselves prone to overfitting.
+Instead, we adopt the more conservative approach of retaining all coefficients within a certain range of subband indices $\gamma_1$ and $\gamma_2$.
+In comparison with adaptive algorithms, the subset of selected features $\gamma$ has a suboptimal statistical relevance, but it is easily interpretable in the realm of audio signal processing, as it is equivalent to a coarse band-pass filtering of $\boldsymbol{x}(t)$ and $\boldsymbol{x_1}(t,\gamma_1)$ over the time dimension $t$.
 
-This is inspired by the state-of-the-art results of \cite{roma2013}, which rely on cepstral descriptors computed up to a cutoff frequency as small as 900 Hz.
+The conservation equations underlying the physics of sound production constrain the waveform $\boldsymbol{x}(t)$ to be regular in the time domain, that is, to have a polynomial decay in the Fourier domain.
+This decay is only made sharper by the specific properties of acoustic environments, especially indoor.
+As a result, the amount of energy in the wavelet subband $\gamma_1$ is roughly proportional to the center frequency $2^{-\gamma_1/Q_1}$.
+
+This is consistent with the results of \cite{roma2013}, who secured the first place in the 2013 edition of the DCASE challenge, by relying on cepstral descriptors computed up to a cutoff frequency as small as 900 Hz.
 

### 2016-07-08 paper/structureScene.tex
finish writing up IV.B

@@
 
-\subsection{Feature selection}
+\subsection{Feature selection by restricting the frequency range}
 In a multivariate time series of descriptors $\mathbf{S}\boldsymbol{x}(t,\gamma)$, some coefficients $\gamma$ may consistently bear negligible information for the task of interest.
@@
 This decay is only made sharper by the specific properties of acoustic environments, especially indoor.
-As a result, the amount of energy in the wavelet subband $\gamma_1$ is roughly proportional to the center frequency $2^{-\gamma_1/Q_1}$.
+Consequently, the amount of energy in the wavelet subband $\gamma_1$ is roughly inversely proportional to the center frequency $2^{-\gamma_1/Q_1}$.
+In the topmost frequencies, this amount is so faint that the corresponding coefficients $\boldsymbol{x_1}(t,\gamma_1)$ and $\boldsymbol{x_2}(t,\gamma_1,\gamma_2)$ no longer bear discriminative information for the classification task at hand.
 
-This is consistent with the results of \cite{roma2013}, who secured the first place in the 2013 edition of the DCASE challenge, by relying on cepstral descriptors computed up to a cutoff frequency as small as 900 Hz.
-
-The starting scale $\gamma_{1,\min}$ corresponding to some upper bound $\xi_{1,\max}$ on the frequency axis is given by
+In the sequel, we evaluate the impact of discarding the scattering coefficients above a certain acoustic frequency $\xi_{1,\max}$, which corresponds to discarding subbands $\gamma_1$ below a certain scale
 \begin{equation}
 \gamma_{1,\min} =
-\left\lfloor Q_1 \log_2 \frac{\xi_1}{\xi_{1,\max}} \right\rceil
+\left\lfloor Q_1 \log_2 \frac{\xi_1}{\xi_{1,\max}} \right\rceil,
 \end{equation}
 where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.
-% Advantage over PCA: we keep nonnegativity and the features remain interpretable.
+
+This strategy is already widespread in the literature.
+For instance, we should note that \cite{roma2013}, who secured the first place in the 2013 edition of the DCASE challenge, rely on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max}  = 900\,\mathrm{Hz}$ instead of the more conventional value of $16\,\mathrm{kHz}$.
 

### 2016-07-08 paper/structureScene.tex
raw conclusion

--- 2016-07-08 paper/structureScene.tex
@@
 
-\ml{ML 1/2 colonne}
+In this paper, the benefit of considering 1) the scattering transform for building discriminant features and 2) unsupervised clustering for the modeling of large time span environmental acoustic scenes.
+Experiments shows that the time/frequency scattering achieves better discriminability than the time scattering one or MFCCs as they probably better isolate elements of interest in different information channels, thus allowing an efficient and scalable early integration process.
+This leads to improvement both for the ASSR and ASC tasks.
+
+Abstracting the scene as a reduced set of exemplars with adapted similarity strategies allows us to improve ASSR performance when the BOF approaches based on GMMs fails with reference to an early integration scheme.
+Those results advocate for a hierarchical model of the scene, where small time scale structure  (below a second) shall be decorelated using convolutional network such as the ones of the scattering transforms and larger time scale (over a second) structure shall be modeled as adaptive aggregates that can be identified in the time domain in an unsupervised setting and in the features domain in the supervised case (provided that the dimensionality is sufficiently high).
+
+This assertion shall be challenged in future work by considering larger databases and more diverse application scenarios, such as pleasantness prediction \cite{acta} or ecoacoustics tasks.
 

### 2016-07-08 figs/compression/compression.m
two subfigs rather than one fig with subplots.

@@ -18,15 +18,15 @@ X = X(:, :, 3, :);
 X = X(:);
 X = max(X, 0);
 
-clf();
-subplot(121);
+%%
+figure(1);
 histogram(X, nBins, 'EdgeColor', 'none', 'FaceColor', color, ...
     'FaceAlpha', 1.0);
 axis off
-subplot(122);
+export_fig -transparent feature_histogram.png
+
+figure(2);
 histogram(log(X), nBins, 'EdgeColor', 'none', 'FaceColor', color, ...
     'FaceAlpha', 1.0);
 axis off
-
-%%
-export_fig -transparent compressed_histogram.png
+export_fig -transparent feature_histogram_logcompressed.png

### 2016-07-08 paper/structureScene.tex
include draft histogram fig in section IV.3
update number of instances with new train/test split

@@
 
+\begin{figure}
+\begin{center}
+\setlength{\unitlength}{1cm}
+\begin{picture}(5,2)
+ \put(-1.5,0.0){\includegraphics[height=2cm,width=3.5cm]{feature_histogram.png}}
+ \put(2.5,0.0){\includegraphics[height=2cm,width=3.5cm]{feature_histogram_logcompressed.png}}
+\end{picture}
+\caption{
+\label{fig:histograms}
+\vl{we should discuss color schemes and axis labels}}
+\end{center}
+\end{figure}
+
 In particular, logarithmic compression is ubiquitous in audio signal processing, as it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
@@
 \subsection{Early vs. late temporal integration}
-Owing to the scarcity of salient events in many natural \ml{sound scenes soundscapes est un modele de perception de la scene}, fine-grained classification of auditory scenes is only made
+Owing to the scarcity of salient events in many natural scenes,
+fine-grained classification is only made
 possible by integrating signal information over a long temporal context.
@@
 The principal drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene -- if data augmentation is not considered.
-In the context of the DCASE 2013 dataset, this corresponds to merely $10$ training instances per class, and $100$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
-On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene, resulting in $12800$ instances overall.
-\vl{Update number of instances with new train / test split}
+In the context of the DCASE 2013 dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
+On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene, resulting in $10240$ instances overall.
 However, many of these instances may be silent or lack any salient properties of the category they are assigned to, hence an increase in bias and a risk of underfitting.

### 2016-07-08 paper/structureScene.tex
update IV. C

@@
 For instance, we should note that \cite{roma2013}, who secured the first place in the 2013 edition of the DCASE challenge, rely on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max}  = 900\,\mathrm{Hz}$ instead of the more conventional value of $16\,\mathrm{kHz}$.
+\vl{Qu'en penses-tu ML ?}
 
@@
 Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
-Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design, and their distribution is typically skewed towards the right \ml{c'est quoi la droite, sans figure c'est pas facile de voir de quoi tu parles}.
-This issue can however be addressed by applying to all coefficients a pointwise concave transformation, \eg logarithmic, hence reducing skewness.
+Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design.
+It appears empirically that their distribution is skewed towards the right, which means that the tail towards greater values is longer than the tail towards lower values.
+However, skewness can be reduced by applying to all coefficients a pointwise concave transformation, \eg logarithmic.
+Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
+\vl{Qu'en penses-tu ML ?}
 
@@
 
-In particular, logarithmic compression is ubiquitous in audio signal processing, as it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
+The application of the pointwise logarithm to magnitude spectra is ubiquitous in audio signal processing.
+Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
 In the context of auditory scenes, we must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone, which is a spurious factor a variability to the task of auditory scene classification.

### 2016-07-08 paper/structureScene.tex
\newcommand*{\vs}{vs.\@\xspace}

@@
 \newcommand*{\ie}{i.e.\@\xspace}
+\newcommand*{\vs}{vs.\@\xspace}
 
@@
 
-\subsection{Early vs. late temporal integration}
+\subsection{Early \vs late temporal integration}
 Owing to the scarcity of salient events in many natural scenes,
@@
 
-\ml{VL 2 colonnes}
+\subsection{Role of logarithmic compression}
 
-\subsection{Protocol}
-
-\subsection{Results}
+\subsection{Time scattering \vs time-frequency scattering}
 

### 2016-07-08 paper/structureScene.tex
subsections of Section VII
reformulate section on  log compression

@@
 Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
-In the context of auditory scenes, we must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone, which is a spurious factor a variability to the task of auditory scene classification.
+We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone, which is a spurious factor of variability to the task of auditory scene classification.
 Logarithmic compression can linearize this dependency, which arguably facilitates the construction of a powerful invariant at the classifier stage.
@@
 a musical instrument, or a genre, it may require up to 30 seconds
-to disambiguate auditory scenes which share part of their semantic content, \eg a train from a subway station or a quiet street from a park.
+to disambiguate two classes auditory scenes which share part of their semantic content, \eg a train from a subway station or a quiet street from a park.
 Depending on whether aggregation is performed in feature space or in decision space, the corresponding method is referred to as early or late integration.
@@
 On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene, resulting in $10240$ instances overall.
-However, many of these instances may be silent or lack any salient properties of the category they are assigned to, hence an increase in bias and a risk of underfitting.
+However, many of these instances may be silent or lack any salient properties of the class they are assigned to, hence an increase in bias and a risk of underfitting.
 In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff.
@@
 
+\subsection{Early \vs late integration}
+
+\subsection{Role of data augmentation}
+
+\subsection{Role of frequency range}
+
 \section{Conclusion}

### 2016-07-08 experiments/supervised/dcase_script.py

@@ -19,17 +19,14 @@ np.set_printoptions(precision=2)
 
 
 
-# Evaluate role of compression
-exp1 = []
-for C in 2**np.arange(-1, 4).astype(np.float):
-    exp1.append(dcase_svm(0, 12, False, "time", "none", "early", C))
-    exp1.append(dcase_svm(0, 12, False, "time", "log", "early", C))
-    exp1.append(dcase_svm(0, 12, False, "time", "logmedian", "early", C))
+C = 1
+exp1.append(dcase_svm(0, 12, False, "time", "logmedian", "early", C))
+exp1.append(dcase_svm(0, 12, True, "time", "logmedian", "early", C))
+exp1.append(dcase_svm(0, 12, False, "timefrequency", "logmedian", "early", C))
+exp1.append(dcase_svm(0, 12, True, "timefrequency", "logmedian", "early", C))
 
-# Evaluate role of data augmentation
-exp2 = []
-for C in 2**np.arange(-1, 4).astype(np.float):
-    exp2.append(dcase_svm(0, 12, False, "time", "log", "early", C))
-    exp2.append(dcase_svm(0, 12, False, "time", "logmedian", "late", C))
-    exp2.append(dcase_svm(0, 12, True, "time", "log", "early", C))
-    exp2.append(dcase_svm(0, 12, True, "time", "logm", "late", C))
+
+exp1.append(dcase_svm(0, 11, True, "timefrequency", "logmedian", "early", C))
+exp1.append(dcase_svm(0, 10, True, "timefrequency", "logmedian", "early", C))
+exp1.append(dcase_svm(0,  9, True, "timefrequency", "logmedian", "early", C))
+exp1.append(dcase_svm(0,  8, True, "timefrequency", "logmedian", "early", C))

### 2016-07-08 experiments/supervised/dcase_svm.py

@@ -37,9 +37,6 @@ def dcase_svm(octmin, octmax, augmentation, scattering, compression, integration
     freq_indices = (freqs > fmin) & (freqs < fmax)
     X = X[:, :, :, freq_indices]
 
-    # Rectify
-    X = np.maximum(0.0, X)
-
     # Split folds
     X = np.reshape(X,
         (X.shape[0] / 5,   #  20 scenes per fold
@@ -95,8 +92,8 @@ def dcase_svm(octmin, octmax, augmentation, scattering, compression, integration
             X_training = np.log1p(1e2 * X_training / medians)
             X_test = np.log1p(1e2 * X_test / medians)
         if compression == "log":
-            X_training = np.log(1e-6 + X_training)
-            X_test = np.log1p(1e-6 + X_test)
+            X_training = np.log(1e-6 + np.maximum(X_training, 0))
+            X_test = np.log1p(1e-6 + np.maximum(X_test, 0))
 
         # Standardize features
         scaler = sklearn.preprocessing.StandardScaler().fit(X_training)

### 2016-07-08 figs/compression/compression.m

@@ -5,9 +5,9 @@ load(features_path)
 
 %% Draw a histogram
 % Pick a scattering subband index
-scattering_index = 50;
-nBins = 50;
-freq = dcase2013_timeQ8_test.freqs(scattering_index)
+scattering_index = 90;
+nBins = 100;
+freq = dcase2013_timeQ8_test.freqs(scattering_index);
 color = [0 51 204] / 256;
 
 X = dcase2013_timeQ8_test.X_test;
@@ -16,17 +16,22 @@ X = X(scattering_index, :, :, :);
 % Center azimuth only
 X = X(:, :, 3, :);
 X = X(:);
-X = max(X, 0);
 
-%%
+%
 figure(1);
 histogram(X, nBins, 'EdgeColor', 'none', 'FaceColor', color, ...
     'FaceAlpha', 1.0);
 axis off
-export_fig -transparent feature_histogram.png
+%export_fig -transparent feature_histogram.png
 
 figure(2);
-histogram(log(X), nBins, 'EdgeColor', 'none', 'FaceColor', color, ...
+histogram(log(max(X,0)), nBins, 'EdgeColor', 'none', 'FaceColor', color, ...
     'FaceAlpha', 1.0);
 axis off
-export_fig -transparent feature_histogram_logcompressed.png
+%export_fig -transparent feature_histogram_logcompressed.png
+
+figure(3);
+histogram(log1p(X / median(X)), ...
+    nBins, 'EdgeColor', 'none', 'FaceColor', color, ...
+    'FaceAlpha', 1.0);
+axis off

### 2016-07-08 paper/structureScene.tex
remove section III. C

@@
 
-\subsection{Time-frequency scattering}
-The most salient events among auditory scenes draw highly structured patterns in the time-frequency plane.
-\vl{missing sentence here ? oui, en lien avec les strf, tu peux donner des exmeples d'orientation de modulation frequentielle}
-Despite its invariance and stability properties, the temporal scattering transform fails to characterize these patterns, because it extracts the amplitude modulations of each wavelet subband separately.
-
-This issue was remedied by time-frequency scattering \cite{Anden2015}, which performs a wavelet decomposition over the log-frequency variable $\gamma_1$ after the second-order wavelet decomposition over time.
-While temporal scattering merely extracts the average amplitude modulation spectrum, time-frequency scattering computes spectral modulations at various temporal rates and spectral scales.
-% link with GBFB and STRF which have proven useful for acoustic scene classification.
-
-Since it operates over a complex-valued input, the filterbank over log-frequencies $\gamma_1$ must cover the whole Fourier spectrum, including negative frequencies and the null frequency, to ensure invertibility and energy conservation.
-Therefore, wavelets on $\gamma_1$ have center frequencies of the form
-$\theta_{\mathrm{fr}}  2^{-\gamma_{\mathrm{fr}}/Q_{\mathrm{fr}}}$, where $\gamma_{\mathrm{fr}}$ takes integer values between $0$ and $(J_\mathrm{fr} \times Q_\mathrm{fr} - 1)$ and $\theta_{\mathrm{fr}} = \pm 1$.
-\begin{equation}
-\boldsymbol{x_2}(t,\gamma_1,\gamma_2,\gamma_{\mathrm{fr}},\theta_{\mathrm{fr}}) =
-\vert \boldsymbol{x_1}
-\overset{t}{\ast} \boldsymbol{\psi_{\gamma_2}}
-\overset{\gamma_1}{\ast} \boldsymbol{\psi_{\gamma_{\mathrm{fr}}, \theta_{\mathrm{fr}}}}
-\vert (t,\gamma_1)
-\end{equation}
-Again, we use the generic literal $\gamma$ as a shorthand for $\gamma_1$ or $(\gamma_1, \gamma_2, \gamma_{\mathrm{fr}}, \theta_{\mathrm{fr}})$.
-
 \section{Feature design}

### 2016-07-08 paper/structureScene.tex
remove subsection on binaural data augmentation

@@
 
-\subsection{Binaural data augmentation}
-Most acoustic scene datasets are recorded according to a binaural protocol, \ie with a pair of in-ear microphones \cite{Wang2006}.
-This protocol provides a realistic description of the spatial auditory environment, as it reproduces the natural listening conditions of humans.
-In particular, the interaural level difference (ILD) between left and right channel conveys the approximate azimuth of each sound source that make up the scene with respect to the listener \cite{Blauert2003}.
-Yet, the location of the recordist may vary across instances of the same class.
-Therefore, any global rotations of all azimuths, corresponding to a rotation of the head of the recordist, should be discarded as a spurious source of variability in the observed data.
-
-For classification purposes, averaging the left and right channels into a monophonic signal is by far the most widespread approach, because the resulting monophonic signal is approximately invariant to azimuth.
-However, it favors the center of the scene while attenuating lateral cues.
-Moreover, sources that are on distinct azimuths get mixed, which may cause bias at training time.
-\vl{qu'en penses-tu ML ?}
-
-Instead, so as to leverage stereophonic data, we perform multiple combinations of the left channel $\boldsymbol{x_{\mathsf{L}}}(t)$ and right channel $\boldsymbol{x_{\mathsf{R}}}(t)$ into monophonic signals $\boldsymbol{x_{\alpha}}(t)$ according to the equation
-\begin{equation}  
-\boldsymbol{x_{\alpha}}(t) =
-\dfrac{1+\alpha}{2} \boldsymbol{x_\mathsf{L}}(t) +
-\dfrac{1-\alpha}{2} \boldsymbol{x_\mathsf{R}}(t),
-\end{equation}
-where $\alpha$ is a panoramic mixing weight, taking values between $-1$ and $+1$, which could be deterministic or random.
-
-In following experiments, we deterministically set $\alpha$ to the values $-1$, $\frac{1}{2}$, $0$, $+\frac{1}{2}$, and $+1$.
-This is a form of data augmentation, because $5$ monophonic signals are derived from each binaural recording in the training set.
-At test time, only the center mix ($\alpha = 0$) is used to predict the class.
-
 \subsection{Feature selection by restricting the frequency range}

### 2016-07-08 paper/structureScene.tex
At a longer scale, they also account for some higher-level attributes of sound, \eg prosody in speech or rhythm in music.

@@
 Among auditory scenes, amplitude modulations may be caused by a broad variety of mechanical interactions, including collision, friction, and turbulent flow.
-They also account for some higher-level cognitive attributes of sound, \eg prosody in speech or rhythm in music.
-\ml{je ne parlerai pas de cognition a ce stade, je dirais plutot que les modulations s'expriment aussi bien au niveau physique à l'échelle fine (interactions mecaniques) qu'a une echelle temporelle plus large (prososdie, rhythme).
-Apres, si tu as des citations, tu peux dire que ce sont des indices importants pour la perception, ce qui est le cas}
+At a longer scale, they also account for some higher-level attributes of sound, \eg prosody in speech or rhythm in music.
 Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a translation-invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
@@
 In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Finally, \ml{finnaly a un sens particulier en Anglais, il implique une sorte de soulagement.
-Dans des documents techniques on met plutot lastly} the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation, yielding
+Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation, yielding
 \begin{equation}

### 2016-07-09 paper/structureScene.tex
begin the redaction of the object based approach

@@
 
+It has been shown the perception of complex acoustic scene rely on the identification of putative sound sources.
 \ml{GL 2 colonnes}
@@
 
+To take into account this object-based representation, we propose the following 3-steps algorithm:
+
+\begin{itemize}
+\item step 1: features extraction.
+\item step 2: scene-wise clustering step.
+A clustering is performed on each scene using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
+Each scene is then describe with a set of centroids.
+\item step 3: Centroids similarity.
+The similarity between the centroids is computed using a radial basis function kernel $K^c$ as well as the local scaling method proposed in \cite{selfTuneManor2004}: \\
+\begin{equation}
+\begin{split}
+K^c_{ij} & =exp\left( - \dfrac{\parallel x_i - x_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
+\sigma_{i} & =\parallel x_i - x_N \parallel
+\end{split}
+\end{equation}
+
+with $x_N$ the $N$'th nearest neihgbor of the centroid $x_i$. \\
+
+\item step 4: scene similarity \\
+\end{itemize}
+
+To compute the scene similarity based on the extracted centroids, we propose the three following strategies:
+
+\begin{itemize}
+\item \emph{closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
+\item \emph{averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
+\item \emph{weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster. 
+\end{itemize}
+
+To compute the  weighted similarity between two scenes by taking into account both theirs clusters weights and theirs centroids similarities, we use a variant of the Earth Moving Distance ($EMD$) adapted for non-normalized histograms known as $\widehat{EMD}$ \cite{pele2008linear} with the implementation proposed in  \cite{pele2009fast}.
+Considering two histograms $h_i$ and $h_j$, represented the centroids weights of each scene, the $\widehat{EMD}$ is computed as follow:
+
+\begin{equation}
+\begin{split}
+\widehat{EMD}(h_i,h_j) &=( \min\limits_{f_{ij}} \sum\limits_{i,j} f_{ij}D_{ij} )  \\ 
+&+ \mid \sum\limits_{i} h_i - \sum\limits_{j} h_j  \mid \alpha \max\limits_{i,j} D_{ij}
+\end{split}
+\end{equation}
+
+\begin{equation*}
+s.t. \quad f_{ij}\geq0\quad \sum\limits_{j} f_{ij} \leq h_{i}\quad \sum\limits_{i} f_{ij} \leq h_{j} 
+\end{equation*}
+\begin{equation*}
+\sum\limits_{i,j}f_{ij} = \min( \sum\limits_{i} h_{i} ,\sum\limits_{j} h_{j} )
+\end{equation*}
+
+where $f_{ij}$ is the flow, and $D_{ij}$ is the ground distance. In our case the ground distance is $1-K^c_{ij}$
+
+To get the final similarity measure between two scenes $s_i$ and $s_j$, we use an extended Gaussian kernels $K^s$ (Chapelle et al., 1999; Jing et al., 2003).
+
+\begin{equation*}
+K^s(s_i,s_j) = exp\left( - \dfrac{\widehat{EMD}(h_i,h_j)}{A} \right) \\
+\end{equation*}
+
+with $A$ a scaling parameter, set to the mean value of the $\widehat{EMD}$. \\
+
+The resulting kernel is an EMD kernel. It has to be noted that there is no proof that such kernel is positive definite.
+
+\section{Experimental Evaluation}
+
+\subsection{Acoustic Scene Similarity and Classification}
+
+The object-based approach is assessed considering both an ASS and an ASC scenario:
+
+\begin{itemize}
+\item \emph{ASS}: It is an unsupervised scenario.
+Evaluation is performed on the private dataset of the DCASE 2013 challenge, using the precision at rank $k$ ($k=[3,\ldots,9]$) metrics.
+Similarity between scenes features is then computed using a Gaussian kernel with local scaling.
+
+\item \emph{ASC}: It is a supervised scenario.
+The systems are evaluated  with five-fold  stratified cross validation on the  private dataset of the DCASE 2013 challenge.
+The folds are identical to those used in the original challenge.
+A linear Support Vector Machine (SVM) is used as classifier.
+For object-based approach, we only consider the step 2.
+Consequently, each scene is represented by a set of centroids and each centroid is classified separately.
+The final scene classification is based on a majority vote.
+In this case, the object-based approach may be seen as a mid integration strategy.
+\end{itemize}
+
+The object-based approach is compared to standard approaches using early integration strategy for the ASS scenario, and both early and late integration strategies for the ASC scenario (see Section \ref{XX}).
+
+\subsection{Features}
+
+Two features are used to describe the scenes: temporal-scattering and Mel Frequency Cesptral Coefficients (MFCCs).
+For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients.
+A logarithm compression (see Section\ref{XX}) is applied to the temporal-scattering coefficient.
+Different set of parameters were tested to compute the MFFCs, we here present only the results for the best setting, that is, 40 MFCCs (including the 0 energy coefficient) computed for windows of 50ms and hops of 25ms, with a limited frequency range of $27.5-1000$Hz.
+This setting gives us 600 feature vectors per scenes.
+To reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices.
+Each $n$ MFCC slices is then averaged over time.
+To ensure that each scene is described by circa the same number of vectors for both the temporal-scattering and the MFCCs, we set $t$ to 250ms, giving us 120 MFCCs vectors per scenes.
+
 \section{Acoustic Scene Similarity}
 
-\ml{GL 1 colonne}
+\subsection{Role of logarithmic compression}
 
-\subsection{Protocol}
+\subsection{MFCC \vs Time scattering}
 
-\subsection{Results}
+\subsection{Object-based \vs early integration}
 
@@
 
-\subsection{Time scattering \vs time-frequency scattering}
+\subsection{MFCC \vs Time scattering}
 
-\subsection{Early \vs late integration}
-
-\subsection{Role of data augmentation}
-
-\subsection{Role of frequency range}
+\subsection{Object-based \vs early/late integration}
 

### 2016-07-09 generateFig/functions/printFigures.m
add scripts to generate figures

@@ -1,0 +1,36 @@
+function [] = printFigures(nbFig,name,opt)
+
+for jj=1:length(nbFig)
+    set(figure(nbFig(jj)),'PaperUnits','centimeters')
+    set(figure(nbFig(jj)),'PaperPositionMode','manual')
+    set(figure(nbFig(jj)),'papersize',[opt.width,opt.height])
+    set(figure(nbFig(jj)),'paperposition',[0,0,opt.width,opt.height])
+    
+    set(findall(figure(nbFig(jj)),'-property','FontSize'),'FontSize',opt.fontsize)
+    set(findall(figure(nbFig(jj)),'-property','FontName'),'FontName','Arial')
+    
+    %%
+    
+    if isfield(opt,'markersize')
+        set(findall(figure(nbFig(jj)),'-property','markersize'),'markersize',opt.markersize)
+    end
+    
+    if isfield(opt,'linewidth')
+        set(findall(figure(nbFig(jj)),'-property','linewidth'),'linewidth',opt.linewidth)
+    end
+    
+    %% title
+    
+    if isfield(opt,'axesTitle')
+        allAxesInFigure = findall(figure(nbFig(jj)),'type','axes');
+        for ll=1:length(allAxesInFigure)
+            allAxesInFigure(ll).Title.String=opt.axesTitle{jj}{length(allAxesInFigure)-ll+1}; % reverse order of axes
+        end
+    end
+    
+    print(['-f' num2str(nbFig(jj))],[name num2str(nbFig(jj))],'-depsc')
+end
+
+
+end
+

### 2016-07-09 generateFig/unsupervised.m

@@ -1,0 +1,81 @@
+clear all
+addpath(genpath('functions'))
+
+
+
+%% settings
+
+dataset='train'; % test;
+dataPath={['~/projets/ass_features/report/figures/unsupervised_earlyInt_' dataset]};
+freqRange = {'27,5_1000','27,5_2000','27,5_4000','27,5_8000','27,5_16000','27,5_22050'};
+
+linewidth=2;
+markersize=10;
+
+%% load data
+data=cell(1,length(dataPath));
+leg=cell(1,length(dataPath));
+
+for jj=1:length(dataPath)
+    d=load(dataPath{jj});
+    [data{jj},leg{jj}]=parseTable(d.data(1).tables{end});
+end
+
+pak_early=str2double(data{1}(9:15,:));
+
+scat=strcmp('scatT',data{1}(1,:));
+mfcc=strcmp('mfcc',data{1}(1,:));
+
+%% best p@9
+
+ind_best_mfcc=find( mfcc & pak_early(end,:)==max(pak_early(end,mfcc)));
+ind_best_scat=find( scat & pak_early(end,:)==max(pak_early(end,scat)));
+
+best_setting_mfcc=data{1}(:,ind_best_mfcc);
+best_setting_scat=data{1}(:,ind_best_scat(1));
+
+freq=strcmp('27,5_1000',data{1}(2,:));
+mfcc40=strcmp('40',data{1}(3,:));
+c0=strcmp('1',data{1}(4,:));
+scat_log=strcmp('1',data{1}(5,:));
+stand=strcmp('stand',data{1}(6,:));
+gaussian=strcmp('gaussian_full',data{1}(7,:));
+nn_5=strcmp('0.05',data{1}(8,:));
+nn_15=strcmp('0.15',data{1}(8,:));
+
+
+pak_early_mfcc=pak_early(:,mfcc & gaussian & ~stand & mfcc40 & c0 & nn_15 & freq);
+pak_early_scat=pak_early(:,scat & gaussian & ~stand  & nn_15 & ~scat_log);
+pak_early_scat_log=pak_early(:,scat & gaussian & ~stand  & nn_15 & scat_log);
+
+%%
+figure(1)
+clf
+plot(pak_early_mfcc,'o-','markersize',markersize,'linewidth',linewidth)
+hold on
+plot(pak_early_scat_log,'+-','markersize',markersize,'linewidth',linewidth)
+hold off
[…]

### 2016-07-09 paper/structureScene.tex
add fig drafts for the ASS task using the test set

@@
 To compute the  weighted similarity between two scenes by taking into account both theirs clusters weights and theirs centroids similarities, we use a variant of the Earth Moving Distance ($EMD$) adapted for non-normalized histograms known as $\widehat{EMD}$ \cite{pele2008linear} with the implementation proposed in  \cite{pele2009fast}.
-Considering two histograms $h_i$ and $h_j$, represented the centroids weights of each scene, the $\widehat{EMD}$ is computed as follow:
+Considering $h_i$ and $h_j$, the histograms of the centroids weights of the scenes $s_i$ and $s_j$ respectively, the $\widehat{EMD}$ is computed as follow:
 
@@
 
-where $f_{ij}$ is the flow, and $D_{ij}$ is the ground distance. In our case the ground distance is $1-K^c_{ij}$
+where $f_{ij}$ is the flow, and $D_{ij}$ is the ground distance. In our case, the ground distance is $D_{ij}=1-K^c_{ij}$.
 
-To get the final similarity measure between two scenes $s_i$ and $s_j$, we use an extended Gaussian kernels $K^s$ (Chapelle et al., 1999; Jing et al., 2003).
+To get the final similarity measure between two scenes $s_i$ and $s_j$, we use an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support}.
 
@@
 
-with $A$ a scaling parameter, set to the mean value of the $\widehat{EMD}$. \\
+with $A$ a scaling parameter, set to the mean value of the $\widehat{EMD}$.
+The resulting kernel is called an EMD kernel, and it has to be noted that there is no proof that such kernel is positive definite.
+\\
 
-The resulting kernel is an EMD kernel. It has to be noted that there is no proof that such kernel is positive definite.
+Other articles using EMD in vision \cite{zhang2007local}
 

### 2016-07-09 generateFig/unsupervised.m

@@ -5,7 +5,9 @@ addpath(genpath('functions'))
 
 %% settings
 
-dataset='train'; % test;
+% dataset='train'; % test;
+dataset='test'; 
+
 dataPath={['~/projets/ass_features/report/figures/unsupervised_earlyInt_' dataset]};
 freqRange = {'27,5_1000','27,5_2000','27,5_4000','27,5_8000','27,5_16000','27,5_22050'};
 

### 2016-07-09 paper/structureScene.tex
update figs

@@
 
-\section{Acoustic Scene Similarity}
+\section{Results: Acoustic Scene Similarity}
 
-\subsection{Role of logarithmic compression}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{gfx/unsupervised_test1.eps}
+\caption{...}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{gfx/unsupervised_test2.eps}
+\caption{...}
+\end{center}
+\end{figure}
 
@@
 
-\section{Acoustic Scene Classification}
+\subsection{Role of logarithmic compression}
+
+\section{Results: Acoustic Scene Classification}
 Support vector machines are large-margin binary discriminators with tunable regularization of misclassified examples.
 
+\begin{table}
+\begin{center}
+\begin{tabular}{c|cc}
+             & time-scattering & mfcc \\
+             \hline
+early int.   & XX (XX)    & XX (XX)   \\
+late int.    & XX (XX)    & XX (XX)   \\
+ob-closest.  & XX (XX)    & XX (XX)   \\
+ob-averaged. & XX (XX)    & XX (XX)   \\
+ob-weighted. & XX (XX)    & XX (XX)   \\
+\end{tabular}
+\caption{...}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{c|cc}
+             & w/ log comp. & w/o log comp. \\
+             \hline
+early int.   & XX (XX)      & XX (XX)   \\
+late int.    & XX (XX)      & XX (XX)   \\
+ob-closest.  & XX (XX)      & XX (XX)   \\
+ob-averaged. & XX (XX)      & XX (XX)   \\
+ob-weighted. & XX (XX)      & XX (XX)   \\
+\end{tabular}
+\caption{...}
+\end{center}
+\end{table}
 \subsection{Role of logarithmic compression}
@@
 \subsection{Object-based \vs early/late integration}
+
+\section{Results: DCASE 2016}
 

### 2016-07-09 generateFig/unsupervised.m

@@ -58,9 +58,11 @@ hold on
 plot(pak_early_scat_log,'+-','markersize',markersize,'linewidth',linewidth)
 hold off
 legend({'ei, mfcc','ei, time-scattering'},'interpreter','none')
+legend('boxoff')
 set(gca,'xtick',1:7,'xticklabel',3:9)
 xlabel('k')
 ylabel('p@k')
+box off
 disp('')
 
 figure(2)
@@ -70,14 +72,16 @@ hold on
 plot(pak_early_scat_log,'+-','markersize',markersize,'linewidth',linewidth)
 hold off
 legend({'ei, w/o log','ei, w/ log'},'interpreter','none')
+legend('boxoff')
 set(gca,'xtick',1:7,'xticklabel',3:9)
 xlabel('k')
 ylabel('p@k')
+box off
 disp('')
 
 %% print
 figOpt.fontsize=16;
-figOpt.height=35;
+figOpt.height=15;
 figOpt.width=20;
 
 printFigures([1 2],['~/papers/paperStructureScene16/paper/gfx/unsupervised_' dataset],figOpt)

### 2016-07-09 paper/structureScene.tex
done with abs intro and motivations

@@
 \begin{abstract}
-This paper introduces a generic hierarchical modeling of acoustic scenes using time/frequency scattering features for time spans below the second and object-based segmentation for larger time spans that effectiely model the scene with almost no requirement for optimization.
-Application of the proposed approach to the acoustic scene similarity and acoustic scene classification tasks demonstrates its potential.
 
-More precisely, the proposed approach achieves 
+This paper introduces a generic hierarchical modeling of acoustic scenes using features computed using the time scattering transform  for time spans below the second and object-based segmentation for larger time spans that effectively model the scene.
+
+Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets and state-of-the-art classifier demonstrates its potential.
+More precisely, features derived from the time scattering transform consistently outperform Mel-frequency cepstral coefficients (MFCCs) for both tasks.
+
+Representing the scene at larger time scale as a set of examplars of coherent segments is an effective approach for similarity retrieval as it allows the retrieval system to focus on discriminative features.
+This benefit is reduced in supervised settings, probably because the much of the discriminability potential of the approach is already learned by the classifier.
+
 \end{abstract}
@@
 \IEEEPARstart{O}{ver} the past decades, the amount of audio data recorded from our sonic environment has considerably grown.
-Recent research areas such as eco-acoustics \cite{ECOACOUSTICS2014, krause} start to consider massive deployment of acoustic sensors around the world in order to measure potential animal biodiversity modification over large temporal scales due to human activity or climate change \cite{NessSST13, stowell13a, stowell13b}.
+Recent research areas such as eco-acoustics \cite{ECOACOUSTICS2014, krause} start to consider massive deployment of acoustic sensors around the world in order to measure potential animal biodiversity modification over large temporal scales due to human activity or climate change \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
 
-Also, some recent works demonstrate the interest of such approaches for the characterization of human acoustic pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} and the prediction of annoyance due to the traffic \cite{gloaguen}.
+Also, some recent works demonstrate the interest of such approaches for the characterization of human acoustic pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} and the prediction of annoyance due to traffic \cite{gloaguen}.
 We believe that those case study are of interest for the signal processing community as they have strong societal impacts and raises interesting research avenues.
@@
 Despite the differences, the following question is valid both for the above cited future application and the acoustic scene classification (ASC) task.
-How can we efficiently and effectively process large amount of audio data using representations that are both compact, generic in their design and with a large application range ?
+How can we efficiently and effectively process large amount of audio data using representations that are both compact, generic in their design and with a large potential application range ?
 
-The research described in this paper tackles that question by demonstrating the usefulness of two main contributions: 1) the use of the scattering transform allows us to extract discriminative features with almost no parameter optimization and 2) object-based modeling of the acoustic scene based on individual unsupervised clustering of each acoustic scene  allows us to effectively focus the description on areas of interest.
-After studying the state of the art in the field of acoustic scene modeling in Section \ref{sec:soa}, those contributions are extensively described, respectively in Section \ref{sec:scattering} and \ref{sec:object} and evaluated both in an unsupervised setting, \textit{i.e.} acoustic scene similarity (ASS) and in a supervised setting, \textit{i.e.}  acoustic scene classification (ASC), respectively in Section \ref{sec:similarity} and Section \ref{sec:classification}.
+The research described in this paper attempt to provide answers to that question by demonstrating the usefulness of two main contributions: 1) the use of the time scattering transform allows us to extract discriminative features with very few parameter optimization and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene  allows us to effectively focus the description of the scene on areas of interest.
+
+The proposed approach is motivated in Section \ref{sec:motivations}  and a brief review the state of the art in the field of acoustic scene modeling is performed in Section \ref{sec:soa}.
+Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively.
+Experiments carried out in an unsupervised setting, \textit{i.e.} acoustic scene similarity (ASS) and in a supervised setting, \textit{i.e.}  acoustic scene classification (ASC) are then reported in Section \ref{sec:experiments}.
+
+\ml{blurb about results}
+
+\section{Motivations} \label{sec:motivations}
+
+As stated in the introduction, the field of eco-acoustics is relatively young, but significant works are worth mentioning.
+One avenue of research issued from bio acoustic studies, consider the audio stream as a set of events produced animals, and is processed to retrieve the precise species that produced the sound and potentially the type of vocalization in order to predict some behavior of the animal under study \cite{au2008principles, stowell2014large}.
+Another, more recent, avenue of research, tackles problems associated with long duration recordings of the environment, in particular, building recognizers for species of interest, extracting acoustic indices to aid navigation and visualisation.
+It typically process the audio stream in an holistic manner in order to infer large time-scale properties of the bio-acoustic scene for example to estimate which part of the stream need to be analyzed more in details, \textit{i.e.} by a human \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+
+The same dichotomy applies for the field of audio scene analysis, where the interest can be whether 1) to isolate and identify precise events, \textit{i.e.} acoustic event detection (AED) or 2) to predict global attributes assigned to large time spans of audio, \textit{i.e.} ASC.
+For the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag} where the audio is modelled as high level statistics computed from a set of local features.
+Even though this approach as the advantage of high genericity and follows some evidence found by psycho perception research \cite{macdermott}, for the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere point averaging of the features  \cite{lagrange:hal-01082501}.
+
+Intuitively, this result indicates that not all the features distribution is not of equal interest while modeling acoustic scenes, which is coherent with the typical morphology of acoustic scene which are relatively scarce in terms of events, even if those events can trigger fast perceptual recognition for humans.
+We thus could consider the acoustic scene as a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}.
+
+If the precise list of events that occur in a given time is unachievable to get even for trained humans, research shows that using only a few events, so called markers \cite{} is enough to reliably predict high level attributes.
+For example, the presence of a few bird calls in a smooth traffic hubhub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
 
 \section{Previous work} \label{sec:soa}
-
-As stated in the introduction, the field of eco-acoustics is relatively young, but significant works are worth mentioning.
-One avenue of research issued from bio acoustic studies, consider the audio stream as a set of events produced animals, and is processed to retrieve the precise species that produced the sound and potentially the type of vocalization in order to predict some behavior of the animal under study \cite{dan}.
-Another, more recent, avenue of research, process the audio stream in an holistic manner in order to infer large time-scale properties of the bio-acoustic scene for example to estimate which part of the stream need to be analyzed in details, \textit{i.e.} by a human \cite{australians}.
-
-The same dichotomy applies for the field of audio scene analysis, where the interest can be whether 1) to isolate and identify precise events, \textit{i.e.} acoustic event detection (AED) or 2) to predict global attributes assigned to large time spans of audio, \textit{i.e.} ASC.
-For the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag} where the audio is modelled as high level statistics comptued from a set of local features.
-Even though this approach as the advantage of high genericity and follows some evidence found by psycho perception research \cite{macdermott}, for the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be equivalent to  \cite{lagrange:hal-01082501}.
-
-citer papier event based de l'intro de jj
-
- there is evidence that modelling the scene as a set of events allows us to predict high level semantic information such as pleasantness
-
-Eventhough the precise list of events that occur in a given time is unachievable to get even for trained human, research shows that using only a few events, so called markers \cite{} is enough to reliably predict high level attributes.
-For example, the presence of a few bird calls in a smooth traffic hubhub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
 
@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test1.eps}
-\caption{...}
+\caption{... \ml{fusionner les deux figures et commencer x a 1 par pas de 1}}
 \end{center}

### 2016-07-09 paper/structureScene.tex
checking VL questions

@@
 For the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag} where the audio is modelled as high level statistics computed from a set of local features.
-Even though this approach as the advantage of high genericity and follows some evidence found by psycho perception research \cite{macdermott}, for the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere point averaging of the features  \cite{lagrange:hal-01082501}.
+Even though this approach as the advantage of high genericity and follows some evidence found by psycho perception research \cite{mcdermott2013summary}, for the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere point averaging of the features  \cite{lagrange:hal-01082501}.
 
@@
 Right: Gammatone wavelet $\boldsymbol{\psi_{\gamma_2}}(t)$ with a quality factor of $Q_2=1$.
-\vl{we should discuss color schemes and axis labels}}
+\vl{we should discuss color schemes and axis labels} \ml{yes we should but for the second submission, go ahead}}
 \end{center}
@@
 
-This strategy is already widespread in the literature.
-For instance, we should note that \cite{roma2013}, who secured the first place in the 2013 edition of the DCASE challenge, rely on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max}  = 900\,\mathrm{Hz}$ instead of the more conventional value of $16\,\mathrm{kHz}$.
+Reducing the frequency range is an effective way to enhance invariance, widely used in the literature.
+For instance, \cite{roma2013} achieved the best performance in the 2013 edition of the DCASE challenge, rely on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max}  = 900\,\mathrm{Hz}$ instead of the more conventional value of $16\,\mathrm{kHz}$.
 \vl{Qu'en penses-tu ML ?}
+\ml{j'ai un peu reformuler}
 
@@
 \vl{Qu'en penses-tu ML ?}
+\ml{tres bien, mettre les deux distributions de couleurs differentes}
 

### 2016-07-09 paper/structureScene.tex
done with datasets

@@
 
-\section{Previous work} \label{sec:soa}
+\section{Background} \label{sec:soa}
+
+Apart from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
+
+To this end, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this task.
 
@@
 
-\gl{Other paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, barchiesi2015acoustic, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory, bisot2016acoustic}
+%\gl{Other paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory}
 
-\subsection{Describing a scene (Feature extraction)}
+%\subsection{Datasets}
 
-\subsection{Classifying a scene}
-
-\subsection{Datasets}
-
-defreville \cite{aucouturier2007bag}
-
-\cite{lagrange:hal-01082501}
-
-DCASE 2013 challenge reference papers. \cite{giannoulis2013database, 7100934} \\
-
-For each scene type, three different recordists (DG, DS,
-EB) visited a wide variety of locations in Greater London over
-a period of months (Summer and Autumn 2012), and in each
-scene recorded a few minutes of audio. We ensured that no
+For most of the experiments reported in this paper, the DCASE2013 datatset is considered \cite{giannoulis2013database, 7100934}.
+To build this dataset, three different recordists visited a wide variety of locations in Greater London over
+a period of months and in each
+scene recorded a few minutes of audio for each scene type. No
 systematic variations in the recordings covaried with scene
@@
 recorded each scene type.
+The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a 5 fold cross validation scheme.
+The folds used in this paper are the same as the ones used during the challenge.
 
-Rouen \cite{rakotomamonjy2015histogram}
+As of today, there exist larger publicly available dataset, the
+Rouen dataset \cite{rakotomamonjy2015histogram} and the dCase2016 \cite{Mesaros2016_EUSIPCO} (only the public part).
+The latter is also considered in this paper as a validation of the proposed approach.
 
-\url{https://sites.google.com/site/alainrakotomamonjy/home/audio-scene}
-
-dCase2016 \cite{Mesaros2016_EUSIPCO}
-
-advantages of dcase2013:
-
-controled intra class diversity
-
-despite its size, it is still challenging as state-of-the-art systems achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
+As studied in \cite{lagrange:hal-01082501}, the DCASE2013 dataset has interesting intra class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices.
+Also, it is still challenging as state-of-the-art systems based on feqtures design and support vector machines (SVMs) achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
 For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively.
+On the former, recent advances using hierarchical classifiers architectures reports 84 to 87 \% of accuracy depending on the prior knowledge considered for building the classifier architecture \cite{phan2016label}.
 On the latter, the state-of-the-art is currently 95 \% \cite{bisot2016acoustic}.
@@
 \section{Scattering representations \label{sec:scattering}}
+
 Scattering transforms are translation-invariant representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.

### 2016-07-09 paper/structureScene.tex
update feature section, add supervised fig, merge tables, update tables

@@
 As studied in \cite{lagrange:hal-01082501}, the DCASE2013 dataset has interesting intra class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices.
-Also, it is still challenging as state-of-the-art systems based on feqtures design and support vector machines (SVMs) achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
+Also, it is still challenging as state-of-the-art systems based on features design and support vector machines (SVMs) achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
 For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively.
@@
 \item \emph{ASS}: It is an unsupervised scenario.
-Evaluation is performed on the private dataset of the DCASE 2013 challenge, using the precision at rank $k$ ($k=[3,\ldots,9]$) metrics.
-Similarity between scenes features is then computed using a Gaussian kernel with local scaling.
+Evaluation is performed on the private dataset of the DCASE 2013 challenge, using the precision at rank $k$ ($k=[1,\ldots,9]$) metrics.
+Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the points to classify.
 
@@
 
+\subsection{Comparison with early and late integration}
+
+ASC, ASS:   Similarity between scenes features is then computed using a Gaussian kernel with local scaling.
+
 The object-based approach is compared to standard approaches using early integration strategy for the ASS scenario, and both early and late integration strategies for the ASC scenario (see Section \ref{XX}).
@@
 
-Two features are used to describe the scenes: temporal-scattering and Mel Frequency Cesptral Coefficients (MFCCs).
-For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients.
-A logarithm compression (see Section\ref{XX}) is applied to the temporal-scattering coefficient.
-Different set of parameters were tested to compute the MFFCs, we here present only the results for the best setting, that is, 40 MFCCs (including the 0 energy coefficient) computed for windows of 50ms and hops of 25ms, with a limited frequency range of $27.5-1000$Hz.
+Two types features are used to describe the scenes:
+
+\begin{itemize}
+\item \emph{temporal-scattering}: Each scene is described by 128 vectors of scattering coefficients. 
+\item \emph{Mel Frequency Cesptral Coefficients} (MFCCs): Different set of parameters were tested to compute the MFFCs, we here present only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
+We use 40 MFCCs (including the 0 energy coefficient) computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz.
 This setting gives us 600 feature vectors per scenes.
@@
 Each $n$ MFCC slices is then averaged over time.
-To ensure that each scene is described by circa the same number of vectors for both the temporal-scattering and the MFCCs, we set $t$ to 250ms, giving us 120 MFCCs vectors per scenes.
+As the object-based approach (see \ref{XX}) involves a clustering step, we want to ensure that each scene is described by circa the same number of vectors prior to the clustering, for both the temporal-scattering and the MFCCs.
+To do so, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.
+\end{itemize}
+
+For the ASC scenario, the features are standardized prior to the SVM classification step.
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{gfx/supervised_linear_test1.eps}
+\caption{...}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{gfx/supervised_rbf_test2.eps}
+\caption{...}
+\end{center}
+\end{figure}
+
 \begin{table}
 \begin{center}
-\begin{tabular}{c|cc}
-             & time-scattering & mfcc \\
+\begin{tabular}{cccc}
+                    & time-scattering & time-scattering & mfcc \\
+                    & w/ log comp.    & w/o log comp.   &      \\
              \hline
-early int.   & XX (XX)    & XX (XX)   \\
-late int.    & XX (XX)    & XX (XX)   \\
-ob-closest.  & XX (XX)    & XX (XX)   \\
-ob-averaged. & XX (XX)    & XX (XX)   \\
-ob-weighted. & XX (XX)    & XX (XX)   \\
+early int. (linear)  & 66 (11)        & 62 (6)  & 46 (8)    \\
+early int. (RBF)     & 60 (7)         & 57 (8)  & 60 (8)    \\
+late int.  (linear)  & XX (XX)        & XX (XX) & XX (XX)   \\
+late int.  (RBF)     & XX (XX)        & XX (XX) & XX (XX)   \\
+ob-closest.          & XX (XX)        & XX (XX) & XX (XX)   \\
+ob-averaged.         & XX (XX)        & XX (XX) & XX (XX)   \\
+ob-weighted.         & XX (XX)        & XX (XX) & XX (XX)   \\
 \end{tabular}
@@
 
-\begin{table}
-\begin{center}
-\begin{tabular}{c|cc}
-             & w/ log comp. & w/o log comp. \\
-             \hline
-early int.   & XX (XX)      & XX (XX)   \\
-late int.    & XX (XX)      & XX (XX)   \\
-ob-closest.  & XX (XX)      & XX (XX)   \\
-ob-averaged. & XX (XX)      & XX (XX)   \\
-ob-weighted. & XX (XX)      & XX (XX)   \\
-\end{tabular}
-\caption{...}
-\end{center}
-\end{table}
 \subsection{Role of logarithmic compression}

### 2016-07-09 generateFig/unsupervised.m

@@ -1,15 +1,12 @@
 clear all
 addpath(genpath('functions'))
 
-
-
 %% settings
 
 % dataset='train'; % test;
 dataset='test'; 
 
 dataPath={['~/projets/ass_features/report/figures/unsupervised_earlyInt_' dataset]};
-freqRange = {'27,5_1000','27,5_2000','27,5_4000','27,5_8000','27,5_16000','27,5_22050'};
 
 linewidth=2;
 markersize=10;
@@ -42,13 +39,12 @@ c0=strcmp('1',data{1}(4,:));
 scat_log=strcmp('1',data{1}(5,:));
 stand=strcmp('stand',data{1}(6,:));
 gaussian=strcmp('gaussian_full',data{1}(7,:));
-nn_5=strcmp('0.05',data{1}(8,:));
-nn_15=strcmp('0.15',data{1}(8,:));
+nn_10=strcmp('0.10',data{1}(8,:));
 
 
-pak_early_mfcc=pak_early(:,mfcc & gaussian & ~stand & mfcc40 & c0 & nn_15 & freq);
-pak_early_scat=pak_early(:,scat & gaussian & ~stand  & nn_15 & ~scat_log);
-pak_early_scat_log=pak_early(:,scat & gaussian & ~stand  & nn_15 & scat_log);
+pak_early_mfcc=pak_early(:,mfcc & gaussian & ~stand & mfcc40 & c0 & nn_10 & freq);
+pak_early_scat=pak_early(:,scat & gaussian & ~stand  & nn_10 & ~scat_log);
+pak_early_scat_log=pak_early(:,scat & gaussian & ~stand  & nn_10 & scat_log);
 
 %%
 figure(1)

### 2016-07-09 paper/structureScene.tex
finish Fig 1

@@
 \usetikzlibrary{arrows,positioning}
+\usetikzlibrary{arrows.meta}
 \usepackage{url}
@@
 % correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\hyphenation{wave-let}
 
@@
 \begin{center}
-\setlength{\unitlength}{1cm}
-\begin{picture}(5,2)
- \put(-1.5,0.0){\includegraphics[height=2cm,width=3.5cm]{gammatone_Q8.png}}
- \put(2.5,0.0){\includegraphics[height=2cm,width=3.5cm]{gammatone_Q1.png}}
-\end{picture}
+\begin{tikzpicture}
+\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+\node (omlabel1) at (3.5, 1) {$t$};
+\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
+\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
+\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
+\node at (2, -0.25) {(a)};
+\node (omlabel2) at (8, 1) {$t$};
+\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
+\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
+\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
+\node at (6.5, -0.25) {(b)};
+\end{tikzpicture}
 \caption{
 \label{fig:gammatones}
-Left: Gammatone wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$ with a quality factor of $Q_1=8$.
-Right: Gammatone wavelet $\boldsymbol{\psi_{\gamma_2}}(t)$ with a quality factor of $Q_2=1$.
-\vl{we should discuss color schemes and axis labels} \ml{yes we should but for the second submission, go ahead}}
+Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
+Blue and red oscillations represent the real and imaginary parts.
+The orange envelope represents the complex modulus.}
 \end{center}
@@
 For instance, \cite{roma2013} achieved the best performance in the 2013 edition of the DCASE challenge, rely on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max}  = 900\,\mathrm{Hz}$ instead of the more conventional value of $16\,\mathrm{kHz}$.
-\vl{Qu'en penses-tu ML ?}
-\ml{j'ai un peu reformuler}
 

### 2016-07-09 figs/wavelets/gammatone.m
better line widths in Fig 1

@@ -1,5 +1,5 @@
 nfo = 8;
-line_width = 1.5;
+line_width = 2;
 
 opts{1}.time.nFilters_per_octave = nfo;
 opts{1}.time.size = 2^19;
@@ -20,16 +20,27 @@ psi2_fts = display_bank(archs{2}.banks{1});
 psi1_ft = psi1_fts(:, 40);
 psi1_ift = ifftshift(ifft(psi1_ft));
 psi1_ift = psi1_ift((end/2) + (-300:1000));
-psi1 = [real(psi1_ift),imag(psi1_ift),abs(psi1_ift)];
+psi1 = [real(psi1_ift),imag(psi1_ift)];
+abspsi1 = abs(psi1_ift);
 
 psi2_ft = psi2_fts(:, 11);
 psi2_ift = ifftshift(ifft(psi2_ft));
 psi2_ift = psi2_ift((end/2) + (-3000:10000));
-psi2 = [real(psi2_ift),imag(psi2_ift),abs(psi2_ift)];
+psi2 = [real(psi2_ift),imag(psi2_ift)];
+abspsi2 = abs(psi2_ift);
 
+figure(1);
 plot(psi1, 'LineWidth', line_width)
 axis off
+hold on
+plot(abspsi1, 'LineWidth', 2 * line_width);
+hold off;
 export_fig -transparent gammatone_Q8.png
+
+figure(2);
 plot(psi2, 'LineWidth', line_width);
 axis off
+hold on
+plot(abspsi2, 'LineWidth', 2 * line_width);
+hold off;
 export_fig -transparent gammatone_Q1.png

### 2016-07-10 paper/structureScene.tex
better line widths in Fig 1
add script figure generation supervised, update figures unsupervised (final results)

--- 2016-07-09 paper/structureScene.tex
@@
 \usepackage{pgf}
-\usepackage{tikz}
-\usetikzlibrary{arrows,positioning}
+%\usepackage{tikz}
+%\usetikzlibrary{arrows,positioning}
+%\usetikzlibrary{arrows.meta}
 \usepackage{url}
@@
 
-\begin{figure}
-\begin{center}
-\setlength{\unitlength}{1cm}
-\begin{picture}(5,2)
- \put(-1.5,0.0){\includegraphics[height=2cm,width=3.5cm]{gammatone_Q8.png}}
- \put(2.5,0.0){\includegraphics[height=2cm,width=3.5cm]{gammatone_Q1.png}}
-\end{picture}
-\caption{
-\label{fig:gammatones}
-Left: Gammatone wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$ with a quality factor of $Q_1=8$.
-Right: Gammatone wavelet $\boldsymbol{\psi_{\gamma_2}}(t)$ with a quality factor of $Q_2=1$.
-\vl{we should discuss color schemes and axis labels} \ml{yes we should but for the second submission, go ahead}}
-\end{center}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\begin{tikzpicture}
+%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+%\node (omlabel1) at (3.5, 1) {$t$};
+%\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
+%\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
+%\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
+%\node at (2, -0.25) {(a)};
+%\node (omlabel2) at (8, 1) {$t$};
+%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
+%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
+%\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
+%\node at (6.5, -0.25) {(b)};
+%\end{tikzpicture}
+%\caption{
+%\label{fig:gammatones}
+%Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
+Blue and red oscillations represent the real and imaginary parts.
+The orange envelope represents the complex modulus.}
+%\end{center}
+%\end{figure}
 
@@
 This operation may also help making the dataset linearly separable in feature space.
-\vl{Check this claim in DCASE 2013 time scattering features.}
 To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we apply the affine transformation

### 2016-07-10 generateFig/supervised_test.m

@@ -1,0 +1,94 @@
+clear all
+addpath(genpath('functions'))
+
+%% settings
+
+dataset='test'; 
+dataPath={['~/projets/ass_features/report/figures/supervised_earlyInt_' dataset]};
+
+linewidth=2;
+markersize=10;
+markerLeg={'diamond','o','*','x','^'};
+%% load data
+
+data=cell(1,length(dataPath));
+leg=cell(1,length(dataPath));
+
+for jj=1:length(dataPath)
+    d=load(dataPath{jj});
+    [data{jj},leg{jj}]=parseTable(d.data(1).tables{end});
+end
+
+acc=str2double(data{1}(10:14,:));
+m_acc=mean(acc,1);
+std_acc=std(acc,[],1);
+
+scat=strcmp('scatT',data{1}(1,:));
+mfcc=strcmp('mfcc',data{1}(1,:));
+mfcc_rank=strcmp('40',data{1}(3,:));
+mfcc_c0=strcmp('1',data{1}(4,:));
+fr=strcmp('27,5_1000',data{1}(2,:));
+stand=strcmp('stand',data{1}(6,:));
+gaussian=strcmp('gaussian_full',data{1}(7,:));
+inner=strcmp('inner_full',data{1}(7,:));
+nn_10_mfcc=strcmp('0.10',data{1}(8,:));
+nn_40_scat=strcmp('0.40',data{1}(8,:));
+scat_log=strcmp('1',data{1}(5,:));
+
+%% best acc
+leg2={'ei-mfcc-linear','ei-scat-linear','ei-scat-logComp-linear',...
+    'ei-mfcc-rbf','ei-scat-rbf','ei-scat-logComp-rbf'};
+
+sel{1}= mfcc_rank & mfcc_c0  & stand & fr & mfcc & inner;
+sel{2}= ~scat_log & stand & scat & inner;
+sel{3}= scat_log & stand & scat & inner;
+sel{4}= mfcc_rank & mfcc_c0  & stand & fr & mfcc & gaussian & nn_10_mfcc;
+sel{5}= ~scat_log & stand & scat & gaussian & nn_40_scat;
+sel{6}= scat_log & stand & scat & gaussian & nn_40_scat;
+
+acc_f=[];
+
+for jj=1:length(leg2)
+    disp([leg2{jj} ' : acc=' num2str(m_acc(sel{jj})) ', std=' num2str(std_acc(sel{jj})) ])
+    acc_f=[acc_f acc(:,sel{jj})];
+end
[…]

### 2016-07-10 generateFig/supervised_train.m

@@ -1,0 +1,44 @@
+clear all
+addpath(genpath('functions'))
+
+%% settings
+
+dataset='train'; 
+dataPath={['~/projets/ass_features/report/figures/supervised_earlyInt_' dataset]};
+
+linewidth=2;
+markersize=10;
+
+%% load data
+
+data=cell(1,length(dataPath));
+leg=cell(1,length(dataPath));
+
+for jj=1:length(dataPath)
+    d=load(dataPath{jj});
+    [data{jj},leg{jj}]=parseTable(d.data(1).tables{end});
+end
+
+acc=str2double(data{1}(10:14,:));
+m_acc=mean(acc,1);
+
+scat=strcmp('scatT',data{1}(1,:));
+mfcc=strcmp('mfcc',data{1}(1,:));
+fr=strcmp('27,5_1000',data{1}(2,:));
+stand=strcmp('stand',data{1}(6,:));
+gaussian=strcmp('gaussian_full',data{1}(7,:));
+inner=strcmp('inner_full',data{1}(7,:));
+scat_log=strcmp('1',data{1}(5,:));
+
+%% best acc
+
+ind_best_mfcc_inner=find(stand & inner & mfcc & fr  & m_acc==max(m_acc(stand & inner & mfcc & fr)));
+ind_best_gaussian=find(stand & gaussian & mfcc & fr  & m_acc==max(m_acc(stand & gaussian & mfcc & fr)));
+ind_best_scat_inner=find(stand & inner & scat & m_acc==max(m_acc(stand & inner & scat)));
+ind_best_scat_gaussian=find(stand & gaussian & scat & m_acc==max(m_acc(stand & gaussian & scat)));
+
+best_setting_mfcc_inner=data{1}(:,ind_best_mfcc_inner);
+best_setting_mfcc_gaussian=data{1}(:,ind_best_gaussian);
+best_setting_scat_inner=data{1}(:,ind_best_scat_inner);
+best_setting_scat_gaussian=data{1}(:,ind_best_scat_gaussian);
+disp('')

### 2016-07-10 paper/structureScene.tex
add results for the ASC task

--- 2016-07-09 paper/structureScene.tex
@@
 early int. (RBF)     & 60 (7)         & 57 (8)  & 60 (8)    \\
-late int.  (linear)  & XX (XX)        & XX (XX) & XX (XX)   \\
-late int.  (RBF)     & XX (XX)        & XX (XX) & XX (XX)   \\
-ob-closest.          & XX (XX)        & XX (XX) & XX (XX)   \\
-ob-averaged.         & XX (XX)        & XX (XX) & XX (XX)   \\
-ob-weighted.         & XX (XX)        & XX (XX) & XX (XX)   \\
+late int.  (linear)  & 75 (5)         & 70 (8)  & 63 (8)   \\
+late int.  (RBF)     & 78 (6)         & 73 (8)  & 72 (10)   \\
+ob-closest.          & 66 (7)         & 58 (6)  & 56 (9)   \\
+ob-averaged.         & 61 (7)         & 54 (8)  & 61 (9)   \\
+ob-weighted.         & 54 (11)        & 43 (10) & 51 (15)   \\
 \end{tabular}

### 2016-07-10 paper/structureScene.tex
done sith systems

--- 2016-07-10 paper/structureScene.tex
@@
 
-Apart from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
+Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
 
-To this end, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this task.
+Thus, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this latter task.
+The table \ref{tab:dcase} summarizes the different algorithmic choices taken by the teams.
+Unless otherwise stated, scene predictions are achieved by majority voting of predictions \ml{gregoire ?}.
 
-\gl{DCASE challenge:} \cite{chum2013ieee}: CHR system, using HMM and frame based classification.
-\cite{elizalde2013vector} ELF system, I vector approach using probabilistic LDA as classifier.
-\cite{geiger2013large} GSR system, basically mel/MFCC using SVM as classifier.
-\cite{krijnders2013tone} KH system, tone fit features computed from cochleogram (but similar results are obtained with MFCCs) using SVM as classifier.
-\cite{li2013auditory} LTT system, Tree bagger classifier used on segments (MFCCs) before a majority vote.
-\cite{lee2013acoustic} NHL system, unsupervised feature learning approach on mel frequency spectrogram with selective Max-pooling.
-Supervised training is performed using SVM.
-\cite{nogueira2013sound} NR1 system, MFCC and spatial features (inter-aural differences) with SVM as classifier.
-\cite{olivetti2013wonder} OE system, normalized compression distance (I honestly did not understand a thing about it, but this the worst method of the challenge anyway) using  Random Forest  classifier.
-\cite{patil2013multiresolution} PE system, 2D Gabor filters applied on 1sec segments with SVM as classifier.
-\cite{rakotomamonjy2015histogram} RG system, histogram of gradient representation, computed from a constant Q transform with SVM as classifier.
-\cite{roma2013recurrence} RNH system, RQA stacked with MFCC, SVM as classifier.
-\\
+\begin{table}
+\begin{center}
+\caption{Designs submitted to the DCASE2013 ASC challenge. RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning.  \label{tab:dcase}}
+\begin{tabular}{lcl}
+& Acc. \%  &  \multicolumn{1}{c}{Design}  \\
+\cite{olivetti2013wonder} OE & 17 & compression distance, Random Forests \\
+\cite{elizalde2013vector} ELF & 55 & I vector approach, LDA \\
+\cite{krijnders2013tone} KH & 55 & cochleogram, SVMs \\
+\cite{patil2013multiresolution} PE & 58 & 2D Gabor filters, SVMs \\
+\cite{lee2013acoustic} NHL & 60 & UL of Mel spectra, SVMs \\
+\cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
+\cite{chum2013ieee} CHR & 65 & \ml{gregoire ?}, HMMs  \\
+\cite{geiger2013large} GSR & 69 & Mel and MFCCs, SVMs \\
+\cite{rakotomamonjy2015histogram} RG & 69  & histogram of gradient (HOG), SVMs \\
+\cite{li2013auditory} LTT & 72 & MFCCs, Tree baggers \\
+\cite{roma2013recurrence} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
+\end{tabular}
+\end{center}
+\end{table}
 
-%\gl{Other paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory}
+In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
+Also, we resort to the use of SVMs with default parameters (C=1) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
+Moreover, since the MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.
+
+%\gl{Othe this r paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory}
 
@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test1.eps}
-\caption{... \ml{fusionner les deux figures et commencer x a 1 par pas de 1}}
+\caption{... \ml{super, c'est propre}}
 \end{center}
@@
 \includegraphics[width=\columnwidth]{gfx/supervised_rbf_test2.eps}
-\caption{...}
+\caption{\ml{simplifier les xticklabels en mettant des acronymes expliquer dans la caption}}
 \end{center}

### 2016-07-10 paper/structureScene.tex

@@
 \usepackage{pgf}
-%\usepackage{tikz}
-%\usetikzlibrary{arrows,positioning}
-%\usetikzlibrary{arrows.meta}
+\usepackage{tikz}
+\usetikzlibrary{arrows,positioning}
+\usetikzlibrary{arrows.meta}
 \usepackage{url}
@@
 
-%\begin{figure}
-%\begin{center}
-%\begin{tikzpicture}
-%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-%\node (omlabel1) at (3.5, 1) {$t$};
-%\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
-%\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
-%\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
-%\node at (2, -0.25) {(a)};
-%\node (omlabel2) at (8, 1) {$t$};
-%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
-%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
-%\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
-%\node at (6.5, -0.25) {(b)};
-%\end{tikzpicture}
-%\caption{
-%\label{fig:gammatones}
-%Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+\node (omlabel1) at (3.5, 1) {$t$};
+\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
+\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
+\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
+\node at (2, -0.25) {(a)};
+\node (omlabel2) at (8, 1) {$t$};
+\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
+\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
+\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
+\node at (6.5, -0.25) {(b)};
+\end{tikzpicture}
+\caption{
+\label{fig:gammatones}
+Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
 Blue and red oscillations represent the real and imaginary parts.
 The orange envelope represents the complex modulus.}
-%\end{center}
-%\end{figure}
+\end{center}
+\end{figure}
 

### 2016-07-10 paper/structureScene.tex
align axes for Fig 2 on log compression

@@
 Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
-\vl{Qu'en penses-tu ML ?}
-\ml{tres bien, mettre les deux distributions de couleurs differentes}
 
@@
 \begin{center}
-\setlength{\unitlength}{1cm}
-\begin{picture}(5,2)
- \put(-1.5,0.0){\includegraphics[height=2cm,width=3.5cm]{feature_histogram.png}}
- \put(2.5,0.0){\includegraphics[height=2cm,width=3.5cm]{feature_histogram_logcompressed.png}}
-\end{picture}
+\begin{tikzpicture}
+\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+\node[anchor=south west,inner sep=0] at (-0.01, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
+\node (histlabel1) at (3, 0.01) {$\mathbf{S}\boldsymbol{x}$};
+\draw[-{Latex[length=2mm]}] (0.05, 0.01) -- (histlabel1);
+\draw[-{Latex[length=2mm]}] (0.05, 0.01) -- (0.05, 2.5);
+\node at (1.5, -0.25) {(a)};
+\node (histlabel2) at (8, 0.01) {$\log \mathbf{S}\boldsymbol{x}$};
+\node[anchor=south west,inner sep=0] at (4, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
+\draw[-{Latex[length=2mm]}] (3.8, 0.01) -- (histlabel2);
+\draw[-{Latex[length=2mm]}] (5, 0.01) -- (5, 2.5);
+\node at (6, -0.25) {(b)};
+\end{tikzpicture}
 \caption{
-\label{fig:histograms}
-\vl{we should discuss color schemes and axis labels}}
+\label{fig:gammatones}
+Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
+Blue and red oscillations represent the real and imaginary parts.
+The orange envelope represents the complex modulus.}
 \end{center}

### 2016-07-10 paper/structureScene.tex
change title: Scattering Transforms -> Wavelet Scattering
RFC
Finish tikzing Fig 2

--- 2016-07-10 paper/structureScene.tex
@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Object-based Auditory Scenes Similarity Retrieval and Classification With Scattering Transforms}
+\title{Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering}
 
@@
 \begin{center}
-\setlength{\unitlength}{1cm}
-\begin{picture}(5,2)
- \put(-1.5,0.0){\includegraphics[height=2cm,width=3.5cm]{feature_histogram.png}}
- \put(2.5,0.0){\includegraphics[height=2cm,width=3.5cm]{feature_histogram_logcompressed.png}}
-\end{picture}
+\begin{tikzpicture}
+\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
+\node (histlabel1) at (2.5, 0.01) {$\mathbf{S}\boldsymbol{x}$};
+\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
+\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
+\node at (1, -0.35) {(a)};
+\node (histlabel2) at (7.8, 0.01) {$\log \mathbf{S}\boldsymbol{x}$};
+\node[anchor=south west,inner sep=0] at (3.3, 0) {\includegraphics[height=2cm,width=3.5cm]{feature_histogram_logcompressed.png}};
+\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
+\draw[-{Latex[length=2mm]}] (5, 0.01) -- (5, 2.5);
+\node at (5.5, -0.35) {(b)};
+\end{tikzpicture}
 \caption{
-\label{fig:histograms}
-\vl{we should discuss color schemes and axis labels}}
+\label{fig:gammatones}
+Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
+Blue and red oscillations represent the real and imaginary parts.
+The orange envelope represents the complex modulus.}
 \end{center}

### 2016-07-10 paper/structureScene.tex
caption Fig 2

@@
 \node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
-\node (histlabel1) at (2.5, 0.01) {$\mathbf{S}\boldsymbol{x}$};
+\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
 \draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
 \draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
-\node at (1, -0.35) {(a)};
-\node (histlabel2) at (7.8, 0.01) {$\log \mathbf{S}\boldsymbol{x}$};
-\node[anchor=south west,inner sep=0] at (3.3, 0) {\includegraphics[height=2cm,width=3.5cm]{feature_histogram_logcompressed.png}};
+\node at (0.8, -0.35) {(a)};
+\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
+\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
 \draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
-\draw[-{Latex[length=2mm]}] (5, 0.01) -- (5, 2.5);
-\node at (5.5, -0.35) {(b)};
+\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
+\node at (4.9, -0.35) {(b)};
 \end{tikzpicture}
 \caption{
-\label{fig:gammatones}
-Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
-Blue and red oscillations represent the real and imaginary parts.
-The orange envelope represents the complex modulus.}
+\label{fig:histograms}
+Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
+(a) before and (b) after logarithmic compression.}
 \end{center}

### 2016-07-10 paper/structureScene.tex
revise abstract

@@
 
-This paper introduces a generic hierarchical modeling of acoustic scenes using features computed using the time scattering transform  for time spans below the second and object-based segmentation for larger time spans that effectively model the scene.
+This paper introduces a generic hierarchical modeling of acoustic scenes using the wavelet scattering features for time spans below the second and object-based segmentation for larger time spans that effectively model the scene.
 
 Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets and state-of-the-art classifier demonstrates its potential.
-More precisely, features derived from the time scattering transform consistently outperform Mel-frequency cepstral coefficients (MFCCs) for both tasks.
+More precisely, features derived from the scattering transform consistently outperform Mel-frequency cepstral coefficients (MFCC) for both tasks.
 
 Representing the scene at larger time scale as a set of examplars of coherent segments is an effective approach for similarity retrieval as it allows the retrieval system to focus on discriminative features.
-This benefit is reduced in supervised settings, probably because the much of the discriminability potential of the approach is already learned by the classifier.
+This benefit is reduced in supervised settings, probably because much of the discriminability potential of the approach is already learned by the classifier.
 

### 2016-07-10 paper/structureScene.tex
reformulate first paragraph
* avoid repetitions of « interest », « area »
* the subordinate clause precedes the principal clause
* fix grammar

@@
 
-\IEEEPARstart{O}{ver} the past decades, the amount of audio data recorded from our sonic environment has considerably grown.
-Recent research areas such as eco-acoustics \cite{ECOACOUSTICS2014, krause} start to consider massive deployment of acoustic sensors around the world in order to measure potential animal biodiversity modification over large temporal scales due to human activity or climate change \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
-
-Also, some recent works demonstrate the interest of such approaches for the characterization of human acoustic pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} and the prediction of annoyance due to traffic \cite{gloaguen}.
-We believe that those case study are of interest for the signal processing community as they have strong societal impacts and raises interesting research avenues.
+\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has considerably grown over the past decades.
+In order to measure the modification of animal biodiversity due to human activity or climate change, the research field of eco-acoustics \cite{ECOACOUSTICS2014, krause} is recently undertaking the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
+In addition, some recent works demonstrate the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
+Because they bear a strong societal impact and raise many scientific challenges, we believe that these case studies are of interest for the signal processing community.
 

### 2016-07-10 paper/structureScene.tex
revise introduction
above cited -> aforementioned
then -> subsequently
well designed -> well-designed

@@
 
-Unfortunately, those fields of research are still in infancy and consequently very few well built datasets are readily available for evaluation purposes for the signal processing community.
-A closely related field of investigation that is relatively more mature is the classification of acoustic scenes whose aim is to predict labels by processing audio data.
-Despite being narrower in terms of range of applications, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and been considered in the data processing field for a while, with availability of some well designed evaluation datasets.
+Unfortunately, the field of signal-based eco-acoustics is still in infancy.
+As a consequence, very few well built datasets are readily available for evaluation purposes.
+A closely related field of investigation, which is relatively more mature, is the classification of acoustic scenes whose aim is to predict labels by processing audio data.
+Despite being narrower in terms of range of applications, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and been considered in the data mining community for a while, with availability of a few well-designed datasets.
 
-Despite the differences, the following question is valid both for the above cited future application and the acoustic scene classification (ASC) task.
-How can we efficiently and effectively process large amount of audio data using representations that are both compact, generic in their design and with a large potential application range ?
+Despite the differences, the following question is valid both for the aforementioned future application and the acoustic scene classification (ASC) task: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and with a large range of potential applications ?
 
-The research described in this paper attempt to provide answers to that question by demonstrating the usefulness of two main contributions: 1) the use of the time scattering transform allows us to extract discriminative features with very few parameter optimization and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene  allows us to effectively focus the description of the scene on areas of interest.
+The research described in this paper attempt to provide answers to this question by demonstrating the usefulness of two main contributions: 1) the use of the scattering transform allows us to extract discriminative features with few parameter optimization and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene allows to effectively focus the description of the scene on areas of interest.
 
-The proposed approach is motivated in Section \ref{sec:motivations}  and a brief review the state of the art in the field of acoustic scene modeling is performed in Section \ref{sec:soa}.
+The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}.
 Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively.
-Experiments carried out in an unsupervised setting, \textit{i.e.} acoustic scene similarity (ASS) and in a supervised setting, \textit{i.e.}  acoustic scene classification (ASC) are then reported in Section \ref{sec:experiments}.
+Experiments carried out in an unsupervised setting, \textit{i.e.} acoustic scene similarity (ASS) and in a supervised setting, \textit{i.e.}  acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:experiments}.
 

### 2016-07-10 paper/structureScene.tex
\textit{i.e.} -> \ie (not italic)

@@
 Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively.
-Experiments carried out in an unsupervised setting, \textit{i.e.} acoustic scene similarity (ASS) and in a supervised setting, \textit{i.e.}  acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:experiments}.
+Experiments carried out in an unsupervised setting, \ie acoustic scene similarity (ASS) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:experiments}.
 

### 2016-07-10 paper/structureScene.tex
revise II. paragraph 1

@@
 
-Unfortunately, the field of signal-based eco-acoustics is still in infancy.
+Unfortunately, the field of signal-based ecoacoustics is still in infancy.
 As a consequence, very few well built datasets are readily available for evaluation purposes.
@@
 
-As stated in the introduction, the field of eco-acoustics is relatively young, but significant works are worth mentioning.
-One avenue of research issued from bio acoustic studies, consider the audio stream as a set of events produced animals, and is processed to retrieve the precise species that produced the sound and potentially the type of vocalization in order to predict some behavior of the animal under study \cite{au2008principles, stowell2014large}.
-Another, more recent, avenue of research, tackles problems associated with long duration recordings of the environment, in particular, building recognizers for species of interest, extracting acoustic indices to aid navigation and visualisation.
-It typically process the audio stream in an holistic manner in order to infer large time-scale properties of the bio-acoustic scene for example to estimate which part of the stream need to be analyzed more in details, \textit{i.e.} by a human \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+As stated in the introduction, the field of ecoacoustics is relatively young, but significant works are worth mentioning.
+One avenue of research, originated in bioacoustics, intends to retrieve the species that produced a given sonic event of relatively short duration, in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
+A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording.
+Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices, migration patterns, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
 

### 2016-07-10 paper/structureScene.tex
reformulate section II

@@
 A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording.
-Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices, migration patterns, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \vl{ref}, migration patterns \vl{ref}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
 
-The same dichotomy applies for the field of audio scene analysis, where the interest can be whether 1) to isolate and identify precise events, \textit{i.e.} acoustic event detection (AED) or 2) to predict global attributes assigned to large time spans of audio, \textit{i.e.} ASC.
-For the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag} where the audio is modelled as high level statistics computed from a set of local features.
-Even though this approach as the advantage of high genericity and follows some evidence found by psycho perception research \cite{mcdermott2013summary}, for the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere point averaging of the features  \cite{lagrange:hal-01082501}.
+The same dichotomy applies to the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
+In the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag}, which consists in modeling audio by high-level summary statistics computed from a set of local features.
+As it enjoys a high genericity, this approach is found in many audio classification tasks besides ASC.
+Yet, the specific implementation with gaussian mixture models (GMM) of Mel-frequency cepstral coefficients (MFCC) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
+Indeed, the typical morphology of acoustic scenes is of a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}, \ie few impulsive sounds superimposed upon a stationary background.
+Therefore, it is arguably more beneficial to focus on the temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
+Interestingly, this statement is validated by auditory psychology as well as sound synthesis based  on summary statistics \cite{mcdermott2013summary}.
 
-Intuitively, this result indicates that not all the features distribution is not of equal interest while modeling acoustic scenes, which is coherent with the typical morphology of acoustic scene which are relatively scarce in terms of events, even if those events can trigger fast perceptual recognition for humans.
-We thus could consider the acoustic scene as a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}.
-
-If the precise list of events that occur in a given time is unachievable to get even for trained humans, research shows that using only a few events, so called markers \cite{} is enough to reliably predict high level attributes.
-For example, the presence of a few bird calls in a smooth traffic hubhub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
+Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{} \vl{ref}.
+For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
+\vl{Missing transition}
 

### 2016-07-10 paper/structureScene.tex
redaction of section VI

@@
 \item step 3: Centroids similarity.
-The similarity between the centroids is computed using a radial basis function kernel $K^c$ as well as the local scaling method proposed in \cite{selfTuneManor2004}: \\
+The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
 \begin{equation}
 \begin{split}
-K^c_{ij} & =exp\left( - \dfrac{\parallel x_i - x_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
-\sigma_{i} & =\parallel x_i - x_N \parallel
+K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
+\sigma_{i} & =\parallel c_i - c_N \parallel
 \end{split}
@@
 
-with $x_N$ the $N$'th nearest neihgbor of the centroid $x_i$. \\
+with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$. \\
 
@@
 
-To compute the scene similarity based on the extracted centroids, we propose the three following strategies:
+The scene similarity is computed based on the extracted centroids. We propose the three following strategies:
 
@@
 \item \emph{averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
-\item \emph{weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster. 
+\item \emph{weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
 \end{itemize}
 
-To compute the  weighted similarity between two scenes by taking into account both theirs clusters weights and theirs centroids similarities, we use a variant of the Earth Moving Distance ($EMD$) adapted for non-normalized histograms known as $\widehat{EMD}$ \cite{pele2008linear} with the implementation proposed in  \cite{pele2009fast}.
-Considering $h_i$ and $h_j$, the histograms of the centroids weights of the scenes $s_i$ and $s_j$ respectively, the $\widehat{EMD}$ is computed as follow:
+Compared to the closest or averaged approach, the  weighted similarity between two scenes takes into account both the clusters weights and the centroids similarities.
+To do so, each scene $s_u$ is viewed as a signature $S_u$ of $m$ clusters ($S_u=\lbrace(c_1^u,w_1^u),\ldots,(c_m^u,w_m^u)\rbrace$), with $c_i^u$ and $w_i^u$ being the centroid and the weight of the $i$th cluster of the scene $s_u$ respectively.
+
+To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ introduced in \cite{pele2008linear}.
+$EMD$ is a cross-bin histogram distance, widely used in computer vision (\cite{zhang2007local}), which compute the minimal cost to be paid to transform one histogram into the other.
+The advantage of using the $EMD$ is that it does not require bins to be aligned.
+The alignment problem is solved by using a ``\,ground distance\,'', being the distance between the bins representatives of the two histograms.
+In our case, the histograms are the weights $w$, and the bins representative the centroids $c$.\\
+
+$\widehat{EMD}$ is a variant version of the $EMD$ adapted for non-normalized histograms.
+To compute the $\widehat{EMD}$, we use the implementation proposed in \cite{pele2009fast}.
+Given two signatures $S_u$ and $S_v$, the $\widehat{EMD}$ is computed by solving the following linear programming problem:
 
@@
 \begin{split}
-\widehat{EMD}(h_i,h_j) &=( \min\limits_{f_{ij}} \sum\limits_{i,j} f_{ij}D_{ij} )  \\ 
-&+ \mid \sum\limits_{i} h_i - \sum\limits_{j} h_j  \mid \alpha \max\limits_{i,j} D_{ij}
+\widehat{EMD}(S_u,S_v) &=( \min\limits_{\lbrace f_{ij}\rbrace} \sum\limits_{i,j} f_{ij}D_{ij} )  \\ 
+&+ \mid \sum\limits_{i} w_i^u - \sum\limits_{j} w_j^v  \mid  \max\limits_{i,j}\lbrace  D_{ij}\rbrace
 \end{split}
@@
 \begin{equation*}
-s.t. \quad f_{ij}\geq0\quad \sum\limits_{j} f_{ij} \leq h_{i}\quad \sum\limits_{i} f_{ij} \leq h_{j} 
-\end{equation*}
-\begin{equation*}
-\sum\limits_{i,j}f_{ij} = \min( \sum\limits_{i} h_{i} ,\sum\limits_{j} h_{j} )
+s.t. \quad f_{ij}\geq0 \quad \sum\limits_{j} f_{ij} \leq w_i^u \quad \sum\limits_{i} f_{ij} \leq w_j^v 
 \end{equation*}
 
-where $f_{ij}$ is the flow, and $D_{ij}$ is the ground distance. In our case, the ground distance is $D_{ij}=1-K^c_{ij}$.
+\begin{equation*}
+\sum\limits_{i,j}f_{ij} = \min( \sum\limits_{i} w_i^u ,\sum\limits_{j} w_j^v )
+\end{equation*}
 
-To get the final similarity measure between two scenes $s_i$ and $s_j$, we use an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support}.
+where $f_{ij}$ is the optimal flow between the clusters weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to the $j$th bin.
+$D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, which is equal to $D_{ij}=1-K^c_{ij}$.\\
+
+To obtain the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support} is computed:
 
 \begin{equation*}
-K^s(s_i,s_j) = exp\left( - \dfrac{\widehat{EMD}(h_i,h_j)}{A} \right) \\
+K_{uv}^s(s_u,s_v) = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
 \end{equation*}
 
-with $A$ a scaling parameter, set to the mean value of the $\widehat{EMD}$.
-The resulting kernel is called an EMD kernel, and it has to be noted that there is no proof that such kernel is positive definite.
+with $A$ a scaling parameter, set to the mean value of the $\widehat{EMD}$ between all the scenes.
+The resulting kernel $K^s$ is called an EMD kernel, and it has to be noted that there is no proof that such kernel is positive definite.
 \\
-
-Other articles using EMD in vision \cite{zhang2007local}
 

### 2016-07-10 paper/structureScene.tex
write section VII

@@
 \usepackage{pgf}
-\usepackage{tikz}
-\usetikzlibrary{arrows,positioning}
-\usetikzlibrary{arrows.meta}
+%\usepackage{tikz}
+%\usetikzlibrary{arrows,positioning}
+%\usetikzlibrary{arrows.meta}
 \usepackage{url}
@@
 
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-\node (omlabel1) at (3.5, 1) {$t$};
-\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
-\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
-\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
-\node at (2, -0.25) {(a)};
-\node (omlabel2) at (8, 1) {$t$};
-\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
-\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
-\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
-\node at (6.5, -0.25) {(b)};
-\end{tikzpicture}
-\caption{
-\label{fig:gammatones}
-Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
+%\begin{figure}
+%\begin{center}
+%\begin{tikzpicture}
+%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+%\node (omlabel1) at (3.5, 1) {$t$};
+%\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
+%\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
+%\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
+%\node at (2, -0.25) {(a)};
+%\node (omlabel2) at (8, 1) {$t$};
+%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
+%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
+%\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
+%\node at (6.5, -0.25) {(b)};
+%\end{tikzpicture}
+%\caption{
+%\label{fig:gammatones}
+%Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
 Blue and red oscillations represent the real and imaginary parts.
 The orange envelope represents the complex modulus.}
-\end{center}
-\end{figure}
+%\end{center}
+%\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
-\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
-\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
-\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
-\node at (0.8, -0.35) {(a)};
-\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
-\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
-\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
-\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
-\node at (4.9, -0.35) {(b)};
-\end{tikzpicture}
-\caption{
-\label{fig:histograms}
-Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
-(a) before and (b) after logarithmic compression.}
-\end{center}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\begin{tikzpicture}
+%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+%\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
+%\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
+%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
+%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
+%\node at (0.8, -0.35) {(a)};
+%\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
+%\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
+%\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
+%\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
+%\node at (4.9, -0.35) {(b)};
+%\end{tikzpicture}
+%\caption{
+%\label{fig:histograms}
+%Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
+%(a) before and (b) after logarithmic compression.}
+%\end{center}
+%\end{figure}
 
@@
 \begin{equation}
+\label{eq:kc}
 \begin{split}
@@
 \section{Experimental Evaluation}
+\label{sec:ExpEval}
 
@@
 
-The object-based approach is assessed considering both an ASS and an ASC scenario:
+The influence of the temporal-scattering transform and the object-based approach (\emph{ob}) are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
 
-\begin{itemize}
-\item \emph{ASS}: It is an unsupervised scenario.
-Evaluation is performed on the private dataset of the DCASE 2013 challenge, using the precision at rank $k$ ($k=[1,\ldots,9]$) metrics.
-Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the points to classify.
+For the ASS setting, evaluation is performed on the private dataset of the DCASE 2013 challenge.
+The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items.
+The $p@k$ is computed for $k=[1,\ldots,9]$.
+Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
 
-\item \emph{ASC}: It is a supervised scenario.
-The systems are evaluated  with five-fold  stratified cross validation on the  private dataset of the DCASE 2013 challenge.
+For the ASC setting, the systems are evaluated  with five-fold  stratified cross validation on the private dataset of the DCASE 2013 challenge.
 The folds are identical to those used in the original challenge.
-A linear Support Vector Machine (SVM) is used as classifier.
-For object-based approach, we only consider the step 2.
-Consequently, each scene is represented by a set of centroids and each centroid is classified separately.
-The final scene classification is based on a majority vote.
-In this case, the object-based approach may be seen as a mid integration strategy.
-\end{itemize}
+Optimal parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approach and the scaling parameters $N$ of the RBF kernels (see Eq.
+\ref{eq:kc}), are learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.
 
@@
 
-ASC, ASS:   Similarity between scenes features is then computed using a Gaussian kernel with local scaling.
-
-The object-based approach is compared to standard approaches using early integration strategy for the ASS scenario, and both early and late integration strategies for the ASC scenario (see Section \ref{XX}).
+The \emph{ob} approach is compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASS setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{XX}).
 
@@
 
-Two types features are used to describe the scenes:
+Experiments are carried out using temporal-scattering coefficients as well as Mel Frequency Cesptral Coefficients (MFCCs) as features.
+For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients \gl{vincent: windows and hops ?}.
+Different sets of parameters were tested to compute the MFFCs.
+We here report only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
+40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz.
+This setting gives us 600 feature vectors per scenes.
+It has been reported that classifying the full set of frames directly may not be the best approach.
+In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices.
+Each $n$ MFCC slices is then averaged over time.
+As the \emph{ob} approach involves a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering.
+Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.
 
-\begin{itemize}
-\item \emph{temporal-scattering}: Each scene is described by 128 vectors of scattering coefficients. 
-\item \emph{Mel Frequency Cesptral Coefficients} (MFCCs): Different set of parameters were tested to compute the MFFCs, we here present only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
-We use 40 MFCCs (including the 0 energy coefficient) computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz.
-This setting gives us 600 feature vectors per scenes.
-To reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices.
-Each $n$ MFCC slices is then averaged over time.
-As the object-based approach (see \ref{XX}) involves a clustering step, we want to ensure that each scene is described by circa the same number of vectors prior to the clustering, for both the temporal-scattering and the MFCCs.
-To do so, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.
-\end{itemize}
+\subsection{Classifier}
 
-For the ASC scenario, the features are standardized prior to the SVM classification step.
+As most of the 2013 DCASE challenge participant did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting.
+For the \emph{ob} approach, we use XX.
+For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used.
+As for the \emph{ob} approach, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}.
+All features, that is the frames, the pooled frames or the centroids depending on the approach considered, are standardized prior to be filled into the classifier.
+The standardization is computed with respect to the train/test set split of the cross validation scheme (see \ref{}).
 
@@
 \section{Results: Acoustic Scene Classification}
-Support vector machines are large-margin binary discriminators with tunable regularization of misclassified examples.
 

### 2016-07-10 paper/structureScene.tex
add  ref to section stand and eraly/late int

@@
 \subsection{Early \vs late temporal integration}
+\label{sec:eili}
+
 Owing to the scarcity of salient events in many natural scenes,
@@
 \subsection{Standardization}
+\label{sec:stand}
+
 Let $\mathbf{S}\boldsymbol{X}(\gamma,n)$ be a dataset, where the indices $\gamma$ and $n$ respectively denotes features and examples.
@@
 
-To take into account this object-based representation, we propose the following 3-steps algorithm:
+To take into account this object-based representation, we propose the following 4-steps algorithm:
 
@@
 
-The \emph{ob} approach is compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASS setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{XX}).
+The \emph{ob} approach is compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASS setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}).
 
@@
 All features, that is the frames, the pooled frames or the centroids depending on the approach considered, are standardized prior to be filled into the classifier.
-The standardization is computed with respect to the train/test set split of the cross validation scheme (see \ref{}).
+The standardization is computed with respect to the train/test set split of the cross validation scheme (see \ref{sec:stand}).
 

### 2016-07-10 paper/structureScene.tex
update features section

@@
 \subsection{Logarithmic compression}
+\label{sec:logcomp}
+
 Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
@@
 
-%\begin{figure}
-%\begin{center}
-%\begin{tikzpicture}
-%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-%\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
-%\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
-%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
-%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
-%\node at (0.8, -0.35) {(a)};
-%\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
-%\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
-%\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
-%\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
-%\node at (4.9, -0.35) {(b)};
-%\end{tikzpicture}
-%\caption{
-%\label{fig:histograms}
-%Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
-%(a) before and (b) after logarithmic compression.}
-%\end{center}
-%\end{figure}
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
+\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
+\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
+\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
+\node at (0.8, -0.35) {(a)};
+\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
+\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
+\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
+\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
+\node at (4.9, -0.35) {(b)};
+\end{tikzpicture}
+\caption{
+\label{fig:histograms}
+Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
+(a) before and (b) after logarithmic compression.}
+\end{center}
+\end{figure}
 
@@
 \begin{itemize}
-\item \emph{closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
-\item \emph{averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
-\item \emph{weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
+\item \emph{ob-closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
+\item \emph{ob-averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
+\item \emph{ob-weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
 \end{itemize}
@@
 \widehat{EMD}(S_u,S_v) &=( \min\limits_{\lbrace f_{ij}\rbrace} \sum\limits_{i,j} f_{ij}D_{ij} )  \\ 
-&+ \mid \sum\limits_{i} w_i^u - \sum\limits_{j} w_j^v  \mid  \max\limits_{i,j}\lbrace  D_{ij}\rbrace
+&+ \mid \sum\limits_{i} w_i^u - \sum\limits_{j} w_j^v  \mid \alpha \max\limits_{i,j}\lbrace  D_{ij}\rbrace
 \end{split}
@@
 
-where $f_{ij}$ is the optimal flow between the clusters weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to the $j$th bin.
-$D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, which is equal to $D_{ij}=1-K^c_{ij}$.\\
+where $\lbrace f_{ij} \rbrace$ is the optimal flow between the clusters weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to the $j$th bin.
+$D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, which is equal to $D_{ij}=1-K^c_{ij}$.
+As suggested in \cite{pele2009fast}, the $\alpha$ parameter is set to one.\\
 
@@
 \begin{equation*}
-K_{uv}^s(s_u,s_v) = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
+K_{uv}^s = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
 \end{equation*}
@@
 
-The influence of the temporal-scattering transform and the object-based approach (\emph{ob}) are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
+The influence of the temporal-scattering transform and the \emph{ob} are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
 
@@
 For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients \gl{vincent: windows and hops ?}.
+Experiments are conducted with and without the log compression (see Section \ref{sec:logcomp}).
+
 Different sets of parameters were tested to compute the MFFCs.
@@
 
-As most of the 2013 DCASE challenge participant did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting.
-For the \emph{ob} approach, we use XX.
+As most of the 2013 DCASE challenge participants did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting.
+For the \emph{ob} approach, \gl{TODO}.
 For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used.
 As for the \emph{ob} approach, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}.
-All features, that is the frames, the pooled frames or the centroids depending on the approach considered, are standardized prior to be filled into the classifier.
-The standardization is computed with respect to the train/test set split of the cross validation scheme (see \ref{sec:stand}).
+All features, \emph{i.e.} the frames, the pooled frames or the centroids depending on the approach considered, are standardized prior to be filled into the classifier.
+The standardization is computed with respect to the train/test set split of the cross validation scheme (see Section \ref{sec:stand}).
 

### 2016-07-10 paper/structureScene.tex
update section VI

@@
 
-\subsection{Acoustic Scene Similarity and Classification}
-
-The influence of the temporal-scattering transform and the \emph{ob} are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
-
-For the ASS setting, evaluation is performed on the private dataset of the DCASE 2013 challenge.
-The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items.
-The $p@k$ is computed for $k=[1,\ldots,9]$.
-Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
-
-For the ASC setting, the systems are evaluated  with five-fold  stratified cross validation on the private dataset of the DCASE 2013 challenge.
-The folds are identical to those used in the original challenge.
-Optimal parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approach and the scaling parameters $N$ of the RBF kernels (see Eq.
-\ref{eq:kc}), are learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.
-
-\subsection{Comparison with early and late integration}
-
-The \emph{ob} approach is compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASS setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}).
-
 \subsection{Features}
@@
 
+\subsection{Acoustic Scene Similarity and Classification}
+
+The influence of the temporal-scattering transform and the \emph{ob} are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
+
+For the ASS setting, evaluation is performed on the private dataset of the DCASE 2013 challenge.
+The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items.
+The $p@k$ is computed for $k=[1,\ldots,9]$.
+Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
+For each experimental setting, we report the results for the set of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approach and the scaling parameters $N$ of the RBF kernels (see Eq.
+\ref{eq:kc}) giving the best $p@9$.
+
+For the ASC setting, the systems are evaluated  with five-fold  stratified cross validation on the private dataset of the DCASE 2013 challenge.
+The folds are identical to those used in the original challenge.
+Optimal parameters (see above) are this time learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.
+
+\subsection{Comparison with early and late integration}
+
+The \emph{ob} approach is compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASS setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}) \gl{TODO: maj vote}.
+
 \subsection{Classifier}
@@
 As most of the 2013 DCASE challenge participants did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting.
-For the \emph{ob} approach, \gl{TODO}.
+For the \emph{ob} approach, \gl{TODO: kernel used}.
 For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used.

### 2016-07-11 paper/structureScene.tex
update EMD description and object based approach description

--- 2016-07-10 paper/structureScene.tex
@@
 Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively.
-Experiments carried out in an unsupervised setting, \ie acoustic scene similarity (ASS) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:experiments}.
+Experiments carried out in an unsupervised setting, \ie acoustic scene similarity (ASS) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:ExpEval}.
 
@@
 
-Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{} \vl{ref}.
+Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
 For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
@@
 \section{Object-based modeling of the scene}
-
-It has been shown the perception of complex acoustic scene rely on the identification of putative sound sources.
-\ml{GL 2 colonnes}
-
-describe in terms of high level attributes
-
-be they cognitive ones such as pleasantness, or descriptive ones such as the type of environment where the
+\label{sec:object}
 
@@
 
-Whilst the range of applications are very large, so is the need
+Studies in cognitive psychology investigating urban sound environment quality pointed out that low level holistic descriptors (perceived or measured) such as global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness jugdment \cite{kuwano_memory_2003}  relies upon high level attributes such as the identification and assessment of the sound source which composed the scene.
+It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafayPartI}.
+Thus recent researches in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}.
+\gl{debate about the kind of process underlying perception: holistic / event based perception ?}.
+\\
 
-holistic / event based
+From a numerical data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
+Instead the outcomes in sound perception cited above encourage us to adopt an object-based  representation of the scene.
+\gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}.
+\\
 
-skeleton of events on a bed of texture.
+In a first attempt to make use of an object-based (\emph{ob}) representation of the scene, we propose the follow a clustering approach.
+After a features extraction step,  a clustering is performed on each scene features using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
+\\
 
-debate about the kind of process underlying perception
-
-strong indication that the detection of specific events, named markers suffice to trigger class prediction
-
-From a numerical data processing overview, the holistic scheme as a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
-
-If available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas.
-
-To take into account this object-based representation, we propose the following 4-steps algorithm:
+Each scene is then described with a set of centroids. To compute the similarities between the scene based on their extracted centroids, we propose three strategies:
 
 \begin{itemize}
-\item step 1: features extraction.
-\item step 2: scene-wise clustering step.
-A clustering is performed on each scene using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
-Each scene is then describe with a set of centroids.
-\item step 3: Centroids similarity.
+\item \emph{ob-closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
+\item \emph{ob-averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
+\item \emph{ob-weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
+\end{itemize}
+
 The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
@@
 
-\item step 4: scene similarity \\
-\end{itemize}
+Compared to the closest or averaged approach, the  weighted similarity between two scenes takes into account both the centroids similarities and the clusters weights.
+Each scene $s_u$ is described by a signature $S_u$ of $m$ clusters ($S_u=\lbrace(c_1^u,w_1^u),\ldots,(c_m^u,w_m^u)\rbrace$), with $c_i^u$ and $w_i^u$ being the centroid and the weight of the $i$th cluster of the scene $s_u$ respectively.
 
-The scene similarity is computed based on the extracted centroids. We propose the three following strategies:
-
-\begin{itemize}
-\item \emph{closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
-\item \emph{averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
-\item \emph{weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
-\end{itemize}
-
-Compared to the closest or averaged approach, the  weighted similarity between two scenes takes into account both the clusters weights and the centroids similarities.
-To do so, each scene $s_u$ is viewed as a signature $S_u$ of $m$ clusters ($S_u=\lbrace(c_1^u,w_1^u),\ldots,(c_m^u,w_m^u)\rbrace$), with $c_i^u$ and $w_i^u$ being the centroid and the weight of the $i$th cluster of the scene $s_u$ respectively.
-
-To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ introduced in \cite{pele2008linear}.
-$EMD$ is a cross-bin histogram distance, widely used in computer vision (\cite{zhang2007local}), which compute the minimal cost to be paid to transform one histogram into the other.
+To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ and introduced in \cite{pele2008linear}.
+$EMD$ is a cross-bin histogram distance, widely used in computer vision \cite{zhang2007local}, which compute the minimal cost to be paid to transform one histogram into the other.
 The advantage of using the $EMD$ is that it does not require bins to be aligned.
-The alignment problem is solved by using a ``\,ground distance\,'', being the distance between the bins representatives of the two histograms.
-In our case, the histograms are the weights $w$, and the bins representative the centroids $c$.\\
+The alignment problem is solved by using a ``\,ground distance\,'', which is the distance between the bins representatives of the two histograms.
+In our case, the histograms are the cluster weights $w$, and the bins representative the cluster centroids $c$.\\
 
@@
 
-where $f_{ij}$ is the optimal flow between the clusters weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to the $j$th bin.
-$D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, which is equal to $D_{ij}=1-K^c_{ij}$.\\
+where $\lbrace f_{ij} \rbrace$ is the optimal flow between the cluster weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to ``\,supply to the demand\,'' of the $j$th bin.
+$D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, and is computed from the kernel $K^c$:
+
+\begin{equation*}
+D_{ij}=1-K^{c(uv)}_{ij}
+\end{equation*}
+
+with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ belonging to the scenes $s_u$ and $s_v$ respectively.
+As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.\\
 
@@
 
-\begin{equation*}
-K_{uv}^s(s_u,s_v) = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
-\end{equation*}
+\begin{equation}
+\label{eq:ks}
+K_{uv}^s = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
+\end{equation}
 
@@
 
+\subsection{Features}
+
+Experiments are carried out using temporal-scattering coefficients as well as Mel Frequency Cesptral Coefficients (MFCCs) as features.
+For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients \gl{vincent: windows and hops ?}.
+Experiments are conducted with and without the log compression (see Section \ref{sec:logcomp}).
+
+Different sets of parameters were tested to compute the MFFCs.
+We here report only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
+40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz.
+This setting gives us 600 feature vectors per scenes.
+It has been reported that classifying the full set of frames directly may not be the best approach \cite{rakotomamonjy2015histogram}.
+In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices.
+Each $n$ MFCC slices is then averaged over time.
+As the \emph{ob} approach involves a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering.
+Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.
+
 \subsection{Acoustic Scene Similarity and Classification}
 
-The influence of the temporal-scattering transform and the object-based approach (\emph{ob}) are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
+The influence of the temporal-scattering transform and the \emph{ob} are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
 
@@
 Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
+For each experimental setting, we report the results for the set of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approach and the scaling parameters $N$ of the RBF kernels (see Eq.
+\ref{eq:kc}) giving the best $p@9$.
 
@@
 The folds are identical to those used in the original challenge.
-Optimal parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approach and the scaling parameters $N$ of the RBF kernels (see Eq.
-\ref{eq:kc}), are learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.
-
-\subsection{Comparison with early and late integration}
+Optimal parameters (see above) are this time learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.
 
@@
 
-\subsection{Features}
-
-Experiments are carried out using temporal-scattering coefficients as well as Mel Frequency Cesptral Coefficients (MFCCs) as features.
-For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients \gl{vincent: windows and hops ?}.
-Different sets of parameters were tested to compute the MFFCs.
-We here report only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
-40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz.
-This setting gives us 600 feature vectors per scenes.
-It has been reported that classifying the full set of frames directly may not be the best approach.
-In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices.
-Each $n$ MFCC slices is then averaged over time.
-As the \emph{ob} approach involves a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering.
-Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.
-
 \subsection{Classifier}
 
-As most of the 2013 DCASE challenge participant did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting.
-For the \emph{ob} approach, we use XX.
+As most of the 2013 DCASE challenge participants did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting.
+The parameter $C$ is here set to one.
+
+For the \emph{ob} approaches (\emph{weighted}, \emph{closest} and \emph{average}), we use the precomputed kernels described in Section \ref{sec:object}.
 For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used.
-As for the \emph{ob} approach, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}.
-All features, that is the frames, the pooled frames or the centroids depending on the approach considered, are standardized prior to be filled into the classifier.
-The standardization is computed with respect to the train/test set split of the cross validation scheme (see \ref{sec:stand}).
+As for the \emph{ob} approaches, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}.
+All feature vectors, being frames, pooled frames or centroids, depending on the approach considered, are standardized prior to be filled into the classifier.
+The standardization is computed with respect to the train/test set splits of the cross validation scheme (see Section \ref{sec:stand}).
 

### 2016-07-11 paper/structureScene.tex
update od approach => approaches

@@
 
-with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ belonging to the scenes $s_u$ and $s_v$ respectively.
+with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ of the scenes $s_u$ and $s_v$ respectively.
 As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.\\
@@
 Each $n$ MFCC slices is then averaged over time.
-As the \emph{ob} approach involves a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering.
+As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering.
 Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.
@@
 
-The influence of the temporal-scattering transform and the \emph{ob} are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
+The influence of the temporal-scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
 
@@
 Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
-For each experimental setting, we report the results for the set of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approach and the scaling parameters $N$ of the RBF kernels (see Eq.
+For each experimental setting, we report the results for the set of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameters $N$ of the RBF kernels (see Eq.
 \ref{eq:kc}) giving the best $p@9$.
@@
 
-The \emph{ob} approach is compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASS setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}).
+The \emph{ob} approaches are compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASS setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}).
 

### 2016-07-11 paper/structureScene.tex
update section VII A

@@
 \gl{debate about the kind of process underlying perception: holistic / event based perception ?}.
-\\
 
@@
 \gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}.
-\\
 
@@
 After a features extraction step,  a clustering is performed on each scene features using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
-\\
 
@@
 \item \emph{ob-averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
-\item \emph{ob-weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
+\item \emph{ob-weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
 \end{itemize}
@@
 
-with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$. \\
+with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$.
 
@@
 The alignment problem is solved by using a ``\,ground distance\,'', which is the distance between the bins representatives of the two histograms.
-In our case, the histograms are the cluster weights $w$, and the bins representative the cluster centroids $c$.\\
+In our case, the histograms are the cluster weights $w$, and the bins representative the cluster centroids $c$.
 
@@
 with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ of the scenes $s_u$ and $s_v$ respectively.
-As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.\\
+As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.
 
-To obtain the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support} is computed:
+To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support} is computed:
 
@@
 Each $n$ MFCC slices is then averaged over time.
-As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering.
+As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors prior to running the clustering, for both the temporal-scattering and the MFCCs.
 Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.
@@
 Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
-For each experimental setting, we report the results for the set of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameters $N$ of the RBF kernels (see Eq.
-\ref{eq:kc}) giving the best $p@9$.
+For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameter $N$ of the RBF kernels (see Eq.
+\ref{eq:kc}) leading to the best $p@9$.
 

### 2016-07-11 paper/structureScene.tex
update ASS figs

@@
 \begin{itemize}
-\item \emph{ob-closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
-\item \emph{ob-averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
-\item \emph{ob-weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
+\item \emph{ob-closest} (ob-c): the similarity between two scenes is equal to the largest similarity between their centroids.
+\item \emph{ob-averaged} (ob-a): the similarity between two scenes is equal to the average of their centroids similarities.
+\item \emph{ob-weighted} (ob-w): for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
 \end{itemize}
@@
 
-For the \emph{ob} approaches (\emph{weighted}, \emph{closest} and \emph{average}), we use the precomputed kernels described in Section \ref{sec:object}.
+For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}.
 For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used.
@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test1.eps}
-\caption{... \ml{super, c'est propre}}
+\caption{ASS: $p@k$ obtained for the MFCC and the time-scattering with log compression}
+\label{fig:ASS_1}
 \end{center}
@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test2.eps}
-\caption{...}
+\caption{ASS: $p@k$ obtained for the time-scattering with and without log compression (w/o log)}
+\label{fig:ASS_1}
 \end{center}
@@
 
+What
 \subsection{Object-based \vs early integration}
@@
 \begin{center}
-\begin{tabular}{cccc}
+\begin{tabular}{llll}
                     & time-scattering & time-scattering & mfcc \\
@@
              \hline
-early int. (linear)  & 66 (11)        & 62 (6)  & 46 (8)    \\
-early int. (RBF)     & 60 (7)         & 57 (8)  & 60 (8)    \\
-late int.  (linear)  & 75 (5)         & 70 (8)  & 63 (8)   \\
-late int.  (RBF)     & 78 (6)         & 73 (8)  & 72 (10)   \\
-ob-closest.          & 66 (7)         & 58 (6)  & 56 (9)   \\
-ob-averaged.         & 61 (7)         & 54 (8)  & 61 (9)   \\
-ob-weighted.         & 54 (11)        & 43 (10) & 51 (15)   \\
+ei (linear)  & 66 (11)        & 62 (6)  & 46 (8)    \\
+ei (RBF)     & 60 (7)         & 57 (8)  & 60 (8)    \\
+li (linear)  & 75 (5)         & 70 (8)  & 63 (8)   \\
+li (RBF)     & 78 (6)         & 73 (8)  & 72 (10)   \\
+ob-c         & 66 (7)         & 58 (6)  & 56 (9)   \\
+ob-a         & 61 (7)         & 54 (8)  & 61 (9)   \\
+ob-w         & 54 (11)        & 43 (10) & 51 (15)   \\
 \end{tabular}

### 2016-07-11 generateFig/unsupervised.m

@@ -7,7 +7,7 @@ addpath(genpath('functions'))
 dataset='test'; 
 
 
-linewidth=2;
+linewidth=1.6;
 markersize=10;
 
 %% load data
@@ -104,7 +104,7 @@ plot(pak_ei_scat_log,'d-','markersize',markersize,'linewidth',linewidth)
 plot(pak_emd_scat_log,'^-','markersize',markersize,'linewidth',linewidth)
 plot(pak_closest_scat_log,'v-','markersize',markersize,'linewidth',linewidth)
 hold off
-legend({'ei, mfcc','emd, mfcc','closest, mfcc','ei, time-scattering','emd, time-scattering','closest, time-scattering'},'interpreter','none')
+legend({'ei, mfcc','ob-w, mfcc','ob-c, mfcc','ei, time-scattering','ob-w, time-scattering','ob-c, time-scattering'},'interpreter','none')
 legend('boxoff')
 xlabel('k')
 ylabel('p@k')
@@ -115,13 +115,13 @@ figure(2)
 clf
 plot(pak_ei_scat,'o--','markersize',markersize,'linewidth',linewidth)
 hold on
-plot(pak_emd_scat,'x--','markersize',markersize,'linewidth',linewidth)
-plot(pak_closest_scat,'*--','markersize',markersize,'linewidth',linewidth)
+plot(pak_emd_scat,'*--','markersize',markersize,'linewidth',linewidth)
+plot(pak_closest_scat,'x--','markersize',markersize,'linewidth',linewidth)
 plot(pak_ei_scat_log,'d-','markersize',markersize,'linewidth',linewidth)
-plot(pak_emd_scat_log,'v-','markersize',markersize,'linewidth',linewidth)
-plot(pak_closest_scat_log,'^-','markersize',markersize,'linewidth',linewidth)
+plot(pak_emd_scat_log,'^-','markersize',markersize,'linewidth',linewidth)
+plot(pak_closest_scat_log,'v-','markersize',markersize,'linewidth',linewidth)
 hold off
-legend({'ei, w/o log','emd, w/o log','closest, w/o log','ei, w/ log','emd, w/ log','closest, w/ log'},'interpreter','none')
+legend({'ei, w/o log','ob-w, w/o log','ob-c, w/o log','ei','ob-w','ob-c'},'interpreter','none')
 legend('boxoff')
 xlabel('k')
 ylabel('p@k')
[…]

### 2016-07-11 paper/structureScene.tex
write ASS results, update ASS figs

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/unsupervised_test1.eps}
+\includegraphics[width=\columnwidth]{gfx/unsupervised_test3.eps}
 \caption{ASS: $p@k$ obtained for the MFCC and the time-scattering with log compression}
@@
 \caption{ASS: $p@k$ obtained for the time-scattering with and without log compression (w/o log)}
-\label{fig:ASS_1}
+\label{fig:ASS_2}
 \end{center}
@@
 
+This section present evaluation results for the ASS settings. $p@k$ for different settings are showed in the Figures \ref{fig:ASS_1} and \ref{fig:ASS_2}.
+
 \subsection{MFCC \vs Time scattering}
 
-What
+Irrespective of the rank $k$ considered, best result is achieved for the time-scattering with log compression using the \emph{ob-c} approach.
+Overall, the time scattering with the log compression outperform the MFCCs systematically.
+
 \subsection{Object-based \vs early integration}
 
+For both the MFCCs and the time-scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{ei}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
+However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{ei}, which shows that the discriminant information is not carried by all the centroids.
+To take advantage of an object-based representation, a selection step has to be run on the representatives of the scenes.
+
+Lastly, we see that while \emph{ob-c} and \emph{ob-w} perform equivalently for the MFCCs, \emph{ob-c} seems to be able to better characterized the classes than \emph{ob-w} for the time scattering.
+This last observation suggests that weighted a centroid according to the number of frames belonging to it may prove to be a limited solution.
+Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
+On the contrary, two similar environments  \gl{find examples} may shared a lot of similar sound sources \gl{reformulate}.
+
 \subsection{Role of logarithmic compression}
+
+Outcomes for time-scattering with and without log compression are showed in Figure~\ref{fig:ASS_2}.
+It can be seen that the log compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{ei} without the log compression.
+\gl{une idée du pourquoi ?}
 

### 2016-07-11 generateFig/unsupervised.m

@@ -113,15 +113,38 @@ disp('')
 
 figure(2)
 clf
-plot(pak_ei_scat,'o--','markersize',markersize,'linewidth',linewidth)
+plot(pak_ei_scat,'d--','markersize',markersize,'linewidth',linewidth)
 hold on
 plot(pak_emd_scat,'*--','markersize',markersize,'linewidth',linewidth)
 plot(pak_closest_scat,'x--','markersize',markersize,'linewidth',linewidth)
-plot(pak_ei_scat_log,'d-','markersize',markersize,'linewidth',linewidth)
-plot(pak_emd_scat_log,'^-','markersize',markersize,'linewidth',linewidth)
-plot(pak_closest_scat_log,'v-','markersize',markersize,'linewidth',linewidth)
+plot(pak_ei_scat_log,'o-','markersize',markersize,'linewidth',linewidth)
+plot(pak_emd_scat_log,'v-','markersize',markersize,'linewidth',linewidth)
+plot(pak_closest_scat_log,'^-','markersize',markersize,'linewidth',linewidth)
 hold off
 legend({'ei, w/o log','ob-w, w/o log','ob-c, w/o log','ei','ob-w','ob-c'},'interpreter','none')
+legend('boxoff')
+xlabel('k')
+ylabel('p@k')
+box off
+disp('')
+
+figure(3)
+clf
+plot(pak_ei_mfcc,'o--','markersize',markersize,'linewidth',linewidth)
+hold on
+plot(pak_average_mfcc,'*--','markersize',markersize,'linewidth',linewidth)
+plot(pak_emd_mfcc,'v--','markersize',markersize,'linewidth',linewidth)
+plot(pak_closest_mfcc,'^--','markersize',markersize,'linewidth',linewidth)
+
+
+plot(pak_ei_scat_log,'o-','markersize',markersize,'linewidth',linewidth)
+plot(pak_average_scat_log,'*-','markersize',markersize,'linewidth',linewidth)
+plot(pak_emd_scat_log,'v-','markersize',markersize,'linewidth',linewidth)
+plot(pak_closest_scat_log,'^-','markersize',markersize,'linewidth',linewidth)
+
+hold off
+legend({'ei, mfcc','ob-a, mfcc','ob-w, mfcc','ob-c, mfcc', ...
[…]

### 2016-07-11 paper/structureScene.tex
up
update EMD description

@@
 
-with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$.
+with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$, and $\parallel \bullet \parallel$ the euclidean norm.
 
@@
 
-This section present evaluation results for the ASS settings. $p@k$ for different settings are showed in the Figures \ref{fig:ASS_1} and \ref{fig:ASS_2}.
+This section present evaluation results for the ASS settings. The $p@k$ for different settings are showed in the Figures~\ref{fig:ASS_1} and \ref{fig:ASS_2}.
 
@@
 
-For both the MFCCs and the time-scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{ei}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
+For both MFCCs and time-scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{ei}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
 However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{ei}, which shows that the discriminant information is not carried by all the centroids.
@@
 
-Outcomes for time-scattering with and without log compression are showed in Figure~\ref{fig:ASS_2}.
+Outcomes for the time-scattering with and without log compression are showed in Figure~\ref{fig:ASS_2}.
 It can be seen that the log compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{ei} without the log compression.
@@
 
-\subsection{Role of logarithmic compression}
-
 \subsection{MFCC \vs Time scattering}
@@
 \subsection{Object-based \vs early/late integration}
+
+\subsection{Role of logarithmic compression}
 

### 2016-07-11 paper/structureScene.tex
log compression => logarithmic compression

@@
 For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients \gl{vincent: windows and hops ?}.
-Experiments are conducted with and without the log compression (see Section \ref{sec:logcomp}).
+Experiments are conducted with and without the logarithmic compression (see Section \ref{sec:logcomp}).
 
@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test3.eps}
-\caption{ASS: $p@k$ obtained for the MFCC and the time-scattering with log compression}
+\caption{ASS: $p@k$ obtained for the MFCC and the time-scattering with logarithmic compression}
 \label{fig:ASS_1}
@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test2.eps}
-\caption{ASS: $p@k$ obtained for the time-scattering with and without log compression (w/o log)}
+\caption{ASS: $p@k$ obtained for the time-scattering with and without logarithmic compression (w/o log)}
 \label{fig:ASS_2}
@@
 
-Irrespective of the rank $k$ considered, best result is achieved for the time-scattering with log compression using the \emph{ob-c} approach.
-Overall, the time scattering with the log compression outperform the MFCCs systematically.
+Irrespective of the rank $k$ considered, best result is achieved for the time-scattering with logarithmic compression using the \emph{ob-c} approach.
+Overall, the time scattering with the logarithmic compression outperform the MFCCs systematically.
 
@@
 
-Outcomes for the time-scattering with and without log compression are showed in Figure~\ref{fig:ASS_2}.
-It can be seen that the log compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{ei} without the log compression.
+Outcomes for the time-scattering with and without logarithmic compression are showed in Figure~\ref{fig:ASS_2}.
+It can be seen that the logarithmic compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{ei} without the logarithmic compression.
 \gl{une idée du pourquoi ?}

### 2016-07-11 paper/structureScene.tex
update ASC results
some revisions before Eq 12

@@
 
-with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$, and $\parallel \bullet \parallel$ the euclidean norm.
+where $c_N$ is defined as the $N^{\textrm{th}}$ nearest neighbor of the centroid $c_i$, and $\parallel \cdot \parallel$ denotes the Euclidean norm.
 
@@
 \end{tabular}
-\caption{...}
+\caption{FYW: best results for object-based approach with majority vote (on the centroids / on the centroids + on the frames) is circa 75\%}
 \end{center}

### 2016-07-11 paper/structureScene.tex
cite 0brist 2010, Bardeli 2010

@@
 A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording.
-Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \vl{ref}, migration patterns \vl{ref}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \cite{Obrist2010}, trends in population sizes \cite{Bardeli2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
 

### 2016-07-11 paper/structureScene.tex
what does ASSR mean ?

@@
 
-Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
-
+Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR \vl{acronym ?}, much of the work dedicated to acoustic scene modeling is applied to ASC tasks.
 Thus, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this latter task.

### 2016-07-11 paper/structureScene.tex
latexify $C=1$, spell out SVM in first appearance

@@
 In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
-Also, we resort to the use of SVMs with default parameters (C=1) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
+Also, we resort to the use of support vector machines (SVMs) with default parameters ($C=1$) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
 Moreover, since the MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.

### 2016-07-11 paper/structureScene.tex
revise section III

@@
 Also, we resort to the use of support vector machines (SVMs) with default parameters ($C=1$) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
-Moreover, since the MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.
+Moreover, since MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.
 
@@
 
-For most of the experiments reported in this paper, the DCASE2013 datatset is considered \cite{giannoulis2013database, 7100934}.
+Most of the experiments reported in this paper rely on the DCASE2013 \cite{giannoulis2013database, 7100934}.
 To build this dataset, three different recordists visited a wide variety of locations in Greater London over
@@
 systematic variations in the recordings covaried with scene
-type: all recordings were made in moderate weather condi-
-tions, and varying times of day and week, and each recordist
+type: all recordings were made in moderate weather conditions, and varying times of day and week, and each recordist
 recorded each scene type.
-The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a 5 fold cross validation scheme.
-The folds used in this paper are the same as the ones used during the challenge.
+The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a $5$-fold cross validation scheme.
+The folds used throughout this paper are the same as the ones used during the challenge.
 
-As of today, there exist larger publicly available dataset, the
-Rouen dataset \cite{rakotomamonjy2015histogram} and the dCase2016 \cite{Mesaros2016_EUSIPCO} (only the public part).
-The latter is also considered in this paper as a validation of the proposed approach.
+As of today, there exist some larger publicly available datasets, namely the
+Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of dCase2016 \cite{Mesaros2016_EUSIPCO}. The latter is also considered in this paper as a validation of the proposed approach.
 
-As studied in \cite{lagrange:hal-01082501}, the DCASE2013 dataset has interesting intra class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices.
-Also, it is still challenging as state-of-the-art systems based on features design and support vector machines (SVMs) achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
-For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively.
+As studied in \cite{lagrange:hal-01082501}, the DCASE2013 dataset has interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices.
+Moreover, is still highly challenging, because state-of-the-art systems based on feature design and support vector machines (SVMs) achieve 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
+For the sake of comparison, the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves an accuracy of 75 \% on DCASE 2013 while topping at 92 \% on the Rouen dataset.
 On the former, recent advances using hierarchical classifiers architectures reports 84 to 87 \% of accuracy depending on the prior knowledge considered for building the classifier architecture \cite{phan2016label}.

### 2016-07-11 paper/structureScene.tex
better intro of section 4

--- 2016-07-11 paper/structureScene.tex
@@
 \usetikzlibrary{arrows,positioning}
-\usetikzlibrary{arrows.meta}
+%\usetikzlibrary{arrows.meta}
 \usepackage{url}
@@
 A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording.
-Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \cite{Obrist2010}, trends in population sizes \cite{Bardeli2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \vl{ref}, migration patterns \vl{ref}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
 
@@
 Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
-For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
+For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \ml{Gregoire : citations ?}.
 \vl{Missing transition}
+
+\ml{Gregoire, ce paragraph est a merger avec la section motivation: Studies in cognitive psychology investigating urban sound environment quality pointed out that low level features such as (perceived or measured) global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003}  relies upon high level cognitive attributes such as the identification and assessment of the sound source which compose the scene.
+It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafay:hal-01111782}.
+As a consequence, research in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}.
+\gl{debate about the kind of process underlying perception: holistic / event based perception ?}.
 
@@
 
-Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR \vl{acronym ?}, much of the work dedicated to acoustic scene modeling is applied to ASC tasks.
+Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
+
 Thus, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this latter task.
@@
 \begin{center}
-\begin{tikzpicture}
-\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-\node (omlabel1) at (3.5, 1) {$t$};
-\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
-\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
-\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
-\node at (2, -0.25) {(a)};
-\node (omlabel2) at (8, 1) {$t$};
-\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
-\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
-\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
-\node at (6.5, -0.25) {(b)};
-\end{tikzpicture}
+%\begin{tikzpicture}
+%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+%\node (omlabel1) at (3.5, 1) {$t$};
+%\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
+%\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
+%\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
+%\node at (2, -0.25) {(a)};
+%\node (omlabel2) at (8, 1) {$t$};
+%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
+%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
+%\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
+%\node at (6.5, -0.25) {(b)};
+%\end{tikzpicture}
 \caption{
@@
 $\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
-wavelets with one vanishing moment \cite{Venkitaraman2014}, as shown in Figure \ref{fig:gammatones}.
+wavelets with one vanishing moment \cite{Venkitaraman2014} \ml{missing ref}, as shown in Figure \ref{fig:gammatones}.
 In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the widespread Morlet wavelets.
@@
 \begin{center}
-\begin{tikzpicture}
-\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
-\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
-\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
-\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
-\node at (0.8, -0.35) {(a)};
-\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
-\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
-\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
-\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
-\node at (4.9, -0.35) {(b)};
-\end{tikzpicture}
+%\begin{tikzpicture}
+%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+%\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
+%\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
+%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
+%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
+%\node at (0.8, -0.35) {(a)};
+%\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
+%\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
+%\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
+%\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
+%\node at (4.9, -0.35) {(b)};
+%\end{tikzpicture}
 \caption{
@@
 
-Studies in cognitive psychology investigating urban sound environment quality pointed out that low level holistic descriptors (perceived or measured) such as global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness jugdment \cite{kuwano_memory_2003}  relies upon high level attributes such as the identification and assessment of the sound source which composed the scene.
-It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafayPartI}.
-Thus recent researches in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}.
-\gl{debate about the kind of process underlying perception: holistic / event based perception ?}.
+%From a data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
 
-From a numerical data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
-Instead the outcomes in sound perception cited above encourage us to adopt an object-based  representation of the scene.
+\ml{gregoire: une figure avec un clustering et une annotation ?}
+
 \gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}.
 
-In a first attempt to make use of an object-based (\emph{ob}) representation of the scene, we propose the follow a clustering approach.
-After a features extraction step,  a clustering is performed on each scene features using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
+\ml{detail: put in experimental part using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
+}
 
-Each scene is then described with a set of centroids. To compute the similarities between the scene based on their extracted centroids, we propose three strategies:
+As discussed in Section \ref{sec:motivations}, several outcomes in sound perception encourage us to evaluate the validity of an object-based representation of the scene for predicting high-level properties.
+Though, a perfect knowledge of which event occurred at a given time is not feasible.
+In this paper, we thus focus on a simple quantization scheme in order to favor genericity, implemented using a clustering approach where coherent regions of the scene are identified.
+
+One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic} that performs some variances estimates over the whole dataset and not solely on each scene, and reproject each scene in the resulting enhanced features space.
+Here, the aim of this process is to better balance the impact of salient events and texture like sounds on final decision by considering several ways of computing the similarity between scenes based on their centroids :
 
@@
 
-The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
+The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
 \begin{equation}
@@
 
-where $c_N$ is defined as the $N^{\textrm{th}}$ nearest neighbor of the centroid $c_i$, and $\parallel \cdot \parallel$ denotes the Euclidean norm.
+with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$, and $\parallel \bullet \parallel$ the euclidean norm.
 
@@
 
-\section{Experimental Evaluation}
+\ml{Grégoire : if the matrix is not psd, then this is NOT a kernel, reformulate}
+
+\section{Experiments}
 \label{sec:ExpEval}
+
+blurb
 
@@
 
-This assertion shall be challenged in future work by considering larger databases and more diverse application scenarios, such as pleasantness prediction \cite{acta} or ecoacoustics tasks.
+This assertion shall be studied further in future work by considering more complex convolutional networks, larger databases and more diverse application scenarios, such as pleasantness prediction \cite{lafaySoundscapePilot, lafay:hal-01111782} or ecoacoustics tasks.
 

### 2016-07-11 paper/structureScene.tex

@@
 In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
-Also, we resort to the use of SVMs with default parameters (C=1) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
-Moreover, since the MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.
+Also, we resort to the use of support vector machines (SVMs) with default parameters ($C=1$) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
+Moreover, since MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.
 

### 2016-07-11 paper/structureScene.tex
first pass on sections 5 and 6

--- 2016-07-11 paper/structureScene.tex
@@
 
-Different sets of parameters were tested to compute the MFFCs.
-We here report only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
+Different sets of parameters were tested to compute the MFCCs.
+Results here are only reported for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
 40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz.
@@
 It has been reported that classifying the full set of frames directly may not be the best approach \cite{rakotomamonjy2015histogram}.
-In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices.
-Each $n$ MFCC slices is then averaged over time.
-As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors prior to running the clustering, for both the temporal-scattering and the MFCCs.
-Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.
+In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is divided into $n$ non-overlapping $t$-seconds long "texture" windows.
+Each $n$ texture windows are then averaged over time.
+As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors prior to the clustering step, for both the temporal-scattering and the MFCCs.
+Thus, $t$ is set to 250ms, resulting in 120 MFCCs texture windows per scene.
 
-\subsection{Acoustic Scene Similarity and Classification}
+\subsection{Methods}
 
-The influence of the temporal-scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).
+The influence of the temporal-scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (\emph{i.e.} ASSR) and in a supervised setting (\emph{i.e.} ASC).
 
-For the ASS setting, evaluation is performed on the private dataset of the DCASE 2013 challenge.
+For the ASSR setting, evaluation is performed on the private dataset of the DCASE 2013 challenge.
 The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items.
-The $p@k$ is computed for $k=[1,\ldots,9]$.
+The $p@k$ is computed for $k=[1,\ldots,9]$, $9$ being the number of elements that are of the same class than the seed item.
 Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
@@
 
-The \emph{ob} approaches are compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASS setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}).
-
-\subsection{Classifier}
+The \emph{ob} approaches are compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASSR setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}).
+In thsi context early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.
 
@@
 As for the \emph{ob} approaches, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}.
-All feature vectors, being frames, pooled frames or centroids, depending on the approach considered, are standardized prior to be filled into the classifier.
+All feature vectors, being frames, texture windows or centroids are standardized prior to be given to the classifier.
 The standardization is computed with respect to the train/test set splits of the cross validation scheme (see Section \ref{sec:stand}).
 
-\section{Results: Acoustic Scene Similarity}
+\section{Acoustic Scene Similarity Retrieval}
 
@@
 
-\section{Results: Acoustic Scene Classification}
+\section{Acoustic Scene Classification}
 

### 2016-07-11 paper/structureScene.tex
fixing compilation issues

@@
 \usetikzlibrary{arrows,positioning}
-%\usetikzlibrary{arrows.meta}
+\usetikzlibrary{arrows.meta}
 \usepackage{url}
@@
 As a consequence, research in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}.
-\gl{debate about the kind of process underlying perception: holistic / event based perception ?}.
+\gl{debate about the kind of process underlying perception: holistic / event based perception ?}}.
 
@@
 \begin{center}
-%\begin{tikzpicture}
-%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-%\node (omlabel1) at (3.5, 1) {$t$};
-%\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
-%\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
-%\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
-%\node at (2, -0.25) {(a)};
-%\node (omlabel2) at (8, 1) {$t$};
-%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
-%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
-%\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
-%\node at (6.5, -0.25) {(b)};
-%\end{tikzpicture}
+\begin{tikzpicture}
+\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+\node (omlabel1) at (3.5, 1) {$t$};
+\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
+\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
+\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
+\node at (2, -0.25) {(a)};
+\node (omlabel2) at (8, 1) {$t$};
+\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
+\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
+\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
+\node at (6.5, -0.25) {(b)};
+\end{tikzpicture}
 \caption{
@@
 \begin{center}
-%\begin{tikzpicture}
-%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-%\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
-%\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
-%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
-%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
-%\node at (0.8, -0.35) {(a)};
-%\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
-%\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
-%\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
-%\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
-%\node at (4.9, -0.35) {(b)};
-%\end{tikzpicture}
+\begin{tikzpicture}
+\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
+\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
+\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
+\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
+\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
+\node at (0.8, -0.35) {(a)};
+\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
+\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
+\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
+\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
+\node at (4.9, -0.35) {(b)};
+\end{tikzpicture}
 \caption{

### 2016-07-11 paper/structureScene.tex
write transition IV -> V

@@
 
+\vl{missing transition}
+
 \section{Scattering representations \label{sec:scattering}}
@@
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
-\ml{tu peux citer le travail publié a Nature de lewicki, qui montrent que ces formes se retrouvent par ICA de sons naturels http://www.cs.cmu.edu/~learning/talks-2002/02-10-21.lewicki.pdf}
+As such, they can be discovered by unsupervised encoding of natural sounds \cite{Smith2006}.
+
+The next section introduces some techniques in feature design to improve auditory scene classification with scattering coefficients.
 

### 2016-07-11 paper/structureScene.tex
width of fig should be ~8cm, not \columnwidth
add missing ref to Venkitaraman2014

@@
 $\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
-wavelets with one vanishing moment \cite{Venkitaraman2014} \ml{missing ref}, as shown in Figure \ref{fig:gammatones}.
+wavelets with one vanishing moment \cite{Venkitaraman2014}, as shown in Figure \ref{fig:gammatones}.
 In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the widespread Morlet wavelets.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/unsupervised_test3.eps}
+\includegraphics[width=8cm]{gfx/unsupervised_test3.eps}
 \caption{ASS: $p@k$ obtained for the MFCC and the time-scattering with logarithmic compression}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/unsupervised_test2.eps}
+\includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
 \caption{ASS: $p@k$ obtained for the time-scattering with and without logarithmic compression (w/o log)}

### 2016-07-11 paper/structureScene.tex
@ Grégoire: specify T in subsection VII. A

@@
 Experiments are carried out using temporal-scattering coefficients as well as Mel Frequency Cesptral Coefficients (MFCCs) as features.
-For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients \gl{vincent: windows and hops ?}.
+For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients, with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=188\,\mathrm{ms}$.
 Experiments are conducted with and without the logarithmic compression (see Section \ref{sec:logcomp}).

### 2016-07-11 paper/structureScene.tex
remove mention to time-frequency scattering
time-scattering -> scattering

@@
 
-\subsection{MFCC \vs Time scattering}
+\subsection{MFCC \vs scattering transform}
 
-Irrespective of the rank $k$ considered, best result is achieved for the time-scattering with logarithmic compression using the \emph{ob-c} approach.
-Overall, the time scattering with the logarithmic compression outperform the MFCCs systematically.
+Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach.
+Overall, log-compressed scattering coefficients systematically outperform MFCCs.
 
@@
 
-For both MFCCs and time-scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{ei}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
+For both MFCCs and scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{ei}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
 However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{ei}, which shows that the discriminant information is not carried by all the centroids.
@@
 
-Lastly, we see that while \emph{ob-c} and \emph{ob-w} perform equivalently for the MFCCs, \emph{ob-c} seems to be able to better characterized the classes than \emph{ob-w} for the time scattering.
+Lastly, we see that while \emph{ob-c} and \emph{ob-w} perform equivalently for the MFCCs, \emph{ob-c} seems to be able to better characterize the classes than \emph{ob-w} for scattering coefficients.
 This last observation suggests that weighted a centroid according to the number of frames belonging to it may prove to be a limited solution.
@@
 In this paper, the benefit of considering 1) the scattering transform for building discriminant features and 2) unsupervised clustering for the modeling of large time span environmental acoustic scenes.
-Experiments shows that the time/frequency scattering achieves better discriminability than the time scattering one or MFCCs as they probably better isolate elements of interest in different information channels, thus allowing an efficient and scalable early integration process.
 This leads to improvement both for the ASSR and ASC tasks.

### 2016-07-11 paper/structureScene.tex
punctuation and spelling in conclusion

@@
 Abstracting the scene as a reduced set of exemplars with adapted similarity strategies allows us to improve ASSR performance when the BOF approaches based on GMMs fails with reference to an early integration scheme.
-Those results advocate for a hierarchical model of the scene, where small time scale structure  (below a second) shall be decorelated using convolutional network such as the ones of the scattering transforms and larger time scale (over a second) structure shall be modeled as adaptive aggregates that can be identified in the time domain in an unsupervised setting and in the features domain in the supervised case (provided that the dimensionality is sufficiently high).
+Those results advocate for a hierarchical model of the scene, where small time scale structure -- \ie below a second -- shall be decorrelated using ca onvolutional network such as in the scattering transform, whereas larger time scales -- \ie over a second -- structure shall be modeled as adaptive aggregates that can be identified in the time domain in an unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.
 

### 2016-07-11 paper/structureScene.tex
time-scattering -> scattering

@@
 \includegraphics[width=8cm]{gfx/unsupervised_test3.eps}
-\caption{ASS: $p@k$ obtained for the MFCC and the time-scattering with logarithmic compression}
+\caption{ASS: $p@k$ obtained for MFCCs and scattering with logarithmic compression}
 \label{fig:ASS_1}
@@
 \includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
-\caption{ASS: $p@k$ obtained for the time-scattering with and without logarithmic compression (w/o log)}
+\caption{ASS: $p@k$ obtained for scattering with and without logarithmic compression (w/o log)}
 \label{fig:ASS_2}
@@
 
-Outcomes for the time-scattering with and without logarithmic compression are showed in Figure~\ref{fig:ASS_2}.
+Outcomes for the scattering with and without logarithmic compression are showed in Figure~\ref{fig:ASS_2}.
 It can be seen that the logarithmic compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{ei} without the logarithmic compression.
@@
 \begin{tabular}{llll}
-                    & time-scattering & time-scattering & mfcc \\
+                    & scattering & scattering & MFCCs \\
                     & w/ log comp.    & w/o log comp.   &      \\
@@
 
-\subsection{MFCC \vs Time scattering}
+\subsection{MFCC \vs scattering}
 

### 2016-07-11 figs/wavelets/gammatone.tex
write a standalone TEX file for exporting gammatones to PDF

--- 2016-07-11 paper/structureScene.tex
@@
 \documentclass[journal]{IEEEtran}
-\usepackage{cite}
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
-\fi
-%
-\usepackage[cmex10]{amsmath}
-
-% *** ALIGNMENT PACKAGES ***
-%
-\usepackage{array}
-\usepackage{fixltx2e}
-
-\usepackage{stfloats}
-% LaTeX2e). It also provides a command:
-\fnbelowfloat
-
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{pgf}
 \usepackage{tikz}
@@
 \usetikzlibrary{arrows.meta}
-\usepackage{url}
-\usepackage{color}
-\usepackage{graphicx}
-\usepackage{xspace}
-\newcommand*{\eg}{e.g.\@\xspace}
-\newcommand*{\ie}{i.e.\@\xspace}
-\newcommand*{\vs}{vs.\@\xspace}
-
-% correct bad hyphenation here
-\hyphenation{wave-let}
-
-\usepackage{color}
-\newcommand{\vl}[1]{\textcolor{orange}{Vincent : #1}}
-\newcommand{\gl}[1]{\textcolor{red}{Gr\'egoire : #1}}
-\newcommand{\ja}[1]{\textcolor{magenta}{Joakim : #1}}
-\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu : #1}}
 
 \begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-% Do not put math or special symbols in the title.
-%\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering}
-
-\author{Vincent Lostanlen, Gr\'egoire Lafay, Joakim And\'en, and Mathieu Lagrange}
-
-% make the title area
-\maketitle
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
-\begin{abstract}
-
-This paper introduces a generic hierarchical modeling of acoustic scenes using the wavelet scattering features for time spans below the second and object-based segmentation for larger time spans that effectively model the scene.
-
-Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets and state-of-the-art classifier demonstrates its potential.
-More precisely, features derived from the scattering transform consistently outperform Mel-frequency cepstral coefficients (MFCC) for both tasks.
-
-Representing the scene at larger time scale as a set of examplars of coherent segments is an effective approach for similarity retrieval as it allows the retrieval system to focus on discriminative features.
-This benefit is reduced in supervised settings, probably because much of the discriminability potential of the approach is already learned by the classifier.
-
-\end{abstract}
-
-% Note that keywords are not normally used for peerreview paper.
-\begin{IEEEkeywords}
-Acoustic scene classification, acoustic scene segmentation.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-\section{Introduction}
-
-\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has considerably grown over the past decades.
-In order to measure the modification of animal biodiversity due to human activity or climate change, the research field of eco-acoustics \cite{ECOACOUSTICS2014, krause} is recently undertaking the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
-In addition, some recent works demonstrate the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
-Because they bear a strong societal impact and raise many scientific challenges, we believe that these case studies are of interest for the signal processing community.
-
-Unfortunately, the field of signal-based ecoacoustics is still in infancy.
-As a consequence, very few well built datasets are readily available for evaluation purposes.
-A closely related field of investigation, which is relatively more mature, is the classification of acoustic scenes whose aim is to predict labels by processing audio data.
-Despite being narrower in terms of range of applications, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and been considered in the data mining community for a while, with availability of a few well-designed datasets.
-
-Despite the differences, the following question is valid both for the aforementioned future application and the acoustic scene classification (ASC) task: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and with a large range of potential applications ?
-
-The research described in this paper attempt to provide answers to this question by demonstrating the usefulness of two main contributions: 1) the use of the scattering transform allows us to extract discriminative features with few parameter optimization and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene allows to effectively focus the description of the scene on areas of interest.
-
-The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}.
-Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively.
-Experiments carried out in an unsupervised setting, \ie acoustic scene similarity (ASS) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:ExpEval}.
-
-\ml{blurb about results}
-
-\section{Motivations} \label{sec:motivations}
-
-As stated in the introduction, the field of ecoacoustics is relatively young, but significant works are worth mentioning.
-One avenue of research, originated in bioacoustics, intends to retrieve the species that produced a given sonic event of relatively short duration, in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
-A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording.
-Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \vl{ref}, migration patterns \vl{ref}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
-
-The same dichotomy applies to the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-In the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag}, which consists in modeling audio by high-level summary statistics computed from a set of local features.
-As it enjoys a high genericity, this approach is found in many audio classification tasks besides ASC.
-Yet, the specific implementation with gaussian mixture models (GMM) of Mel-frequency cepstral coefficients (MFCC) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
-Indeed, the typical morphology of acoustic scenes is of a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}, \ie few impulsive sounds superimposed upon a stationary background.
-Therefore, it is arguably more beneficial to focus on the temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
-Interestingly, this statement is validated by auditory psychology as well as sound synthesis based  on summary statistics \cite{mcdermott2013summary}.
-
-Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
-For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \ml{Gregoire : citations ?}.
-\vl{Missing transition}
-
-\ml{Gregoire, ce paragraph est a merger avec la section motivation: Studies in cognitive psychology investigating urban sound environment quality pointed out that low level features such as (perceived or measured) global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003}  relies upon high level cognitive attributes such as the identification and assessment of the sound source which compose the scene.
-It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafay:hal-01111782}.
-As a consequence, research in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}.
-\gl{debate about the kind of process underlying perception: holistic / event based perception ?}}.
-
-\section{Background} \label{sec:soa}
-
-Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
-
-Thus, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this latter task.
-The table \ref{tab:dcase} summarizes the different algorithmic choices taken by the teams.
-Unless otherwise stated, scene predictions are achieved by majority voting of predictions \ml{gregoire ?}.
-
-\begin{table}
-\begin{center}
-\caption{Designs submitted to the DCASE2013 ASC challenge. RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning.  \label{tab:dcase}}
-\begin{tabular}{lcl}
-& Acc. \%  &  \multicolumn{1}{c}{Design}  \\
-\cite{olivetti2013wonder} OE & 17 & compression distance, Random Forests \\
-\cite{elizalde2013vector} ELF & 55 & I vector approach, LDA \\
-\cite{krijnders2013tone} KH & 55 & cochleogram, SVMs \\
-\cite{patil2013multiresolution} PE & 58 & 2D Gabor filters, SVMs \\
-\cite{lee2013acoustic} NHL & 60 & UL of Mel spectra, SVMs \\
-\cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
-\cite{chum2013ieee} CHR & 65 & \ml{gregoire ?}, HMMs  \\
-\cite{geiger2013large} GSR & 69 & Mel and MFCCs, SVMs \\
-\cite{rakotomamonjy2015histogram} RG & 69  & histogram of gradient (HOG), SVMs \\
-\cite{li2013auditory} LTT & 72 & MFCCs, Tree baggers \\
-\cite{roma2013recurrence} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
-\end{tabular}
-\end{center}
-\end{table}
-
-In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
-Also, we resort to the use of SVMs with default parameters (C=1) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
-Moreover, since the MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.
-
-%\gl{Othe this r paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory}
-
-%\subsection{Datasets}
-
-For most of the experiments reported in this paper, the DCASE2013 datatset is considered \cite{giannoulis2013database, 7100934}.
-To build this dataset, three different recordists visited a wide variety of locations in Greater London over
-a period of months and in each
-scene recorded a few minutes of audio for each scene type. No
-systematic variations in the recordings covaried with scene
-type: all recordings were made in moderate weather condi-
-tions, and varying times of day and week, and each recordist
-recorded each scene type.
-The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a 5 fold cross validation scheme.
-The folds used in this paper are the same as the ones used during the challenge.
-
-As of today, there exist larger publicly available dataset, the
-Rouen dataset \cite{rakotomamonjy2015histogram} and the dCase2016 \cite{Mesaros2016_EUSIPCO} (only the public part).
-The latter is also considered in this paper as a validation of the proposed approach.
-
-As studied in \cite{lagrange:hal-01082501}, the DCASE2013 dataset has interesting intra class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices.
-Also, it is still challenging as state-of-the-art systems based on features design and support vector machines (SVMs) achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
-For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively.
-On the former, recent advances using hierarchical classifiers architectures reports 84 to 87 \% of accuracy depending on the prior knowledge considered for building the classifier architecture \cite{phan2016label}.
-On the latter, the state-of-the-art is currently 95 \% \cite{bisot2016acoustic}.
-
-\vl{missing transition}
-
-\section{Scattering representations \label{sec:scattering}}
-
-Scattering transforms are translation-invariant representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
-Depending upon the chosen architecture, the modulation filterbank may either be performed solely over the time variable or on both time and frequency variables.
-In this section, we present both the temporal scattering transform and the time-frequency scattering transform.
-
-\subsection{Invariance and stability in acoustic scene classification}
-The notion of invariance to translation plays an essential role in auditory scene classification.
-Indeed, the starting time of the recordings are chosen arbitrarily, and thus do not convey any information about the class.
-To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space before training the classifier.
-From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ of  a signal $\boldsymbol{x}(t)$ --
-a translation-invariant representation up to time shifts of duration $T$ can be simply obtained
-by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
-set to $1/T$:
-\begin{equation}
-\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \boldsymbol{\phi}) (t).
-\end{equation}
-As a downside, the transient information in $\boldsymbol{x_1}$ at finer scales than $T$ are lost by this low-pass filtering, hence a lack of discriminability in feature space.
-To address this issue, the temporal scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$, and subsequently applying complex modulus.
-
-By resorting to wavelet transform modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to small time warps,
-in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}.
-Besides invariance to translation, this stability property is crucial in signal classification, as it entails a guarantee of robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.
-
-The scattering transform is defined mathematically as an infinite cascade of nonlinear layers, each consisting of a wavelet transform modulus operator.
-In practice, to achieve invariance to translation up to $T = 186\,\mathrm{ms}$, \ie of the order of the minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ suffice.
-
-The next subsection describes the operations involved in temporal scattering.
-
-\subsection{Temporal scattering}
-Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
-center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$.
-A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
-according to a geometric sequence of scales $2^{\gamma_1/Q_1}$.
-The variable $\gamma_1$, akin to a log-frequency, takes integer values between $0$ and $(J_1 \times Q_1 - 1)$.
-In the sequel, we set $\xi_1$ to 20 kHz, \ie close to the Nyquist frequency of audio recordings ; the number of octaves $J_1$ to $10$, \ie the maximum range of human hearing ; and the number of wavelets per octave $Q_1$ to $8$.
-We denote by
-\begin{equation}
-\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(2^{-\gamma_1/Q_1} t)
-\end{equation}
-the wavelets resulting from the dilation of $\boldsymbol{\psi}$(t).
-For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
-has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$, a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$, and a quality factor of $Q_1$.
-
-The wavelet transform $\boldsymbol{}$ of an audio signal
-$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets.
-Applying pointwise complex modulus to $\boldsymbol{y_1}$ yields
-the wavelet scalogram
-\begin{equation}
-\boldsymbol{x_1}(t, \gamma_1)
-= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t).
-\end{equation}
-The wavelet scalogram bears resemblance to the constant-Q transform (CQT),
-which is derived from the short-term Fourier transform (STFT) by averaging the frequency
-axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}\xi_1$.
-Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
-However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
-theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the temporal resolution of the CQT is fixed by the support of the STFT analyzing window.
-Therefore, the wavelet scalogram has a better temporal localization at high
-frequencies than the CQT, at the expense of a greater computational cost.
-This property allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
-
-Among auditory scenes, amplitude modulations may be caused by a broad variety of mechanical interactions, including collision, friction, and turbulent flow.
-At a longer scale, they also account for some higher-level attributes of sound, \eg prosody in speech or rhythm in music.
-Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a translation-invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
-The amplitude modulation spectrum resulting from this operator is
-\begin{equation}
-\boldsymbol{x_2}(t,\gamma_1,\gamma_2) =
-\vert \boldsymbol{x_1} \ast \boldsymbol{\psi_{\gamma_2}} \vert(t,\gamma_1),
-\end{equation}
-where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order log-frequency index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
-In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation, yielding
-\begin{equation}
-\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
-(\boldsymbol{x_2} \ast \phi)(t,\gamma_1,\gamma_2).
-\end{equation}
-The scattering transform consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
 
@@
 
-Wavelets
-$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
-wavelets with one vanishing moment \cite{Venkitaraman2014}, as shown in Figure \ref{fig:gammatones}.
-In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the widespread Morlet wavelets.
-Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \cite{Fastl2007}.
-Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
-As such, they can be discovered by unsupervised encoding of natural sounds \cite{Smith2006}.
-
-The next section introduces some techniques in feature design to improve auditory scene classification with scattering coefficients.
-
-\section{Feature design}
-Before supervised or unsupervised classification, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power. 
-In this section, we review several ways of achieving these properties, namely binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.
-The former has, to the best of our knowledge, not being considered for ASC and is specific to multichannel recordings, whereas the next ones would also apply to many other kinds of data.
-
-\subsection{Feature selection by restricting the frequency range}
-In a multivariate time series of descriptors $\mathbf{S}\boldsymbol{x}(t,\gamma)$, some coefficients $\gamma$ may consistently bear negligible information for the task of interest.
-This fact may cause the classifier to overfit the training set, especially when the number of features is large and the number of training instances is small, as is the case here.
-Selecting features appropriately before the training stage may circumvent this problem.
-However, the parameters of the chosen feature selection algorithm are themselves prone to overfitting.
-Instead, we adopt the more conservative approach of retaining all coefficients within a certain range of subband indices $\gamma_1$ and $\gamma_2$.
-In comparison with adaptive algorithms, the subset of selected features $\gamma$ has a suboptimal statistical relevance, but it is easily interpretable in the realm of audio signal processing, as it is equivalent to a coarse band-pass filtering of $\boldsymbol{x}(t)$ and $\boldsymbol{x_1}(t,\gamma_1)$ over the time dimension $t$.
-
-The conservation equations underlying the physics of sound production constrain the waveform $\boldsymbol{x}(t)$ to be regular in the time domain, that is, to have a polynomial decay in the Fourier domain.
-This decay is only made sharper by the specific properties of acoustic environments, especially indoor.
-Consequently, the amount of energy in the wavelet subband $\gamma_1$ is roughly inversely proportional to the center frequency $2^{-\gamma_1/Q_1}$.
-In the topmost frequencies, this amount is so faint that the corresponding coefficients $\boldsymbol{x_1}(t,\gamma_1)$ and $\boldsymbol{x_2}(t,\gamma_1,\gamma_2)$ no longer bear discriminative information for the classification task at hand.
-
-In the sequel, we evaluate the impact of discarding the scattering coefficients above a certain acoustic frequency $\xi_{1,\max}$, which corresponds to discarding subbands $\gamma_1$ below a certain scale
-\begin{equation}
-\gamma_{1,\min} =
-\left\lfloor Q_1 \log_2 \frac{\xi_1}{\xi_{1,\max}} \right\rceil,
-\end{equation}
-where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.
-
-Reducing the frequency range is an effective way to enhance invariance, widely used in the literature.
-For instance, \cite{roma2013} achieved the best performance in the 2013 edition of the DCASE challenge, rely on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max}  = 900\,\mathrm{Hz}$ instead of the more conventional value of $16\,\mathrm{kHz}$.
-
-\subsection{Logarithmic compression}
-\label{sec:logcomp}
-
-Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
-Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design.
-It appears empirically that their distribution is skewed towards the right, which means that the tail towards greater values is longer than the tail towards lower values.
-However, skewness can be reduced by applying to all coefficients a pointwise concave transformation, \eg logarithmic.
-Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
-
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
-\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
-\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
-\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
-\node at (0.8, -0.35) {(a)};
-\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
-\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
-\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
-\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
-\node at (4.9, -0.35) {(b)};
-\end{tikzpicture}
-\caption{
-\label{fig:histograms}
-Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
-(a) before and (b) after logarithmic compression.}
-\end{center}
-\end{figure}
-
-The application of the pointwise logarithm to magnitude spectra is ubiquitous in audio signal processing.
-Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
-We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone, which is a spurious factor of variability to the task of auditory scene classification.
-Logarithmic compression can linearize this dependency, which arguably facilitates the construction of a powerful invariant at the classifier stage.
-
-Given a task of musical genre recognition, \cite{Anden2014} has advocated for the renormalization of second-order scattering coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ by the corresponding first-order scattering coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, as it provably decorrelates the amplitudes of their activations.
-Interestingly, taking the logarithm of renormalized coefficients would yield
-\begin{equation}
-\log \dfrac{\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)}{\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)} =
-\log \mathbf{S_2}\boldsymbol{x}(t, \gamma_1, \gamma_2) -
-\log \mathbf{S_1}\boldsymbol{x}(t, \gamma_1),
-\end{equation}
-\ie a linear combination of the logarithms of first- and second-order coefficients.
-Therefore, the theoretical insight brought by \cite{Anden2014} in favor of renormalized scattering also applies to log-scattering up to a linear transformation in feature space, to which affine classifiers are not sensitive.
-
-\subsection{Early \vs late temporal integration}
-\label{sec:eili}
-
-Owing to the scarcity of salient events in many natural scenes,
-fine-grained classification is only made
-possible by integrating signal information over a long temporal context.
-Indeed, whereas a few seconds are often sufficient to recognize a speaker,
-a musical instrument, or a genre, it may require up to 30 seconds
-to disambiguate two classes auditory scenes which share part of their semantic content, \eg a train from a subway station or a quiet street from a park.
-Depending on whether aggregation is performed in feature space or in decision space, the corresponding method is referred to as early or late integration.
-
-A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene, by retaining their average values only.
-Going back to the definition of the scattering transform given in section \ref{sec:scattering}, this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
-With a slight abuse of notation, we denote by
-\begin{equation}
-\mathbf{S}\boldsymbol{x}(\gamma) =
-\int_{-\infty}^{+\infty} \mathbf{S}\boldsymbol{x}(t,\gamma)\;\mathrm{d}t
-\end{equation}
-the summarized features.
-
-Conversely, a late integration scheme relies on probabilistic estimates over short-term windows of length $T$, which are subsequently aggregated to produce a final decision
-\begin{equation}
-\hat{y} = \arg \max_{y} \rho\Big(\big\{ \mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right] \big\}_{t} \Big),
-\end{equation}
-where $\hat{y}$ is the estimated class label and $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.
-
-The principal drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene -- if data augmentation is not considered.
-In the context of the DCASE 2013 dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
-On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene, resulting in $10240$ instances overall.
-However, many of these instances may be silent or lack any salient properties of the class they are assigned to, hence an increase in bias and a risk of underfitting.
-In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff.
-We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
-
-\subsection{Standardization}
-\label{sec:stand}
-
-Let $\mathbf{S}\boldsymbol{X}(\gamma,n)$ be a dataset, where the indices $\gamma$ and $n$ respectively denotes features and examples.
-It is commonly acknowledged that support vector machines should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges.
-This operation may also help making the dataset linearly separable in feature space.
-To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we apply the affine transformation
-\begin{equation}
-\widetilde{\mathbf{S}}\boldsymbol{X}(\gamma, n) =
-\dfrac{ \mathbf{S}\boldsymbol{X}(\gamma, n) -
-\mathbb{E}[ \mathbf{S}\boldsymbol{X}](\gamma)}{\sigma[ \mathbf{S}\boldsymbol{X}](\gamma)}
-\end{equation}
-where $\mathbb{E}[ \mathbf{S}\boldsymbol{X}](\gamma) = \frac{1}{N} \sum_{n=1}^{N} \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the expected value of $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and
-\begin{equation}
-\sigma[\mathbf{S}\boldsymbol{X}] (\gamma) =
-\sqrt{\frac{1}{N-1} \sum_{n=1}^{N}
-\left( \mathbf{S}\boldsymbol{X}(\gamma,n) - \mathbb{E}[\mathbf{S}\boldsymbol{X}] \right)^2}
-\end{equation}
- is the sample standard deviation.
-
- The vectors $\mathbb{E}[\mathbf{S}\boldsymbol{X}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{X}](\gamma)$ are estimated from the training set only, and the same affine transformation is subsequently applied to all samples in the training set and the test set.
-
-The next section discusses adaptive methods for the temporal integration of scattering coefficients in an unsupervised setting.
-
-\section{Object-based modeling of the scene}
-\label{sec:object}
-
-%As part of the aforementioned research areas and applications, the emerging field of \emph{Acoustic Scene Analysis} (also called \emph{Sound Scene Analysis}) \cite{Stowell15} aims to develop approaches and systems for the automatic analysis of environmental sounds and soundscapes (originating both from urban or nature environments).
-While research methodologies in related fields such as Automatic Speech Recognition (ASR) \cite{Rabiner93} and Music Information Retrieval (MIR) \cite{Muller07} are now well established, research addressing Acoustic Scene Analysis remains relatively young.
-
-%From a data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
-
-\ml{gregoire: une figure avec un clustering et une annotation ?}
-
-\gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}.
-
-\ml{detail: put in experimental part using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
-}
-
-As discussed in Section \ref{sec:motivations}, several outcomes in sound perception encourage us to evaluate the validity of an object-based representation of the scene for predicting high-level properties.
-Though, a perfect knowledge of which event occurred at a given time is not feasible.
-In this paper, we thus focus on a simple quantization scheme in order to favor genericity, implemented using a clustering approach where coherent regions of the scene are identified.
-
-One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic} that performs some variances estimates over the whole dataset and not solely on each scene, and reproject each scene in the resulting enhanced features space.
-Here, the aim of this process is to better balance the impact of salient events and texture like sounds on final decision by considering several ways of computing the similarity between scenes based on their centroids :
-
-\begin{itemize}
-\item \emph{ob-closest} (ob-c): the similarity between two scenes is equal to the largest similarity between their centroids.
-\item \emph{ob-averaged} (ob-a): the similarity between two scenes is equal to the average of their centroids similarities.
-\item \emph{ob-weighted} (ob-w): for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
-\end{itemize}
-
-The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
-\begin{equation}
-\label{eq:kc}
-\begin{split}
-K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
-\sigma_{i} & =\parallel c_i - c_N \parallel
-\end{split}
-\end{equation}
-
-with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$, and $\parallel \bullet \parallel$ the euclidean norm.
-
-Compared to the closest or averaged approach, the  weighted similarity between two scenes takes into account both the centroids similarities and the clusters weights.
-Each scene $s_u$ is described by a signature $S_u$ of $m$ clusters ($S_u=\lbrace(c_1^u,w_1^u),\ldots,(c_m^u,w_m^u)\rbrace$), with $c_i^u$ and $w_i^u$ being the centroid and the weight of the $i$th cluster of the scene $s_u$ respectively.
-
-To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ and introduced in \cite{pele2008linear}.
-$EMD$ is a cross-bin histogram distance, widely used in computer vision \cite{zhang2007local}, which compute the minimal cost to be paid to transform one histogram into the other.
-The advantage of using the $EMD$ is that it does not require bins to be aligned.
-The alignment problem is solved by using a ``\,ground distance\,'', which is the distance between the bins representatives of the two histograms.
-In our case, the histograms are the cluster weights $w$, and the bins representative the cluster centroids $c$.
-
-$\widehat{EMD}$ is a variant version of the $EMD$ adapted for non-normalized histograms.
-To compute the $\widehat{EMD}$, we use the implementation proposed in \cite{pele2009fast}.
-Given two signatures $S_u$ and $S_v$, the $\widehat{EMD}$ is computed by solving the following linear programming problem:
-
-\begin{equation}
-\begin{split}
-\widehat{EMD}(S_u,S_v) &=( \min\limits_{\lbrace f_{ij}\rbrace} \sum\limits_{i,j} f_{ij}D_{ij} )  \\ 
-&+ \mid \sum\limits_{i} w_i^u - \sum\limits_{j} w_j^v  \mid \alpha \max\limits_{i,j}\lbrace  D_{ij}\rbrace
-\end{split}
-\end{equation}
-
-\begin{equation*}
-s.t. \quad f_{ij}\geq0 \quad \sum\limits_{j} f_{ij} \leq w_i^u \quad \sum\limits_{i} f_{ij} \leq w_j^v 
-\end{equation*}
-
-\begin{equation*}
-\sum\limits_{i,j}f_{ij} = \min( \sum\limits_{i} w_i^u ,\sum\limits_{j} w_j^v )
-\end{equation*}
-
-where $\lbrace f_{ij} \rbrace$ is the optimal flow between the cluster weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to ``\,supply to the demand\,'' of the $j$th bin.
-$D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, and is computed from the kernel $K^c$:
-
-\begin{equation*}
-D_{ij}=1-K^{c(uv)}_{ij}
-\end{equation*}
-
-with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ of the scenes $s_u$ and $s_v$ respectively.
-As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.
-
-To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support} is computed:
-
-\begin{equation}
-\label{eq:ks}
-K_{uv}^s = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
-\end{equation}
-
-with $A$ a scaling parameter, set to the mean value of the $\widehat{EMD}$ between all the scenes.
-The resulting kernel $K^s$ is called an EMD kernel, and it has to be noted that there is no proof that such kernel is positive definite.
-\\
-
-\ml{Grégoire : if the matrix is not psd, then this is NOT a kernel, reformulate}
-
-\section{Experiments}
-\label{sec:ExpEval}
-
-blurb
-
-\subsection{Features}
-
-Experiments are carried out using temporal-scattering coefficients as well as Mel Frequency Cesptral Coefficients (MFCCs) as features.
-For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients, with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=188\,\mathrm{ms}$.
-Experiments are conducted with and without the logarithmic compression (see Section \ref{sec:logcomp}).
-
-Different sets of parameters were tested to compute the MFCCs.
-Results here are only reported for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
-40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz.
-This setting gives us 600 feature vectors per scenes.
-It has been reported that classifying the full set of frames directly may not be the best approach \cite{rakotomamonjy2015histogram}.
-In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is divided into $n$ non-overlapping $t$-seconds long "texture" windows.
-Each $n$ texture windows are then averaged over time.
-As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors prior to the clustering step, for both the temporal-scattering and the MFCCs.
-Thus, $t$ is set to 250ms, resulting in 120 MFCCs texture windows per scene.
-
-\subsection{Methods}
-
-The influence of the temporal-scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (\emph{i.e.} ASSR) and in a supervised setting (\emph{i.e.} ASC).
-
-For the ASSR setting, evaluation is performed on the private dataset of the DCASE 2013 challenge.
-The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items.
-The $p@k$ is computed for $k=[1,\ldots,9]$, $9$ being the number of elements that are of the same class than the seed item.
-Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
-For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameter $N$ of the RBF kernels (see Eq.
-\ref{eq:kc}) leading to the best $p@9$.
-
-For the ASC setting, the systems are evaluated  with five-fold  stratified cross validation on the private dataset of the DCASE 2013 challenge.
-The folds are identical to those used in the original challenge.
-Optimal parameters (see above) are this time learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.
-
-The \emph{ob} approaches are compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASSR setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}).
-In thsi context early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.
-
-As most of the 2013 DCASE challenge participants did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting.
-The parameter $C$ is here set to one.
-
-For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}.
-For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used.
-As for the \emph{ob} approaches, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}.
-All feature vectors, being frames, texture windows or centroids are standardized prior to be given to the classifier.
-The standardization is computed with respect to the train/test set splits of the cross validation scheme (see Section \ref{sec:stand}).
-
-\section{Acoustic Scene Similarity Retrieval}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=8cm]{gfx/unsupervised_test3.eps}
-\caption{ASS: $p@k$ obtained for MFCCs and scattering with logarithmic compression}
-\label{fig:ASS_1}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
-\caption{ASS: $p@k$ obtained for scattering with and without logarithmic compression (w/o log)}
-\label{fig:ASS_2}
-\end{center}
-\end{figure}
-
-This section present evaluation results for the ASS settings. The $p@k$ for different settings are showed in the Figures~\ref{fig:ASS_1} and \ref{fig:ASS_2}.
-
-\subsection{MFCC \vs scattering transform}
-
-Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach.
-Overall, log-compressed scattering coefficients systematically outperform MFCCs.
-
-\subsection{Object-based \vs early integration}
-
-For both MFCCs and scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{ei}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
-However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{ei}, which shows that the discriminant information is not carried by all the centroids.
-To take advantage of an object-based representation, a selection step has to be run on the representatives of the scenes.
-
-Lastly, we see that while \emph{ob-c} and \emph{ob-w} perform equivalently for the MFCCs, \emph{ob-c} seems to be able to better characterize the classes than \emph{ob-w} for scattering coefficients.
-This last observation suggests that weighted a centroid according to the number of frames belonging to it may prove to be a limited solution.
-Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
-On the contrary, two similar environments  \gl{find examples} may shared a lot of similar sound sources \gl{reformulate}.
-
-\subsection{Role of logarithmic compression}
-
-Outcomes for the scattering with and without logarithmic compression are showed in Figure~\ref{fig:ASS_2}.
-It can be seen that the logarithmic compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{ei} without the logarithmic compression.
-\gl{une idée du pourquoi ?}
-
-\section{Acoustic Scene Classification}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{gfx/supervised_linear_test1.eps}
-\caption{...}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{gfx/supervised_rbf_test2.eps}
-\caption{\ml{simplifier les xticklabels en mettant des acronymes expliquer dans la caption}}
-\end{center}
-\end{figure}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{llll}
-                    & scattering & scattering & MFCCs \\
-                    & w/ log comp.    & w/o log comp.   &      \\
-             \hline
-ei (linear)  & 66 (11)        & 62 (6)  & 46 (8)    \\
-ei (RBF)     & 60 (7)         & 57 (8)  & 60 (8)    \\
-li (linear)  & 75 (5)         & 70 (8)  & 63 (8)   \\
-li (RBF)     & 78 (6)         & 73 (8)  & 72 (10)   \\
-ob-c         & 66 (7)         & 58 (6)  & 56 (9)   \\
-ob-a         & 61 (7)         & 54 (8)  & 61 (9)   \\
-ob-w         & 54 (11)        & 43 (10) & 51 (15)   \\
-\end{tabular}
-\caption{FYW: best results for object-based approach with majority vote (on the centroids / on the centroids + on the frames) is circa 75\%}
-\end{center}
-\end{table}
-
-\subsection{MFCC \vs scattering}
-
-\subsection{Object-based \vs early/late integration}
-
-\subsection{Role of logarithmic compression}
-
-\section{Results: DCASE 2016}
-
-\section{Conclusion}
-
-In this paper, the benefit of considering 1) the scattering transform for building discriminant features and 2) unsupervised clustering for the modeling of large time span environmental acoustic scenes.
-This leads to improvement both for the ASSR and ASC tasks.
-
-Abstracting the scene as a reduced set of exemplars with adapted similarity strategies allows us to improve ASSR performance when the BOF approaches based on GMMs fails with reference to an early integration scheme.
-Those results advocate for a hierarchical model of the scene, where small time scale structure -- \ie below a second -- shall be decorrelated using ca onvolutional network such as in the scattering transform, whereas larger time scales -- \ie over a second -- structure shall be modeled as adaptive aggregates that can be identified in the time domain in an unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.
-
-This assertion shall be studied further in future work by considering more complex convolutional networks, larger databases and more diverse application scenarios, such as pleasantness prediction \cite{lafaySoundscapePilot, lafay:hal-01111782} or ecoacoustics tasks.
-
-\bibliographystyle{unsrt}
-\bibliography{biblio}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-%\begin{IEEEbiography}{Mathieu Lagrange}
-%Biography text here.
-%\end{IEEEbiography}
-
-% if you will not have a photo at all:
-%\begin{IEEEbiographynophoto}{John Doe}
-%Biography text here.
-%\end{IEEEbiographynophoto}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-%\begin{IEEEbiographynophoto}{Jane Doe}
-%Biography text here.
-%\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
 \end{document}

### 2016-07-11 paper/structureScene.tex
minor fix in sec 4
new sectionning of results

@@
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
-As such, they can be discovered by unsupervised encoding of natural sounds \cite{Smith2006}.
+As such, they can be discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
 
@@
 
-\section{Acoustic Scene Similarity Retrieval}
+\section{Results}
 
@@
 
+\subsection{Acoustic Scene Similarity Retrieval}
+
 This section present evaluation results for the ASS settings. The $p@k$ for different settings are showed in the Figures~\ref{fig:ASS_1} and \ref{fig:ASS_2}.
 
-\subsection{MFCC \vs scattering transform}
+\subsection*{MFCC \vs scattering transform}
 
@@
 
-\subsection{Object-based \vs early integration}
+\subsection*{Object-based \vs early integration}
 
@@
 
-\subsection{Role of logarithmic compression}
+\subsection*{Role of logarithmic compression}
 
@@
 
-\section{Acoustic Scene Classification}
+\subsection{Acoustic Scene Classification}
 
@@
 
-\subsection{MFCC \vs scattering}
+MFCC \vs scattering
 
-\subsection{Object-based \vs early/late integration}
+Object-based \vs early/late integration
 
-\subsection{Role of logarithmic compression}
+Role of logarithmic compression
 
-\section{Results: DCASE 2016}
+Results: DCASE 2016
 

### 2016-07-11 paper/structureScene.tex
ASS -> ASSR

@@
 Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively.
-Experiments carried out in an unsupervised setting, \ie acoustic scene similarity (ASS) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:ExpEval}.
+Experiments carried out in an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:ExpEval}.
 
@@
 As it enjoys a high genericity, this approach is found in many audio classification tasks besides ASC.
-Yet, the specific implementation with gaussian mixture models (GMM) of Mel-frequency cepstral coefficients (MFCC) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
+Yet, the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
 Indeed, the typical morphology of acoustic scenes is of a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}, \ie few impulsive sounds superimposed upon a stationary background.
@@
 \begin{tabular}{llll}
-                    & scattering & scattering & MFCCs \\
-                    & w/ log comp.    & w/o log comp.   &      \\
+                    & log-scattering & scattering & MFCCs \\
              \hline
@@
 
-Results: DCASE 2016
+\vl{Results: DCASE 2016}
 

### 2016-07-11 paper/structureScene.tex
cosmetic changes around Eq 12
\parallel -> \Vert ; \bullet -> \cdot in Eq 12
refs Bardeli and Obrist

@@
 A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording.
-Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \vl{ref}, migration patterns \vl{ref}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
 
@@
 
-The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
+The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
 \begin{equation}
-\label{eq:kc}
-\begin{split}
-K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
-\sigma_{i} & =\parallel c_i - c_N \parallel
-\end{split}
+K^c_{ij} = exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_i - c_N \Vert}
 \end{equation}
 
-with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$, and $\parallel \bullet \parallel$ the euclidean norm.
+where $c_N$ is the $N^{\mathrm{th}}$ nearest neihgbor of the centroid $c_i$, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
 

### 2016-07-11 paper/structureScene.tex
latexify

@@
 Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets and state-of-the-art classifier demonstrates its potential.
-More precisely, features derived from the scattering transform consistently outperform Mel-frequency cepstral coefficients (MFCC) for both tasks.
+More precisely, features derived from the scattering transform consistently outperform Mel-frequency cepstral coefficients (MFCCs) for both tasks.
 
@@
 In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
-Also, we resort to the use of SVMs with default parameters (C=1) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
+Also, we resort to the use of SVMs with default parameters ($C=1$) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
 Moreover, since the MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.
@@
 
-\ml{detail: put in experimental part using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
+\ml{detail: put in experimental part using an enhanced version of the popular K-means algorithm known as $K$-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
 }

### 2016-07-11 paper/structureScene.tex
update section III

@@
 systematic variations in the recordings covaried with scene
-type: all recordings were made in moderate weather condi-
-tions, and varying times of day and week, and each recordist
-recorded each scene type.
+type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
 The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a 5 fold cross validation scheme.

### 2016-07-11 paper/structureScene.tex
ref in intro

--- 2016-07-11 paper/structureScene.tex
@@
 Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively.
-Experiments carried out in an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:ExpEval}.
-
-\ml{blurb about results}
+Experiments carried out in an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and in a supervised setting, \ie acoustic scene classification (ASC) are described in \ref{sec:experiments}.
+Results are subsequently reported in Section \ref{sec:results}.
 
@@
 As it enjoys a high genericity, this approach is found in many audio classification tasks besides ASC.
-Yet, the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
+Yet, the specific implementation with gaussian mixture models (GMM) of Mel-frequency cepstral coefficients (MFCC) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
 Indeed, the typical morphology of acoustic scenes is of a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}, \ie few impulsive sounds superimposed upon a stationary background.
@@
 
-\vl{missing transition}
+In this paper, we shall demonstrate that the use of features based on the wavelet scattering together with object-based representation of the acoustic scene allows us to achieve interesting performance both in the ASSR and ASC paradigms.
 
@@
 
-The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
+The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
 \begin{equation}
-K^c_{ij} = exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_i - c_N \Vert}
+\label{eq:kc}
+\begin{split}
+K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
+\sigma_{i} & =\parallel c_i - c_N \parallel
+\end{split}
 \end{equation}
 
-where $c_N$ is the $N^{\mathrm{th}}$ nearest neihgbor of the centroid $c_i$, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$, and $\parallel \bullet \parallel$ the euclidean norm.
 
@@
 \section{Experiments}
-\label{sec:ExpEval}
+\label{sec:experiments}
 
-blurb
-
-\subsection{Features}
+\subsection{Feature design}
 
@@
 
-\section{Results}
+\section{Results \label{sec:results}}
 
@@
 \includegraphics[width=8cm]{gfx/unsupervised_test3.eps}
-\caption{ASS: $p@k$ obtained for MFCCs and scattering with logarithmic compression}
+\caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression}
 \label{fig:ASS_1}
@@
 \includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
-\caption{ASS: $p@k$ obtained for scattering with and without logarithmic compression (w/o log)}
+\caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression (w/o log)}
 \label{fig:ASS_2}
@@
 
-This section present evaluation results for the ASS settings. The $p@k$ for different settings are showed in the Figures~\ref{fig:ASS_1} and \ref{fig:ASS_2}.
+This section present evaluation results for the ASSR settings. The $p@k$ for different settings are showed in the Figures~\ref{fig:ASS_1} and \ref{fig:ASS_2}.
 
@@
 \begin{tabular}{llll}
-                    & log-scattering & scattering & MFCCs \\
+                    & scattering & scattering & MFCCs \\
+                    & w/ log comp.    & w/o log comp.   &      \\
              \hline
@@
 
-\vl{Results: DCASE 2016}
+Results: DCASE 2016
 

### 2016-07-11 paper/structureScene.tex
add \right)
solving gregoire issues

@@
 Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
-For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \ml{Gregoire : citations ?}.
-\vl{Missing transition}
+For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas.
 
@@
 Thus, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this latter task.
-The table \ref{tab:dcase} summarizes the different algorithmic choices taken by the teams.
-Unless otherwise stated, scene predictions are achieved by majority voting of predictions \ml{gregoire ?}.
+The table \ref{tab:dcase} summarizes the different algorithmic choices taken by the teams where most of scene predictions are achieved by majority voting of predictions over features.
 
@@
 \cite{olivetti2013wonder} OE & 17 & compression distance, Random Forests \\
-\cite{elizalde2013vector} ELF & 55 & I vector approach, LDA \\
+\cite{elizalde2013vector} ELF & 55 & I vector, LDA \\
 \cite{krijnders2013tone} KH & 55 & cochleogram, SVMs \\
@@
 \cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
-\cite{chum2013ieee} CHR & 65 & \ml{gregoire ?}, HMMs  \\
+\cite{chum2013ieee} CHR & 65 & sparsity features, HMMs  \\
 \cite{geiger2013large} GSR & 69 & Mel and MFCCs, SVMs \\
@@
 \begin{equation}
-K^c_{ij} = exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_i - c_N \Vert}
+K^c_{ij} = exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_j - c_N \Vert} \right) 
 \end{equation}

### 2016-07-11 paper/structureScene.tex
refactor Fig 1
thank ERC InvariantClass grant
upadte results with full range mfcc, move kmeans++ part

@@
 
+\usepackage{booktabs}
 \usepackage{amssymb}
@@
 
-\author{Vincent Lostanlen, Gr\'egoire Lafay, Joakim And\'en, and Mathieu Lagrange}
+\author{Vincent Lostanlen\thanks{This work is supported by the ERC InvariantClass grant 320959.}, Gr\'egoire Lafay, Joakim And\'en, and Mathieu Lagrange}
 
@@
 \begin{center}
-%\begin{tikzpicture}
-%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-%\node (omlabel1) at (3.5, 1) {$t$};
-%\draw[-{Latex[length=2mm]}] (0, 1) -- (omlabel1);
-%\draw[-{Latex[length=2mm]}] (0, 1) -- (0, 2);
-%\node[anchor=south west,inner sep=0] at (0, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q8.png}};
-%\node at (2, -0.25) {(a)};
-%\node (omlabel2) at (8, 1) {$t$};
-%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (omlabel2);
-%\draw[-{Latex[length=2mm]}] (4.5, 1) -- (4.5, 2);
-%\node[anchor=south west,inner sep=0] at (4.5, 0) {\includegraphics[height=2cm,width=3cm]{gammatone_Q1.png}};
-%\node at (6.5, -0.25) {(b)};
-%\end{tikzpicture}
+\includegraphics[width=\columnwidth]{gammatones.png}
 \caption{
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/supervised_linear_test1.eps}
+\includegraphics[width=\columnwidth]{gfx/supervised_linear_test3.eps}
 \caption{...}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{gfx/supervised_rbf_test2.eps}
-\caption{\ml{simplifier les xticklabels en mettant des acronymes expliquer dans la caption}}
 \end{center}
@@
 \begin{tabular}{llll}
-                    & log-scattering & scattering & MFCCs \\
+              MFCCs &  scattering & log-scattering  \\
              \hline
-ei (linear)  & 66 (11)        & 62 (6)  & 46 (8)    \\
-ei (RBF)     & 60 (7)         & 57 (8)  & 60 (8)    \\
-li (linear)  & 75 (5)         & 70 (8)  & 63 (8)   \\
-li (RBF)     & 78 (6)         & 73 (8)  & 72 (10)   \\
-ob-c         & 66 (7)         & 58 (6)  & 56 (9)   \\
-ob-a         & 61 (7)         & 54 (8)  & 61 (9)   \\
-ob-w         & 54 (11)        & 43 (10) & 51 (15)   \\
+ei (linear)  & 54 (20)       & 62 (6)  & 66 (11)     \\
+ei (RBF)     & 47 (16)       & 56 (7)  & 63 (6)    \\
+li (linear)  & 60  (9)       & 70 (8)  & 75 (5)   \\
+li (RBF)     & 68 (10)       & 73 (8)  & 78 (6)   \\
 \end{tabular}

### 2016-07-11 generateFig/functions/parseTable.m

@@ -1,15 +1,0 @@
-function [ data,leg ] = parseTable(d)
-
-disp('')
-leg=fieldnames(d.rows{1});
-
-tmp=cellfun(@(x) struct2cell(x),d.rows,'UniformOutput',false);
-
-data=cell(length(tmp{1}),length(tmp));
-
-for ii=1:size(data,1)
-    for jj=1:size(data,2)
-        data{ii,jj}=tmp{jj}{ii};
-    end
-end
-

### 2016-07-11 generateFig/unsupervised.m

@@ -1,165 +1,0 @@
-clear all
-addpath(genpath('functions'))
-
-%% settings
-
-% dataset='train'; % test;
-dataset='test'; 
-
-
-linewidth=1.6;
-markersize=10;
-
-%% load data
-d=load(['~/projets/ass_features/report/figures/unsupervised_earlyInt_' dataset]);
-[data{1},leg{1}]=parseTable(d.data.tables{4});
-
-d=load(['~/projets/ass_clustering_features/report/figures/unsupervised_object_' dataset]);
-[data{2},leg{2}]=parseTable(d.data.tables{2});
-
-
-pak_ei=str2double(data{1}(9:17,:));
-pak_obj=str2double(data{2}(8:16,:));
-
-scat_ei=strcmp('scatT',data{1}(1,:));
-mfcc_ei=strcmp('mfcc',data{1}(1,:));
-
-scat_obj=strcmp('scatT',data{2}(1,:));
-mfcc_obj=strcmp('mfcc',data{2}(1,:));
-
-emd=strcmp('emd',data{2}(7,:));
-closest=strcmp('closest',data{2}(7,:));
-average=strcmp('average',data{2}(7,:));
-
-%% best p@9
-
-ind_best_mfcc_ei=find( mfcc_ei & pak_ei(end,:)==max(pak_ei(end,mfcc_ei)));
-ind_best_scat_ei=find( scat_ei & pak_ei(end,:)==max(pak_ei(end,scat_ei)));
-
-best_mfcc_ei=data{1}(:,ind_best_mfcc_ei);
-best_scat_ei=data{1}(:,ind_best_scat_ei);
-
-ind_best_mfcc_emd=find(emd &  mfcc_obj & pak_obj(end,:)==max(pak_obj(end,emd & mfcc_obj)));
-ind_best_mfcc_closest=find(closest & mfcc_obj & pak_obj(end,:)==max(pak_obj(end,closest & mfcc_obj)));
-ind_best_mfcc_average=find(average & mfcc_obj & pak_obj(end,:)==max(pak_obj(end,average & mfcc_obj)));
-
-ind_best_scat_emd=find( emd & scat_obj & pak_obj(end,:)==max(pak_obj(end,emd & scat_obj)));
-ind_best_scat_closest=find(closest & scat_obj & pak_obj(end,:)==max(pak_obj(end,closest & scat_obj)));
[…]

### 2016-07-11 paper/structureScene.tex
features description

@@
 Different sets of parameters were tested to compute the MFCCs.
-Results here are only reported for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}.
-40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz.
+Results here are only reported for the best setting.
+40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a full frequency range.
 This setting gives us 600 feature vectors per scenes.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/supervised_linear_test3.eps}
+\includegraphics[width=\columnwidth]{gfx/supervised_rbf_test3.eps}
 \caption{...}

### 2016-07-11 paper/structureScene.tex
clean ml comments

@@
 
-\ml{gregoire: une figure avec un clustering et une annotation ?}
-
 \gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}.
-
-\ml{detail: put in experimental part using an enhanced version of the popular K-means algorithm known as $K$-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
-}
 
@@
 
-\ml{Grégoire : if the matrix is not psd, then this is NOT a kernel, reformulate}
+%\ml{Grégoire : if the matrix is not psd, then this is NOT a kernel, reformulate}
 
@@
 Optimal parameters (see above) are this time learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.
+Clustering is done using an enhanced version of the popular K-means algorithm known as $K$-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
 

### 2016-07-11 paper/structureScene.tex
clean table II
add label in kc eq.
clean some comments

@@
 \begin{equation}
+\label{eq:kc}
 K^c_{ij} = exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_j - c_N \Vert} \right) 
@@
 Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
-On the contrary, two similar environments  \gl{find examples} may shared a lot of similar sound sources \gl{reformulate}.
+On the contrary, two similar environments may shared a lot of similar sound sources.
 
@@
 \begin{tabular}{llll}
-              MFCCs &  scattering & log-scattering  \\
+             & MFCCs         & scattering & log-scattering  \\
              \hline
@@
 \end{tabular}
-\caption{FYW: best results for object-based approach with majority vote (on the centroids / on the centroids + on the frames) is circa 75\%}
+\caption{Accuracies for the best settings}
 \end{center}

### 2016-07-11 paper/structureScene.tex
update table II entries

@@
              \hline
-ei (linear)  & 54 (20)       & 62 (6)  & 66 (11)     \\
-ei (RBF)     & 47 (16)       & 56 (7)  & 63 (6)    \\
-li (linear)  & 60  (9)       & 70 (8)  & 75 (5)   \\
-li (RBF)     & 68 (10)       & 73 (8)  & 78 (6)   \\
+early (linear)  & 54 (20)       & 62 (6)  & 66 (11)     \\
+early (RBF)     & 47 (16)       & 56 (7)  & 63 (6)    \\
+late (linear)  & 60  (9)       & 70 (8)  & 75 (5)   \\
+late (RBF)     & 68 (10)       & 73 (8)  & 78 (6)   \\
 \end{tabular}

### 2016-07-11 paper/structureScene.tex
ei => early / li => late

@@
 
-The \emph{ob} approaches are compared to commonly used approaches using an early integration strategy (\emph{ei}) for the ASSR setting, and both an early and a late integration (\emph{li}) strategy for the ASC setting (see Section \ref{sec:eili}).
+The \emph{ob} approaches are compared to commonly used approaches using an early integration strategy (\emph{early}) for the ASSR setting, and both an early and a late integration (\emph{late}) strategy for the ASC setting (see Section \ref{sec:eili}).
 In thsi context early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.
@@
 For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}.
-For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used.
+For both the \emph{early} and \emph{late} approaches, a linear kernel as well as a RBF kernel are used.
 As for the \emph{ob} approaches, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}.
@@
 
-For both MFCCs and scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{ei}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
-However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{ei}, which shows that the discriminant information is not carried by all the centroids.
+For both MFCCs and scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
+However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{early}, which shows that the discriminant information is not carried by all the centroids.
 To take advantage of an object-based representation, a selection step has to be run on the representatives of the scenes.
@@
 Outcomes for the scattering with and without logarithmic compression are showed in Figure~\ref{fig:ASS_2}.
-It can be seen that the logarithmic compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{ei} without the logarithmic compression.
+It can be seen that the logarithmic compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{early} without the logarithmic compression.
 \gl{une idée du pourquoi ?}

### 2016-07-11 paper/structureScene.tex
minor changes
merge motivation

@@
 Therefore, it is arguably more beneficial to focus on the temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
+
 Interestingly, this statement is validated by auditory psychology as well as sound synthesis based  on summary statistics \cite{mcdermott2013summary}.
+Indeed studies in cognitive psychology investigating urban sound environment quality pointed out that low level features such as (perceived or measured) global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003}  relies upon high level cognitive attributes such as the identification and assessment of the sound source which compose the scene.
+It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafay:hal-01111782}.
+As a consequence, research in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
 
@@
 For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas.
-
-\ml{Gregoire, ce paragraph est a merger avec la section motivation: Studies in cognitive psychology investigating urban sound environment quality pointed out that low level features such as (perceived or measured) global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003}  relies upon high level cognitive attributes such as the identification and assessment of the sound source which compose the scene.
-It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafay:hal-01111782}.
-As a consequence, research in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}.
-\gl{debate about the kind of process underlying perception: holistic / event based perception ?}}.
 
@@
 \end{tabular}
-\caption{Accuracies for the best settings}
+\caption{ASC: Accuracies for the best settings}
 \end{center}

### 2016-07-11 paper/structureScene.tex
update section VIII and ASR figs with correct MFCCs frequency range

@@
 
-For both MFCCs and scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
+For the scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
 However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{early}, which shows that the discriminant information is not carried by all the centroids.
 To take advantage of an object-based representation, a selection step has to be run on the representatives of the scenes.
-
-Lastly, we see that while \emph{ob-c} and \emph{ob-w} perform equivalently for the MFCCs, \emph{ob-c} seems to be able to better characterize the classes than \emph{ob-w} for scattering coefficients.
+Furthermore It appears that \emph{ob-c} is be able to better characterize the classes than \emph{ob-w}.
 This last observation suggests that weighted a centroid according to the number of frames belonging to it may prove to be a limited solution.
@@
 On the contrary, two similar environments may shared a lot of similar sound sources.
+
+For the MFCCs, there is no clear evidence that the \emph{ob} approaches improve the results.
 

### 2016-07-11 paper/structureScene.tex
add compression.png
improve appearance of Table 2 + add a caption

@@
 \usetikzlibrary{arrows,positioning}
-%\usetikzlibrary{arrows.meta}
 \usepackage{url}
@@
 \begin{center}
-%\begin{tikzpicture}
-%\useasboundingbox[red] (-0.25, -0.5) rectangle (8.25, 2);
-%\node[anchor=south west,inner sep=0] at (-0.36, 0) {\includegraphics[height=2cm,width=5cm]{feature_histogram.png}};
-%\node (histlabel1) at (2.3, 0.01) {$\mathbf{S}\boldsymbol{x}(\gamma)$};
-%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (histlabel1);
-%\draw[-{Latex[length=2mm]}] (-0.3, 0.01) -- (-0.3, 2.5);
-%\node at (0.8, -0.35) {(a)};
-%\node (histlabel2) at (7.6, 0.01) {$\log \mathbf{S}\boldsymbol{x}(\gamma)$};
-%\node[anchor=south west,inner sep=0] at (3.45, 0) {\includegraphics[height=2cm,width=3cm]{feature_histogram_logcompressed.png}};
-%\draw[-{Latex[length=2mm]}] (3.2, 0.01) -- (histlabel2);
-%\draw[-{Latex[length=2mm]}] (4.9, 0.01) -- (4.9, 2.5);
-%\node at (4.9, -0.35) {(b)};
-%\end{tikzpicture}
+\includegraphics[width=\columnwidth]{compression.png}
 \caption{
@@
              \hline
-early (linear)  & 54 (20)       & 62 (6)  & 66 (11)     \\
-early (RBF)     & 47 (16)       & 56 (7)  & 63 (6)    \\
-late (linear)  & 60  (9)       & 70 (8)  & 75 (5)   \\
-late (RBF)     & 68 (10)       & 73 (8)  & 78 (6)   \\
+early, linear  & $54\pm20$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
+early, RBF     & $47\pm16$  & $56\pm\phantom{0}7$  & $63\pm\phantom{0}6$   \\
+late, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
+late, RBF     & $68\pm10$ & $73\pm\phantom{0}8$  & $\mathbf{78}\pm\phantom{0}6$   \\
 \end{tabular}
-\caption{ASC: Accuracies for the best settings}
+\caption{Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.}
 \end{center}

### 2016-07-11 paper/structureScene.tex
rephrase Section VI, paragraph 2

@@
 
-One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic} that performs some variances estimates over the whole dataset and not solely on each scene, and reproject each scene in the resulting enhanced features space.
-Here, the aim of this process is to better balance the impact of salient events and texture like sounds on final decision by considering several ways of computing the similarity between scenes based on their centroids :
+One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, which performs some multivariate estimates over the whole dataset and not solely on each scene, and re-projects each scene in the resulting enhanced feature space.
+Here, with the aim of balancing better the influence of salient events and texture-like sounds on the final decision, we consider several centroid-based similarity metrics between centroids:
 
@@
 \end{equation}
-
-where $c_N$ is the $N^{\mathrm{th}}$ nearest neihgbor of the centroid $c_i$, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+where $c_N$ is the $N^{\textrm{th}}$ nearest neighbor to the centroid $c_i$, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
 

### 2016-07-11 paper/structureScene.tex
exp -> \exp in Eq 12

@@
 \gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}. 
+\vl{this section should be better segmented into subsections. The words centroid and cluster are defined after they are used. The order of paragraphs could be 1-4-3-2-5.}
 
@@
 One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, which performs some multivariate estimates over the whole dataset and not solely on each scene, and re-projects each scene in the resulting enhanced feature space.
-Here, with the aim of balancing better the influence of salient events and texture-like sounds on the final decision, we consider several centroid-based similarity metrics between centroids:
+Here, with the aim of balancing better the influence of salient events and texture-like sounds on the final decision, we consider several centroid-based similarity metrics:
 
@@
 \label{eq:kc}
-K^c_{ij} = exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_j - c_N \Vert} \right) 
+K^c_{ij} = \exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_j - c_N \Vert} \right) 
 \end{equation}
@@
 \\
-
-%\ml{Grégoire : if the matrix is not psd, then this is NOT a kernel, reformulate}
 

### 2016-07-12 paper/structureScene.tex
Edit abstract & keywords

@@
 
-This paper introduces a generic hierarchical modeling of acoustic scenes using the wavelet scattering features for time spans below the second and object-based segmentation for larger time spans that effectively model the scene.
+This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second timescales with object-based segmentation at larger timescales.
+Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates its potential, consistently outperforming baseline mel-frequency cepstral coefficients (MFCCs).
 
-Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets and state-of-the-art classifier demonstrates its potential.
-More precisely, features derived from the scattering transform consistently outperform Mel-frequency cepstral coefficients (MFCCs) for both tasks.
-
-Representing the scene at larger time scale as a set of examplars of coherent segments is an effective approach for similarity retrieval as it allows the retrieval system to focus on discriminative features.
-This benefit is reduced in supervised settings, probably because much of the discriminability potential of the approach is already learned by the classifier.
-
+Representing the scene at larger time scale as a set of examplars of coherent segments is an effective approach for scene similarity retrieval as it allows the retrieval system to focus on discriminative features.
+This benefit is reduced in the supervised scene classification, probably because much of the discriminability potential of the approach is already learned by the classifier.
+\ja{This last part is very vague and hard for me to understand.
+Why is this the case?}
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Acoustic scene classification, acoustic scene segmentation.
+acoustic scene classification, acoustic scene segmentation, scattering, wavelets, earth mover's distance
 \end{IEEEkeywords}

### 2016-07-12 paper/structureScene.tex
Edit intro

@@
 
-\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has considerably grown over the past decades.
-In order to measure the modification of animal biodiversity due to human activity or climate change, the research field of eco-acoustics \cite{ECOACOUSTICS2014, krause} is recently undertaking the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
-In addition, some recent works demonstrate the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
-Because they bear a strong societal impact and raise many scientific challenges, we believe that these case studies are of interest for the signal processing community.
+\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
+In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{ECOACOUSTICS2014, krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
+In addition, recent work demonstrates the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
+Because they bear a strong societal impact and raise many scientific challenges, we believe that these case studies are of considerable interest for the signal processing community.
 
 Unfortunately, the field of signal-based ecoacoustics is still in infancy.
-As a consequence, very few well built datasets are readily available for evaluation purposes.
-A closely related field of investigation, which is relatively more mature, is the classification of acoustic scenes whose aim is to predict labels by processing audio data.
-Despite being narrower in terms of range of applications, it has the advantage of being rooted by numerous works in cognitive psychology on categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and been considered in the data mining community for a while, with availability of a few well-designed datasets.
+\ja{We still haven't strictly defined what ``ecouacoustics'' is.
+Is there a simple, short definition we can use somewhere before this?}
+As a consequence, few well-built datasets are readily available for evaluation purposes.
+A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
+\cite{So what's the difference between ecoacoustics and acoustic scene classification?}
+While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of several well-designed datasets.
 
-Despite the differences, the following question is valid both for the aforementioned future application and the acoustic scene classification (ASC) task: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and with a large range of potential applications ?
+Despite the differences, ecoacoustics \ja{?} and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?
 
-The research described in this paper attempt to provide answers to this question by demonstrating the usefulness of two main contributions: 1) the use of the scattering transform allows us to extract discriminative features with few parameter optimization and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene allows to effectively focus the description of the scene on areas of interest.
+The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the scattering transform to extract discriminative features subject to a small number of parameters and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on specific areas of interest.
+\ja{Again, the last part is a little vague to me.}
 
 The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}.
-Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively.
-Experiments carried out in an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and in a supervised setting, \ie acoustic scene classification (ASC) are described in \ref{sec:experiments}.
+We describe the scattering transform in Section \ref{sec:scattering}, discuss post-processing transformations in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
+In Section \ref{sec:experiments} several experiments are proposed for an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and for a supervised setting, \ie acoustic scene classification (ASC).
 Results are subsequently reported in Section \ref{sec:results}.
@@
 \section{Feature design}
+\label{sec:design}
+
 Before supervised or unsupervised classification, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power. 

### 2016-07-12 paper/structureScene.tex
Edit motivations

@@
 
-As stated in the introduction, the field of ecoacoustics is relatively young, but significant works are worth mentioning.
-One avenue of research, originated in bioacoustics, intends to retrieve the species that produced a given sonic event of relatively short duration, in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
-A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording.
-Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify the species that produced a specific sonic event of short duration in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large} \ja{Is this the only reason?}.
+A more recent paradigm processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
+Automated systems belonging to this paradigm aim to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating} \ja{What does this mean?}.
 
-The same dichotomy applies to the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-In the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag}, which consists in modeling audio by high-level summary statistics computed from a set of local features.
-As it enjoys a high genericity, this approach is found in many audio classification tasks besides ASC.
-Yet, the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
-Indeed, the typical morphology of acoustic scenes is of a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}, \ie few impulsive sounds superimposed upon a stationary background.
-Therefore, it is arguably more beneficial to focus on the temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
+The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
+In the latter \ja{What about former?
+Why are we discussing both if we are going to focus on one?
+Shouldn't we just say that at some point?}, most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Acouturier et al.
+\cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
+Being a very generic method, this approach has been applied to many audio classification tasks besides ASC.
+Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated comparably to a mere averaging of the features \cite{lagrange:hal-01082501}.
+Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken_ear_2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
+Therefore, it is arguably more beneficial to focus on the large-scale temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
 
-Interestingly, this statement is validated by auditory psychology as well as sound synthesis based  on summary statistics \cite{mcdermott2013summary}.
-Indeed studies in cognitive psychology investigating urban sound environment quality pointed out that low level features such as (perceived or measured) global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003}  relies upon high level cognitive attributes such as the identification and assessment of the sound source which compose the scene.
-It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafay:hal-01111782}.
-As a consequence, research in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
+This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}.
+Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that low-level features such as (perceived or measured) global sound level \ja{Other things like spectral envelope too?}, are not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene.
+\ja{This argument is semantically a little tricky, since technically the low-level features already contain all this high-level information, it's just that the simple models we use to describe them fail to capture this high-level structure} It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafay:hal-01111782}.
+Consequently, research in sound perception is now focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
 
-Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
+Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes \cite{lafayPartI}.
 For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas.
+\ja{This is very similar to the last few sentences of the last paragraph.
+What are we trying to say differently here?}
 

### 2016-07-12 paper/structureScene.tex
Edit background section

@@
 Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
-
-Thus, studying the first DCASE challenge \cite{barchiesi2015acoustic} in 2013 is convenient as it concentrates much of the effort for this latter task.
-The table \ref{tab:dcase} summarizes the different algorithmic choices taken by the teams where most of scene predictions are achieved by majority voting of predictions over features.
+\ja{Since we go into so much detail on ASC, shouldn't we at lest describe what these works do?
+Some of this is already described in the motivations section, so perhaps we should move some over in here?}
+An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare differrent approaches to scene classification.
+Table \ref{tab:dcase} summarizes the results of the submitted algorithms where most \ja{?} of scene predictions are achieved by majority voting of predictions over features.
+\ja{What about the other algorithms?
+Why are we excluding them here?}
 
@@
 \begin{center}
-\caption{Designs submitted to the DCASE2013 ASC challenge. RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning.  \label{tab:dcase}}
 \begin{tabular}{lcl}
 & Acc. \%  &  \multicolumn{1}{c}{Design}  \\
-\cite{olivetti2013wonder} OE & 17 & compression distance, Random Forests \\
-\cite{elizalde2013vector} ELF & 55 & I vector, LDA \\
+\cite{olivetti2013wonder} OE & 17 & compression distance, random forests \\
+\cite{elizalde2013vector} ELF & 55 & i-vector, LDA \\
 \cite{krijnders2013tone} KH & 55 & cochleogram, SVMs \\
 \cite{patil2013multiresolution} PE & 58 & 2D Gabor filters, SVMs \\
-\cite{lee2013acoustic} NHL & 60 & UL of Mel spectra, SVMs \\
+\cite{lee2013acoustic} NHL & 60 & UL of mel spectra, SVMs \\
 \cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
 \cite{chum2013ieee} CHR & 65 & sparsity features, HMMs  \\
-\cite{geiger2013large} GSR & 69 & Mel and MFCCs, SVMs \\
+\cite{geiger2013large} GSR & 69 & mel \ja{?} and MFCCs, SVMs \\
 \cite{rakotomamonjy2015histogram} RG & 69 & histogram of gradient (HOG), SVMs \\
-\cite{li2013auditory} LTT & 72 & MFCCs, Tree baggers \\
+\cite{li2013auditory} LTT & 72 & MFCCs, tree baggers \\
 \cite{roma2013recurrence} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
@@
 \end{center}
+\caption{Designs submitted to the DCASE2013 ASC challenge. RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning. \label{tab:dcase}}
 \end{table}
@@
 In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
-Also, we resort to the use of SVMs with default parameters ($C=1$) for computing classification accuracy for the ASC task, as SVMs are widely used in the community and allows us to better root our work and focus our performance analysis on the earlier processing stages.
-Moreover, since the MFCCs are well established features and perform well in this task, they will be thoroughly used as baselines features in the experiments described in this paper.
+\ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked linear SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages.
+Moreover, since MFCCs are well-established features which perform well for these task, we shall include them baseline features in all experiments described in this paper.
 
@@
 
-For most of the experiments reported in this paper, the DCASE2013 datatset is considered \cite{giannoulis2013database, 7100934}.
+For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}.
 To build this dataset, three different recordists visited a wide variety of locations in Greater London over
-a period of months and in each
-scene recorded a few minutes of audio for each scene type. No
+a period of several months and in each
+scene recorded a few minutes of audio. No
 systematic variations in the recordings covaried with scene
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a 5 fold cross validation scheme.
+The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme.
 The folds used in this paper are the same as the ones used during the challenge.
+\ja{Why are we describing the dataset here in the background section?
+Wouldn't it be more appropriate under experiments?}
 
-As of today, there exist larger publicly available dataset, the
-Rouen dataset \cite{rakotomamonjy2015histogram} and the dCase2016 \cite{Mesaros2016_EUSIPCO} (only the public part).
+There are also two larger datasets: the
+Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge \cite{Mesaros2016_EUSIPCO}.
 The latter is also considered in this paper as a validation of the proposed approach.
+\ja{What does this last sentence mean?}
 
-As studied in \cite{lagrange:hal-01082501}, the DCASE2013 dataset has interesting intra class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices.
-Also, it is still challenging as state-of-the-art systems based on features design and support vector machines (SVMs) achieves 76 \% \cite{roma2013} (winner of the 2013 challenge RQA+SVM).
-For the sake of comparison the HOG+SVM approach \cite{rakotomamonjy2015histogram} achieves 75 \% and 92 \% on the dcase2013 and the Rouen datasets respectively.
-On the former, recent advances using hierarchical classifiers architectures reports 84 to 87 \% of accuracy depending on the prior knowledge considered for building the classifier architecture \cite{phan2016label}.
-On the latter, the state-of-the-art is currently 95 \% \cite{bisot2016acoustic}.
+The DCASE2013 dataset has significant intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+In addition, it is still challenging, with a state-of-the-art system based on feature design \ja{?} and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge).
+Another system, using a HOG+SVM approach \cite{rakotomamonjy2015histogram}, achieves $75\%$ and $92\%$ on the DCASE2013 and Rouen datasets, respectively.
+Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
+On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
 
-In this paper, we shall demonstrate that the use of features based on the wavelet scattering together with object-based representation of the acoustic scene allows us to achieve interesting performance both in the ASSR and ASC paradigms.
+In this paper, we shall demonstrate that the use of features based on the wavelet scattering transform together with object-based representation of the acoustic scene allows us to achieve state-of-the-art \ja{?} performance in both the ASSR and ASC paradigms.
 

### 2016-07-12 paper/structureScene.tex
Edit scattering section

@@
 
-Scattering transforms are translation-invariant representations of audio signals which cascade an auditory filterbank and a modulation filterbank, interspersed with complex modulus nonlinearities.
-Depending upon the chosen architecture, the modulation filterbank may either be performed solely over the time variable or on both time and frequency variables.
-In this section, we present both the temporal scattering transform and the time-frequency scattering transform.
+Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filterbanks interspersed with complex modulus nonlinearities.
+In this section, we present the standard temporal scattering transform.
 
 \subsection{Invariance and stability in acoustic scene classification}
-The notion of invariance to translation plays an essential role in auditory scene classification.
-Indeed, the starting time of the recordings are chosen arbitrarily, and thus do not convey any information about the class.
-To cancel this superfluous source of variability, signals must be mapped to a translation-invariant feature space before training the classifier.
-From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ of  a signal $\boldsymbol{x}(t)$ --
-a translation-invariant representation up to time shifts of duration $T$ can be simply obtained
+The notion of invariance to time shifts plays an essential role in auditory scene classification.
+Indeed, recordings may be shifted locally without affecting their perception and therefore such shifts do not convey any information about the class.
+To discard this superfluous source of variability, signals can be mapped to a time-shift invariant feature space before training the classifier, eliminating the need for the classifier to explicitly learn this invariance.
+From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ \ja{$x_1$, $\gamma_1$, etc have not been introduced yet.} of a signal $\boldsymbol{x}(t)$ --
+a representation invariant to time shifts shorter in duration than $T$ can be obtained
 by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
-set to $1/T$:
+set to $1/T$ \ja{How are we defining frequency here, as $\omega$ or $2\pi\omega$?
+In the former case, we should have $2\pi/T$.
+Probably we should define our Fourier transform somewhere to nail this down}:
 \begin{equation}
@@
 \end{equation}
-As a downside, the transient information in $\boldsymbol{x_1}$ at finer scales than $T$ are lost by this low-pass filtering, hence a lack of discriminability in feature space.
-To address this issue, the temporal scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$, and subsequently applying complex modulus.
+A downside is that the transient information in $\boldsymbol{x_1}$ at scales finer than $T$ are lost by this low-pass filtering, reducing discriminability in feature space.
+To address this issue, the temporal scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$ and then applying complex modulus.
 
-By resorting to wavelet transform modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to small time warps,
+By resorting to wavelet transform modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to time-warping deformation,
 in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}.
-Besides invariance to translation, this stability property is crucial in signal classification, as it entails a guarantee of robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.
+In addition time-shift invariance, this stability property is crucial to signal classification, since it guarantees robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.
 
-The scattering transform is defined mathematically as an infinite cascade of nonlinear layers, each consisting of a wavelet transform modulus operator.
-In practice, to achieve invariance to translation up to $T = 186\,\mathrm{ms}$, \ie of the order of the minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ suffice.
+The scattering transform is defined as an infinite cascade of wavelet transform and modulus operators.
+However, to achieve invariance to translation up to $T = 186\,\mathrm{ms}$, \ie the approximately minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ often suffice.
 
@@
 
+\ja{I'm not sure that this whole section makes sense right here.
+It can be a bit daunting to jump right in and talk about spectrograms $x_1$ and wavelelets, etc.
+before we've defined them.
+Could we put this after the next subsection, explaining all the nice properties of the scattering transform and why we need them after we've defined it fully?}
+
 \subsection{Temporal scattering}
 Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
-center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$.
+center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$. \ja{I mentioned this already, but it would be good to define the Fourier transform so that we can make these statements more specific}
 A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
-according to a geometric sequence of scales $2^{\gamma_1/Q_1}$.
-The variable $\gamma_1$, akin to a log-frequency, takes integer values between $0$ and $(J_1 \times Q_1 - 1)$.
-In the sequel, we set $\xi_1$ to 20 kHz, \ie close to the Nyquist frequency of audio recordings ; the number of octaves $J_1$ to $10$, \ie the maximum range of human hearing ; and the number of wavelets per octave $Q_1$ to $8$.
-We denote by
+according to a geometric sequence of scales $2^{\gamma_1/Q_1}$, obtaining
 \begin{equation}
-\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(2^{-\gamma_1/Q_1} t)
+\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(2^{-\gamma_1/Q_1} t)\mbox{.}
 \end{equation}
-the wavelets resulting from the dilation of $\boldsymbol{\psi}$(t).
+The variable $\gamma_1$ is an inverted log-frequency, taking integer values between $0$ and $(J_1 \times Q_1 - 1)$.
+In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$ (close to the Nyquist frequency of the audio recordings) the number of octaves $J_1$ to $10$ (the lower end of human hearing range) and the number of wavelets per octave $Q_1$ to $8$.
 For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
@@
 
-The wavelet transform $\boldsymbol{}$ of an audio signal
+The wavelet transform of an audio signal
 $\boldsymbol{x}(t)$ is obtained by convolution with all wavelets.
-Applying pointwise complex modulus to $\boldsymbol{y_1}$ yields
+Applying pointwise complex modulus the transform yields
 the wavelet scalogram
@@
 \boldsymbol{x_1}(t, \gamma_1)
-= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t).
+= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t)\mbox{.}
 \end{equation}
@@
 which is derived from the short-term Fourier transform (STFT) by averaging the frequency
-axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}\xi_1$.
+axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q_1}\xi_1$.
 Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
@@
 theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the temporal resolution of the CQT is fixed by the support of the STFT analyzing window.
+frequency range, whereas the temporal resolution of the CQT is fixed by the support of the STFT analyzing window. \ja{This is not true. See for example Sch\"orkhuber and Klapuri 2010.}
 Therefore, the wavelet scalogram has a better temporal localization at high
-frequencies than the CQT, at the expense of a greater computational cost.
+frequencies than the CQT, at the expense of a greater computational cost. \ja{Why is the cost greater?}
 This property allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
@@
 At a longer scale, they also account for some higher-level attributes of sound, \eg prosody in speech or rhythm in music.
-Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a translation-invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
+Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
 The amplitude modulation spectrum resulting from this operator is
@@
 where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order log-frequency index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
+Note that these second-order wavelets are dilated versions of a second mother wavelet $\boldsymbol{\psi}$, with a different center frequency $\xi_2$ and quality factor $Q_2$.
+The identity of the wavelet will be clear from context.
 In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to translation, yielding
+Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to time-shifting, which yields
 \begin{equation}
@@
 wavelets with one vanishing moment \cite{Venkitaraman2014}, as shown in Figure \ref{fig:gammatones}.
-In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the widespread Morlet wavelets.
-Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \cite{Fastl2007}.
+In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
+Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \ja{What does this mean?} \cite{Fastl2007}.
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
+\ja{Why is this relevant for our classification task?
+Sparsity?}
 As such, they can be discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
-
-The next section introduces some techniques in feature design to improve auditory scene classification with scattering coefficients.
 

### 2016-07-12 paper/structureScene.tex
Edit design section

@@
 \label{sec:design}
+\ja{It seems that what we are doing in this section is more post-processing than feature design, which would be the scattering transform itself.}
 
 Before supervised or unsupervised classification, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power.
-In this section, we review several ways of achieving these properties, namely binaural data augmentation, feature selection, logarithmic compression, standardization, and temporal integration.
-The former has, to the best of our knowledge, not being considered for ASC and is specific to multichannel recordings, whereas the next ones would also apply to many other kinds of data.
+In this section, we review several ways of achieving these properties, namely feature selection, logarithmic compression, standardization, and temporal integration.
 
@@
 This fact may cause the classifier to overfit the training set, especially when the number of features is large and the number of training instances is small, as is the case here.
-Selecting features appropriately before the training stage may circumvent this problem.
+Appropriately selecting features before the training stage may circumvent this problem.
 However, the parameters of the chosen feature selection algorithm are themselves prone to overfitting.
 Instead, we adopt the more conservative approach of retaining all coefficients within a certain range of subband indices $\gamma_1$ and $\gamma_2$.
+\ja{These indices are already used to determine the first- and second-order log-frequency indices, respectively.
+Perhaps $\gamma_\mathrm{min}$ and $\gamma_\mathrm{max}$ would be better?}
 In comparison with adaptive algorithms, the subset of selected features $\gamma$ has a suboptimal statistical relevance, but it is easily interpretable in the realm of audio signal processing, as it is equivalent to a coarse band-pass filtering of $\boldsymbol{x}(t)$ and $\boldsymbol{x_1}(t,\gamma_1)$ over the time dimension $t$.
+\ja{Are we filering $x_1$ too?
+The selection is only in $\gamma_1$, correct?}
 
@@
 In the topmost frequencies, this amount is so faint that the corresponding coefficients $\boldsymbol{x_1}(t,\gamma_1)$ and $\boldsymbol{x_2}(t,\gamma_1,\gamma_2)$ no longer bear discriminative information for the classification task at hand.
+\ja{The problem is that we will magnify these using a logarithm in the next section, so this argument is not as strong as that.
+Is there some other reason we would like to discard the high frequencies?}
 
-In the sequel, we evaluate the impact of discarding the scattering coefficients above a certain acoustic frequency $\xi_{1,\max}$, which corresponds to discarding subbands $\gamma_1$ below a certain scale
+In the sequel, we evaluate the impact of discarding the scattering coefficients above a certain acoustic frequency $\xi_{1,\max}$, which corresponds to discarding subbands $\gamma_1$ below a certain scale \ja{It can be a little confusing for the reader to switch between log-frequency and scale, especially since we use both words to refer to $\gamma_1$ in some sense.
+Technically, there is a lower bound on the scale, which turns into a lower bound of $\gamma_1$, but it's not clear from first glance.}
 \begin{equation}
@@
 Reducing the frequency range is an effective way to enhance invariance, widely used in the literature.
-For instance, \cite{roma2013} achieved the best performance in the 2013 edition of the DCASE challenge, rely on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max}  = 900\,\mathrm{Hz}$ instead of the more conventional value of $16\,\mathrm{kHz}$.
+For instance, \cite{roma2013} achieved the best performance in the DCASE2013 challenge and relied on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max} = 900\,\mathrm{Hz}$ instead of including all frequencies up to the Nyquist rate.
 
@@
 
-Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
-Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design.
-It appears empirically that their distribution is skewed towards the right, which means that the tail towards greater values is longer than the tail towards lower values.
-However, skewness can be reduced by applying to all coefficients a pointwise concave transformation, \eg logarithmic.
-Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
+Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances \ja{Do we have a reference for this?}.
+Yet the distribution of the scattering coefficients is skewed towards larger values. We can reduce this skewness by applying a pointwise concave transformation, like a logarithm, to all coefficients.
+Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE2013 dataset, before and after logarithmic compression.
 
@@
 
-The application of the pointwise logarithm to magnitude spectra is ubiquitous in audio signal processing.
-Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
-We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the binaural microphone, which is a spurious factor of variability to the task of auditory scene classification.
-Logarithmic compression can linearize this dependency, which arguably facilitates the construction of a powerful invariant at the classifier stage.
+Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
+Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. \ja{Cite?}
+We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone, a source of spurious variability in scene classification.
+Logarithmic compression can linearize this dependency, facilitating the construction of a powerful invariant at the classifier stage.
 
-Given a task of musical genre recognition, \cite{Anden2014} has advocated for the renormalization of second-order scattering coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ by the corresponding first-order scattering coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, as it provably decorrelates the amplitudes of their activations.
-Interestingly, taking the logarithm of renormalized coefficients would yield
+For the task of musical genre recognition, second-order scattering coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ are sometimes normalized by the corresponding first-order scattering coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, since this decorrelates them from one another \cite{Anden2014}.
+We note that taking the logarithm of such renormalized coefficients yields
 \begin{equation}
@@
 \ie a linear combination of the logarithms of first- and second-order coefficients.
-Therefore, the theoretical insight brought by \cite{Anden2014} in favor of renormalized scattering also applies to log-scattering up to a linear transformation in feature space, to which affine classifiers are not sensitive.
+As such, a non-linear renormalization becomes a linear transformation, which can be learned by a linearly discriminative classifier.
 
@@
 possible by integrating signal information over a long temporal context.
-Indeed, whereas a few seconds are often sufficient to recognize a speaker,
-a musical instrument, or a genre, it may require up to 30 seconds
-to disambiguate two classes auditory scenes which share part of their semantic content, \eg a train from a subway station or a quiet street from a park.
+Indeed, whereas a few seconds is often sufficient to recognize a speaker,
+a musical instrument, or a genre, up to $30$ seconds may be required
+to disambiguate two classes of auditory scenes with overlapping semantic content, \eg a train from a subway station or a quiet street from a park.
 Depending on whether aggregation is performed in feature space or in decision space, the corresponding method is referred to as early or late integration.
@@
 A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene, by retaining their average values only.
-Going back to the definition of the scattering transform given in section \ref{sec:scattering}, this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
+Going back to the definition of the scattering transform given in Section \ref{sec:scattering}, this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
 With a slight abuse of notation, we denote by
@@
 
-Conversely, a late integration scheme relies on probabilistic estimates over short-term windows of length $T$, which are subsequently aggregated to produce a final decision
+Conversely, a late integration scheme relies on probabilistic assignments $\mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right]$ over short-term windows of length $T$, which are subsequently aggregated to produce a final decision
 \begin{equation}
-\hat{y} = \arg \max_{y} \rho\Big(\big\{ \mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right] \big\}_{t} \Big),
+\hat{y} = \arg \max_{y} \rho\Big(\big\{ \mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right] \big\}_{t} \Big)\mbox{,}
 \end{equation}
@@
 
-The principal drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene -- if data augmentation is not considered.
+A drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene.
 In the context of the DCASE 2013 dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
-On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene, resulting in $10240$ instances overall.
+\ja{That may be, but by averaging, we also reduce the variability of our data, which may be a good thing.
+If we're dealing with Gaussian processes, this is in fact a very good thing to do.}
+On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene \ja{How long is each recording?
+Should we specify this somewhere?}, resulting in $10240$ instances overall.
 However, many of these instances may be silent or lack any salient properties of the class they are assigned to, hence an increase in bias and a risk of underfitting.
@@
 We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
+\ja{Another important difference is that early integration makes the classifier objective more appropriate, since the SVM will maximize the final classification error on the training data, as opposed to some intermediate frame-level label.}
 
@@
 Let $\mathbf{S}\boldsymbol{X}(\gamma,n)$ be a dataset, where the indices $\gamma$ and $n$ respectively denotes features and examples.
-It is commonly acknowledged that support vector machines should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges.
-This operation may also help making the dataset linearly separable in feature space.
+It is commonly acknowledged that SVMs should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges.
+\ja{Again, it would be really good if we could cite something for this.}
+This operation may also help making the dataset linearly separable in feature space. \ja{Why?}
 To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we apply the affine transformation
@@
 \dfrac{ \mathbf{S}\boldsymbol{X}(\gamma, n) -
-\mathbb{E}[ \mathbf{S}\boldsymbol{X}](\gamma)}{\sigma[ \mathbf{S}\boldsymbol{X}](\gamma)}
+\mu[ \mathbf{S}\boldsymbol{X}](\gamma)}{\sigma[ \mathbf{S}\boldsymbol{X}](\gamma)}
 \end{equation}
-where $\mathbb{E}[ \mathbf{S}\boldsymbol{X}](\gamma) = \frac{1}{N} \sum_{n=1}^{N} \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the expected value of $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and
+where $\mu[ \mathbf{S}\boldsymbol{X}](\gamma) = \frac{1}{N} \sum_{n=1}^{N} \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the sample mean of $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and
 \begin{equation}
@@
 \sqrt{\frac{1}{N-1} \sum_{n=1}^{N}
-\left( \mathbf{S}\boldsymbol{X}(\gamma,n) - \mathbb{E}[\mathbf{S}\boldsymbol{X}] \right)^2}
+\left( \mathbf{S}\boldsymbol{X}(\gamma,n) - \mu[\mathbf{S}\boldsymbol{X}] \right)^2}
 \end{equation}
@@
 
- The vectors $\mathbb{E}[\mathbf{S}\boldsymbol{X}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{X}](\gamma)$ are estimated from the training set only, and the same affine transformation is subsequently applied to all samples in the training set and the test set.
-
-The next section discusses adaptive methods for the temporal integration of scattering coefficients in an unsupervised setting.
+The vectors $\mu[\mathbf{S}\boldsymbol{X}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{X}](\gamma)$ are estimated from the training set only, and the same affine transformation is subsequently applied to all samples in the training set and the test set.
 

### 2016-07-12 paper/structureScene.tex
It's \ja, not \cite

@@
 A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
-\cite{So what's the difference between ecoacoustics and acoustic scene classification?}
+\ja{So what's the difference between ecoacoustics and acoustic scene classification?}
 While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of several well-designed datasets.

### 2016-07-12 paper/structureScene.tex
Edit objects section

@@
 \newcommand*{\vs}{vs.\@\xspace}
+
+\newcommand*{\EMD}{\mathrm{EMD}}
 
@@
 
-As discussed in Section \ref{sec:motivations}, several outcomes in sound perception encourage us to evaluate the validity of an object-based representation of the scene for predicting high-level properties.
-Though, a perfect knowledge of which event occurred at a given time is not feasible.
-In this paper, we thus focus on a simple quantization scheme in order to favor genericity, implemented using a clustering approach where coherent regions of the scene are identified.
+As discussed in Section \ref{sec:motivations}, results in sound perception suggest the appropriateness of an object-based representation of auditory scenes for predicting high-level properties.
+That being said, perfect knowledge of which event occurred at a given time is not feasible.
+\ja{What does this mean?} In this paper, we therefore focus on a simple quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
 
-One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, which performs some multivariate estimates over the whole dataset and not solely on each scene, and re-projects each scene in the resulting enhanced feature space.
-Here, with the aim of balancing better the influence of salient events and texture-like sounds on the final decision, we consider several centroid-based similarity metrics:
+\ja{We need to describe the clustering here and what it gives as output. Otherwise it becomes difficult to describe the way we compare scenes below.}
+
+One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, which clusters not over a single scene, but over the whole dataset, and re-projects each scene in the resulting enhanced feature space.
+\ja{This sentence is also a big unclear.}
+Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, we consider several centroid-based similarity metrics:
 
 \begin{itemize}
-\item \emph{ob-closest} (ob-c): the similarity between two scenes is equal to the largest similarity between their centroids.
-\item \emph{ob-averaged} (ob-a): the similarity between two scenes is equal to the average of their centroids similarities.
-\item \emph{ob-weighted} (ob-w): for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
+\item \emph{ob-closest} (ob-c): the similarity between two scenes is equal to the largest similarity between their centroids,
+\item \emph{ob-averaged} (ob-a): the similarity between two scenes is equal to the average of their centroids similarities, and
+\item \emph{ob-weighted} (ob-w): for each scene, each centroid is weighted according to the number of frames belonging to its cluster. \ja{And then we compute the EMD?}
 \end{itemize}
@@
 K^c_{ij} = \exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_j - c_N \Vert} \right) 
-\end{equation}
+\end{equation} \ja{To simplify the analysis later, can we already specify that we're looking at a specific pair $(u, v)$ and that the kernel $K^c$ is defined only for $i$ in $u$ and $j$ in $v$?}
 where $c_N$ is the $N^{\textrm{th}}$ nearest neighbor to the centroid $c_i$, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+\ja{How does the $c_N$ notation distinguish between $c_i$ and $c_j$?
+Do we have $c_{i,N}$ and $c_{j,N}$?}
 
@@
 Each scene $s_u$ is described by a signature $S_u$ of $m$ clusters ($S_u=\lbrace(c_1^u,w_1^u),\ldots,(c_m^u,w_m^u)\rbrace$), with $c_i^u$ and $w_i^u$ being the centroid and the weight of the $i$th cluster of the scene $s_u$ respectively.
+\ja{Everything that follows only refers to the ob-w metric, correct?
+If so, we should specify it here, and also say how the similarity is explicitly calculated for the other options}
 
-To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ and introduced in \cite{pele2008linear}.
-$EMD$ is a cross-bin histogram distance, widely used in computer vision \cite{zhang2007local}, which compute the minimal cost to be paid to transform one histogram into the other.
-The advantage of using the $EMD$ is that it does not require bins to be aligned.
-The alignment problem is solved by using a ``\,ground distance\,'', which is the distance between the bins representatives of the two histograms.
-In our case, the histograms are the cluster weights $w$, and the bins representative the cluster centroids $c$.
+To compute the distance between two signatures $S_u$ and $S_v$, we use a variant of the earth mover's distance ($\EMD$), known as $\widehat{\EMD}$ and introduced in \cite{pele2008linear}.
+The $\EMD$ distance is a cross-bin histogram distance, widely used in computer vision \cite{zhang2007local}, which computes the minimal cost to be paid to transform one histogram into the other.
+An advantage of $\EMD$ is that automatically aligns the histogram bins by using a ``ground distance,'' which is the distance between the bin centers in the two histograms.
+In our case, the histograms are the cluster weights $w$ defined over the bins formed by cluster centroids $c$.
 
-$\widehat{EMD}$ is a variant version of the $EMD$ adapted for non-normalized histograms.
-To compute the $\widehat{EMD}$, we use the implementation proposed in \cite{pele2009fast}.
+The $\widehat{\EMD}$ variant of $\EMD$ is adapted for non-normalized histograms.
+To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
 Given two signatures $S_u$ and $S_v$, the $\widehat{EMD}$ is computed by solving the following linear programming problem:
@@
 \begin{split}
-\widehat{EMD}(S_u,S_v) &=( \min\limits_{\lbrace f_{ij}\rbrace} \sum\limits_{i,j} f_{ij}D_{ij} )  \\ 
-&+ \mid \sum\limits_{i} w_i^u - \sum\limits_{j} w_j^v  \mid \alpha \max\limits_{i,j}\lbrace  D_{ij}\rbrace
+\widehat{\EMD}(S_u,S_v) &=\left( \min\limits_{\lbrace f_{ij}\rbrace} \sum\limits_{i,j} f_{ij}D_{ij} \right) \\
+&+ \left|\sum\limits_{i} w_i^u - \sum\limits_{j} w_j^v  \right| \alpha \max\limits_{i,j}\lbrace  D_{ij}\rbrace
 \end{split}
@@
 \begin{equation*}
-s.t. \quad f_{ij}\geq0 \quad \sum\limits_{j} f_{ij} \leq w_i^u \quad \sum\limits_{i} f_{ij} \leq w_j^v 
+\mathrm{s.t.} \quad f_{ij}\geq0 \quad \sum\limits_{j} f_{ij} \leq w_i^u \quad \sum\limits_{i} f_{ij} \leq w_j^v
 \end{equation*}
@@
 \begin{equation*}
-\sum\limits_{i,j}f_{ij} = \min( \sum\limits_{i} w_i^u ,\sum\limits_{j} w_j^v )
-\end{equation*}
-
-where $\lbrace f_{ij} \rbrace$ is the optimal flow between the cluster weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to ``\,supply to the demand\,'' of the $j$th bin.
-$D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, and is computed from the kernel $K^c$:
-
+\sum\limits_{i,j}f_{ij} = \min\left( \sum\limits_{i} w_i^u ,\sum\limits_{j} w_j^v \right)
+\end{equation*} \ja{Can we provide some interpretation for the different terms and constraints here?}
+where $\lbrace f_{ij} \rbrace$ is the flow between the cluster weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to ``supply the demand'' of the $j$th bin.
+We denote by $D_{ij}$ the ground distance between the centroids $c_i^u$ and $c_j^v$ computed from the kernel $K^c$:
 \begin{equation*}
@@
 \end{equation*}
-
 with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ of the scenes $s_u$ and $s_v$ respectively.
-As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.
+As suggested in \cite{pele2009fast}, we set the tradeoff parameter $\alpha$ to $1$.
 
 To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support} is computed:
-
 \begin{equation}
 \label{eq:ks}
-K_{uv}^s = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
+K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(S_u,S_v)}{A} \right) \\
 \end{equation}
-
-with $A$ a scaling parameter, set to the mean value of the $\widehat{EMD}$ between all the scenes.
-The resulting kernel $K^s$ is called an EMD kernel, and it has to be noted that there is no proof that such kernel is positive definite.
-\\
+with $A$ a scaling parameter, set to the mean value of the $\widehat{\EMD}$ between all the scenes.
+The resulting kernel $K^s$ is known as an EMD kernel, and it should be noted that there is no guarantee that this kernel is positive definite.
 

### 2016-07-12 paper/structureScene.tex
Edit experiments section

@@
 
-Experiments are carried out using temporal-scattering coefficients as well as Mel Frequency Cesptral Coefficients (MFCCs) as features.
-For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients, with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=188\,\mathrm{ms}$.
+Experiments are carried out using temporal scattering coefficients as well as mel-frequency cepstral coefficients (MFCCs) as features.
+For the temporal scattering, each scene is described by $128$ vectors of scattering coefficients computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=188\,\mathrm{ms}$.
 Experiments are conducted with and without the logarithmic compression (see Section \ref{sec:logcomp}).
@@
 Different sets of parameters were tested to compute the MFCCs.
-Results here are only reported for the best setting.
-40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a full frequency range.
-This setting gives us 600 feature vectors per scenes.
-It has been reported that classifying the full set of frames directly may not be the best approach \cite{rakotomamonjy2015histogram}.
-In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is divided into $n$ non-overlapping $t$-seconds long "texture" windows.
-Each $n$ texture windows are then averaged over time.
-As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors prior to the clustering step, for both the temporal-scattering and the MFCCs.
-Thus, $t$ is set to 250ms, resulting in 120 MFCCs texture windows per scene.
+Results here are only reported for the best setting, which consisted of $40$ MFCCs (including the first energy coefficient) computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$, with full frequency range.
+This setting gives us $600$ feature vectors per scene.
+It has been reported that classifying the each individual frame directly may not be the best approach \cite{rakotomamonjy2015histogram}.
+\ja{Why?} In order to reduce the amount of data, a pooling step is performed on the MFCCs: each recording is divided into $n$ non-overlapping $t$-seconds long ``texture'' windows.
+The MFCC vectors within each texture window are then averaged in time.
+As the \emph{ob} approaches involve a clustering step, we need to ensure that both MFCCs and scattering coefficients have a comparable number of vectors per recording.
+\ja{Why do we need to do this?} Thus, $t$ is set to $250\,\mathrm{ms}$, resulting in $120$ MFCCs texture windows per scene.
+
+\ja{We've tested the impact of the logarithm here, but not of the feature selection or standardization.
+Should we do this as well?
+For the feature selection, we did say earlier that we would test its impact.}
 
@@
 
-The influence of the temporal-scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (\emph{i.e.} ASSR) and in a supervised setting (\emph{i.e.} ASC).
+The influence of the temporal scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (ASSR) and in a supervised setting (ASC).
 
-For the ASSR setting, evaluation is performed on the private dataset of the DCASE 2013 challenge.
-The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items.
-The $p@k$ is computed for $k=[1,\ldots,9]$, $9$ being the number of elements that are of the same class than the seed item.
-Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.
+For ASSR, evaluation is performed on the private part of the DCASE2013 dataset.
+The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averagin over all query items.
+The $p@k$ is computed for $k=\{1,\ldots,9\}$, since each class only has $10$ items \ja{?}.
+Note that a $p@1$ is equivalent to the classification accuracy obtained by choosing the label of the closest neighbor for a given item.
 For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameter $N$ of the RBF kernels (see Eq.
 \ref{eq:kc}) leading to the best $p@9$.
+\ja{I'm not sure what this means.
+Do we fix $m$ and $N$ here for all experiments or do we optimize them over the test data each time?
+If so, this will bias the results, no?}
 
-For the ASC setting, the systems are evaluated  with five-fold  stratified cross validation on the private dataset of the DCASE 2013 challenge.
+For ASC, the systems are again evaluated on the private part of DCASE2013, but here with five-fold stratified cross validation.
 The folds are identical to those used in the original challenge.
-Optimal parameters (see above) are this time learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.
+Optimal parameters (see above) are this time learned on the public part of the DCASE2013 dataset, also using five-fold stratified cross validation.
 Clustering is done using an enhanced version of the popular K-means algorithm known as $K$-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
+\ja{How are they selected differently?}
 
-The \emph{ob} approaches are compared to commonly used approaches using an early integration strategy (\emph{early}) for the ASSR setting, and both an early and a late integration (\emph{late}) strategy for the ASC setting (see Section \ref{sec:eili}).
-In thsi context early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.
+The \emph{ob} approaches are compared to commonly used early integration approach (\emph{early}) for ASSR, while both early and late integration (\emph{late}) strategy are evaluated for the ASC setting (see Section \ref{sec:eili}).
+In this context early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.
+\ja{We should perhaps mention this in that particular section instead of over here.}
 
-As most of the 2013 DCASE challenge participants did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting.
-The parameter $C$ is here set to one.
+In the same way as the majority of the DCASE2013 participants, we use an SVM for the ASC task, which is a large-margin binary discriminator with tunable regularization of misclassified examples.
+The parameter $C$ is here set to $1$.
 
@@
 For both the \emph{early} and \emph{late} approaches, a linear kernel as well as a RBF kernel are used.
-As for the \emph{ob} approaches, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}.
-All feature vectors, being frames, texture windows or centroids are standardized prior to be given to the classifier.
-The standardization is computed with respect to the train/test set splits of the cross validation scheme (see Section \ref{sec:stand}).
+As in the RBF kernels used in the \emph{ob} approaches, the RBF kernel in the SVM is scaled locally using the method proposed in \cite{selfTuneManor2004}.
+All feature vectors, be they frames, texture windows or centroids, are standardized prior to be given to the classifier.
+The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
 

### 2016-07-12 paper/structureScene.tex
Edit results section

@@
 \includegraphics[width=8cm]{gfx/unsupervised_test3.eps}
-\caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression}
+\caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression.}
 \label{fig:ASS_1}
@@
 \includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
-\caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression (w/o log)}
+\caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression.}
 \label{fig:ASS_2}
@@
 
-This section present evaluation results for the ASSR settings. The $p@k$ for different settings are showed in the Figures~\ref{fig:ASS_1} and \ref{fig:ASS_2}.
+This section presents evaluation results for the ASSR settings.
+The $p@k$ for different settings are shown in Figures \ref{fig:ASS_1} and \ref{fig:ASS_2}, illustrating the effect of the different similarity metrics and the influence of the logarithm, respectively.
 
-\subsection*{MFCC \vs scattering transform}
+\subsubsection*{MFCC \vs scattering transform}
 
@@
 Overall, log-compressed scattering coefficients systematically outperform MFCCs.
+This is to be expected since the scattering coefficients capture larger-scale modulations as opposed to the MFCCs which only describe the short-time spectral envelope.
 
-\subsection*{Object-based \vs early integration}
+\subsubsection*{Object-based \vs early integration}
 
-For the scattering, \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
-However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{early}, which shows that the discriminant information is not carried by all the centroids.
-To take advantage of an object-based representation, a selection step has to be run on the representatives of the scenes.
-Furthermore It appears that \emph{ob-c} is be able to better characterize the classes than \emph{ob-w}.
-This last observation suggests that weighted a centroid according to the number of frames belonging to it may prove to be a limited solution.
+For the scattering transform, \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
+However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{early}, which shows that the discriminant information is destroyed by averaging the contributions from all centroids.
+To take advantage of an object-based representation, we need to select certain representative centroids when comparing objects.
+Furthermore, it appears that \emph{ob-c} is be able to better characterize the classes than \emph{ob-w}.
+This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
 Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
-On the contrary, two similar environments may shared a lot of similar sound sources.
+On the contrary, two similar environments may shared a lot of similar sound sources with only a few sources discriminating between them.
 
-For the MFCCs, there is no clear evidence that the \emph{ob} approaches improve the results.
+For the MFCCs, there is no clear evidence that the \emph{ob} approaches improve the results. \ja{Why?}
 
-\subsection*{Role of logarithmic compression}
+\subsubsection*{Role of logarithmic compression}
 
-Outcomes for the scattering with and without logarithmic compression are showed in Figure~\ref{fig:ASS_2}.
-It can be seen that the logarithmic compression strongly improves the results, specially for the \emph{ob} approaches, which perform lesser than \emph{early} without the logarithmic compression.
+Outcomes for the scattering with and without logarithmic compression are shown in Figure \ref{fig:ASS_2}.
+It can be seen that the logarithmic compression strongly improves the results, especially for the \emph{ob} approaches, which perform worse than \emph{early} without the logarithmic compression.
 \gl{une idée du pourquoi ?}
@@
 \includegraphics[width=\columnwidth]{gfx/supervised_rbf_test3.eps}
-\caption{...}
+\caption{... \ja{What is this?}}
 \end{center}

### 2016-07-12 paper/structureScene.tex
new abstract

--- 2016-07-12 paper/structureScene.tex
@@
 This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second timescales with object-based segmentation at larger timescales.
-Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates its potential, consistently outperforming baseline mel-frequency cepstral coefficients (MFCCs).
+Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates the potential of featurs designed using the wavelet scattering transform, consistently outperforming baseline mel-frequency cepstral coefficients (MFCCs).
 
-Representing the scene at larger time scale as a set of examplars of coherent segments is an effective approach for scene similarity retrieval as it allows the retrieval system to focus on discriminative features.
-This benefit is reduced in the supervised scene classification, probably because much of the discriminability potential of the approach is already learned by the classifier.
-\ja{This last part is very vague and hard for me to understand.
-Why is this the case?}
+Also, representing the scene at larger time scale as a set of representative features of coherent segments is an effective approach for scene similarity retrieval as it allows the retrieval system to focus on discriminative features.
 \end{abstract}

### 2016-07-12 paper/structureScene.tex
Edit conclusion

--- 2016-07-12 paper/structureScene.tex
@@
 
-In this paper, the benefit of considering 1) the scattering transform for building discriminant features and 2) unsupervised clustering for the modeling of large time span environmental acoustic scenes.
-This leads to improvement both for the ASSR and ASC tasks.
+In this paper we consider the benefit of 1) the scattering transform for building discriminant features and 2) unsupervised clustering for the modeling of large-scale environmental acoustic scenes.
+This leads to improvements in both the ASSR and ASC tasks.
 
-Abstracting the scene as a reduced set of exemplars with adapted similarity strategies allows us to improve ASSR performance when the BOF approaches based on GMMs fails with reference to an early integration scheme.
-Those results advocate for a hierarchical model of the scene, where small time scale structure -- \ie below a second -- shall be decorrelated using ca onvolutional network such as in the scattering transform, whereas larger time scales -- \ie over a second -- structure shall be modeled as adaptive aggregates that can be identified in the time domain in an unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.
+Representign a scene using a reduced set of exemplars, in the form of centroids, with adapted similarity strategies allows us to improve ASSR performance compared to the insufficient BOF approaches based on GMMs.
+These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- are represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.
+\ja{I'm not sure what this last part means.}
 
-This assertion shall be studied further in future work by considering more complex convolutional networks, larger databases and more diverse application scenarios, such as pleasantness prediction \cite{lafaySoundscapePilot, lafay:hal-01111782} or ecoacoustics tasks.
+This assertion will be studied further in future work by considering more complex convolutional networks, larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot, lafay:hal-01111782} and other ecoacoustics tasks.
 

### 2016-07-12 paper/structureScene.tex
revise up to motivations

--- 2016-07-12 paper/structureScene.tex
@@
 In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{ECOACOUSTICS2014, krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
+
 In addition, recent work demonstrates the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
@@
 
-Unfortunately, the field of signal-based ecoacoustics is still in infancy.
-\ja{We still haven't strictly defined what ``ecouacoustics'' is.
-Is there a simple, short definition we can use somewhere before this?}
+Formerly put, ecoacoustics is an interdisciplinary science that investigates natural and anthropogenic sounds and their relationship with the environment over a wide range of study scales at the intersection between ecology, acoustics and computer science.
+Despite identified needs, the field of signal-based ecoacoustics is still in infancy.
 As a consequence, few well-built datasets are readily available for evaluation purposes.
 A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
-\ja{So what's the difference between ecoacoustics and acoustic scene classification?}
+
 While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of several well-designed datasets.
 
-Despite the differences, ecoacoustics \ja{?} and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?
+Despite the differences, ecoacoustics and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?
 
-The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the scattering transform to extract discriminative features subject to a small number of parameters and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on specific areas of interest.
+The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the wavelet scattering transform to extract discriminative features subject to a small number of parameters and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on specific areas of interest.
 \ja{Again, the last part is a little vague to me.}
@@
 
-One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify the species that produced a specific sonic event of short duration in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large} \ja{Is this the only reason?}.
-A more recent paradigm processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
-Automated systems belonging to this paradigm aim to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating} \ja{What does this mean?}.
+%One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify species properties of vocalizations, for example to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
 
-The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-In the latter \ja{What about former?
-Why are we discussing both if we are going to focus on one?
-Shouldn't we just say that at some point?}, most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Acouturier et al.
+A recent paradigm processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
+Automated systems belonging to this paradigm aim to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time regions of particular interest requesting detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+
+%The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
+In the latter, most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Acouturier et al.
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
@@
 Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated comparably to a mere averaging of the features \cite{lagrange:hal-01082501}.
-Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken_ear_2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
+Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
 Therefore, it is arguably more beneficial to focus on the large-scale temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
@@
 This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}.
-Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that low-level features such as (perceived or measured) global sound level \ja{Other things like spectral envelope too?}, are not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that (perceived or measured) global sound level, are not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
 Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene.
-\ja{This argument is semantically a little tricky, since technically the low-level features already contain all this high-level information, it's just that the simple models we use to describe them fail to capture this high-level structure} It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafay:hal-01111782}.
-Consequently, research in sound perception is now focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
-
-Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes \cite{lafayPartI}.
-For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas.
-\ja{This is very similar to the last few sentences of the last paragraph.
-What are we trying to say differently here?}
+\ja{This argument is semantically a little tricky, since technically the low-level features already contain all this high-level information, it's just that the simple models we use to describe them fail to capture this high-level structure} It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafayPartI}.
+Consequently, research in sound perception is now strongly focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
+Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes.
 

### 2016-07-12 paper/structureScene.tex
renew Fig 5 + caption

@@
 
+\subsubsection*{MFCCs \vs scattering}
+As shown in Figure, log-scattering coefficients consistently outperform $MFCCs$ over all folds of the DCASE 2013 private dataset.
+The average accuracy
+
+\subsubsection*{Early \vs late integration}
+
+\subsubsection*{Role of logarithmic compression}
+
+\subsubsection*{Participation to the DCASE 2016 challenge}
+In order to evaluate how general these results are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
+
 \begin{figure}
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/supervised_rbf_test3.eps}
-\caption{... \ja{What is this?}}
+\includegraphics[width=\columnwidth]{folds.png}
+\caption{Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies. 
+The chosen classifier is a support vector machine (SVM) with a Gaussian radial kernel.
+See text for details.}
 \end{center}
+\ref{fig:folds}
 \end{figure}
@@
 
-MFCC \vs scattering
-
-Object-based \vs early/late integration
-
-Role of logarithmic compression
-
-\vl{Results: DCASE 2016}
-
 \section{Conclusion}
@@
 
-Representign a scene using a reduced set of exemplars, in the form of centroids, with adapted similarity strategies allows us to improve ASSR performance compared to the insufficient BOF approaches based on GMMs.
+Representing a scene using a reduced set of exemplars, in the form of centroids, with adapted similarity strategies allows us to improve ASSR performance compared to the insufficient BOF approaches based on GMMs.
 These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- are represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.

### 2016-07-12 paper/structureScene.tex
compare MFCC with scattering in ASC, cite Salamon 2015

@@
 \subsubsection*{MFCCs \vs scattering}
-As shown in Figure, log-scattering coefficients consistently outperform $MFCCs$ over all folds of the DCASE 2013 private dataset.
-The average accuracy
+As shown in Figure, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
+Not only the average accuracy is improved, but the variance between fold-wise accuracies is also reduced, \eg from $16\%$ to $7\%$ in the case of early integration.
+This is in accordance with the previous results on similarity retrieval, as well as with other papers comparing mel-frequency coefficients with scattering coefficients on a variety of tasks, ranging from musical genre recognition \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.
 
@@
 \subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general these results are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
+In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
 

### 2016-07-12 paper/structureScene.tex
finer stroke for Fig 5

@@
 \subsubsection*{MFCCs \vs scattering}
-As shown in Figure, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
+As shown in Figure \cite{folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
 Not only the average accuracy is improved, but the variance between fold-wise accuracies is also reduced, \eg from $16\%$ to $7\%$ in the case of early integration.
@@
 \caption{Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies. 
-The chosen classifier is a support vector machine (SVM) with a Gaussian radial kernel.
+The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
+The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.
 See text for details.}
@@
 \end{center}
+\ref{table:dcase2013}
 \end{table}

### 2016-07-12 paper/structureScene.tex
revise conclusion and reauest to vincent

@@
 The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.
-See text for details.}
+See text for details.\ml{je mettrai des lignes fines entre les points 1 et 3, et 4 et 6, je zoomerai un peu et l'axe des y n'est pas numerote ni labelise }}
 \end{center}
@@
 In this paper we consider the benefit of 1) the scattering transform for building discriminant features and 2) unsupervised clustering for the modeling of large-scale environmental acoustic scenes.
-This leads to improvements in both the ASSR and ASC tasks.
+%This leads to improvements in both the ASSR and ASC tasks.
 
-Representing a scene using a reduced set of exemplars, in the form of centroids, with adapted similarity strategies allows us to improve ASSR performance compared to the insufficient BOF approaches based on GMMs.
-These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- are represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.
+Representing a scene using a reduced set of representatives, with adapted similarity strategies allows us to improve ASSR performance compared to the BOF approaches based on GMMs.
+These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- are represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates.
+
+\ml{Vincent: un par pour resumer la partie classif}
+%, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.
 \ja{I'm not sure what this last part means.}
 
-This assertion will be studied further in future work by considering more complex convolutional networks, larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot, lafay:hal-01111782} and other ecoacoustics tasks.
+This assertion shall be studied further in future work by considering more complex convolutional networks, larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot, lafayPartI} and other ecoacoustics tasks.
 

### 2016-07-12 paper/structureScene.tex
write on log compression

@@
 The scattering transform is defined as an infinite cascade of wavelet transform and modulus operators.
-However, to achieve invariance to translation up to $T = 186\,\mathrm{ms}$, \ie the approximately minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ often suffice.
+However, to achieve invariance to translation up to $T = 372\,\mathrm{ms}$, \ie the approximately minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ often suffice.
 
@@
 (a) before and (b) after logarithmic compression.}
+\label{fig:compression}
 \end{center}
@@
 If we're dealing with Gaussian processes, this is in fact a very good thing to do.}
-On the contrary, a late integration scheme for $T=188\textrm{ ms}$ would yield $128$ instances per auditory scene \ja{How long is each recording?
+On the contrary, a late integration scheme for $T=372\textrm{ ms}$ would yield $128$ instances per auditory scene \ja{How long is each recording?
 Should we specify this somewhere?}, resulting in $10240$ instances overall.
@@
 Experiments are carried out using temporal scattering coefficients as well as mel-frequency cepstral coefficients (MFCCs) as features.
-For the temporal scattering, each scene is described by $128$ vectors of scattering coefficients computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=188\,\mathrm{ms}$.
+For the temporal scattering, each scene is described by $128$ vectors of scattering coefficients computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$.
 Experiments are conducted with and without the logarithmic compression (see Section \ref{sec:logcomp}).
@@
 \subsubsection*{MFCCs \vs scattering}
-As shown in Figure \cite{folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
+as shown in Figure \cite{folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
 Not only the average accuracy is improved, but the variance between fold-wise accuracies is also reduced, \eg from $16\%$ to $7\%$ in the case of early integration.
@@
 \subsubsection*{Early \vs late integration}
+late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
 
 \subsubsection*{Role of logarithmic compression}
+logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
+Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
+This transformation improves the discriminative power of support vector machines (SVM) trained on scattering coefficients, whatever be the chosen temporal integration strategy (early or late) or the inner product kernel (linear or Gaussian RBF).
+Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.
 

### 2016-07-12 paper/structureScene.tex
horz lines in Fig 5
reformulate EMD section VI

@@
 
-\gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}. 
-\vl{this section should be better segmented into subsections. The words centroid and cluster are defined after they are used. The order of paragraphs could be 1-4-3-2-5.}
-
 As discussed in Section \ref{sec:motivations}, results in sound perception suggest the appropriateness of an object-based representation of auditory scenes for predicting high-level properties.
@@
 \ja{What does this mean?} In this paper, we therefore focus on a simple quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
+The clustering is done using the centroid-based clustering algorithm KMeans.
 
-\ja{We need to describe the clustering here and what it gives as output. Otherwise it becomes difficult to describe the way we compare scenes below.}
+Given a set a $d$-dimensional feature vectors $x_l^u\in X_u$, $l=[1,2,\ldots,L]$, extracted from the scene $s_u$, $u=[1,2,\ldots,U]$, the goal of KMeans is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=[1,2,\ldots,M]$.
+The partitioning is done by minimizing for each $m$ cluster the squared error between its empirical means (centroid) and the points it contains.
+Considering $\mu_m^u$ the centroid of the cluster $c_m^u$, KMeans minimizes the following objective function:
 
+\begin{equation}
+J(C_u)=\sum\limits_{m} \sum_{x^u_l\in c^u_m} \vert \vert x_l^u - \mu_m^u \vert \vert^2
+\end{equation}
+
+At the end, each scene $s_u$ is described by a set of clusters $C_u$.
 One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, which clusters not over a single scene, but over the whole dataset, and re-projects each scene in the resulting enhanced feature space.
 \ja{This sentence is also a big unclear.}
-Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, we consider several centroid-based similarity metrics:
+
+Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on their centroids similarities.
+
+The similarity between all the scene centroids $\mu_i$, $i=[1,2,\ldots,I]$ with $I=UM$, is computed using a radial basis function (RBF) kernel $K$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
+\begin{equation}
+\label{eq:kc}
+K_{ij} = \exp\left( - \dfrac{\Vert \mu_i - \mu_j \Vert^2}{\Vert \mu_i - \mu_{k,i} \Vert \Vert \mu_j - \mu_{k,j} \Vert} \right) 
+\end{equation}
+
+where $\mu_{k,i}$ and $\mu_{k,j}$ are the $k^{\textrm{th}}$ nearest neighbor to the centroids $\mu_i$ and $\mu_j$ respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+
+To compute the similarity between two scenes, we then consider several centroid-based similarity metrics:
 
@@
 \item \emph{ob-averaged} (ob-a): the similarity between two scenes is equal to the average of their centroids similarities, and
-\item \emph{ob-weighted} (ob-w): for each scene, each centroid is weighted according to the number of frames belonging to its cluster. \ja{And then we compute the EMD?}
+\item \emph{ob-weighted} (ob-w): for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
+Each scene $s_u$ is thus described by a signature $p_u$ of $M$ clusters ($p_u=\lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace$), with $\mu_m^u$ and $w_m^u$ being the centroid and the weight of the $m$th cluster respectively.
+A cross-bin histogram distance, taking into account both the clusters weights of the scenes and the distances between their cluster centroids is then used.
 \end{itemize}
 
-The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function (RBF) kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
-\begin{equation}
-\label{eq:kc}
-K^c_{ij} = \exp\left( - \dfrac{\Vert c_i - c_j \Vert^2}{\Vert c_i - c_N \Vert \Vert c_j - c_N \Vert} \right) 
-\end{equation} \ja{To simplify the analysis later, can we already specify that we're looking at a specific pair $(u, v)$ and that the kernel $K^c$ is defined only for $i$ in $u$ and $j$ in $v$?}
-where $c_N$ is the $N^{\textrm{th}}$ nearest neighbor to the centroid $c_i$, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
-\ja{How does the $c_N$ notation distinguish between $c_i$ and $c_j$?
-Do we have $c_{i,N}$ and $c_{j,N}$?}
-
-Compared to the closest or averaged approach, the weighted similarity between two scenes takes into account both the centroids similarities and the clusters weights.
-Each scene $s_u$ is described by a signature $S_u$ of $m$ clusters ($S_u=\lbrace(c_1^u,w_1^u),\ldots,(c_m^u,w_m^u)\rbrace$), with $c_i^u$ and $w_i^u$ being the centroid and the weight of the $i$th cluster of the scene $s_u$ respectively.
-\ja{Everything that follows only refers to the ob-w metric, correct?
-If so, we should specify it here, and also say how the similarity is explicitly calculated for the other options}
-
-To compute the distance between two signatures $S_u$ and $S_v$, we use a variant of the earth mover's distance ($\EMD$), known as $\widehat{\EMD}$ and introduced in \cite{pele2008linear}.
-The $\EMD$ distance is a cross-bin histogram distance, widely used in computer vision \cite{zhang2007local}, which computes the minimal cost to be paid to transform one histogram into the other.
-An advantage of $\EMD$ is that automatically aligns the histogram bins by using a ``ground distance,'' which is the distance between the bin centers in the two histograms.
-In our case, the histograms are the cluster weights $w$ defined over the bins formed by cluster centroids $c$.
+For \emph{ob-w}, the cross-bin histogram distance used is a variant of the earth mover's distance ($\EMD$), known as $\widehat{\EMD}$ and introduced in \cite{pele2008linear}.
+$\EMD$ computes the distance between two histograms by finding the minimal cost to be paid to transform one histogram into the other.
+An advantage of $\EMD$ is that it automatically aligns the histogram bins by using a ``ground distance'' which is the distance between the bin centers in the two histograms.
+In our case, the histograms are the cluster weights $w^u$ defined over the bins formed by cluster centroids $\mu^u$.
 
@@
 To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
-Given two signatures $S_u$ and $S_v$, the $\widehat{EMD}$ is computed by solving the following linear programming problem:
+Given two signatures $p_u$ and $p_v$ made of $n$ and $m$ clusters respectively, the $\widehat{EMD}$ is computed by solving the following linear programming problem:
 
@@
 \begin{split}
-\widehat{\EMD}(S_u,S_v) &=\left( \min\limits_{\lbrace f_{ij}\rbrace} \sum\limits_{i,j} f_{ij}D_{ij} \right) \\
-&+ \left|\sum\limits_{i} w_i^u - \sum\limits_{j} w_j^v  \right| \alpha \max\limits_{i,j}\lbrace  D_{ij}\rbrace
+\widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
+&+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \alpha \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace
 \end{split}
@@
 \begin{equation*}
-\mathrm{s.t.} \quad f_{ij}\geq0 \quad \sum\limits_{j} f_{ij} \leq w_i^u \quad \sum\limits_{i} f_{ij} \leq w_j^v
+\mathrm{s.t.} \quad f_{nm}\geq0 \quad \sum\limits_{m} f_{nm} \leq w_n^u \quad \sum\limits_{n} f_{nm} \leq w_m^v
 \end{equation*}
@@
 \begin{equation*}
-\sum\limits_{i,j}f_{ij} = \min\left( \sum\limits_{i} w_i^u ,\sum\limits_{j} w_j^v \right)
+\sum\limits_{n,m}f_{nm} = \min\left( \sum\limits_{n} w_n^u ,\sum\limits_{m} w_m^v \right)
 \end{equation*} \ja{Can we provide some interpretation for the different terms and constraints here?}
-where $\lbrace f_{ij} \rbrace$ is the flow between the cluster weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to ``supply the demand'' of the $j$th bin.
-We denote by $D_{ij}$ the ground distance between the centroids $c_i^u$ and $c_j^v$ computed from the kernel $K^c$:
+where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n$th bin to ``supply the demand'' of the $m$th bin.
+We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroid sets $\mu^u$ and $\mu^v$.
+$D^{uv}$ is computed from the kernel $K$:
 \begin{equation*}
-D_{ij}=1-K^{c(uv)}_{ij}
+D^{uv}=1-K^{(uv)}
 \end{equation*}
-with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ of the scenes $s_u$ and $s_v$ respectively.
+with $K^{(uv)}$ being the slice of $K$ containing the pairwise similarities between the centroids sets $\mu^u$ and $\mu^v$ of the scenes $s_u$ and $s_v$ respectively.
 As suggested in \cite{pele2009fast}, we set the tradeoff parameter $\alpha$ to $1$.
@@
 \label{eq:ks}
-K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(S_u,S_v)}{A} \right) \\
+K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(p_u,p_v)}{A} \right) \\
 \end{equation}
@@
 The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.
-See text for details.\ml{je mettrai des lignes fines entre les points 1 et 3, et 4 et 6, je zoomerai un peu et l'axe des y n'est pas numerote ni labelise }}
+See text for details}
 \end{center}

### 2016-07-12 paper/structureScene.tex
upload table III

@@
 Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
-This transformation improves the discriminative power of support vector machines (SVM) trained on scattering coefficients, whatever be the chosen temporal integration strategy (early or late) or the inner product kernel (linear or Gaussian RBF).
+This transformation improves the discriminative power of support vector machines (SVM) trained on scattering coefficients, whatever be the chosen temporal integration strategy -- \ie early or late -- or the inner product kernel -- \ie linear or Gaussian RBF.
 Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.
-
-\subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
 
@@
 The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.
-See text for details}
+See text for details.}
 \end{center}
-\ref{fig:folds}
+\label{fig:folds}
 \end{figure}
@@
 \end{tabular}
-\caption{Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.}
+\caption{Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.
+See text for details.}
 \end{center}
-\ref{table:dcase2013}
+\label{table:dcase2013}
+\end{table}
+
+\subsubsection*{Participation to the DCASE 2016 challenge}
+In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{lcc}
+Features & MFCCs & log-scattering \\
+Classifier  & GMMs    & linear SVMs \\
+Temporal integration & late (product rule) & late (majority voting) \\
+\midrule
+beach & $74.6 \pm 18.9$ & $83.5 \pm \phantom{0}7.1$ \\
+bus & $58.2 \pm 18.3$ & $88.8 \pm 11.4$ \\
+caf\'{e} / restaurant & $85.1 \pm 10.8$ & $64.5 \pm \phantom{0}8.3$ \\
+car & $69.1 \pm 20.8$ & $94.9 \pm \phantom{0}6.1$ \\
+city center & $89.6 \pm \phantom{0}9.2$ & $91.7 \pm \phantom{0}9.4$ \\
+forest path & $72.4 \pm 23.8$ & $93.8 \pm \phantom{0}7.8$ \\
+grocery store & $74.1 \pm 13.7$ & $90.9 \pm \phantom{0}6.9$ \\
+home & $78.1 \pm 17.7$ & $56.2 \pm 23.9$ \\
+library & $65.1 \pm 23.3$ & $82.0 \pm 13.0$ \\
+metro station & $85.2 \pm 15.6$ & $96.0 \pm \phantom{0}2.3$ \\
+office & $90.8 \pm 16.0$ & $87.5 \pm 21.7$ \\
+park & $25.6 \pm 11.0$ & $75.4 \pm \phantom{0}7.4$ \\
+residential area & $75.1 \pm 19.4$ & $44.2 \pm 15.0$ \\
+train & $34.1 \pm \phantom{0}7.2$ & $58.3 \pm 10.0$ \\
+tram & $85.4 \pm 13.9$ & $83.7 \pm 12.2$ \\
+\bottomrule
+Average & $70.8 \pm \phantom{0}2.6$ & $79.4 \pm \phantom{0}3.0$
+\end{tabular}
+\end{center}
+\label{table:dcase2016}
+\caption{Class-wise accuracies on the DCASE 2016 dataset for two pipelines: a baseline of mel-frequency cepstral coefficients (MFCCs), Gaussian mixture models (GMMs), and late temporal integration by product of likelihoods ; and our proposed algorithm of log-scattering coefficients, support vector machines (SVMs) with a linear kernel, and late temporal integration by majority voting.
+See text for details.}
 \end{table}

### 2016-07-12 paper/structureScene.tex
update section VI

--- 2016-07-12 paper/structureScene.tex
@@
 An advantage of $\EMD$ is that it automatically aligns the histogram bins by using a ``ground distance'' which is the distance between the bin centers in the two histograms.
-In our case, the histograms are the cluster weights $w^u$ defined over the bins formed by cluster centroids $\mu^u$.
+In our case, the histograms are the clusters weights $w^u$ defined over the bins formed by clusters centroids $\mu^u$.
 
@@
 where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n$th bin to ``supply the demand'' of the $m$th bin.
-We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroid sets $\mu^u$ and $\mu^v$.
+We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$.
 $D^{uv}$ is computed from the kernel $K$:

### 2016-07-12 paper/structureScene.tex
address joakim's comments in IV. B

@@
 To discard this superfluous source of variability, signals can be mapped to a time-shift invariant feature space before training the classifier, eliminating the need for the classifier to explicitly learn this invariance.
-From any set of descriptors -- \eg the time-frequency representation $\boldsymbol{x_1}(t,\gamma_1)$ \ja{$x_1$, $\gamma_1$, etc have not been introduced yet.} of a signal $\boldsymbol{x}(t)$ --
-a representation invariant to time shifts shorter in duration than $T$ can be obtained
-by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency
-set to $1/T$ \ja{How are we defining frequency here, as $\omega$ or $2\pi\omega$?
-In the former case, we should have $2\pi/T$.
-Probably we should define our Fourier transform somewhere to nail this down}:
+From any set of descriptors $\boldsymbol{x_1}(t,\gamma)$, where $\gamma$ denotes a descriptor index, a representation invariant to time shifts shorter in duration than $T$ can be obtained by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency set to $1/T$, as measured in Hertz:
 \begin{equation}
@@
 in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}.
-In addition time-shift invariance, this stability property is crucial to signal classification, since it guarantees robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.
+In addition to invariance, this stability property is crucial to signal classification, since it guarantees robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.
 
@@
 \subsection{Temporal scattering}
+Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int x(t) \exp(\mathrm{i} 2\pi t \omega) \; \mathrm{d}\omega$.
+Note that the frequency variable $\omega$ is expressed in Hertz, not in radians per second.
+
 Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
-center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$. \ja{I mentioned this already, but it would be good to define the Fourier transform so that we can make these statements more specific}
+center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$.
 A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
@@
 \end{equation}
-The variable $\gamma_1$ is an inverted log-frequency, taking integer values between $0$ and $(J_1 \times Q_1 - 1)$.
+The variable $\gamma_1$ is the opposite number of a log-frequency, taking integer values between $0$ and $(J_1 \times Q_1 - 1)$.
 In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$ (close to the Nyquist frequency of the audio recordings) the number of octaves $J_1$ to $10$ (the lower end of human hearing range) and the number of wavelets per octave $Q_1$ to $8$.

### 2016-07-12 paper/structureScene.tex
update KMeans++ description

@@
 Optimal parameters (see above) are this time learned on the public part of the DCASE2013 dataset, also using five-fold stratified cross validation.
-Clustering is done using an enhanced version of the popular K-means algorithm known as $K$-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.
-\ja{How are they selected differently?}
+Clustering is done using KMeans$++$ \cite{arthur2007k}, an augmented version of K-means with a robust initialization procedure.
 

### 2016-07-12 paper/structureScene.tex
address ja comments in VI

@@
 At the end, each scene $s_u$ is described by a set of clusters $C_u$.
-One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, which clusters not over a single scene, but over the whole dataset, and re-projects each scene in the resulting enhanced feature space.
-\ja{This sentence is also a big unclear.}
+One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
+
+%which clusters not over a single scene, but over the whole dataset, and re-projects each scene in the resulting enhanced feature space. \ja{This sentence is also a big unclear.}
 

### 2016-07-12 paper/structureScene.tex
bis. Cite Brown 1992

@@
 theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the temporal resolution of the CQT is fixed by the support of the STFT analyzing window. \ja{This is not true. See for example Sch\"orkhuber and Klapuri 2010.}
+frequency range, whereas the temporal resolution of the CQT, at least in its original implementation \cite{Brown1992}, is fixed by the support of the STFT analyzing window.
 Therefore, the wavelet scalogram has a better temporal localization at high
-frequencies than the CQT, at the expense of a greater computational cost. \ja{Why is the cost greater?}
-This property allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+frequencies than the CQT.
+This increase in time-frequency localization comes at the expense of a greater computational cost, the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
+Yet, it allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
 

### 2016-07-12 paper/structureScene.tex
no need to cite Weber-Fechner

@@
 Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
-Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. \ja{Cite?}
+Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
 We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone, a source of spurious variability in scene classification.

### 2016-07-12 paper/structureScene.tex
small corrections section VI

--- 2016-07-12 paper/structureScene.tex
@@
 Given a set a $d$-dimensional feature vectors $x_l^u\in X_u$, $l=[1,2,\ldots,L]$, extracted from the scene $s_u$, $u=[1,2,\ldots,U]$, the goal of KMeans is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=[1,2,\ldots,M]$.
-The partitioning is done by minimizing for each $m$ cluster the squared error between its empirical means (centroid) and the points it contains.
-Considering $\mu_m^u$ the centroid of the cluster $c_m^u$, KMeans minimizes the following objective function:
+The partitioning is done by minimizing for each $m$ cluster the squared error between its empirical mean (centroid) and the points it contains.
+Given $\mu_m^u$ the centroid of the cluster $c_m^u$, KMeans minimizes the following objective function:
 

### 2016-07-12 paper/structureScene.tex
re-upload table III

@@
 
-\subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
-
 \begin{figure}
@@
 \end{center}
-\ref{fig:folds}
+\label{fig:folds}
 \end{figure}
@@
 \end{center}
-\ref{table:dcase2013}
+\label{table:dcase2013}
+\end{table}
+
+\subsubsection*{Participation to the DCASE 2016 challenge}
+In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
+
+The results on the evaluation dataset are awaiting publication by the DCASE organizing committee.
+Our results on the development dataset are reported in Table \ref{table:dcase2016}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{lcc}
+Features & MFCCs & log-scattering \\
+Classifier  & GMMs    & linear SVMs \\
+Temporal integration & late (product rule) & late (majority voting) \\
+\midrule
+beach & $74.6 \pm 18.9$ & $83.5 \pm \phantom{0}7.1$ \\
+bus & $58.2 \pm 18.3$ & $88.8 \pm 11.4$ \\
+caf\'{e} / restaurant & $85.1 \pm 10.8$ & $64.5 \pm \phantom{0}8.3$ \\
+car & $69.1 \pm 20.8$ & $94.9 \pm \phantom{0}6.1$ \\
+city center & $89.6 \pm \phantom{0}9.2$ & $91.7 \pm \phantom{0}9.4$ \\
+forest path & $72.4 \pm 23.8$ & $93.8 \pm \phantom{0}7.8$ \\
+grocery store & $74.1 \pm 13.7$ & $90.9 \pm \phantom{0}6.9$ \\
+home & $78.1 \pm 17.7$ & $56.2 \pm 23.9$ \\
+library & $65.1 \pm 23.3$ & $82.0 \pm 13.0$ \\
+metro station & $85.2 \pm 15.6$ & $96.0 \pm \phantom{0}2.3$ \\
+office & $90.8 \pm 16.0$ & $87.5 \pm 21.7$ \\
+park & $25.6 \pm 11.0$ & $75.4 \pm \phantom{0}7.4$ \\
+residential area & $75.1 \pm 19.4$ & $44.2 \pm 15.0$ \\
+train & $34.1 \pm \phantom{0}7.2$ & $58.3 \pm 10.0$ \\
+tram & $85.4 \pm 13.9$ & $83.7 \pm 12.2$ \\
+\bottomrule
+Average & $70.8 \pm \phantom{0}2.6$ & $79.4 \pm \phantom{0}3.0$
+\end{tabular}
+\end{center}
+\label{table:dcase2016}
+\caption{Class-wise accuracies on the DCASE 2016 dataset for two pipelines: a baseline of mel-frequency cepstral coefficients (MFCCs), Gaussian mixture models (GMMs), and late temporal integration by product of likelihoods ; and our proposed algorithm of log-scattering coefficients, support vector machines (SVMs) with a linear kernel, and late temporal integration by majority voting.
+See text for details.}
 \end{table}

### 2016-07-12 paper/structureScene.tex
brush up IV.A

@@
 
-Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filterbanks interspersed with complex modulus nonlinearities.
+Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filter banks interspersed with complex modulus nonlinearities.
 In this section, we present the standard temporal scattering transform.
@@
 \begin{equation}
-\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = (\boldsymbol{x_1} \ast \boldsymbol{\phi}) (t).
+\mathbf{S_1}\boldsymbol{x}(t, \gamma) = (\boldsymbol{x_1} \ast \boldsymbol{\phi}) (t).
 \end{equation}
@@
 
-The scattering transform is defined as an infinite cascade of wavelet transform and modulus operators.
-However, to achieve invariance to translation up to $T = 372\,\mathrm{ms}$, \ie the approximately minimal duration between non-overlapping acoustic events, two layers of scattering transform of the audio signal $\boldsymbol{x}(t)$ often suffice.
+Starting from a monophonic waveform $\boldsymbol{x}(t)$, the scattering transform is defined as an infinite cascade of wavelet transform and modulus operators.
+However, to achieve invariance to translation up to $T = 372\,\mathrm{ms}$, \ie the approximately minimal duration between non-overlapping acoustic events, two layers of scattering transform.
 
-The next subsection describes the operations involved in temporal scattering.
+The next subsection describes the operations involved in temporal scattering, and in particular the construction of wavelet filter banks.
 
 \ja{I'm not sure that this whole section makes sense right here.
-It can be a bit daunting to jump right in and talk about spectrograms $x_1$ and wavelelets, etc.
+It can be a bit daunting to jump right in and talk about spectrograms $x_1$ and wavelets, etc.
 before we've defined them.
@@
 The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
-The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.
-See text for details}
+The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}}
 \end{center}
@@
 \label{table:dcase2016}
-\caption{Class-wise accuracies on the DCASE 2016 dataset for two pipelines: a baseline of mel-frequency cepstral coefficients (MFCCs), Gaussian mixture models (GMMs), and late temporal integration by product of likelihoods ; and our proposed algorithm of log-scattering coefficients, support vector machines (SVMs) with a linear kernel, and late temporal integration by majority voting.
-See text for details.}
+\caption{Class-wise accuracies on the DCASE 2016 dataset for two pipelines: a baseline of mel-frequency cepstral coefficients (MFCCs), Gaussian mixture models (GMMs), and late temporal integration by product of likelihoods ; and our proposed algorithm of log-scattering coefficients, support vector machines (SVMs) with a linear kernel, and late temporal integration by majority voting.}
 \end{table}

### 2016-07-12 paper/structureScene.tex
more in VIII.B on DCASE 2016

@@
 
-The results on the evaluation dataset are awaiting publication by the DCASE organizing committee.
-Our results on the development dataset are reported in Table \ref{table:dcase2016}
+Our results on the development dataset are reported in Table \ref{table:dcase2016}, in comparison with a baseline of $MFCCs$ coefficients and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
+The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
+It appears that log-scattering features compare favorably to 
+The results on the evaluation dataset are awaiting publication by the DCASE organizing committee, and will be disclosed in August 2016.
 
@@
 
-This assertion shall be studied further in future work by considering more complex convolutional networks, larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot, lafayPartI} and other ecoacoustics tasks.
+This assertion shall be studied further in future work by considering data-driven convolutional networks, larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot, lafayPartI} as well as other tasks in ecoacoustics.
 

### 2016-07-12 paper/structureScene.tex
no need to cite Weber-Fechner, bis

@@
 Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
-Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. \ja{Cite?}
+Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. 
 We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone, a source of spurious variability in scene classification.

### 2016-07-12 paper/structureScene.tex
put back my changes reverted by gregoire

@@
 theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the temporal resolution of the CQT is fixed by the support of the STFT analyzing window. \ja{This is not true. See for example Sch\"orkhuber and Klapuri 2010.}
+frequency range, whereas the temporal resolution of the CQT, at least in its original implementation \cite{Brown1992}, is fixed by the support of the STFT analyzing window.
 Therefore, the wavelet scalogram has a better temporal localization at high
-frequencies than the CQT, at the expense of a greater computational cost. \ja{Why is the cost greater?}
-This property allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+frequencies than the CQT, at the expense of a greater computational cost.
+This increase in time-frequency localization comes at the expense of a greater computational cost, the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
+Yet, it allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
 

### 2016-07-12 paper/structureScene.tex
more on VIII.B

@@
 late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
+This means that, despite its discriminative power at fine scales, the scattering transform is unable to segregate distinct acoustic events beyond the scale of about one second.
 
@@
 \subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
+In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge, whose results are forthcoming.
+This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude
 

### 2016-07-12 paper/structureScene.tex
cite Mesaros

@@
 In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge, whose results are forthcoming.
-This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude
+This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
 

### 2016-07-12 paper/structureScene.tex
finish writing on DCASE 2016

@@
 This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
+Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
+Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel.
+Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
 
@@
 The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
-It appears that log-scattering features compare favorably to 
 The results on the evaluation dataset are awaiting publication by the DCASE organizing committee, and will be disclosed in August 2016.

### 2016-07-12 paper/structureScene.tex
write on the failure of object-based integration for supervised classification

@@
 late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
-This means that, despite its discriminative power at fine scales, the scattering transform is unable to segregate distinct acoustic events beyond the scale of about one second.
+This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to segregate distinct acoustic events beyond the scale of about one second.
+Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
+Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none of them managed to outperform the baseline strategy in temporal integration, namely majority voting.
 

### 2016-07-12 paper/structureScene.tex
address JA comments section Experiment

@@
 This setting gives us $600$ feature vectors per scene.
-It has been reported that classifying the each individual frame directly may not be the best approach \cite{rakotomamonjy2015histogram}.
-\ja{Why?} In order to reduce the amount of data, a pooling step is performed on the MFCCs: each recording is divided into $n$ non-overlapping $t$-seconds long ``texture'' windows.
+To get a more robust representation of lower dimensionality a pooling step is performed on the MFCCs: each recording is divided into $n$ non-overlapping $t$-seconds long ``texture'' windows.
 The MFCC vectors within each texture window are then averaged in time.

### 2016-07-12 paper/structureScene.tex
moving dataset to experiments

--- 2016-07-12 paper/structureScene.tex
@@
 
-The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}.
+Motivations of the proposed approach and a brief review of the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}.
 We describe the scattering transform in Section \ref{sec:scattering}, discuss post-processing transformations in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
@@
 
-\section{Motivations} \label{sec:motivations}
+\section{Background} \label{sec:soa}
 
@@
 
-\section{Background} \label{sec:soa}
+BOF approaches treats each observations equally, whereas the above cited results show that only a few events scarcely occurring may have be more important than events or texture present throughout the whole scene duration, thus advocating for a more selective approach such as the one presented in Section \ref{sec:object} for the ASSR task.
+
+%\section{Background} \label{sec:soa}
 
@@
 Some of this is already described in the motivations section, so perhaps we should move some over in here?}
-An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare differrent approaches to scene classification.
+An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
 Table \ref{tab:dcase} summarizes the results of the submitted algorithms where most \ja{?} of scene predictions are achieved by majority voting of predictions over features.
@@
 \cite{chum2013ieee} CHR & 65 & sparsity features, HMMs  \\
-\cite{geiger2013large} GSR & 69 & mel \ja{?} and MFCCs, SVMs \\
+\cite{geiger2013large} GSR & 69 & Mel-frequency Spectra and MFCCs, SVMs \\
 \cite{rakotomamonjy2015histogram} RG & 69 & histogram of gradient (HOG), SVMs \\
@@
 In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
-\ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked linear SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages.
+\ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages.
 Moreover, since MFCCs are well-established features which perform well for these task, we shall include them baseline features in all experiments described in this paper.
@@
 
-For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}.
-To build this dataset, three different recordists visited a wide variety of locations in Greater London over
-a period of several months and in each
-scene recorded a few minutes of audio. No
-systematic variations in the recordings covaried with scene
-type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme.
-The folds used in this paper are the same as the ones used during the challenge.
-\ja{Why are we describing the dataset here in the background section?
-Wouldn't it be more appropriate under experiments?}
+ % The latter is also considered in this paper as a validation of the proposed approach. \ja{What does this last sentence mean?}
 
-There are also two larger datasets: the
-Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge \cite{Mesaros2016_EUSIPCO}.
-The latter is also considered in this paper as a validation of the proposed approach.
-\ja{What does this last sentence mean?}
-
-The DCASE2013 dataset has significant intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
-In addition, it is still challenging, with a state-of-the-art system based on feature design \ja{?} and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge).
-Another system, using a HOG+SVM approach \cite{rakotomamonjy2015histogram}, achieves $75\%$ and $92\%$ on the DCASE2013 and Rouen datasets, respectively.
-Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
-On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
-
-In this paper, we shall demonstrate that the use of features based on the wavelet scattering transform together with object-based representation of the acoustic scene allows us to achieve state-of-the-art \ja{?} performance in both the ASSR and ASC paradigms.
+%In this paper, we shall demonstrate that the use of features based on the wavelet scattering transform together with object-based representation of the acoustic scene allows us to achieve state-of-the-art \ja{?} performance in both the ASSR and ASC paradigms.
 
@@
 theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the temporal resolution of the CQT, at least in its original implementation \cite{Brown1992}, is fixed by the support of the STFT analyzing window.
+frequency range, whereas the temporal resolution of the CQT is fixed by the support of the STFT analyzing window. \ja{This is not true. See for example Sch\"orkhuber and Klapuri 2010.}
 Therefore, the wavelet scalogram has a better temporal localization at high
-frequencies than the CQT, at the expense of a greater computational cost.
-This increase in time-frequency localization comes at the expense of a greater computational cost, the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
-Yet, it allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+frequencies than the CQT, at the expense of a greater computational cost. \ja{Why is the cost greater?}
+This property allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
 
@@
 Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
-Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. 
+Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. \ja{Cite?}
 We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone, a source of spurious variability in scene classification.
@@
 
+\subsection{Datasets}
+
+%For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}. Larger datasets exists: the
+%Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
+
+In the experiments reported in this paper, the DCASE2013 \cite{giannoulis2013database, 7100934} and the public part of the DCASE2016 datasets  \cite{Mesaros2016_EUSIPCO} are considered.
+To build the DCASE2013 dataset, three different recordists visited a wide variety of locations in Greater London over
+a period of several months and in each
+scene recorded a few minutes of audio. No
+systematic variations in the recordings covaried with scene
+type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
+The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme.
+The folds used in this paper are the same as the ones used during the challenge.
+
+The DCASE2013 dataset has significant intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+In addition, it is still challenging, with a state-of-the-art system based on feature design \ja{?} and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge).
+Another system, using a HOG+SVM approach \cite{rakotomamonjy2015histogram}, achieves $75\%$ and $92\%$ on the DCASE2013 and Rouen datasets, respectively.
+Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
+On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
+
 \subsection{Feature design}
@@
 late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
-This means that, despite its discriminative power at fine scales, the scattering transform is unable to segregate distinct acoustic events beyond the scale of about one second.
 
@@
 \subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge, whose results are forthcoming.
-This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
-Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
-Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel.
-Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
+In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge.
 
@@
 The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
+It appears that log-scattering features compare favorably to 
 The results on the evaluation dataset are awaiting publication by the DCASE organizing committee, and will be disclosed in August 2016.

### 2016-07-12 paper/structureScene.tex
cite Hsu2003 remove section III.A

@@
 In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
-Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \ja{What does this mean?} \cite{Fastl2007}.
+Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
@@
 Before supervised or unsupervised classification, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power.
-In this section, we review several ways of achieving these properties, namely feature selection, logarithmic compression, standardization, and temporal integration.
-
-\subsection{Feature selection by restricting the frequency range}
-In a multivariate time series of descriptors $\mathbf{S}\boldsymbol{x}(t,\gamma)$, some coefficients $\gamma$ may consistently bear negligible information for the task of interest.
-This fact may cause the classifier to overfit the training set, especially when the number of features is large and the number of training instances is small, as is the case here.
-Appropriately selecting features before the training stage may circumvent this problem.
-However, the parameters of the chosen feature selection algorithm are themselves prone to overfitting.
-Instead, we adopt the more conservative approach of retaining all coefficients within a certain range of subband indices $\gamma_1$ and $\gamma_2$.
-\ja{These indices are already used to determine the first- and second-order log-frequency indices, respectively.
-Perhaps $\gamma_\mathrm{min}$ and $\gamma_\mathrm{max}$ would be better?}
-In comparison with adaptive algorithms, the subset of selected features $\gamma$ has a suboptimal statistical relevance, but it is easily interpretable in the realm of audio signal processing, as it is equivalent to a coarse band-pass filtering of $\boldsymbol{x}(t)$ and $\boldsymbol{x_1}(t,\gamma_1)$ over the time dimension $t$.
-\ja{Are we filering $x_1$ too?
-The selection is only in $\gamma_1$, correct?}
-
-The conservation equations underlying the physics of sound production constrain the waveform $\boldsymbol{x}(t)$ to be regular in the time domain, that is, to have a polynomial decay in the Fourier domain.
-This decay is only made sharper by the specific properties of acoustic environments, especially indoor.
-Consequently, the amount of energy in the wavelet subband $\gamma_1$ is roughly inversely proportional to the center frequency $2^{-\gamma_1/Q_1}$.
-In the topmost frequencies, this amount is so faint that the corresponding coefficients $\boldsymbol{x_1}(t,\gamma_1)$ and $\boldsymbol{x_2}(t,\gamma_1,\gamma_2)$ no longer bear discriminative information for the classification task at hand.
-\ja{The problem is that we will magnify these using a logarithm in the next section, so this argument is not as strong as that.
-Is there some other reason we would like to discard the high frequencies?}
-
-In the sequel, we evaluate the impact of discarding the scattering coefficients above a certain acoustic frequency $\xi_{1,\max}$, which corresponds to discarding subbands $\gamma_1$ below a certain scale \ja{It can be a little confusing for the reader to switch between log-frequency and scale, especially since we use both words to refer to $\gamma_1$ in some sense.
-Technically, there is a lower bound on the scale, which turns into a lower bound of $\gamma_1$, but it's not clear from first glance.}
-\begin{equation}
-\gamma_{1,\min} =
-\left\lfloor Q_1 \log_2 \frac{\xi_1}{\xi_{1,\max}} \right\rceil,
-\end{equation}
-where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.
-
-Reducing the frequency range is an effective way to enhance invariance, widely used in the literature.
-For instance, \cite{roma2013} achieved the best performance in the DCASE2013 challenge and relied on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max} = 900\,\mathrm{Hz}$ instead of including all frequencies up to the Nyquist rate.
+In this section, we review three ways of achieving these properties, namely logarithmic compression, and temporal integration, and standardization.
 
@@
 
-Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances \ja{Do we have a reference for this?}.
+Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
 Yet the distribution of the scattering coefficients is skewed towards larger values. We can reduce this skewness by applying a pointwise concave transformation, like a logarithm, to all coefficients.
@@
 \subsection{Acoustic Scene Classification}
+We report experiments on acoustic scene classification (ASC), following the original methodology of the DCASE 2013 challenge, and extend our approach to the its 2016 edition.
 
@@
 Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
-Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel.
+Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel \cite{Hsu2003}.
 Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.

### 2016-07-12 paper/structureScene.tex
refactor beginning of section III

@@
 Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filter banks interspersed with complex modulus nonlinearities.
-In this section, we present the standard temporal scattering transform.
+This section highlights the importance of invariance to time shifts and stability to time warps in the representation of acoustic scenes, and explains how the scattering transform is designed to satisfy these properties while having a high discriminative power.
 
-\subsection{Invariance and stability in acoustic scene classification}
-The notion of invariance to time shifts plays an essential role in auditory scene classification.
+\subsection{Invariance and stability in acoustic scenes}
+The notion of invariance to time shifts plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
 Indeed, recordings may be shifted locally without affecting their perception and therefore such shifts do not convey any information about the class.

### 2016-07-12 paper/structureScene.tex
move early int and late int description in section IV.B

@@
 \ja{Another important difference is that early integration makes the classifier objective more appropriate, since the SVM will maximize the final classification error on the training data, as opposed to some intermediate frame-level label.}
-
+In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions. 
+%\ja{We should perhaps mention this in that particular section instead of over here.}
 \subsection{Standardization}
@@
 For ASSR, evaluation is performed on the private part of the DCASE2013 dataset.
-The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averagin over all query items.
+The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
 The $p@k$ is computed for $k=\{1,\ldots,9\}$, since each class only has $10$ items \ja{?}.
@@
 The \emph{ob} approaches are compared to commonly used early integration approach (\emph{early}) for ASSR, while both early and late integration (\emph{late}) strategy are evaluated for the ASC setting (see Section \ref{sec:eili}).
-In this context early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.
-\ja{We should perhaps mention this in that particular section instead of over here.}
 

### 2016-07-12 paper/structureScene.tex
address Joacim's comments on section III + new subtitles

@@
 
-\section{Scattering representations \label{sec:scattering}}
+\section{Wavelet scattering \label{sec:scattering}}
 
@@
 
-\ja{I'm not sure that this whole section makes sense right here.
-It can be a bit daunting to jump right in and talk about spectrograms $x_1$ and wavelets, etc.
-before we've defined them.
-Could we put this after the next subsection, explaining all the nice properties of the scattering transform and why we need them after we've defined it fully?}
-
-\subsection{Temporal scattering}
+\subsection{Wavelet scalogram}
 Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int x(t) \exp(\mathrm{i} 2\pi t \omega) \; \mathrm{d}\omega$.
@@
 
+\subsection{Extracting modulations with second-order scattering}
 Among auditory scenes, amplitude modulations may be caused by a broad variety of mechanical interactions, including collision, friction, and turbulent flow.
@@
 
+\subsection{Gammatone wavelets}
 \begin{figure}
@@
 \end{figure}
-
 Wavelets
@@
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
-\ja{Why is this relevant for our classification task?
-Sparsity?}
 As such, they can be discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
+This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparse time-frequency representation of acoustic scenes.
 

### 2016-07-12 paper/structureScene.tex
fix first question in sec obect

--- 2016-07-12 paper/structureScene.tex
@@
 
-Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
-\ja{Since we go into so much detail on ASC, shouldn't we at lest describe what these works do?
-Some of this is already described in the motivations section, so perhaps we should move some over in here?}
+Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC task. 
 An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
@@
 \cite{roma2013recurrence} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
+\hline
+
 \end{tabular}
 \end{center}
-\caption{Designs submitted to the DCASE2013 ASC challenge. RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning. \label{tab:dcase}}
+\caption{Algorithms evaluated on the DCASE2013 ASC private DCASE2013 dataset.
+RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning.
+Algorithms above the horizontal line participated to the DCASE challenge.
+Others used the private dataset afterward.
+\label{tab:dcase}}
 \end{table}
@@
 Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filter banks interspersed with complex modulus nonlinearities.
-This section highlights the importance of invariance to time shifts and stability to time warps in the representation of acoustic scenes, and explains how the scattering transform is designed to satisfy these properties while having a high discriminative power.
+In this section, we present the standard temporal scattering transform.
 
-\subsection{Invariance and stability in acoustic scenes}
-The notion of invariance to time shifts plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
+\subsection{Invariance and stability in acoustic scene classification}
+The notion of invariance to time shifts plays an essential role in auditory scene classification.
 Indeed, recordings may be shifted locally without affecting their perception and therefore such shifts do not convey any information about the class.
@@
 In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
-Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
+Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking \ja{What does this mean?} \cite{Fastl2007}.
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
@@
 Before supervised or unsupervised classification, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power.
-In this section, we review three ways of achieving these properties, namely logarithmic compression, and temporal integration, and standardization.
+In this section, we review several ways of achieving these properties, namely feature selection, logarithmic compression, standardization, and temporal integration.
+
+\subsection{Feature selection by restricting the frequency range}
+In a multivariate time series of descriptors $\mathbf{S}\boldsymbol{x}(t,\gamma)$, some coefficients $\gamma$ may consistently bear negligible information for the task of interest.
+This fact may cause the classifier to overfit the training set, especially when the number of features is large and the number of training instances is small, as is the case here.
+Appropriately selecting features before the training stage may circumvent this problem.
+However, the parameters of the chosen feature selection algorithm are themselves prone to overfitting.
+Instead, we adopt the more conservative approach of retaining all coefficients within a certain range of subband indices $\gamma_1$ and $\gamma_2$.
+\ja{These indices are already used to determine the first- and second-order log-frequency indices, respectively.
+Perhaps $\gamma_\mathrm{min}$ and $\gamma_\mathrm{max}$ would be better?}
+In comparison with adaptive algorithms, the subset of selected features $\gamma$ has a suboptimal statistical relevance, but it is easily interpretable in the realm of audio signal processing, as it is equivalent to a coarse band-pass filtering of $\boldsymbol{x}(t)$ and $\boldsymbol{x_1}(t,\gamma_1)$ over the time dimension $t$.
+\ja{Are we filering $x_1$ too?
+The selection is only in $\gamma_1$, correct?}
+
+The conservation equations underlying the physics of sound production constrain the waveform $\boldsymbol{x}(t)$ to be regular in the time domain, that is, to have a polynomial decay in the Fourier domain.
+This decay is only made sharper by the specific properties of acoustic environments, especially indoor.
+Consequently, the amount of energy in the wavelet subband $\gamma_1$ is roughly inversely proportional to the center frequency $2^{-\gamma_1/Q_1}$.
+In the topmost frequencies, this amount is so faint that the corresponding coefficients $\boldsymbol{x_1}(t,\gamma_1)$ and $\boldsymbol{x_2}(t,\gamma_1,\gamma_2)$ no longer bear discriminative information for the classification task at hand.
+\ja{The problem is that we will magnify these using a logarithm in the next section, so this argument is not as strong as that.
+Is there some other reason we would like to discard the high frequencies?}
+
+In the sequel, we evaluate the impact of discarding the scattering coefficients above a certain acoustic frequency $\xi_{1,\max}$, which corresponds to discarding subbands $\gamma_1$ below a certain scale \ja{It can be a little confusing for the reader to switch between log-frequency and scale, especially since we use both words to refer to $\gamma_1$ in some sense.
+Technically, there is a lower bound on the scale, which turns into a lower bound of $\gamma_1$, but it's not clear from first glance.}
+\begin{equation}
+\gamma_{1,\min} =
+\left\lfloor Q_1 \log_2 \frac{\xi_1}{\xi_{1,\max}} \right\rceil,
+\end{equation}
+where $\lfloor \cdot \rceil$ denotes a rounding to the nearest integer.
+
+Reducing the frequency range is an effective way to enhance invariance, widely used in the literature.
+For instance, \cite{roma2013} achieved the best performance in the DCASE2013 challenge and relied on cepstral descriptors computed up to a cutoff frequency as small as $\xi_{1,\max} = 900\,\mathrm{Hz}$ instead of including all frequencies up to the Nyquist rate.
 
@@
 
-Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
+Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances \ja{Do we have a reference for this?}.
 Yet the distribution of the scattering coefficients is skewed towards larger values. We can reduce this skewness by applying a pointwise concave transformation, like a logarithm, to all coefficients.
@@
 
-As discussed in Section \ref{sec:motivations}, results in sound perception suggest the appropriateness of an object-based representation of auditory scenes for predicting high-level properties.
-That being said, perfect knowledge of which event occurred at a given time is not feasible.
-\ja{What does this mean?} In this paper, we therefore focus on a simple quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
+As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of an object-based representation of auditory scenes for predicting high-level properties.
+As the detection of events is still an open problem \cite{7100934},  we consider in this paper a simple quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
 The clustering is done using the centroid-based clustering algorithm KMeans.
@@
 \subsection{Acoustic Scene Classification}
-We report experiments on acoustic scene classification (ASC), following the original methodology of the DCASE 2013 challenge, and extend our approach to the its 2016 edition.
 
@@
 Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
-Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel \cite{Hsu2003}.
+Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel.
 Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.

### 2016-07-12 paper/structureScene.tex
refactor start of section IV

@@
 \label{sec:design}
-\ja{It seems that what we are doing in this section is more post-processing than feature design, which would be the scattering transform itself.}
 
@@
 In this section, we review three ways of achieving these properties, namely logarithmic compression, and temporal integration, and standardization.
+Whether we adopt a time-frequency perspective or a machine learning perspective on the problem of acoustic scene classification, such operations would either be qualifies as post-processing or pre-processing.
 

### 2016-07-12 paper/structureScene.tex
remove gl comments

--- 2016-07-12 paper/structureScene.tex
@@
 
-Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC task. 
+Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC tasks.
+\ja{Since we go into so much detail on ASC, shouldn't we at lest describe what these works do?
+Some of this is already described in the motivations section, so perhaps we should move some over in here?}
 An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
@@
 \cite{roma2013recurrence} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
-\hline
-
 \end{tabular}
 \end{center}
-\caption{Algorithms evaluated on the DCASE2013 ASC private DCASE2013 dataset.
-RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning.
-Algorithms above the horizontal line participated to the DCASE challenge.
-Others used the private dataset afterward.
-\label{tab:dcase}}
+\caption{Designs submitted to the DCASE2013 ASC challenge. RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning. \label{tab:dcase}}
 \end{table}
@@
 
-As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of an object-based representation of auditory scenes for predicting high-level properties.
-As the detection of events is still an open problem \cite{7100934},  we consider in this paper a simple quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
+As discussed in Section \ref{sec:motivations}, results in sound perception suggest the appropriateness of an object-based representation of auditory scenes for predicting high-level properties.
+That being said, perfect knowledge of which event occurred at a given time is not feasible.
+\ja{What does this mean?} In this paper, we therefore focus on a simple quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
 The clustering is done using the centroid-based clustering algorithm KMeans.
@@
 It can be seen that the logarithmic compression strongly improves the results, especially for the \emph{ob} approaches, which perform worse than \emph{early} without the logarithmic compression.
-\gl{une idée du pourquoi ?}
 

### 2016-07-12 paper/structureScene.tex
move Kmeans++
explain that 372 ms is a power of 2

@@
 Starting from a monophonic waveform $\boldsymbol{x}(t)$, the scattering transform is defined as an infinite cascade of wavelet transform and modulus operators.
-However, to achieve invariance to translation up to $T = 372\,\mathrm{ms}$, \ie the approximately minimal duration between non-overlapping acoustic events, two layers of scattering transform.
+However, to achieve invariance to translation up to $T = 372\,\mathrm{ms}$, \ie the approximately minimal duration between non-overlapping acoustic events, corresponding to $2^14 = 16384$ discrete-time samples at $44100\,\mathrm{Hz}$, two layers of scattering transform.
 
@@
 Optimal parameters (see above) are this time learned on the public part of the DCASE2013 dataset, also using five-fold stratified cross validation.
-Clustering is done using KMeans$++$ \cite{arthur2007k}, an augmented version of K-means with a robust initialization procedure.
 
@@
 
+Clustering is done using KMeans$++$ \cite{arthur2007k}, an augmented version of K-means with a robust initialization procedure.
 \section{Results \label{sec:results}}

### 2016-07-12 paper/structureScene.tex
add changes removed by commit 5b26520
write concluding paragraph

--- 2016-07-12 paper/structureScene.tex
@@
 \hline
-
 \end{tabular}
@@
 \label{sec:design}
+\ja{It seems that what we are doing in this section is more post-processing than feature design, which would be the scattering transform itself.}
 
@@
 Whether we adopt a time-frequency perspective or a machine learning perspective on the problem of acoustic scene classification, such operations would either be qualifies as post-processing or pre-processing.
-
 \subsection{Logarithmic compression}
@@
 
-\ml{Vincent: un par pour resumer la partie classif}
+Within a supervised setting of acoustic scene classification (ASC), we have outlined the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{ms}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine (SVM) classification.
 %, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.

### 2016-07-12 paper/structureScene.tex
outlined -> shown

@@
 
-Within a supervised setting of acoustic scene classification (ASC), we have outlined the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{ms}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine (SVM) classification.
+Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{ms}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine (SVM) classification.
 %, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.

### 2016-07-12 paper/structureScene.tex
reformulate MFCCs description

@@
 
-Different sets of parameters were tested to compute the MFCCs.
-Results here are only reported for the best setting, which consisted of $40$ MFCCs (including the first energy coefficient) computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$, with full frequency range.
+MFCCs are computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$, with full frequency range.
+Different ranks of coefficients were tested to compute the MFCCs, results here are only reported for the best setting, which consisted of $40$ MFCCs, including the first coefficient related to the average energy.
 This setting gives us $600$ feature vectors per scene.
@@
 
-Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{ms}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine (SVM) classification.
+\ml{Vincent: un par pour resumer la partie classif}
 %, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.

### 2016-07-12 paper/structureScene.tex

@@
 
-\ml{Vincent: un par pour resumer la partie classif}
+Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{ms}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine (SVM) classification.
 %, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.

### 2016-07-12 paper/structureScene.tex
new soa algos

--- 2016-07-12 paper/structureScene.tex
@@
 An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
-Table \ref{tab:dcase} summarizes the results of the submitted algorithms where most \ja{?} of scene predictions are achieved by majority voting of predictions over features.
+Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that considered the private datset for their experiments.
+% where most \ja{?} of scene predictions are achieved by majority voting of predictions over features.
 \ja{What about the other algorithms?
@@
 \hline
+\cite{phan2016label} & 84 to 87 & Label Tree Embedding of Gammatone, SVMs \\
+\cite{7362846} & 85$^*$ & AMS, LDA \\
 \end{tabular}
 \end{center}
-\caption{Algorithms evaluated on the DCASE2013 ASC private DCASE2013 dataset.
-RQA stands for recurrence quantification analysis, ILD for inter-aural differences and UL for unsupervised learning.
+\caption{Algorithms evaluated on the DCASE2013 ASC DCASE2013 dataset.
+RQA stands for recurrence quantification analysis, ILD for inter-aural differences, AMS for amplitude Modulated Spectrum and UL for unsupervised learning.
 Algorithms above the horizontal line participated to the DCASE challenge.
-Others used the private dataset afterward.
+Others used the private dataset afterward and the last used the public dataset.
 \label{tab:dcase}}
@@
 
-In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
-\ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages.
-Moreover, since MFCCs are well-established features which perform well for these task, we shall include them baseline features in all experiments described in this paper.
+In addition, it is still challenging, with a state-of-the-art system based on feature design \ja{?} and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge).
+Another system, using a HOG+SVM approach \cite{rakotomamonjy2015histogram}, achieves $75\%$ and $92\%$ on the DCASE2013 and Rouen datasets, respectively.
+Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
+On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
 
@@
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
+As a consequence, DCASE2013 dataset has interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
 The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme.
@@
 
-The DCASE2013 dataset has significant intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
-In addition, it is still challenging, with a state-of-the-art system based on feature design \ja{?} and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge).
-Another system, using a HOG+SVM approach \cite{rakotomamonjy2015histogram}, achieves $75\%$ and $92\%$ on the DCASE2013 and Rouen datasets, respectively.
-Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
-On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
+%In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
+\ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages.
+Moreover, since MFCCs are well-established features which perform well for these task, we shall include them baseline features in all experiments described in this paper.
 

### 2016-07-12 paper/structureScene.tex

@@
 
-Given a set a $d$-dimensional feature vectors $x_l^u\in X_u$, $l=[1,2,\ldots,L]$, extracted from the scene $s_u$, $u=[1,2,\ldots,U]$, the goal of KMeans is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=[1,2,\ldots,M]$.
+Given a set of $d$-dimensional feature vectors $x_l^u\in X_u$, $l=[1,2,\ldots,L]$, extracted from the scene $s_u$, $u=[1,2,\ldots,U]$, the goal of KMeans is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=[1,2,\ldots,M]$.
 The partitioning is done by minimizing for each $m$ cluster the squared error between its empirical mean (centroid) and the points it contains.

### 2016-07-12 paper/structureScene.tex
replace cite to ref

--- 2016-07-12 paper/structureScene.tex
@@
 \subsubsection*{MFCCs \vs scattering}
-as shown in Figure \cite{folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
+as shown in Figure \ref{folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
 Not only the average accuracy is improved, but the variance between fold-wise accuracies is also reduced, \eg from $16\%$ to $7\%$ in the case of early integration.

### 2016-07-12 paper/structureScene.tex
replace cite to ref

@@
 \subsubsection*{MFCCs \vs scattering}
-as shown in Figure \ref{folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
+
+As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
 Not only the average accuracy is improved, but the variance between fold-wise accuracies is also reduced, \eg from $16\%$ to $7\%$ in the case of early integration.

### 2016-07-12 paper/structureScene.tex
cite Hsu 2003

@@
 Let $\mathbf{S}\boldsymbol{X}(\gamma,n)$ be a dataset, where the indices $\gamma$ and $n$ respectively denotes features and examples.
-It is commonly acknowledged that SVMs should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges.
-\ja{Again, it would be really good if we could cite something for this.}
+It is commonly acknowledged that SVMs should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
 This operation may also help making the dataset linearly separable in feature space. \ja{Why?}

### 2016-07-12 paper/structureScene.tex
update algos table

--- 2016-07-12 paper/structureScene.tex
@@
 \cite{patil2013multiresolution} PE & 58 & 2D Gabor filters, SVMs \\
-\cite{lee2013acoustic} NHL & 60 & UL of mel spectra, SVMs \\
+\cite{lee2013acoustic} NHL & 60 & unsupervised learning over Mel spectra, SVMs \\
 \cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
@@
 \hline
-\cite{phan2016label} & 84 to 87 & Label Tree Embedding of Gammatone, SVMs \\
-\cite{7362846} & 85$^*$ & AMS, LDA \\
+\cite{phan2016label} & 84/87 & label tree embedding of Gammatone, SVMs \\
+\cite{7362846} & 85$^*$ & amplitude modulated spectra, LDA \\
 \end{tabular}
@@
 \caption{Algorithms evaluated on the DCASE2013 ASC DCASE2013 dataset.
-RQA stands for recurrence quantification analysis, ILD for inter-aural differences, AMS for amplitude Modulated Spectrum and UL for unsupervised learning.
+RQA stands for recurrence quantification analysis and ILD for inter-aural differences.
 Algorithms above the horizontal line participated to the DCASE challenge.
-Others used the private dataset afterward and the last used the public dataset.
+Others used the datasets afterward and the last used only the public dataset ($^*$).
 \label{tab:dcase}}

### 2016-07-12 paper/structureScene.tex
update III. C
upper case in sec 6

--- 2016-07-12 paper/structureScene.tex
@@
 \subsubsection*{Early \vs late integration}
-late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
+Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
 This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to segregate distinct acoustic events beyond the scale of about one second.
@@
 
+\ml{explain why ?}
+
 \subsubsection*{Role of logarithmic compression}
-logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
+
+Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
 Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.

### 2016-07-12 paper/structureScene.tex
update datasets

@@
 Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene.
-\ja{This argument is semantically a little tricky, since technically the low-level features already contain all this high-level information, it's just that the simple models we use to describe them fail to capture this high-level structure} It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafayPartI}.
+It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafayPartI}.
 Consequently, research in sound perception is now strongly focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
@@
 An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
-Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that considered the private datset for their experiments.
-% where most \ja{?} of scene predictions are achieved by majority voting of predictions over features.
+Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that considered the private dataset for their experiments.
+Most of scene predictions are achieved by majority voting of predictions over features.
 \ja{What about the other algorithms?
@@
 \end{table}
-
-In addition, it is still challenging, with a state-of-the-art system based on feature design \ja{?} and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge).
-Another system, using a HOG+SVM approach \cite{rakotomamonjy2015histogram}, achieves $75\%$ and $92\%$ on the DCASE2013 and Rouen datasets, respectively.
-Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
-On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
 
@@
 As a consequence, DCASE2013 dataset has interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+
+In addition, it is still a challenging dataset, with a state-of-the-art system based on feature design and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieves  between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
+
+%Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
+On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
+
 The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme.

### 2016-07-12 paper/structureScene.tex
write on VI.B about boundary artifacts

@@
 A downside is that the transient information in $\boldsymbol{x_1}$ at scales finer than $T$ are lost by this low-pass filtering, reducing discriminability in feature space.
-To address this issue, the temporal scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$ and then applying complex modulus.
+To address this issue, the scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$ and then applying complex modulus.
 
@@
 
-The next subsection describes the operations involved in temporal scattering, and in particular the construction of wavelet filter banks.
+The next subsection describes the operations involved in the scattering transform, and in particular the construction of wavelet filter banks.
 
@@
 A drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene.
-In the context of the DCASE 2013 dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
+In the context of the DCASE 2013 private dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
 \ja{That may be, but by averaging, we also reduce the variability of our data, which may be a good thing.
@@
 
-Experiments are carried out using temporal scattering coefficients as well as mel-frequency cepstral coefficients (MFCCs) as features.
-For the temporal scattering, each scene is described by $128$ vectors of scattering coefficients computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$.
+Experiments are carried out using scattering coefficients as well as mel-frequency cepstral coefficients (MFCCs) as features.
+For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$.
+Indeed, $3$ seconds are discarded at the beginning and at the end of the scene so as to avoid boundary artifacts.
 Experiments are conducted with and without the logarithmic compression (see Section \ref{sec:logcomp}).

### 2016-07-12 paper/structureScene.tex
put caption in top

--- 2016-07-12 paper/structureScene.tex
@@
 \begin{center}
+\caption{Algorithms evaluated on the DCASE2013 ASC DCASE2013 dataset.
+RQA stands for recurrence quantification analysis and ILD for inter-aural differences.
+Algorithms above the horizontal line participated to the DCASE challenge.
+Others used the datasets afterward and the last used only the public dataset ($^*$).
+\label{tab:dcase}}
 \begin{tabular}{lcl}
@@
 \end{center}
-\caption{Algorithms evaluated on the DCASE2013 ASC DCASE2013 dataset.
-RQA stands for recurrence quantification analysis and ILD for inter-aural differences.
-Algorithms above the horizontal line participated to the DCASE challenge.
-Others used the datasets afterward and the last used only the public dataset ($^*$).
-\label{tab:dcase}}
 \end{table}
@@
 This setting gives us $600$ feature vectors per scene.
-To get a more robust representation of lower dimensionality a pooling step is performed on the MFCCs: each recording is divided into $n$ non-overlapping $t$-seconds long ``texture'' windows.
+To get a more robust representation of lower dimensionality a pooling step is performed on the MFCCs: each recording is divided into $n$ non-overlapping 250 seconds long ``texture'' windows \cite{1021072}, achieved a comparable number of vectors than while using the scattering ones.
+
+%.
 The MFCC vectors within each texture window are then averaged in time.
 As the \emph{ob} approaches involve a clustering step, we need to ensure that both MFCCs and scattering coefficients have a comparable number of vectors per recording.
-\ja{Why do we need to do this?} Thus, $t$ is set to $250\,\mathrm{ms}$, resulting in $120$ MFCCs texture windows per scene.
+Thus, $t$ is set to $250\,\mathrm{ms}$, resulting in $120$ MFCCs texture windows per scene.
 
-\ja{We've tested the impact of the logarithm here, but not of the feature selection or standardization.
+%\ja{We've tested the impact of the logarithm here, but not of the feature selection or standardization.
 Should we do this as well?
@@
 The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
-The $p@k$ is computed for $k=\{1,\ldots,9\}$, since each class only has $10$ items \ja{?}.
+The $p@k$ is computed for $k=\{1,\ldots,9\}$, since each class only has $10$ items.
 Note that a $p@1$ is equivalent to the classification accuracy obtained by choosing the label of the closest neighbor for a given item.
@@
 \ref{eq:kc}) leading to the best $p@9$.
-\ja{I'm not sure what this means.
-Do we fix $m$ and $N$ here for all experiments or do we optimize them over the test data each time?
-If so, this will bias the results, no?}
+
+%\ja{I'm not sure what this means. Do we fix $m$ and $N$ here for all experiments or do we optimize them over the test data each time? If so, this will bias the results, no?}
 
@@
 
-For the MFCCs, there is no clear evidence that the \emph{ob} approaches improve the results. \ja{Why?}
+%For the MFCCs, there is no clear evidence that the \emph{ob} approaches improve the results. \ja{Why?}
 
@@
 \begin{center}
+\caption{Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.}
 \begin{tabular}{llll}
@@
 \end{tabular}
-\caption{Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.}
 \end{center}
@@
 \begin{center}
+\caption{Class-wise accuracies on the DCASE 2016 dataset for two pipelines: a baseline of mel-frequency cepstral coefficients (MFCCs), Gaussian mixture models (GMMs), and late temporal integration by product of likelihoods ; and our proposed algorithm of log-scattering coefficients, support vector machines (SVMs) with a linear kernel, and late temporal integration by majority voting.}
 \begin{tabular}{lcc}
@@
 \label{table:dcase2016}
-\caption{Class-wise accuracies on the DCASE 2016 dataset for two pipelines: a baseline of mel-frequency cepstral coefficients (MFCCs), Gaussian mixture models (GMMs), and late temporal integration by product of likelihoods ; and our proposed algorithm of log-scattering coefficients, support vector machines (SVMs) with a linear kernel, and late temporal integration by majority voting.}
 \end{table}

### 2016-07-12 paper/structureScene.tex
fix first sentence in section II

@@
 
-A recent paradigm processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
+A recent paradigm in computational bioacoustics processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
 Automated systems belonging to this paradigm aim to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time regions of particular interest requesting detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.

### 2016-07-12 paper/structureScene.tex
DCASE2013 -> DCASE 2013

@@
 Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC task. 
-An overview of the efforts for the latter task is provided by the DCASE2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
+An overview of the efforts for the latter task is provided by the DCASE 2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
 Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that considered the private dataset for their experiments.

### 2016-07-12 paper/structureScene.tex
update caption of Fig 1
\textit{e.g.} -> \eg

@@
 Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene.
-It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafayPartI}.
+It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \eg the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafayPartI}.
 Consequently, research in sound perception is now strongly focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
@@
 \begin{center}
-\caption{Algorithms evaluated on the DCASE2013 ASC DCASE2013 dataset.
+\caption{Algorithms evaluated on the DCASE 2013 private dataset of acoustic scene classification.
 RQA stands for recurrence quantification analysis and ILD for inter-aural differences.

### 2016-07-12 paper/structureScene.tex
Joakim's comment on Section IV intro was addressed
minor change in Section III

@@
 \end{equation}
-The scattering transform consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
+The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
 

### 2016-07-12 paper/structureScene.tex
re-cite Hsu 2003

@@
 Let $\mathbf{S}\boldsymbol{X}(\gamma,n)$ be a dataset, where the indices $\gamma$ and $n$ respectively denotes features and examples.
-It is commonly acknowledged that SVMs should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges.
-\ja{Again, it would be really good if we could cite something for this.}
-This operation may also help making the dataset linearly separable in feature space. \ja{Why?}
+It is commonly acknowledged that SVMs should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
+
 To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we apply the affine transformation

### 2016-07-12 paper/structureScene.tex
addresses caption of Table III

@@
 Our results on the development dataset are reported in Table \ref{table:dcase2016}, in comparison with a baseline of $MFCCs$ coefficients and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
+The baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
 The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
@@
 \begin{center}
-\caption{Class-wise accuracies on the DCASE 2016 dataset for two pipelines: a baseline of mel-frequency cepstral coefficients (MFCCs), Gaussian mixture models (GMMs), and late temporal integration by product of likelihoods ; and our proposed algorithm of log-scattering coefficients, support vector machines (SVMs) with a linear kernel, and late temporal integration by majority voting.}
+\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline and proposed algorithms.}
 \begin{tabular}{lcc}

### 2016-07-12 paper/structureScene.tex
bis
MFCCs should not be between dollars

@@
 
-Our results on the development dataset are reported in Table \ref{table:dcase2016}, in comparison with a baseline of $MFCCs$ coefficients and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
+Our results on the development dataset are reported in Table \ref{table:dcase2016}, in comparison with a baseline of MFCCs coefficients and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
 The baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
@@
 \begin{center}
-\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline and proposed algorithms.}
+\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline algorithm (left) and proposed algorithm (right).}
 \begin{tabular}{lcc}

### 2016-07-12 paper/structureScene.tex
fix typo in conclusion

@@
 
-Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{ms}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine (SVM) classification.
+Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{s}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine (SVM) classification.
 %, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.

### 2016-07-12 paper/structureScene.tex
typo in VII. B

@@
 \subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task 2016 edition of the DCASE challenge, whose results are forthcoming.
+In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, whose results are forthcoming.
 This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.

### 2016-07-12 paper/structureScene.tex
add scene duration, add cluter range
add new figs for ASSR, with correct p@k; rm JA comments in EMD description

@@
 \sum\limits_{n,m}f_{nm} = \min\left( \sum\limits_{n} w_n^u ,\sum\limits_{m} w_m^v \right)
-\end{equation*} \ja{Can we provide some interpretation for the different terms and constraints here?}
+\end{equation*} 
+%\ja{Can we provide some interpretation for the different terms and constraints here?}
 where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n$th bin to ``supply the demand'' of the $m$th bin.
@@
 In the experiments reported in this paper, the DCASE2013 \cite{giannoulis2013database, 7100934} and the public part of the DCASE2016 datasets  \cite{Mesaros2016_EUSIPCO} are considered.
-To build the DCASE2013 dataset, three different recordists visited a wide variety of locations in Greater London over
-a period of several months and in each
-scene recorded a few minutes of audio. No
+
+The DCASE2013  dataset consist of 2 parts, namely public and private, each made up of 100 30sec recordings of various acoustic scenes.
+The 100 recordings are equally divided into 10 acoustic scene classes.
+To build the DCASE2013 dataset, three different recordists visited a wide variety of locations in Greater London over a period of several months and in each scene recorded a few minutes of audio.
+No
 systematic variations in the recordings covaried with scene
@@
 
-Clustering is done using KMeans$++$ \cite{arthur2007k}, an augmented version of K-means with a robust initialization procedure.
+Clustering is done using KMeans$++$ \cite{arthur2007k}, an augmented version of K-means with a robust initialization procedure. Three numbers of clusters are tested: 8, 16 and 32.
+
 \section{Results \label{sec:results}}
@@
 \begin{center}
-\includegraphics[width=8cm]{gfx/unsupervised_test3.eps}
+\includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
 \caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression.}
@@
 \begin{center}
-\includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
+\includegraphics[width=8cm]{gfx/unsupervised_test1.eps}
 \caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression.}

### 2016-07-12 paper/structureScene.tex
ASSR: write results for MFCCs

@@
 
-%For the MFCCs, there is no clear evidence that the \emph{ob} approaches improve the results. \ja{Why?}
+The same observations are made for the MFCCs for a $k\leq5$.
+However for a rank $k$ greater than 5, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
+The latter even outperform \emph{ob-a} for $k\geq5$.
+
+From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
+This is may due to the fact that, at some point, the MFCCs fail to capture the discriminant information contained in the larger-scale modulations, thus making a centroids-based description useless, or even counterproductive when the centroids are not weighted.
 

### 2016-07-12 paper/structureScene.tex
ASSR: reformulate results for MFCCs, small correction to section VI, add best result for ob-c in section VII.B

@@
 
-Given a set of $d$-dimensional feature vectors $x_l^u\in X_u$, $l=[1,2,\ldots,L]$, extracted from the scene $s_u$, $u=[1,2,\ldots,U]$, the goal of KMeans is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=[1,2,\ldots,M]$.
+Given a set of $d$-dimensional feature vectors $x_l^u\in X_u$, $l=\lbrace 1,2,\ldots,L\rbrace$, extracted from the scene $s_u$, $u=\lbrace1,2,\ldots,U\rbrace$, the goal of KMeans is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=\lbrace 1,2,\ldots,M\rbrace$.
 The partitioning is done by minimizing for each $m$ cluster the squared error between its empirical mean (centroid) and the points it contains.
@@
 
-The similarity between all the scene centroids $\mu_i$, $i=[1,2,\ldots,I]$ with $I=UM$, is computed using a radial basis function (RBF) kernel $K$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
+The similarity between all the scene centroids $\mu_i$, $i=\lbrace1,2,\ldots,I \rbrace$ with $I=UM$, is computed using a radial basis function (RBF) kernel $K$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
 \begin{equation}
@@
 From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
-This is may due to the fact that, at some point, the MFCCs fail to capture the discriminant information contained in the larger-scale modulations, thus making a centroids-based description useless, or even counterproductive when the centroids are not weighted.
+This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.
 
@@
 Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none of them managed to outperform the baseline strategy in temporal integration, namely majority voting.
+The best result was achieved by the \emph{ob-a} approach using log-compressed scattering, and was of 66\% (0.7).
 

### 2016-07-12 paper/structureScene.tex
give the duration of each recording in section III

@@
 In the context of the DCASE 2013 private dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
-\ja{That may be, but by averaging, we also reduce the variability of our data, which may be a good thing.
-If we're dealing with Gaussian processes, this is in fact a very good thing to do.}
-On the contrary, a late integration scheme for $T=372\textrm{ ms}$ would yield $128$ instances per auditory scene \ja{How long is each recording?
-Should we specify this somewhere?}, resulting in $10240$ instances overall.
+On the contrary, once the scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$ has been unpadded of any boundary artifacts, a late integration scheme for $T=372\textrm{ ms}$ would yield $128$ instances per $30$-second auditory scene, resulting in $10240$ instances overall.
 However, many of these instances may be silent or lack any salient properties of the class they are assigned to, hence an increase in bias and a risk of underfitting.

### 2016-07-12 paper/structureScene.tex
integrate Joakim's comment in IV.B
I did not mention the reduction of variability in the data, though

@@
 However, many of these instances may be silent or lack any salient properties of the class they are assigned to, hence an increase in bias and a risk of underfitting.
+Moreover, when applying late integration, the classifier optimizes accuracy at the frame level, not at the scene level, so its prediction might not yield optimal training error under the final evaluation metric.
 In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff.
 We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
-\ja{Another important difference is that early integration makes the classifier objective more appropriate, since the SVM will maximize the final classification error on the training data, as opposed to some intermediate frame-level label.}
 In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions. 

### 2016-07-12 paper/structureScene.tex
we found T=372 ms to be optimal in DCASE 2016
typo section Vii.B, ob-a => ob-c
Reintroduce commits d56dd7..b1645a

@@
 Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none of them managed to outperform the baseline strategy in temporal integration, namely majority voting.
-The best result was achieved by the \emph{ob-a} approach using log-compressed scattering, and was of 66\% (0.7).
+The best result was achieved by the \emph{ob-c} approach using log-compressed scattering, and was of 66\% (0.7).
 
@@
 The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
+Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
 The results on the evaluation dataset are awaiting publication by the DCASE organizing committee, and will be disclosed in August 2016.

### 2016-07-12 paper/structureScene.tex
reformulate Grégoire's sentence
Edit abstract & intro

@@
 This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second timescales with object-based segmentation at larger timescales.
-Application of the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates the potential of featurs designed using the wavelet scattering transform, consistently outperforming baseline mel-frequency cepstral coefficients (MFCCs).
+Applying the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates the potential of the features designed using the wavelet scattering transform, consistently outperforming baseline mel-frequency cepstral coefficients (MFCCs).
 
-Also, representing the scene at larger time scale as a set of representative features of coherent segments is an effective approach for scene similarity retrieval as it allows the retrieval system to focus on discriminative features.
+In addition, representing the scene at larger time scale as a set of representative features from coherent segments is an effective approach for scene similarity retrieval as it allows the retrieval system to focus on discriminative features.
 \end{abstract}
@@
 
-Formerly put, ecoacoustics is an interdisciplinary science that investigates natural and anthropogenic sounds and their relationship with the environment over a wide range of study scales at the intersection between ecology, acoustics and computer science.
+Formally, ecoacoustics is the interdisciplinary study of natural and anthropogenic sounds and their relationship with the environment over a wide range of scales at the intersection between ecology, acoustics and computer science.
 Despite identified needs, the field of signal-based ecoacoustics is still in infancy.
 As a consequence, few well-built datasets are readily available for evaluation purposes.
+
 A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
-
 While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of several well-designed datasets.
@@
 
-The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the wavelet scattering transform to extract discriminative features subject to a small number of parameters and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on specific areas of interest.
-\ja{Again, the last part is a little vague to me.}
+The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the wavelet scattering transform to extract discriminative features subject to a small number of parameters and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.
 
 Motivations of the proposed approach and a brief review of the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}.
-We describe the scattering transform in Section \ref{sec:scattering}, discuss post-processing transformations in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
-In Section \ref{sec:experiments} several experiments are proposed for an unsupervised setting, \ie acoustic scene similarity retrieval (ASSR) and for a supervised setting, \ie acoustic scene classification (ASC).
+We describe the scattering transform in Section \ref{sec:scattering}, discuss transformations of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
+In Section \ref{sec:experiments}, several experiments are proposed for an unsupervised setting, acoustic scene similarity retrieval (ASSR), and for a supervised setting, acoustic scene classification (ASC).
 Results are subsequently reported in Section \ref{sec:results}.

### 2016-07-12 paper/structureScene.tex
Make includegraphics format-agnostic

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gammatones.png}
+\includegraphics[width=\columnwidth]{gammatones}
 \caption{
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{compression.png}
+\includegraphics[width=\columnwidth]{compression}
 \caption{
@@
 \begin{center}
-\includegraphics[width=8cm]{gfx/unsupervised_test2.eps}
+\includegraphics[width=8cm]{gfx/unsupervised_test2}
 \caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression.}
@@
 \begin{center}
-\includegraphics[width=8cm]{gfx/unsupervised_test1.eps}
+\includegraphics[width=8cm]{gfx/unsupervised_test1}
 \caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression.}
@@
 Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none of them managed to outperform the baseline strategy in temporal integration, namely majority voting.
-The best result was achieved by the \emph{ob-c} approach using log-compressed scattering, and was of 66\% (0.7).
+Among object-based methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, and reaches a accuracy of $66\% \pm0.7$.
 

### 2016-07-12 paper/structureScene.tex
Edit background

@@
 A recent paradigm in computational bioacoustics processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
-Automated systems belonging to this paradigm aim to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time regions of particular interest requesting detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
 
 %The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-In the latter, most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Acouturier et al.
+Most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Aucouturier et al.
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
 Being a very generic method, this approach has been applied to many audio classification tasks besides ASC.
-Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated comparably to a mere averaging of the features \cite{lagrange:hal-01082501}.
+Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a mere averaging of the features \cite{lagrange:hal-01082501}.
 Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
@@
 This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}.
-Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that (perceived or measured) global sound level, are not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene.
+Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of the sound sources which compose the scene.
 It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \eg the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafayPartI}.
@@
 
-BOF approaches treats each observations equally, whereas the above cited results show that only a few events scarcely occurring may have be more important than events or texture present throughout the whole scene duration, thus advocating for a more selective approach such as the one presented in Section \ref{sec:object} for the ASSR task.
+The BOF approaches treat each observations equally, whereas the results cited above show that only a few distinct events may be more important than events or texture present throughout the whole scene duration, thus advocating for a more selective approach such as the one presented in Section \ref{sec:object} for the ASSR task.
 
@@
 An overview of the efforts for the latter task is provided by the DCASE 2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
-Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that considered the private dataset for their experiments.
-Most of scene predictions are achieved by majority voting of predictions over features.
-\ja{What about the other algorithms?
-Why are we excluding them here?}
+Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that use the private part of the dataset for their experiments.
+Most of the scene prediction results are achieved by majority voting of predictions over features.
 
@@
 & Acc. \%  &  \multicolumn{1}{c}{Design}  \\
+\hline
 \cite{olivetti2013wonder} OE & 17 & compression distance, random forests \\
@@
 \cite{patil2013multiresolution} PE & 58 & 2D Gabor filters, SVMs \\
-\cite{lee2013acoustic} NHL & 60 & unsupervised learning over Mel spectra, SVMs \\
+\cite{lee2013acoustic} NHL & 60 & unsupervised learning over mel spectra, SVMs \\
 \cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
 \cite{chum2013ieee} CHR & 65 & sparsity features, HMMs  \\
-\cite{geiger2013large} GSR & 69 & Mel-frequency Spectra and MFCCs, SVMs \\
+\cite{geiger2013large} GSR & 69 & mel spectra and MFCCs, SVMs \\
 \cite{rakotomamonjy2015histogram} RG & 69 & histogram of gradient (HOG), SVMs \\

### 2016-07-12 paper/structureScene.tex
Edit scattering section

@@
 Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filter banks interspersed with complex modulus nonlinearities.
-This section highlights the importance of invariance to time shifts and stability to time warps in the representation of acoustic scenes, and explains how the scattering transform is designed to satisfy these properties while having a high discriminative power.
+This section highlights the importance of invariance to time-shifts and stability to time-warping in the representation of acoustic scenes, and explains how the scattering transform is designed to satisfy these properties while having a high discriminative power.
 
@@
 Indeed, recordings may be shifted locally without affecting their perception and therefore such shifts do not convey any information about the class.
-To discard this superfluous source of variability, signals can be mapped to a time-shift invariant feature space before training the classifier, eliminating the need for the classifier to explicitly learn this invariance.
-From any set of descriptors $\boldsymbol{x_1}(t,\gamma)$, where $\gamma$ denotes a descriptor index, a representation invariant to time shifts shorter in duration than $T$ can be obtained by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency set to $1/T$, as measured in Hertz:
+To discard this superfluous source of variability, signals can be mapped to a time-shift invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance.
+From any time-varying set of descriptors $\boldsymbol{x_1}(t,\gamma)$, where $\gamma$ denotes a descriptor index, a representation invariant to time-shifts shorter in duration than $T$ can be obtained by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency set to $1/T$, as measured in Hertz:
 \begin{equation}
@@
 Starting from a monophonic waveform $\boldsymbol{x}(t)$, the scattering transform is defined as an infinite cascade of wavelet transform and modulus operators.
-However, to achieve invariance to translation up to $T = 372\,\mathrm{ms}$, \ie the approximately minimal duration between non-overlapping acoustic events, two layers of scattering transform.
+However, to achieve invariance to translation up to $T = 372\,\mathrm{ms}$, \ie the approximate minimal duration between non-overlapping acoustic events, two layers of scattering transform often suffice.
 
@@
 \end{equation}
-The variable $\gamma_1$ is the opposite number of a log-frequency, taking integer values between $0$ and $(J_1 \times Q_1 - 1)$.
-In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$ (close to the Nyquist frequency of the audio recordings) the number of octaves $J_1$ to $10$ (the lower end of human hearing range) and the number of wavelets per octave $Q_1$ to $8$.
+The variable $\gamma_1$ is a scale, or an inverse log-frequency, taking integer values between $0$ and $(J_1 \times Q_1 - 1)$.
+In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$ (close to the Nyquist frequency of the audio recordings), the number of octaves $J_1$ to $10$ (the lower end of human hearing range), and the number of wavelets per octave $Q_1$ to $8$.
 For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
@@
 theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the temporal resolution of the CQT, at least in its original implementation \cite{Brown1992}, is fixed by the support of the STFT analyzing window.
+frequency range, whereas the temporal resolution of the traditional CQT is fixed by the support of the STFT analyzing window \cite{Brown1992}.
 Therefore, the wavelet scalogram has a better temporal localization at high
-frequencies than the CQT, at the expense of a greater computational cost.
-This increase in time-frequency localization comes at the expense of a greater computational cost, the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
-Yet, it allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+frequencies than the CQT, at the expense of a greater computational cost
+since the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
+However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
 
@@
 Among auditory scenes, amplitude modulations may be caused by a broad variety of mechanical interactions, including collision, friction, and turbulent flow.
-At a longer scale, they also account for some higher-level attributes of sound, \eg prosody in speech or rhythm in music.
+At longer scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
 Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
@@
 \end{equation}
-where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order log-frequency index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
+where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order scale index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
 Note that these second-order wavelets are dilated versions of a second mother wavelet $\boldsymbol{\psi}$, with a different center frequency $\xi_2$ and quality factor $Q_2$.
@@
 $\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
-wavelets with one vanishing moment \cite{Venkitaraman2014}, as shown in Figure \ref{fig:gammatones}.
+wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
 In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.

### 2016-07-12 paper/structureScene.tex
Edit design section

@@
 In this section, we review three ways of achieving these properties, namely logarithmic compression, and temporal integration, and standardization.
-Whether we adopt a time-frequency perspective or a machine learning perspective on the problem of acoustic scene classification, such operations would either be qualifies as post-processing or pre-processing.
+Whether we adopt a time-frequency perspective or a machine learning perspective on the problem of acoustic scene classification, such operations would either qualify as pre- or post-processing.
+
 \subsection{Logarithmic compression}
@@
 
-Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
-Yet the distribution of the scattering coefficients is skewed towards larger values. We can reduce this skewness by applying a pointwise concave transformation, like a logarithm, to all coefficients.
-Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE2013 dataset, before and after logarithmic compression.
+Many algorithms in pattern recognition, including nearest-neighbor classifiers and SVMs, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
+Yet the distribution of the scattering coefficients is skewed towards larger values.
+We can reduce this skewness by applying a pointwise concave transformation, such as a logarithm, to all coefficients.
+Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
 
@@
 Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. 
-We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone, a source of spurious variability in scene classification.
+We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone -- a source of spurious variability in scene classification.
 Logarithmic compression can linearize this dependency, facilitating the construction of a powerful invariant at the classifier stage.
@@
 
-A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene, by retaining their average values only.
+A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene by only storing their average values.
 Going back to the definition of the scattering transform given in Section \ref{sec:scattering}, this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
-With a slight abuse of notation, we denote by
+With a slight abuse of notation, we denote the summarized features by
 \begin{equation}
 \mathbf{S}\boldsymbol{x}(\gamma) =
-\int_{-\infty}^{+\infty} \mathbf{S}\boldsymbol{x}(t,\gamma)\;\mathrm{d}t
+\int_{-\infty}^{+\infty} \mathbf{S}\boldsymbol{x}(t,\gamma)\;\mathrm{d}t\mbox{.}
 \end{equation}
-the summarized features.
 
@@
 In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.
-%\ja{We should perhaps mention this in that particular section instead of over here.}
+
 \subsection{Standardization}

### 2016-07-12 paper/structureScene.tex
done up to background

@@
 
-In addition, representing the scene at larger time scale as a set of representative features from coherent segments is an effective approach for scene similarity retrieval as it allows the retrieval system to focus on discriminative features.
+In addition, representing the scene at larger time scale as a set of representative features from coherent segments is an effective approach for acoustic scene similarity retrieval as it allows the retrieval system to focus on discriminative features.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-acoustic scene classification, acoustic scene segmentation, scattering, wavelets, earth mover's distance
+acoustic scene classification, acoustic scene segmentation, scattering, wavelets
 \end{IEEEkeywords}
@@
 A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
-While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of several well-designed datasets.
+While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of well-designed datasets.
 
@@
 %The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-Most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Aucouturier et al.
+To do so, a popular approach for acoustic scene modeling is the bag-of-frames (BOF) approach introduced by Aucouturier et al.
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
@@
 The BOF approaches treat each observations equally, whereas the results cited above show that only a few distinct events may be more important than events or texture present throughout the whole scene duration, thus advocating for a more selective approach such as the one presented in Section \ref{sec:object} for the ASSR task.
+Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC task.
 
-%\section{Background} \label{sec:soa}
-
-Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC task.
 An overview of the efforts for the latter task is provided by the DCASE 2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
-Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that use the private part of the dataset for their experiments.
-Most of the scene prediction results are achieved by majority voting of predictions over features.
+Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that use the private part of the DCASE 2013 dataset for their experiments.
+Most of the scene prediction results are achieved by majority voting of predictions over local features.
 
@@
 Algorithms above the horizontal line participated to the DCASE challenge.
-Others used the datasets afterward and the last used only the public dataset ($^*$).
+Others used the datasets afterward and the last considered only the public dataset ($^*$).
 \label{tab:dcase}}

### 2016-07-12 paper/structureScene.tex
Edit objects section

@@
 As the detection of events is still an open problem \cite{7100934},  we consider in this paper a simple quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
-The clustering is done using the centroid-based clustering algorithm KMeans.
+The clustering is done using the centroid-based $k$-means clustering algorithm.
 
-Given a set of $d$-dimensional feature vectors $x_l^u\in X_u$, $l=\lbrace 1,2,\ldots,L\rbrace$, extracted from the scene $s_u$, $u=\lbrace1,2,\ldots,U\rbrace$, the goal of KMeans is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=\lbrace 1,2,\ldots,M\rbrace$.
-The partitioning is done by minimizing for each $m$ cluster the squared error between its empirical mean (centroid) and the points it contains.
-Given $\mu_m^u$ the centroid of the cluster $c_m^u$, KMeans minimizes the following objective function:
-
+Given a set of $d$-dimensional feature vectors $x_l^u\in X_u$, $l=\lbrace 1,2,\ldots,L\rbrace$, extracted from the scene $s_u$, $u=\lbrace 1,2,\ldots,U\rbrace$, the goal of $k$-means is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=\lbrace 1,2,\ldots,M\rbrace$.
+The partitioning is done by minimizing for each cluster the squared error between its empirical mean (centroid) and the contained points.
+Given $\mu_m^u$ the centroid of the cluster $c_m^u$, $k$-means attempts to minimize the following objective function:
 \begin{equation}
-J(C_u)=\sum\limits_{m} \sum_{x^u_l\in c^u_m} \vert \vert x_l^u - \mu_m^u \vert \vert^2
+J(C_u)=\sum\limits_{m} \sum_{x^u_l\in c^u_m} \Vert x_l^u - \mu_m^u \Vert^2\mbox{.}
 \end{equation}
-
-At the end, each scene $s_u$ is described by a set of clusters $C_u$.
+Each scene $s_u$ is then described by a set of clusters $C_u$.
 One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
@@
 
-Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on their centroids similarities.
+Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroid.
 
@@
 \end{equation} 
-
-where $\mu_{k,i}$ and $\mu_{k,j}$ are the $k^{\textrm{th}}$ nearest neighbor to the centroids $\mu_i$ and $\mu_j$ respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+where $\mu_{k,i}$ and $\mu_{k,j}$ are the $k^{\textrm{th}}$ nearest neighbor to the centroids $\mu_i$ and $\mu_j$, respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
 
 To compute the similarity between two scenes, we then consider several centroid-based similarity metrics:
-
 \begin{itemize}
 \item \emph{ob-closest} (ob-c): the similarity between two scenes is equal to the largest similarity between their centroids,
-\item \emph{ob-averaged} (ob-a): the similarity between two scenes is equal to the average of their centroids similarities, and
+\item \emph{ob-averaged} (ob-a): the similarity between two scenes is equal to the average of their centroid similarities, and
 \item \emph{ob-weighted} (ob-w): for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
 Each scene $s_u$ is thus described by a signature $p_u$ of $M$ clusters ($p_u=\lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace$), with $\mu_m^u$ and $w_m^u$ being the centroid and the weight of the $m$th cluster respectively.
-A cross-bin histogram distance, taking into account both the clusters weights of the scenes and the distances between their cluster centroids is then used.
+The similarity between scenes is then given by a cross-bin histogram distance, taking into account both the cluster weights and the distances between their centroids.
 \end{itemize}
@@
 For \emph{ob-w}, the cross-bin histogram distance used is a variant of the earth mover's distance ($\EMD$), known as $\widehat{\EMD}$ and introduced in \cite{pele2008linear}.
-$\EMD$ computes the distance between two histograms by finding the minimal cost to be paid to transform one histogram into the other.
+The $\EMD$ computes the distance between two histograms by finding the minimal cost to be paid to transform one histogram into the other.
 An advantage of $\EMD$ is that it automatically aligns the histogram bins by using a ``ground distance'' which is the distance between the bin centers in the two histograms.
@@
 To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
-Given two signatures $p_u$ and $p_v$ made of $n$ and $m$ clusters respectively, the $\widehat{EMD}$ is computed by solving the following linear programming problem:
-
+Given two signatures $p_u$ and $p_v$ made of $n$ and $m$ clusters respectively, the $\widehat{\EMD}$ is computed by solving the following linear programming problem:
 \begin{equation}
@@
 \begin{equation*}
-D^{uv}=1-K^{(uv)}
+D^{uv}=\boldsymbol{1}-K^{(uv)}
 \end{equation*}
-with $K^{(uv)}$ being the slice of $K$ containing the pairwise similarities between the centroids sets $\mu^u$ and $\mu^v$ of the scenes $s_u$ and $s_v$ respectively.
+with $K^{(uv)}$ being the slice of $K$ containing the pairwise similarities between the centroids sets $\mu^u$ and $\mu^v$ of the scenes $s_u$ and $s_v$, respectively.
 As suggested in \cite{pele2009fast}, we set the tradeoff parameter $\alpha$ to $1$.
 
-To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support} is computed:
+To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ \cite{chapelle1999support,jing2003support} is computed:
 \begin{equation}
@@
 with $A$ a scaling parameter, set to the mean value of the $\widehat{\EMD}$ between all the scenes.
-The resulting kernel $K^s$ is known as an EMD kernel, and it should be noted that there is no guarantee that this kernel is positive definite.
+The resulting kernel $K^s$ is known as an $\EMD$ kernel, and it should be noted that there is no guarantee that this kernel is positive definite.
 

### 2016-07-12 paper/structureScene.tex
done up to conclusion

--- 2016-07-12 paper/structureScene.tex
@@
 The latter even outperform \emph{ob-a} for $k\geq5$.
-
 From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
@@
 Among object-based methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, and reaches a accuracy of $66\% \pm0.7$.
-
-\ml{explain why ?}
+This might be due to the fact that the discriminability gained by the object-based approach is already gained by the classifiers during the training and the fact that considering the object-base approach resort to considering an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size, see discussion in Section \ref{sec:eili}.
 

### 2016-07-12 paper/structureScene.tex
done with first final version !!

@@
 
-Representing a scene using a reduced set of representatives, with adapted similarity strategies allows us to improve ASSR performance compared to the BOF approaches based on GMMs.
+Representing a scene using a reduced set of representatives, with adapted similarity strategies allows us to improve ASSR performance compared to the BOF approaches based on GMMs which have been demonstrated to be equivalent to early integrations scheme on well balanced datasets \cite{lagrange:hal-01082501}.
 These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- are represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates.
@@
 
-This assertion shall be studied further in future work by considering data-driven convolutional networks, larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot, lafayPartI} as well as other tasks in ecoacoustics.
+Those outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot, lafayPartI} as well as emerging tasks in ecoacoustics.
 

### 2016-07-12 paper/structureScene.tex
well balanced -> well-balanced

@@
 
-Representing a scene using a reduced set of representatives, with adapted similarity strategies allows us to improve ASSR performance compared to the BOF approaches based on GMMs which have been demonstrated to be equivalent to early integrations scheme on well balanced datasets \cite{lagrange:hal-01082501}.
+Representing a scene using a reduced set of representatives, with adapted similarity strategies allows us to improve ASSR performance compared to the BOF approaches based on GMMs which have been demonstrated to be equivalent to early integrations scheme on well-balanced datasets \cite{lagrange:hal-01082501}.
 These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- are represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates.

### 2016-07-12 paper/structureScene.tex
Edit results section

--- 2016-07-12 paper/structureScene.tex
@@
 
-In the experiments reported in this paper, the DCASE2013 \cite{giannoulis2013database, 7100934} and the public part of the DCASE2016 datasets  \cite{Mesaros2016_EUSIPCO} are considered.
+In the experiments reported in this paper, we consider the DCASE 2013 \cite{giannoulis2013database, 7100934} and the public part of the DCASE 2016 datasets  \cite{Mesaros2016_EUSIPCO}.
 
-The DCASE2013  dataset consist of 2 parts, namely public and private, each made up of 100 30sec recordings of various acoustic scenes.
-The 100 recordings are equally divided into 10 acoustic scene classes.
-To build the DCASE2013 dataset, three different recordists visited a wide variety of locations in Greater London over a period of several months and in each scene recorded a few minutes of audio.
+The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes.
+The $100$ recordings are evenly divided into $10$ acoustic scene classes.
+To build the DCASE 2013 dataset, three different recordists visited a wide variety of locations in Greater London over a period of several months and in each scene recorded a few minutes of audio.
 No
@@
 
-In addition, it is still a challenging dataset, with a state-of-the-art system based on feature design and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieves  between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
+In addition, it is still a challenging dataset, with a state-of-the-art system based on feature design and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieves between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
 
@@
 
-The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme.
+The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a five-fold cross validation scheme.
 The folds used in this paper are the same as the ones used during the challenge.
@@
 For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$.
-Indeed, $3$ seconds are discarded at the beginning and at the end of the scene so as to avoid boundary artifacts.
-Experiments are conducted with and without the logarithmic compression (see Section \ref{sec:logcomp}).
+Indeed, $3$ seconds are discarded at the beginning and end of the scene to avoid boundary artifacts.
+Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}).
 
@@
 This setting gives us $600$ feature vectors per scene.
-To get a more robust representation of lower dimensionality a pooling step is performed on the MFCCs: each recording is divided into $n$ non-overlapping 250 seconds long ``texture'' windows \cite{1021072}, achieved a comparable number of vectors than while using the scattering ones.
+To get a more robust representation, a pooling step is performed on the MFCCs: each recording is divided into $250$-second long non-overlapping ``texture'' windows \cite{1021072}, over which the feature vectors are averaged.
+This yields number of vectors comparable to that of the scattering transform.
 
@@
 
-For ASSR, evaluation is performed on the private part of the DCASE2013 dataset.
+For ASSR, evaluation is performed on the private part of the DCASE 2013 dataset.
 The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
-The $p@k$ is computed for $k=\{1,\ldots,9\}$, since each class only has $10$ items.
-Note that a $p@1$ is equivalent to the classification accuracy obtained by choosing the label of the closest neighbor for a given item.
-For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameter $N$ of the RBF kernels (see Eq.
+The $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
+Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooeses the label of the closest neighbor for a given item.
+For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $M$ for the \emph{ob} approaches and the scaling parameter $k$ of the RBF kernels (see Eq.
 \ref{eq:kc}) leading to the best $p@9$.
@@
 
-For ASC, the systems are again evaluated on the private part of DCASE2013, but here with five-fold stratified cross validation.
+For ASC, the systems are again evaluated on the private part of DCASE 2013, but here with five-fold stratified cross validation.
 The folds are identical to those used in the original challenge.
-Optimal parameters (see above) are this time learned on the public part of the DCASE2013 dataset, also using five-fold stratified cross validation.
+Optimal parameters (see above) are this time learned on the public part of the DCASE 2013 dataset, also using five-fold stratified cross validation.
 
@@
 For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}.
-For both the \emph{early} and \emph{late} approaches, a linear kernel as well as a RBF kernel are used.
-As in the RBF kernels used in the \emph{ob} approaches, the RBF kernel in the SVM is scaled locally using the method proposed in \cite{selfTuneManor2004}.
+For both the \emph{early} and \emph{late} approaches, a linear kernel as well as an RBF kernel are used.
+As with the RBF kernels used in the \emph{ob} approaches, the RBF kernel in the SVM is scaled locally using the method proposed in \cite{selfTuneManor2004}.
 All feature vectors, be they frames, texture windows or centroids, are standardized prior to be given to the classifier.
@@
 
-Clustering is done using KMeans$++$ \cite{arthur2007k}, an augmented version of K-means with a robust initialization procedure. Three numbers of clusters are tested: 8, 16 and 32.
+Clustering is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure. Three numbers of clusters are tested: $8$, $16$ and $32$.
 
@@
 Overall, log-compressed scattering coefficients systematically outperform MFCCs.
-This is to be expected since the scattering coefficients capture larger-scale modulations as opposed to the MFCCs which only describe the short-time spectral envelope.
+This is to be expected since the scattering coefficients capture larger-scale modulations, as opposed to MFCCs which only describe the short-time spectral envelope.
 
@@
 
-For the scattering transform, \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
-However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{early}, which shows that the discriminant information is destroyed by averaging the contributions from all centroids.
+For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
+However, it is worth noticing that \emph{ob-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
 To take advantage of an object-based representation, we need to select certain representative centroids when comparing objects.
-Furthermore, it appears that \emph{ob-c} is be able to better characterize the classes than \emph{ob-w}.
+Furthermore, it appears that \emph{ob-c} is better able to characterize the classes than \emph{ob-w}.
 This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
@@
 The same observations are made for the MFCCs for a $k\leq5$.
-However for a rank $k$ greater than 5, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
-The latter even outperform \emph{ob-a} for $k\geq5$.
+However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
+The latter even outperforms \emph{ob-a} for $k\geq5$.
+
 From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
@@
 Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
-This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to segregate distinct acoustic events beyond the scale of about one second.
+This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to separately characterize distinct acoustic events beyond the scale of about one second.
 Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
-Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none of them managed to outperform the baseline strategy in temporal integration, namely majority voting.
-Among object-based methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, and reaches a accuracy of $66\% \pm0.7$.
-This might be due to the fact that the discriminability gained by the object-based approach is already gained by the classifiers during the training and the fact that considering the object-base approach resort to considering an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size, see discussion in Section \ref{sec:eili}.
+Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in temporal integration, namely majority voting.
+Among object-based methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$.
+
+%\ml{explain why ?}
 
@@
 Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
-This transformation improves the discriminative power of support vector machines (SVM) trained on scattering coefficients, whatever be the chosen temporal integration strategy (early or late) or the inner product kernel (linear or Gaussian RBF).
+This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
 Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.
@@
 \subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, whose results are forthcoming.
+In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
 This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
@@
 
-Our results on the development dataset are reported in Table \ref{table:dcase2016}, in comparison with a baseline of MFCCs coefficients and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
-The baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
-The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
+Our results on the public part of the DCASE 2016 dataset are reported in Table \ref{table:dcase2016}, compared with a baseline using MFCCs and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
+This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
+The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016}.
 Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
-The results on the evaluation dataset are awaiting publication by the DCASE organizing committee, and will be disclosed in August 2016.
 

### 2016-07-12 paper/structureScene.tex
Edit conclusion

@@
 Representing a scene using a reduced set of representatives, with adapted similarity strategies allows us to improve ASSR performance compared to the BOF approaches based on GMMs.
-These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- are represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates.
+These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- is represented using a convolutional network such as the scattering transform, while larger time scales -- \ie over a second -- are modeled as adaptive aggregates.
 
-Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{s}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine (SVM) classification.
+Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{s}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine classification.
 %, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.

### 2016-07-12 paper/structureScene.tex
Move last table

@@
 
-The same observations are made for the MFCCs for a $k\leq5$.
+The same observations are made for the MFCCs for $k\leq5$.
 However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
 The latter even outperforms \emph{ob-a} for $k\geq5$.
-
 From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
@@
 Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
-Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in temporal integration, namely majority voting.
+
+Surprisingly, although we ran a large number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in temporal integration, namely majority voting.
 Among object-based methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$.
-
-%\ml{explain why ?}
+This might be due to the fact that the discriminability gained by the object-based approach is already gained by the classifiers during the training and the fact the object-based approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
 
@@
 
-\subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
-This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
-Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
-Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel \cite{Hsu2003}.
-Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
-
-Our results on the public part of the DCASE 2016 dataset are reported in Table \ref{table:dcase2016}, compared with a baseline using MFCCs and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
-This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
-The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016}.
-Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
-
 \begin{table}
@@
 
+\subsubsection*{Participation to the DCASE 2016 challenge}
+In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
+This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
+Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
+Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel \cite{Hsu2003}.
+Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
+
+Our results on the public part of the DCASE 2016 dataset are reported in Table \ref{table:dcase2016}, compared with a baseline using MFCCs and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
+This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
+The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016}.
+Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
+
 \section{Conclusion}
@@
 
-Representing a scene using a reduced set of representatives, with adapted similarity strategies allows us to improve ASSR performance compared to the BOF approaches based on GMMs.
-These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- is represented using a convolutional network such as the scattering transform, while larger time scales -- \ie over a second -- are modeled as adaptive aggregates.
+Representing a scene using a reduced set of representatives, with adapted similarity strategies, allows us to improve ASSR performance compared to the BOF approaches based on GMMs which have been demonstrated to be equivalent to early integrations scheme on well-balanced datasets \cite{lagrange:hal-01082501}.
+These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- is represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates.
 

### 2016-07-12 paper/structureScene.tex
mitigating floating figures issue
fix figs
review beginning of section VI

@@
 \newcommand{\ml}[1]{\textcolor{blue}{ Mathieu : #1}}
+
+%\renewcommand{\baselinestretch}{2}
 
@@
 %Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
-
-In the experiments reported in this paper, we consider the DCASE 2013 \cite{giannoulis2013database, 7100934} and the public part of the DCASE 2016 datasets  \cite{Mesaros2016_EUSIPCO}.
+The experiments in this paper are carried out on the private part of the DCASE 2013 dataset\cite{giannoulis2013database, 7100934} and the public part of the DCASE 2016 dataset \cite{Mesaros2016_EUSIPCO}.
 
@@
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-As a consequence, DCASE2013 dataset has interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+As a consequence, DCASE2013 dataset enjoys an interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
 
-In addition, it is still a challenging dataset, with a state-of-the-art system based on feature design and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieves between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
+In addition, it is still a challenging dataset, with a state-of-the-art system based on feature design and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
 
@@
 
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
+\caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression.}
+\label{fig:ASS_1}
+\end{center}
+\end{figure}
+
 For ASC, the systems are again evaluated on the private part of DCASE 2013, but here with five-fold stratified cross validation.
@@
 \section{Results \label{sec:results}}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=8cm]{gfx/unsupervised_test2}
-\caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression.}
-\label{fig:ASS_1}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=8cm]{gfx/unsupervised_test1}
-\caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression.}
-\label{fig:ASS_2}
-\end{center}
-\end{figure}
 
@@
 It can be seen that the logarithmic compression strongly improves the results, especially for the \emph{ob} approaches, which perform worse than \emph{early} without the logarithmic compression.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=\columnwidth]{gfx/unsupervised_test1}
+\caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression.}
+\label{fig:ASS_2}
+\end{center}
+\end{figure}
 

### 2016-07-12 paper/structureScene.tex
last change

--- 2016-07-12 paper/structureScene.tex
@@
 
-The same observations are made for the MFCCs for $k\leq5$.
-However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
+%However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
 The latter even outperforms \emph{ob-a} for $k\geq5$.

### 2016-07-12 paper/structureScene.tex
minor changes to abstract

@@
 This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second timescales with object-based segmentation at larger timescales.
-Applying the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates the potential of the features designed using the wavelet scattering transform, consistently outperforming baseline mel-frequency cepstral coefficients (MFCCs).
+Applying the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates the potential of the features designed using the wavelet scattering transform, consistently outperforming a baseline of mel-frequency cepstral coefficients (MFCCs).
 
-In addition, representing the scene at larger time scale as a set of representative features from coherent segments is an effective approach for acoustic scene similarity retrieval as it allows the retrieval system to focus on discriminative features.
+In addition, representing the scene at larger time scale as a set of representative features from coherent segments is an effective approach for acoustic scene similarity retrieval, because it allows the retrieval system to focus on discriminative features.
 \end{abstract}

### 2016-07-12 paper/structureScene.tex
minor changes to section II

@@
 A recent paradigm in computational bioacoustics processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
-Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time intervals of particular interest befor detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
 
@@
 Consequently, research in sound perception is now strongly focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
-Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes.
+Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so called markers, suffice to reliably predict many high-level attributes.
 

### 2016-07-12 paper/structureScene.tex
minor changes to section V

@@
 
-Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroid.
+Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroids.
 
@@
 To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
-Given two signatures $p_u$ and $p_v$ made of $n$ and $m$ clusters respectively, the $\widehat{\EMD}$ is computed by solving the following linear programming problem:
+Given two signatures $p_u$ and $p_v$ made of $n$ and $m$ clusters respectively, the $\widehat{\EMD}$ is computed by solving the following linear program:
 \begin{equation}
@@
 \widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
-&+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \alpha \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace
+&+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \alpha \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
 \end{split}
@@
 %\ja{Can we provide some interpretation for the different terms and constraints here?}
-where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n$th bin to ``supply the demand'' of the $m$th bin.
+where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin.
 We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$.

### 2016-07-12 paper/structureScene.tex
self-contained captions for Fig. 3 and Fig. 4

@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
-\caption{ASSR: $p@k$ obtained for MFCCs and scattering with logarithmic compression.}
+\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
 \label{fig:ASS_1}
@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test1}
-\caption{ASSR: $p@k$ obtained for scattering with and without logarithmic compression.}
+\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.}
 \label{fig:ASS_2}

### 2016-07-12 paper/structureScene.tex
punctuation at the end of footnote

@@
 This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
-The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016}.
+The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.}.
 Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.

### 2016-07-12 paper/structureScene.tex
fixing figs sizes
fixing floats, grr

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
+\includegraphics[width=.9\columnwidth]{gfx/unsupervised_test2}
 \caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/unsupervised_test1}
+\includegraphics[width=.9\columnwidth]{gfx/unsupervised_test1}
 \caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.}
@@
 \subsection{Acoustic Scene Classification}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{folds.png}
+\caption{Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies. 
+The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
+The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}}
+\end{center}
+\label{fig:folds}
+\end{figure}
+
 We report experiments on acoustic scene classification (ASC), following the original methodology of the DCASE 2013 challenge, and extend our approach to the its 2016 edition.
@@
 Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{folds.png}
-\caption{Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies. 
-The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
-The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}}
-\end{center}
-\label{fig:folds}
-\end{figure}
 

### 2016-07-12 paper/structureScene.tex
small change to mfcc interpretation (ASSR)

@@
 
+The same observations are made for the MFCCs for $k\leq5$.
+However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently.
+This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
+
 %However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.

### 2016-11-25 paper/responseToReviewers.tex
addition of response to reviewers

\documentclass[10pt]{article}

\title{Response to Reviewers: \\ T-ASL-05872-2016 "Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering"}

\usepackage{color}
\newcommand{\gl}[1]{\textcolor{red}{Gregoire : #1}}
\newcommand{\eb}[1]{\textcolor{magenta}{Emmanouil : #1}}
\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu : #1}}

\begin{document}

\maketitle

We would like to thank the editor and the reviewers for their comments and suggestions. Following these comments, we made several changes to the article, which are summarized here. The next sections list our answers to each of the reviewer’s comments, with references to the revised manuscript where appropriate.

\section{Responses to the Associate Editor}

\begin{enumerate}

\item \emph{Line 40, alpha may appear in previous equations}

\item \emph{The argumentation toward the log compression is well conducted, and meaningful in this issue. The authors may simply swap the order of fig 3 and 4}

\item \emph{The class accuracy Table III for Home and Residential area are singular. Authors shall discuss that.}

\item \emph{Complete legend fig 5.}

\end{enumerate}

\section{Responses to Reviewer 1}

\begin{enumerate}

\item \emph{Object-based is inappropriate, This paper introduces: no this paper uses the scattering network to try to tackle a specific sound classification task.}

\item \emph{The second paragraph of the abstract is vague.}

\item \emph{No novel technique is presented}

\item \emph{the whole paper architecture is poor}

\item \emph{all the justifications arise from psychoacoustic, bioacoustic,... when the challenge is to classify ambient non natural stationary sounds. For example the whole Gammatone wavelet justification over Morlet is biased}

\item \emph{the classification accuracy is not suited for comparison since the dimension of the feature vectors are not the same for all the techniques }

\item \emph{most importantly the used classifier is not the same (from unsupervised GMM to supervised SVM)}

\item \emph{the paper reads like a recipe for the specific task without justifications, insights or proofs of the different given definitions and statements.}

\item \emph{The dataset is not well presented since the classes seem to represent stationary noises such as car/tram/train}

\item \emph{"recordings may be shifted locally 
without affecting their perception and therefore such shifts do not convey any information about the class" this is false. Rather say that application of time shift should not change the representation since it doesn't change the class belonging in this context and thus imposing time shift invariance will improve robustness.}

\item \emph{the used descriptors are only locally time shift invariant and not globally which is never pointed out}

\item \emph{The function phi is hardly considered as a low-pass since a low-pass is only used in coordination with a high-pass for quadratic mirror filters. Here phi is a scaling function.}

\item \emph{Line 34: "wavelet transform modulus" is not clear at all.}

\item \emph{It is then explained that the features are stable to time warping yet the justification is for pitch variations which are frequency-warping}

\item \emph{The scattering network is defined with an infinite cascade of transform yet it is enough to compute two layers, why?}

\item \emph{It is never specified what are the signals frequency sampling which we have to deduce to be 44100 Hz from line 9 page 3 "20kHz close to the Nyquist frequency of the audio recordings".}

\item \emph{"wavelet scalogram" is used along all the paper but scalogram is only defined for wavelets since 
it is a scale-o-gram.}

\item \emph{line 60: auditory and modulation filter – nonsense.}

\item \emph{"unsupervised classification" it is clustering then.}

\item \emph{"time-frequency perspective or a machine learning perspective" is it imposed versus learned representation? }

\item \emph{Equations 9 and 10 are superfluous.}

\end{enumerate}

\section{Responses to Reviewer 2}

\begin{enumerate}

\item \emph{The late versus early integration of the features is well presented and evaluated. But a graphic representation of the whole process would arise the novelty of the model.}

\item \emph{Authors may precise  how the Gammatone wavelets allow
to reproduce the asymmetry of temporal masking and the asymmetry of spectral masking.}

\item \emph{Setting of Q1 shall be motivated.}

\item \emph{Reproducibility:  data sets are publicly available, the author do not mention if they plan to make their model available.}

\item \emph{In the last experience (Table III) GMM(MFCC) may be compared to SVM(MFCC) (in addition to SVM(logscat)). There is somewhere a high dimension effect that shall be clarified.}

\item \emph{Confusion matrix of the two best methods will be interesting.}

\item \emph{The cross-validation analyses confirmed that the value T = 372 ms for the support of the low-pass filter is optimal when applying late integration by majority voting, but shall it adapted per classes  ?}

\item \emph{A table with the 2016 CASE private results may be added (or replace Tab III).}

\item \emph{The bibliography is huge, it may be reduced by 10 %. Check typos in the references (Museum, Histoire nat...). 
“befor”
Bregman shall be cited in the discussion about  binding of acoustic cues.}

\end{enumerate}

\end{document}

### 2016-12-01 paper/structureScene.tex
typo: befor -> before

@@
 A recent paradigm in computational bioacoustics processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
-Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time intervals of particular interest befor detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time intervals of particular interest before detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
 

### 2016-12-01 paper/structureScene.tex
typo: rather -> rather than

@@
 Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
-Therefore, it is arguably more beneficial to focus on the large-scale temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.
+Therefore, it is arguably more beneficial to focus on the large-scale temporal evolution of auditory scenes rather than on the joint probability distribution of short-term features.
 

### 2016-12-01 paper/responseToReviewers.tex
R1-21: Equations 9 and 10 are superfluous

@@
 Following these comments, we made several changes to the article, which are summarized here.
-The next sections list our answers to each of the reviewer’s comments, with references to the revised manuscript where appropriate.
+The next sections list our answers to each of the reviewer's comments, with references to the revised manuscript where appropriate.
 
@@
 \item \emph{Equations 9 and 10 are superfluous.}
+These equations were removed from the paper.
 

### 2016-12-01 paper/structureScene.tex
R1-21: Equations 9 and 10 are superfluous

@@
 
-To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we apply the affine transformation
-\begin{equation}
-\widetilde{\mathbf{S}}\boldsymbol{X}(\gamma, n) =
-\dfrac{ \mathbf{S}\boldsymbol{X}(\gamma, n) -
-\mu[ \mathbf{S}\boldsymbol{X}](\gamma)}{\sigma[ \mathbf{S}\boldsymbol{X}](\gamma)}
-\end{equation}
-where $\mu[ \mathbf{S}\boldsymbol{X}](\gamma) = \frac{1}{N} \sum_{n=1}^{N} \mathbf{S}\boldsymbol{X}(\gamma,n)$ is the sample mean of $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and
-\begin{equation}
-\sigma[\mathbf{S}\boldsymbol{X}] (\gamma) =
-\sqrt{\frac{1}{N-1} \sum_{n=1}^{N}
-\left( \mathbf{S}\boldsymbol{X}(\gamma,n) - \mu[\mathbf{S}\boldsymbol{X}] \right)^2}
-\end{equation}
-is the sample standard deviation.
-
+To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we substract the sample means $\mu[\mathbf{S}\boldsymbol{X}(\gamma)]$ to $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and divide the result by the sample standard deviations $\sigma[\mathbf{S}\boldsymbol{X}] (\gamma)$.
 The vectors $\mu[\mathbf{S}\boldsymbol{X}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{X}](\gamma)$ are estimated from the training set only, and the same affine transformation is subsequently applied to all samples in the training set and the test set.

### 2016-12-01 paper/structureScene.tex
new title
Auditory Scene Similarity Retrieval and Classification With Wavelet
Scattering
bis

@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering}
+\title{Auditory Scene Similarity Retrieval and Classification with Wavelet Scattering}
 
@@
 \subsection{Wavelet scalogram}
-Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int x(t) \exp(\mathrm{i} 2\pi t \omega) \; \mathrm{d}\omega$.
+Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int x(t) \exp(- \mathrm{i} 2\pi \omega t) \; \mathrm{d}t$.
 Note that the frequency variable $\omega$ is expressed in Hertz, not in radians per second.

### 2016-12-06 paper/structureScene.tex
relevance based quantization is now in the manuscript

@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Auditory Scene Similarity Retrieval and Classification with Wavelet Scattering}
+\title{Auditory Scene Similarity Retrieval and Classification with Relevance based Quantization of Scattering Features}
 
@@
 
-This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second timescales with object-based segmentation at larger timescales.
+This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second timescales with relevance based quantization at larger timescales.
 Applying the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates the potential of the features designed using the wavelet scattering transform, consistently outperforming a baseline of mel-frequency cepstral coefficients (MFCCs).
 
-In addition, representing the scene at larger time scale as a set of representative features from coherent segments is an effective approach for acoustic scene similarity retrieval, because it allows the retrieval system to focus on discriminative features.
+In addition, representing the scene at larger time scale as a set of representative features from coherent segments is an effective approach for acoustic scene similarity retrieval, as it is found that it allows the retrieval system to focus on discriminative features.
 \end{abstract}
@@
 
-The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the wavelet scattering transform to extract discriminative features subject to a small number of parameters and 2) object-based modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.
+The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the wavelet scattering transform to extract discriminative features subject to a small number of parameters and 2) large time scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to hopefully focus the scene description on a few salient sound objects that carry most of the semantic information.
 
@@
 
-\section{Object-based modeling of the scene}
+\section{Relevance based quantization of features}
 \label{sec:object}
@@
 
-\subsubsection*{Object-based \vs early integration}
+\subsubsection*{Relevance-based quantization \vs early integration}
 
-For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes.
+For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization to improve the similarity measures between the scenes.
 However, it is worth noticing that \emph{ob-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
-To take advantage of an object-based representation, we need to select certain representative centroids when comparing objects.
+To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.
 Furthermore, it appears that \emph{ob-c} is better able to characterize the classes than \emph{ob-w}.
@@
 
-Surprisingly, although we ran a large number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in temporal integration, namely majority voting.
-Among object-based methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$.
-This might be due to the fact that the discriminability gained by the object-based approach is already gained by the classifiers during the training and the fact the object-based approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
+Surprisingly, although we ran a large number of experiments to formulate the relevance-based quantization scheme into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in temporal integration, namely majority voting.
+Among such methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$.
+This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already gained by the classifiers during the training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
 

### 2016-12-06 experiments/sceneSvm/scsvReport.m

@@ -6,7 +6,7 @@ function config = scsvReport(config)
 % Copyright: Mathieu Lagrange
 % Date: 29-Jun-2016
 
-if nargin==0, sceneSvm('report', 'rcv'); return; end
+if nargin==0, sceneSvm('report', 'rh'); return; end
 
 %% label
 
@@ -15,24 +15,24 @@ setVal={'train','test'};
 %% get best setting
 
 set=1;
-config = expExpose(config, 't','fontSize','scriptsize','step', 1, 'mask', {0 0 0 set 0 0},'obs',1,'precision', 2);
-
+% config = expExpose(config, 't','fontSize','scriptsize','step', 1, 'mask', {0 0 1 set 0 0},'obs',1,'precision', 2);
+% 
 cut=1;
 select=2;
 gcaStyle='set(gca,''xtick'',1:6,''xticklabel'',[])';
 
-config = expExpose(config, 'p','fontSize','scriptsize','step', 1, 'mask', {0 0 cut set 0 select},'obs',1,'precision', 2, ...
-    'plotAxisProperties',{'fontsize',16},...
-    'plotCommand',{'legend(''mfcc, objBased, closest'',''mfcc, objBased, EMD'',''Scat, objBased, closest'',''Scat, objBased, EMD'',''mfcc, baseline'',''Scat, baseline'')', ...
-    'ylabel(''accuracy'')','ylim([.3 1])',gcaStyle,['title(''MFCC vs. Scattering using SVM on ' setVal{set} ''')']});
+% config = expExpose(config, 'p','fontSize','scriptsize','step', 1, 'mask', {0 0 cut set 0 select},'obs',1,'precision', 2, ...
+%     'plotAxisProperties',{'fontsize',16},...
+%     'plotCommand',{'legend(''mfcc, objBased, closest'',''mfcc, objBased, EMD'',''Scat, objBased, closest'',''Scat, objBased, EMD'',''mfcc, baseline'',''Scat, baseline'')', ...
+%     'ylabel(''accuracy'')','ylim([.3 1])',gcaStyle,['title(''MFCC vs. Scattering using SVM on ' setVal{set} ''')']});
 
 %% run test set with best setting
 set=2;
 
-config = expExpose(config, 'p','fontSize','scriptsize','step', 1, 'mask', {0 0 cut set 0 select},'obs',1,'precision', 2, ...
-    'plotAxisProperties',{'fontsize',16},...
[…]

### 2016-12-06 paper/responseToReviewers.tex
improving bib

@@
 \usepackage{color}
-\newcommand{\gl}[1]{\textcolor{red}{Gregoire : #1}}
-\newcommand{\eb}[1]{\textcolor{magenta}{Emmanouil : #1}}
+\newcommand{\vl}[1]{\textcolor{red}{Vincent : #1}}
+\newcommand{\ja}[1]{\textcolor{magenta}{Joakim : #1}}
 \newcommand{\ml}[1]{\textcolor{blue}{ Mathieu : #1}}
@@
 
-\item \emph{The bibliography is huge, it may be reduced by 10 %. Check typos in the references (Museum, Histoire nat...). 
-“befor”
-Bregman shall be cited in the discussion about  binding of acoustic cues.}
+\item \emph{The bibliography is huge, it may be reduced by 10 \%.}
+
+\item \emph{Check typos in the references (Museum, Histoire nat...). “befor”}
+
+\item \emph{Bregman shall be cited in the discussion about  binding of acoustic cues.}
+
+\ja{can you spot where we speak oabout binding acoustic cues ?, I can't ;(}
 

### 2016-12-06 paper/structureScene.tex
improving bib

@@
 \IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{ECOACOUSTICS2014, krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
+In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
 
-In addition, recent work demonstrates the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{lafayPartI, guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
+In addition, recent work demonstrates the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
 Because they bear a strong societal impact and raise many scientific challenges, we believe that these case studies are of considerable interest for the signal processing community.
@@
 A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
-While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, maffiolo_caracterisation_1999, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of well-designed datasets.
+While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of well-designed datasets.
 
@@
 Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of the sound sources which compose the scene.
-It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \eg the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafayPartI}.
-Consequently, research in sound perception is now strongly focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
-Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so called markers, suffice to reliably predict many high-level attributes.
+It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \eg the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators.
+Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
+Although the complete set of events occurring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so called markers, suffice to reliably predict many high-level attributes.
 
@@
 \cite{li2013auditory} LTT & 72 & MFCCs, tree baggers \\
-\cite{roma2013recurrence} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
+\cite{roma2013} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
 \hline
@@
 
-Those outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot, lafayPartI} as well as emerging tasks in ecoacoustics.
+Those outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as emerging tasks in ecoacoustics.
 

### 2016-12-06 paper/responseToReviewers.tex
done with biblio

@@
 
+The sentence have been rewritten to be more explicit.
+
 \item \emph{"time-frequency perspective or a machine learning perspective" is it imposed versus learned representation? }
@@
 
+Experiments have been performed in the Matlab environment using open source libraries an all the authors are committed to comply to the reproducible research statements.
+Unfortunately, due to time constraints, we will not be able to deliver the code at the time of submission of the revision.
+Nevertheless, it will be fully available at the time of publication if the revision is accepted for publication.
+
 \item \emph{In the last experience (Table III) GMM(MFCC) may be compared to SVM(MFCC) (in addition to SVM(logscat)). There is somewhere a high dimension effect that shall be clarified.}
@@
 
+The bibliography section is now more tight.
+
 \item \emph{Check typos in the references (Museum, Histoire nat...). “befor”}
+
+The references have been carefully curated.
 
@@
 
-\ja{can you spot where we speak oabout binding acoustic cues ?, I can't ;(}
+\ja{can you spot where we speak about binding acoustic cues ?, I can't ;(}
 

### 2016-12-06 paper/structureScene.tex
done with biblio

@@
 \IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13a, stowell13b}.
+In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
 
@@
 A recent paradigm in computational bioacoustics processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
-Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time intervals of particular interest before detailed human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.
+Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time intervals of particular interest before detailed human inspection \cite{rosenstock2002landbird}.
 
@@
 
-The BOF approaches treat each observations equally, whereas the results cited above show that only a few distinct events may be more important than events or texture present throughout the whole scene duration, thus advocating for a more selective approach such as the one presented in Section \ref{sec:object} for the ASSR task.
+The BOF approaches treat each observations equally, whereas the results cited above show that only a few distinct events may be more far important than events or texture present throughout the whole scene duration, thus advocating for a more selective approach such as the one presented in Section \ref{sec:object} for the ASSR task.
 Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC task.
 
-An overview of the efforts for the latter task is provided by the DCASE 2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
+An overview of the efforts for the latter task is provided by the IEEE ASSP detection and classification of acoustic scenes and events (DCASE) 2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
 Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that use the private part of the DCASE 2013 dataset for their experiments.
@@
 \caption{Algorithms evaluated on the DCASE 2013 private dataset of acoustic scene classification.
-RQA stands for recurrence quantification analysis and ILD for inter-aural differences.
+RQA stands for recurrence quantification analysis, ILD for inter-aural differences and SVMs for support vector machines.
 Algorithms above the horizontal line participated to the DCASE challenge.
@@
 theoretical limit of optimal time-frequency localization across the whole
-frequency range, whereas the temporal resolution of the traditional CQT is fixed by the support of the STFT analyzing window \cite{Brown1992}.
+frequency range, whereas the temporal resolution of the traditional CQT is fixed by the support of the STFT analyzing window. %\cite{Brown1992}
 Therefore, the wavelet scalogram has a better temporal localization at high
@@
 
-Before supervised or unsupervised classification, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power.
+Before classification or clustering, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power.
 In this section, we review three ways of achieving these properties, namely logarithmic compression, and temporal integration, and standardization.
@@
 
-Many algorithms in pattern recognition, including nearest-neighbor classifiers and SVMs, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
+Many algorithms in pattern recognition, including nearest-neighbor classifiers and support vector machines (SVM)s, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
 Yet the distribution of the scattering coefficients is skewed towards larger values.
@@
 As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of an object-based representation of auditory scenes for predicting high-level properties.
-As the detection of events is still an open problem \cite{7100934},  we consider in this paper a simple quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
+As the detection of events is still an open problem \cite{7100934},  we consider in this paper a quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
 The clustering is done using the centroid-based $k$-means clustering algorithm.
@@
 
-To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ \cite{chapelle1999support,jing2003support} is computed:
+To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ is computed: %\cite{chapelle1999support,jing2003support}
 \begin{equation}
@@
 %Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
-The experiments in this paper are carried out on the private part of the DCASE 2013 dataset\cite{giannoulis2013database, 7100934} and the public part of the DCASE 2016 dataset \cite{Mesaros2016_EUSIPCO}.
+The experiments in this paper are carried out on the private part of the DCASE 2013 dataset\cite{7100934} and the public part of the DCASE 2016 dataset \cite{Mesaros2016_EUSIPCO}.
 
@@
 
-Clustering is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure. Three numbers of clusters are tested: $8$, $16$ and $32$.
+Clustering is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure.
+Three numbers of clusters are tested: $8$, $16$ and $32$.
+The implementation of tte presented methods as well as the one of the experimental protocolis available\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}}.
 
@@
 
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in Paris, France, in 1990.
+He received in 2011 the B.S.
+degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, France, and the B.S.
+degree in Musicology from the Sorbonne University, Paris, France.
+He received his M.S.
+degree in acoustics, signal processing and musical informatics (ATIAM) from the University Pierre and Marie Curie and the IRCAM laboratory, Paris, France, in 2013.
+Since 2013, he is a Ph.D student at Irccyn, a French laboratory dedicated to cybernetics.
+His current research interests include acoustic scene similarity and classification, as well as acoustic scene perception.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at Irccyn, a French laboratory dedicated to cybernetics.
+He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions, in Canada (University of Victoria, McGill University) and in France (Orange Labs, Telecom ParisTech, Ircam).
+His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.
+\end{IEEEbiography}
+
 % biography section

### 2016-12-06 paper/structureScene.tex
moving to monochrome. Sad but true.

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gammatones}
+\includegraphics[width=\columnwidth]{bw/gammatones}
 \caption{
@@
 Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$. 
+\vl{If you want to keep this level of detail you have to adapt the figure so that it is readable in greyscale.
 Blue and red oscillations represent the real and imaginary parts.
 The orange envelope represents the complex modulus.}
+}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{compression}
+\includegraphics[width=\columnwidth]{bw/compression}
 \caption{
@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{gfx/unsupervised_test2}
+\includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
 \caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{gfx/unsupervised_test1}
+\includegraphics[width=.9\columnwidth]{bw/unsupervised_test1}
 \caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{folds.png}
+\includegraphics[width=\columnwidth]{bw/folds.png}
 \caption{Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.
+Performance on each fold is depicted using a given symbol.
 The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
-The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}}
+The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.}
 \end{center}

### 2016-12-12 paper/structureScene.tex
Edit abstract

@@
 
-This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second timescales with relevance based quantization at larger timescales.
-Applying the proposed approach to the acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets using a state-of-the-art classifier demonstrates the potential of the features designed using the wavelet scattering transform, consistently outperforming a baseline of mel-frequency cepstral coefficients (MFCCs).
+This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second time-scales with relevance-based quantization at larger time-scales.
+Instead of representing a scene using summary statistics, a method popularized by bag-of-features models, each scene is represented by a subset of its most representative features, allowing the retrieval system to focus on particular discriminative features in the scene.
+We demonstrate the potential of the proposed approach for both acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets, consistently outperforming baseline mel-frequency cepstral coefficient (MFCC) methods.
 
-In addition, representing the scene at larger time scale as a set of representative features from coherent segments is an effective approach for acoustic scene similarity retrieval, as it is found that it allows the retrieval system to focus on discriminative features.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-acoustic scene classification, acoustic scene segmentation, scattering, wavelets
+acoustic scene classification, acoustic scene segmentation, acoustic scene similarity retrieval, scattering, wavelets, invariance
 \end{IEEEkeywords}

### 2016-12-12 paper/responseToReviewers.tex
moving to monochrome. Sad but true.
Insert responses regarding title & abstract

@@
 
+The legend have been completed to help the readability of the figure.
+
 \end{enumerate}
@@
 \item \emph{Object-based is inappropriate, This paper introduces: no this paper uses the scattering network to try to tackle a specific sound classification task.}
+The title has been changed to remove the "object-based" formulation.
+The particular approach introduced in the paper is the hierarchical representation of a scene as a partition of its distribution of scattering features.
+While the scattering transform, which has been introduced by previous works, forms the low-level layer of this representation, the hierarchical structure is novel and we believe generic enough to represent auditory scenes for a variety of information retrieval tasks.
 
 \item \emph{The second paragraph of the abstract is vague.}
+This paragraph has been reworded and worked into the rest of the abstract in a more fluid manner.
 
 \item \emph{No novel technique is presented}
+Again, the hierarchical representation proposed is novel and we provide adequate evidence, in our opinion, for its power in the proposed similarity retrieval task.
 
 \item \emph{the whole paper architecture is poor}
+Without any specific points of criticism, it is hard to address this particular statement.
 
@@
 
-\item \emph{Authors may precise  how the Gammatone wavelets allow
-to reproduce the asymmetry of temporal masking and the asymmetry of spectral masking.}
+\item \emph{Authors may precise  how the Gammatone wavelets allow to reproduce the asymmetry of temporal masking and the asymmetry of spectral masking.}
 

### 2016-12-12 paper/structureScene.tex
Edit intro

@@
 This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second time-scales with relevance-based quantization at larger time-scales.
-Instead of representing a scene using summary statistics, a method popularized by bag-of-features models, each scene is represented by a subset of its most representative features, allowing the retrieval system to focus on particular discriminative features in the scene.
+Instead of representing a scene using summary statistics, a method popularized by bag-of-features models, each scene is represented by a subset of its most representative features, allowing the retrieval system to focus on discriminative features.
 We demonstrate the potential of the proposed approach for both acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets, consistently outperforming baseline mel-frequency cepstral coefficient (MFCC) methods.
@@
 In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
+In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
+Because they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.
 
-In addition, recent work demonstrates the interest of acoustic monitoring for the characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound} as well as the prediction of annoyance due to traffic \cite{gloaguen}.
-Because they bear a strong societal impact and raise many scientific challenges, we believe that these case studies are of considerable interest for the signal processing community.
-
-Formally, ecoacoustics is the interdisciplinary study of natural and anthropogenic sounds and their relationship with the environment over a wide range of scales at the intersection between ecology, acoustics and computer science.
-Despite identified needs, the field of signal-based ecoacoustics is still in infancy.
-As a consequence, few well-built datasets are readily available for evaluation purposes.
+Ecoacoustics is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science.
+Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
+As a consequence, few well-designed datasets are readily available for evaluation purposes.
 
 A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
-While narrower in its range of applications, it has the advantage of being rooted in cognitive psychology of categorization \cite{dubois2006cognitive, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of well-designed datasets.
+While narrower in its range of applications, it has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of well-designed datasets.
 
 Despite the differences, ecoacoustics and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?
-
-The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) using the wavelet scattering transform to extract discriminative features subject to a small number of parameters and 2) large time scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to hopefully focus the scene description on a few salient sound objects that carry most of the semantic information.
+The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) application of the wavelet scattering transform to extract discriminative features and 2) large time-scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.
+\ja{Technically, the 1st concept has already been demonstrated in previous papers.
+Do we really want to claim this here?}
 
 Motivations of the proposed approach and a brief review of the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}.
-We describe the scattering transform in Section \ref{sec:scattering}, discuss transformations of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
-In Section \ref{sec:experiments}, several experiments are proposed for an unsupervised setting, acoustic scene similarity retrieval (ASSR), and for a supervised setting, acoustic scene classification (ASC).
+We describe the scattering transform in Section \ref{sec:scattering}, discuss processing of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
+In Section \ref{sec:experiments}, several experiments are proposed for the unsupervised setting of acoustic scene similarity retrieval (ASSR) and for the supervised setting of acoustic scene classification (ASC).
 Results are subsequently reported in Section \ref{sec:results}.

### 2016-12-12 paper/structureScene.tex
Edit background

@@
 
-A recent paradigm in computational bioacoustics processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording.
-Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as to detect time intervals of particular interest before detailed human inspection \cite{rosenstock2002landbird}.
+A recent paradigm in computational bioacoustics processes the audio stream in a holistic way over large time-scales, without assuming that a single species is present throughout the recording.
+\ja{Should we define somewhere what ``computational bioacoustics'' is?} Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as marking time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.
 
 %The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-To do so, a popular approach for acoustic scene modeling is the bag-of-frames (BOF) approach introduced by Aucouturier et al.
+For these tasks, a popular approach is the bag-of-frames (BOF) approach for acoustic scene modeling introduced by Aucouturier et al.
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
 Being a very generic method, this approach has been applied to many audio classification tasks besides ASC.
-Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a mere averaging of the features \cite{lagrange:hal-01082501}.
-Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
-Therefore, it is arguably more beneficial to focus on the large-scale temporal evolution of auditory scenes rather than on the joint probability distribution of short-term features.
+Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
+Indeed, the typical morphology of acoustic scenes is a ``skeleton of events on a bed of textures'' \cite{nelken2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
+Therefore, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on the joint probability distribution of short-term features.
 
 This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}.
-Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of the sound sources which compose the scene.
-It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \eg the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators.
-Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.
-Although the complete set of events occurring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so called markers, suffice to reliably predict many high-level attributes.
+For example, studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene.
+It has been shown that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes.
+For example, the presence of birds is very likely to be heard in parks in urban areas and so are strong pleasantness indicators.
+Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments \cite{ricciardi2015sound,lavandier2006contribution}.
+Although the complete set of events occurring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes.
 
-The BOF approaches treat each observations equally, whereas the results cited above show that only a few distinct events may be more far important than events or texture present throughout the whole scene duration, thus advocating for a more selective approach such as the one presented in Section \ref{sec:object} for the ASSR task.
+From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation equally when computing summary statistics.
+A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes.
+
 Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC task.
-
-An overview of the efforts for the latter task is provided by the IEEE ASSP detection and classification of acoustic scenes and events (DCASE) 2013 challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
+An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
 Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that use the private part of the DCASE 2013 dataset for their experiments.
@@
 \caption{Algorithms evaluated on the DCASE 2013 private dataset of acoustic scene classification.
-RQA stands for recurrence quantification analysis, ILD for inter-aural differences and SVMs for support vector machines.
 Algorithms above the horizontal line participated to the DCASE challenge.
-Others used the datasets afterward and the last considered only the public dataset ($^*$).
-\label{tab:dcase}}
+The algorithm marked with $^*$ considered only the public dataset.
+\label{tab:dcase} \ja{What are the initials OE, ELF, etc.
+for?
+Do we need them?}}
 \begin{tabular}{lcl}

### 2016-12-12 paper/structureScene.tex
Edit scattering section

@@
 
-Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filter banks interspersed with complex modulus nonlinearities.
-This section highlights the importance of invariance to time-shifts and stability to time-warping in the representation of acoustic scenes, and explains how the scattering transform is designed to satisfy these properties while having a high discriminative power.
+Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filter banks alternated with complex modulus non-linearities.
+This section first explains the importance of local invariance to time-shifting and stability to time-warping in the representation of acoustic scenes, then describes how the scattering transform is designed to satisfy these properties while having a high discriminative power.
 
 \subsection{Invariance and stability in acoustic scenes}
-The notion of invariance to time shifts plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
-Indeed, recordings may be shifted locally without affecting their perception and therefore such shifts do not convey any information about the class.
-To discard this superfluous source of variability, signals can be mapped to a time-shift invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance.
-From any time-varying set of descriptors $\boldsymbol{x_1}(t,\gamma)$, where $\gamma$ denotes a descriptor index, a representation invariant to time-shifts shorter in duration than $T$ can be obtained by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency set to $1/T$, as measured in Hertz:
+The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
+Indeed, recordings may be shifted locally without affecting their original class memberships.
+To discard this superfluous source of variability, signals can be mapped to a time-shifting invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance and increasing the robustness of its predictions.
+From any time-varying set of descriptors $\boldsymbol{x_1}(t,\gamma)$, where $\gamma$ denotes a descriptor index, a representation invariant to time-shifts shorter than $T$ can be obtained by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency set to $1/T$, which gives
 \begin{equation}
-\mathbf{S_1}\boldsymbol{x}(t, \gamma) = (\boldsymbol{x_1} \ast \boldsymbol{\phi}) (t).
+\mathbf{S_1}\boldsymbol{x}(t, \gamma) = (\boldsymbol{x_1}(\cdot,\gamma) \ast \boldsymbol{\phi}) (t).
 \end{equation}
 A downside is that the transient information in $\boldsymbol{x_1}$ at scales finer than $T$ are lost by this low-pass filtering, reducing discriminability in feature space.
-To address this issue, the scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$ and then applying complex modulus.
+To address this issue, the scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$ and then applying a complex modulus.
+\ja{Why are we doing this for a generic set of descriptors $\boldsymbol{x_1}(t, \gamma)$ here?
+Seems like we're going forward, then backwards in the next section.
+In addition, many results will not hold unless $\boldsymbol{x_1}$ is a scalogram.
+We do specify here that higher-order coefficients have to be obtained with wavelets, so why not $\boldsymbol{x_1}$?}
 
 By resorting to wavelet transform modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to time-warping deformation,
-in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}.
-In addition to invariance, this stability property is crucial to signal classification, since it guarantees robustness to small variations in pitch, reverberation, and rhythmic organization of events, which make up an important part of the intra-class variability among natural sounds.
+in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}. \ja{This will not be true if $\boldsymbol{x_1}$ is an STFT, for example.}
+This stability property is crucial to signal classification, since it guarantees robustness to small variations in pitch, reverberation, and rhythmic organization of events, which can all be described as small time-warping transformations.
+These make up an important part of the intra-class variability among natural sounds, so a representation that has some insensitivity to them has an advantage.
 
-Starting from a monophonic waveform $\boldsymbol{x}(t)$, the scattering transform is defined as an infinite cascade of wavelet transform and modulus operators.
-However, to achieve invariance to translation up to $T = 372\,\mathrm{ms}$, \ie the approximate minimal duration between non-overlapping acoustic events, two layers of scattering transform often suffice.
-
-The next subsection describes the operations involved in the scattering transform, and in particular the construction of wavelet filter banks.
+Starting from a monophonic waveform $\boldsymbol{x}(t)$, the scattering transform is defined as an infinite cascade of wavelet transforms alternated with modulus operators.
+However, to achieve invariance to translation up to $T \approx 500\,\mathrm{ms}$, (\ie the approximate minimal duration between non-overlapping acoustic events in our signals), two layers of scattering transform often suffice.
+Higher layers of scattering coefficients will represent larger-scale structure that is better captured by the cluster-based representation presented below.
 
 \subsection{Wavelet scalogram}
-Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int x(t) \exp(- \mathrm{i} 2\pi \omega t) \; \mathrm{d}t$.
-Note that the frequency variable $\omega$ is expressed in Hertz, not in radians per second.
-
+Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int_{-\infty}^{+\infty} x(t) \exp(- \mathrm{i} 2\pi \omega t) \; \mathrm{d}t$.
 Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
-center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$.
+center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$, where $Q_1$ is the quality factor of the filter.
 A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
@@
 \end{equation}
-The variable $\gamma_1$ is a scale, or an inverse log-frequency, taking integer values between $0$ and $(J_1 \times Q_1 - 1)$.
-In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$ (close to the Nyquist frequency of the audio recordings), the number of octaves $J_1$ to $10$ (the lower end of human hearing range), and the number of wavelets per octave $Q_1$ to $8$.
+The variable $\gamma_1$ is a scale (an inverse log-frequency) taking integer values between $0$ and $(J_1 \times Q_1 - 1)$, where $J_1$ is the number of octaves spanned by the filter bank.
 For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
-has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$, a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$, and a quality factor of $Q_1$.
+has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$ and a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$ resulting in the same quality factor $Q_1$ as $\boldsymbol{\psi}$.
+In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$, $J_1$ to $10$, and the quality factor $Q_1$, which doubles as the number of wavelets per octave, to $8$.
+This results in the wavelet filters covering the whole range of human hearing, from $20~\mathrm{Hz}$ to $20~\mathrm{kHz}$.
 
 The wavelet transform of an audio signal
-$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets.
+$\boldsymbol{x}(t)$ is obtained by convolution with all wavelet filters
 Applying pointwise complex modulus the transform yields
@@
 \end{equation}
-The wavelet scalogram bears resemblance to the constant-Q transform (CQT),
+The scalogram bears resemblance to the constant-Q transform (CQT),
 which is derived from the short-term Fourier transform (STFT) by averaging the frequency
@@
 frequencies than the CQT, at the expense of a greater computational cost
-since the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
-However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to the minimum scale $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+since the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank. \ja{Is the cost really greater?}
+However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
+\ja{Where does the formula $2Q_1/\xi_1$ come from?}
 
 \subsection{Extracting modulations with second-order scattering}
-Among auditory scenes, amplitude modulations may be caused by a broad variety of mechanical interactions, including collision, friction, and turbulent flow.
-At longer scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
+In auditory scenes, amplitude modulations may be caused by a broad variety of rapid mechanical interactions, including collision, friction, turbulent flow, and so on.
+At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
 Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
-The amplitude modulation spectrum resulting from this operator is
+
+We define second-order wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ in the same way as the first-order wavelets, but with parameters $\xi_2$, $J_2$, and $Q_2$ and so have center frequencies $2^{-\gamma_2/Q_2}\xi_2$ for $\gamma_2$ taking values between $0$ and $(J_2 \times Q_2 - 1)$.
+While this abuses notation slightly, the identity of the wavelets should be clear from context.
+The amplitude modulation spectrum resulting from a wavelet modulus decomposition using these second-order wavelets is then
 \begin{equation}
 \boldsymbol{x_2}(t,\gamma_1,\gamma_2) =
-\vert \boldsymbol{x_1} \ast \boldsymbol{\psi_{\gamma_2}} \vert(t,\gamma_1),
+\vert \boldsymbol{x_1} \ast \boldsymbol{\psi_{\gamma_2}} \vert(t,\gamma_1).
 \end{equation}
-where the center frequencies of the wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ are of the form $2^{-\gamma_2/Q_2} \xi_2$, and the second-order scale index $\gamma_2$ takes integer values between $0$ and $(J_2 \times Q_2 - 1)$.
-Note that these second-order wavelets are dilated versions of a second mother wavelet $\boldsymbol{\psi}$, with a different center frequency $\xi_2$ and quality factor $Q_2$.
-The identity of the wavelet will be clear from context.
 In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to time-shifting, which yields
+\ja{Why is $\xi_2 \neq \xi_1$?} Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to time-shifting, which yields
 \begin{equation}
 \mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
-(\boldsymbol{x_2} \ast \phi)(t,\gamma_1,\gamma_2).
+(\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \phi)(t).
 \end{equation}
+
 The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
+While higher-order scattering coefficients can be calculated, for the purposes of our current, work, the first and second order are sufficient.
 
@@
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
-As such, they can be discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
-This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparse time-frequency representation of acoustic scenes.
+As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
+This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.
 

### 2016-12-12 paper/responseToReviewers.tex
Add some more responses on the scattering section

@@
 Rather say that application of time shift should not change the representation since it doesn't change the class belonging in this context and thus imposing time shift invariance will improve robustness.}
+The reviewer is correct that this is not true as stated. Our meaning is indeed what he/she describes and the text has been changed to reflect this.
 
 \item \emph{the used descriptors are only locally time shift invariant and not globally which is never pointed out}
+As formulated, the scattering transforms do indeed have only local time-shifting invariance, which has been clarified.
+However, the processing corresponding to ``early'' integrating achieves more invariance through complete averaging of the scattering transform along time and so does have that global time-shifting invariance.
 
 \item \emph{The function phi is hardly considered as a low-pass since a low-pass is only used in coordination with a high-pass for quadratic mirror filters. Here phi is a scaling function.}
+The function $\phi$ is very much a low-pass filter with a frequency support of approximately $1/T$ and a time bandwidth of $T$, where $T$ is of the order of hundreds of milliseconds.
+It does not correspond to the traditional scaling function used in quadratic mirror filters.
 
 \item \emph{Line 34: "wavelet transform modulus" is not clear at all.}
+We have updated the text to clarify this expression.
 
 \item \emph{It is then explained that the features are stable to time warping yet the justification is for pitch variations which are frequency-warping}
+The justification of pitch variation is indeed a frequency-warping, but can be written as a time-warping.
+Indeed, a signal that is slightly compressed or dilated locally will have its pitch transposed up or down.
 
 \item \emph{The scattering network is defined with an infinite cascade of transform yet it is enough to compute two layers, why?}
+For the time scales considered in this paper, the first two orders are sufficient to describe the sound events present in the signals. This has been made clearer in the text.
 
@@
 it is a scale-o-gram.}
+This is true, but for readers less familiar with the terminology, we have opted to use this more redundant expression to express the fact that these derive from wavelet decompositions.
 

### 2016-12-12 paper/structureScene.tex
More edits to scattering section
Actually make some of the changes we promised reviewers.

@@
 Starting from a monophonic waveform $\boldsymbol{x}(t)$, the scattering transform is defined as an infinite cascade of wavelet transforms alternated with modulus operators.
-However, to achieve invariance to translation up to $T \approx 500\,\mathrm{ms}$, (\ie the approximate minimal duration between non-overlapping acoustic events in our signals), two layers of scattering transform often suffice.
+However, to achieve local invariance to translation up to $T \approx 500\,\mathrm{ms}$, (\ie the approximate minimal duration between non-overlapping acoustic events in our signals), two layers of scattering transform often suffice.
 Higher layers of scattering coefficients will represent larger-scale structure that is better captured by the cluster-based representation presented below.
@@
 At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
-Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform modulus operator.
+Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a local time-shifting invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform followed by a pointwise modulus operator.
 
@@
 In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-\ja{Why is $\xi_2 \neq \xi_1$?} Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee invariance to time-shifting, which yields
+\ja{Why is $\xi_2 \neq \xi_1$?} Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee local invariance to time-shifting, which yields
 \begin{equation}

### 2016-12-12 paper/structureScene.tex
Edit feature design section

@@
 
-Before classification or clustering, it is beneficial to process scattering coefficients with feature transformations that improve invariance, normality, or generalization power.
-In this section, we review three ways of achieving these properties, namely logarithmic compression, and temporal integration, and standardization.
-Whether we adopt a time-frequency perspective or a machine learning perspective on the problem of acoustic scene classification, such operations would either qualify as pre- or post-processing.
+Before classification or clustering, it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
+In this section, we review three transformations which achieve these properties, namely logarithmic compression, temporal integration, and standardization.
 
@@
 
-Many algorithms in pattern recognition, including nearest-neighbor classifiers and support vector machines (SVM)s, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
+Many algorithms in pattern recognition, including nearest-neighbor classifiers and support vector machines (SVMs), tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
 Yet the distribution of the scattering coefficients is skewed towards larger values.
-We can reduce this skewness by applying a pointwise concave transformation, such as a logarithm, to all coefficients.
+We can reduce this skewness by applying a pointwise concave transformation to all coefficients.
+In particular, we find that the logarithm performs particularly well in this respect.
 Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
@@
 Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure. 
-We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone -- a source of spurious variability in scene classification.
-Logarithmic compression can linearize this dependency, facilitating the construction of a powerful invariant at the classifier stage.
+We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone--a source of spurious variability in scene classification.
+Logarithmic compression linearizes this dependency, facilitating the construction of powerful invariants at the classifier stage.
 
@@
 As such, a non-linear renormalization becomes a linear transformation, which can be learned by a linearly discriminative classifier.
+\ja{Do we do this here or not?
+What is the point of this paragraph?}
 
@@
 to disambiguate two classes of auditory scenes with overlapping semantic content, \eg a train from a subway station or a quiet street from a park.
-Depending on whether aggregation is performed in feature space or in decision space, the corresponding method is referred to as early or late integration.
+Depending on whether aggregation is performed in feature space or in decision space (the output of a classifier) the corresponding method is referred to as early or late integration.
 
-A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene by only storing their average values.
-Going back to the definition of the scattering transform given in Section \ref{sec:scattering}, this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
+A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene by computing their average values over time.
+In the definition of the scattering transform (see Section \ref{sec:scattering}), this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
 With a slight abuse of notation, we denote the summarized features by
@@
 
-Conversely, a late integration scheme relies on probabilistic assignments $\mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right]$ over short-term windows of length $T$, which are subsequently aggregated to produce a final decision
+Conversely, a late integration scheme relies on probabilistic assignments $\mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right]$ over short-term windows of length $T$ obtained from some supervised model, such as a classifier.
+These are subsequently aggregated to produce a final decision
 \begin{equation}
@@
 where $\hat{y}$ is the estimated class label and $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.
+\ja{Do we really sum probabilities?
+Seems like a strange thing to do.}
 
 A drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene.
-In the context of the DCASE 2013 private dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, hence an increase in variance in statistical estimation and a risk of overfitting.
+In the context of the DCASE 2013 private dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, which can increase the variance of the estimated model and run the risk of overfitting.
 On the contrary, once the scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$ has been unpadded of any boundary artifacts, a late integration scheme for $T=372\textrm{ ms}$ would yield $128$ instances per $30$-second auditory scene, resulting in $10240$ instances overall.
-However, many of these instances may be silent or lack any salient properties of the class they are assigned to, hence an increase in bias and a risk of underfitting.
+However, many of these instances may be silent or lack any salient properties of the class they are assigned to, which may increase the bias in the estimated model and run the risk of underfitting.
 Moreover, when applying late integration, the classifier optimizes accuracy at the frame level, not at the scene level, so its prediction might not yield optimal training error under the final evaluation metric.
 In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff.
+The appropriateness of either approach depends therefore on the nature of the signal that is classified.
 We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
-In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority voting of the predictions.
+In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority vote of the predictions given by an SVM.
 
@@
 
-Let $\mathbf{S}\boldsymbol{X}(\gamma,n)$ be a dataset, where the indices $\gamma$ and $n$ respectively denotes features and examples.
-It is commonly acknowledged that SVMs should be trained on scaled features with null mean and unit variance, so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
-
-To standardize $\mathbf{S}\boldsymbol{X}(\gamma,n)$, we substract the sample means $\mu[\mathbf{S}\boldsymbol{X}(\gamma)]$ to $\mathbf{S}\boldsymbol{X}(\gamma,n)$ and divide the result by the sample standard deviations $\sigma[\mathbf{S}\boldsymbol{X}] (\gamma)$.
-The vectors $\mu[\mathbf{S}\boldsymbol{X}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{X}](\gamma)$ are estimated from the training set only, and the same affine transformation is subsequently applied to all samples in the training set and the test set.
+Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where the $\gamma$ and $n$ respectively denotes feature and sample indices, respectively.
+It has been found that SVMs should be trained on scaled features with null mean and unit variance so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
+To standardize $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
+The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the training set only, and the same affine transformation is then applied to all samples in both the training and the test sets.
 

### 2016-12-12 paper/structureScene.tex
Edit objects section

@@
 
-As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of an object-based representation of auditory scenes for predicting high-level properties.
+As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of object-based representations of auditory scenes for predicting high-level properties.
 As the detection of events is still an open problem \cite{7100934}, we consider in this paper a quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
-The clustering is done using the centroid-based $k$-means clustering algorithm.
+\ja{This last sentence is a bit hard to parse.
+Are we segmenting the time-series and identifying coherent regions?}
 
-Given a set of $d$-dimensional feature vectors $x_l^u\in X_u$, $l=\lbrace 1,2,\ldots,L\rbrace$, extracted from the scene $s_u$, $u=\lbrace 1,2,\ldots,U\rbrace$, the goal of $k$-means is to partition $X_u$ into $M$ clusters $c^u_m\in C_u$, $m=\lbrace 1,2,\ldots,M\rbrace$.
-The partitioning is done by minimizing for each cluster the squared error between its empirical mean (centroid) and the contained points.
-Given $\mu_m^u$ the centroid of the cluster $c_m^u$, $k$-means attempts to minimize the following objective function:
+Given a set of $d$-dimensional feature vectors $X_u = \{x_1^u, \ldots, x_L^u\}$, extracted from the scene $s_u$, where $u=\lbrace 1,2,\ldots,U\rbrace$, we would like to partition $X_u$ into a set $C_u = \{c^u_1, \ldots, c^u_M\}$ of $M$ clusters.
+The partitioning is done by minimizing squared error between the empirical mean, or centroid, of each cluster and the vectors belonging to it.
+Letting $\mu_m^u$ denote the centroid of the cluster $c_m^u$, we attempt to minimize the following objective function:
 \begin{equation}
@@
 \end{equation}
+This is known as $k$-means clustering.
 Each scene $s_u$ is then described by a set of clusters $C_u$.
@@
 
-The similarity between all the scene centroids $\mu_i$, $i=\lbrace1,2,\ldots,I \rbrace$ with $I=UM$, is computed using a radial basis function (RBF) kernel $K$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
+The similarity between the scene centroids $\mu_i$ in the entire dataset, where $i=\lbrace1,2,\ldots,I \rbrace$ for $I=U \times M$, is computed using a radial basis function (RBF) kernel $K$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
 \begin{equation}
 \label{eq:kc}
-K_{ij} = \exp\left( - \dfrac{\Vert \mu_i - \mu_j \Vert^2}{\Vert \mu_i - \mu_{k,i} \Vert \Vert \mu_j - \mu_{k,j} \Vert} \right) 
+K_{ij} = \exp\left( - \dfrac{\Vert \mu_i - \mu_j \Vert^2}{\Vert \mu_i - \mu_{k,i} \Vert \Vert \mu_j - \mu_{k,j} \Vert} \right).
 \end{equation} 
-where $\mu_{k,i}$ and $\mu_{k,j}$ are the $k^{\textrm{th}}$ nearest neighbor to the centroids $\mu_i$ and $\mu_j$, respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+Here, $\mu_{k,i}$ and $\mu_{k,j}$ are the $k^{\textrm{th}}$ nearest neighbor to the centroids $\mu_i$ and $\mu_j$, respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+\ja{Wouldn't it be clearer to define $K$ using scene indices $u$ and $v$ and feature indices $m$ and $n$?}
 
@@
 \begin{itemize}
-\item \emph{ob-closest} (ob-c): the similarity between two scenes is equal to the largest similarity between their centroids,
-\item \emph{ob-averaged} (ob-a): the similarity between two scenes is equal to the average of their centroid similarities, and
-\item \emph{ob-weighted} (ob-w): for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
-Each scene $s_u$ is thus described by a signature $p_u$ of $M$ clusters ($p_u=\lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace$), with $\mu_m^u$ and $w_m^u$ being the centroid and the weight of the $m$th cluster respectively.
-The similarity between scenes is then given by a cross-bin histogram distance, taking into account both the cluster weights and the distances between their centroids.
+\item \emph{ob-closest} (\emph{ob-c}): the similarity between two scenes is equal to the largest similarity between their centroids,
+\item \emph{ob-averaged} (\emph{ob-a}): the similarity between two scenes is equal to the average of their centroid similarities, and
+\item \emph{ob-weighted} (\emph{ob-w}): the similarity between two scenes is computed using a variant of the earth mover's distance applied to the set of centroids weighted by the number of frames assigned to its cluster.
 \end{itemize}
 
-For \emph{ob-w}, the cross-bin histogram distance used is a variant of the earth mover's distance ($\EMD$), known as $\widehat{\EMD}$ and introduced in \cite{pele2008linear}.
-The $\EMD$ computes the distance between two histograms by finding the minimal cost to be paid to transform one histogram into the other.
-An advantage of $\EMD$ is that it automatically aligns the histogram bins by using a ``ground distance'' which is the distance between the bin centers in the two histograms.
-In our case, the histograms are the clusters weights $w^u$ defined over the bins formed by clusters centroids $\mu^u$.
+For \emph{ob-w}, each centroid is weighted according to the number of frames belonging to its cluster.
+Each scene $s_u$ is thus described by a signature $p_u$ of $M$ clusters, where $p_u=\lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace$) and $\mu_m^u$ and $w_m^u$ are the centroid and the weight of the $m$th cluster, respectively.
+The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance ($\widehat{\EMD}$) introduced by \cite{pele2008linear}.
+The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts times the ``ground distance'' moved in terms of the histogram bins.
+In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the distance over these cluster centroids.
 
-The $\widehat{\EMD}$ variant of $\EMD$ is adapted for non-normalized histograms.
 To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
-Given two signatures $p_u$ and $p_v$ made of $n$ and $m$ clusters respectively, the $\widehat{\EMD}$ is computed by solving the following linear program:
+Given two signatures $p_u$ and $p_v$, the $\widehat{\EMD}$ is computed by solving the following linear program:
 \begin{equation}
@@
 %\ja{Can we provide some interpretation for the different terms and constraints here?}
-where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin.
+where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin, and $\alpha$ is a parameter which trades off between \ja{?}.
 We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$.

### 2016-12-12 paper/structureScene.tex
Edit methods section

@@
 %Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
-The experiments in this paper are carried out on the private part of the DCASE 2013 dataset\cite{7100934} and the public part of the DCASE 2016 dataset \cite{Mesaros2016_EUSIPCO}.
-
+The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} and DCASE 2016 \cite{Mesaros2016_EUSIPCO} datasets.
 The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes.
@@
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-As a consequence, DCASE2013 dataset enjoys an interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
-
-In addition, it is still a challenging dataset, with a state-of-the-art system based on feature design and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
+As a consequence, DCASE 2013 dataset enjoys an interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+In addition, it is still a challenging dataset, with a state-of-the-art system based using handcrafted features as input to SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
 
@@
 
-The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a five-fold cross validation scheme.
-The folds used in this paper are the same as the ones used during the challenge.
+The public part of the dataset can used for optimizing the ASC system and the private part used for computing the resulting accuracy using a five-fold cross validation scheme.
+For this paper, we shall perform our experiments on the private part of the dataset.
+The folds used are the same as the ones used during the challenge.
+
+The DCASE 2016 dataset has a similar organization, but here we conduct our experiments on the public part of the dataset. \ja{Any more information on this dataset?}
 
@@
 
-Experiments are carried out using scattering coefficients as well as mel-frequency cepstral coefficients (MFCCs) as features.
+Experiments are carried out using scattering coefficients as well as baseline mel-frequency cepstral coefficients (MFCCs) as features.
 For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$.
-Indeed, $3$ seconds are discarded at the beginning and end of the scene to avoid boundary artifacts.
+In this case, $3$ seconds are discarded at the beginning and end of the scene to avoid boundary artifacts.
 Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}).
@@
 MFCCs are computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$, with full frequency range.
-Different ranks of coefficients were tested to compute the MFCCs, results here are only reported for the best setting, which consisted of $40$ MFCCs, including the first coefficient related to the average energy.
-This setting gives us $600$ feature vectors per scene.
-To get a more robust representation, a pooling step is performed on the MFCCs: each recording is divided into $250$-second long non-overlapping ``texture'' windows \cite{1021072}, over which the feature vectors are averaged.
-This yields number of vectors comparable to that of the scattering transform.
+The standard configuration of $39$ coefficients coupled with an average-energy measure performs best in preliminary tests, so we use this in the following.
+The coefficients are averaged using $250\,\mathrm{ms}$ long non-overlapping windows so that each window represents structures of similar scale to the scattering coefficients.
 
@@
 
-The influence of the temporal scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (ASSR) and in a supervised setting (ASC).
+The performance of the temporal scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (ASSR) and in a supervised setting (ASC).
+\ja{The \emph{ob} approaches are not really assessed for ASC, though.}
 
@@
 The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
-The $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
-Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooeses the label of the closest neighbor for a given item.
+\ja{The variable $k$ is used previously, so we should change that.} The $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
+Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item.
 For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $M$ for the \emph{ob} approaches and the scaling parameter $k$ of the RBF kernels (see Eq.
 \ref{eq:kc}) leading to the best $p@9$.
+The \emph{ob} approaches are compared to commonly used early integration approach \emph{early}.
 
@@
 The folds are identical to those used in the original challenge.
+In the same way as the majority of the DCASE 2013 participants, we use an SVM for the ASC task, which is a large-margin binary discriminator with tunable regularization of misclassified examples.
+The slack parameter $C$ is here set to $1$.
 Optimal parameters (see above) are this time learned on the public part of the DCASE 2013 dataset, also using five-fold stratified cross validation.
-
-The \emph{ob} approaches are compared to commonly used early integration approach (\emph{early}) for ASSR, while both early and late integration (\emph{late}) strategy are evaluated for the ASC setting (see Section \ref{sec:eili}).
-
-In the same way as the majority of the DCASE2013 participants, we use an SVM for the ASC task, which is a large-margin binary discriminator with tunable regularization of misclassified examples.
-The parameter $C$ is here set to $1$.
+\ja{Which parameters are these?} For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).
 
 For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}.
-For both the \emph{early} and \emph{late} approaches, a linear kernel as well as an RBF kernel are used.
+For both the \emph{early} and \emph{late} methods, a linear kernel as well as an RBF kernel are used.
 As with the RBF kernels used in the \emph{ob} approaches, the RBF kernel in the SVM is scaled locally using the method proposed in \cite{selfTuneManor2004}.
-All feature vectors, be they frames, texture windows or centroids, are standardized prior to be given to the classifier.
+All feature vectors, that is scattering frames, MFCCs and centroids, are standardized prior to be given to the classifier.
 The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
 
-Clustering is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure.
+The clustering for the \emph{ob} methods is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure.
 Three numbers of clusters are tested: $8$, $16$ and $32$.
-The implementation of tte presented methods as well as the one of the experimental protocolis available\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}}.
+The implementation of the presented methods as well as the one of the experimental protocol is available.\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}} \ja{What is ``the one''?}
 

### 2016-12-12 paper/structureScene.tex
Edit results section

@@
 This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
+\ja{What is the ``at some point'' referring to?
+In other words, why do we expect this to happen for larger $k$ but not smaller?}
 
@@
 \includegraphics[width=.9\columnwidth]{bw/unsupervised_test1}
-\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.}
+\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.
+\ja{Why not \emph{ob-a} here?
+Also, the relationship between the different approaches for the regular scattering coefficients seems to be inverted with respect to the log-scattering and the previous figure.
+That is, here we have \emph{early} $>$ \emph{ob-w} $\approx$ \emph{ob-c} instead of \emph{ob-w} $>$ \emph{ob-w} $>$ \emph{early}.}}
 \label{fig:ASS_2}
@@
 \includegraphics[width=\columnwidth]{bw/folds.png}
-\caption{Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.
+\caption{\label{fig:folds}
+Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.
 Performance on each fold is depicted using a given symbol.
@@
 \end{center}
-\label{fig:folds}
 \end{figure}
 
-We report experiments on acoustic scene classification (ASC), following the original methodology of the DCASE 2013 challenge, and extend our approach to the its 2016 edition.
+We report experiments on acoustic scene classification (ASC), following the original methodology of the DCASE 2013 challenge, and extend our approach to its 2016 edition.
 
@@
 As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
-Not only the average accuracy is improved, but the variance between fold-wise accuracies is also reduced, \eg from $16\%$ to $7\%$ in the case of early integration.
-This is in accordance with the previous results on similarity retrieval, as well as with other papers comparing mel-frequency coefficients with scattering coefficients on a variety of tasks, ranging from musical genre recognition \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.
+Not only is the average accuracy improved, but the variance between fold-wise accuracies is also reduced, such as from $16\%$ to $7\%$ in the case of early integration. \ja{Why does this happen?}
+This is in accordance with previous results on similarity retrieval, as well as comparisons between mel-frequency coefficients and scattering coefficients for tasks ranging from musical genre recognition and phone segment classification \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.
 
@@
 This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to separately characterize distinct acoustic events beyond the scale of about one second.
+\ja{This is not suggested since the scattering transform is computed at $T=372\,\mathrm{ms}$ and so cannot be expected to capture larger-scale structures.
+Rather, this says something about the non-stationarity of the audio signals and the fact that we cannot characterize them by some set of summary statistics.}
 Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
 
-Surprisingly, although we ran a large number of experiments to formulate the relevance-based quantization scheme into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in temporal integration, namely majority voting.
+Surprisingly, although we ran a large number of experiments to formulate the relevance-based quantization scheme into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in \emph{late} temporal integration, namely majority voting.
 Among such methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$.
-This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already gained by the classifiers during the training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
+\ja{Why don't we display these results?
+Also how is the variance so much smaller here?} This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already achieved by the classifiers during training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
+\ja{This sentence is very confusing.}
 
@@
 Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
+\ja{Isn't this why we also standardize the coefficients?}
 This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
@@
 \begin{center}
-\caption{Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.}
+\caption{
+\label{table:dcase2013}
+Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.
+\ja{The relationship between linear and RBF is inverted for the \emph{early} integration scheme compared to \emph{late}. Is this correct?}
+}
 \begin{tabular}{llll}
@@
              \hline
-early, linear  & $54\pm20$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
-early, RBF     & $47\pm16$  & $56\pm\phantom{0}7$  & $63\pm\phantom{0}6$   \\
-late, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
-late, RBF     & $68\pm10$ & $73\pm\phantom{0}8$  & $\mathbf{78}\pm\phantom{0}6$   \\
+\emph{early}, linear  & $54\pm20$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
+\emph{early}, RBF     & $47\pm16$  & $56\pm\phantom{0}7$  & $63\pm\phantom{0}6$   \\
+\emph{late}, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
+\emph{late}, RBF     & $68\pm10$ & $73\pm\phantom{0}8$  & $\mathbf{78}\pm\phantom{0}6$   \\
 \end{tabular}
 \end{center}
-\label{table:dcase2013}
 \end{table}
@@
 \caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline algorithm (left) and proposed algorithm (right).}
-\begin{tabular}{lcc}
-Features & MFCCs & log-scattering \\
-Classifier  & GMMs    & linear SVMs \\
-Temporal integration & late (product rule) & late (majority voting) \\
+\begin{tabular}{lccc}
+Features & MFCCs & MFCCs & log-scattering \\
+Classifier  & GMMs & linear SVMs & linear SVMs \\
+Variant of \emph{late} integration & product & majority vote & majority vote \\
 \midrule
-beach & $74.6 \pm 18.9$ & $83.5 \pm \phantom{0}7.1$ \\
-bus & $58.2 \pm 18.3$ & $88.8 \pm 11.4$ \\
-caf\'{e} / restaurant & $85.1 \pm 10.8$ & $64.5 \pm \phantom{0}8.3$ \\
-car & $69.1 \pm 20.8$ & $94.9 \pm \phantom{0}6.1$ \\
-city center & $89.6 \pm \phantom{0}9.2$ & $91.7 \pm \phantom{0}9.4$ \\
-forest path & $72.4 \pm 23.8$ & $93.8 \pm \phantom{0}7.8$ \\
-grocery store & $74.1 \pm 13.7$ & $90.9 \pm \phantom{0}6.9$ \\
-home & $78.1 \pm 17.7$ & $56.2 \pm 23.9$ \\
-library & $65.1 \pm 23.3$ & $82.0 \pm 13.0$ \\
-metro station & $85.2 \pm 15.6$ & $96.0 \pm \phantom{0}2.3$ \\
-office & $90.8 \pm 16.0$ & $87.5 \pm 21.7$ \\
-park & $25.6 \pm 11.0$ & $75.4 \pm \phantom{0}7.4$ \\
-residential area & $75.1 \pm 19.4$ & $44.2 \pm 15.0$ \\
-train & $34.1 \pm \phantom{0}7.2$ & $58.3 \pm 10.0$ \\
-tram & $85.4 \pm 13.9$ & $83.7 \pm 12.2$ \\
+beach & $74.6 \pm 18.9$ & & $83.5 \pm \phantom{0}7.1$ \\
+bus & $58.2 \pm 18.3$ & & $88.8 \pm 11.4$ \\
+caf\'{e} / restaurant & $85.1 \pm 10.8$ & & $64.5 \pm \phantom{0}8.3$ \\
+car & $69.1 \pm 20.8$ & & $94.9 \pm \phantom{0}6.1$ \\
+city center & $89.6 \pm \phantom{0}9.2$ & & $91.7 \pm \phantom{0}9.4$ \\
+forest path & $72.4 \pm 23.8$ & & $93.8 \pm \phantom{0}7.8$ \\
+grocery store & $74.1 \pm 13.7$ & & $90.9 \pm \phantom{0}6.9$ \\
+home & $78.1 \pm 17.7$ & & $56.2 \pm 23.9$ \\
+library & $65.1 \pm 23.3$ & & $82.0 \pm 13.0$ \\
+metro station & $85.2 \pm 15.6$ & & $96.0 \pm \phantom{0}2.3$ \\
+office & $90.8 \pm 16.0$ & & $87.5 \pm 21.7$ \\
+park & $25.6 \pm 11.0$ & & $75.4 \pm \phantom{0}7.4$ \\
+residential area & $75.1 \pm 19.4$ & & $44.2 \pm 15.0$ \\
+train & $34.1 \pm \phantom{0}7.2$ & & $58.3 \pm 10.0$ \\
+tram & $85.4 \pm 13.9$ & & $83.7 \pm 12.2$ \\
 \bottomrule
-Average & $70.8 \pm \phantom{0}2.6$ & $79.4 \pm \phantom{0}3.0$
+Average & $70.8 \pm \phantom{0}2.6$ & & $79.4 \pm \phantom{0}3.0$
 \end{tabular}
@@
 \subsubsection*{Participation to the DCASE 2016 challenge}
-In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
+In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
 This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
-Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
-Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel \cite{Hsu2003}.
-Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
+Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$, we therefore use a linear kernel instead.
+Our experiments on the smaller DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
 
@@
 This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
-The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.}.
-Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
+The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$.\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.}
+Furthermore, our cross-validation analysis confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
 

### 2016-12-12 paper/structureScene.tex
Edit conclusion

@@
 
-Representing a scene using a reduced set of representatives, with adapted similarity strategies, allows us to improve ASSR performance compared to the BOF approaches based on GMMs which have been demonstrated to be equivalent to early integrations scheme on well-balanced datasets \cite{lagrange:hal-01082501}.
-These results advocate for a hierarchical model of the scene, where short-time structure -- \ie below a second -- is represented using a convolutional network such as the scattering transform, whereas larger time scales -- \ie over a second -- are modeled as adaptive aggregates.
+Representing a scene using a reduced set of representatives, with adapted similarity strategies, allows us to improve ASSR performance compared to the early-integration scheme which have been demonstrated to be equivalent to BOF approaches using GMMs on well-balanced datasets \cite{lagrange:hal-01082501}.
+These results advocate for a hierarchical model of the scene, where short-time structure--\ie below a second--is represented using a convolutional network such as the scattering transform, whereas larger time scales--\ie over a second--are modeled as adaptive aggregates.
 
-Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{s}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine classification.
+Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{s}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine classification for the tasks we considered.
 %, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.
@@
 
-Those outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as emerging tasks in ecoacoustics.
+These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in ecoacoustics.
 

### 2016-12-12 paper/responseToReviewers.tex
Add more responses

@@
 \item \emph{the whole paper architecture is poor}
-Without any specific points of criticism, it is hard to address this particular statement.
+Without any specific points of criticism, it is hard to address this particular statement. Nonetheless, we have reworked the paper as a whole, trying to clarify the overall message.
 
@@
 \item \emph{the classification accuracy is not suited for comparison since the dimension of the feature vectors are not the same for all the techniques }
+While the dimension of the features vectors are different, this does not necessarily have to impact the performance of the classification.
+For example, if we were classifiying stationary Gaussian signals, a spectral envelope feature, like the MFCCs, would prove to be just as accurate, if not better, than the richer second-order scattering transform.
+The important relationship to consider is whether the feature transform provides an accurate description of the relevant information content in the signal.
+In addition, most of our experiments deal with the feature vectors through an RBF kernel, which is equivalent to considering the features embedded in an infinite-dimensional space, so in this sense the original dimensionality of the feature vectors carries less importance.
 
 \item \emph{most importantly the used classifier is not the same (from unsupervised GMM to supervised SVM)}
+This is an important point and we have included results for MFCCs using a linear SVM.
+However, we point out that the GMM is also supervised, although it is trained using a generative objective as opposed to a discriminative one, which is the case for the SVM.
 
@@
 \item \emph{The dataset is not well presented since the classes seem to represent stationary noises such as car/tram/train}
+While these scenes definitely have a stationary component, such as for the car, tram, train, metro station, and bus classes.
+This is not sufficient to classify them, since this stationary signal is very similar for these classes.
+This is why the cluster-based hierarchical representation that we propose is necessary in that it allows the system to pick out the distinct non-stationary components that discriminate between the classes.
 
@@
 \item \emph{"time-frequency perspective or a machine learning perspective" is it imposed versus learned representation? }
+The intention was to situate the feature processing step either as post-processing of the scattering features or as pre-processing for the classifier stage. We have removed it during the revision.
 
@@
 \item \emph{In the last experience (Table III) GMM(MFCC) may be compared to SVM(MFCC) (in addition to SVM(logscat)). There is somewhere a high dimension effect that shall be clarified.}
+We have added results for SVM on MFCCs.
+As mentioned in the response to the other reviewer, we believe that the high dimensionality should not have a significant effect on these results.
+\ja{Although perhaps we should indeed talk about this in the paper.}
 

### 2016-12-13 paper/structureScene.tex
Change kernel K definition per #4

@@
 
-The similarity between the scene centroids $\mu_i$ in the entire dataset, where $i=\lbrace1,2,\ldots,I \rbrace$ for $I=U \times M$, is computed using a radial basis function (RBF) kernel $K$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
+The similarity between the scene centroids $\mu_m^u$ over the entire dataset, is computed using a radial basis function (RBF) kernel $K$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
 \begin{equation}
 \label{eq:kc}
-K_{ij} = \exp\left( - \dfrac{\Vert \mu_i - \mu_j \Vert^2}{\Vert \mu_i - \mu_{k,i} \Vert \Vert \mu_j - \mu_{k,j} \Vert} \right).
+K_{mn}^{uv} = \exp\left( - \dfrac{\Vert \mu_m^u - \mu_n^v \Vert^2}{\Vert \mu_m^u - \mu_{m,k}^u \Vert \Vert \mu_n^v - \mu_{n,k}^v \Vert} \right).
 \end{equation} 
-Here, $\mu_{k,i}$ and $\mu_{k,j}$ are the $k^{\textrm{th}}$ nearest neighbor to the centroids $\mu_i$ and $\mu_j$, respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
-\ja{Wouldn't it be clearer to define $K$ using scene indices $u$ and $v$ and feature indices $m$ and $n$?}
+Here, $\mu_{m,k}^u$ and $\mu_{n,k}^v$ are the $k^{\textrm{th}}$ nearest neighbors to the centroids $\mu_m^u$ and $\mu_n^v$, respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
 
@@
 \begin{itemize}
-\item \emph{ob-closest} (\emph{ob-c}): the similarity between two scenes is equal to the largest similarity between their centroids,
-\item \emph{ob-averaged} (\emph{ob-a}): the similarity between two scenes is equal to the average of their centroid similarities, and
+\item \emph{ob-closest} (\emph{ob-c}): the similarity between two scenes $s_u$ and $s_v$ is equal to the largest similarity between their centroids, that is
+\begin{equation}
+	\max_{m,n} K_{mn}^{uv},
+\end{equation}
+\item \emph{ob-averaged} (\emph{ob-a}): the similarity between two scenes $s_u$ and $s_v$ is equal to the average of their centroid similarities, that is
+\begin{equation}
+	\frac{1}{M^2} \sum_{m,n} K_{mn}^{uv}
+\end{equation}
+and,
 \item \emph{ob-weighted} (\emph{ob-w}): the similarity between two scenes is computed using a variant of the earth mover's distance applied to the set of centroids weighted by the number of frames assigned to its cluster.
@@
 \begin{equation*}
-D^{uv}=\boldsymbol{1}-K^{(uv)}
+D_{mn}^{uv}=1-K_{mn}^{uv}.
 \end{equation*}
-with $K^{(uv)}$ being the slice of $K$ containing the pairwise similarities between the centroids sets $\mu^u$ and $\mu^v$ of the scenes $s_u$ and $s_v$, respectively.
 As suggested in \cite{pele2009fast}, we set the tradeoff parameter $\alpha$ to $1$.

### 2016-12-13 paper/structureScene.tex
Reorganize scattering section per #3

@@
 To discard this superfluous source of variability, signals can be mapped to a time-shifting invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance and increasing the robustness of its predictions.
-From any time-varying set of descriptors $\boldsymbol{x_1}(t,\gamma)$, where $\gamma$ denotes a descriptor index, a representation invariant to time-shifts shorter than $T$ can be obtained by convolving $\boldsymbol{x_1}$ with a low-pass filter $\boldsymbol{\phi}(t)$ of cutoff frequency set to $1/T$, which gives
-\begin{equation}
-\mathbf{S_1}\boldsymbol{x}(t, \gamma) = (\boldsymbol{x_1}(\cdot,\gamma) \ast \boldsymbol{\phi}) (t).
-\end{equation}
-A downside is that the transient information in $\boldsymbol{x_1}$ at scales finer than $T$ are lost by this low-pass filtering, reducing discriminability in feature space.
-To address this issue, the scattering transform recovers this information by convolving $\boldsymbol{x_1}$ with wavelets whose center frequencies are above $1/T$ and then applying a complex modulus.
-\ja{Why are we doing this for a generic set of descriptors $\boldsymbol{x_1}(t, \gamma)$ here?
-Seems like we're going forward, then backwards in the next section.
-In addition, many results will not hold unless $\boldsymbol{x_1}$ is a scalogram.
-We do specify here that higher-order coefficients have to be obtained with wavelets, so why not $\boldsymbol{x_1}$?}
 
-By resorting to wavelet transform modulus, as opposed to Fourier transform modulus, the resulting features are provably stable to time-warping deformation,
-in the sense of Lipschitz regularity with respect to diffeomorphisms \cite{Mallat2012}. \ja{This will not be true if $\boldsymbol{x_1}$ is an STFT, for example.}
-This stability property is crucial to signal classification, since it guarantees robustness to small variations in pitch, reverberation, and rhythmic organization of events, which can all be described as small time-warping transformations.
-These make up an important part of the intra-class variability among natural sounds, so a representation that has some insensitivity to them has an advantage.
+Formally, given a signal $\boldsymbol{x}(t)$, we would like its translation $\boldsymbol{x_c}(t) = \boldsymbol{x}(t-c)$ to be mapped to the same feature vector provided that $|c| \ll T$ for some maximum duration $T$ that specifies the extent of the time-shifting invariance.
+We can also define more complicated transformations by letting $c$ vary with $t$.
+In this case, we have $\boldsymbol{x_\tau}(t) = \boldsymbol{x}(t-\tau(t))$ for some function $\tau$, which performs a time-warping of $\boldsymbol{x}(t)$ to obtain $\boldsymbol{x_\tau}(t)$.
+Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events.
+These make up an important part of intra-class variability among natural sounds, so a representation that has some insensitivity to such transformations would be very advantageous.
 
-Starting from a monophonic waveform $\boldsymbol{x}(t)$, the scattering transform is defined as an infinite cascade of wavelet transforms alternated with modulus operators.
-However, to achieve local invariance to translation up to $T \approx 500\,\mathrm{ms}$, (\ie the approximate minimal duration between non-overlapping acoustic events in our signals), two layers of scattering transform often suffice.
-Higher layers of scattering coefficients will represent larger-scale structure that is better captured by the cluster-based representation presented below.
+The wavelet scattering transform, described below, has both of these desired properties: invariance to time-shifting and stability to time-warping.
+The stability condition can be formulated as a Lipschitz continuity property, which guarantees that the feature transforms of $\boldsymbol{x}(t)$ and $\boldsymbol{x_\tau}(t)$ are close together if $|\tau'(t)|$ is bounded by a small constant.
 
@@
 \ja{Where does the formula $2Q_1/\xi_1$ come from?}
+
+To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\phi(t)$ with cut-off frequency $1/T$ (and so a duration of approximately $T$), to get
+\begin{equation}
+\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \phi(t),
+\end{equation}
+which is known as the set of first-order scattering coefficients.
+They capture the average spectral envelope of $\boldsymbol{x}$ over scales of duration $T$, with the spectral resolution varying with constant Q, as described above.
+In this way, they are very closely related to the mel-frequency spectrogram and related features, such as MFCCs.
 

### 2016-12-13 paper/structureScene.tex
Add MFCC + SVM results

@@
 \midrule
-beach & $74.6 \pm 18.9$ & & $83.5 \pm \phantom{0}7.1$ \\
-bus & $58.2 \pm 18.3$ & & $88.8 \pm 11.4$ \\
-caf\'{e} / restaurant & $85.1 \pm 10.8$ & & $64.5 \pm \phantom{0}8.3$ \\
-car & $69.1 \pm 20.8$ & & $94.9 \pm \phantom{0}6.1$ \\
-city center & $89.6 \pm \phantom{0}9.2$ & & $91.7 \pm \phantom{0}9.4$ \\
-forest path & $72.4 \pm 23.8$ & & $93.8 \pm \phantom{0}7.8$ \\
-grocery store & $74.1 \pm 13.7$ & & $90.9 \pm \phantom{0}6.9$ \\
-home & $78.1 \pm 17.7$ & & $56.2 \pm 23.9$ \\
-library & $65.1 \pm 23.3$ & & $82.0 \pm 13.0$ \\
-metro station & $85.2 \pm 15.6$ & & $96.0 \pm \phantom{0}2.3$ \\
-office & $90.8 \pm 16.0$ & & $87.5 \pm 21.7$ \\
-park & $25.6 \pm 11.0$ & & $75.4 \pm \phantom{0}7.4$ \\
-residential area & $75.1 \pm 19.4$ & & $44.2 \pm 15.0$ \\
-train & $34.1 \pm \phantom{0}7.2$ & & $58.3 \pm 10.0$ \\
-tram & $85.4 \pm 13.9$ & & $83.7 \pm 12.2$ \\
+beach & $74.6 \pm 18.9$ & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ \\
+bus & $58.2 \pm 18.3$ & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ \\
+cafe/restaurant & $85.1 \pm 10.8$ & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ \\
+car & $69.1 \pm 20.8$ & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ \\
+city\_center & $89.6 \pm \phantom{0}9.2$ & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ \\
+forest\_path & $72.4 \pm 23.8$ & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ \\
+grocery\_store & $74.1 \pm 13.7$ & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ \\
+home & $78.1 \pm 17.7$ & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ \\
+library & $65.1 \pm 23.3$ & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ \\
+metro\_station & $85.2 \pm 15.6$ & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ \\
+office & $90.8 \pm 16.0$ & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ \\
+park & $25.6 \pm 11.0$ & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ \\
+residential\_area & $75.1 \pm 19.4$ & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ \\
+train & $34.1 \pm \phantom{0}7.2$ & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ \\
+tram & $85.4 \pm 13.9$ & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ \\
 \bottomrule
-Average & $70.8 \pm \phantom{0}2.6$ & & $79.4 \pm \phantom{0}3.0$
+Average & $70.8 \pm \phantom{0}2.6$ & $65.8 \pm \phantom{0}8.0$ & $79.4 \pm \phantom{0}3.0$ \\
 \end{tabular}

### 2016-12-13 paper/structureScene.tex
Add short discussion of MFCC + SVM results

@@
 This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
-The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$.\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.}
+For comparison, we also provide the results for the same MFCC features coupled with a linear SVM classifier.
+Results here are slightly worse since the SVM can only discriminate between classes that are linearly separable, a limitation not present for the GMMs.
+With the scattering transforms, however, the average accuracy across classes is brought from $70.8\%$ to $79.4\%$.\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.}
 Furthermore, our cross-validation analysis confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.

### 2016-12-13 paper/structureScene.tex
Swap figures 3 and 4

@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
-\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
-\label{fig:ASS_1}
+\includegraphics[width=.9\columnwidth]{bw/unsupervised_test1}
+\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.
+\ja{Why not \emph{ob-a} here?
+Also, the relationship between the different approaches for the regular scattering coefficients seems to be inverted with respect to the log-scattering and the previous figure.
+That is, here we have \emph{early} $>$ \emph{ob-w} $\approx$ \emph{ob-c} instead of \emph{ob-w} $>$ \emph{ob-w} $>$ \emph{early}.}}
+\label{fig:ASS_2}
 \end{center}
@@
 This section presents evaluation results for the ASSR settings.
-The $p@k$ for different settings are shown in Figures \ref{fig:ASS_1} and \ref{fig:ASS_2}, illustrating the effect of the different similarity metrics and the influence of the logarithm, respectively.
+The $p@k$ for different settings are shown in Figures \ref{fig:ASS_2} and \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics and the influence of the logarithm, respectively.
+
+\subsubsection*{Role of logarithmic compression}
+
+Outcomes for the scattering with and without logarithmic compression are shown in Figure \ref{fig:ASS_2}.
+It can be seen that the logarithmic compression strongly improves the results, especially for the \emph{ob} approaches, which perform worse than \emph{early} without the logarithmic compression.
 
@@
 
-\subsubsection*{Role of logarithmic compression}
-
-Outcomes for the scattering with and without logarithmic compression are shown in Figure \ref{fig:ASS_2}.
-It can be seen that the logarithmic compression strongly improves the results, especially for the \emph{ob} approaches, which perform worse than \emph{early} without the logarithmic compression.
-
 \begin{figure}[t]
 \begin{center}
-\includegraphics[width=.9\columnwidth]{bw/unsupervised_test1}
-\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.
-\ja{Why not \emph{ob-a} here?
-Also, the relationship between the different approaches for the regular scattering coefficients seems to be inverted with respect to the log-scattering and the previous figure.
-That is, here we have \emph{early} $>$ \emph{ob-w} $\approx$ \emph{ob-c} instead of \emph{ob-w} $>$ \emph{ob-w} $>$ \emph{early}.}}
-\label{fig:ASS_2}
+\includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
+\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
+\label{fig:ASS_1}
 \end{center}

### 2016-12-13 paper/responseToReviewers.tex
Add some responses to reviewer 1

@@
 \item \emph{Line 40, alpha may appear in previous equations}
+As far as we can tell, $\alpha$ only appears during the definition of the $\widehat{\mathrm{EMD}}$ distance and not anywhere else.
 
 \item \emph{The argumentation toward the log compression is well conducted, and meaningful in this issue. The authors may simply swap the order of fig 3 and 4}
+The order of the figures and their corresponding text have been reversed.
 

### 2016-12-13 paper/structureScene.tex
Add discussion of odd classes

@@
 Results here are slightly worse since the SVM can only discriminate between classes that are linearly separable, a limitation not present for the GMMs.
-With the scattering transforms, however, the average accuracy across classes is brought from $70.8\%$ to $79.4\%$.\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.}
+
+With the scattering transforms, however, the average accuracy across classes is brought from $70.8\%$ to $79.4\%$.\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.} It is interesting to note, however, that two classes, ``home'' and ``residential\_area'' both drop significantly in accuracy from the baseline GMM model to the scattering SVM classifier.
+This can be partially explained by the substitution of the GMM with an SVM, which shows a similar drop for the MFCCs.
+However, another reason could be that these classes are well-characterized by the average spectral envelope of the signal and the second-order modulation information captured by the scattering coefficients is therefore superfluous.
 Furthermore, our cross-validation analysis confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.

### 2016-12-13 paper/structureScene.tex
Edit filter bank phrase

@@
 
-Scattering transforms are time-shift invariant representations of audio signals consisting of auditory and modulation filter banks alternated with complex modulus non-linearities.
+Scattering transforms are time-shift invariant representations of audio signals computed by applying auditory and modulation filter banks alternated with complex modulus non-linearities.
 This section first explains the importance of local invariance to time-shifting and stability to time-warping in the representation of acoustic scenes, then describes how the scattering transform is designed to satisfy these properties while having a high discriminative power.

### 2016-12-13 paper/responseToReviewers.tex
More responses to the AE
Add more responses to reviewer 1

@@
 \item \emph{The class accuracy Table III for Home and Residential area are singular. Authors shall discuss that.}
+The text has been modified to discuss these results.
 
@@
 For example the whole Gammatone wavelet justification over Morlet is biased}
+While some of the sounds are ``non-natural'', that is human-caused, that does not make differentiating between them an impossible task for a biological system evolved under exposure to other, more ``natural'' sounds.
+In addition, the psychoacoustic results cited have been obtained by conducting experiments where subjects listened to the same type of ``non-natural'' sounds.
+As such, we feel that the justifications are well-adapted to our method and tasks.
 
@@
 \item \emph{the paper reads like a recipe for the specific task without justifications, insights or proofs of the different given definitions and statements.}
+We justify the cluster-based representation with results from cognitive psychology.
+The use of the scattering transform is justified with invariance conditions of the tasks.
+Where necessary, the proofs and evidence are provided in references.
+Again, the lack of specificity of this point makes it hard to address, but as stated above, we have reworked the whole paper to hopefully make our argument clearer.
 
@@
 \item \emph{line 60: auditory and modulation filter – nonsense.}
+While we are not certain what the intent of the reviewer is here, we have modified the sentence to clarify our meaning.
 

### 2016-12-13 paper/responseToReviewers.tex
Add responses to reviewer 2

@@
 \item \emph{Authors may precise  how the Gammatone wavelets allow to reproduce the asymmetry of temporal masking and the asymmetry of spectral masking.}
+\ja{Vincent, you explained this much better than I can. Would you mind adding this part?}
 
 \item \emph{Setting of Q1 shall be motivated.}
+This has been added to the text.
 
@@
 \item \emph{The cross-validation analyses confirmed that the value T = 372 ms for the support of the low-pass filter is optimal when applying late integration by majority voting, but shall it adapted per classes  ?}
+While it is true that different values of $T$ could be appropriate for different classes (that is, the different sound objects present could be of different scale), we do not expect results to change significantly.
+This is because the sounds are of a similar nature (ambient, non-directed) and so any distinct object shouldn't be much longer or shorter in duration in one class compared to another.
+In addtion, $T = 372\,\mathrm{ms}$ is at the longer end of the useful range of window sizes for which the scattering transform is useful, so it is unlikely that larger sizes would improve.
 
@@
 \ja{can you spot where we speak about binding acoustic cues ?, I can't ;(}
+\ja{Can't find it...}
 

### 2016-12-13 paper/responseToReviewers.tex
fix typos in response to reviewers

@@
 
-\item \emph{No novel technique is presented}
+\item \emph{No novel technique is presented.}
 Again, the hierarchical representation proposed is novel and we provide adequate evidence, in our opinion, for its power in the proposed similarity retrieval task.
 
-\item \emph{the whole paper architecture is poor}
+\item \emph{The whole paper architecture is poor.}
 Without any specific points of criticism, it is hard to address this particular statement. Nonetheless, we have reworked the paper as a whole, trying to clarify the overall message.
 
-\item \emph{all the justifications arise from psychoacoustic, bioacoustic,...
+\item \emph{All the justifications arise from psychoacoustics, bioacoustics ...
 when the challenge is to classify ambient non natural stationary sounds.
-For example the whole Gammatone wavelet justification over Morlet is biased}
+For example, the whole Gammatone wavelet justification over Morlet is biased.}
 While some of the sounds are ``non-natural'', that is human-caused, that does not make differentiating between them an impossible task for a biological system evolved under exposure to other, more ``natural'' sounds.
@@
 
-\item \emph{the classification accuracy is not suited for comparison since the dimension of the feature vectors are not the same for all the techniques }
+\item \emph{The classification accuracy is not suited for comparison since the dimension of the feature vectors are not the same for all the techniques.}
 While the dimension of the features vectors are different, this does not necessarily have to impact the performance of the classification.
@@
 
-\item \emph{most importantly the used classifier is not the same (from unsupervised GMM to supervised SVM)}
+\item \emph{The used classifier is not the same (from unsupervised GMM to supervised SVM)}
 This is an important point and we have included results for MFCCs using a linear SVM.
@@
 
-\item \emph{the paper reads like a recipe for the specific task without justifications, insights or proofs of the different given definitions and statements.}
+\item \emph{The paper reads like a recipe for the specific task without justifications, insights or proofs of the different given definitions and statements.}
 We justify the cluster-based representation with results from cognitive psychology.
@@
 
-\item \emph{the used descriptors are only locally time shift invariant and not globally which is never pointed out}
+\item \emph{The used descriptors are only locally time shift invariant and not globally which is never pointed out.}
 As formulated, the scattering transforms do indeed have only local time-shifting invariance, which has been clarified.
@@
 
-\item \emph{The function phi is hardly considered as a low-pass since a low-pass is only used in coordination with a high-pass for quadratic mirror filters. Here phi is a scaling function.}
+\item \emph{The function $\phi$ is hardly considered as a low-pass since a low-pass is only used in coordination with a high-pass for quadratic mirror filters. Here $\phi$ is a scaling function.}
 The function $\phi$ is very much a low-pass filter with a frequency support of approximately $1/T$ and a time bandwidth of $T$, where $T$ is of the order of hundreds of milliseconds.
@@
 \item \emph{"unsupervised classification" it is clustering then.}
-
 The sentence have been rewritten to be more explicit.
@@
 \item \emph{The cross-validation analyses confirmed that the value T = 372 ms for the support of the low-pass filter is optimal when applying late integration by majority voting, but shall it adapted per classes  ?}
-While it is true that different values of $T$ could be appropriate for different classes (that is, the different sound objects present could be of different scale), we do not expect results to change significantly.
+Although it is true that different values of $T$ could be appropriate for different classes (that is, the different sound objects present could be of different scale), we do not expect results to change significantly.
 This is because the sounds are of a similar nature (ambient, non-directed) and so any distinct object shouldn't be much longer or shorter in duration in one class compared to another.
@@
 
-The bibliography section is now more tight.
+The bibliography section has been made shorter.
 
-\item \emph{Check typos in the references (Museum, Histoire nat...). “befor”}
+\item \emph{Check typos in the references}
 

### 2016-12-14 paper/structureScene.tex
Add Q_1 = 8 motivation
move to biblatex

@@
 \documentclass[journal]{IEEEtran}
-\usepackage{cite}
+%\usepackage{cite}
 % *** GRAPHICS RELATED PACKAGES ***
@@
 
+\usepackage[backend=bibtex,style=numeric-comp,sorting=none,firstinits=true]{biblatex}
+\bibliography{biblio}
 %\renewcommand{\baselinestretch}{2}
@@
 Despite the differences, ecoacoustics and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?
-The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts: 1) application of the wavelet scattering transform to extract discriminative features and 2) large time-scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.
-\ja{Technically, the 1st concept has already been demonstrated in previous papers.
-Do we really want to claim this here?}
+The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts for environmental audio processing: 1) application of the wavelet scattering transform to extract discriminative features and 2) large time-scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.
+%\ja{Technically, the 1st concept has already been demonstrated in previous papers.
+Do we really want to claim this here?} ML narrowed the scope
 
@@
 
-A recent paradigm in computational bioacoustics processes the audio stream in a holistic way over large time-scales, without assuming that a single species is present throughout the recording.
-\ja{Should we define somewhere what ``computational bioacoustics'' is?} Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as marking time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.
+Computational bioacoustics usually refers to the investigation of sound production, dispersion and reception in animals (including humans) by computational means.
+A recent paradigm  processes the audio stream in a holistic way over large time-scales, without assuming that a single species is present throughout the recording.
+Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as marking time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.
 
@@
 The algorithm marked with $^*$ considered only the public dataset.
-\label{tab:dcase} \ja{What are the initials OE, ELF, etc.
+\label{tab:dcase} \ml{need to define RQA ILD LDA or remove table ?} %\ja{What are the initials OE, ELF, etc.
 for?
@@
 \hline
-\cite{olivetti2013wonder} OE & 17 & compression distance, random forests \\
-\cite{elizalde2013vector} ELF & 55 & i-vector, LDA \\
-\cite{krijnders2013tone} KH & 55 & cochleogram, SVMs \\
-\cite{patil2013multiresolution} PE & 58 & 2D Gabor filters, SVMs \\
-\cite{lee2013acoustic} NHL & 60 & unsupervised learning over mel spectra, SVMs \\
-\cite{nogueira2013sound} NR1 & 60 & MFCCs and ILD features, SVMs \\
-\cite{chum2013ieee} CHR & 65 & sparsity features, HMMs  \\
-\cite{geiger2013large} GSR & 69 & mel spectra and MFCCs, SVMs \\
-\cite{rakotomamonjy2015histogram} RG & 69 & histogram of gradient (HOG), SVMs \\
-\cite{li2013auditory} LTT & 72 & MFCCs, tree baggers \\
-\cite{roma2013} RNH & 76 & RQA features stacked with MFCCs, SVMs \\
+\cite{olivetti2013wonder} & 17 & compression distance, random forests \\
+\cite{elizalde2013vector} & 55 & i-vector, LDA \\
+\cite{krijnders2013tone} & 55 & cochleogram, SVMs \\
+\cite{patil2013multiresolution} & 58 & 2D Gabor filters, SVMs \\
+\cite{lee2013acoustic} & 60 & unsupervised learning over mel spectra, SVMs \\
+\cite{nogueira2013sound} & 60 & MFCCs and ILD features, SVMs \\
+\cite{chum2013ieee} & 65 & sparsity features, HMMs  \\
+\cite{geiger2013large} & 69 & mel spectra and MFCCs, SVMs \\
+\cite{rakotomamonjy2015histogram} & 69 & histogram of gradient (HOG), SVMs \\
+\cite{li2013auditory} & 72 & MFCCs, tree baggers \\
+\cite{roma2013} & 76 & RQA features stacked with MFCCs, SVMs \\
 \hline
@@
 This results in the wavelet filters covering the whole range of human hearing, from $20~\mathrm{Hz}$ to $20~\mathrm{kHz}$.
+Setting $Q_1 = 8$ results in filters whose bandwidth approximates an equivalent rectangular bandwidth (ERB) scale.
 
@@
 
-As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of object-based representations of auditory scenes for predicting high-level properties.
-As the detection of events is still an open problem \cite{7100934}, we consider in this paper a quantization scheme that is as generic as possible, implemented using a clustering approach where coherent regions of the scene are identified.
+As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
+As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
 \ja{This last sentence is a bit hard to parse.
-Are we segmenting the time-series and identifying coherent regions?}
+Are we segmenting the time-series and identifying coherent regions?} \ml{better ?}
 
@@
 The performance of the temporal scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (ASSR) and in a supervised setting (ASC).
-\ja{The \emph{ob} approaches are not really assessed for ASC, though.}
+\ja{The \emph{ob} approaches are not really assessed for ASC, though.} \ml{well, we say that this does not work well, no ?}
 
@@
 \ja{Why don't we display these results?
-Also how is the variance so much smaller here?} This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already achieved by the classifiers during training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
+Also how is the variance so much smaller here?} \ml{I agree, we should put the results} This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already achieved by the classifiers during training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
 \ja{This sentence is very confusing.}
@@
 
-\bibliographystyle{unsrt}
-\bibliography{biblio}
+%\bibliographystyle{unsrt}
+%\bibliography{biblio}
+\printbibliography
 

### 2016-12-14 paper/responseToReviewers.tex
capitalization in response to reviewers

@@
 
-\item \emph{"recordings may be shifted locally 
+\item \emph{"Recordings may be shifted locally 
 without affecting their perception and therefore such shifts do not convey any information about the class" this is false.
@@
 
-\item \emph{"wavelet scalogram" is used along all the paper but scalogram is only defined for wavelets since 
+\item \emph{"Wavelet scalogram" is used along all the paper but scalogram is only defined for wavelets since 
 it is a scale-o-gram.}

### 2016-12-14 paper/structureScene.tex
answer to ja

--- 2016-12-13 paper/structureScene.tex
@@
 In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-\ja{Why is $\xi_2 \neq \xi_1$?} Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee local invariance to time-shifting, which yields
+\ja{Why is $\xi_2 \neq \xi_1$?} \vl{$\xi_2 = \xi_1 / 8$ because $Q_1=8$}Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee local invariance to time-shifting, which yields
 \begin{equation}
@@
 Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
-\ja{Isn't this why we also standardize the coefficients?}
+\ja{Isn't this why we also standardize the coefficients?} \vl{No, standardization does not affect skewness, unlike log compression}
 This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).

### 2016-12-14 paper/structureScene.tex
upload biography

@@
 
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lostanlen}}]{Vincent Lostanlen} was born in 1992.
+He received his M.S.
+degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie and the Ircam in Paris, France.
+Since 2013, he is a PhD student at ENS in Paris.
+\end{IEEEbiography}
+
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in Paris, France, in 1990.
@@
 He received his M.S.
-degree in acoustics, signal processing and musical informatics (ATIAM) from the University Pierre and Marie Curie and the IRCAM laboratory, Paris, France, in 2013.
-Since 2013, he is a Ph.D student at Irccyn, a French laboratory dedicated to cybernetics.
+degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre and Marie Curie and the Ircam in Paris, France, in 2013.
+Since 2013, he is a Ph.D student at IRCCyN, a French laboratory dedicated to cybernetics.
 His current research interests include acoustic scene similarity and classification, as well as acoustic scene perception.
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at Irccyn, a French laboratory dedicated to cybernetics.
-He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions, in Canada (University of Victoria, McGill University) and in France (Orange Labs, Telecom ParisTech, Ircam).
+\begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
+He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, Telecom ParisTech, Ircam).
 His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.
-\end{IEEEbiography}
 

### 2016-12-14 paper/structureScene.tex
forgotten \end{IEEEbography}

--- 2016-12-14 paper/structureScene.tex
@@
 
-%\bibliographystyle{unsrt}
-%\bibliography{biblio}
+\bibliographystyle{unsrt}
+\bibliography{biblio}
 \printbibliography
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lostanlen}}]{Vincent Lostanlen} was born in 1992.
+He received his M.S.
+degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie and the Ircam in Paris, France.
+Since 2013, he is a PhD student at ENS in Paris.
+\end{IEEEbiography}
 
@@
 His current research interests include acoustic scene similarity and classification, as well as acoustic scene perception.
+>>>>>>> origin/master
 \end{IEEEbiography}
 
-\begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at Irccyn, a French laboratory dedicated to cybernetics.
-He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions, in Canada (University of Victoria, McGill University) and in France (Orange Labs, Telecom ParisTech, Ircam).
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in Paris, France, in 1990.
+He received in 2011 the B.S.
+degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, France, and the B.S.
+degree in Musicology from the Sorbonne University, Paris, France.
+He received his M.S.
+degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre and Marie Curie and the Ircam in Paris, France, in 2013.
+Since 2013, he is a Ph.D student at IRCCyN, a French laboratory dedicated to cybernetics.
+His current research interests include acoustic scene similarity and classification, as well as acoustic scene perception.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
+He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, Telecom ParisTech, Ircam).
 His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.

### 2016-12-14 paper/structureScene.tex
forgotten >>>>>>> origin/master
leaf's biography appears twice
Fix curly brace
Fix biblio
Commands that we had in ef8812
hyphen in title: relevance-based instead of relevance based

@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Auditory Scene Similarity Retrieval and Classification with Relevance based Quantization of Scattering Features}
+\title{Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features}
 
@@
 for?
-Do we need them?}}
+Do we need them?}
+}
 \begin{tabular}{lcl}
@@
 
-\bibliographystyle{unsrt}
-\bibliography{biblio}
+%\bibliographystyle{unsrt}
+%\bibliography{biblio}
 \printbibliography
@@
 Since 2013, he is a PhD student at ENS in Paris.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in Paris, France, in 1990.
-He received in 2011 the B.S.
-degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, France, and the B.S.
-degree in Musicology from the Sorbonne University, Paris, France.
-He received his M.S.
-degree in acoustics, signal processing and musical informatics (ATIAM) from the University Pierre and Marie Curie and the IRCAM laboratory, Paris, France, in 2013.
-Since 2013, he is a Ph.D student at Irccyn, a French laboratory dedicated to cybernetics.
-His current research interests include acoustic scene similarity and classification, as well as acoustic scene perception.
->>>>>>> origin/master
 \end{IEEEbiography}

### 2016-12-14 paper/structureScene.tex
update caption in Figure 1

@@
 Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$. 
-\vl{If you want to keep this level of detail you have to adapt the figure so that it is readable in greyscale.
-Blue and red oscillations represent the real and imaginary parts.
-The orange envelope represents the complex modulus.}
-}
+Oscillations represent the real and imaginary parts. The orange envelope represents the complex modulus.}
 \end{center}

### 2016-12-14 paper/structureScene.tex
remove dialog between ja and me on xi_1 and xi_2

@@
 In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-\ja{Why is $\xi_2 \neq \xi_1$?} \vl{$\xi_2 = \xi_1 / 8$ because $Q_1=8$}Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee local invariance to time-shifting, which yields
+Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee local invariance to time-shifting, which yields
 \begin{equation}

### 2016-12-14 paper/structureScene.tex
TODO on asymmetry of masking in Gammatones

@@
 In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
-Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
+Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.\vl{TODO}
 Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.

### 2016-12-14 paper/structureScene.tex
remove ja's comment on sum of probabilities
minor changes in biographies

@@
 where $\hat{y}$ is the estimated class label and $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.
-\ja{Do we really sum probabilities?
-Seems like a strange thing to do.}
 
@@
 He received his M.S.
-degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie and the Ircam in Paris, France.
+degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie and the Ircam in Paris, France, in 2013.
 Since 2013, he is a PhD student at ENS in Paris.
@@
 \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
-He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, Telecom ParisTech, Ircam).
+He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, TELECOM ParisTech, Ircam).
 His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.

### 2016-12-14 paper/structureScene.tex
explain asymmetry of temporal and spectral masking

@@
 In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
-Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.\vl{TODO}
-Moreover, it should be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
+Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
+The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
+Likewise, because critical bands are skewed towards higher frequencies, a masker tone has to be louder if it is above the stimulus in frequency rather than below.
+It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
 As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.

### 2016-12-14 paper/structureScene.tex
"wavelet scalogram" -> "scalogram" in two places

@@
 Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
-However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
+However, contrary to the CQT, the scalogram reaches the Heisenberg
 theoretical limit of optimal time-frequency localization across the whole
 frequency range, whereas the temporal resolution of the traditional CQT is fixed by the support of the STFT analyzing window. %\cite{Brown1992}
-Therefore, the wavelet scalogram has a better temporal localization at high
+Therefore, the scalogram has a better temporal localization at high
 frequencies than the CQT, at the expense of a greater computational cost

### 2016-12-14 paper/structureScene.tex
write more of my biography + punctuation in gl's biography

@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lostanlen}}]{Vincent Lostanlen} was born in 1992.
-He received his M.S.
-degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie and the Ircam in Paris, France, in 2013.
-Since 2013, he is a PhD student at ENS in Paris.
+He received an engineering degree from TELECOM ParisTech in 2013 and M.S.
+degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie and the Ircam (UPMC) in Paris, France, in 2013.
+Since 2013, he is a Ph.D.
+student at \'{E}cole normale sup\'{e}rieure in Paris, working under the supervision of St\'{e}phane Mallat.
 \end{IEEEbiography}
@@
 degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre and Marie Curie and the Ircam in Paris, France, in 2013.
-Since 2013, he is a Ph.D student at IRCCyN, a French laboratory dedicated to cybernetics.
-His current research interests include acoustic scene similarity and classification, as well as acoustic scene perception.
+Since 2013, he is a Ph.D.
+student at IRCCyN, a French laboratory dedicated to cybernetics.
+His research interests include acoustic scene similarity and classification as well as acoustic scene perception.
 \end{IEEEbiography}

### 2016-12-14 paper/structureScene.tex
remove ja's comment on feature renormalization

@@
 \ie a linear combination of the logarithms of first- and second-order coefficients.
-As such, a non-linear renormalization becomes a linear transformation, which can be learned by a linearly discriminative classifier.
-\ja{Do we do this here or not?
-What is the point of this paragraph?}
+As such, a nonlinear renormalization becomes a linear transformation, which can be learned by a linearly discriminative classifier.
 

### 2016-12-14 paper/responseToReviewers.tex
response 2 to reviewer 2

@@
 \item \emph{Authors may precise how the Gammatone wavelets allow to reproduce the asymmetry of temporal masking and the asymmetry of spectral masking.}
-\ja{Vincent, you explained this much better than I can. Would you mind adding this part?}
+We have written a short explanation of the role of Gammatone wavelets in the modeling of the asymmetry of temporal masking and the asymmetry of spectral masking.
 

### 2016-12-15 paper/responseToReviewers.tex
some comments on the response to reviewer 2
polishing response

--- 2016-12-14 paper/responseToReviewers.tex
@@
 \item \emph{Line 40, alpha may appear in previous equations}
+
 As far as we can tell, $\alpha$ only appears during the definition of the $\widehat{\mathrm{EMD}}$ distance and not anywhere else.
@@
 \item \emph{The argumentation toward the log compression is well conducted, and meaningful in this issue. The authors may simply swap the order of fig 3 and 4}
+
 The order of the figures and their corresponding text have been reversed.
@@
 \item \emph{The class accuracy Table III for Home and Residential area are singular. Authors shall discuss that.}
+
 The text has been modified to discuss these results.
@@
 \item \emph{Object-based is inappropriate, This paper introduces: no this paper uses the scattering network to try to tackle a specific sound classification task.}
+
 The title has been changed to remove the "object-based" formulation.
@@
 \item \emph{The second paragraph of the abstract is vague.}
+
 This paragraph has been reworded and worked into the rest of the abstract in a more fluid manner.
@@
 \item \emph{No novel technique is presented.}
-Again, the hierarchical representation proposed is novel and we provide adequate evidence, in our opinion, for its power in the proposed similarity retrieval task.
+
+Again, the hierarchical representation proposed is novel and we provide adequate evidence, in our opinion, for its power in the proposed similarity retrieval task in the case of environmental audio data processing.
 
 \item \emph{The whole paper architecture is poor.}
-Without any specific points of criticism, it is hard to address this particular statement. Nonetheless, we have reworked the paper as a whole, trying to clarify the overall message.
+
+We have reworked the paper as a whole, trying to clarify the overall message.
+
+\ml{I think we could do it better, see my issue}
 
@@
 For example, the whole Gammatone wavelet justification over Morlet is biased.}
+
 While some of the sounds are ``non-natural'', that is human-caused, that does not make differentiating between them an impossible task for a biological system evolved under exposure to other, more ``natural'' sounds.
@@
 \item \emph{The classification accuracy is not suited for comparison since the dimension of the feature vectors are not the same for all the techniques.}
+
 While the dimension of the features vectors are different, this does not necessarily have to impact the performance of the classification.
@@
 \item \emph{The used classifier is not the same (from unsupervised GMM to supervised SVM)}
-This is an important point and we have included results for MFCCs using a linear SVM.
+
+We have included results for MFCCs using a linear SVM.
 However, we point out that the GMM is also supervised, although it is trained using a generative objective as opposed to a discriminative one, which is the case for the SVM.
@@
 \item \emph{The paper reads like a recipe for the specific task without justifications, insights or proofs of the different given definitions and statements.}
+
 We justify the cluster-based representation with results from cognitive psychology.
@@
 Where necessary, the proofs and evidence are provided in references.
-Again, the lack of specificity of this point makes it hard to address, but as stated above, we have reworked the whole paper to hopefully make our argument clearer.
+The lack of specificity of this point makes it hard to address, but as stated above, we have reworked the whole paper to hopefully make our argument clearer.
 
 \item \emph{The dataset is not well presented since the classes seem to represent stationary noises such as car/tram/train}
+
 While these scenes definitely have a stationary component, such as for the car, tram, train, metro station, and bus classes.
@@
 Rather say that application of time shift should not change the representation since it doesn't change the class belonging in this context and thus imposing time shift invariance will improve robustness.}
-The reviewer is correct that this is not true as stated. Our meaning is indeed what he/she describes and the text has been changed to reflect this.
+
+That was indeed our meaning, the text has been changed to reflect this.
 
 \item \emph{The used descriptors are only locally time shift invariant and not globally which is never pointed out.}
+
 As formulated, the scattering transforms do indeed have only local time-shifting invariance, which has been clarified.
@@
 \item \emph{The function $\phi$ is hardly considered as a low-pass since a low-pass is only used in coordination with a high-pass for quadratic mirror filters. Here $\phi$ is a scaling function.}
+
 The function $\phi$ is very much a low-pass filter with a frequency support of approximately $1/T$ and a time bandwidth of $T$, where $T$ is of the order of hundreds of milliseconds.
@@
 \item \emph{Line 34: "wavelet transform modulus" is not clear at all.}
+
 We have updated the text to clarify this expression.
@@
 \item \emph{It is then explained that the features are stable to time warping yet the justification is for pitch variations which are frequency-warping}
+
 The justification of pitch variation is indeed a frequency-warping, but can be written as a time-warping.
@@
 \item \emph{The scattering network is defined with an infinite cascade of transform yet it is enough to compute two layers, why?}
+
 For the time scales considered in this paper, the first two orders are sufficient to describe the sound events present in the signals. This has been made clearer in the text.
@@
 it is a scale-o-gram.}
+
 This is true, but for readers less familiar with the terminology, we have opted to use this more redundant expression to express the fact that these derive from wavelet decompositions.
@@
 \item \emph{line 60: auditory and modulation filter – nonsense.}
-While we are not certain what the intent of the reviewer is here, we have modified the sentence to clarify our meaning.
+
+We have modified the sentence to clarify our meaning.
 
 \item \emph{"unsupervised classification" it is clustering then.}
+
 The sentence have been rewritten to be more explicit.
@@
 \item \emph{"time-frequency perspective or a machine learning perspective" is it imposed versus learned representation? }
+
 The intention was to situate the feature processing step either as post-processing of the scattering features or as pre-processing for the classifier stage. We have removed it during the revision.
@@
 \item \emph{Equations 9 and 10 are superfluous.}
+
 These equations were removed from the paper.
@@
 
+We acknowldege that such a diagram would help the reader, we are facing page limit contraints due to the extensive experiments decribed in the paper.
+
 \item \emph{Authors may precise how the Gammatone wavelets allow to reproduce the asymmetry of temporal masking and the asymmetry of spectral masking.}
-\ja{Vincent, you explained this much better than I can. Would you mind adding this part?}
+
+We have written a short explanation of the role of Gammatone wavelets in the modeling of the asymmetry of temporal masking and the asymmetry of spectral masking.
 
 \item \emph{Setting of Q1 shall be motivated.}
+
 This has been added to the text.
@@
 \item \emph{In the last experience (Table III) GMM(MFCC) may be compared to SVM(MFCC) (in addition to SVM(logscat)). There is somewhere a high dimension effect that shall be clarified.}
+
 We have added results for SVM on MFCCs.
-As mentioned in the response to the other reviewer, we believe that the high dimensionality should not have a significant effect on these results.
-\ja{Although perhaps we should indeed talk about this in the paper.}
+As mentioned in the response to the other reviewer, we believe that the high dimensionality should not have a significant effect on these results, even-though we acknowledge that this effect would be interesting to study.
+Our justification have been added to the manuscript.
 
@@
 
+We have computed the confusion matrices and they are indeed of interest as the two methods do not confuse the same classes.
+Though, due to page limit constraints, they can not be added to the manuscript.
+
+%They are added as supplementary material in manuscript central and will be available on the code repository.
+
 \item \emph{The cross-validation analyses confirmed that the value T = 372 ms for the support of the low-pass filter is optimal when applying late integration by majority voting, but shall it adapted per classes  ?}
+
 Although it is true that different values of $T$ could be appropriate for different classes (that is, the different sound objects present could be of different scale), we do not expect results to change significantly.
 This is because the sounds are of a similar nature (ambient, non-directed) and so any distinct object shouldn't be much longer or shorter in duration in one class compared to another.
-In addtion, $T = 372\,\mathrm{ms}$ is at the longer end of the useful range of window sizes for which the scattering transform is useful, so it is unlikely that larger sizes would improve.
+In addition, $T = 372\,\mathrm{ms}$ is at the longer end of the useful range of window sizes for which the scattering transform is useful, so it is unlikely that larger sizes would improve.
 
@@
 
-\ja{can you spot where we speak about binding acoustic cues ?, I can't ;(}
-\ja{Can't find it...}
+We sought for such a discussion in the manuscript and were unable to find one.
 

### 2016-12-15 paper/structureScene.tex
removing uneeded package
fix caption of fig 1
polishing response

@@
 \usepackage{array}
-\usepackage{fixltx2e}
+%\usepackage{fixltx2e}
 
@@
 Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$. 
-Oscillations represent the real and imaginary parts. The orange envelope represents the complex modulus.}
+Oscillations represent the real and imaginary parts whose complex modulus is also shown.}
 \end{center}
@@
 \begin{center}
-\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline algorithm (left) and proposed algorithm (right).}
+\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline algorithm (left) and proposed algorithm (right).\label{table:dcase2016}}
 \begin{tabular}{lccc}
@@
 \end{center}
-\label{table:dcase2016}
+
 \end{table}
@@
 Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$, we therefore use a linear kernel instead.
-Our experiments on the smaller DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
+Our experiments on the smaller DCASE 2013 dataset, as reported in Table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
 

### 2016-12-16 paper/structureScene.tex
update caption of Fig 1
Update abstract and keywords

@@
 
-This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second time-scales with relevance-based quantization at larger time-scales.
-Instead of representing a scene using summary statistics, a method popularized by bag-of-features models, each scene is represented by a subset of its most representative features, allowing the retrieval system to focus on discriminative features.
-We demonstrate the potential of the proposed approach for both acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets, consistently outperforming baseline mel-frequency cepstral coefficient (MFCC) methods.
+An important task in computational bioacoustics and related fields, such as ecouacoustics, is to determine the similarity of two audio scenes.
+This is often done using bag-of-frames models and other representations based on summary statistics.
+While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background.
+To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure.
+To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs).
+Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.
 
@@
 \begin{IEEEkeywords}
-acoustic scene classification, acoustic scene segmentation, acoustic scene similarity retrieval, scattering, wavelets, invariance
+bioacoustics, ecoacoustics, acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
 \end{IEEEkeywords}
@@
 Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$. 
-Oscillations represent the real and imaginary parts whose complex modulus is also shown.}
+Oscillations represent the real and imaginary parts. The envelope represents the complex modulus.}
 \end{center}

### 2016-12-16 paper/structureScene.tex
Remove addressed comments

--- 2016-12-15 paper/structureScene.tex
@@
 The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts for environmental audio processing: 1) application of the wavelet scattering transform to extract discriminative features and 2) large time-scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.
-%\ja{Technically, the 1st concept has already been demonstrated in previous papers.
-Do we really want to claim this here?} ML narrowed the scope
 
@@
 The algorithm marked with $^*$ considered only the public dataset.
-\label{tab:dcase} \ml{need to define RQA ILD LDA or remove table ?} %\ja{What are the initials OE, ELF, etc.
-for?
-Do we need them?}
+\label{tab:dcase} \ml{need to define RQA ILD LDA or remove table ?}
 }
@@
 As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
-\ja{This last sentence is a bit hard to parse.
-Are we segmenting the time-series and identifying coherent regions?} \ml{better ?}
 
@@
 The performance of the temporal scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (ASSR) and in a supervised setting (ASC).
-\ja{The \emph{ob} approaches are not really assessed for ASC, though.} \ml{well, we say that this does not work well, no ?}
 
@@
 Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
-\ja{Isn't this why we also standardize the coefficients?} \vl{No, standardization does not affect skewness, unlike log compression}
 This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
@@
 Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.
-\ja{The relationship between linear and RBF is inverted for the \emph{early} integration scheme compared to \emph{late}. Is this correct?}
 }

### 2016-12-16 paper/structureScene.tex
Change k to q for NN parameter

--- 2016-12-16 paper/structureScene.tex
@@
 \label{eq:kc}
-K_{mn}^{uv} = \exp\left( - \dfrac{\Vert \mu_m^u - \mu_n^v \Vert^2}{\Vert \mu_m^u - \mu_{m,k}^u \Vert \Vert \mu_n^v - \mu_{n,k}^v \Vert} \right).
+K_{mn}^{uv} = \exp\left( - \dfrac{\Vert \mu_m^u - \mu_n^v \Vert^2}{\Vert \mu_m^u - \mu_{m,q}^u \Vert \Vert \mu_n^v - \mu_{n,q}^v \Vert} \right).
 \end{equation} 
-Here, $\mu_{m,k}^u$ and $\mu_{n,k}^v$ are the $k^{\textrm{th}}$ nearest neighbors to the centroids $\mu_m^u$ and $\mu_n^v$, respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+Here, $\mu_{m,q}^u$ and $\mu_{n,q}^v$ are the $q^{\textrm{th}}$ nearest neighbors to the centroids $\mu_m^u$ and $\mu_n^v$, respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
 
@@
 The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
-\ja{The variable $k$ is used previously, so we should change that.} The $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
+The $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
 Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item.
-For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $M$ for the \emph{ob} approaches and the scaling parameter $k$ of the RBF kernels (see Eq.
+For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $M$ for the \emph{ob} approaches and the scaling parameter $q$ of the RBF kernels (see Eq.
 \ref{eq:kc}) leading to the best $p@9$.

### 2016-12-16 paper/structureScene.tex
Remove variance phrase & comment

@@
 As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
-Not only is the average accuracy improved, but the variance between fold-wise accuracies is also reduced, such as from $16\%$ to $7\%$ in the case of early integration. \ja{Why does this happen?}
 This is in accordance with previous results on similarity retrieval, as well as comparisons between mel-frequency coefficients and scattering coefficients for tasks ranging from musical genre recognition and phone segment classification \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.

### 2016-12-16 paper/structureScene.tex
Fix 0.7 standard deviation

@@
 Surprisingly, although we ran a large number of experiments to formulate the relevance-based quantization scheme into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in \emph{late} temporal integration, namely majority voting.
-Among such methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$.
-\ja{Why don't we display these results?
-Also how is the variance so much smaller here?} \ml{I agree, we should put the results} This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already achieved by the classifiers during training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
+Among such methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm 7$.
+\ja{Why don't we display these results?} \ml{I agree, we should put the results} This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already achieved by the classifiers during training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
 \ja{This sentence is very confusing.}

### 2016-12-16 paper/structureScene.tex
Bioacoustics -> ecoacoustics

@@
 
-An important task in computational bioacoustics and related fields, such as ecouacoustics, is to determine the similarity of two audio scenes.
+un important task in computational ecoacoustics is to determine the similarity of two audio scenes.
 This is often done using bag-of-frames models and other representations based on summary statistics.
@@
 \begin{IEEEkeywords}
-bioacoustics, ecoacoustics, acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
+ecoacoustics, acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
 \end{IEEEkeywords}

### 2016-12-16 paper/structureScene.tex
New intro
This is currently too long, I think, so will need to be condensed
to relevant size.

@@
 \IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
+In order to measure the effect of human activity and climate change on animal biodiversity researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
 In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
@@
 
-Ecoacoustics is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science.
+This works falls in the realm of ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
 Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
-As a consequence, few well-designed datasets are readily available for evaluation purposes.
+Consequently, few well-designed datasets are readily available for evaluation purposes.
 
-A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
-While narrower in its range of applications, it has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of well-designed datasets.
+An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation.
+One application of this is acoustic scene similarity retrieval (ASSR), where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset.
+This requires the auditory scenes to be represented in some way that captures their distinctive qualities.
+Often, this has been achieved by bag-of-frames approaches, which attempt to describe the auditory scene recording using summary statistics.
+Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain rare, distinct sound events that discriminate the scene from others.Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes.
+\ja{Refs?}
 
-Despite the differences, ecoacoustics and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?
-The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts for environmental audio processing: 1) application of the wavelet scattering transform to extract discriminative features and 2) large time-scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.
+To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}.
+In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.
+
+The first question to be answered when confronted with this task is how is the signal to be represented at a low level.
+In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
+A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years \cite{logan}.
+\ja{Other MFCC works?} A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
+
+The next question is the choice of a classifier.
+Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM), to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision.
+This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes.
+The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it completely.
+Combining scattering transforms for low-level representation with late-integrated SVMs used for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.
+
+Returning to scene retrieval setting, we cannot use supervised models such as SVMs, but we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
+We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a number of clusters.
+We then use these to define a set of distances that can subsequently be used for similarity retrieval.Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.
 

### 2016-12-16 paper/structureScene.tex
Edit background section

@@
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
-Being a very generic method, this approach has been applied to many audio classification tasks besides ASC.
-Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
-Indeed, the typical morphology of acoustic scenes is a ``skeleton of events on a bed of textures'' \cite{nelken2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
+Interestingly, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
+This contrasts with the typical morphology of acoustic scenes, which is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events superimposed upon a stationary acoustic background.
 Therefore, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on the joint probability distribution of short-term features.
@@
 An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
-Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that use the private part of the DCASE 2013 dataset for their experiments.
-Most of the scene prediction results are achieved by majority voting of predictions over local features.
 
-\begin{table}
-\begin{center}
-\caption{Algorithms evaluated on the DCASE 2013 private dataset of acoustic scene classification.
-Algorithms above the horizontal line participated to the DCASE challenge.
-The algorithm marked with $^*$ considered only the public dataset.
-\label{tab:dcase} \ml{need to define RQA ILD LDA or remove table ?}
-}
-\begin{tabular}{lcl}
-& Acc. \%  &  \multicolumn{1}{c}{Design}  \\
-\hline
-\cite{olivetti2013wonder} & 17 & compression distance, random forests \\
-\cite{elizalde2013vector} & 55 & i-vector, LDA \\
-\cite{krijnders2013tone} & 55 & cochleogram, SVMs \\
-\cite{patil2013multiresolution} & 58 & 2D Gabor filters, SVMs \\
-\cite{lee2013acoustic} & 60 & unsupervised learning over mel spectra, SVMs \\
-\cite{nogueira2013sound} & 60 & MFCCs and ILD features, SVMs \\
-\cite{chum2013ieee} & 65 & sparsity features, HMMs  \\
-\cite{geiger2013large} & 69 & mel spectra and MFCCs, SVMs \\
-\cite{rakotomamonjy2015histogram} & 69 & histogram of gradient (HOG), SVMs \\
-\cite{li2013auditory} & 72 & MFCCs, tree baggers \\
-\cite{roma2013} & 76 & RQA features stacked with MFCCs, SVMs \\
-\hline
-\cite{phan2016label} & 84/87 & label tree embedding of Gammatone, SVMs \\
-\cite{7362846} & 85$^*$ & amplitude modulated spectra, LDA \\
-\end{tabular}
-\end{center}
-\end{table}
+The methods that performed the best all followed a similar structure, where a low-level feature representation, such as MFCCs, were calculated over short windows and fed separately to a classifier, such as an SVM, for training.
+During the classification stage, the recording to be classified was similarly processed into short windows represented by feature vectors which were classified individually by the classifier and the classifier output was combined over the recording through majority vote.
+This last step is known as late integration and allows the classifier to work on individual, shorter-scale sound events instead of trying to classify the recording as a whole.
+By doing this, the classifier is able to identify and exploit the distinct events that characterize the scene instead of trying to model its average structure.
+This is in agreement with the psychoacoustic results related to the perception of auditory scenes presented above.
 

### 2016-12-16 paper/structureScene.tex
Subtitle in scattering section
Discussion is relevant to any audio signals, not just auditory
scenes.

@@
 
-\subsection{Invariance and stability in acoustic scenes}
+\subsection{Invariance and stability in audio signals}
 The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.

### 2016-12-16 paper/structureScene.tex
Edit feature design; break out integration
New section on supervised system, talking about SVM and
integration approaches.

@@
 
-Before classification or clustering, it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
-In this section, we review three transformations which achieve these properties, namely logarithmic compression, temporal integration, and standardization.
+Prior to constructing models (for classification or similarity estimation), it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
+In this section, we review three transformations which achieve these properties, namely logarithmic compression and standardization.
 
@@
 
+\subsection{Standardization}
+\label{sec:stand}
+
+Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where the $\gamma$ and $n$ respectively denotes feature and sample indices, respectively.
+It has been found that SVMs should be trained on scaled features with null mean and unit variance so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
+To standardize $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
+The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the training set only, and the same affine transformation is then applied to all samples in both the training and the test sets.
+
+\section{Supervised classification system}
+\label{sec:supervised}
+
+We are now ready to present the supervised method that we will apply to the acoustic scene classification task using scattering transforms.
+It consists of two parts, a classifier and a temporal integration method.
+While the classifier is an SVM, we shall consider two integration approaches, early and late.
+
+\subsection{Support vector machine classifier}
+\label{sec:svm}
+
+A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets\cite{cortes-vapnik}.
+From another point of view, it attemts to find a direction on which to project in order to maximize separation between two point clouds.
+Since it is not a generative, but discriminative model, it is able to pick out those feature vectors that best characterize a particular class of auditory scenes compared to another.
+This property, coupled with its ease of use and efficient implementations \cite{libsvm} have rendered SVMs a popular choice for many classification systems.
+
+At the classification stage, the SVM gives a class label as output, which is one of the two classes it models.
+For multi-class settings, a one-\vs-one approach is often used, where an SVM is trained for each pair of classes.
+During classification, the outputs of all pairwise SVMs is computed and the class selected the largest number of times is chosen.
+
 \subsection{Early \vs late temporal integration}
@@
 
-A drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene.
-In the context of the DCASE 2013 private dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, which can increase the variance of the estimated model and run the risk of overfitting.
-On the contrary, once the scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$ has been unpadded of any boundary artifacts, a late integration scheme for $T=372\textrm{ ms}$ would yield $128$ instances per $30$-second auditory scene, resulting in $10240$ instances overall.
-However, many of these instances may be silent or lack any salient properties of the class they are assigned to, which may increase the bias in the estimated model and run the risk of underfitting.
-Moreover, when applying late integration, the classifier optimizes accuracy at the frame level, not at the scene level, so its prediction might not yield optimal training error under the final evaluation metric.
-In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff.
-The appropriateness of either approach depends therefore on the nature of the signal that is classified.
-We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
-In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority vote of the predictions given by an SVM.
-
-\subsection{Standardization}
-\label{sec:stand}
-
-Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where the $\gamma$ and $n$ respectively denotes feature and sample indices, respectively.
-It has been found that SVMs should be trained on scaled features with null mean and unit variance so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
-To standardize $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
-The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the training set only, and the same affine transformation is then applied to all samples in both the training and the test sets.
+A potential drawback of early integration is that it reduces a sequence of feature vectors to a single average.
+If the underlying signal is static in nature, this may be advantageous, as the temporal variation of the feature vectors constitutes an unwanted source of variability.
+For many audio signals, and in particular auditory scenes, this is very far from the truth, and prevents us from identifying the distinct sound events that characterize the signal as these may be drowned out by the background during the averaging step.
+In this case, a late integration approach would perform better, since it operates on the individual feature vectors and only the classification results are aggregated over time.
+Using experiments in acoustic scene classification below, we will demonstrate that the late integration approach does indeed succeed over early integration when scattering coefficients are used.
 

### 2016-12-16 paper/structureScene.tex
Edit object section
Edit methods
Reorganize to separate the two experiments and their intentions.

@@
 
-\section{Relevance based quantization of features}
+\section{Relevance-based quantization of features}
 \label{sec:object}
@@
 As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
+While this can be addressed in the supervised case using late integration of discriminative classifiers (see Section \ref{sec:eili}), this is not possible in the unsupervised case.
 As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
@@
 
-\subsection{Methods}
+\subsection{Experiment \#1: acoustic scene classification}
 
-The performance of the temporal scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (ASSR) and in a supervised setting (ASC).
+Our first task is to compare the performance of the standard scattering transform, the logarithmically compressed scattering (log-scattering) transform, and baseline MFCCs.
+Secondly, we evaluate the two temporal integration approaches presented above, \emph{early} and \emph{late}.
 
-For ASSR, evaluation is performed on the private part of the DCASE 2013 dataset.
+The different systems are evaluated on the private part of the DCASE 2013 dataset using five-fold stratified cross validation.
+The folds are identical to those used in the original challenge.
+The slack parameter $C$ of the SVM is here set to $1$.
+Both linear and Gaussian RBF kernels are used.
+For the latter, a local scaling method is applied \cite{selfTuneManor2004}.
+All feature vectors, that is scattering frames, MFCCs and centroids, are standardized prior to be given to the classifier.
+The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
+Optimal parameters (see above) are this time learned on the public part of the DCASE 2013 dataset, also using five-fold stratified cross validation.
+\ja{Which parameters are these?} For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).
+
+\subsection{Experiment \#2: acoustic scene similarity retrieval}
+
+Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval.
+
+The evaluation is performed on the private part of the DCASE 2013 dataset.
 The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
@@
 The \emph{ob} approaches are compared to commonly used early integration approach \emph{early}.
+\ja{And BOF?}
+
+The clustering for the \emph{ob} methods is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure.
+Three numbers of clusters are tested: $8$, $16$ and $32$.
+The implementation of the presented methods as well as the one of the experimental protocol is available.\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}} \ja{What is ``the one''?}
 
@@
 \end{figure}
-
-For ASC, the systems are again evaluated on the private part of DCASE 2013, but here with five-fold stratified cross validation.
-The folds are identical to those used in the original challenge.
-In the same way as the majority of the DCASE 2013 participants, we use an SVM for the ASC task, which is a large-margin binary discriminator with tunable regularization of misclassified examples.
-The slack parameter $C$ is here set to $1$.
-Optimal parameters (see above) are this time learned on the public part of the DCASE 2013 dataset, also using five-fold stratified cross validation.
-\ja{Which parameters are these?} For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).
-
-For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}.
-For both the \emph{early} and \emph{late} methods, a linear kernel as well as an RBF kernel are used.
-As with the RBF kernels used in the \emph{ob} approaches, the RBF kernel in the SVM is scaled locally using the method proposed in \cite{selfTuneManor2004}.
-All feature vectors, that is scattering frames, MFCCs and centroids, are standardized prior to be given to the classifier.
-The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
-
-The clustering for the \emph{ob} methods is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure.
-Three numbers of clusters are tested: $8$, $16$ and $32$.
-The implementation of the presented methods as well as the one of the experimental protocol is available.\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}} \ja{What is ``the one''?}
 

### 2016-12-16 paper/structureScene.tex
Edit results section
I changed the order, per the ordering of the rest of the paper. I
also removed the log-compression part of the ASSR experiment. The
advantage of the log is already in a sense established by the ASC
experiments, so we can gain some space (and repeat ourselves a
little less) but removing this part. Of course, we can always put
it back in.

@@
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{bw/unsupervised_test1}
-\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.
-\ja{Why not \emph{ob-a} here?
-Also, the relationship between the different approaches for the regular scattering coefficients seems to be inverted with respect to the log-scattering and the previous figure.
-That is, here we have \emph{early} $>$ \emph{ob-w} $\approx$ \emph{ob-c} instead of \emph{ob-w} $>$ \emph{ob-w} $>$ \emph{early}.}}
-\label{fig:ASS_2}
-\end{center}
-\end{figure}
-
 \section{Results \label{sec:results}}
 
-\subsection{Acoustic Scene Similarity Retrieval}
+The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation.
+Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering features and MFCCs.
 
-This section presents evaluation results for the ASSR settings.
-The $p@k$ for different settings are shown in Figures \ref{fig:ASS_2} and \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics and the influence of the logarithm, respectively.
-
-\subsubsection*{Role of logarithmic compression}
-
-Outcomes for the scattering with and without logarithmic compression are shown in Figure \ref{fig:ASS_2}.
-It can be seen that the logarithmic compression strongly improves the results, especially for the \emph{ob} approaches, which perform worse than \emph{early} without the logarithmic compression.
-
-\subsubsection*{MFCC \vs scattering transform}
-
-Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach.
-Overall, log-compressed scattering coefficients systematically outperform MFCCs.
-This is to be expected since the scattering coefficients capture larger-scale modulations, as opposed to MFCCs which only describe the short-time spectral envelope.
-
-\subsubsection*{Relevance-based quantization \vs early integration}
-
-For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization to improve the similarity measures between the scenes.
-However, it is worth noticing that \emph{ob-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
-To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.
-Furthermore, it appears that \emph{ob-c} is better able to characterize the classes than \emph{ob-w}.
-This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
-Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
-On the contrary, two similar environments may shared a lot of similar sound sources with only a few sources discriminating between them.
-
-The same observations are made for the MFCCs for $k\leq5$.
-However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently.
-This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
-\ja{What is the ``at some point'' referring to?
-In other words, why do we expect this to happen for larger $k$ but not smaller?}
-
-%However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
-The latter even outperforms \emph{ob-a} for $k\geq5$.
-From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
-This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
-\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
-\label{fig:ASS_1}
-\end{center}
-\end{figure}
-
-\subsection{Acoustic Scene Classification}
+\subsection{Experiment \#1: acoustic scene classification}
 
@@
 
+\subsubsection*{Role of logarithmic compression}
+
+Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
+Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
+This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
+
 \subsubsection*{Early \vs late integration}
@@
 Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
-
-Surprisingly, although we ran a large number of experiments to formulate the relevance-based quantization scheme into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in \emph{late} temporal integration, namely majority voting.
-Among such methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm 7$.
-\ja{Why don't we display these results?} \ml{I agree, we should put the results} This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already achieved by the classifiers during training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
-\ja{This sentence is very confusing.}
-
-\subsubsection*{Role of logarithmic compression}
-
-Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
-Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
-This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
-Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.
 
@@
 Furthermore, our cross-validation analysis confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
+
+\subsection{Experiment \#2: acoustic scene similarity retrieval}
+
+This section presents evaluation results for the ASSR settings. The $p@k$ for different settings are shown in Figure \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics.
+
+\subsubsection*{MFCC \vs scattering transform}
+
+Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach.
+Overall, log-compressed scattering coefficients systematically outperform MFCCs.
+This is to be expected since the scattering coefficients capture larger-scale modulations, as opposed to MFCCs which only describe the short-time spectral envelope.
+
+\subsubsection*{Relevance-based quantization \vs early integration}
+
+For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization to improve the similarity measures between the scenes.
+However, it is worth noticing that \emph{ob-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
+To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.
+Furthermore, it appears that \emph{ob-c} is better able to characterize the classes than \emph{ob-w}.
+This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
+Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
+On the contrary, two similar environments may shared a lot of similar sound sources with only a few sources discriminating between them.
+
+The same observations are made for the MFCCs for $k\leq5$.
+However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently.
+This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
+\ja{What is the ``at some point'' referring to?
+In other words, why do we expect this to happen for larger $k$ but not smaller?}
+
+%However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
+The latter even outperforms \emph{ob-a} for $k\geq5$.
+From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
+This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
+\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
+\label{fig:ASS_1}
+\end{center}
+\end{figure}
 

### 2016-12-16 paper/structureScene.tex
Add non-linear SVM paragraph

@@
 
-A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets\cite{cortes-vapnik}.
+A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{cortes-vapnik}.
 From another point of view, it attemts to find a direction on which to project in order to maximize separation between two point clouds.
@@
 During classification, the outputs of all pairwise SVMs is computed and the class selected the largest number of times is chosen.
+
+Finally, the SVM can also be extended to handle non-linear decision surfaces.
+This is obtained using the ``kernel trick,'' which exploits the fact that all calculations in the SVM rely only on the inner product between vectors.
+By replacing this standard inner product by some other positive definite symmetric mapping, such as a Gaussian radial basis function (RBF), this is equivalent to performing a linear SVM after mapping the feature vectors into some high-dimensional space.
+A non-linear mapping then results in a non-linear decision surface in the orignal space.
 

### 2016-12-16 paper/structureScene.tex
New conclusion

@@
 
-In this paper we consider the benefit of 1) the scattering transform for building discriminant features and 2) unsupervised clustering for the modeling of large-scale environmental acoustic scenes.
-%This leads to improvements in both the ASSR and ASC tasks.
-
-Representing a scene using a reduced set of representatives, with adapted similarity strategies, allows us to improve ASSR performance compared to the early-integration scheme which have been demonstrated to be equivalent to BOF approaches using GMMs on well-balanced datasets \cite{lagrange:hal-01082501}.
-These results advocate for a hierarchical model of the scene, where short-time structure--\ie below a second--is represented using a convolutional network such as the scattering transform, whereas larger time scales--\ie over a second--are modeled as adaptive aggregates.
-
-Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{s}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine classification for the tasks we considered.
-%, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high.
-\ja{I'm not sure what this last part means.}
-
+This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales.
+Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature.
+We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification.
+In order to extend the ideas of late integration to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval.
+For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
 These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in ecoacoustics.

### 2016-12-16 paper/structureScene.tex
Add Andén biography

--- 2016-12-16 paper/structureScene.tex
@@
 
-un important task in computational ecoacoustics is to determine the similarity of two audio scenes.
-This is often done using bag-of-frames models and other representations based on summary statistics.
-While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background.
-To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure.
-To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs).
-Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.
+This paper introduces a generic hierarchical modeling scheme for acoustic scenes using the wavelet scattering transform at sub-second time-scales with relevance-based quantization at larger time-scales.
+Instead of representing a scene using summary statistics, a method popularized by bag-of-features models, each scene is represented by a subset of its most representative features, allowing the retrieval system to focus on discriminative features.
+We demonstrate the potential of the proposed approach for both acoustic scene similarity retrieval and acoustic scene classification tasks on publicly available datasets, consistently outperforming baseline mel-frequency cepstral coefficient (MFCC) methods.
 
@@
 \begin{IEEEkeywords}
-ecoacoustics, acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
+acoustic scene classification, acoustic scene segmentation, acoustic scene similarity retrieval, scattering, wavelets, invariance
 \end{IEEEkeywords}
@@
 \IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
+In order to measure the effect of human activity and climate change on animal biodiversity, researchers in the field of ecoacoustics \cite{krause} have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
 In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
@@
 
-This works falls in the realm of ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
+Ecoacoustics is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science.
 Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
-Consequently, few well-designed datasets are readily available for evaluation purposes.
+As a consequence, few well-designed datasets are readily available for evaluation purposes.
 
-An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation.
-One application of this is acoustic scene similarity retrieval (ASSR), where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset.
-This requires the auditory scenes to be represented in some way that captures their distinctive qualities.
-Often, this has been achieved by bag-of-frames approaches, which attempt to describe the auditory scene recording using summary statistics.
-Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain rare, distinct sound events that discriminate the scene from others.Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes.
-\ja{Refs?}
+A closely related, and more mature, field of investigation is the classification of acoustic scenes, where the aim is to predict labels by processing audio data.
+While narrower in its range of applications, it has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive, guastavino_ideal_2006} and has received the attention of the data mining community for some time, resulting in the availability of well-designed datasets.
 
-To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}.
-In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.
-
-The first question to be answered when confronted with this task is how is the signal to be represented at a low level.
-In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
-A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years \cite{logan}.
-\ja{Other MFCC works?} A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
-
-The next question is the choice of a classifier.
-Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM), to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision.
-This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes.
-The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it completely.
-Combining scattering transforms for low-level representation with late-integrated SVMs used for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.
-
-Returning to scene retrieval setting, we cannot use supervised models such as SVMs, but we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
-We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a number of clusters.
-We then use these to define a set of distances that can subsequently be used for similarity retrieval.Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.
+Despite the differences, ecoacoustics and acoustic scene classification (ASC) task both raise the following question: how can we efficiently and effectively process a large amount of audio data using representations that are compact, generic in their design, and flexible in their use?
+The work described in this paper attempts to answer this question by demonstrating the usefulness of two concepts for environmental audio processing: 1) application of the wavelet scattering transform to extract discriminative features and 2) large time-scale modeling of the acoustic scene based on unsupervised clustering of each acoustic scene, allowing us to focus the scene description on a few salient sound objects that carry most of the semantic information.
 
@@
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
-Interestingly, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
-This contrasts with the typical morphology of acoustic scenes, which is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events superimposed upon a stationary acoustic background.
+Being a very generic method, this approach has been applied to many audio classification tasks besides ASC.
+Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
+Indeed, the typical morphology of acoustic scenes is a ``skeleton of events on a bed of textures'' \cite{nelken2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
 Therefore, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on the joint probability distribution of short-term features.
@@
 An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
+Table \ref{tab:dcase} summarizes the results of the submitted algorithms as well as newer systems that use the private part of the DCASE 2013 dataset for their experiments.
+Most of the scene prediction results are achieved by majority voting of predictions over local features.
 
-The methods that performed the best all followed a similar structure, where a low-level feature representation, such as MFCCs, were calculated over short windows and fed separately to a classifier, such as an SVM, for training.
-During the classification stage, the recording to be classified was similarly processed into short windows represented by feature vectors which were classified individually by the classifier and the classifier output was combined over the recording through majority vote.
-This last step is known as late integration and allows the classifier to work on individual, shorter-scale sound events instead of trying to classify the recording as a whole.
-By doing this, the classifier is able to identify and exploit the distinct events that characterize the scene instead of trying to model its average structure.
-This is in agreement with the psychoacoustic results related to the perception of auditory scenes presented above.
+\begin{table}
+\begin{center}
+\caption{Algorithms evaluated on the DCASE 2013 private dataset of acoustic scene classification.
+Algorithms above the horizontal line participated to the DCASE challenge.
+The algorithm marked with $^*$ considered only the public dataset.
+\label{tab:dcase} \ml{need to define RQA ILD LDA or remove table ?}
+}
+\begin{tabular}{lcl}
+& Acc. \%  &  \multicolumn{1}{c}{Design}  \\
+\hline
+\cite{olivetti2013wonder} & 17 & compression distance, random forests \\
+\cite{elizalde2013vector} & 55 & i-vector, LDA \\
+\cite{krijnders2013tone} & 55 & cochleogram, SVMs \\
+\cite{patil2013multiresolution} & 58 & 2D Gabor filters, SVMs \\
+\cite{lee2013acoustic} & 60 & unsupervised learning over mel spectra, SVMs \\
+\cite{nogueira2013sound} & 60 & MFCCs and ILD features, SVMs \\
+\cite{chum2013ieee} & 65 & sparsity features, HMMs  \\
+\cite{geiger2013large} & 69 & mel spectra and MFCCs, SVMs \\
+\cite{rakotomamonjy2015histogram} & 69 & histogram of gradient (HOG), SVMs \\
+\cite{li2013auditory} & 72 & MFCCs, tree baggers \\
+\cite{roma2013} & 76 & RQA features stacked with MFCCs, SVMs \\
+\hline
+\cite{phan2016label} & 84/87 & label tree embedding of Gammatone, SVMs \\
+\cite{7362846} & 85$^*$ & amplitude modulated spectra, LDA \\
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-\subsection{Invariance and stability in audio signals}
+\subsection{Invariance and stability in acoustic scenes}
 The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
@@
 
-Prior to constructing models (for classification or similarity estimation), it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
-In this section, we review three transformations which achieve these properties, namely logarithmic compression and standardization.
+Before classification or clustering, it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
+In this section, we review three transformations which achieve these properties, namely logarithmic compression, temporal integration, and standardization.
 
@@
 
-\subsection{Standardization}
-\label{sec:stand}
-
-Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where the $\gamma$ and $n$ respectively denotes feature and sample indices, respectively.
-It has been found that SVMs should be trained on scaled features with null mean and unit variance so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
-To standardize $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
-The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the training set only, and the same affine transformation is then applied to all samples in both the training and the test sets.
-
-\section{Supervised classification system}
-\label{sec:supervised}
-
-We are now ready to present the supervised method that we will apply to the acoustic scene classification task using scattering transforms.
-It consists of two parts, a classifier and a temporal integration method.
-While the classifier is an SVM, we shall consider two integration approaches, early and late.
-
-\subsection{Support vector machine classifier}
-\label{sec:svm}
-
-A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets\cite{cortes-vapnik}.
-From another point of view, it attemts to find a direction on which to project in order to maximize separation between two point clouds.
-Since it is not a generative, but discriminative model, it is able to pick out those feature vectors that best characterize a particular class of auditory scenes compared to another.
-This property, coupled with its ease of use and efficient implementations \cite{libsvm} have rendered SVMs a popular choice for many classification systems.
-
-At the classification stage, the SVM gives a class label as output, which is one of the two classes it models.
-For multi-class settings, a one-\vs-one approach is often used, where an SVM is trained for each pair of classes.
-During classification, the outputs of all pairwise SVMs is computed and the class selected the largest number of times is chosen.
-
 \subsection{Early \vs late temporal integration}
@@
 
-A potential drawback of early integration is that it reduces a sequence of feature vectors to a single average.
-If the underlying signal is static in nature, this may be advantageous, as the temporal variation of the feature vectors constitutes an unwanted source of variability.
-For many audio signals, and in particular auditory scenes, this is very far from the truth, and prevents us from identifying the distinct sound events that characterize the signal as these may be drowned out by the background during the averaging step.
-In this case, a late integration approach would perform better, since it operates on the individual feature vectors and only the classification results are aggregated over time.
-Using experiments in acoustic scene classification below, we will demonstrate that the late integration approach does indeed succeed over early integration when scattering coefficients are used.
+A drawback of early integration is that it drastically reduces the number of training instances at the classifier stage, down to one per auditory scene.
+In the context of the DCASE 2013 private dataset, this corresponds to merely $8$ training instances per class, and $80$ instances overall, which can increase the variance of the estimated model and run the risk of overfitting.
+On the contrary, once the scattering representation $\mathbf{S}\boldsymbol{x}(t,\gamma)$ has been unpadded of any boundary artifacts, a late integration scheme for $T=372\textrm{ ms}$ would yield $128$ instances per $30$-second auditory scene, resulting in $10240$ instances overall.
+However, many of these instances may be silent or lack any salient properties of the class they are assigned to, which may increase the bias in the estimated model and run the risk of underfitting.
+Moreover, when applying late integration, the classifier optimizes accuracy at the frame level, not at the scene level, so its prediction might not yield optimal training error under the final evaluation metric.
+In short, early and late integration methods lie at opposite ends of the bias-versus-variance statistical tradeoff.
+The appropriateness of either approach depends therefore on the nature of the signal that is classified.
+We refer to \cite{Joder2009} for a comprehensive review of this problematic in the context of musical instrument recognition.
+In the context of this paper, early integration refers to an averaging of the features and late integration refers to a majority vote of the predictions given by an SVM.
 
-\section{Relevance-based quantization of features}
+\subsection{Standardization}
+\label{sec:stand}
+
+Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where the $\gamma$ and $n$ respectively denotes feature and sample indices, respectively.
+It has been found that SVMs should be trained on scaled features with null mean and unit variance so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
+To standardize $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
+The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the training set only, and the same affine transformation is then applied to all samples in both the training and the test sets.
+
+\section{Relevance based quantization of features}
 \label{sec:object}
@@
 As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
-While this can be addressed in the supervised case using late integration of discriminative classifiers (see Section \ref{sec:eili}), this is not possible in the unsupervised case.
 As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
@@
 
-\subsection{Experiment \#1: acoustic scene classification}
+\subsection{Methods}
 
-Our first task is to compare the performance of the standard scattering transform, the logarithmically compressed scattering (log-scattering) transform, and baseline MFCCs.
-Secondly, we evaluate the two temporal integration approaches presented above, \emph{early} and \emph{late}.
+The performance of the temporal scattering transform and the \emph{ob} approaches are assessed in an unsupervised setting (ASSR) and in a supervised setting (ASC).
 
-The different systems are evaluated on the private part of the DCASE 2013 dataset using five-fold stratified cross validation.
-The folds are identical to those used in the original challenge.
-The slack parameter $C$ of the SVM is here set to $1$.
-Both linear and Gaussian RBF kernels are used.
-For the latter, a local scaling method is applied \cite{selfTuneManor2004}.
-All feature vectors, that is scattering frames, MFCCs and centroids, are standardized prior to be given to the classifier.
-The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
-Optimal parameters (see above) are this time learned on the public part of the DCASE 2013 dataset, also using five-fold stratified cross validation.
-\ja{Which parameters are these?} For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).
-
-\subsection{Experiment \#2: acoustic scene similarity retrieval}
-
-Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval.
-
-The evaluation is performed on the private part of the DCASE 2013 dataset.
+For ASSR, evaluation is performed on the private part of the DCASE 2013 dataset.
 The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
@@
 The \emph{ob} approaches are compared to commonly used early integration approach \emph{early}.
-\ja{And BOF?}
+
+%\ja{I'm not sure what this means. Do we fix $m$ and $N$ here for all experiments or do we optimize them over the test data each time? If so, this will bias the results, no?}
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{bw/unsupervised_test1}
+\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.
+\ja{Why not \emph{ob-a} here?
+Also, the relationship between the different approaches for the regular scattering coefficients seems to be inverted with respect to the log-scattering and the previous figure.
+That is, here we have \emph{early} $>$ \emph{ob-w} $\approx$ \emph{ob-c} instead of \emph{ob-w} $>$ \emph{ob-w} $>$ \emph{early}.}}
+\label{fig:ASS_2}
+\end{center}
+\end{figure}
+
+For ASC, the systems are again evaluated on the private part of DCASE 2013, but here with five-fold stratified cross validation.
+The folds are identical to those used in the original challenge.
+In the same way as the majority of the DCASE 2013 participants, we use an SVM for the ASC task, which is a large-margin binary discriminator with tunable regularization of misclassified examples.
+The slack parameter $C$ is here set to $1$.
+Optimal parameters (see above) are this time learned on the public part of the DCASE 2013 dataset, also using five-fold stratified cross validation.
+\ja{Which parameters are these?} For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).
+
+For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}.
+For both the \emph{early} and \emph{late} methods, a linear kernel as well as an RBF kernel are used.
+As with the RBF kernels used in the \emph{ob} approaches, the RBF kernel in the SVM is scaled locally using the method proposed in \cite{selfTuneManor2004}.
+All feature vectors, that is scattering frames, MFCCs and centroids, are standardized prior to be given to the classifier.
+The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
 
@@
 
-%\ja{I'm not sure what this means. Do we fix $m$ and $N$ here for all experiments or do we optimize them over the test data each time? If so, this will bias the results, no?}
-
 \section{Results \label{sec:results}}
 
-The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation.
-Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering features and MFCCs.
+\subsection{Acoustic Scene Similarity Retrieval}
 
-\subsection{Experiment \#1: acoustic scene classification}
+This section presents evaluation results for the ASSR settings.
+The $p@k$ for different settings are shown in Figures \ref{fig:ASS_2} and \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics and the influence of the logarithm, respectively.
+
+\subsubsection*{Role of logarithmic compression}
+
+Outcomes for the scattering with and without logarithmic compression are shown in Figure \ref{fig:ASS_2}.
+It can be seen that the logarithmic compression strongly improves the results, especially for the \emph{ob} approaches, which perform worse than \emph{early} without the logarithmic compression.
+
+\subsubsection*{MFCC \vs scattering transform}
+
+Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach.
+Overall, log-compressed scattering coefficients systematically outperform MFCCs.
+This is to be expected since the scattering coefficients capture larger-scale modulations, as opposed to MFCCs which only describe the short-time spectral envelope.
+
+\subsubsection*{Relevance-based quantization \vs early integration}
+
+For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization to improve the similarity measures between the scenes.
+However, it is worth noticing that \emph{ob-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
+To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.
+Furthermore, it appears that \emph{ob-c} is better able to characterize the classes than \emph{ob-w}.
+This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
+Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
+On the contrary, two similar environments may shared a lot of similar sound sources with only a few sources discriminating between them.
+
+The same observations are made for the MFCCs for $k\leq5$.
+However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently.
+This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
+\ja{What is the ``at some point'' referring to?
+In other words, why do we expect this to happen for larger $k$ but not smaller?}
+
+%However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
+The latter even outperforms \emph{ob-a} for $k\geq5$.
+From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
+This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
+\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
+\label{fig:ASS_1}
+\end{center}
+\end{figure}
+
+\subsection{Acoustic Scene Classification}
 
@@
 
-\subsubsection*{Role of logarithmic compression}
-
-Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
-Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
-This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
-
 \subsubsection*{Early \vs late integration}
@@
 Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
+
+Surprisingly, although we ran a large number of experiments to formulate the relevance-based quantization scheme into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in \emph{late} temporal integration, namely majority voting.
+Among such methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm 7$.
+\ja{Why don't we display these results?} \ml{I agree, we should put the results} This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already achieved by the classifiers during training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).
+\ja{This sentence is very confusing.}
+
+\subsubsection*{Role of logarithmic compression}
+
+Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
+Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
+This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
+Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.
 
@@
 
-\subsection{Experiment \#2: acoustic scene similarity retrieval}
-
-This section presents evaluation results for the ASSR settings. The $p@k$ for different settings are shown in Figure \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics.
-
-\subsubsection*{MFCC \vs scattering transform}
-
-Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach.
-Overall, log-compressed scattering coefficients systematically outperform MFCCs.
-This is to be expected since the scattering coefficients capture larger-scale modulations, as opposed to MFCCs which only describe the short-time spectral envelope.
-
-\subsubsection*{Relevance-based quantization \vs early integration}
-
-For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization to improve the similarity measures between the scenes.
-However, it is worth noticing that \emph{ob-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
-To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.
-Furthermore, it appears that \emph{ob-c} is better able to characterize the classes than \emph{ob-w}.
-This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
-Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
-On the contrary, two similar environments may shared a lot of similar sound sources with only a few sources discriminating between them.
-
-The same observations are made for the MFCCs for $k\leq5$.
-However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently.
-This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
-\ja{What is the ``at some point'' referring to?
-In other words, why do we expect this to happen for larger $k$ but not smaller?}
-
-%However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
-The latter even outperforms \emph{ob-a} for $k\geq5$.
-From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
-This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
-\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
-\label{fig:ASS_1}
-\end{center}
-\end{figure}
-
 \section{Conclusion}
@@
 His research interests include acoustic scene similarity and classification as well as acoustic scene perception.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/anden}}]{Joakim And\'en}
+received the M.Sc.
+degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the PhD degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014.
+He is currently a postdoctoral researcher with the Program in Applied and Computational mathematics at Princeton University, Princeton, USA.
+His research interests include signal processing, machine learning, and statistical data analysis.
+He is a member of the IEEE.
+
 \end{IEEEbiography}

### 2016-12-16 paper/structureScene.tex

--- 2016-12-16 paper/structureScene.tex
@@
 
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/anden}}]{Joakim And\'en}
+received the M.Sc.
+degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the PhD degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014.
+He is currently a postdoctoral researcher with the Program in Applied and Computational mathematics at Princeton University, Princeton, USA.
+His research interests include signal processing, machine learning, and statistical data analysis.
+He is a member of the IEEE.
+
+\end{IEEEbiography}
+
 \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.

### 2016-12-16 paper/structureScene.tex
polishing abs and intro

@@
 
-un important task in computational ecoacoustics is to determine the similarity of two audio scenes.
+An important task in computational ecoacoustics is to determine the similarity of two audio scenes.
 This is often done using bag-of-frames models and other representations based on summary statistics.
@@
 This requires the auditory scenes to be represented in some way that captures their distinctive qualities.
-Often, this has been achieved by bag-of-frames approaches, which attempt to describe the auditory scene recording using summary statistics.
-Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain rare, distinct sound events that discriminate the scene from others.Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes.
-\ja{Refs?}
+Often, this has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics.
+Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain rare, distinct sound events that discriminate the scene from others.
+Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes.
+\gl{put some Refs}
 
@@
 
-The first question to be answered when confronted with this task is how is the signal to be represented at a low level.
+Adressing this task requires first to be able to capture meanningful aspects within the signal at a small time scale.
 In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
-A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years \cite{logan}.
-\ja{Other MFCC works?} A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
+A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environemental audio processing \cite{aucouturier2007bag}.
+A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
 
 The next question is the choice of a classifier.
-Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM), to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision.
+Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM) \cite{cristianini2000introduction}, to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision.
 This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes.
-The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it completely.
+The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it almost completely.
 Combining scattering transforms for low-level representation with late-integrated SVMs used for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.
@@
 Returning to scene retrieval setting, we cannot use supervised models such as SVMs, but we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
-We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a number of clusters.
-We then use these to define a set of distances that can subsequently be used for similarity retrieval.Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.
+We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters.
+We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval.
+Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.
 

### 2016-12-16 paper/structureScene.tex
clarify between bio and eco acoustics

@@
 
-An important task in computational ecoacoustics is to determine the similarity of two audio scenes.
-This is often done using bag-of-frames models and other representations based on summary statistics.
+An important task in the emerging field of computational ecoacoustics is to determine the similarity of two audio scenes to be able to get a high level understanding of large amount of audio recordings coming from large deployment of acoustics sensors that are continuously recordings the sonic environment.
+A convenient method considers bag-of-frames models and other representations based on summary statistics.
 While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background.
@@
 \begin{IEEEkeywords}
-ecoacoustics, acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
+acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
 \end{IEEEkeywords}
@@
 
-This works falls in the realm of ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
+This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
 Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
@@
 For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
-These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in ecoacoustics.
+These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in computational bioacoustics.
 

### 2016-12-16 paper/structureScene.tex
done with polishing

@@
 Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes.
-\gl{put some Refs}
+\gl{put some Refs that are already cited if possible}
 
@@
 We describe the scattering transform in Section \ref{sec:scattering}, discuss processing of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
-In Section \ref{sec:experiments}, several experiments are proposed for the unsupervised setting of acoustic scene similarity retrieval (ASSR) and for the supervised setting of acoustic scene classification (ASC).
+In Section \ref{sec:experiments}, several experiments are described for the acoustic scene classification task and the acoustic scene similarity retrieval task.
 Results are subsequently reported in Section \ref{sec:results}.
@@
 %The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-For these tasks, a popular approach is the bag-of-frames (BOF) approach for acoustic scene modeling introduced by Aucouturier et al.
+In a closely related field, namely the analysis of the urban sound environment, a popular approach is the bag-of-frames (BOF) introduced in the field by Aucouturier et al.
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
-Interestingly, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
-This contrasts with the typical morphology of acoustic scenes, which is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events superimposed upon a stationary acoustic background.
-Therefore, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on the joint probability distribution of short-term features.
+Though, recently, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
+This contrasts with the typical morphology of acoustic scenes, which we believe is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events are superimposed upon a stationary acoustic background.
+As those distinct and scarce events are most of the time highly relevant for the task at hand, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on the joint probability distribution of short-term features.
 
@@
 
-Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of ASSR, much of the work dedicated to acoustic scene modeling is applied to the ASC task.
+Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of acoustic scene similarity retrieval, much of the work dedicated to acoustic scene modeling is applied to the classification task.
 An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
@@
 
-\section{Relevance-based quantization of features}
+\section{Acoustic scene similarity retrieval}
 \label{sec:object}
@@
 As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
-While this can be addressed in the supervised case using late integration of discriminative classifiers (see Section \ref{sec:eili}), this is not possible in the unsupervised case.
+While this can be addressed in the supervised case using late integration of discriminative classifiers (see Section \ref{sec:eili}), this is not direclty feasible in the unsupervised case.
 As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
@@
 One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
-
-%which clusters not over a single scene, but over the whole dataset, and re-projects each scene in the resulting enhanced feature space. \ja{This sentence is also a big unclear.}
-
 Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroids.
@@
 
+The implementations of the presented methods and the experimental protocol are available online\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}}. % \ja{What is ``the one''?}
+
 \subsection{Datasets}
@@
 
-The public part of the dataset can used for optimizing the ASC system and the private part used for computing the resulting accuracy using a five-fold cross validation scheme.
+The public part of the dataset can used for optimizing the acoustic scene classification system and the private part used for computing the resulting accuracy using a five-fold cross validation scheme.
 For this paper, we shall perform our experiments on the private part of the dataset.
@@
 
-The DCASE 2016 dataset has a similar organization, but here we conduct our experiments on the public part of the dataset. \ja{Any more information on this dataset?}
+The DCASE 2016 dataset has a similar organization, with $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
+
+%, but here we conduct our experiments on the public part of the dataset. \ja{Any more information on this dataset?}
 
@@
 The \emph{ob} approaches are compared to commonly used early integration approach \emph{early}.
-\ja{And BOF?}
+%\ja{And BOF?}
 
@@
 Three numbers of clusters are tested: $8$, $16$ and $32$.
-The implementation of the presented methods as well as the one of the experimental protocol is available.\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}} \ja{What is ``the one''?}
 
@@
 The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation.
-Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering features and MFCCs.
+Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering and MFCCs features.
 
@@
 
-We report experiments on acoustic scene classification (ASC), following the original methodology of the DCASE 2013 challenge, and extend our approach to its 2016 edition.
+We report experiments on acoustic scene classification, following the original methodology of the DCASE 2013 challenge, and extend our approach to its 2016 edition.
 
@@
 \subsubsection*{Participation to the DCASE 2016 challenge}
+
 In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
-This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
+This edition relies on the TUT 2016 acoustic scenes dataset.
+
 Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$, we therefore use a linear kernel instead.
@@
 
+\ml{this section takes one column and is I believe a good candidate for being shrunk or even merged with the previous subsection.
+By statement like, similar findings ...
+or those results are confirmed using the 2016 dataset ...}
+
 \subsection{Experiment \#2: acoustic scene similarity retrieval}
 
-This section presents evaluation results for the ASSR settings. The $p@k$ for different settings are shown in Figure \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics.
+This section presents evaluation results for the acoustic scene similarity retrieval task.
+The $p@k$ for different settings are shown in Figure \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics.
+The metrics is shown for values of $k$ up to 9 which corresponds to the number of elements minus one within each class of scene of the dataset.
+This allows us to also evaluate the recall capabilities of the system under evaluation.
 
@@
 However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently.
-This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
+As such, one can conclude that the use of MFCCs allows us to retrieve closely related scenes, but their lack of discriminative capabilities making the cluster selection/weighting step useless for not so closely related scenes of the same class
+
+%This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
 \ja{What is the ``at some point'' referring to?
@@
 \includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
-\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
 \label{fig:ASS_1}
@@
 We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification.
-In order to extend the ideas of late integration to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval.
+In order to extend the selectivity potential of late integration using SVMs to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval.
 For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.

### 2016-12-17 paper/structureScene.tex
Edit discussion on late integration results

@@
 Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
-This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to separately characterize distinct acoustic events beyond the scale of about one second.
-\ja{This is not suggested since the scattering transform is computed at $T=372\,\mathrm{ms}$ and so cannot be expected to capture larger-scale structures.
-Rather, this says something about the non-stationarity of the audio signals and the fact that we cannot characterize them by some set of summary statistics.}
-Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
+This suggests that there is relevant variability in scales beyond the window size $T = 372~\mathrm{ms}$ that is not captured by averaging the scattering transform in time (early integration).
+One potential solution to this is to increase the scale of the second-order modulation wavelets from $J_2 = 12$ octaves to some larger value.
+This does not result in any performance gain, however (results not shown).
+What does help is the late integration, since it allows the SVM to successfully classify the few short extracts that characterize the scene which are then used to classify the whole scene through the majority vote.
+Since the scattering transform, even with larger values of $J_2$, is not able to separate out these short, discriminative sound events, delegating this task to the late-integrated results in a more powerful classification system.
 

### 2016-12-17 paper/structureScene.tex
Restructure expt #1 section
Change order of log and MFCCs vs scattering sections. Include 2016
results in the MFCCs vs scattering section. Remove GMM results for
MFCCs.

@@
 
-We report experiments on acoustic scene classification, following the original methodology of the DCASE 2013 challenge, and extend our approach to its 2016 edition.
-
-\subsubsection*{MFCCs \vs scattering}
-
-As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
-This is in accordance with previous results on similarity retrieval, as well as comparisons between mel-frequency coefficients and scattering coefficients for tasks ranging from musical genre recognition and phone segment classification \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.
-
-\subsubsection*{Role of logarithmic compression}
-
-Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
-Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
-This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
-
-\subsubsection*{Early \vs late integration}
-Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
-This suggests that there is relevant variability in scales beyond the window size $T = 372~\mathrm{ms}$ that is not captured by averaging the scattering transform in time (early integration).
-One potential solution to this is to increase the scale of the second-order modulation wavelets from $J_2 = 12$ octaves to some larger value.
-This does not result in any performance gain, however (results not shown).
-What does help is the late integration, since it allows the SVM to successfully classify the few short extracts that characterize the scene which are then used to classify the whole scene through the majority vote.
-Since the scattering transform, even with larger values of $J_2$, is not able to separate out these short, discriminative sound events, delegating this task to the late-integrated results in a more powerful classification system.
-
 \begin{table}
@@
 \begin{center}
-\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline algorithm (left) and proposed algorithm (right).\label{table:dcase2016}}
-\begin{tabular}{lccc}
-Features & MFCCs & MFCCs & log-scattering \\
-Classifier  & GMMs & linear SVMs & linear SVMs \\
-Variant of \emph{late} integration & product & majority vote & majority vote \\
+\caption{Class-wise accuracies on the development part of the DCASE 2016 dataset.\label{table:dcase2016}}
+\begin{tabular}{lcc}
+Features & MFCCs & log-scattering \\
 \midrule
-beach & $74.6 \pm 18.9$ & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ \\
-bus & $58.2 \pm 18.3$ & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ \\
-cafe/restaurant & $85.1 \pm 10.8$ & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ \\
-car & $69.1 \pm 20.8$ & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ \\
-city\_center & $89.6 \pm \phantom{0}9.2$ & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ \\
-forest\_path & $72.4 \pm 23.8$ & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ \\
-grocery\_store & $74.1 \pm 13.7$ & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ \\
-home & $78.1 \pm 17.7$ & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ \\
-library & $65.1 \pm 23.3$ & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ \\
-metro\_station & $85.2 \pm 15.6$ & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ \\
-office & $90.8 \pm 16.0$ & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ \\
-park & $25.6 \pm 11.0$ & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ \\
-residential\_area & $75.1 \pm 19.4$ & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ \\
-train & $34.1 \pm \phantom{0}7.2$ & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ \\
-tram & $85.4 \pm 13.9$ & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ \\
+beach & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ \\
+bus & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ \\
+cafe/restaurant & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ \\
+car & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ \\
+city\_center & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ \\
+forest\_path & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ \\
+grocery\_store & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ \\
+home & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ \\
+library & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ \\
+metro\_station & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ \\
+office & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ \\
+park & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ \\
+residential\_area & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ \\
+train & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ \\
+tram & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ \\
 \bottomrule
-Average & $70.8 \pm \phantom{0}2.6$ & $65.8 \pm \phantom{0}8.0$ & $79.4 \pm \phantom{0}3.0$ \\
+Average & $65.8 \pm \phantom{0}8.0$ & $79.4 \pm \phantom{0}3.0$ \\
 \end{tabular}
@@
 
-\subsubsection*{Participation to the DCASE 2016 challenge}
+We report experiments on acoustic scene classification, following the original methodology of the DCASE 2013 challenge, and extend our approach to its 2016 edition.
 
-In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
-This edition relies on the TUT 2016 acoustic scenes dataset.
+\subsubsection*{Role of logarithmic compression}
 
-Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$, we therefore use a linear kernel instead.
-Our experiments on the smaller DCASE 2013 dataset, as reported in Table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.
+Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
+Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
+This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
 
-Our results on the public part of the DCASE 2016 dataset are reported in Table \ref{table:dcase2016}, compared with a baseline using MFCCs and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model.
-This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
-For comparison, we also provide the results for the same MFCC features coupled with a linear SVM classifier.
-Results here are slightly worse since the SVM can only discriminate between classes that are linearly separable, a limitation not present for the GMMs.
+\subsubsection*{MFCCs \vs scattering}
 
-With the scattering transforms, however, the average accuracy across classes is brought from $70.8\%$ to $79.4\%$.\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.} It is interesting to note, however, that two classes, ``home'' and ``residential\_area'' both drop significantly in accuracy from the baseline GMM model to the scattering SVM classifier.
-This can be partially explained by the substitution of the GMM with an SVM, which shows a similar drop for the MFCCs.
-However, another reason could be that these classes are well-characterized by the average spectral envelope of the signal and the second-order modulation information captured by the scattering coefficients is therefore superfluous.
-Furthermore, our cross-validation analysis confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
+As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
+This is in accordance with previous results on similarity retrieval, as well as comparisons between mel-frequency coefficients and scattering coefficients for tasks ranging from musical genre recognition and phone segment classification \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.
 
-\ml{this section takes one column and is I believe a good candidate for being shrunk or even merged with the previous subsection.
-By statement like, similar findings ...
-or those results are confirmed using the 2016 dataset ...}
+In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, which was evaluated on a much larger dataset compared to DCASE 2013.
+For this reason, training support vector machines with the Gaussian RBF kernel proved computationally too demanding and we instead used a linear kernel.
+Results on DCASE 2013 (see Table \ref{table:dcase2013}) suggests that this provides comparable, albeit slightly worse results.
+
+We compare our results with baseline classifier using linear SVMs and MFCC features, which achieves an average accuracy of $65.5\%$.
+With the scattering transform, however, the average accuracy increases to $79.4\%$.
+This accuracy increases slightly to $80.8\%$ when calculated on the held-out evaluation dataset.\footnote{See \url{http://www.cs.tut.fi/sgn/arg/dcase2016/task-results-acoustic-scene-classification}} The small change in accuracy suggests that there is little overfitting in the model, an indication of good generalization power.
+
+\subsubsection*{Early \vs late integration}
+On the DCASE 2013 dataset, late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
+This suggests that there is relevant variability in scales beyond the window size $T = 372~\mathrm{ms}$ that is not captured by averaging the scattering transform in time (early integration).
+One potential solution to this is to increase the scale of the second-order modulation wavelets from $J_2 = 12$ octaves to some larger value.
+This does not result in any performance gain, however (results not shown).
+What does help is the late integration, since it allows the SVM to successfully classify the few short extracts that characterize the scene which are then used to classify the whole scene through the majority vote.
+Since the scattering transform, even with larger values of $J_2$, is not able to separate out these short, discriminative sound events, delegating this task to the late-integrated results in a more powerful classification system.
 

### 2016-12-17 paper/structureScene.tex
Formatting DCASE 2016 results URL

@@
 \usetikzlibrary{arrows,positioning}
-\usepackage{url}
+\usepackage[hyphens]{url}
 \usepackage{color}
@@
 With the scattering transform, however, the average accuracy increases to $79.4\%$.
-This accuracy increases slightly to $80.8\%$ when calculated on the held-out evaluation dataset.\footnote{See \url{http://www.cs.tut.fi/sgn/arg/dcase2016/task-results-acoustic-scene-classification}} The small change in accuracy suggests that there is little overfitting in the model, an indication of good generalization power.
+This accuracy increases slightly to $80.8\%$ when calculated on the held-out evaluation dataset.\footnote{\url{http://www.cs.tut.fi/sgn/arg/dcase2016/task-results-acoustic-scene-classification}} The small change in accuracy suggests that there is little overfitting in the model, an indication of good generalization power.
 

### 2016-12-17 paper/structureScene.tex
remove alpha from equation 13 and description below

@@
 \widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
-&+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \alpha \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
+&+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
 \end{split}
@@
 %\ja{Can we provide some interpretation for the different terms and constraints here?}
-where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin, and $\alpha$ is a parameter which trades off between \ja{?}.
+where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin.
 We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$.
@@
 The implementations of the presented methods and the experimental protocol are available online\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}}. % \ja{What is ``the one''?}
+
+% , and $\alpha$ is a parameter which trades off between \ja{?}
 

### 2016-12-17 paper/structureScene.tex
part C experiment 1 : clarrify which parameter is learned

@@
 The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
-Optimal parameters (see above) are this time learned on the public part of the DCASE 2013 dataset, also using five-fold stratified cross validation.
-\ja{Which parameters are these?} For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).
+Optimal scaling parameter $q$ of the RBF kernels is  learned on the public part of the DCASE 2013 dataset, using five-fold stratified cross validation.
+For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).
 

### 2016-12-18 paper/structureScene.tex
polishing part C exp 1 : remove "centroids"
rename of repo

@@
 
-The implementations of the presented methods and the experimental protocol are available online\footnote{\url{https://github.com/mathieulagrange/paperStructureScene16}}. % \ja{What is ``the one''?}
+The implementations of the presented methods and the experimental protocol are available online\footnote{\url{https://github.com/mathieulagrange/taslp16}}. % \ja{What is ``the one''?}
 
@@
 For the latter, a local scaling method is applied \cite{selfTuneManor2004}.
-All feature vectors, that is scattering frames, MFCCs and centroids, are standardized prior to be given to the classifier.
+All feature vectors, that is scattering frames and MFCCs, are standardized prior to be given to the classifier.
 The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).

### 2016-12-18 paper/structureScene.tex
update Introduction, citation and "rare"=>"episodic"

@@
 Often, this has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics.
-Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain rare, distinct sound events that discriminate the scene from others.
-Indeed, experiments in cognitive psychology suggests that this matches more closely the way that humans and other animals perceive acoustic scenes.
-\gl{put some Refs that are already cited if possible}
+Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others.
+Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes.
 

### 2016-12-19 paper/structureScene.tex
punctuation in introduction

@@
 \IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
+In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
 In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
@@
 However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
-\ja{Where does the formula $2Q_1/\xi_1$ come from?}
 

### 2016-12-19 paper/structureScene.tex
PhD -> Ph.D. in Joakim's and Mathieu's biographies

@@
 received the M.Sc.
-degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the PhD degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014.
+degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the Ph.D.
+degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014.
 He is currently a postdoctoral researcher with the Program in Applied and Computational mathematics at Princeton University, Princeton, USA.
@@
 \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
-He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, TELECOM ParisTech, Ircam).
+He obtained his Ph.D.
+in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, TELECOM ParisTech, Ircam).
 His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.

### 2016-12-19 experiments/generateFig/nonExposed/parseTable.m
++

@@ -1,0 +1,15 @@
+function [ data,leg ] = parseTable(d)
+
+disp('')
+leg=fieldnames(d.rows{1});
+
+tmp=cellfun(@(x) struct2cell(x),d.rows,'UniformOutput',false);
+
+data=cell(length(tmp{1}),length(tmp));
+
+for ii=1:size(data,1)
+    for jj=1:size(data,2)
+        data{ii,jj}=tmp{jj}{ii};
+    end
+end
+

### 2016-12-19 experiments/generateFig/nonExposed/printFigures.m

@@ -1,0 +1,41 @@
+function [] = printFigures(nbFig,name,opt)
+
+for jj=1:length(nbFig)
+    
+    set(figure(nbFig(jj)),'PaperUnits','centimeters')
+    set(figure(nbFig(jj)),'Units','centimeters')
+    set(figure(nbFig(jj)),'PaperPositionMode','manual')
+    set(figure(nbFig(jj)),'papersize',[opt.width,opt.height])
+    set(figure(nbFig(jj)),'paperposition',[0,0,opt.width,opt.height])
+  
+    
+    
+    set(findall(figure(nbFig(jj)),'-property','FontSize'),'FontSize',opt.fontsize)
+    set(findall(figure(nbFig(jj)),'-property','FontName'),'FontName','Arial')
+    
+    %%
+    
+    if isfield(opt,'markersize')
+        set(findall(figure(nbFig(jj)),'-property','markersize'),'markersize',opt.markersize)
+    end
+    
+    if isfield(opt,'linewidth')
+        set(findall(figure(nbFig(jj)),'-property','linewidth'),'linewidth',opt.linewidth)
+    end
+    
+    %% title
+    
+    if isfield(opt,'axesTitle')
+        allAxesInFigure = findall(figure(nbFig(jj)),'type','axes');
+        for ll=1:length(allAxesInFigure)
+            allAxesInFigure(ll).Title.String=opt.axesTitle{jj}{length(allAxesInFigure)-ll+1}; % reverse order of axes
+        end
+    end
+
+    print(['-f' num2str(nbFig(jj))],[name num2str(nbFig(jj))],'-depsc')
+    print(['-f' num2str(nbFig(jj))],[name num2str(nbFig(jj))],'-dpdf')
+end
+
+
+end
+

### 2016-12-19 experiments/generateFig/supervised.m

@@ -1,0 +1,143 @@
+clear all
+addpath([fileparts(mfilename('fullpath')) '/nonExposed'])
+
+%% settings
+
+dataPath='../talspStruct2016_supervised_earlyLate/report/figures/';
+linewidth=4;
+markersize=10;
+
+%% load data
+
+dataTrainLab={'train_mfcc_early_linear','train_mfcc_late_linear',...
+    'train_scatT_early_linear','train_scatT_late_linear',...
+    'train_logscatT_early_linear','train_logscatT_late_linear',...
+    'train_mfcc_early_rbf','train_mfcc_late_rbf',...
+    'train_scatT_early_rbf','train_scatT_late_rbf',...
+    'train_logscatT_early_rbf','train_logscatT_late_rbf'};
+
+dataTestLab={'test_mfcc_early_linear','test_mfcc_late_linear',...
+    'test_scatT_early_linear','test_scatT_late_linear',...
+    'test_logscatT_early_linear','test_logscatT_late_linear',...
+    'test_mfcc_early_rbf','test_mfcc_late_rbf',...
+    'test_scatT_early_rbf','test_scatT_late_rbf',...
+    'test_logscatT_early_rbf','test_logscatT_late_rbf'};
+
+
+for ii=1:length(dataTrainLab)
+    load([dataPath 'supervised_' dataTrainLab{ii}]);
+    eval([dataTrainLab{ii} '.mean=data.meanData;']);
+    eval([dataTrainLab{ii} '.std=data.stdData;']);
+end
+
+for ii=1:length(dataTestLab)
+    load([dataPath 'supervised_' dataTestLab{ii}]);
+    eval([dataTestLab{ii} '.mean=data.meanData;']);
+    eval([dataTestLab{ii} '.std=data.stdData;']);
+end
+
+disp('')
+
+%% best accuracy train
+
+dataTrainLab={'train_mfcc_early_linear','train_mfcc_late_linear',...
+    'train_scatT_early_linear','train_scatT_late_linear',...
+    'train_mfcc_early_rbf','train_mfcc_late_rbf',...
+    'train_scatT_early_rbf','train_scatT_late_rbf'};
+
+dataTestLab={'test_mfcc_early_linear','test_mfcc_late_linear',...
+    'test_scatT_early_linear','test_scatT_late_linear',...
+    'test_mfcc_early_rbf','test_mfcc_late_rbf',...
[…]

### 2016-12-19 experiments/generateFig/unsupervised.m

@@ -1,0 +1,86 @@
+clear all
+addpath([fileparts(mfilename('fullpath')) '/nonExposed'])
+
+%% settings
+
+dataPath='../talspStruct2016_unsupervised/report/figures/';
+linewidth=1.6;
+markersize=10;
+
+%% load data
+
+dataLab={'mfcc_early','mfcc_clustering_closest','mfcc_clustering_average','mfcc_clustering_emd',...
+    'scatT_early','scatT_clustering_closest','scatT_clustering_average','scatT_clustering_emd',...
+    'logscatT_early','logscatT_clustering_closest','logscatT_clustering_average','logscatT_clustering_emd'};
+
+for ii=1:length(dataLab)
+    load([dataPath 'unsupervised_test_' dataLab{ii}]); 
+    eval([dataLab{ii} '.mean=data.meanData;']); 
+end
+
+disp('')
+
+%% best p@9
+
+for ii=1:length(dataLab)
+    eval(['indMax=find(' dataLab{ii} '.mean(:,end) == max(' dataLab{ii} '.mean(:,end)));']);
+    if length(indMax)>1
+        eval(['indMax=indMax(find(' dataLab{ii} '.mean(indMax,1) == max(' dataLab{ii} '.mean(indMax,1))));'])
+        if length(indMax)>1
+           error('') 
+        end
+    end
+    eval([dataLab{ii} '.meanBest=' dataLab{ii} '.mean(indMax,:);']); 
+end
+
+disp('')
+
+%% figure 
+
+figure(1)
+clf
+plot(scatT_early.meanBest,'d--','markersize',markersize,'linewidth',linewidth)
+hold on
+plot(scatT_clustering_emd.meanBest,'*--','markersize',markersize,'linewidth',linewidth)
+plot(scatT_clustering_closest.meanBest,'x--','markersize',markersize,'linewidth',linewidth)
+plot(logscatT_early.meanBest,'o-','markersize',markersize,'linewidth',linewidth)
+plot(logscatT_clustering_emd.meanBest,'v-','markersize',markersize,'linewidth',linewidth)
+plot(logscatT_clustering_closest.meanBest,'^-','markersize',markersize,'linewidth',linewidth)
+hold off
+legend({'early, scattering','ob-w, scattering','ob-c, scattering','early, log-scattering','ob-w, log-scattering','ob-c, log-scattering'},'interpreter','none')
+legend('boxoff')
[…]

### 2016-12-19 experiments/talspStruct2016_supervised_earlyLate/talspStruct2016_supervised_earlyLate.m

@@ -1,0 +1,190 @@
+function talspStruct2016_supervised_earlyLate(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
+% Welcome to the main entry point of talspStruct2016_supervised_earlyLate                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       
+% Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
+shortExperimentName = 'tasuea';                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
+[p, experimentName] = fileparts(mfilename('fullpath'));                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
[…]

### 2016-12-19 experiments/talspStruct2016_supervised_earlyLate/tasuea1features.m

@@ -1,0 +1,121 @@
+function [config, store, obs] = tasuea1features(config, setting, data)                                     
+% tasuea1features FEATURES step of the expLanes experiment talspStruct2016_supervised_earlyLate            
+%    [config, store, obs] = tasuea1features(config, setting, data)                                         
+%      - config : expLanes configuration state                                                             
+%      - setting   : set of factors to be evaluated                                                        
+%      - data   : processing data stored during the previous step                                          
+%      -- store  : processing data to be saved for the other steps                                         
+%      -- obs    : observations to be saved for analysis                                                   
+                                                                                                           
+% Copyright: gregoirelafay                                                                                 
+% Date: 16-Dec-2016                                                                                        
+                                                                                                           
+% Set behavior for debug mode                                                                              
+if nargin==0, talspStruct2016_supervised_earlyLate('do', 3, 'mask', {0 [1 2] 0 0 0 0}); return; else store=[]; obs=[]; end
+                                                                                                           
+%% setting
+store.xp_settings.sr=44100;
+store.xp_settings.classes = {'bus','busystreet','office','openairmarket','park','quietstreet','restaurant','supermarket','tube','tubestation'};
+
+switch setting.features
+    case 'scatT'
+        
+        fileId = fopen([config.inputPath 'sampleList_' setting.dataset '.txt']);
+        store.xp_settings.sounds=textscan(fileId,'%s');store.xp_settings.sounds=store.xp_settings.sounds{1};
+        fclose(fileId);
+        
+        store.xp_settings.sounds=cellfun(@(x) x(strfind(x,'/')+1:end),store.xp_settings.sounds,'UniformOutput',false);
+        
+        store.soundIndex=[];
+        store.indSample=[];
+        store.class=[];
+        store.dataset=[];
+        eval(['store.features.' setting.features '=[];']);
+        
+                load([config.inputPath  'scattering/dcase2013_timeQ8_' setting.dataset '.mat']);
[…]

### 2016-12-19 experiments/talspStruct2016_supervised_earlyLate/tasuea2preprocessing.m

@@ -1,0 +1,66 @@
+function [config, store, obs] = tasuea2preprocessing(config, setting, data)
+% tasuea2normalization NORMALIZATION step of the expLanes experiment talspStruct2016_supervised_earlyLate
+%    [config, store, obs] = tasuea2preprocessing(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 16-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, talspStruct2016_supervised_earlyLate('do', 3, 'mask', {1 1 1 1 1 1 1}); return; else store=[]; obs=[]; end
+
+%% store
+
+store.xp_settings=data.xp_settings;
+store.soundIndex=data.soundIndex;
+store.class=data.class;
+store.dataset=data.dataset;
+
+%% get features
+
+X=[];
+features2use=strsplit(setting.features,'_');
+for jj=1:length(features2use)
+    eval(['X=[X ; data.features.' features2use{jj} '];']);
+end
+
+%% pooling
+
+switch setting.features
+    case 'mfcc'
+        store.features=[];
+        store.indSample=[];
+        for jj=unique(data.indSample)
+            X_tmp=X(:,data.indSample==jj);
+            step=round(0.250/store.xp_settings.hoptime);
+            onsets=1:step:size(X_tmp,2);
+            offsets=[onsets(2:end)-1 size(X_tmp,2)];
+            X_tmp=featuresPooling(X_tmp,'mean',onsets,offsets,0,0,store.xp_settings.hoptime);
+            store.features = [store.features X_tmp];
+            store.indSample=[store.indSample ones(1,size(X_tmp,2))*jj];
+        end
+    otherwise
+        store.features=X;
+        store.indSample=data.indSample;
+end
+
+clearvars X
+
+%% Rectify
+
+switch setting.features
+    case 'scatT'
+        store.features(store.features<0)=0;
+end
+
+%% integration
+
+switch setting.integration
+    case 'early'
+        store.features=cell2mat(arrayfun(@(x) mean(store.features(:,store.indSample==x),2),unique(store.indSample),'UniformOutput',false));
+        store.indSample=unique(store.indSample);
+end

### 2016-12-19 experiments/talspStruct2016_supervised_earlyLate/tasuea3classifier.m

@@ -1,0 +1,133 @@
+function [config, store, obs] = tasuea3classifier(config, setting, data)
+% tasuea3classifier CLASSIFIER step of the expLanes experiment talspStruct2016_supervised_earlyLate
+%    [config, store, obs] = tasuea3classifier(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 16-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, talspStruct2016_supervised_earlyLate('do', 3, 'mask', {0 [1 2] 1 0 11 0 0}); return; else store=[]; obs=[]; end
+
+
+%% store
+
+store.xp_settings=data.xp_settings;
+store.soundIndex=data.soundIndex;
+store.class=data.class;
+store.dataset=data.dataset;
+
+%% Load folds
+
+fileId = fopen([config.inputPath  'dcase2013_task1_filenamekey.txt']);
+folds=textscan(fileId,'%s %s %s %s','Delimiter',',');
+fclose(fileId);
+
+foldsSoundName=cellfun(@(x) x(1:end-4),folds{4}(2:end),'UniformOutput',false);
+foldsIndex=cellfun(@(x) str2double(x(5)),folds{1}(2:end));
+
+[~,indSort]=sort(foldsSoundName);
+foldsIndex=foldsIndex(indSort);
+foldsLabs=unique(foldsIndex);
+
+switch setting.integration
+    case 'late'
+        
+        foldsIndexSample=repmat(foldsIndex(:)',sum(data.indSample==1),1);
+        sampleClass=repmat(store.class(:)',sum(data.indSample==1),1);
+        
+        foldsIndexSample=foldsIndexSample(:)';
+        sampleClass=sampleClass(:)';
+        
+    case 'early'
+        
+        sampleClass=store.class(:)';
+        
+end
+
+%% svm ftrs
+
+acc_svm_ftrs_train= zeros(1,5);
+acc_svm_ftrs_test= zeros(1,5);
+
+for jj=1:length(foldsLabs)
+    
+    switch setting.integration
+        case 'early'
+            foldIndSampleTest=data.indSample(foldsIndex==foldsLabs(jj));
[…]

### 2016-12-19 experiments/talspStruct2016_supervised_earlyLate/tasueaInit.m

@@ -1,0 +1,10 @@
+function [config, store] = tasueaInit(config)                                              
+% tasueaInit INITIALIZATION of the expLanes experiment talspStruct2016_supervised_earlyLate
+%    [config, store] = tasueaInit(config)                                                  
+%      - config : expLanes configuration state                                             
+%      -- store  : processing data to be saved for the other steps                         
+                                                                                           
+% Copyright: gregoirelafay                                                                 
+% Date: 16-Dec-2016                                                                        
+                                                                                           
+if nargin==0, talspStruct2016_supervised_earlyLate(); return; else store=[];  end          

### 2016-12-19 experiments/talspStruct2016_supervised_earlyLate/tasueaReport.m

@@ -1,0 +1,35 @@
+function config = tasueaReport(config)
+% tasueaReport REPORTING of the expLanes experiment talspStruct2016_supervised_earlyLate
+%    config = tasueaInitReport(config)
+%       config : expLanes configuration state
+
+% Copyright: gregoirelafay
+% Date: 16-Dec-2016
+
+if nargin==0, talspStruct2016_supervised_earlyLate('report', 'r'); return; end
+
+dataset={'train','test'};
+features = {'mfcc','scatT'};
+integration = {'early','late'};
+scat_log = {'','log'};
+similarity = {'linear','rbf'};
+
+for a=1:length(dataset)
+    for b=1:length(features)
+        for c=1:length(integration)
+            for e=1:length(similarity)
+                
+                switch features{b}
+                    case 'mfcc'
+                        config = expExpose(config, 't','fontSize','scriptsize','step', 3, 'mask',{a b c 0 e [1 3 5 7]},'obs',[1 2 3 4 5 6],'precision', 2,'save',1,'name',...
+                            ['supervised_' dataset{a} '_' features{b} '_' integration{c} '_' similarity{e}]);
+                    case 'scatT'
+                        for d=1:length(scat_log)
+                         config = expExpose(config, 't','fontSize','scriptsize','step', 3, 'mask',{a b c d e [1 3 5 7]},'obs',[1 2 3 4 5 6],'precision', 2,'save',1,'name',...
+                            ['supervised_' dataset{a}  '_' scat_log{d} features{b} '_' integration{c} '_' similarity{e}]);                           
+                        end
+                end
+            end
+        end
+    end
+end

### 2016-12-19 experiments/talspStruct2016_unsupervised/talspStruct2016_unsupervised.m

@@ -1,0 +1,190 @@
+function talspStruct2016_unsupervised(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
+% Welcome to the main entry point of talspStruct2016_unsupervised                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               
+% Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
+shortExperimentName = 'taun';                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                   
+[p, experimentName] = fileparts(mfilename('fullpath'));                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
[…]

### 2016-12-19 experiments/talspStruct2016_unsupervised/taun1features.m

@@ -1,0 +1,123 @@
+function [config, store, obs] = taun1features(config, setting, data)
+% taun1features FEATURES step of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store, obs] = taun1features(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, talspStruct2016_unsupervised('do', 1, 'mask', {}); return; else store=[]; obs=[]; end
+
+%% setting
+store.xp_settings.sr=44100;
+store.xp_settings.classes = {'bus','busystreet','office','openairmarket','park','quietstreet','restaurant','supermarket','tube','tubestation'};
+% TODO remove this
+
+
+switch setting.features
+    case 'scatT'
+        
+        fileId = fopen([config.inputPath 'sampleList_' setting.dataset '.txt']);
+        store.xp_settings.sounds=textscan(fileId,'%s');store.xp_settings.sounds=store.xp_settings.sounds{1};
+        fclose(fileId);
+        
+        store.xp_settings.sounds=cellfun(@(x) x(strfind(x,'/')+1:end),store.xp_settings.sounds,'UniformOutput',false);
+        
+        store.soundIndex=[];
+        store.indSample=[];
+        store.class=[];
+        store.dataset=[];
+        eval(['store.features.' setting.features '=[];']);
+        
+                load([config.inputPath 'scattering/dcase2013_timeQ8_' setting.dataset '.mat']);
+                load([config.inputPath 'scattering/dcase2013_timeQ8_freqs.mat']);
+                eval(['Y=dcase2013_timeQ8_' setting.dataset '.Y_' setting.dataset ';']);
+                eval(['X=dcase2013_timeQ8_' setting.dataset '.X_' setting.dataset ';']);
+                eval(['clearvars dcase2013_timeQ8_' setting.dataset ';']);
+                
+                store.xp_settings.Xfreqs=dcase2013_timeQ8_freqs;
+                X=squeeze(X(:,:,3,:));
+                
+        store.class=Y(:)'+1;
+        store.soundIndex=1:length(Y);
+        
+        nbTrame=size(X,2);
+        
+        for ii=1:size(X,3)
[…]

### 2016-12-19 experiments/talspStruct2016_unsupervised/taun2norm.m

@@ -1,0 +1,69 @@
+function [config, store, obs] = taun2norm(config, setting, data)
+% taun2norm NORM step of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store, obs] = taun2norm(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, talspStruct2016_unsupervised('do', 2, 'mask', {}); return; else store=[]; obs=[]; end
+
+%% store
+
+store.xp_settings=data.xp_settings;
+store.soundIndex=data.soundIndex;
+store.class=data.class;
+store.dataset=data.dataset;
+
+%% get features
+
+X=[];
+features2use=strsplit(setting.features,'_');
+for jj=1:length(features2use)
+    eval(['X=[X ; data.features.' features2use{jj} '];']);
+end
+
+%% pooling
+
+switch setting.features
+    case 'mfcc'
+        store.features=[];
+        store.indSample=[];
+        for jj=unique(data.indSample)
+            X_tmp=X(:,data.indSample==jj);
+            step=round(0.250/store.xp_settings.hoptime);
+            onsets=1:step:size(X_tmp,2);
+            offsets=[onsets(2:end)-1 size(X_tmp,2)];
+            X_tmp=featuresPooling(X_tmp,'mean',onsets,offsets,0,0,store.xp_settings.hoptime);
+            store.features = [store.features X_tmp];
+            store.indSample=[store.indSample ones(1,size(X_tmp,2))*jj];
+        end
+    otherwise
+        store.features=X;
+        store.indSample=data.indSample;
+end
+
+clearvars X
+
+%% Rectify
+
+switch setting.features
+    case 'scatT'
+        store.features(store.features<0)=0;
+end
+
+%% Normalisation
+
+switch setting.features
+    case 'scatT'
+        params.type='scattering';
+        params.ftrsNorm_scat_threshold = 1e2;
+        params.ftrsNorm_scat_selection = 1;
+        params.ftrsNorm_scat_log =  setting.scat_log;
+        [store.features,~] = normFtrs(store.features,params);
+end

### 2016-12-19 experiments/talspStruct2016_unsupervised/taun3clustering.m

@@ -1,0 +1,75 @@
+function [config, store, obs] = taun3clustering(config, setting, data)
+% taun3clustering CLUSTERING step of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store, obs] = taun3clustering(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, talspStruct2016_unsupervised('do', 3, 'mask', {}); return; else store=[]; obs=[]; end
+
+%% seed
+
+rng(0);
+
+%% store
+
+store.xp_settings=data.xp_settings;
+store.soundIndex=data.soundIndex;
+store.class=data.class;
+store.dataset=data.dataset;
+
+store.weight=[];
+store.centroid=[];
+store.indSample=[];
+store.centroidClass=[];
+store.prediction=[];
+
+%% integration
+
+switch setting.integration
+    case 'clustering'
+        
+        params.clustering='kmeans';
+
+        for jj=unique(data.indSample)
+            
+            params.similarity='sqeuclidean';
+            params.nbc=setting.clustering_nbc;
+            params.rep=1;
+            params.emptyAction='singleton';
+            
+            [prediction,centroid] = kmeans(full(data.features(:,data.indSample==jj))',params.nbc,'maxiter',1000,'replicates',params.rep,'start','plus','Distance',params.similarity,'EmptyAction',params.emptyAction);
+            centroid=centroid';
+            %[prediction,centroid,params] = featuresBasedClustering(data.features(:,data.indSample==jj),params);
+            
+            %% weight
+            
+            params.histType='';
+            params.nbc=size(centroid,2);
+            [weight] = histFeatures(prediction,ones(1,length(prediction)),params);
+            
+            %% Norm
+            
+            % weight=weight./sum(weight);
+            
+            %% store
+            store.prediction=[store.prediction prediction+setting.clustering_nbc*(jj-1)];
+            store.weight=[store.weight weight];
+            store.centroid=[store.centroid centroid];
+            store.indSample=[store.indSample ones(1,size(centroid,2))*jj];
[…]

### 2016-12-19 experiments/talspStruct2016_unsupervised/taun4similarity.m

@@ -1,0 +1,62 @@
+function [config, store, obs] = taun4similarity(config, setting, data)
+% taun4similarity SIMILARITY step of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store, obs] = taun4similarity(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, talspStruct2016_unsupervised('do', 4, 'mask', {}); return; else store=[]; obs=[]; end
+
+%% store
+
+store.xp_settings=data.xp_settings;
+store.soundIndex=data.soundIndex;
+store.class=data.class;
+store.dataset=data.dataset;
+
+store.weight=data.weight;
+store.indSample=data.indSample;
+store.centroid=data.centroid;
+store.centroidClass=data.centroidClass;
+store.prediction=data.prediction;
+
+%% get clusters Similarity
+
+params.mode='centroid';
+params.similarity='rbf';
+params.nn=setting.similarity_nn;
+params.class=store.centroidClass;
+
+simClus= computeSimilarity(store.centroid,params);
+
+switch setting.integration
+    
+    case 'clustering'
+        
+        %% reshape weights
+        
+        onsets=1:size(data.weight,1):size(simClus,2)-size(data.weight,1)+1;
+        offsets=[onsets(2:end)-1 size(simClus,2)];
+        weight=zeros(size(simClus,2),size(data.weight,2));
+        
+        for jj=1:size(weight,2)
+            weight(onsets(jj):offsets(jj),jj)=data.weight(:,jj);
+        end
+        
+        %% scenes similarity
+        
+        params.histDist=setting.similarity_dist;
+        params.indSample=store.indSample;
+        
+        [store.A,store.params] = histSimilarity(weight,params,simClus);
+        
+    otherwise
+        
+        store.A = simClus;
+end

### 2016-12-19 experiments/talspStruct2016_unsupervised/taun5metrics.m

@@ -1,0 +1,22 @@
+function [config, store, obs] = taun5metrics(config, setting, data)                                
+% taun5metrics METRICS step of the expLanes experiment talspStruct2016_unsupervised                
+%    [config, store, obs] = taun5metrics(config, setting, data)                                    
+%      - config : expLanes configuration state                                                     
+%      - setting   : set of factors to be evaluated                                                
+%      - data   : processing data stored during the previous step                                  
+%      -- store  : processing data to be saved for the other steps                                 
+%      -- obs    : observations to be saved for analysis                                           
+                                                                                                   
+% Copyright: gregoirelafay                                                                         
+% Date: 17-Dec-2016                                                                                
+                                                                                                   
+% Set behavior for debug mode                                                                      
+if nargin==0, talspStruct2016_unsupervised('do', 5, 'mask', {2 [1 2] 0 11 2 0 0 3}); return; else store=[]; obs=[]; end
+                                                                                                   
+D=1-data.A/max(data.A(:));
+k2look=[1 2 3 4 5 6 7 8 9];
+
+for jj=k2look
+    rm = rankingMetrics(D, data.class,jj);
+    eval(['obs.pa' num2str(jj) '=rm.precisionAt' num2str(jj) ';']);
+end

### 2016-12-19 experiments/talspStruct2016_unsupervised/taunInit.m

@@ -1,0 +1,10 @@
+function [config, store] = taunInit(config)                                      
+% taunInit INITIALIZATION of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store] = taunInit(config)                                          
+%      - config : expLanes configuration state                                   
+%      -- store  : processing data to be saved for the other steps               
+                                                                                 
+% Copyright: gregoirelafay                                                       
+% Date: 17-Dec-2016                                                              
+                                                                                 
+if nargin==0, talspStruct2016_unsupervised(); return; else store=[];  end        

### 2016-12-19 experiments/talspStruct2016_unsupervised/taunReport.m

@@ -1,0 +1,45 @@
+function config = taunReport(config)
+% taunReport REPORTING of the expLanes experiment talspStruct2016_unsupervised
+%    config = taunInitReport(config)
+%       config : expLanes configuration state
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+if nargin==0, talspStruct2016_unsupervised('report', 'r'); return; end
+
+features = {'mfcc','scatT'};
+scat_log = {'','log'};
+integration = {'early','clustering'};
+similarity_dist = {'emd','average','closest'};
+
+for a=1:length(features)
+    for c=1:length(integration)   
+        switch integration{c} 
+            case 'early'
+                switch features{a}
+                    case 'mfcc'
+                        config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a 0 c 0 [1 2 3] 0},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                            ['unsupervised_test_' features{a} '_' integration{c}]);
+                    case 'scatT'
+                        for b=1:length(scat_log)
+                            config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a b c 0 [1 2 3] 0},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                                ['unsupervised_test_' scat_log{b} features{a} '_' integration{c}]);
+                        end
+                end  
+            case 'clustering'
+                for d=1:length(similarity_dist)
+                    switch features{a}
+                        case 'mfcc'
+                            config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a 0 c 0 [1 2 3] d},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                                ['unsupervised_test_' features{a} '_' integration{c} '_' similarity_dist{d}]);
+                        case 'scatT'
+                            for b=1:length(scat_log)
+                                config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a b c 0 [1 2 3] d},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                                    ['unsupervised_test_' scat_log{b} features{a} '_' integration{c} '_' similarity_dist{d}]);
+                            end
+                    end  
+                end
+        end
+    end
+end

### 2016-12-19 paper/structureScene.tex
finish Vincent's biography

@@
 \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lostanlen}}]{Vincent Lostanlen} was born in 1992.
-He received an engineering degree from TELECOM ParisTech in 2013 and M.S.
-degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie and the Ircam (UPMC) in Paris, France, in 2013.
+He received an engineering degree from TELECOM ParisTech in 2013 and his M.S.
+degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie (UPMC) and the Ircam in Paris, France, in 2013.
 Since 2013, he is a Ph.D.
 student at \'{E}cole normale sup\'{e}rieure in Paris, working under the supervision of St\'{e}phane Mallat.
+His research focuses on defining convolutional operators in the time-frequency domain with applications in audio classification.
 \end{IEEEbiography}

### 2016-12-19 paper/responseToReviewers.tex
done with response

@@
 
-Again, the hierarchical representation proposed is novel and we provide adequate evidence, in our opinion, for its power in the proposed similarity retrieval task in the case of environmental audio data processing.
+Again, the hierarchical representation proposed is, to the best of our knowledge, novel and we provide adequate evidence, in our opinion, for its power in the proposed similarity retrieval task in the case of environmental audio data processing.
 
@@
 We have reworked the paper as a whole, trying to clarify the overall message.
-
-\ml{I think we could do it better, see my issue}
+In particular, we reorganized the paper towards a unique claim, that is supported by experimental evidence.
+the contribution now explicitly target the acoustic similarity retrieval task.
+Complementary experiments in the more widely used classification task are used to validate the first step of representation using the scattering transform.
 
@@
 
-We acknowldege that such a diagram would help the reader, we are facing page limit contraints due to the extensive experiments decribed in the paper.
+We acknowledge that such a diagram would help the reader, we are facing page limit constraints due to the extensive experiments described in the paper.
 
@@
 Experiments have been performed in the Matlab environment using open source libraries an all the authors are committed to comply to the reproducible research statements.
-Unfortunately, due to time constraints, we will not be able to deliver the code at the time of submission of the revision.
-Nevertheless, it will be fully available at the time of publication if the revision is accepted for publication.
+The replication code is now available in beta, and will be doubled checked very soon.
 

### 2016-12-19 paper/structureScene.tex
dot at the end of keywords
start improving the abstract

@@
 \begin{abstract}
-
-An important task in the emerging field of computational ecoacoustics is to determine the similarity of two audio scenes to be able to get a high level understanding of large amount of audio recordings coming from large deployment of acoustics sensors that are continuously recordings the sonic environment.
+The emerging field of computational ecoacoustics aims at retrieving high-level information from the continuous, environmental recordings of acoustic sensors.
+In this context, a central task is to cluster auditory scenes by similarity.
 A convenient method considers bag-of-frames models and other representations based on summary statistics.
@@
 To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure.
-To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional feature transforms like mel-frequency cepstral coefficients (MFCCs).
+To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional features, such as mel-frequency cepstral coefficients (MFCCs).
 Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.
@@
 \begin{IEEEkeywords}
-acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance
+acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance.
 \end{IEEEkeywords}

### 2016-12-19 paper/structureScene.tex
cleaning outputs

--- 2016-12-19 paper/structureScene.tex
@@
 \IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
+In order to measure the effect of human activity and climate change on animal biodiversity researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
 In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
@@
 
-A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{cortes-vapnik}.
+A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{Cortes1995}.
 From another point of view, it attemts to find a direction on which to project in order to maximize separation between two point clouds.
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{bw/folds.png}
-\caption{\label{fig:folds}
-Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=\columnwidth]{bw/folds.png}
+%\caption{\label{fig:folds}
+%Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.
 Performance on each fold is depicted using a given symbol.
-The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
-The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.}
-\end{center}
-\end{figure}
+%The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
+%The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.}
+%\end{center}
+%\end{figure}
 
@@
 
-As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
+%As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
+
+As shown on Table \ref{table:dcase2013}, scattering coefficients outperform MFCCs on the DCASE 2013 private dataset when applying early and integration.
 This is in accordance with previous results on similarity retrieval, as well as comparisons between mel-frequency coefficients and scattering coefficients for tasks ranging from musical genre recognition and phone segment classification \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.
@@
 received the M.Sc.
-degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the Ph.D.
-degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014.
+degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the PhD degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014.
 He is currently a postdoctoral researcher with the Program in Applied and Computational mathematics at Princeton University, Princeton, USA.
@@
 \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
-He obtained his Ph.D.
-in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, TELECOM ParisTech, Ircam).
+He obtained his PhD in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, TELECOM ParisTech, Ircam).
 His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.

### 2016-12-19 experiments/generateFig/nonExposed/printFigures.m

@@ -1,7 +1,7 @@
 function [] = printFigures(nbFig,name,opt)
 
 for jj=1:length(nbFig)
-    
+    colormap('gray')
     set(figure(nbFig(jj)),'PaperUnits','centimeters')
     set(figure(nbFig(jj)),'Units','centimeters')
     set(figure(nbFig(jj)),'PaperPositionMode','manual')

### 2016-12-19 experiments/generateFig/unsupervisedFig.m

@@ -3,7 +3,7 @@ addpath([fileparts(mfilename('fullpath')) '/nonExposed'])
 
 %% settings
 
-dataPath='../talspStruct2016_unsupervised/report/figures/';
+dataPath='../unsupervised/report/figures/';
 linewidth=1.6;
 markersize=10;
 
@@ -31,6 +31,7 @@ for ii=1:length(dataLab)
         end
     end
     eval([dataLab{ii} '.meanBest=' dataLab{ii} '.mean(indMax,:);']); 
+     eval([dataLab{ii} '.meanBest=' dataLab{ii} '.meanBest(1, :);']); % ML
 end
 
 disp('')
@@ -47,7 +48,7 @@ plot(logscatT_early.meanBest,'o-','markersize',markersize,'linewidth',linewidth)
 plot(logscatT_clustering_emd.meanBest,'v-','markersize',markersize,'linewidth',linewidth)
 plot(logscatT_clustering_closest.meanBest,'^-','markersize',markersize,'linewidth',linewidth)
 hold off
-legend({'early, scattering','ob-w, scattering','ob-c, scattering','early, log-scattering','ob-w, log-scattering','ob-c, log-scattering'},'interpreter','none')
+legend({'early, scattering','RbQ-w, scattering','RbQ-c, scattering','early, log-scattering','RbQ-w, log-scattering','RbQ-c, log-scattering'},'interpreter','none')
 legend('boxoff')
 xlabel('k')
 ylabel('p@k')
@@ -57,6 +58,7 @@ disp('')
 
 figure(2)
 clf
+
 plot(mfcc_early.meanBest,'o--','markersize',markersize,'linewidth',linewidth)
 hold on
 plot(mfcc_clustering_average.meanBest,'*--','markersize',markersize,'linewidth',linewidth)
@@ -70,8 +72,8 @@ plot(logscatT_clustering_emd.meanBest,'v-','markersize',markersize,'linewidth',l
 plot(logscatT_clustering_closest.meanBest,'^-','markersize',markersize,'linewidth',linewidth)
 
 hold off
-legend({'early, mfcc','ob-a, mfcc','ob-w, mfcc','ob-c, mfcc', ...
-    'early, log-scattering','ob-a, log-scattering','ob-w, log-scattering','ob-c, log-scattering'},'interpreter','none')
+legend({'early, mfcc','RbQ-a, mfcc','RbQ-w, mfcc','RbQ-c, mfcc', ...
[…]

### 2016-12-19 experiments/supervised/supervised.m

@@ -1,8 +1,8 @@
-function talspStruct2016_unsupervised(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
-% Welcome to the main entry point of talspStruct2016_unsupervised                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               
+function supervised(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
+% Welcome to the main entry point of talspStruct2016_supervised_earlyLate                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       
 % Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
[…]

### 2016-12-19 experiments/unsupervised/taun1features.m

@@ -22,7 +22,7 @@ store.xp_settings.classes = {'bus','busystreet','office','openairmarket','park',
 switch setting.features
     case 'scatT'
         
-        fileId = fopen([config.inputPath 'sampleList_' setting.dataset '.txt']);
+        fileId = fopen([config.inputPath 'sampleList_test.txt']);
         store.xp_settings.sounds=textscan(fileId,'%s');store.xp_settings.sounds=store.xp_settings.sounds{1};
         fclose(fileId);
         
@@ -31,14 +31,13 @@ switch setting.features
         store.soundIndex=[];
         store.indSample=[];
         store.class=[];
-        store.dataset=[];
         eval(['store.features.' setting.features '=[];']);
         
-                load([config.inputPath 'scattering/dcase2013_timeQ8_' setting.dataset '.mat']);
+                load([config.inputPath 'scattering/dcase2013_timeQ8_test.mat']);
                 load([config.inputPath 'scattering/dcase2013_timeQ8_freqs.mat']);
-                eval(['Y=dcase2013_timeQ8_' setting.dataset '.Y_' setting.dataset ';']);
-                eval(['X=dcase2013_timeQ8_' setting.dataset '.X_' setting.dataset ';']);
-                eval(['clearvars dcase2013_timeQ8_' setting.dataset ';']);
+                eval(['Y=dcase2013_timeQ8_test.Y_test;']);
+                eval(['X=dcase2013_timeQ8_test.X_test;']);
+                eval(['clearvars dcase2013_timeQ8_test ;']);
                 
                 store.xp_settings.Xfreqs=dcase2013_timeQ8_freqs;
                 X=squeeze(X(:,:,3,:));
@@ -72,7 +71,7 @@ switch setting.features
         
         %% select sound
         
-        fileId = fopen([config.inputPath  'sampleList_' setting.dataset '.txt']);
+        fileId = fopen([config.inputPath  'sampleList_test.txt']);
         store.xp_settings.sounds=textscan(fileId,'%s');store.xp_settings.sounds=store.xp_settings.sounds{1};
         fclose(fileId);
         
@@ -85,7 +84,6 @@ switch setting.features
         
         store.soundIndex=zeros(1,length(store.xp_settings.soundIndex));
         store.class=zeros(1,length(store.xp_settings.soundIndex));
-        store.dataset=zeros(1,length(store.xp_settings.soundIndex));
         store.indSample=[];
         
         for jj=1:length(store.xp_settings.soundIndex)

### 2016-12-19 experiments/unsupervised/taun2norm.m

@@ -18,7 +18,6 @@ if nargin==0, talspStruct2016_unsupervised('do', 2, 'mask', {}); return; else st
 store.xp_settings=data.xp_settings;
 store.soundIndex=data.soundIndex;
 store.class=data.class;
-store.dataset=data.dataset;
 
 %% get features
 

### 2016-12-19 experiments/unsupervised/taun3clustering.m

@@ -22,7 +22,6 @@ rng(0);
 store.xp_settings=data.xp_settings;
 store.soundIndex=data.soundIndex;
 store.class=data.class;
-store.dataset=data.dataset;
 
 store.weight=[];
 store.centroid=[];

### 2016-12-19 experiments/unsupervised/taun4similarity.m

@@ -18,7 +18,6 @@ if nargin==0, talspStruct2016_unsupervised('do', 4, 'mask', {}); return; else st
 store.xp_settings=data.xp_settings;
 store.soundIndex=data.soundIndex;
 store.class=data.class;
-store.dataset=data.dataset;
 
 store.weight=data.weight;
 store.indSample=data.indSample;

### 2016-12-19 experiments/unsupervised/taun5metrics.m

@@ -11,7 +11,7 @@ function [config, store, obs] = taun5metrics(config, setting, data)
 % Date: 17-Dec-2016                                                                                
                                                                                                    
 % Set behavior for debug mode                                                                      
-if nargin==0, talspStruct2016_unsupervised('do', 5, 'mask', {2 [1 2] 0 11 2 0 0 3}); return; else store=[]; obs=[]; end
+if nargin==0, unsupervised('do', 5, 'mask', {}); return; else store=[]; obs=[]; end
                                                                                                    
 D=1-data.A/max(data.A(:));
 k2look=[1 2 3 4 5 6 7 8 9];

### 2016-12-19 experiments/unsupervised/taunReport.m

@@ -6,7 +6,7 @@ function config = taunReport(config)
 % Copyright: gregoirelafay
 % Date: 17-Dec-2016
 
-if nargin==0, talspStruct2016_unsupervised('report', 'r'); return; end
+if nargin==0, unsupervised('report', 'r'); return; end
 
 features = {'mfcc','scatT'};
 scat_log = {'','log'};
@@ -19,11 +19,11 @@ for a=1:length(features)
             case 'early'
                 switch features{a}
                     case 'mfcc'
-                        config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a 0 c 0 [1 2 3] 0},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                        config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{a 0 c 0 [1 2 3] 0},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
                             ['unsupervised_test_' features{a} '_' integration{c}]);
                     case 'scatT'
                         for b=1:length(scat_log)
-                            config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a b c 0 [1 2 3] 0},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                            config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{a b c 0 [1 2 3] 0},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
                                 ['unsupervised_test_' scat_log{b} features{a} '_' integration{c}]);
                         end
                 end  
@@ -31,11 +31,11 @@ for a=1:length(features)
                 for d=1:length(similarity_dist)
                     switch features{a}
                         case 'mfcc'
-                            config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a 0 c 0 [1 2 3] d},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                            config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{a 0 c 0 [1 2 3] d},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
                                 ['unsupervised_test_' features{a} '_' integration{c} '_' similarity_dist{d}]);
[…]

### 2016-12-19 experiments/unsupervised/unsupervised.m

@@ -1,8 +1,8 @@
-function talspStruct2016_supervised_earlyLate(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
-% Welcome to the main entry point of talspStruct2016_supervised_earlyLate                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       
+function taslpStruct2016_unsupervised(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
+% Welcome to the main entry point of talspStruct2016_unsupervised                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               
 % Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
[…]

### 2016-12-19 paper/structureScene.tex
moving from ob* to RbQ*

@@
 \begin{itemize}
-\item \emph{ob-closest} (\emph{ob-c}): the similarity between two scenes $s_u$ and $s_v$ is equal to the largest similarity between their centroids, that is
+\item Relevance-based Quantization closest similarity (\emph{RbQ-c}): the similarity between two scenes $s_u$ and $s_v$ is equal to the largest similarity between their centroids, that is
 \begin{equation}
@@
 \end{equation}
-\item \emph{ob-averaged} (\emph{ob-a}): the similarity between two scenes $s_u$ and $s_v$ is equal to the average of their centroid similarities, that is
+\item Relevance-based Quantization average similarity  (\emph{RbQ-a}): the similarity between two scenes $s_u$ and $s_v$ is equal to the average of their centroid similarities, that is
 \begin{equation}
@@
 and,
-\item \emph{ob-weighted} (\emph{ob-w}): the similarity between two scenes is computed using a variant of the earth mover's distance applied to the set of centroids weighted by the number of frames assigned to its cluster.
+\item Relevance-based Quantization weighted similarity  (\emph{RbQ-w}): the similarity between two scenes is computed using a variant of the earth mover's distance applied to the set of centroids weighted by the number of frames assigned to its cluster.
 \end{itemize}
 
-For \emph{ob-w}, each centroid is weighted according to the number of frames belonging to its cluster.
+For \emph{RbQ-w}, each centroid is weighted according to the number of frames belonging to its cluster.
 Each scene $s_u$ is thus described by a signature $p_u$ of $M$ clusters, where $p_u=\lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace$) and $\mu_m^u$ and $w_m^u$ are the centroid and the weight of the $m$th cluster, respectively.
@@
 
-Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach.
+Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{RbQ-c} approach.
 Overall, log-compressed scattering coefficients systematically outperform MFCCs.
@@
 
-For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization to improve the similarity measures between the scenes.
-However, it is worth noticing that \emph{ob-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
+For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (RbQ) to improve the similarity measures between the scenes.
+However, it is worth noticing that \emph{RbQ-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
 To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.
-Furthermore, it appears that \emph{ob-c} is better able to characterize the classes than \emph{ob-w}.
+Furthermore, it appears that \emph{RbQ-c} is better able to characterize the classes than \emph{RbQ-w}.
 This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
@@
 The same observations are made for the MFCCs for $k\leq5$.
-However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently.
+However for a rank $k$ greater than $5$, all the \emph{RbQ} approaches perform equivalently.
 As such, one can conclude that the use of MFCCs allows us to retrieve closely related scenes, but their lack of discriminative capabilities making the cluster selection/weighting step useless for not so closely related scenes of the same class

### 2016-12-19 paper/structureScene.tex
rephrase abstract

@@
 \begin{abstract}
-The emerging field of computational ecoacoustics aims at retrieving high-level information from the continuous, environmental recordings of acoustic sensors.
-In this context, a central task is to cluster auditory scenes by similarity.
+The emerging field of computational ecoacoustics aims at retrieving high-level information from the acoustic scenes recorded by some network of sensors.
+In this context, a central task is to cluster acoustic scenes by similarity.
 A convenient method considers bag-of-frames models and other representations based on summary statistics.

### 2016-12-19 paper/structureScene.tex
finish refactoring abstract

@@
 The emerging field of computational ecoacoustics aims at retrieving high-level information from the acoustic scenes recorded by some network of sensors.
-In this context, a central task is to cluster acoustic scenes by similarity.
-A convenient method considers bag-of-frames models and other representations based on summary statistics.
-While these successfully characterize static scenes with little variability in sound content, they are unable to accurately differentiate scenes characterized by a few distinct sound events superimposed over a static background.
-To address this problem, we propose a two-scale representation, which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure.
-To motivate this, we validate the use of scattering transforms for supervised acoustic scene classification (ASC), illustrating its state-of-the-art performance compared to traditional features, such as mel-frequency cepstral coefficients (MFCCs).
-Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) task, where we again significantly outperform more traditional approaches based on summary statistics of MFCCs.
-
+In this paper, we address the problem of acoustic scene classification (ASC) according to the taxonomy of the DCASE 2013 challenge, both in supervised and unsupervised settings.
+Traditional approaches compute a "bag-of-frames", i.e. the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
+Although they successfully characterize static scenes with little variability in auditory content, they are unable to accurately discriminate scenes with few salient events superimposed over a textured background.
+To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure.
+We validate the use of the scattering transform for supervised acoustic scene classification by reporting a better accuracy than MFCCs when using a support vector machine (SVM) classifier.
+Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) unsupervised task, where we significantly outperform more traditional approaches based on summary statistics of MFCCs.
 \end{abstract}
@@
 This requires the auditory scenes to be represented in some way that captures their distinctive qualities.
-Often, this has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics.
+Often, this has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics.
 Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others.

### 2016-12-19 paper/responseToReviewers.tex
rephrase reproducibility status
cosmetic changes to response to reviewers

@@
 
-\item \emph{Line 40, alpha may appear in previous equations}
+\item \emph{Line 40, $\alpha$ may appear in previous equations}
 
@@
 
-\item \emph{The argumentation toward the log compression is well conducted, and meaningful in this issue. The authors may simply swap the order of fig 3 and 4}
+\item \emph{The argumentation toward the log compression is well conducted, and meaningful in this issue. The authors may simply swap the order of Fig. 3 and 4.}
 
@@
 
-\item \emph{Complete legend fig 5.}
+\item \emph{Complete caption of Fig. 5.}
 
-The legend have been completed to help the readability of the figure.
+The caption have been completed to help the readability of the figure.
 
@@
 
-\item \emph{line 60: auditory and modulation filter – nonsense.}
+\item \emph{Line 60: auditory and modulation filter – nonsense.}
 
@@
 
-\item \emph{Setting of Q1 shall be motivated.}
+\item \emph{Setting of $Q1$ shall be motivated.}
 
@@
 Experiments have been performed in the Matlab environment using open source libraries an all the authors are committed to comply to the reproducible research statements.
-The replication code is now available in beta, and will be doubled checked very soon.
+The replication code is now available in beta, and will be soon stable.
 
-\item \emph{In the last experience (Table III) GMM(MFCC) may be compared to SVM(MFCC) (in addition to SVM(logscat)). There is somewhere a high dimension effect that shall be clarified.}
+\item \emph{In the last experience (Table III) GMM (MFCC) may be compared to SVM (MFCC) (in addition to SVM (logscat)). There is somewhere a high dimension effect that shall be clarified.}
 
@@
 
-\item \emph{A table with the 2016 CASE private results may be added (or replace Tab III).}
+\item \emph{A table with the 2016 DCASE private results may be added (or replace Tab III).}
 
@@
 
-\item \emph{Check typos in the references}
+\item \emph{Check typos in the references.}
 

### 2016-12-19 paper/structureScene.tex
doubles as -> is also

@@
 has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$ and a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$ resulting in the same quality factor $Q_1$ as $\boldsymbol{\psi}$.
-In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$, $J_1$ to $10$, and the quality factor $Q_1$, which doubles as the number of wavelets per octave, to $8$.
+In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$, $J_1$ to $10$, and the quality factor $Q_1$, which is also the number of wavelets per octave, to $8$.
 This results in the wavelet filters covering the whole range of human hearing, from $20~\mathrm{Hz}$ to $20~\mathrm{kHz}$.

### 2016-12-19 paper/structureScene.tex
remove ja's comment on computational cost

@@
 frequencies than the CQT, at the expense of a greater computational cost
-since the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank. \ja{Is the cost really greater?}
+since the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
 However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.

### 2016-12-19 paper/structureScene.tex
sampled at 44100 Hz

@@
 The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} and DCASE 2016 \cite{Mesaros2016_EUSIPCO} datasets.
-The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes.
+The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes, sampled at $44100\,\mathrm{Hz}$.
 The $100$ recordings are evenly divided into $10$ acoustic scene classes.

### 2016-12-19 paper/structureScene.tex
paper -> article

@@
 The emerging field of computational ecoacoustics aims at retrieving high-level information from the acoustic scenes recorded by some network of sensors.
-In this paper, we address the problem of acoustic scene classification (ASC) according to the taxonomy of the DCASE 2013 challenge, both in supervised and unsupervised settings.
+In this article, we address the problem of acoustic scene classification (ASC) according to the taxonomy of the DCASE 2013 challenge, both in supervised and unsupervised settings.
 Traditional approaches compute a "bag-of-frames", i.e. the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).

### 2016-12-20 paper/structureScene.tex
Remove references to ASC and ASSR

--- 2016-12-19 paper/structureScene.tex
@@
 The emerging field of computational ecoacoustics aims at retrieving high-level information from the acoustic scenes recorded by some network of sensors.
-In this paper, we address the problem of acoustic scene classification (ASC) according to the taxonomy of the DCASE 2013 challenge, both in supervised and unsupervised settings.
+In this article, we address the problem of acoustic scene classification according to the taxonomy of the DCASE 2013 challenge, both in supervised and unsupervised settings.
 Traditional approaches compute a "bag-of-frames", i.e. the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
@@
 We validate the use of the scattering transform for supervised acoustic scene classification by reporting a better accuracy than MFCCs when using a support vector machine (SVM) classifier.
-Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval (ASSR) unsupervised task, where we significantly outperform more traditional approaches based on summary statistics of MFCCs.
+Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval unsupervised task, where we significantly outperform more traditional approaches based on summary statistics of MFCCs.
 \end{abstract}
@@
 An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation.
-One application of this is acoustic scene similarity retrieval (ASSR), where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset.
+One application of this is acoustic scene similarity retrieval, where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset.
 This requires the auditory scenes to be represented in some way that captures their distinctive qualities.

### 2016-12-20 paper/structureScene.tex
Remove references to ``ob'' approaches

@@
 Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item.
-For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $M$ for the \emph{ob} approaches and the scaling parameter $q$ of the RBF kernels (see Eq.
+For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $M$ for the \emph{RbQ} approaches and the scaling parameter $q$ of the RBF kernels (see Eq.
 \ref{eq:kc}) leading to the best $p@9$.
-The \emph{ob} approaches are compared to commonly used early integration approach \emph{early}.
+The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}.
 %\ja{And BOF?}
 
-The clustering for the \emph{ob} methods is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure.
+The clustering for the \emph{RbQ} methods is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure.
 Three numbers of clusters are tested: $8$, $16$ and $32$.

### 2016-12-20 paper/structureScene.tex
Edit abstract

@@
 We validate the use of the scattering transform for supervised acoustic scene classification by reporting a better accuracy than MFCCs when using a support vector machine (SVM) classifier.
-Furthermore, we illustrate the power of the full cluster-based representation by applying it to an acoustic scene similarity retrieval unsupervised task, where we significantly outperform more traditional approaches based on summary statistics of MFCCs.
+Furthermore, we illustrate the power of the full cluster-based representation by applying it to an unsupervised acoustic scene similarity retrieval task, where we significantly outperform more traditional approaches based on summary statistics of MFCCs.
 \end{abstract}

### 2016-12-20 paper/structureScene.tex
Edit intro

@@
 
-To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}.
+To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.
+This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}.
 In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.
 
-Adressing this task requires first to be able to capture meanningful aspects within the signal at a small time scale.
+Adressing this task requires first to be able to capture meaningful aspects within the signal at small time scales.
 In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
-A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environemental audio processing \cite{aucouturier2007bag}.
+A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
 A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.

### 2016-12-20 paper/structureScene.tex
Shorten intro slightly

@@
 The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it almost completely.
-Combining scattering transforms for low-level representation with late-integrated SVMs used for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.
+Combining scattering transforms for low-level representation with late-integrated SVMs for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.
 

### 2016-12-20 paper/structureScene.tex
Edit background section

@@
 %The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-In a closely related field, namely the analysis of the urban sound environment, a popular approach is the bag-of-frames (BOF) introduced in the field by Aucouturier et al.
+In the closely related field of urban sound environment analysis, a popular approach is the bag-of-frames (BOF) introduced in this context by Aucouturier et al.
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
-Though, recently, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
+Recently, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
 This contrasts with the typical morphology of acoustic scenes, which we believe is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events are superimposed upon a stationary acoustic background.
@@
 
-The methods that performed the best all followed a similar structure, where a low-level feature representation, such as MFCCs, were calculated over short windows and fed separately to a classifier, such as an SVM, for training.
+The methods that performed the best all followed a similar structure, where low-level feature representations, such as MFCCs, were calculated over short windows and fed separately to a classifier, such as an SVM, for training.
 During the classification stage, the recording to be classified was similarly processed into short windows represented by feature vectors which were classified individually by the classifier and the classifier output was combined over the recording through majority vote.

### 2016-12-20 paper/structureScene.tex
Edit scattering section

@@
 Indeed, recordings may be shifted locally without affecting their original class memberships.
-To discard this superfluous source of variability, signals can be mapped to a time-shifting invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance and increasing the robustness of its predictions.
+To discard this superfluous source of variability, signals are mapped into a time-shifting invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance and increasing the robustness of its predictions.
 
@@
 Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events.
-These make up an important part of intra-class variability among natural sounds, so a representation that has some insensitivity to such transformations would be very advantageous.
+These make up an important part of intra-class variability among natural sounds, so it is important for representation to be robust to such transformations.
 
@@
 
-To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\phi(t)$ with cut-off frequency $1/T$ (and so a duration of approximately $T$), to get
+To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\phi(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
 \begin{equation}
@@
 which is known as the set of first-order scattering coefficients.
-They capture the average spectral envelope of $\boldsymbol{x}$ over scales of duration $T$, with the spectral resolution varying with constant Q, as described above.
-In this way, they are very closely related to the mel-frequency spectrogram and related features, such as MFCCs.
+They capture the average spectral envelope of $\boldsymbol{x}$ over scales of duration $T$, with spectral resolution varying with constant Q.
+In this way, they are closely related to the mel-frequency spectrogram and related features, such as MFCCs.
 
 \subsection{Extracting modulations with second-order scattering}
-In auditory scenes, amplitude modulations may be caused by a broad variety of rapid mechanical interactions, including collision, friction, turbulent flow, and so on.
+In auditory scenes, short-time amplitude modulations may be caused by a variety of rapid mechanical interactions, including collision, friction, turbulent flow, and so on.
 At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
@@
 The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
-While higher-order scattering coefficients can be calculated, for the purposes of our current, work, the first and second order are sufficient.
+While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
 

### 2016-12-20 paper/structureScene.tex
Edit supervised section

@@
 
-A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{Cortes1995}.
-From another point of view, it attemts to find a direction on which to project in order to maximize separation between two point clouds.
+/ support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{Cortes1995}.
+From another point of view, it attempts to find a direction on which to project in order to maximize separation between two point clouds.
 Since it is not a generative, but discriminative model, it is able to pick out those feature vectors that best characterize a particular class of auditory scenes compared to another.
@@
 Finally, the SVM can also be extended to handle non-linear decision surfaces.
-This is obtained using the ``kernel trick,'' which exploits the fact that all calculations in the SVM rely only on the inner product between vectors.
+This is obtained using the ``kernel trick'', which exploits the fact that all calculations in the SVM rely only on the inner product between vectors.
 By replacing this standard inner product by some other positive definite symmetric mapping, such as a Gaussian radial basis function (RBF), this is equivalent to performing a linear SVM after mapping the feature vectors into some high-dimensional space.

### 2016-12-20 paper/structureScene.tex
Edit ASSR section

@@
 As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
-While this can be addressed in the supervised case using late integration of discriminative classifiers (see Section \ref{sec:eili}), this is not direclty feasible in the unsupervised case.
+While this can be addressed in the supervised case using late integration of discriminative classifiers (see Section \ref{sec:eili}), this is not directly feasible in the unsupervised case.
 As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
@@
 
-The similarity between the scene centroids $\mu_m^u$ over the entire dataset, is computed using a radial basis function (RBF) kernel $K$ combined with the local scaling method proposed in \cite{selfTuneManor2004}:
+The similarity between the scene centroids $\mu_m^u$ over the entire dataset, is computed using a radial basis function (RBF) kernel $K$ combined with a local scaling method \cite{selfTuneManor2004}:
 \begin{equation}
@@
 and,
-\item Relevance-based Quantization weighted similarity  (\emph{RbQ-w}): the similarity between two scenes is computed using a variant of the earth mover's distance applied to the set of centroids weighted by the number of frames assigned to its cluster.
+\item Relevance-based Quantization weighted similarity  (\emph{RbQ-w}): the similarity between two scenes is computed using a variant of the earth mover's distance applied to the set of centroids each weighted by the number of frames assigned to its cluster.
 \end{itemize}
@@
 where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin.
-We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$.
-$D^{uv}$ is computed from the kernel $K$:
+We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$ which is computed from $K$:
 \begin{equation*}
@@
 \end{equation*}
-As suggested in \cite{pele2009fast}, we set the tradeoff parameter $\alpha$ to $1$.
 

### 2016-12-20 paper/structureScene.tex
Edit experiments section

@@
 
-The implementations of the presented methods and the experimental protocol are available online\footnote{\url{https://github.com/mathieulagrange/taslp16}}. % \ja{What is ``the one''?}
+The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/taslp16}} % \ja{What is ``the one''?}
 
@@
 The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
-Optimal scaling parameter $q$ of the RBF kernels is  learned on the public part of the DCASE 2013 dataset, using five-fold stratified cross validation.
+The optimal scaling parameter $q$ of the RBF kernels is  learned on the public part of the DCASE 2013 dataset, using five-fold stratified cross validation.
 For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).

### 2016-12-20 paper/structureScene.tex
Edit results

@@
 However for a rank $k$ greater than $5$, all the \emph{RbQ} approaches perform equivalently.
-As such, one can conclude that the use of MFCCs allows us to retrieve closely related scenes, but their lack of discriminative capabilities making the cluster selection/weighting step useless for not so closely related scenes of the same class
+As such, one can conclude that the use of MFCCs allows us to retrieve closely related scenes, but their lack of discriminative capabilities making the cluster selection/weighting step useless for not so closely related scenes of the same class.
 

### 2016-12-20 paper/structureScene.tex
Add comment about dimensionality

@@
 
+While the scattering coefficients have higher dimension compared to the MFCCs, this does not necessarily imply higher performance.
+Indeed, if the sounds to be classified were completely classified by their spectral envelopes, MFCCs would be sufficient and may even outperform the scattering transforms, since in this case the latter would contain superfluous modulation information.
+The results above, however, show the opposite tendency, where this additional information does in fact improve results.
+
 In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, which was evaluated on a much larger dataset compared to DCASE 2013.

### 2016-12-20 paper/responseToReviewers.tex
response 16 to reviewer 1
Add comment about dimensionality
Edit response

@@
 In addition, most of our experiments deal with the feature vectors through an RBF kernel, which is equivalent to considering the features embedded in an infinite-dimensional space, so in this sense the original dimensionality of the feature vectors carries less importance.
+A paragraph has been added to discuss the influence of dimensionality.
 
@@
 The function $\phi$ is very much a low-pass filter with a frequency support of approximately $1/T$ and a time bandwidth of $T$, where $T$ is of the order of hundreds of milliseconds.
-It does not correspond to the traditional scaling function used in quadratic mirror filters.
+It does not correspond to the traditional scaling function used in quadrature mirror filters.
 
@@
 \item \emph{It is never specified what are the signals frequency sampling which we have to deduce to be 44100 Hz from line 9 page 3 "20kHz close to the Nyquist frequency of the audio recordings".}
+
+It is true that the audio recordings from the DCASE 2013 and 2016 datasets are sampled at 44100 Hz. This information has been added in Section VII.A.
 
@@
 
-Experiments have been performed in the Matlab environment using open source libraries an all the authors are committed to comply to the reproducible research statements.
+Experiments have been performed in the MATLAB environment using open source libraries an all the authors are committed to comply to the reproducible research statements.
 The replication code is now available in beta, and will be soon stable.
@@
 We have added results for SVM on MFCCs.
-As mentioned in the response to the other reviewer, we believe that the high dimensionality should not have a significant effect on these results, even-though we acknowledge that this effect would be interesting to study.
+As mentioned in the response to the other reviewer, we believe that the high dimensionality should not have a significant effect on these results, even though we acknowledge that this effect would be interesting to study.
 Our justification have been added to the manuscript.
@@
 
+The private results have been reported in terms of the average class accuracy, but due to lack of space, we have not included the per-class results.
+However, we do provide a link to the website where these can be viewed.
+
 \item \emph{The bibliography is huge, it may be reduced by 10 \%.}

### 2016-12-20 paper/structureScene.tex
edit results Table I
edit results supervised: "albeit consistent" => "over all but one"
edit RbQ results

@@
              \hline
-\emph{early}, linear  & $54\pm20$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
-\emph{early}, RBF     & $47\pm16$  & $56\pm\phantom{0}7$  & $63\pm\phantom{0}6$   \\
+\emph{early}, linear  & $54\pm19$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
+\emph{early}, RBF     & $47\pm16$  & $61\pm\phantom{0}9$  & $63\pm\phantom{0}6$   \\
 \emph{late}, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
-\emph{late}, RBF     & $68\pm10$ & $73\pm\phantom{0}8$  & $\mathbf{78}\pm\phantom{0}6$   \\
+\emph{late}, RBF     & $68\pm14$ & $73\pm\phantom{0}8$  & $\mathbf{78}\pm\phantom{0}6$   \\
 \end{tabular}
@@
 
-Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
+Logarithmic compression provides a small boost to acoustic scene classification over all but one fold of the DCASE 2013 private dataset.
 Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
@@
 For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (RbQ) to improve the similarity measures between the scenes.
-However, it is worth noticing that \emph{RbQ-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
+However, it is worth noticing that \emph{RbQ-a} performs worse or comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
 To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.

### 2016-12-20 experiments/experiment_code/nonexposed/computeSimilarity.m
addition of dependency code

@@ -1,0 +1,27 @@
+function [ A ] = computeSimilarity(X,param)
+
+X=full(X');
+
+switch param.similarity
+    
+    case {'cosine'}
+        D=squareform(pdist(X,param.similarity));
+        A=1-D/(max(D(:)));  
+    case 'linear'
+        A = X * X';    
+    case 'rbf'
+        D = dist2(X, X);
+        nn=max([1 round(param.nn*size(D,2))]);
+        [ A ] = rbfKernel(D,'st-1nn',nn); 
+        
+    otherwise
+        error('invalid similarity setting.')
+
+end
+
+clearvars Xnorm d D;
+
+A = (A + A') / 2;
+
+end
+

### 2016-12-20 experiments/experiment_code/nonexposed/histFeatures.m

@@ -1,0 +1,19 @@
+function [ sceneFeatures ] = histFeatures(prediction,indSample,params)
+
+c=1:params.nbc;
+nbc=params.nbc;
+
+sampleLabels=unique(indSample);
+sceneFeatures=zeros(nbc,length(sampleLabels));
+
+for jj=1:length(sampleLabels)
+    sceneFeatures(:,jj)=hist(prediction(indSample==sampleLabels(jj)),c);
+end
+
+switch params.histType
+    case 'pres'
+        sceneFeatures(sceneFeatures>0)=1;
+    case 'thresh'
+        sceneFeatures(sceneFeatures<=params.thresh)=0;      
+end
+disp('')

### 2016-12-20 experiments/experiment_code/nonexposed/normFtrs.m

@@ -1,0 +1,29 @@
+function [features,setting] = normFtrs(features,param)
+
+setting.type=param.type;
+
+
+switch param.type
+    
+    case 'stand'
+        
+        if isfield(param,'mu')
+            
+            features=bsxfun(@minus,features,param.mu)./repmat(param.sd,1,size(features,2));
+            
+        else
+            
+            setting.mu=mean(features,2);
+            setting.sd=std(features,0,2);
+            features=bsxfun(@minus,features,setting.mu)./repmat(setting.sd,1,size(features,2));
+            
+        end
+        
+    case {'scattering'}
+        
+        [ features,setting] = normScattering(features,param);
+        
+end
+
+end
+

### 2016-12-20 experiments/experiment_code/nonexposed/voteMaj.m

@@ -1,0 +1,13 @@
+function [newPred,newPred2] = voteMaj(prediction,indSample)
+
+labSample=unique(indSample);
+predLab=unique(prediction);
+newPred=zeros(1,length(labSample));
+newPred2=zeros(size(prediction));
+
+for jj=1:length(labSample)
+    count=hist(prediction(indSample==labSample(jj)),predLab);
+    [~,newPred(jj)]=max(count);
+    newPred2(indSample==labSample(jj))=newPred(jj);
+end
+

### 2016-12-20 paper/structureScene.tex
edit supervised results
edit GL bio

@@
 \emph{late}, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
-\emph{late}, RBF     & $68\pm14$ & $73\pm\phantom{0}8$  & $\mathbf{78}\pm\phantom{0}6$   \\
+\emph{late}, RBF     & $68\pm14$ & $73\pm\phantom{0}6$  & $\mathbf{78}\pm\phantom{0}6$   \\
 \end{tabular}
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in Paris, France, in 1990.
-He received in 2011 the B.S.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in 1990.
+He received the B.S.
 degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, France, and the B.S.
-degree in Musicology from the Sorbonne University, Paris, France.
+degree in Musicology from the Sorbonne University, Paris, France, in 2011.
 He received his M.S.
-degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre and Marie Curie and the Ircam in Paris, France, in 2013.
+degree in acoustics, signal processing, and musical informatics (ATIAM) from the UPMC and the Ircam in Paris, France, in 2013.
 Since 2013, he is a Ph.D.
-student at IRCCyN, a French laboratory dedicated to cybernetics.
+student at IRCCyN, Nantes, France.
 His research interests include acoustic scene similarity and classification as well as acoustic scene perception.

### 2016-12-20 paper/structureScene.tex
Add DCASE2016 class-wise results

@@
 \begin{center}
-\caption{Class-wise accuracies on the development part of the DCASE 2016 dataset.\label{table:dcase2016}}
-\begin{tabular}{lcc}
-Features & MFCCs & log-scattering \\
+\caption{Class-wise accuracies on the DCASE 2016 dataset.\label{table:dcase2016}}
+\begin{tabular}{lccc}
+Features & MFCCs & log-scattering & log-scattering \\
+Subset & development & development & evaluation \\
 \midrule
-beach & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ \\
-bus & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ \\
-cafe/restaurant & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ \\
-car & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ \\
-city\_center & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ \\
-forest\_path & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ \\
-grocery\_store & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ \\
-home & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ \\
-library & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ \\
-metro\_station & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ \\
-office & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ \\
-park & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ \\
-residential\_area & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ \\
-train & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ \\
-tram & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ \\
+beach & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ & $80.8$\\
+bus & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ & $92.3$ \\
+cafe/restaurant & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ & $50.0$ \\
+car & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ & $96.2$ \\
+city\_center & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ & $84.6$ \\
+forest\_path & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ & $96.2$ \\
+grocery\_store & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ & $84.6$ \\
+home & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ & $80.8$ \\
+library & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ & $65.4$ \\
+metro\_station & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ & $96.2$ \\
+office & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ & $100$ \\
+park & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ & $65.4$ \\
+residential\_area & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ & $69.2$ \\
+train & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ & $53.8$ \\
+tram & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ & $96.2$ \\
 \bottomrule
-Average & $65.8 \pm \phantom{0}8.0$ & $79.4 \pm \phantom{0}3.0$ \\
+Average & $65.8 \pm \phantom{0}8.0$ & $79.4 \pm \phantom{0}3.0$ & $80.8$ \\
 \end{tabular}
@@
 Results on DCASE 2013 (see Table \ref{table:dcase2013}) suggests that this provides comparable, albeit slightly worse results.
+The results on DCASE 2016 (see Table \ref{table:dcase2016}).
 
@@
 With the scattering transform, however, the average accuracy increases to $79.4\%$.
-This accuracy increases slightly to $80.8\%$ when calculated on the held-out evaluation dataset.\footnote{\url{http://www.cs.tut.fi/sgn/arg/dcase2016/task-results-acoustic-scene-classification}} The small change in accuracy suggests that there is little overfitting in the model, an indication of good generalization power.
+This accuracy increases slightly to $80.8\%$ when calculated on the held-out evaluation dataset.\footnote{\url{http://www.cs.tut.fi/sgn/arg/dcase2016/task-results-acoustic-scene-classification}} The small change in overall accuracy and relatively stable class-wise accuracies suggest that there is little overfitting in the model, an indication of good generalization power.
 

### 2016-12-20 experiments/deprecated/supervised/convnet.py
delete convnet.py

@@ -1,104 +1,0 @@
-from keras.models import Sequential
-from keras.layers import Activation, Dense, Dropout, Flatten
-from keras.layers import Convolution2D, MaxPooling2D
-from keras.layers.advanced_activations import LeakyReLU
-from keras.optimizers import Adam
-
-import h5py
-import numpy as np
-import sklearn.preprocessing
-
-method = 'time'
-hdf5_file = h5py.File("memoized_features/" + method + "_Q=08.mat")
-hdf5_file.keys()
-hdf5_group = hdf5_file[u'data']
-
-X_training = hdf5_group[u'X_scattergrams_train']
-X_test = hdf5_group[u'X_scattergrams_test']
-
-X_training = np.transpose(X_training, (1, 0, 2, 4, 3))
-X_training = np.reshape(X_training,
-    (X_training.shape[0]*X_training.shape[1], X_training.shape[2],
-    X_training.shape[3], X_training.shape[4]))
-X_test = np.transpose(X_test, (1, 0, 2, 4, 3))
-X_test = np.reshape(X_test,
-    (X_test.shape[0]*X_test.shape[1], X_test.shape[2],
-    X_test.shape[3], X_test.shape[4]))
-
-X_training = np.log(X_training[:, 0, :, :] + 10.0)
-X_test = np.log(X_test[:, 0, :, :] + 10.0)
-
-# offset = - np.min(X_training[:])
-# X_training = X_training + offset + 0.1
-# X_test = X_test + offset + 0.1
-# _, lambda_ = scipy.stats.boxcox(np.ravel(X_training))
-# X_training = (X_training ** lambda_ - 1.0) / lambda_
-# X_test = (X_test ** lambda_ - 1.0) / lambda_
-
-X_mean = np.mean(X_training)
-X_std = np.std(X_training)
-X_training = ((X_training - X_mean) / (1e-6 + X_std))
-X_test = ((X_test - X_mean) / (1e-6 + X_std))
-
-Y_training = np.repeat(hdf5_group[u'Y_train'], 5)
-Y_training = np.eye(10)[Y_training.astype('int')]
-Y_test = np.repeat(hdf5_group[u'Y_test'], 5)
-Y_test = np.eye(10)[Y_test.astype('int')]
-
-scattering_channels = 1
-conv_height = 5
-conv_width = 5
-conv_channels = 16
-pool_height = 2
-pool_width = 3
-dense_channels = 32
-
-alpha = 0.3  # for LeakyReLU
-
-init = "he_normal"
-batch_size = 32
[…]

### 2016-12-20 experiments/deprecated/supervised/supervised_script.m
update supervised_script

@@ -1,8 +1,7 @@
-% Before running this script, the libraries scattering.m and rastamat must
-% have been added to the path
+% Before running this script, the library scattering.m must have been added
+% to the MATLAB path
 dataset_path = '~/datasets/dcase2013/scenes_stereo';
 Q1 = 8;
-nAzimuths = 5;
 
 dcase_scattering(dataset_path, Q1);
 dcase_scattering([dataset_path, '_testset'], Q1);

### 2016-12-20 experiments/deprecated/supervised/dcase_scattering.m
bugfix dcase_scattering

@@ -19,16 +19,8 @@ opts{1}.time.wavelet_handle = @gammatone_1d;
 opts{1}.time.S_log2_oversampling = 0;
 opts{2}.banks.time.nFilters_per_octave = 1;
 opts{2}.banks.time.wavelet_handle = @gammatone_1d;
-if strcmp(modulations, 'time')
-    opts{2}.banks.time.sibling_mask_factor = 2^6;
-elseif strcmp(modulations, 'timefrequency')
-    opts{2}.banks.time.sibling_mask_factor = 2^3;
-end
+opts{2}.banks.time.sibling_mask_factor = 2^6;
 opts{2}.banks.time.T = 2^17;
-if strcmp(modulations, 'timefrequency')
-    opts{2}.banks.gamma.nFilters_per_octave = 1;
-    opts{2}.banks.gamma.T = 2^nextpow2(Q1 * 4);
-end
 opts{2}.invariants.time.size = 2^19;
 opts{2}.invariants.time.T = 2^14;
 opts{2}.invariants.time.subscripts = 1;
@@ -38,7 +30,7 @@ opts{3}.invariants.time.subscripts = 1;
 archs = sc_setup(opts);
 
 % Prepare paths for memoized features
-out_file_name = [modulations, '_Q=', num2str(Q1, '%0.2d')];
+out_file_name = ['_Q=', num2str(Q1, '%0.2d')];
 if strcmp(dataset_path((end-6):end), 'testset')
     out_file_name = [out_file_name, '_test'];
 else

### 2016-12-20 paper/structureScene.tex
edit Results log compression
Fix standard deviation
more on the point conclusion

@@
 
-This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
+This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause, wimmer2013sampling}.
 Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
@@
 
-Logarithmic compression provides a small boost to acoustic scene classification over all but one fold of the DCASE 2013 private dataset.
+On the DCASE 2013 dataset, logarithmic compression provides a small, albeit consistent boost to acoustic scene classification.
 Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
@@
 For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
-These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in computational bioacoustics.
+These outcomes shall be studied further in future work by considering larger databases and emerging tasks in computational bioacoustics \cite{wimmer2013sampling}.
 

### 2016-12-21 paper/structureScene.tex
move table 1
done with revision

@@
 
+\begin{table}
+\begin{center}
+\caption{
+\label{table:dcase2013}
+Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.
+}
+\begin{tabular}{llll}
+             & MFCCs         & scattering & log-scattering  \\
+             \hline
+\emph{early}, linear  & $54\pm19$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
+\emph{early}, RBF     & $47\pm16$  & $61\pm\phantom{0}9$  & $63\pm\phantom{0}6$   \\
+\emph{late}, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
+\emph{late}, RBF     & $68\pm14$ & $73\pm\phantom{0}6$  & $\mathbf{78}\pm\phantom{0}6$   \\
+\end{tabular}
+\end{center}
+\end{table}
+
 \subsection{Experiment \#1: acoustic scene classification}
@@
 Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item.
-For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $M$ for the \emph{RbQ} approaches and the scaling parameter $q$ of the RBF kernels (see Eq.
-\ref{eq:kc}) leading to the best $p@9$.
 The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}.
@@
 
-The clustering for the \emph{RbQ} methods is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure.
-Three numbers of clusters are tested: $8$, $16$ and $32$.
+The scaling parameter $q$ of the RBF kernels (see Eq.
+\ref{eq:kc}) is set to 10 \% of the number of data points to cluster, being the number of scenes for \emph{early} and the number of cluster for RbQ approaches.
+The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and 200 replications.
+For each method, the numbers of clusters is set to the value among $8$, $16$ and $32$ that leads to the best $p@9$ on the training set.
 
@@
 %\end{figure}
-
-\begin{table}
-\begin{center}
-\caption{
-\label{table:dcase2013}
-Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.
-}
-\begin{tabular}{llll}
-             & MFCCs         & scattering & log-scattering  \\
-             \hline
-\emph{early}, linear  & $54\pm19$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
-\emph{early}, RBF     & $47\pm16$  & $61\pm\phantom{0}9$  & $63\pm\phantom{0}6$   \\
-\emph{late}, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
-\emph{late}, RBF     & $68\pm14$ & $73\pm\phantom{0}6$  & $\mathbf{78}\pm\phantom{0}6$   \\
-\end{tabular}
-\end{center}
-\end{table}
 
@@
 \bottomrule
-Average & $65.8 \pm \phantom{0}8.0$ & $79.4 \pm \phantom{0}3.0$ & $80.8$ \\
+Average & $65.8 \pm 19.3$ & $79.4 \pm 15.6$ & $80.8 \pm 16.4$ \\
 \end{tabular}
@@
 However, it is worth noticing that \emph{RbQ-a} performs worse or comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
+This result is in line with the findings of \cite{lagrange:hal-01082501}.
 To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.
@@
 
-The same observations are made for the MFCCs for $k\leq5$.
+Considering the $p@5$ as in \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} as our metric, the use of the log-scattering transform versus MFCCs increases the performance  from 0.31 to 0.49 and the use of the relevance based quantization approach using the closest similarity (RbQ-c) further improves the performance to 0.54 for a global increase of 0.23.
+
+%The same observations are made for the MFCCs for $k\leq5$.
 However for a rank $k$ greater than $5$, all the \emph{RbQ} approaches perform equivalently.
@@
 \begin{center}
-\includegraphics[width=.9\columnwidth]{bw/unsupervised_test2}
+\includegraphics[width=\columnwidth]{bw/unsupervised_test2}
 \caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}

### 2017-05-01 paper/structureScene.tex
addition of reference pdfs
fix small mistake in section IV intro
typo

@@
-\documentclass[journal]{IEEEtran}
+\documentclass[onecolumn,journal]{IEEEtran}
 %\usepackage{cite}
@@
 \bibliography{biblio}
-%\renewcommand{\baselinestretch}{2}
+\renewcommand{\baselinestretch}{2}
 
@@
 Prior to constructing models (for classification or similarity estimation), it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
-In this section, we review three transformations which achieve these properties, namely logarithmic compression and standardization.
+In this section, we review two transformations which achieve these properties, namely logarithmic compression and standardization.
 
@@
 
-/ support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{Cortes1995}.
+A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{Cortes1995}.
 From another point of view, it attempts to find a direction on which to project in order to maximize separation between two point clouds.

### 2017-05-01 paper/structureScene.tex
non-linear --> nonlinear

@@
 
-Scattering transforms are time-shift invariant representations of audio signals computed by applying auditory and modulation filter banks alternated with complex modulus non-linearities.
+Scattering transforms are time-shift invariant representations of audio signals computed by applying auditory and modulation filter banks alternated with complex modulus nonlinearities.
 This section first explains the importance of local invariance to time-shifting and stability to time-warping in the representation of acoustic scenes, then describes how the scattering transform is designed to satisfy these properties while having a high discriminative power.
@@
 
-Finally, the SVM can also be extended to handle non-linear decision surfaces.
+Finally, the SVM can also be extended to handle nonlinear decision surfaces.
 This is obtained using the ``kernel trick'', which exploits the fact that all calculations in the SVM rely only on the inner product between vectors.
 By replacing this standard inner product by some other positive definite symmetric mapping, such as a Gaussian radial basis function (RBF), this is equivalent to performing a linear SVM after mapping the feature vectors into some high-dimensional space.
-A non-linear mapping then results in a non-linear decision surface in the orignal space.
+A nonlinear mapping then results in a nonlinear decision surface in the orignal space.
 

### 2017-06-04 paper/structureScene.tex
minor change to ref to Table DCASE 2016
update keywords

@@
 \begin{IEEEkeywords}
-acoustic scene classification, acoustic scene similarity retrieval, cluster models, wavelets, scattering, invariance.
+acoustic scene classification, acoustic scene similarity retrieval, scattering transform, clustering, similarity measures.
 \end{IEEEkeywords}
@@
 Results on DCASE 2013 (see Table \ref{table:dcase2013}) suggests that this provides comparable, albeit slightly worse results.
-The results on DCASE 2016 (see Table \ref{table:dcase2016}).
+The results on DCASE 2016 are reported on Table \ref{table:dcase2016}.
 

### 2017-06-07 paper/structureScene.tex
twocolumn
disable double spacing
update keywords

@@
-\documentclass[onecolumn,journal]{IEEEtran}
+\documentclass{IEEEtran}
 %\usepackage{cite}
@@
 \bibliography{biblio}
-\renewcommand{\baselinestretch}{2}
+%\renewcommand{\baselinestretch}{2}
 
@@
 \begin{IEEEkeywords}
-acoustic scene classification, acoustic scene similarity retrieval, scattering transform, clustering, similarity measures.
+unsupervised learning, classification algorithms, clustering algorithms, acoustic signal processing, wavelet transforms, audio databases, content-based retrieval, nearest neighbor searches, acoustic sensors, environmental sensors.
 \end{IEEEkeywords}

### 2017-06-26 paper/structureScene.tex
comment on scattering and end-to-end learning

@@
 A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
+This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network.
+As a result, the scattering transform benefits from some of the assets of end-to-end learning without needing a large amount of annotated training data to converge towards a meaningful representation.
 

### 2017-06-26 paper/structureScene.tex
prior to -> before

@@
 
-Prior to constructing models (for classification or similarity estimation), it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
+Before constructing models for classification or similarity estimation, it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
 In this section, we review two transformations which achieve these properties, namely logarithmic compression and standardization.

### 2017-06-26 paper/structureScene.tex
update link to github repo

@@
 
-The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/taslp16}} % \ja{What is ``the one''?}
+The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/tetci17}} % \ja{What is ``the one''?}
 

### 2017-06-26 paper/structureScene.tex
spelling

@@
 
-Adressing this task requires first to be able to capture meaningful aspects within the signal at small time scales.
+Addressing this task requires first to be able to capture meaningful aspects within the signal at small time scales.
 In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
 A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
-A richer set of features are found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
+A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
 This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network.

### 2017-06-26 paper/structureScene.tex
do not hyphenate nearest neighbor, rendered -> made

@@
 
-Many algorithms in pattern recognition, including nearest-neighbor classifiers and support vector machines (SVMs), tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
+Many algorithms in pattern recognition, including nearest neighbor classifiers and SVMs, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
 Yet the distribution of the scattering coefficients is skewed towards larger values.
@@
 Since it is not a generative, but discriminative model, it is able to pick out those feature vectors that best characterize a particular class of auditory scenes compared to another.
-This property, coupled with its ease of use and efficient implementations \cite{libsvm} have rendered SVMs a popular choice for many classification systems.
+This property, coupled with its ease of use and efficient implementations \cite{libsvm} have made SVMs a popular choice for many classification systems.
 

### 2017-06-26 paper/structureScene.tex
typo
delete paragraph on dimensionality to fit 10 pages

@@
 By replacing this standard inner product by some other positive definite symmetric mapping, such as a Gaussian radial basis function (RBF), this is equivalent to performing a linear SVM after mapping the feature vectors into some high-dimensional space.
-A nonlinear mapping then results in a nonlinear decision surface in the orignal space.
+A nonlinear mapping then results in a nonlinear decision surface in the original space.
 
@@
 
-While the scattering coefficients have higher dimension compared to the MFCCs, this does not necessarily imply higher performance.
-Indeed, if the sounds to be classified were completely classified by their spectral envelopes, MFCCs would be sufficient and may even outperform the scattering transforms, since in this case the latter would contain superfluous modulation information.
-The results above, however, show the opposite tendency, where this additional information does in fact improve results.
-
 In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, which was evaluated on a much larger dataset compared to DCASE 2013.

### 2017-09-12 paper/structureScene.tex
rewrite abstract

@@
 %\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features}
+\title{Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio}
 
@@
 The emerging field of computational ecoacoustics aims at retrieving high-level information from the acoustic scenes recorded by some network of sensors.
-In this article, we address the problem of acoustic scene classification according to the taxonomy of the DCASE 2013 challenge, both in supervised and unsupervised settings.
-Traditional approaches compute a "bag-of-frames", i.e. the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
+Large amount of unlabeled data are gathered using those networks that needs to be analyzed.
+In order to identify which parts needs human inspection, one needs tools that can mine data and identify structural such as recurring patterns or on contrary sparse occurences.
+To do so, one needs to be able to measure the level of similarity among the data without strong assumptions.
+
+State of the art approaches for computing such a similarity is the "bag-of-frames" approach, i.e.
+modeling the audio using the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
 Although they successfully characterize static scenes with little variability in auditory content, they are unable to accurately discriminate scenes with few salient events superimposed over a textured background.
-To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure.
-We validate the use of the scattering transform for supervised acoustic scene classification by reporting a better accuracy than MFCCs when using a support vector machine (SVM) classifier.
-Furthermore, we illustrate the power of the full cluster-based representation by applying it to an unsupervised acoustic scene similarity retrieval task, where we significantly outperform more traditional approaches based on summary statistics of MFCCs.
+To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure that better handle the required modeling of scarce events.
+Experimentation performed using the Acoustic Scene Similarity framework demonstrates the superiority of the proposed approach.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-unsupervised learning, classification algorithms, clustering algorithms, acoustic signal processing, wavelet transforms, audio databases, content-based retrieval, nearest neighbor searches, acoustic sensors, environmental sensors.
+unsupervised learning, data mining, acoustic signal processing, wavelet transforms, audio databases, content-based retrieval, nearest neighbor searches, acoustic sensors, environmental sensors.
 \end{IEEEkeywords}
@@
 This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause, wimmer2013sampling}.
-Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
-Consequently, few well-designed datasets are readily available for evaluation purposes.
+%Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
+%Consequently, few well-designed datasets are readily available for evaluation purposes.
 
 An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation.
-One application of this is acoustic scene similarity retrieval, where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset.
+
+To this aim, one needs to design an algorithm to perform an acoustic scene similarity retrieval, where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset.
 This requires the auditory scenes to be represented in some way that captures their distinctive qualities.
-Often, this has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics.
+This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics.
 Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others.
 Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes.
+We believe that the failure to model such distinct events is one of the reason why this approach is found to be not sufficient when considering \cite{}.
 
@@
 A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
-This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network.
-As a result, the scattering transform benefits from some of the assets of end-to-end learning without needing a large amount of annotated training data to converge towards a meaningful representation.
+This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{} \vl{some relevant refs on dnns 4 audio}.
+
+For the task at at hand, the scattering features have the valuable advantage of not requiring training and thus potentially lead to wider range of application than trained.
+Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be
 

### 2017-09-12 paper/structureScene.tex
rewrite of intro step 1

@@
 
-This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause, wimmer2013sampling}.
+This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
 %Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
@@
 
-An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation.
+An important problem in this field is that analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling} and needs some sort of pre-screening to reduce the need for human expert listening and annotation.
 
-To this aim, one needs to design an algorithm to perform an acoustic scene similarity retrieval, where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset.
-This requires the auditory scenes to be represented in some way that captures their distinctive qualities.
-This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics.
+The most straightforward way is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}.
+By doing so, a period of time, say a day is represented as the number of events within the predefined typology that have been detected in this period of time.
+This approach allows the scientist to reduce drastically the amount of information that needs to be processed.
+Though, this approach have two drawbacks.
+The first is the reliance to trained models whose behavior on unseen data (different sensors, ...) is prone to errors and can hardly be asserted.
+And, more importantly, it is based on \textit{a priori} knowledge, and thus cannot be considered for exploratory analysis where the quantities of interest are yet to be defined.
+
+and annotation to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors.
+The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.
+
+To this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset.
+This requires the audio fragments to be represented in some way that captures their distinctive qualities.
+This has been achieved by bag-of-frames approaches in the litterature\cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics.
 Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others.
 Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes.
-We believe that the failure to model such distinct events is one of the reason why this approach is found to be not sufficient when considering \cite{}.
+We believe that the failure to model such distinct events is one of the reason why this approach is found to be not sufficient when considering decent size evaluation datasets \cite{lagrange:hal-01082501}.
 
-To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.
+%To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.
 This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}.
@@
 
-For the task at at hand, the scattering features have the valuable advantage of not requiring training and thus potentially lead to wider range of application than trained.
-Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be
+For the task at at hand, the scattering features have the valuable advantage of not requiring training and thus potentially lead to wider range of application than trained features.
+Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined.
+Thus leading to an unserpervised setting.
 
-The next question is the choice of a classifier.
+%The next question is the choice of a classifier.
 Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM) \cite{cristianini2000introduction}, to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision.
@@
 
-Returning to scene retrieval setting, we cannot use supervised models such as SVMs, but we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
-We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters.
+%we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
+We also propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters.
 We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval.
@@
 We describe the scattering transform in Section \ref{sec:scattering}, discuss processing of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
-In Section \ref{sec:experiments}, several experiments are described for the acoustic scene classification task and the acoustic scene similarity retrieval task.
+In Section \ref{sec:experiments}, several experiments are described for the acoustic scene similarity retrieval task.
 Results are subsequently reported in Section \ref{sec:results}.
@@
 
-\section{Supervised classification system}
-\label{sec:supervised}
-
-We are now ready to present the supervised method that we will apply to the acoustic scene classification task using scattering transforms.
-It consists of two parts, a classifier and a temporal integration method.
-While the classifier is an SVM, we shall consider two integration approaches, early and late.
-
-\subsection{Support vector machine classifier}
-\label{sec:svm}
-
-A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{Cortes1995}.
-From another point of view, it attempts to find a direction on which to project in order to maximize separation between two point clouds.
-Since it is not a generative, but discriminative model, it is able to pick out those feature vectors that best characterize a particular class of auditory scenes compared to another.
-This property, coupled with its ease of use and efficient implementations \cite{libsvm} have made SVMs a popular choice for many classification systems.
-
-At the classification stage, the SVM gives a class label as output, which is one of the two classes it models.
-For multi-class settings, a one-\vs-one approach is often used, where an SVM is trained for each pair of classes.
-During classification, the outputs of all pairwise SVMs is computed and the class selected the largest number of times is chosen.
-
-Finally, the SVM can also be extended to handle nonlinear decision surfaces.
-This is obtained using the ``kernel trick'', which exploits the fact that all calculations in the SVM rely only on the inner product between vectors.
-By replacing this standard inner product by some other positive definite symmetric mapping, such as a Gaussian radial basis function (RBF), this is equivalent to performing a linear SVM after mapping the feature vectors into some high-dimensional space.
-A nonlinear mapping then results in a nonlinear decision surface in the original space.
-
-\subsection{Early \vs late temporal integration}
-\label{sec:eili}
-
-Owing to the scarcity of salient events in many natural scenes,
-fine-grained classification is only made
-possible by integrating signal information over a long temporal context.
-Indeed, whereas a few seconds is often sufficient to recognize a speaker,
-a musical instrument, or a genre, up to $30$ seconds may be required
-to disambiguate two classes of auditory scenes with overlapping semantic content, \eg a train from a subway station or a quiet street from a park.
-Depending on whether aggregation is performed in feature space or in decision space (the output of a classifier) the corresponding method is referred to as early or late integration.
-
-A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene by computing their average values over time.
-In the definition of the scattering transform (see Section \ref{sec:scattering}), this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity.
-With a slight abuse of notation, we denote the summarized features by
-\begin{equation}
-\mathbf{S}\boldsymbol{x}(\gamma) =
-\int_{-\infty}^{+\infty} \mathbf{S}\boldsymbol{x}(t,\gamma)\;\mathrm{d}t\mbox{.}
-\end{equation}
-
-Conversely, a late integration scheme relies on probabilistic assignments $\mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right]$ over short-term windows of length $T$ obtained from some supervised model, such as a classifier.
-These are subsequently aggregated to produce a final decision
-\begin{equation}
-\hat{y} = \arg \max_{y} \rho\Big(\big\{ \mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right] \big\}_{t} \Big)\mbox{,}
-\end{equation}
-where $\hat{y}$ is the estimated class label and $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.
-
-A potential drawback of early integration is that it reduces a sequence of feature vectors to a single average.
-If the underlying signal is static in nature, this may be advantageous, as the temporal variation of the feature vectors constitutes an unwanted source of variability.
-For many audio signals, and in particular auditory scenes, this is very far from the truth, and prevents us from identifying the distinct sound events that characterize the signal as these may be drowned out by the background during the averaging step.
-In this case, a late integration approach would perform better, since it operates on the individual feature vectors and only the classification results are aggregated over time.
-Using experiments in acoustic scene classification below, we will demonstrate that the late integration approach does indeed succeed over early integration when scattering coefficients are used.
-
 \section{Acoustic scene similarity retrieval}
@@
 
-\begin{table}
-\begin{center}
-\caption{
-\label{table:dcase2013}
-Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.
-}
-\begin{tabular}{llll}
-             & MFCCs         & scattering & log-scattering  \\
-             \hline
-\emph{early}, linear  & $54\pm19$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
-\emph{early}, RBF     & $47\pm16$  & $61\pm\phantom{0}9$  & $63\pm\phantom{0}6$   \\
-\emph{late}, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
-\emph{late}, RBF     & $68\pm14$ & $73\pm\phantom{0}6$  & $\mathbf{78}\pm\phantom{0}6$   \\
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Experiment \#1: acoustic scene classification}
-
-Our first task is to compare the performance of the standard scattering transform, the logarithmically compressed scattering (log-scattering) transform, and baseline MFCCs.
-Secondly, we evaluate the two temporal integration approaches presented above, \emph{early} and \emph{late}.
-
-The different systems are evaluated on the private part of the DCASE 2013 dataset using five-fold stratified cross validation.
-The folds are identical to those used in the original challenge.
-The slack parameter $C$ of the SVM is here set to $1$.
-Both linear and Gaussian RBF kernels are used.
-For the latter, a local scaling method is applied \cite{selfTuneManor2004}.
-All feature vectors, that is scattering frames and MFCCs, are standardized prior to be given to the classifier.
-The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).
-The optimal scaling parameter $q$ of the RBF kernels is learned on the public part of the DCASE 2013 dataset, using five-fold stratified cross validation.
-For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).
-
-\subsection{Experiment \#2: acoustic scene similarity retrieval}
-
 Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval.
@@
 
-The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation.
+%The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation.
 Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering and MFCCs features.
-
-\subsection{Experiment \#1: acoustic scene classification}
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=\columnwidth]{bw/folds.png}
-%\caption{\label{fig:folds}
-%Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.
-Performance on each fold is depicted using a given symbol.
-%The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
-%The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.}
-%\end{center}
-%\end{figure}
-
-\begin{table}
-\begin{center}
-\caption{Class-wise accuracies on the DCASE 2016 dataset.\label{table:dcase2016}}
-\begin{tabular}{lccc}
-Features & MFCCs & log-scattering & log-scattering \\
-Subset & development & development & evaluation \\
-\midrule
-beach & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ & $80.8$\\
-bus & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ & $92.3$ \\
-cafe/restaurant & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ & $50.0$ \\
-car & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ & $96.2$ \\
-city\_center & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ & $84.6$ \\
-forest\_path & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ & $96.2$ \\
-grocery\_store & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ & $84.6$ \\
-home & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ & $80.8$ \\
-library & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ & $65.4$ \\
-metro\_station & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ & $96.2$ \\
-office & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ & $100$ \\
-park & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ & $65.4$ \\
-residential\_area & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ & $69.2$ \\
-train & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ & $53.8$ \\
-tram & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ & $96.2$ \\
-\bottomrule
-Average & $65.8 \pm 19.3$ & $79.4 \pm 15.6$ & $80.8 \pm 16.4$ \\
-\end{tabular}
-\end{center}
-
-\end{table}
-
-We report experiments on acoustic scene classification, following the original methodology of the DCASE 2013 challenge, and extend our approach to its 2016 edition.
-
-\subsubsection*{Role of logarithmic compression}
-
-On the DCASE 2013 dataset, logarithmic compression provides a small, albeit consistent boost to acoustic scene classification.
-Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
-This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
-
-\subsubsection*{MFCCs \vs scattering}
-
-%As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
-
-As shown on Table \ref{table:dcase2013}, scattering coefficients outperform MFCCs on the DCASE 2013 private dataset when applying early and integration.
-This is in accordance with previous results on similarity retrieval, as well as comparisons between mel-frequency coefficients and scattering coefficients for tasks ranging from musical genre recognition and phone segment classification \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.
-
-In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, which was evaluated on a much larger dataset compared to DCASE 2013.
-For this reason, training support vector machines with the Gaussian RBF kernel proved computationally too demanding and we instead used a linear kernel.
-Results on DCASE 2013 (see Table \ref{table:dcase2013}) suggests that this provides comparable, albeit slightly worse results.
-The results on DCASE 2016 are reported on Table \ref{table:dcase2016}.
-
-We compare our results with baseline classifier using linear SVMs and MFCC features, which achieves an average accuracy of $65.5\%$.
-With the scattering transform, however, the average accuracy increases to $79.4\%$.
-This accuracy increases slightly to $80.8\%$ when calculated on the held-out evaluation dataset.\footnote{\url{http://www.cs.tut.fi/sgn/arg/dcase2016/task-results-acoustic-scene-classification}} The small change in overall accuracy and relatively stable class-wise accuracies suggest that there is little overfitting in the model, an indication of good generalization power.
-
-\subsubsection*{Early \vs late integration}
-On the DCASE 2013 dataset, late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
-This suggests that there is relevant variability in scales beyond the window size $T = 372~\mathrm{ms}$ that is not captured by averaging the scattering transform in time (early integration).
-One potential solution to this is to increase the scale of the second-order modulation wavelets from $J_2 = 12$ octaves to some larger value.
-This does not result in any performance gain, however (results not shown).
-What does help is the late integration, since it allows the SVM to successfully classify the few short extracts that characterize the scene which are then used to classify the whole scene through the majority vote.
-Since the scattering transform, even with larger values of $J_2$, is not able to separate out these short, discriminative sound events, delegating this task to the late-integrated results in a more powerful classification system.
-
-\subsection{Experiment \#2: acoustic scene similarity retrieval}
 

### 2018-01-18 paper/structureScene.tex
change styling

@@
-\documentclass{IEEEtran}
+\documentclass[smallextended]{svjour3}
 %\usepackage{cite}
@@
 %
-\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
-\fi
 %
@@
 \title{Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio}
+\titlerunning{Relevance-based Quantization of Scattering Features}
 
@@
 Experimentation performed using the Acoustic Scene Similarity framework demonstrates the superiority of the proposed approach.
+\keywords{
+unsupervised learning \and data mining \and acoustic signal processing \and wavelet transforms \and audio databases \and content-based retrieval \and nearest neighbor searches \and acoustic sensors \and environmental sensors.}
+% \PACS{PACS code1 \and PACS code2 \and more}
 \end{abstract}
-
-% Note that keywords are not normally used for peerreview paper.
-\begin{IEEEkeywords}
-unsupervised learning, data mining, acoustic signal processing, wavelet transforms, audio databases, content-based retrieval, nearest neighbor searches, acoustic sensors, environmental sensors.
-\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
 
@@
 
-\IEEEPARstart{T}{he} amount of audio data recorded from our sonic environment has grown considerably over the past decades.
+The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
 In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
@@
 A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
-This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{} \vl{some relevant refs on dnns 4 audio}.
+This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{} \ja{some relevant refs on dnns 4 audio}.
 
@@
 As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
-While this can be addressed in the supervised case using late integration of discriminative classifiers (see Section \ref{sec:eili}), this is not directly feasible in the unsupervised case.
+While this can be addressed in the supervised case using late integration of discriminative classifiers \ja{provide citation of your work}, this is not directly feasible in the unsupervised case.
 As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
@@
 
-%\bibliographystyle{unsrt}
+%\bibliographystyle{spmpsci}
 %\bibliography{biblio}
-\printbibliography
+%\printbibliography
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lostanlen}}]{Vincent Lostanlen} was born in 1992.
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lostanlen}}]{Vincent Lostanlen} was born in 1992.
 He received an engineering degree from TELECOM ParisTech in 2013 and his M.S.
@@
 His research focuses on defining convolutional operators in the time-frequency domain with applications in audio classification.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in 1990.
+% \end{IEEEbiography}
+%
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in 1990.
 He received the B.S.
@@
 His research interests include acoustic scene similarity and classification as well as acoustic scene perception.
-\end{IEEEbiography}
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/anden}}]{Joakim And\'en}
-received the M.Sc.
+% \end{IEEEbiography}
+%
+%
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/anden}}]{Joakim And\'en}
+% received the M.Sc.
 degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the Ph.D.
@@
 He is a member of the IEEE.
-
-\end{IEEEbiography}
-
-\begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
+%
+% \end{IEEEbiography}
+%
+% \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
 He obtained his Ph.D.
@@
 His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.
-\end{IEEEbiography}
+% \end{IEEEbiography}
 

### 2018-01-18 paper/structureScene.tex
cleaning

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{bw/gammatones}
+\includegraphics[width=\columnwidth]{gfx/gammatones}
 \caption{
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{bw/compression}
+\includegraphics[width=\columnwidth]{gfx/compression}
 \caption{
@@
 
-The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/tetci17}} % \ja{What is ``the one''?}
+The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}} % \ja{What is ``the one''?}
 
@@
 %Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
-The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} and DCASE 2016 \cite{Mesaros2016_EUSIPCO} datasets.
+The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} dataset.
 The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes, sampled at $44100\,\mathrm{Hz}$.
@@
 As a consequence, DCASE 2013 dataset enjoys an interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
-In addition, it is still a challenging dataset, with a state-of-the-art system based using handcrafted features as input to SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
+%In addition, it is still a challenging dataset, with a state-of-the-art system based using handcrafted features as input to SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{bw/unsupervised_test2}
+\includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
 \caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
@@
 %\bibliography{biblio}
-%\printbibliography
+\printbibliography
 

### 2018-02-02 paper/structureScene.tex
Edit abstract

@@
 \begin{abstract}
-The emerging field of computational ecoacoustics aims at retrieving high-level information from the acoustic scenes recorded by some network of sensors.
-Large amount of unlabeled data are gathered using those networks that needs to be analyzed.
-In order to identify which parts needs human inspection, one needs tools that can mine data and identify structural such as recurring patterns or on contrary sparse occurences.
+The emerging field of computational ecoacoustics aims at retrieving high-level information from acoustic scenes recorded by some network of sensors.
+Large amounts of unlabeled data are gathered using such networks that needs to be analyzed.
+In order to identify which parts require human inspection, one needs tools that can mine data and identify structures such as recurring patterns or isolated events.
 To do so, one needs to be able to measure the level of similarity among the data without strong assumptions.
 
-State of the art approaches for computing such a similarity is the "bag-of-frames" approach, i.e.
-modeling the audio using the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
+State of the art approaches for computing such a similarity is the ``bag-of-frames'' approach, i.e., modeling the audio signal using the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
 Although they successfully characterize static scenes with little variability in auditory content, they are unable to accurately discriminate scenes with few salient events superimposed over a textured background.
-To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure that better handle the required modeling of scarce events.
-Experimentation performed using the Acoustic Scene Similarity framework demonstrates the superiority of the proposed approach.
+To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure that better handle the required modeling of sparse events.
+Experiments performed using the acoustic scene similarity framework demonstrates the superiority of the proposed approach.
 \keywords{

### 2018-02-02 paper/structureScene.tex
Edit intro

@@
 The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken the massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
+In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
 In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
-Because they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.
+Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.
 
-This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
+This work falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
 %Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
 %Consequently, few well-designed datasets are readily available for evaluation purposes.
+An important problem in this field is that manually analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling}.
+Some sort of pre-screening is therefore needed to reduce the need for human expert listening and annotation.
 
-An important problem in this field is that analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling} and needs some sort of pre-screening to reduce the need for human expert listening and annotation.
+The most straightforward approach is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}.
+A given time interval, for example a single day, is then represented by the number of events within the predefined typology that have been detected in that period of time.
+This approach allows the scientist to drastically reduce the amount of information that needs to be processed.
+However, this approach has two drawbacks.
+The first is the reliance to trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be always be trusted.
+More importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where the quantities of interest have yet to be defined.
 
-The most straightforward way is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}.
-By doing so, a period of time, say a day is represented as the number of events within the predefined typology that have been detected in this period of time.
-This approach allows the scientist to reduce drastically the amount of information that needs to be processed.
-Though, this approach have two drawbacks.
-The first is the reliance to trained models whose behavior on unseen data (different sensors, ...) is prone to errors and can hardly be asserted.
-And, more importantly, it is based on \textit{a priori} knowledge, and thus cannot be considered for exploratory analysis where the quantities of interest are yet to be defined.
-
+\ml{What's going on here?}
 and annotation to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors.
@@
 
-To this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset.
+With this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset.
 This requires the audio fragments to be represented in some way that captures their distinctive qualities.
-This has been achieved by bag-of-frames approaches in the litterature\cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics.
-Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others.
-Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes.
-We believe that the failure to model such distinct events is one of the reason why this approach is found to be not sufficient when considering decent size evaluation datasets \cite{lagrange:hal-01082501}.
+This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics.
+Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented with scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others.
+Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely human acoustic perception of acoustic scenes.
+We believe that the failure to model such distinct events is one of the reason why this approach is found to be insufficient when considering evaluation datasets of typical sizes \cite{lagrange:hal-01082501}.
 
@@
 
-Addressing this task requires first to be able to capture meaningful aspects within the signal at small time scales.
+Addressing this task first requires the ability to capture meaningful aspects within the signal at small time-scales.
 In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
 A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
-A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014}.
+A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014,chudacek}.
 This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{} \ja{some relevant refs on dnns 4 audio}.
 
-For the task at at hand, the scattering features have the valuable advantage of not requiring training and thus potentially lead to wider range of application than trained features.
-Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined.
-Thus leading to an unserpervised setting.
+For the task at at hand, scattering features have the advantage of not requiring training and thus potentially lead to wider range of application than trained features.
+Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined, thus leading to an unserpervised setting.
 
@@
 We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval.
-Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.
+Testing this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
 
-Motivations of the proposed approach and a brief review of the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}.
-We describe the scattering transform in Section \ref{sec:scattering}, discuss processing of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
-In Section \ref{sec:experiments}, several experiments are described for the acoustic scene similarity retrieval task.
+Motivations of the proposed approach and a brief review of the state of the art in acoustic scene modeling are given in Section \ref{sec:soa}.
+We describe the scattering transform in Section \ref{sec:scattering}, discuss feature post-processing in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
+Section \ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task.
 Results are subsequently reported in Section \ref{sec:results}.

### 2018-02-02 paper/structureScene.tex
Add some references for conv nets

@@
 A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014,chudacek}.
-This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{} \ja{some relevant refs on dnns 4 audio}.
+This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}.
 

### 2018-02-02 paper/structureScene.tex
Edit background section

@@
 
-Computational bioacoustics usually refers to the investigation of sound production, dispersion and reception in animals (including humans) by computational means.
+Computational bioacoustics usually refers to the numerical investigation of sound production, dispersion and reception in animals (including humans).
 A recent paradigm  processes the audio stream in a holistic way over large time-scales, without assuming that a single species is present throughout the recording.
@@
 %The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-In the closely related field of urban sound environment analysis, a popular approach is the bag-of-frames (BOF) introduced in this context by Aucouturier et al.
+In the closely related field of urban sound environment analysis, a popular approach is the bag of frames (BOF) introduced in this context by Aucouturier et al.
 \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
@@
 This contrasts with the typical morphology of acoustic scenes, which we believe is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events are superimposed upon a stationary acoustic background.
-As those distinct and scarce events are most of the time highly relevant for the task at hand, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on the joint probability distribution of short-term features.
+As those distinct and sparse events are often highly relevant for the task at hand, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on summary statistics of short-term features.
 
@@
 Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments \cite{ricciardi2015sound,lavandier2006contribution}.
-Although the complete set of events occurring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes.
+Although the complete set of events occurring within a given auditory stream may not be discernable even to human expects, research has shown that a small set of events (so-called markers) suffice to reliably predict many high-level attributes.
 
-From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation equally when computing summary statistics.
+From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation the same and do not incorporate temporal structure between features.
 A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes.
@@
 
-The methods that performed the best all followed a similar structure, where low-level feature representations, such as MFCCs, were calculated over short windows and fed separately to a classifier, such as an SVM, for training.
-During the classification stage, the recording to be classified was similarly processed into short windows represented by feature vectors which were classified individually by the classifier and the classifier output was combined over the recording through majority vote.
+The methods that performed the best all follow a similar structure, where low-level feature representations, such as MFCCs, are calculated over short windows and fed separately to a classifier, such as a support vector machine (SVM) \cite{cortes-vapnik,cristianini2000introduction}, for training.
+During the classification stage, a given recording is similarly processed into short windows represented by feature vectors which are classified individually by the classifier.
+The individial class assignments are then combined through a majority vote which assigns a class to the whole recording.
 This last step is known as late integration and allows the classifier to work on individual, shorter-scale sound events instead of trying to classify the recording as a whole.
@@
 This is in agreement with the psychoacoustic results related to the perception of auditory scenes presented above.
+
+\ml{These two last paragraphs are a bit out of place here, since we don't do classification in the paper. Perhaps there is some way to cut them down a bit?}
 

### 2018-02-02 paper/structureScene.tex
Edit scattering section

@@
 
-Scattering transforms are time-shift invariant representations of audio signals computed by applying auditory and modulation filter banks alternated with complex modulus nonlinearities.
-This section first explains the importance of local invariance to time-shifting and stability to time-warping in the representation of acoustic scenes, then describes how the scattering transform is designed to satisfy these properties while having a high discriminative power.
+Scattering transforms are time-shift invariant representations of audio signals computed by applying auditory and modulation wavelet filter banks alternated with complex modulus nonlinearities.
+This section first explains the importance of local invariance to time-shifting and stability to time-warping in the representation of acoustic scenes, then describes how the scattering transform is designed to satisfy these properties while retaining high discriminative power.
 
@@
 The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
-Indeed, recordings may be shifted locally without affecting their original class memberships.
-To discard this superfluous source of variability, signals are mapped into a time-shifting invariant feature space before training the classifier, eliminating the need for this classifier to explicitly learn this invariance and increasing the robustness of its predictions.
+Indeed, recordings may be shifted locally without affecting their similarity to other recordings.
+To discard this superfluous source of variability, signals are first mapped into a time-shifting invariant feature space.
+These features are then used to calculate similarities or to train a classifier.
+Since invariance is ensured by the features, it does not have to be learned during this last model construction step.
 
@@
 Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int_{-\infty}^{+\infty} x(t) \exp(- \mathrm{i} 2\pi \omega t) \; \mathrm{d}t$.
-Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
+Let $\boldsymbol{\psi}(t)$ a complex-valued analytic bandpass filter of
 center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$, where $Q_1$ is the quality factor of the filter.
@@
 \end{equation}
-The variable $\gamma_1$ is a scale (an inverse log-frequency) taking integer values between $0$ and $(J_1 \times Q_1 - 1)$, where $J_1$ is the number of octaves spanned by the filter bank.
+The variable $\gamma_1$ is a scale (an inverse log-frequency) taking integer values between $0$ and $(J_1 Q_1 - 1)$, where $J_1$ is the number of octaves spanned by the filter bank.
 For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
 has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$ and a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$ resulting in the same quality factor $Q_1$ as $\boldsymbol{\psi}$.
-In the sequel, we set $\xi_1$ to $20~\mathrm{kHz}$, $J_1$ to $10$, and the quality factor $Q_1$, which is also the number of wavelets per octave, to $8$.
+In the following, we set $\xi_1$ to $20~\mathrm{kHz}$, $J_1$ to $10$, and the quality factor $Q_1$, which is also the number of wavelets per octave, to $8$.
 This results in the wavelet filters covering the whole range of human hearing, from $20~\mathrm{Hz}$ to $20~\mathrm{kHz}$.
-Setting $Q_1 = 8$ results in filters whose bandwidth approximates an equivalent rectangular bandwidth (ERB) scale.
+Setting $Q_1 = 8$ results in filters whose bandwidth approximates an equivalent rectangular bandwidth (ERB) scale \cite{Fastl2007}.
 
 The wavelet transform of an audio signal
-$\boldsymbol{x}(t)$ is obtained by convolution with all wavelet filters
+$\boldsymbol{x}(t)$ is obtained by convolution with all wavelet filters.
 Applying pointwise complex modulus the transform yields
@@
 Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
-However, contrary to the CQT, the scalogram reaches the Heisenberg
-theoretical limit of optimal time-frequency localization across the whole
+However, contrary to the CQT, the scalogram reaches a better time-frequency localization across the whole
 frequency range, whereas the temporal resolution of the traditional CQT is fixed by the support of the STFT analyzing window. %\cite{Brown1992}
@@
 frequencies than the CQT, at the expense of a greater computational cost
-since the inverse fast Fourier transform (IFFT) routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
+since the inverse fast Fourier transform routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
 However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
@@
 which is known as the set of first-order scattering coefficients.
-They capture the average spectral envelope of $\boldsymbol{x}$ over scales of duration $T$, with spectral resolution varying with constant Q.
+They capture the average spectral envelope of $\boldsymbol{x}(t)$ over scales of duration $T$ and where the spectral resolution varying with constant Q.
 In this way, they are closely related to the mel-frequency spectrogram and related features, such as MFCCs.
@@
 At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
-Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into a local time-shifting invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform followed by a pointwise modulus operator.
+Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into the time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform and another complex modulus.
 
-We define second-order wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ in the same way as the first-order wavelets, but with parameters $\xi_2$, $J_2$, and $Q_2$ and so have center frequencies $2^{-\gamma_2/Q_2}\xi_2$ for $\gamma_2$ taking values between $0$ and $(J_2 \times Q_2 - 1)$.
+We define second-order wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ in the same way as the first-order wavelets, but with parameters $\xi_2$, $J_2$, and $Q_2$.
+Consequently, they have center frequencies $2^{-\gamma_2/Q_2}\xi_2$ for $\gamma_2$ taking values between $0$ and $(J_2 Q_2 - 1)$.
 While this abuses notation slightly, the identity of the wavelets should be clear from context.
@@
 \end{equation}
-In the sequel, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}$ to guarantee local invariance to time-shifting, which yields
+In the following, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
+Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}(t, \gamma_1, \gamma_2)$ to guarantee local invariance to time-shifting, which yields the second-order scattering coefficients
 \begin{equation}
@@
 While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
+Indeed, higher-order scattering coefficients have been shown to contain reduced energy and are therefore of limited use \cite{irene}.
 
@@
 Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
-Oscillations represent the real and imaginary parts. The envelope represents the complex modulus.}
+Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
 \end{center}
@@
 wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
-In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
+In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
 Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
@@
 Likewise, because critical bands are skewed towards higher frequencies, a masker tone has to be louder if it is above the stimulus in frequency rather than below.
+
 It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.

### 2018-02-02 paper/structureScene.tex
Edit feature design section

@@
 
-Before constructing models for classification or similarity estimation, it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
-In this section, we review two transformations which achieve these properties, namely logarithmic compression and standardization.
+Before constructing models for similarity estimation, it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
+In this section, we review two transformations which achieve these properties: logarithmic compression and standardisation.
 
@@
 
-For the task of musical genre recognition, second-order scattering coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ are sometimes normalized by the corresponding first-order scattering coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, since this decorrelates them from one another \cite{Anden2014}.
-We note that taking the logarithm of such renormalized coefficients yields
-\begin{equation}
-\log \dfrac{\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)}{\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)} =
-\log \mathbf{S_2}\boldsymbol{x}(t, \gamma_1, \gamma_2) -
-\log \mathbf{S_1}\boldsymbol{x}(t, \gamma_1),
-\end{equation}
-\ie a linear combination of the logarithms of first- and second-order coefficients.
-As such, a nonlinear renormalization becomes a linear transformation, which can be learned by a linearly discriminative classifier.
-
-\subsection{Standardization}
+\subsection{Standardisation}
 \label{sec:stand}
 
-Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where the $\gamma$ and $n$ respectively denotes feature and sample indices, respectively.
-It has been found that SVMs should be trained on scaled features with null mean and unit variance so as to avoid mismatch in numeric ranges \cite{Hsu2003}.
-To standardize $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
-The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the training set only, and the same affine transformation is then applied to all samples in both the training and the test sets.
+Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where $\gamma$ and $n$ denote feature and sample indices, respectively.
+Many algorithms operate better on features which have zero mean and unit variance to avoid mismatch in numeric ranges \cite{Hsu2003}.
+To standardise $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
+The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the entire dataset.
 

### 2018-02-03 paper/structureScene.tex
Update ASSR section

@@
 As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
-While this can be addressed in the supervised case using late integration of discriminative classifiers \ja{provide citation of your work}, this is not directly feasible in the unsupervised case.
+While this can be addressed in the supervised case using late integration of discriminative classifiers \cite{Anden2014}, this is not directly feasible in the unsupervised case.
 As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
@@
 The partitioning is done by minimizing squared error between the empirical mean, or centroid, of each cluster and the vectors belonging to it.
-Letting $\mu_m^u$ denote the centroid of the cluster $c_m^u$, we attempt to minimize the following objective function:
-\begin{equation}
-J(C_u)=\sum\limits_{m} \sum_{x^u_l\in c^u_m} \Vert x_l^u - \mu_m^u \Vert^2\mbox{.}
-\end{equation}
-This is known as $k$-means clustering.
+This is known as $k$-means clustering \cite{lloyd} and has found widespread use for a variety of applications.
 Each scene $s_u$ is then described by a set of clusters $C_u$.
@@
 
-To compute the similarity between two scenes, we then consider several centroid-based similarity metrics:
+To compute the similarity between two scenes, we consider several centroid-based similarity metrics:
 \begin{itemize}
@@
 
-For \emph{RbQ-w}, each centroid is weighted according to the number of frames belonging to its cluster.
-Each scene $s_u$ is thus described by a signature $p_u$ of $M$ clusters, where $p_u=\lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace$) and $\mu_m^u$ and $w_m^u$ are the centroid and the weight of the $m$th cluster, respectively.
+For \emph{RbQ-w}, each scene is represented by a signature
+$$
+p_u = \lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace,
+$$
+where each of the $M$ centroids $\mu_1^u, \ldots, \mu_M^u$ are paired with corresponding weights $w_1^u, \ldots, w_M^u$.
+The weight $w_m^u$ for the $m$th centroid $\mu_m^u$ is the number of frames belonging to a particular cluster.
 The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance ($\widehat{\EMD}$) introduced by \cite{pele2008linear}.
-The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts times the ``ground distance'' moved in terms of the histogram bins.
-In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the distance over these cluster centroids.
+The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied with the ``ground distance'' between the histogram bins.
+In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the centroid distance.
+\ml{From what we have below, the distance is not actually the distance between centroids, though.
+It's given by the kernel $K_{mn}^{uv}$.}
 
+\ml{Do we need the details that follow? Is it not enough to refer to \cite{pele2009fast}?}
 To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
@@
 \end{equation*}
+\ml{Note that this is not actually a distance.}
 
@@
 The resulting kernel $K^s$ is known as an $\EMD$ kernel, and it should be noted that there is no guarantee that this kernel is positive definite.
+\ml{Is this kernel used anywhere?
+For similarity, do we not just use the EMD distance directly?}
 

### 2018-02-03 paper/structureScene.tex
Edit experiments section

@@
 
+To evaluate the representations introduced in the previous section, we apply it to the acoustic scene similarity retrieval task on the DCASE 2013 dataset.
+Results demonstrate the improved performance of the relevance-based quantization of scattering coefficients compared to baseline methods using summary statistics of MFCCs.
 The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}} % \ja{What is ``the one''?}
@@
 
-\subsection{Datasets}
+\subsection{Dataset}
 
@@
 The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} dataset.
-The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes, sampled at $44100\,\mathrm{Hz}$.
-The $100$ recordings are evenly divided into $10$ acoustic scene classes.
-To build the DCASE 2013 dataset, three different recordists visited a wide variety of locations in Greater London over a period of several months and in each scene recorded a few minutes of audio.
+This dataset consists of two parts, namely a public and a private subset, each made up of $100$ acoustic scene recordings, sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
+The dataset is evenly divided into $10$ acoustic scence classes.
+The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
 No
@@
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-As a consequence, DCASE 2013 dataset enjoys an interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+As a result, the DCASE 2013 dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
 %In addition, it is still a challenging dataset, with a state-of-the-art system based using handcrafted features as input to SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
@@
 On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
-
-The public part of the dataset can used for optimizing the acoustic scene classification system and the private part used for computing the resulting accuracy using a five-fold cross validation scheme.
-For this paper, we shall perform our experiments on the private part of the dataset.
-The folds used are the same as the ones used during the challenge.
-
-The DCASE 2016 dataset has a similar organization, with $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
-
 %, but here we conduct our experiments on the public part of the dataset. \ja{Any more information on this dataset?}
@@
 Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}).
+\ml{Results in the next section only seem to contain the logarithmic compression.
+Do we have the other results as well?}
 
-MFCCs are computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$, with full frequency range.
+MFCCs are computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$ with full frequency range.
 The standard configuration of $39$ coefficients coupled with an average-energy measure performs best in preliminary tests, so we use this in the following.
@@
 Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval.
+\ml{What does this mean?}
+
+\subsection{Algorithm}
 
@@
 The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
-The $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
+These closest neighbors are determined by a particular distance, which is one of \emph{RbQ-c}, \emph{RbQ-a}, or \emph{RbQ-w}.
+
+The metric $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
 Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item.
@@
 %\ja{And BOF?}
+\ml{What is the early integration approach here? We should define it.}
 
 The scaling parameter $q$ of the RBF kernels (see Eq.
-\ref{eq:kc}) is set to 10 \% of the number of data points to cluster, being the number of scenes for \emph{early} and the number of cluster for RbQ approaches.
-The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and 200 replications.
+\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster, being the number of scenes for \emph{early} and the number of clusters for RbQ approaches.
+\ml{What does this mean?}
+The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and $200$ replicates.
 For each method, the numbers of clusters is set to the value among $8$, $16$ and $32$ that leads to the best $p@9$ on the training set.
+\ml{What's the training set here?}
 

### 2018-02-03 paper/structureScene.tex
Edit results section

@@
 The $p@k$ for different settings are shown in Figure \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics.
-The metrics is shown for values of $k$ up to 9 which corresponds to the number of elements minus one within each class of scene of the dataset.
 This allows us to also evaluate the recall capabilities of the system under evaluation.
@@
 
-For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (RbQ) to improve the similarity measures between the scenes.
+For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (\emph{RbQ}) to improve the similarity measures between the scenes.
 However, it is worth noticing that \emph{RbQ-a} performs worse or comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
 This result is in line with the findings of \cite{lagrange:hal-01082501}.
-To take advantage of an such a representation, we need to select certain representative centroids when comparing quantized objects.
-Furthermore, it appears that \emph{RbQ-c} is better able to characterize the classes than \emph{RbQ-w}.
+To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects.
+Furthermore, it appears that \emph{RbQ-c} is better able to characterize the classes compared to \emph{RbQ-w}.
 This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
 Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
-On the contrary, two similar environments may shared a lot of similar sound sources with only a few sources discriminating between them.
+On the contrary, two similar environments may share a lot of similar sound sources with only a few sources discriminating between them.
+\ml{This is a bit of a stretch, since it may just be that we used the wrong distance to take the weighting into account.
+Perhaps the non-normalize d EMD is not appropriate here?}
 
-Considering the $p@5$ as in \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} as our metric, the use of the log-scattering transform versus MFCCs increases the performance  from 0.31 to 0.49 and the use of the relevance based quantization approach using the closest similarity (RbQ-c) further improves the performance to 0.54 for a global increase of 0.23.
+Considering $p@5$ as our metric (as in \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), the use of the log-scattering transform instead of MFCCs increases the performance from $0.31$ to $0.49$ and the use of the relevance-based quantization approach using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
 

### 2018-02-08 paper/structureScene.tex
Edit conclusion
answer to JAs comments

@@
 
-\ml{What's going on here?}
-and annotation to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors.
+In order to identify which parts needs human inspection, one needs tools that can mine data and annotate to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors.
 The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.
@@
 From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation the same and do not incorporate temporal structure between features.
-A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes.
+A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes, provided that the representation of the representation of those distinct events is rich enough.
 
-Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of acoustic scene similarity retrieval, much of the work dedicated to acoustic scene modeling is applied to the classification task.
-An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
+%Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of acoustic scene similarity retrieval, much of the work dedicated to acoustic scene modeling is applied to the classification task.
+%An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
 
-The methods that performed the best all follow a similar structure, where low-level feature representations, such as MFCCs, are calculated over short windows and fed separately to a classifier, such as a support vector machine (SVM) \cite{cortes-vapnik,cristianini2000introduction}, for training.
+%The methods that performed the best all follow a similar structure, where low-level feature representations, such as MFCCs, are calculated over short windows and fed separately to a classifier, such as a support vector machine (SVM) \cite{cortes-vapnik,cristianini2000introduction}, for training.
 During the classification stage, a given recording is similarly processed into short windows represented by feature vectors which are classified individually by the classifier.
@@
 
-\ml{These two last paragraphs are a bit out of place here, since we don't do classification in the paper. Perhaps there is some way to cut them down a bit?}
+%\ml{These two last paragraphs are a bit out of place here, since we don't do classification in the paper. Perhaps there is some way to cut them down a bit?}
 
@@
 The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied with the ``ground distance'' between the histogram bins.
-In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the centroid distance.
+%In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the centroid distance.
 \ml{From what we have below, the distance is not actually the distance between centroids, though.
@@
 
-\ml{Do we need the details that follow? Is it not enough to refer to \cite{pele2009fast}?}
-To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
+%\ml{Do we need the details that follow? Is it not enough to refer to \cite{pele2009fast}?}
+%To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
 Given two signatures $p_u$ and $p_v$, the $\widehat{\EMD}$ is computed by solving the following linear program:
-\begin{equation}
-\begin{split}
-\widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
-&+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
-\end{split}
-\end{equation}
+% \begin{equation}
+% \begin{split}
+% \widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
+% &+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
+% \end{split}
+% \end{equation}
+%
+% \begin{equation*}
+% \mathrm{s.t.} \quad f_{nm}\geq0 \quad \sum\limits_{m} f_{nm} \leq w_n^u \quad \sum\limits_{n} f_{nm} \leq w_m^v
+% \end{equation*}
+%
+% \begin{equation*}
+% \sum\limits_{n,m}f_{nm} = \min\left( \sum\limits_{n} w_n^u ,\sum\limits_{m} w_m^v \right)
+% \end{equation*}
+% %\ja{Can we provide some interpretation for the different terms and constraints here?}
+% where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin.
+We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$ which is computed from $K$:
+% \begin{equation*}
+% D_{mn}^{uv}=1-K_{mn}^{uv}.
+% \end{equation*}
+% \ml{Note that this is not actually a distance.}
 
-\begin{equation*}
-\mathrm{s.t.} \quad f_{nm}\geq0 \quad \sum\limits_{m} f_{nm} \leq w_n^u \quad \sum\limits_{n} f_{nm} \leq w_m^v
-\end{equation*}
-
-\begin{equation*}
-\sum\limits_{n,m}f_{nm} = \min\left( \sum\limits_{n} w_n^u ,\sum\limits_{m} w_m^v \right)
-\end{equation*}
-%\ja{Can we provide some interpretation for the different terms and constraints here?}
-where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin.
-We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$ which is computed from $K$:
-\begin{equation*}
-D_{mn}^{uv}=1-K_{mn}^{uv}.
-\end{equation*}
-\ml{Note that this is not actually a distance.}
-
-To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ is computed: %\cite{chapelle1999support,jing2003support}
-\begin{equation}
-\label{eq:ks}
-K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(p_u,p_v)}{A} \right) \\
-\end{equation}
-with $A$ a scaling parameter, set to the mean value of the $\widehat{\EMD}$ between all the scenes.
+% To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ is computed: %\cite{chapelle1999support,jing2003support}
+% \begin{equation}
+% \label{eq:ks}
+% K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(p_u,p_v)}{A} \right) \\
+% \end{equation}
+% with $A$ a scaling parameter, set to the mean value of the $\widehat{\EMD}$ between all the scenes.
 The resulting kernel $K^s$ is known as an $\EMD$ kernel, and it should be noted that there is no guarantee that this kernel is positive definite.
@@
 Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}).
-\ml{Results in the next section only seem to contain the logarithmic compression.
-Do we have the other results as well?}
 
@@
 
-Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval.
+%Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval.
 \ml{What does this mean?}
@@
 Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item.
-The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}.
-%\ja{And BOF?}
-\ml{What is the early integration approach here? We should define it.}
+The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene.
+The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \texit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization algorithm.
+Gaussian models are compared using Monte-Carlo sampling approaximation.
 
 The scaling parameter $q$ of the RBF kernels (see Eq.
-\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster, being the number of scenes for \emph{early} and the number of clusters for RbQ approaches.
+\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.%, being the number of scenes for \emph{early} and the number of clusters for RbQ approaches.
 \ml{What does this mean?}
-The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and $200$ replicates.
-For each method, the numbers of clusters is set to the value among $8$, $16$ and $32$ that leads to the best $p@9$ on the training set.
-\ml{What's the training set here?}
+The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and $200$ replicates. For each method, the numbers of clusters is set to $8$.
 
@@
 This allows us to also evaluate the recall capabilities of the system under evaluation.
+
+\subsubsection*{Logarithmic compression}
+
+As can be seen on Fiugre \ref{fig:ASS_0}, the logarithmic compression is beneficial for precision performance irrespective of the similarity considered.
+The gain is more pronounced when considering alternative similarities.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=\columnwidth]{gfx/unsupervised_test1}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
+\ml{ob a missing ??}}
+\label{fig:ASS_0}
+\end{center}
+\end{figure}
 
@@
 
+\ja{needs to be discussed}
+
 Considering $p@5$ as our metric (as in \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), the use of the log-scattering transform instead of MFCCs increases the performance from $0.31$ to $0.49$ and the use of the relevance-based quantization approach using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
@@
 \includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.
+\ml{add gmms}}
 \label{fig:ASS_1}
@@
 Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature.
-We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification.
 In order to extend the selectivity potential of late integration using SVMs to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval.

### 2018-02-09 paper/structureScene.tex
new results

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/gammatones}
+\includegraphics[width=\columnwidth]{figures/gammatones}
 \caption{
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/compression}
+\includegraphics[width=\columnwidth]{figures/compression}
 \caption{
@@
 The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene.
-The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \texit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization algorithm.
+The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \textit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization algorithm.
 Gaussian models are compared using Monte-Carlo sampling approaximation.
@@
 
+\subsubsection*{Baselines}
+
+As can be seen on Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when considering MFCCs as features.
+Its performance decreases compared to the early approach when considering the scattering features.
+Extensive analysis of those results is out of the scope of this paper, but we can assume that this decrease is to to the dimensionality reduction operated on the scattering features before GMM training in order to ensure the convergence of the EM algorithm.
+
+The early approach being simpler in terms of implementation and runtime complexity, this method is retained as baseline for the remaining.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/baselines}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
+\label{fig:ASS_-1}
+\end{center}
+\end{figure}
+
 \subsubsection*{Logarithmic compression}
 
-As can be seen on Fiugre \ref{fig:ASS_0}, the logarithmic compression is beneficial for precision performance irrespective of the similarity considered.
+As can be seen on Figure \ref{fig:ASS_0}, the logarithmic compression is beneficial for precision performance irrespective of the similarity considered.
 The gain is more pronounced when considering alternative similarities.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/unsupervised_test1}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
-\ml{ob a missing ??}}
+\includegraphics[width=\columnwidth]{figures/log}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
 \label{fig:ASS_0}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
+\includegraphics[width=\columnwidth]{figures/overall}
 \caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.
@@
 
-% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lostanlen}}]{Vincent Lostanlen} was born in 1992.
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/lostanlen}}]{Vincent Lostanlen} was born in 1992.
 He received an engineering degree from TELECOM ParisTech in 2013 and his M.S.
@@
 %
-% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/lafay}}]{Gr\'egoire Lafay} was born in 1990.
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/lafay}}]{Gr\'egoire Lafay} was born in 1990.
 He received the B.S.
@@
 %
-% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{gfx/anden}}]{Joakim And\'en}
+% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/anden}}]{Joakim And\'en}
 % received the M.Sc.
@@
 %
-% \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{gfx/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
+% \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{figures/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
 He obtained his Ph.D.

### 2018-02-09 experiments/deprecated/generateFig/unsupervisedFig.m

@@ -35,7 +35,7 @@ for ii=1:length(dataLab)
         end
     end
     eval([dataLab{ii} '.meanBest=' dataLab{ii} '.mean(indMax,:);']); 
-     eval([dataLab{ii} '.meanBest=' dataLab{ii} '.meanBest(1, :);']); % ML
+%      eval([dataLab{ii} '.meanBest=' dataLab{ii} '.meanBest(1, :);']); % ML
 end
 
 disp('')

### 2018-02-09 experiments/deprecated/supervised/dcase_scattering.m

@@ -1,74 +1,0 @@
-function dcase_scattering(dataset_path, Q1)
-% Load names
-listing = list_dir(dataset_path);
-names = {listing.name};
-
-% Remove hidden files
-names = names(~cellfun(@(x) x(1)=='.', names));
-
-% Keep only WAV extension
-names = names(cellfun(@(x) strcmp(x((end-3):end), '.wav'), names));
-
-% Prepare scattering "architectures", i.e. filter banks and nonlinearities
-opts{1}.time.nFilters_per_octave = Q1;
-opts{1}.time.size = 2^19;
-opts{1}.time.T = 2^14;
-opts{1}.time.is_chunked = false;
-opts{1}.time.gamma_bounds = [0 Q1*11]; % Restrict to top 11 acoustic octaves
-opts{1}.time.wavelet_handle = @gammatone_1d;
-opts{1}.time.S_log2_oversampling = 0;
-opts{2}.banks.time.nFilters_per_octave = 1;
-opts{2}.banks.time.wavelet_handle = @gammatone_1d;
-opts{2}.banks.time.sibling_mask_factor = 2^6;
-opts{2}.banks.time.T = 2^17;
-opts{2}.invariants.time.size = 2^19;
-opts{2}.invariants.time.T = 2^14;
-opts{2}.invariants.time.subscripts = 1;
-opts{3}.invariants.time.size = 2^19;
-opts{3}.invariants.time.T = 2^14;
-opts{3}.invariants.time.subscripts = 1;
-archs = sc_setup(opts);
-
-% Prepare paths for memoized features
-out_file_name = ['Q=', num2str(Q1, '%0.2d')];
-if strcmp(dataset_path((end-6):end), 'testset')
-    out_file_name = [out_file_name, '_test'];
-else
-    out_file_name = [out_file_name, '_train'];
-end
-[~,~] = mkdir('memoized_features'); % [~,~] is to ignore the "already exists" warning
-
-chunk_length = 2^19;
-hop_length = 2^18;
-nChunks = 4;
-chunk_range = 1:chunk_length;
-S_chunks = cell(1, nChunks);
-
-%%
-nFiles = length(names);
-X_features = cell(1, nFiles);
-for file_index = 1:nFiles
-    name = names{file_index};
-    path = fullfile(dataset_path, name);
-    stereo_waveform = audioread(path);
-    waveform = mean(stereo_waveform, 2);
-    chunks = cat(2, ...
-        waveform(0*hop_length + chunk_range, :, :), ...
[…]

### 2018-02-09 experiments/deprecated/supervised/dcase_script.py

@@ -1,32 +1,0 @@
-import collections
-import datetime
-import joblib
-import numpy as np
-import pickle
-import h5py
-import numpy as np
-import os.path
-import scipy.stats
-import sklearn.decomposition
-import sklearn.ensemble
-import sklearn.linear_model
-import sklearn.metrics
-import sklearn.preprocessing
-
-from dcase_svm import cached_dcase_svm, dcase_svm
-
-np.set_printoptions(precision=2)
-
-
-
-C = 1
-exp1.append(dcase_svm(0, 12, False, "time", "logmedian", "early", C))
-exp1.append(dcase_svm(0, 12, True, "time", "logmedian", "early", C))
-exp1.append(dcase_svm(0, 12, False, "timefrequency", "logmedian", "early", C))
-exp1.append(dcase_svm(0, 12, True, "timefrequency", "logmedian", "early", C))
-
-
-exp1.append(dcase_svm(0, 11, True, "timefrequency", "logmedian", "early", C))
-exp1.append(dcase_svm(0, 10, True, "timefrequency", "logmedian", "early", C))
-exp1.append(dcase_svm(0,  9, True, "timefrequency", "logmedian", "early", C))
-exp1.append(dcase_svm(0,  8, True, "timefrequency", "logmedian", "early", C))

### 2018-02-09 experiments/deprecated/supervised/dcase_svm.py

@@ -1,147 +1,0 @@
-import collections
-import datetime
-import joblib
-import numpy as np
-import pickle
-import h5py
-import os.path
-import scipy.stats
-import sklearn.decomposition
-import sklearn.ensemble
-import sklearn.linear_model
-import sklearn.metrics
-import sklearn.preprocessing
-
-def dcase_svm(octmin, octmax, augmentation, scattering, compression, integration, C):
-    if scattering == "time":
-        hdf5_file = h5py.File("memoized_features/dcase2013_timeQ8_test.mat")
-        X = hdf5_file["dcase2013_timeQ8_test"]["X_test"]
-        freqs = np.ravel(hdf5_file["dcase2013_timeQ8_test"]["freqs"])
-        Y = np.ravel(hdf5_file["dcase2013_timeQ8_test"][u'Y_test'])
-    if scattering == "timefrequency":
-        hdf5_file = h5py.File(
-            "memoized_features/dcase2013_timefrequencyQ8_test.mat")
-        X = hdf5_file["dcase2013_timefrequencyQ8_test"]["X_test"]
-        freqs = np.ravel(hdf5_file["dcase2013_timefrequencyQ8_test"]["X_freqs"])
-        Y = np.ravel(hdf5_file["dcase2013_timefrequencyQ8_test"][u'Y_test'])
-
-    method_str = scattering + "_oct" + str(octmin) + "to" + str(octmax)
-    method_str = method_str + "_" + compression
-    if augmentation:
-        method_str = method_str + "_augmentation"
-    method_str = method_str + "_" + integration
-    method_str = method_str + "_C" + str(C)
-
-    fmin = 10.0 * (2**octmin) # in Hz
-    fmax = 10.0 * (2**octmax) # in Hz
-    freq_indices = (freqs > fmin) & (freqs < fmax)
-    X = X[:, :, :, freq_indices]
-
-    # Split folds
-    X = np.reshape(X,
-        (X.shape[0] / 5,   #  20 scenes per fold
-         5,                #   5 folds
-         X.shape[1],       #   5 azimuths
-         X.shape[2],       # 128 time
-         X.shape[3]))      # ~1k features
-    folds = np.mod(np.arange(100), 5)
-
-    # Early integration
-    if integration == "early":
-        X = np.sum(X, 3)[:, :, :, np.newaxis, :]
-
-    Y_predicted = []
-    accuracies = []
-    for fold_id in range(5):
[…]

### 2018-02-09 experiments/deprecated/supervised/get_time_frequencies.m

@@ -1,79 +1,0 @@
-function X_freqs = get_time_frequencies(nfo)
-
-% Prepare scattering "architectures", i.e. filter banks and nonlinearities
-opts{1}.time.nFilters_per_octave = nfo;
-opts{1}.time.size = 2^19;
-opts{1}.time.T = 2^14;
-opts{1}.time.is_chunked = false;
-opts{1}.time.gamma_bounds = [0 nfo*11]; % Restrict to top 11 acoustic octaves
-opts{1}.time.wavelet_handle = @gammatone_1d;
-opts{1}.time.S_log2_oversampling = 0;
-opts{2}.banks.time.nFilters_per_octave = 1;
-opts{2}.banks.time.wavelet_handle = @gammatone_1d;
-opts{2}.banks.time.sibling_mask_factor = 2^6;
-opts{2}.banks.time.T = 2^17;
-opts{2}.invariants.time.size = 2^19;
-opts{2}.invariants.time.T = 2^14;
-opts{2}.invariants.time.subscripts = 1;
-opts{3}.invariants.time.size = 2^19;
-opts{3}.invariants.time.T = 2^14;
-opts{3}.invariants.time.subscripts = 1;
-archs = sc_setup(opts);
-
-x = randn(2^19, 1);
-S = sc_propagate(x, archs);
-
-%%
-S1_refs = generate_refs(S{1+1}.data, 1, S{1+1}.ranges{1+0});
-for S1_ref_index = 1:size(S1_refs, 2)
-    S{1+1}.data = subsasgn(S{1+1}.data, S1_refs(:, S1_ref_index), S1_ref_index);
-end
-S2_refs = generate_refs(S{1+2}.data, 1, S{1+2}.ranges{1+0});
-for S2ref_index = 1:size(S2_refs, 2)
-    S{1+2}.data = subsasgn( ...
-        S{1+2}.data, S2_refs(:, S2ref_index), ...
-        S2ref_index);
-end
-formatted_S = sc_format(S);
-refs = formatted_S(:, 1);
-
-%%
-networks = [ ...
-    repmat(1, size(S1_refs, 2), 1) ; ...
-    repmat(2, size(S2_refs, 2), 1)];
-
-%%
-gamma1_start = S{1+1}.ranges{1}(1,2);
-nS1_refs = length(S1_refs);
-S1_paths(1:nS1_refs) = struct('gamma', [], 'gamma2', []);
-for ref_index = 1:nS1_refs
[…]

### 2018-02-09 experiments/deprecated/supervised/get_timefrequency_frequencies.m

@@ -1,114 +1,0 @@
-function X_freqs = get_timefrequency_frequencies(nfo)
-
-% Prepare scattering "architectures", i.e. filter banks and nonlinearities
-opts{1}.time.nFilters_per_octave = nfo;
-opts{1}.time.size = 2^15;
-opts{1}.time.T = 2^14;
-opts{1}.time.is_chunked = false;
-opts{1}.time.gamma_bounds = [0 nfo*11]; % Restrict to top 11 acoustic octaves
-opts{1}.time.wavelet_handle = @gammatone_1d;
-opts{1}.time.S_log2_oversampling = 0;
-opts{2}.banks.time.nFilters_per_octave = 1;
-opts{2}.banks.time.wavelet_handle = @gammatone_1d;
-opts{2}.banks.time.sibling_mask_factor = 2^3;
-opts{2}.banks.time.T = 2^17;
-opts{2}.banks.gamma.nFilters_per_octave = 1;
-opts{2}.banks.gamma.T = 2^nextpow2(nfo * 4);
-opts{2}.invariants.time.size = 2^15;
-opts{2}.invariants.time.T = 2^14;
-opts{2}.invariants.time.subscripts = 1;
-opts{3}.invariants.time.size = 2^15;
-opts{3}.invariants.time.T = 2^14;
-opts{3}.invariants.time.subscripts = 1;
-archs = sc_setup(opts);
-
-x = randn(2^15, 1);
-S = sc_propagate(x, archs);
-
-%%
-S1_refs = generate_refs(S{1+1}.data, 1, S{1+1}.ranges{1+0});
-for S1_ref_index = 1:size(S1_refs, 2)
-    S{1+1}.data = subsasgn(S{1+1}.data, S1_refs(:, S1_ref_index), S1_ref_index);
-end
-S2psi_refs = generate_refs(S{1+2}{1,1}.data, 1, S{1+2}{1,1}.ranges{1+0});
-for S2psiref_index = 1:size(S2psi_refs, 2)
-    S{1+2}{1,1}.data = subsasgn( ...
-        S{1+2}{1,1}.data, S2psi_refs(:, S2psiref_index), ...
-        S2psiref_index);
-end
-S2phi_refs = generate_refs(S{1+2}{1,2}.data, 1, S{1+2}{1,2}.ranges{1+0});
-for S2phi_ref_index = 1:size(S2phi_refs, 2)
-    S{1+2}{1,2}.data = subsasgn( ...
[…]

### 2018-02-09 experiments/deprecated/supervised/multichannel_scattering.m

@@ -1,36 +1,0 @@
-function features = multichannel_scattering(waveform, archs)
-%% Truncation
-original_length = 30 * 44100;
-truncated_length = 5 * 2^18;
-start = 1 + (original_length - truncated_length) / 2;
-stop = original_length - start + 1;
-waveform = waveform(start:stop, :);
-nAzimuths = size(waveform, 2);
-if nAzimuths > 1
-    waveform = reshape(waveform, size(waveform, 1), 1, size(waveform, 2));
-end
-
-%% Chunking
-chunk_length = 2^19;
-hop_length = 2^18;
-chunk_range = 1:chunk_length;
-chunks = cat(2, ...
-    waveform(0*hop_length + chunk_range, :, :), ...
-    waveform(1*hop_length + chunk_range, :, :), ...
-    waveform(2*hop_length + chunk_range, :, :), ...
-    waveform(3*hop_length + chunk_range, :, :));
-nChunks = 4;
-chunks = reshape(chunks, size(chunks, 1), nChunks * nAzimuths);
-
-%%
-S_chunks = cell(1, nChunks);
-parfor chunk_index = 1:nChunks
-    S_chunks{chunk_index} = sc_propagate(chunks(:, chunk_index), archs);
-end
-
-features = cellfun(@sc_format, S_chunks, 'UniformOutput', false);
-features = cellfun(@(x) x(:, (1+end/4):(3*end/4)), features, ...
-    'UniformOutput', false);
-features = [features{:}];
-end
-

### 2018-02-09 experiments/deprecated/supervised/parfor_save.m

@@ -1,4 +1,0 @@
-function parfor_save(file, data)
-save(file, 'data');
-end
-

### 2018-02-09 experiments/deprecated/supervised/parse_filenamekey.m

@@ -1,9 +1,0 @@
-function sorted_folds = parse_filenamekey(filenamekey_path)
-%%
-table = readtable(filenamekey_path);
-fold_strs = table2cell(table(:,1));
-folds = cellfun(@(x) str2num(x(5)), fold_strs);
-name_strs = table2cell(table(:,4));
-[sorted_names, sorting_indices] = sort(name_strs);
-sorted_folds = folds(sorting_indices)
-end

### 2018-02-09 experiments/deprecated/supervised/planck_taper.m

@@ -1,17 +1,0 @@
-function w = planck_taper(N, epsilon)
-if nargin < 2
-    epsilon = 0.1;
-end
-%%
-x = linspace(0, 1, N);
-Z_plus = epsilon ./ x + epsilon ./ (x - epsilon);
-Z_minus = - epsilon ./ (x - (1-epsilon)) - epsilon ./ (x - 1);
-w = ones(N, 1);
-t2 = round(epsilon * N);
-t3 = round((1-epsilon) * N);
-w(1) = 0;
-w(2:t2) = 1 ./ (exp(Z_plus(2:t2)) + 1);
-w((t3+1):end-1) = 1 ./ (exp(Z_minus((t3+1):end-1)) + 1);
-w(end) = 0;
-plot(w);
-end

### 2018-02-09 experiments/deprecated/supervised/supervised_script.m

@@ -1,9 +1,0 @@
-% Before running this script, the library scattering.m must have been added
-% to the MATLAB path
-%dataset_path = '~/datasets/dcase2013/scenes_stereo';
-dataset_path = ...
-    '/archive/rmb456/private_datasets/DCASE2013_SC/scenes_stereo';
-Q1 = 8;
-
-dcase_scattering(dataset_path, Q1);
-dcase_scattering([dataset_path, '_testset'], Q1);

### 2018-02-09 experiments/deprecated/supervised/wrangling.m

@@ -1,39 +1,0 @@
-X = load('memoized_features/time_Q=08_train.mat');
-
-%%
-X_features = X.X_features;
-X_features = cat(1, X_features{:}).';
-[nAzimuths, nExamples] = size(X_features);
-X_features = cat(3, X_features{:});
-X_features = reshape(X_features, [ ...
-    size(X_features, 1), ...
-    size(X_features, 2), ...
-    nAzimuths, ...
-    nExamples]);
-
-%%
-X_scattergrams = X.X_scattergrams;
-X_scattergrams = cat(1, X_scattergrams{:}).';
-[nAzimuths, nExamples] = size(X_scattergrams);
-X_scattergrams = cat(4, X_scattergrams{:});
-X_scattergrams = reshape(X_scattergrams, [ ...
-    size(X_scattergrams, 1), ...
-    size(X_scattergrams, 2), ...
-    size(X_scattergrams, 3), ...
-    nAzimuths, ...
-    nExamples]);
-
-%%
-class_names = ...
-    {'bus', 'busystreet', 'office', 'openairmarket', 'park', ...
-    'quietstreet', 'restaurant', 'supermarket', 'tube', 'tubestation'};
-
-%%
-filename = 'time_Q=08.h5';
-data = struct( ...
-    'X_features_train', X_features_train, ...
-    'X_features_test', X_features_test, ...
-    'X_scattergrams_train', X_scattergrams_train, ...
-    'X_scattergrams_test', X_scattergrams_test, ...
-    'Y_train', Y_train, ...
-    'Y_test', Y_test);

### 2018-02-09 experiments/unsupervised/boflib/READAIF.M
publishing code

@@ -1,0 +1,201 @@
+function [y,fs,wmode,fidx]=readaif(filename,mode,nmax,nskip)
+%READAIF  Read a .AIF format sound file [Y,FS,WMODE,FIDX]=(FILENAME,MODE,NMAX,NSKIP)
+%
+% Input Parameters:
+%
+%	FILENAME gives the name of the file (with optional .AIF extension) or alternatively
+%                 can be the FIDX output from a previous call to READAIF
+%	MODE		specifies the following (*=default):
+%
+%    Scaling: 's'    Auto scale to make data peak = +-1
+%             'r'    Raw unscaled data (integer values)
+%             'q'    Scaled to make 0dBm0 be unity mean square
+%             'p' *	Scaled to make +-1 equal full scale
+%   File I/O: 'f'    Do not close file on exit
+%
+%	NMAX     maximum number of samples to read (or -1 for unlimited [default])
+%	NSKIP    number of samples to skip from start of file
+%               (or -1 to continue from previous read when FIDX is given instead of FILENAME [default])
+%
+% Output Parameters:
+%
+%	Y        data matrix of dimension (samples,channels)
+%	FS       sample frequency in Hz
+%	WMODE    mode string needed to recreate the data file
+%	FIDX     Information row vector containing the element listed below.
+%
+%           (1)  file id
+%				(2)  current position in file
+%           (3)  dataoff	byte offset in file to start of data
+%           (4)  nsamp	number of samples
+%           (5)  nchan	number of channels
+%           (6)  nbyte	bytes per data value
+%           (7)  bits	number of bits of precision
+%           (8)  code	Data format: 1=PCM, 2=ADPCM, 6=A-law, 7=Mu-law (always 1 for AIFF)
+%           (9)  fs	sample frequency
+%           (10) offset
+%           (11) block size
+%
+%   If no output parameters are specified, header information will be printed.
+%
+%   For stereo data, y(:,1) is the left channel and y(:,2) the right
+
+%  future enhancements: (1) handle LIST and CAT files as well
+%                       (2) deal with offset and blocksize properly
[…]

### 2018-02-09 experiments/unsupervised/boflib/computeMfccs.m

@@ -1,0 +1,31 @@
+function [data l f]= computeMfccs(audio,fs, nbcoef)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ex. to compute 20 mfcc from "data/audio1.wav", call 
+%% data = computeMfccs("data/audio1.wav"1,20);
+%% data is a N * 20 matrix, where N is the number of 2048-pt, 50-percent-overlap frames in the signal
+%% Copyright (C) JJ Aucouturier, 2008
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+if ~exist('nbcoef', 'var'), nbcoef=20; end
+
+% convert to mono
+if(size(audio,2)==2)
+    audio=.5*(audio(:,1)+audio(:,2));
+end
+% compute mfcc
+%fprintf(1,'- compute mfcc\n');
+win=2048; % frame size
+hop=win*0.5;
+w='t'; % triangular filters
+
+% cuts zeros at the beginning of a file, if needed
+i=1;
+while (audio(i)==0)
+    i=i+1;
+end
+audio=audio(i:end);
+
+Z=enframe(audio,hamming(win),hop);
+clear audio;
+[data l f]=mfc(Z,fs,w,nbcoef,floor(3*log(fs)),win,hop,0,0.5);
+clear Z;

### 2018-02-09 experiments/unsupervised/boflib/createGmm.m

@@ -1,0 +1,26 @@
+function mix = createGmm(data, nbgmm)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ex. to create a gmm with 50 states from a mfcc matrix ("mymfcc") and save it to a file ("gmm/mymodel.gmm"), call 
+%% mix = createGmm(mymfcc, 'gmm/mymodel.gmm', 50)
+%% mix is the netlab struct containing gmm parameters
+%% mymodel.gmm is a mat file containing a netlab-packed version of mix (see saveGMM)
+%% Copyright (C) JJ Aucouturier, 2008
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+
+%fprintf(1,'- compute gmm\n');
+nbcoef=size(data,2);
+mix = gmm(nbcoef, nbgmm, 'diag');
+% options = foptions;
+options(14) = 20;	% Just use 20 iterations of k-means in initialisation
+options(1)  = -1;
+mix = gmminit(mix, data, options);
+options = zeros(1, 18);
+options(1)  = -1;		% Prints out error values.
+options(14) = 100;		% Number of iterations.
+[mix, options, errlog] = gmmem(mix, data, options);
+
+% note: if netlab prints a "warning:divide by zero", then the resulting gmm is probably corrupted. 
+% Try calling "createGMM" again with the same arguments. 
+% Such numerical errors often happen in local minima, so likely that a simple rerun will avoid the trap.

### 2018-02-09 experiments/unsupervised/boflib/demo.m

@@ -1,0 +1,51 @@
+%% a demo fo the BOF package
+%% Copyright (C) JJ Aucouturier, 2008
+%% credits: Mike Brookes for MFCC code, Chris Bishop for GMM code
+
+fprintf(1,'compute mfcc for market soundscape\n');
+[audio,fs]=wavread('data/Market 004-a-001-extrait4.mp3.wav');
+data1 = computeMfccs(audio, fs, 20);
+fprintf(1,'compute mfcc for boulevard1 soundscape\n');
+[audio,fs]=wavread('data/Boulevard 002-001-extrait10.mp3.wav');
+data2 = computeMfccs(audio, fs, 20);
+fprintf(1,'compute mfcc for boulevard2 soundscape\n');
+[audio,fs]=wavread('data/Boulevard 006-a-002-extrait6.mp3.wav');
+data3 = computeMfccs(audio,fs, 20);
+
+fprintf(1,'compute gmm for market soundscape\n');
+mix1 = createGmm(data1, 10);
+fprintf(1,'compute gmm for boulevard1 soundscape\n');
+mix2 = createGmm(data2, 10);
+fprintf(1,'compute gmm for boulevard2 soundscape\n');
+mix3 = createGmm(data3, 10);
+fprintf(1,'compute gmm for boulevard class\n');
+mix23 = createGmm([data2;data3], 10);
+
+fprintf(1,'save gmm for market soundscape\n');
+saveGmm(mix1,'gmms/market.gmm');
+fprintf(1,'save gmm for boulevard1 soundscape\n');
+saveGmm(mix2,'gmms/boulevard1.gmm');
+fprintf(1,'save gmm for boulevard2 soundscape\n');
+saveGmm(mix3,'gmms/boulevard2.gmm');
+fprintf(1,'save gmm for boulevard class\n');
+saveGmm(mix23,'gmms/boulevard-class.gmm');
+
+fprintf(1,'load gmm for market soundscape\n');
+mix1=loadGmm('gmms/market.gmm',20,10);
+fprintf(1,'load gmm for boulevard1 soundscape\n');
+mix2=loadGmm('gmms/boulevard1.gmm',20,10);
+fprintf(1,'load gmm for boulevard2 soundscape\n');
+mix3=loadGmm('gmms/boulevard2.gmm',20,10);
+fprintf(1,'load gmm for boulevard class\n');
+mix23=loadGmm('gmms/boulevard-class.gmm',20,10);
+
[…]

### 2018-02-09 experiments/unsupervised/boflib/distanceDataToModel.m

@@ -1,0 +1,20 @@
+function d = distanceDataToModel(data,model)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% computes the averaged log-likelihood of the data given the model
+%% data should be similar to the output of data = computeMfccs(audiofilename, nbcoef) (i.e. a N*M matrix of floats)
+%% model should be similar to the output of mix = createGmm(data, nbgmm) (i.e. a netlab struct with a gmmm)
+%% and the dimensionality of data should be compatible with that of model
+%% outputs a float, scaled between 0 and +inf, which can be interpreted as a distance (i.e. 0 : best match)
+%% what to do with this :
+%% 1) Classification (1)
+%% to do ML classification, compute the distanceDataToModel(data,model_i) of a given data set 
+%% to a set of models trained to represent different classes, and assign the class corresponding to the model which gives the smallest "distance" (i.e. the best likelihood)
+%% 2) Similarity
+%% use this function as in distanceModelToModel to compute a distance score between the models of 2 audio signals (note : this is not a distance in the mathematically sense, i.e. only approximates triangular inequality, etc.)
+%% 3) Classification (2)
+%% use the similarity function above to retrieve all nearest neighbors to a given signal in a dataset, and do nearest neighbor classification, i.e. assign the class which is most represented in the -say 10 - nearest neighbors
+%% Copyright (C) JJ Aucouturier, 2008
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+nsamples = size(data,1);
+d=-sum(log(gmmprob(model, data)))/nsamples;

### 2018-02-09 experiments/unsupervised/boflib/distanceModelToModel.m

@@ -1,0 +1,17 @@
+function d=distanceModelToModel(model1, model2, nsamples)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% computes a monte-carlo estimate of the Kullback-Leibler distance between two gmms
+%% both models in argument should be similar to the output of mix = createGmm(data, nbgmm) (i.e. a netlab struct with a gmmm)
+%% and have compatible dimensionality
+%% nsamples is the number of samples used to estimate the distribution. Recommended value= 2000. (linear cpu dependency, log precision)
+%% can be used to compute a similarity measure between songs, based on their BOF distribution
+%% Copyright (C) JJ Aucouturier, 2008
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% sample from models
+samples1 = gmmsamp(model1,nsamples);
+samples2 = gmmsamp(model2,nsamples);
+% compute symmetric distance
+a=distanceDataToModel(samples2,model1);
+b=distanceDataToModel(samples1,model2);
+d=(a+b)/2;

### 2018-02-09 experiments/unsupervised/boflib/loadGmm.m

@@ -1,0 +1,14 @@
+function mix=loadGmm(gmmfilename, nbcoef,nbgmm)
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ex. to load a packed-and-saved gmm from a file ("gmm/mymodel.gmm") to a natlab struct, 
+%% knowing it's a 50-state gmm constructed from 20-dim mfcc,  call  
+%% mix=loadGmm('gmm/mymodel.gmm', 20,50)
+%% mix is the netlab struct containing gmm parameters
+%% mymodel.gmm is a mat file containing a netlab-packed version of mix (see saveGmm)
+%% Copyright (C) JJ Aucouturier, 2008
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+fprintf(1,'- load gmm from file\n');
+load(gmmfilename,'-mat');
+premix=gmm(nbcoef,nbgmm,'diag');
+mix = gmmunpak(premix, p);

### 2018-02-09 experiments/unsupervised/boflib/netlab - put in path/gmminit.m

@@ -1,0 +1,93 @@
+function mix = gmminit(mix, x, options)
+%GMMINIT Initialises Gaussian mixture model from data
+%
+%	Description
+%	MIX = GMMINIT(MIX, X, OPTIONS) uses a dataset X to initialise the
+%	parameters of a Gaussian mixture model defined by the data structure
+%	MIX.  The k-means algorithm is used to determine the centres. The
+%	priors are computed from the proportion of examples belonging to each
+%	cluster. The covariance matrices are calculated as the sample
+%	covariance of the points associated with (i.e. closest to) the
+%	corresponding centres. For a mixture of PPCA model, the PPCA
+%	decomposition is calculated for the points closest to a given centre.
+%	This initialisation can be used as the starting point for training
+%	the model using the EM algorithm.
+%
+%	See also
+%	GMM
+%
+
+%	Copyright (c) Ian T Nabney (1996-2001)
+
+[ndata, xdim] = size(x);
+
+% Check that inputs are consistent
+errstring = consist(mix, 'gmm', x);
+if ~isempty(errstring)
+  error(errstring);
+end
+
+% Arbitrary width used if variance collapses to zero: make it 'large' so
+% that centre is responsible for a reasonable number of points.
+GMM_WIDTH = 1.0;
+
+% Use kmeans algorithm to set centres
+options(5) = 1;
+[mix.centres, options, post] = nlkmeans(mix.centres, x, options);
+
+% Set priors depending on number of points in each cluster
+cluster_sizes = max(sum(post, 1), 1);  % Make sure that no prior is zero
+mix.priors = cluster_sizes/sum(cluster_sizes); % Normalise priors
+
+switch mix.covar_type
+case 'spherical'
+   if mix.ncentres > 1
+      % Determine widths as distance to nearest centre
+      % (or a constant if this is zero)
+      cdist = dist2(mix.centres, mix.centres);
+      cdist = cdist + diag(ones(mix.ncentres, 1)*realmax);
+      mix.covars = min(cdist);
+      mix.covars = mix.covars + GMM_WIDTH*(mix.covars < eps);
+   else
+      % Just use variance of all data points averaged over all
+      % dimensions
+      mix.covars = mean(diag(cov(x)));
+   end
+  case 'diag'
+    for j = 1:mix.ncentres
[…]

### 2018-02-09 experiments/unsupervised/boflib/netlab - put in path/nlkmeans.m

@@ -1,0 +1,123 @@
+function [centres, options, post, errlog] = nlkmeans(centres, data, options)
+%KMEANS	Trains a k means cluster model.
+%
+%	Description
+%	 CENTRES = KMEANS(CENTRES, DATA, OPTIONS) uses the batch K-means
+%	algorithm to set the centres of a cluster model. The matrix DATA
+%	represents the data which is being clustered, with each row
+%	corresponding to a vector. The sum of squares error function is used.
+%	The point at which a local minimum is achieved is returned as
+%	CENTRES.  The error value at that point is returned in OPTIONS(8).
+%
+%	[CENTRES, OPTIONS, POST, ERRLOG] = KMEANS(CENTRES, DATA, OPTIONS)
+%	also returns the cluster number (in a one-of-N encoding) for each
+%	data point in POST and a log of the error values after each cycle in
+%	ERRLOG.    The optional parameters have the following
+%	interpretations.
+%
+%	OPTIONS(1) is set to 1 to display error values; also logs error
+%	values in the return argument ERRLOG. If OPTIONS(1) is set to 0, then
+%	only warning messages are displayed.  If OPTIONS(1) is -1, then
+%	nothing is displayed.
+%
+%	OPTIONS(2) is a measure of the absolute precision required for the
+%	value of CENTRES at the solution.  If the absolute difference between
+%	the values of CENTRES between two successive steps is less than
+%	OPTIONS(2), then this condition is satisfied.
+%
+%	OPTIONS(3) is a measure of the precision required of the error
+%	function at the solution.  If the absolute difference between the
+%	error functions between two successive steps is less than OPTIONS(3),
+%	then this condition is satisfied. Both this and the previous
+%	condition must be satisfied for termination.
+%
+%	OPTIONS(14) is the maximum number of iterations; default 100.
+%
+%	See also
+%	GMMINIT, GMMEM
+%
+
+%	Copyright (c) Ian T Nabney (1996-2001)
+
+[ndata, data_dim] = size(data);
+[ncentres, dim] = size(centres);
+
+if dim ~= data_dim
+  error('Data dimension does not match dimension of centres')
+end
+
+if (ncentres > ndata)
+  error('More centres than data')
+end
+
+% Sort out the options
+if (options(14))
+  niters = options(14);
+else
+  niters = 100;
+end
+
+store = 0;
+if (nargout > 3)
+  store = 1;
[…]

### 2018-02-09 experiments/unsupervised/boflib/saveGmm.m

@@ -1,0 +1,14 @@
+function saveGmm(mix,gmmfilename)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ex. to save a newly created gmm to a file ("gmm/mymodel.gmm"), call 
+%% saveGmm(mix, 'gmm/mymodel.gmm')
+%% mix is the netlab struct containing gmm parameters
+%% mymodel.gmm is a mat file containing a netlab-packed version of mix (see loadGmm)
+%% Copyright (C) JJ Aucouturier, 2008
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+fprintf(1,'- save gmm in file\n');
+p=gmmpak(mix);
+% note: if filename is a path including directories, all directories should exist before calling saveGmm
+save(gmmfilename,'p');

### 2018-02-09 experiments/unsupervised/boflib/voicebox - put in path/enframe.m

@@ -1,0 +1,39 @@
+function f=enframe(x,win,inc)
+%ENFRAME split signal up into (overlapping) frames: one per row. F=(X,WIN,INC)
+%
+%	F = ENFRAME(X,LEN) splits the vector X up into
+%	frames. Each frame is of length LEN and occupies
+%	one row of the output matrix. The last few frames of X
+%	will be ignored if its length is not divisible by LEN.
+%	It is an error if X is shorter than LEN.
+%
+%	F = ENFRAME(X,LEN,INC) has frames beginning at increments of INC
+%	The centre of frame I is X((I-1)*INC+(LEN+1)/2) for I=1,2,...
+%	The number of frames is fix((length(X)-LEN+INC)/INC)
+%
+%	F = ENFRAME(X,WINDOW) or ENFRAME(X,WINDOW,INC) multiplies
+%	each frame by WINDOW(:)
+
+%	Copyright (C) Mike Brookes 1997
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+nx=length(x);
+nwin=length(win);
+if (nwin == 1)
+   len = win;
+else
+   len = nwin;
+end
+if (nargin < 3)
+   inc = len;
+end
+nf = fix((nx-len+inc)/inc);
+f=zeros(nf,len);
+indf= inc*(0:(nf-1)).';
+inds = (1:len);
+f(:) = x(indf(:,ones(1,len))+inds(ones(nf,1),:));
+if (nwin > 1)
+    w = win(:)';
+    f = f .* w(ones(nf,1),:);
+end

### 2018-02-09 experiments/unsupervised/boflib/voicebox - put in path/melbankm.m

@@ -1,0 +1,87 @@
+function [x,mn,mx]=melbankm(p,n,fs,fl,fh,w)
+%MELBANKM determine matrix for a mel-spaced filterbank [X,MN,MX]=(P,N,FS,FL,FH,W)
+%
+% Inputs:	p   number of filters in filterbank
+%		n   length of fft
+%		fs  sample rate in Hz
+%		fl  low end of the lowest filter as a fraction of fs (default = 0)
+%		fh  high end of highest filter as a fraction of fs (default = 0.5)
+%		w   any sensible combination of the following:
+%		      't'  triangular shaped filters in mel domain (default)
+%		      'n'  hanning shaped filters in mel domain
+%		      'm'  hamming shaped filters in mel domain
+%
+%		      'z'  highest and lowest filters taper down to zero (default)
+%		      'y'  lowest filter remains at 1 down to 0 frequency and
+%			   highest filter remains at 1 up to nyquist freqency
+%
+%		       If 'ty' or 'ny' is specified, the total power in the fft is preserved.
+%
+% Outputs:	x     a sparse matrix containing the filterbank amplitudes
+%		      If x is the only output argument then size(x)=[p,1+floor(n/2)]
+%		      otherwise size(x)=[p,mx-mn+1]
+%		mn   the lowest fft bin with a non-zero coefficient
+%		mx   the highest fft bin with a non-zero coefficient
+%
+% Usage:	f=fft(s);			f=fft(s);
+%		x=melbankm(p,n,fs);		[x,na,nb]=melbankm(p,n,fs);
+%		n2=1+floor(n/2);		z=log(x*(f(na:nb)).*conj(f(na:nb)));
+%		z=log(x*abs(f(1:n2)).^2);
+%		c=dct(z); c(1)=[];
+%
+% To plot filterbanks e.g.	plot(melbankm(20,256,8000)')
+
+% Copyright (C) Mike Brookes 1997
+%
[…]

### 2018-02-09 experiments/unsupervised/boflib/voicebox - put in path/mfc.m

@@ -1,0 +1,102 @@
+function [c,y, f]=mfc(z,fs,w,nc,p,n,inc,fl,fh)
+%MELCEPST Calculate the mel cepstrum of a framed signal C=(Z,FS,W,NC,P,N,INC,FL,FH)
+%
+%
+% Simple use: c=melcepst(Z,fs)	% calculate mel cepstrum with 12 coefs, 256 sample frames
+%				  c=melcepst(s,fs,'e0dD') % include log energy, 0th cepstral coef, delta and delta-delta coefs
+%
+% Inputs:
+%     z	 framed signal 
+%     fs  sample rate in Hz (default 11025)
+%     nc  number of cepstral coefficients excluding 0'th coefficient (default 12)
+%     n   length of frame (default power of 2 <30 ms))
+%     p   number of filters in filterbank (default floor(3*log(fs)) )
+%     inc frame increment (default n/2)
+%     fl  low end of the lowest filter as a fraction of fs (default = 0)
+%     fh  high end of highest filter as a fraction of fs (default = 0.5)
+%
+%		w   any sensible combination of the following:
+%
+%		      't'  triangular shaped filters in mel domain (default)
+%		      'n'  hanning shaped filters in mel domain
+%		      'm'  hamming shaped filters in mel domain
+%
+%				'p'	filters act in the power domain
+%				'a'	filters act in the absolute magnitude domain (default)
+%
+%			   '0'  include 0'th order cepstral coefficient
+%			    'e'  include log energy
+%			
+%		      'z'  highest and lowest filters taper down to zero (default)
+%		      'y'  lowest filter remains at 1 down to 0 frequency and
+%			   	  highest filter remains at 1 up to nyquist freqency
+%
+%		       If 'ty' or 'ny' is specified, the total power in the fft is preserved.
+%
+% Outputs:	c     mel cepstrum output: one frame per row
+%			
+%  Copyright (C) Mike Brookes 1997
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
[…]

### 2018-02-09 experiments/unsupervised/taun1features.m

@@ -1,0 +1,121 @@
+function [config, store, obs] = taun1features(config, setting, data)
+% taun1features FEATURES step of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store, obs] = taun1features(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, unsupervised('do', 1, 'mask', {}); return; else store=[]; obs=[]; end
+
+%% setting
+store.xp_settings.sr=44100;
+store.xp_settings.classes = {'bus','busystreet','office','openairmarket','park','quietstreet','restaurant','supermarket','tube','tubestation'};
+
+switch setting.features
+    case 'scatT'
+        
+        fileId = fopen([config.inputPath 'sampleList_' setting.dataset '.txt']);
+        store.xp_settings.sounds=textscan(fileId,'%s');store.xp_settings.sounds=store.xp_settings.sounds{1};
+        fclose(fileId);
+        
+        store.xp_settings.sounds=cellfun(@(x) x(strfind(x,'/')+1:end),store.xp_settings.sounds,'UniformOutput',false);
+        
+        store.soundIndex=[];
+        store.indSample=[];
+        store.class=[];
+        store.dataset=[];
+        eval(['store.features.' setting.features '=[];']);
+        
+                load([config.inputPath 'scattering/dcase2013_timeQ8_' setting.dataset '.mat']);
+                load([config.inputPath 'scattering/dcase2013_timeQ8_freqs.mat']);
+                eval(['Y=dcase2013_timeQ8_' setting.dataset '.Y_' setting.dataset ';']);
+                eval(['X=dcase2013_timeQ8_' setting.dataset '.X_' setting.dataset ';']);
+                eval(['clearvars dcase2013_timeQ8_' setting.dataset ';']);
+                
+                store.xp_settings.Xfreqs=dcase2013_timeQ8_freqs;
+                X=squeeze(X(:,:,3,:));
+                
+        store.class=Y(:)'+1;
+        store.soundIndex=1:length(Y);
+        
+        nbTrame=size(X,2);
+        
+        for ii=1:size(X,3)
[…]

### 2018-02-09 experiments/unsupervised/taun2norm.m

@@ -1,0 +1,68 @@
+function [config, store, obs] = taun2norm(config, setting, data)
+% taun2norm NORM step of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store, obs] = taun2norm(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, talspStruct2016_unsupervised('do', 2, 'mask', {}); return; else store=[]; obs=[]; end
+
+%% store
+
+store.xp_settings=data.xp_settings;
+store.soundIndex=data.soundIndex;
+store.class=data.class;
+
+%% get features
+
+X=[];
+features2use=strsplit(setting.features,'_');
+for jj=1:length(features2use)
+    eval(['X=[X ; data.features.' features2use{jj} '];']);
+end
+
+%% pooling
+
+switch setting.features
+    case 'mfcc'
+        store.features=[];
+        store.indSample=[];
+        for jj=unique(data.indSample)
+            X_tmp=X(:,data.indSample==jj);
+            step=round(0.250/store.xp_settings.hoptime);
+            onsets=1:step:size(X_tmp,2);
+            offsets=[onsets(2:end)-1 size(X_tmp,2)];
+            X_tmp=featuresPooling(X_tmp,'mean',onsets,offsets,0,0,store.xp_settings.hoptime);
+            store.features = [store.features X_tmp];
+            store.indSample=[store.indSample ones(1,size(X_tmp,2))*jj];
+        end
+    otherwise
+        store.features=X;
+        store.indSample=data.indSample;
+end
+
+clearvars X
+
+%% Rectify
+
+switch setting.features
+    case 'scatT'
+        store.features(store.features<0)=0;
+end
+
+%% Normalisation
+
+switch setting.features
+    case 'scatT'
+        params.type='scattering';
+        params.ftrsNorm_scat_threshold = 1e2;
+        params.ftrsNorm_scat_selection = 1;
+        params.ftrsNorm_scat_log =  setting.scat_log;
+        [store.features,~] = normFtrs(store.features,params);
+end

### 2018-02-09 experiments/unsupervised/taun3clustering.m

@@ -1,0 +1,107 @@
+function [config, store, obs] = taun3clustering(config, setting, data)
+% taun3clustering CLUSTERING step of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store, obs] = taun3clustering(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, unsupervised('do', 3, 'mask', {2 0 2 3 }); return; else store=[]; obs=[]; end
+
+%% seed
+
+rng(0);
+
+%% store
+
+store.xp_settings=data.xp_settings;
+store.soundIndex=data.soundIndex;
+store.class=data.class;
+
+store.weight=[];
+store.centroid=[];
+store.indSample=[];
+store.centroidClass=[];
+store.prediction=[];
+
+%% integration
+
+switch setting.integration
+    case 'clustering'
+        
+        params.clustering='kmeans';
+
+        for jj=unique(data.indSample)
+            
+            params.similarity='sqeuclidean';
+            params.nbc=setting.clustering_nbc;
+            params.rep=1;
+            params.emptyAction='singleton';
+            
+            [prediction,centroid] = kmeans(full(data.features(:,data.indSample==jj))',params.nbc,'maxiter',1000,'replicates',params.rep,'start','plus','Distance',params.similarity,'EmptyAction',params.emptyAction);
+            centroid=centroid';
+            %[prediction,centroid,params] = featuresBasedClustering(data.features(:,data.indSample==jj),params);
+            
+            %% weight
+            
+            params.histType='';
+            params.nbc=size(centroid,2);
+            [weight] = histFeatures(prediction,ones(1,length(prediction)),params);
+            
+            %% Norm
+            
+            % weight=weight./sum(weight);
+            
+            %% store
+            store.prediction=[store.prediction prediction+setting.clustering_nbc*(jj-1)];
+            store.weight=[store.weight weight];
+            store.centroid=[store.centroid centroid];
+            store.indSample=[store.indSample ones(1,size(centroid,2))*jj];
+            store.centroidClass=[store.centroidClass ones(1,size(centroid,2))*store.class(jj)];
+        end
+        
+    case 'early'
+        
[…]

### 2018-02-09 experiments/unsupervised/taun4similarity.m

@@ -1,0 +1,74 @@
+function [config, store, obs] = taun4similarity(config, setting, data)
+% taun4similarity SIMILARITY step of the expLanes experiment talspStruct2016_unsupervised
+%    [config, store, obs] = taun4similarity(config, setting, data)
+%      - config : expLanes configuration state
+%      - setting   : set of factors to be evaluated
+%      - data   : processing data stored during the previous step
+%      -- store  : processing data to be saved for the other steps
+%      -- obs    : observations to be saved for analysis
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+% Set behavior for debug mode
+if nargin==0, unsupervised('do', 4:5, 'mask', {2 0 2 3}); return; else store=[]; obs=[]; end
+
+%% store
+
+store.xp_settings=data.xp_settings;
+store.soundIndex=data.soundIndex;
+store.class=data.class;
+
+store.weight=data.weight;
+store.indSample=data.indSample;
+store.centroid=data.centroid;
+store.centroidClass=data.centroidClass;
+store.prediction=data.prediction;
+
+%% get clusters Similarity
+
+params.mode='centroid';
+params.similarity='rbf';
+params.nn=setting.similarity_nn;
+params.class=store.centroidClass;
+
+
+switch setting.integration
+    
+    case 'clustering'
+        simClus= computeSimilarity(store.centroid,params);
+
+        %% reshape weights
+        
+        onsets=1:size(data.weight,1):size(simClus,2)-size(data.weight,1)+1;
+        offsets=[onsets(2:end)-1 size(simClus,2)];
+        weight=zeros(size(simClus,2),size(data.weight,2));
+        
+        for jj=1:size(weight,2)
+            weight(onsets(jj):offsets(jj),jj)=data.weight(:,jj);
+        end
+        
+        %% scenes similarity
+        
+        params.histDist=setting.similarity_dist;
+        params.indSample=store.indSample;
+        
+        [store.A,store.params] = histSimilarity(weight,params,simClus);
+    case 'gmm'
+        d=[];
+        for k=1:length(data.model)
+            vec=[];
+            parfor l=k+1:length(data.model)
+                vec(l) = distanceModelToModel(data.model{k}, data.model{l}, 2000);
+            end
+            for l=k+1:length(data.model)
+                d(k, l) = vec(l);
+                d(l, k) = d(k, l);
+            end
+        end
+        store.A = 1-d;
+    otherwise
[…]

### 2018-02-09 experiments/unsupervised/taun5metrics.m

@@ -1,0 +1,25 @@
+function [config, store, obs] = taun5metrics(config, setting, data)                                
+% taun5metrics METRICS step of the expLanes experiment talspStruct2016_unsupervised                
+%    [config, store, obs] = taun5metrics(config, setting, data)                                    
+%      - config : expLanes configuration state                                                     
+%      - setting   : set of factors to be evaluated                                                
+%      - data   : processing data stored during the previous step                                  
+%      -- store  : processing data to be saved for the other steps                                 
+%      -- obs    : observations to be saved for analysis                                           
+                                                                                                   
+% Copyright: gregoirelafay                                                                         
+% Date: 17-Dec-2016                                                                                
+                                                                                                   
+% Set behavior for debug mode                                                                      
+if nargin==0, unsupervised('do', 5, 'mask', {2 0 2}); return; else store=[]; obs=[]; end
+                                                                                                   
+D=1-data.A/max(data.A(:));
+% k2look=[1 2 3 4 5 6 7 8 9];
+% 
+% for jj=k2look
+%     rm = rankingMetrics(D, data.class,jj);
+%     eval(['obs.pa' num2str(jj) '=rm.precisionAt' num2str(jj) ';']);
+% end
+
+rm = rankingMetrics(D, data.class,setting.precision);
+eval(['obs.p=rm.precisionAt' num2str(setting.precision) ';']);

### 2018-02-09 experiments/unsupervised/taunReport.m

@@ -1,0 +1,52 @@
+function config = taunReport(config)
+% taunReport REPORTING of the expLanes experiment talspStruct2016_unsupervised
+%    config = taunInitReport(config)
+%       config : expLanes configuration state
+
+% Copyright: gregoirelafay
+% Date: 17-Dec-2016
+
+if nargin==0, unsupervised('report', 'r', 'reportName', 'ml'); return; end
+
+config = expExpose(config, 'p', 'mask', {2 0 2 [1 3] 0 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1, 'save', 'baselines');
+config = expExpose(config, 'p', 'mask', {2 0 2 [1 2] 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1, 'save', 'overall');
+config = expExpose(config, 'p', 'mask', {2 2 0 [1 2] 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1, 'save', 'log');
+return
+
+features = {'mfcc','scatT'};
+scat_log = {'','log'};
+integration = {'early','clustering'};
+similarity_dist = {'emd','average','closest'};
+
+for a=1:length(features)
+    for c=1:length(integration)   
+        switch integration{c} 
+            case 'early'
+                switch features{a}
+                    case 'mfcc'
+                        config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a 0 c 0 [1 2 3] 0},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                            ['unsupervised_test_' features{a} '_' integration{c}]);
+                    case 'scatT'
+                        for b=1:length(scat_log)
+                            config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a b c 0 [1 2 3] 0},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
+                                ['unsupervised_test_' scat_log{b} features{a} '_' integration{c}]);
+                        end
+                end  
+            case 'clustering'
+                for d=1:length(similarity_dist)
+                    switch features{a}
+                        case 'mfcc'
+                            config = expExpose(config, 't','fontSize','scriptsize','step', 5, 'mask',{2 a 0 c 0 [1 2 3] d},'obs',[1 2 3 4 5 6 7 8 9],'precision', 2,'save',1,'name', ...
[…]

### 2018-02-09 experiments/unsupervised/unsupervised.m

@@ -1,0 +1,190 @@
+function taslpStruct2016_unsupervised(varargin)                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                 
+% Welcome to the main entry point of talspStruct2016_unsupervised                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               
+% Please DO NOT modify this file unless you have a precise intent.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
+shortExperimentName = 'taun';                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                   
+[p, experimentName] = fileparts(mfilename('fullpath'));                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
[…]

### 2018-02-09 paper/structureScene.tex
add section about pca

@@
 The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene.
-The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \textit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization algorithm.
+The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \textit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization (EM) algorithm.
 Gaussian models are compared using Monte-Carlo sampling approaximation.
+In order to ensure convergence of the EM algorithm when considering the scattering features as input, the dimension of the feature vector is reduced from 1367 to 30 by projecting the features over the 30 principal components computed over the whole dataset.
 
@@
 \includegraphics[width=\columnwidth]{figures/overall}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.
-\ml{add gmms}}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
 \label{fig:ASS_1}

### 2018-02-15 paper/structureScene.tex
comment on Figure 3

@@
 \includegraphics[width=\columnwidth]{figures/baselines}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
+\vl{The x-axis label is wrong. precision should be replaced by "rank $k$". The y-axis should be "precision ($p@k$)". It would be good to have the y-axis range from $0\%$ to $100\%$.
+The lines should be thicker and follow a color palette of complementary colors (blue vs. orange). The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC. }}
 \label{fig:ASS_-1}

### 2018-02-15 paper/structureScene.tex
the verb utilize should be avoided
comments on writing

@@
 We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval.
-Testing this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
+\vl{The verb use is repeated in the previous sentence.} Testing this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
 
@@
 These closest neighbors are determined by a particular distance, which is one of \emph{RbQ-c}, \emph{RbQ-a}, or \emph{RbQ-w}.
+\vl{Avoid passive voice.}
 
@@
 In order to ensure convergence of the EM algorithm when considering the scattering features as input, the dimension of the feature vector is reduced from 1367 to 30 by projecting the features over the 30 principal components computed over the whole dataset.
+\vl{Too much passive voice, too many instances of "use", "used".}
 
@@
 These outcomes shall be studied further in future work by considering larger databases and emerging tasks in computational bioacoustics \cite{wimmer2013sampling}.
+\vl{The verb utilize should be avoided.}
 

### 2018-02-15 paper/structureScene.tex
more comments on writing
typo

@@
 The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \textit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization (EM) algorithm.
-Gaussian models are compared using Monte-Carlo sampling approaximation.
+Gaussian models are compared using Monte-Carlo sampling approximation.
 In order to ensure convergence of the EM algorithm when considering the scattering features as input, the dimension of the feature vector is reduced from 1367 to 30 by projecting the features over the 30 principal components computed over the whole dataset.
@@
 \ml{This is a bit of a stretch, since it may just be that we used the wrong distance to take the weighting into account.
-Perhaps the non-normalize d EMD is not appropriate here?}
+Perhaps the non-normalized EMD is not appropriate here?}
 
@@
 Considering $p@5$ as our metric (as in \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), the use of the log-scattering transform instead of MFCCs increases the performance from $0.31$ to $0.49$ and the use of the relevance-based quantization approach using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
+\vl{Too many instances of ``use'', ``use'', ``using''.
+It is worth reminding what the baselines are.}
 

### 2018-02-16 paper/structureScene.tex
Edit abstract

@@
 The emerging field of computational ecoacoustics aims at retrieving high-level information from acoustic scenes recorded by some network of sensors.
-Large amounts of unlabeled data are gathered using such networks that needs to be analyzed.
-In order to identify which parts require human inspection, one needs tools that can mine data and identify structures such as recurring patterns or isolated events.
-To do so, one needs to be able to measure the level of similarity among the data without strong assumptions.
+These networks gather large amounts of data requiring analysis.
+To decide which parts to inspect manually, we need tools that automatically mine the data, identifying recurring patterns and isolated events.
+This requires a similarity measure for acoustic scenes that does not impose strong assumptions on the data.
 
-State of the art approaches for computing such a similarity is the ``bag-of-frames'' approach, i.e., modeling the audio signal using the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
-Although they successfully characterize static scenes with little variability in auditory content, they are unable to accurately discriminate scenes with few salient events superimposed over a textured background.
-To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure that better handle the required modeling of sparse events.
-Experiments performed using the acoustic scene similarity framework demonstrates the superiority of the proposed approach.
+The state of the art in audio similarity measurement is the ``bag-of-frames'' approach, which models a recording using summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
+They successfully characterize static scenes with little variability in auditory content, but cannot accurately capture scenes with a few salient events superimposed over static background.
+To overcome this issue, we propose a two-scale representation which describes a recording using clusters of scattering coefficients.
+The scattering coefficients capture short-scale structure, while the cluster model captures longer time scales, allowing for more accurate characterization of sparse events.
+Evaluation within the acoustic scene similarity framework demonstrates the superiority of the proposed approach.
 \keywords{

### 2018-02-17 paper/structureScene.tex
Edit introduction

@@
 In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
-Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.
+Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for to signal processing community.
 
-This work falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
-%Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
-%Consequently, few well-designed datasets are readily available for evaluation purposes.
+These questions are explored in an emerging field of computational bioacoustics, known as ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
 An important problem in this field is that manually analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling}.
-Some sort of pre-screening is therefore needed to reduce the need for human expert listening and annotation.
+Some sort of pre-screening is therefore required to reduce the need for human expert listening and annotation.
 
-The most straightforward approach is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}.
-A given time interval, for example a single day, is then represented by the number of events within the predefined typology that have been detected in that period of time.
-This approach allows the scientist to drastically reduce the amount of information that needs to be processed.
+The most straightforward approach is to specify a closed set of sound classes, such as sounds of animals expected to live near the acoustic sensors.
+We then train models for these classes which are used to automatically annotate our recordings \cite{}.
+\ja{What's the reference here?} A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class.
+This allows the scientist to drastically reduce the amount of information requiring manual processing.
 However, this approach has two drawbacks.
-The first is the reliance to trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be always be trusted.
-More importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where the quantities of interest have yet to be defined.
+First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be trusted.
+Second, and more importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.
 
-In order to identify which parts needs human inspection, one needs tools that can mine data and annotate to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors.
-The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.
+To identify which parts need human inspection, we need tools that can detect both recurring patterns and sparsely distributed events.
+Identifying recurring patterns allows the user to focus on certain time points for manual annotation, while detection of more rare structures enables discovery of unforeseen phenomena.
 
-With this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset.
-This requires the audio fragments to be represented in some way that captures their distinctive qualities.
-This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics.
-Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented with scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others.
-Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely human acoustic perception of acoustic scenes.
-We believe that the failure to model such distinct events is one of the reason why this approach is found to be insufficient when considering evaluation datasets of typical sizes \cite{lagrange:hal-01082501}.
+With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset.
+This requires us to represent an audio recording in a way that captures its distinctive qualities.
+This has previously been attempted using bag-of-frames approaches \cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
+Unfortunately, this only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterized by a few distinct sound events.
+Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neuroscience \cite{nelken2004processing} suggest that this more closely matches human acoustic perception.
+We believe that the failure to model such distinct events is one of the reasons why this approach is insufficient for moderate-sized evaluation datasets \cite{lagrange:hal-01082501}.
 
-%To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.
-This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}.
-In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.
+Solving the acoustic similarity retrieval first requires the ability to capture meaningful signal structure at small time scales.
+This is often achieved using mel-frequency cepstral coefficients (MFCCs).
+Originally developed for speech processing \cite{davis-mermelstein}, MFCCs have recently found wider use in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
+A richer representation, the scattering transform, has enjoyed significant success in various audio \cite{Anden2014} and biomedical \cite{chudacek} signal classification tasks.
+Its structure is that of a convolutional neural network \cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}, but with fixed filters.
+Specifically, it alternates convolutions with wavelet filters and pointwise nonlinearities to ensure time-shift invariance and time-warping stability \cite{Mallat2012}.
 
-Addressing this task first requires the ability to capture meaningful aspects within the signal at small time-scales.
-In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
-A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
-A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014,chudacek}.
-This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}.
+For our task, one advantage of the scattering transform is not requiring a training step, allowing for a wider range of applications compared to learned features.
+Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unserpervised setting.
 
-For the task at at hand, scattering features have the advantage of not requiring training and thus potentially lead to wider range of application than trained features.
-Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined, thus leading to an unserpervised setting.
-
-%The next question is the choice of a classifier.
-Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM) \cite{cristianini2000introduction}, to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision.
-This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes.
-The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it almost completely.
-Combining scattering transforms for low-level representation with late-integrated SVMs for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.
-
-%we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
-We also propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters.
-We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval.
-\vl{The verb use is repeated in the previous sentence.} Testing this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
+In this work, we propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an cluster model.
+This unsupervised model quantizes the scattering coefficients into a given number of clusters.
+These clusters to define a set of distances for acoustic similarity retrieval.
+Evaluating this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
 
@@
 Section \ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task.
-Results are subsequently reported in Section \ref{sec:results}.
+Results are reported in Section \ref{sec:results}.
 

### 2018-02-17 paper/structureScene.tex
Edit background section

@@
 
-%One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify species properties of vocalizations, for example to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
+Computational bioacoustics refers to the numerical investigation of sound production, dispersion and reception in animals, including humans.
+A recent paradigm processes the audio stream in a holistic way over large time scales, without assuming that a single species is present throughout the recording.
+Automated systems within this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010} and migration patterns \cite{Obrist2010}.
+They also mark time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.
 
-Computational bioacoustics usually refers to the numerical investigation of sound production, dispersion and reception in animals (including humans).
-A recent paradigm  processes the audio stream in a holistic way over large time-scales, without assuming that a single species is present throughout the recording.
-Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as marking time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.
-
-%The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
-In the closely related field of urban sound environment analysis, a popular approach is the bag of frames (BOF) introduced in this context by Aucouturier et al.
-\cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
-Recently, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
-This contrasts with the typical morphology of acoustic scenes, which we believe is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events are superimposed upon a stationary acoustic background.
-As those distinct and sparse events are often highly relevant for the task at hand, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on summary statistics of short-term features.
+A closely related field is that of urban sound environment analysis.
+In this context, a popular approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al. \cite{aucouturier2007bag}.
+It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.
+Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
+This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background \cite{nelken2013}.
+Such events are not well-characterized by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
+The latter approach should therefore prove more fruitful in measuring auditory scene similarity.
 
 This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}.
-For example, studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene.
+Studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Instead, cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} seem to rely upon higher-level cognitive attributes.
+These typically include the identities of the sound sources which consitute the scene.
 It has been shown that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes.
-For example, the presence of birds is very likely to be heard in parks in urban areas and so are strong pleasantness indicators.
+For example, in urban areas the presence of birds is likely to be heard in parks and are therefore strong pleasantness indicators.
 Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments \cite{ricciardi2015sound,lavandier2006contribution}.
@@
 
-From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation the same and do not incorporate temporal structure between features.
-A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes, provided that the representation of the representation of those distinct events is rich enough.
-
-%Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of acoustic scene similarity retrieval, much of the work dedicated to acoustic scene modeling is applied to the classification task.
-%An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
-
-%The methods that performed the best all follow a similar structure, where low-level feature representations, such as MFCCs, are calculated over short windows and fed separately to a classifier, such as a support vector machine (SVM) \cite{cortes-vapnik,cristianini2000introduction}, for training.
-During the classification stage, a given recording is similarly processed into short windows represented by feature vectors which are classified individually by the classifier.
-The individial class assignments are then combined through a majority vote which assigns a class to the whole recording.
-This last step is known as late integration and allows the classifier to work on individual, shorter-scale sound events instead of trying to classify the recording as a whole.
-By doing this, the classifier is able to identify and exploit the distinct events that characterize the scene instead of trying to model its average structure.
-This is in agreement with the psychoacoustic results related to the perception of auditory scenes presented above.
-
-%\ml{These two last paragraphs are a bit out of place here, since we don't do classification in the paper. Perhaps there is some way to cut them down a bit?}
-
-%\gl{Othe this r paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory}
-
-%\subsection{Datasets}
-
- % The latter is also considered in this paper as a validation of the proposed approach. \ja{What does this last sentence mean?}
-
-%In this paper, we shall demonstrate that the use of features based on the wavelet scattering transform together with object-based representation of the acoustic scene allows us to achieve state-of-the-art \ja{?} performance in both the ASSR and ASC paradigms.
+From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF models which treat each observation separately and do not capture their temporal structure.
+A method that takes this knowledge into account could therefore have potential for great impact in acoustic scene modeling, given a rich enough representation of these distinct events.
 

### 2018-02-17 paper/structureScene.tex
Edit scattering section

@@
 
-Scattering transforms are time-shift invariant representations of audio signals computed by applying auditory and modulation wavelet filter banks alternated with complex modulus nonlinearities.
-This section first explains the importance of local invariance to time-shifting and stability to time-warping in the representation of acoustic scenes, then describes how the scattering transform is designed to satisfy these properties while retaining high discriminative power.
+Local invariance to time-shifting and stability to time-warping are necessary when representing acoustic scenes for similarity measurement.
+The scattering transform is designed to satisfy these properties while retaining high discriminative power.
+It is computed by applying auditory and modulation wavelet filter banks alternated with complex modulus nonlinearities.
 
 \subsection{Invariance and stability in audio signals}
-The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
-Indeed, recordings may be shifted locally without affecting their similarity to other recordings.
-To discard this superfluous source of variability, signals are first mapped into a time-shifting invariant feature space.
-These features are then used to calculate similarities or to train a classifier.
-Since invariance is ensured by the features, it does not have to be learned during this last model construction step.
+The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval.
+Indeed, recordings may be shifted locally in time without affecting similarity to other recordings.
+To discard this superfluous source of variability, signals are first mapped into a time-shift invariant feature space.
+These features are then used to calculate similarities.
+Since the features ensure invariance, it does not have to be learned when constructing the desired similarity measure.
 
@@
 Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events.
-These make up an important part of intra-class variability among natural sounds, so it is important for representation to be robust to such transformations.
+These make up an important part of intra-class variability among natural sounds, so representations must be robust with respect to such transformations.
 
@@
 $\boldsymbol{x}(t)$ is obtained by convolution with all wavelet filters.
-Applying pointwise complex modulus the transform yields
+Applying a pointwise complex modulus the transform yields
 the wavelet scalogram
@@
 
-To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\phi(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
+To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\boldsymbol{\phi}(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
 \begin{equation}
-\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \phi(t),
+\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \boldsymbol{\phi}(t),
 \end{equation}
@@
 At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
-Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into the time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform and another complex modulus.
+Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into the time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered from $\boldsymbol{x_1}(t,\gamma_1)$ by a second wavelet transform and another complex modulus.
 
@@
 In the following, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}(t, \gamma_1, \gamma_2)$ to guarantee local invariance to time-shifting, which yields the second-order scattering coefficients
+Lastly, the low-pass filter $\boldsymbol{\phi}(t)$ is applied to $\boldsymbol{x_2}(t, \gamma_1, \gamma_2)$ to guarantee local invariance to time-shifting, which yields the second-order scattering coefficients
 \begin{equation}
 \mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
-(\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \phi)(t).
+(\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \boldsymbol{\phi})(t).
 \end{equation}
 
-The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
+The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ denotes either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
 While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.

### 2018-02-17 paper/structureScene.tex
Edit ASSR section

@@
 Given a set of $d$-dimensional feature vectors $X_u = \{x_1^u, \ldots, x_L^u\}$, extracted from the scene $s_u$, where $u=\lbrace 1,2,\ldots,U\rbrace$, we would like to partition $X_u$ into a set $C_u = \{c^u_1, \ldots, c^u_M\}$ of $M$ clusters.
-The partitioning is done by minimizing squared error between the empirical mean, or centroid, of each cluster and the vectors belonging to it.
-This is known as $k$-means clustering \cite{lloyd} and has found widespread use for a variety of applications.
+This partition is obtained by minimizing the variance of each cluster and known as a $k$-means clustering \cite{lloyd}.
 Each scene $s_u$ is then described by a set of clusters $C_u$.
-One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
+Note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
 Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroids.
@@
 \begin{itemize}
-\item Relevance-based Quantization closest similarity (\emph{RbQ-c}): the similarity between two scenes $s_u$ and $s_v$ is equal to the largest similarity between their centroids, that is
+\item Relevance-based Quantization closest similarity (\emph{RbQ-c}): the similarity between two scenes $s_u$ and $s_v$ is equal to the largest similarity between their centroids
 \begin{equation}
@@
 \end{equation}
-\item Relevance-based Quantization average similarity  (\emph{RbQ-a}): the similarity between two scenes $s_u$ and $s_v$ is equal to the average of their centroid similarities, that is
+\item Relevance-based Quantization average similarity  (\emph{RbQ-a}): the similarity between two scenes $s_u$ and $s_v$ is equal to the average of their centroid similarities
 \begin{equation}
@@
 The weight $w_m^u$ for the $m$th centroid $\mu_m^u$ is the number of frames belonging to a particular cluster.
-The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance ($\widehat{\EMD}$) introduced by \cite{pele2008linear}.
-The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied with the ``ground distance'' between the histogram bins.
-%In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the centroid distance.
-\ml{From what we have below, the distance is not actually the distance between centroids, though.
-It's given by the kernel $K_{mn}^{uv}$.}
-
-%\ml{Do we need the details that follow? Is it not enough to refer to \cite{pele2009fast}?}
-%To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
-Given two signatures $p_u$ and $p_v$, the $\widehat{\EMD}$ is computed by solving the following linear program:
-% \begin{equation}
-% \begin{split}
-% \widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
-% &+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
-% \end{split}
-% \end{equation}
-%
-% \begin{equation*}
-% \mathrm{s.t.} \quad f_{nm}\geq0 \quad \sum\limits_{m} f_{nm} \leq w_n^u \quad \sum\limits_{n} f_{nm} \leq w_m^v
-% \end{equation*}
-%
-% \begin{equation*}
-% \sum\limits_{n,m}f_{nm} = \min\left( \sum\limits_{n} w_n^u ,\sum\limits_{m} w_m^v \right)
-% \end{equation*}
-% %\ja{Can we provide some interpretation for the different terms and constraints here?}
-% where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin.
-We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$ which is computed from $K$:
-% \begin{equation*}
-% D_{mn}^{uv}=1-K_{mn}^{uv}.
-% \end{equation*}
-% \ml{Note that this is not actually a distance.}
-
-% To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ is computed: %\cite{chapelle1999support,jing2003support}
-% \begin{equation}
-% \label{eq:ks}
-% K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(p_u,p_v)}{A} \right) \\
-% \end{equation}
-% with $A$ a scaling parameter, set to the mean value of the $\widehat{\EMD}$ between all the scenes.
-The resulting kernel $K^s$ is known as an $\EMD$ kernel, and it should be noted that there is no guarantee that this kernel is positive definite.
-\ml{Is this kernel used anywhere?
-For similarity, do we not just use the EMD distance directly?}
+The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance $\widehat{\EMD}$ introduced by \cite{pele2008linear}.
+The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied by the a dissimilarity measure between the histogram bins.
+Here, that measure is given by $1-K_{mn}^{uv}$.
 

### 2018-02-17 paper/structureScene.tex
Edit experiments section

@@
 
-To evaluate the representations introduced in the previous section, we apply it to the acoustic scene similarity retrieval task on the DCASE 2013 dataset.
+To evaluate the representations introduced in the previous section, we apply it to the acoustic scene similarity retrieval task.
 Results demonstrate the improved performance of the relevance-based quantization of scattering coefficients compared to baseline methods using summary statistics of MFCCs.
-The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}} % \ja{What is ``the one''?}
-
-% , and $\alpha$ is a parameter which trades off between \ja{?}
+The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}}
 
@@
 
-%For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}. Larger datasets exists: the
-%Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
-The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} dataset.
-This dataset consists of two parts, namely a public and a private subset, each made up of $100$ acoustic scene recordings, sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
+The experiments in this paper are carried out on the DCASE 2013 dataset \cite{7100934}.
+This dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
 The dataset is evenly divided into $10$ acoustic scence classes.
@@
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-As a result, the DCASE 2013 dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
-%In addition, it is still a challenging dataset, with a state-of-the-art system based using handcrafted features as input to SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
-
-%Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
-On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
-%, but here we conduct our experiments on the public part of the dataset. \ja{Any more information on this dataset?}
-
-%In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
-\ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages.
-Moreover, since MFCCs are well-established features which perform well for these task, we shall include them baseline features in all experiments described in this paper.
+As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
 
@@
 
-Experiments are carried out using scattering coefficients as well as baseline mel-frequency cepstral coefficients (MFCCs) as features.
-For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$.
-In this case, $3$ seconds are discarded at the beginning and end of the scene to avoid boundary artifacts.
-Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}).
+We perform our experiments using both scattering coefficients and MFCCs.
+For the scattering transform, each $30$-second scene is described by $128$ vectors of dimension $1367$ computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$.
+Here we discard $3$ seconds from the beginning and end of the scene to avoid boundary artifacts.
+We also conduct experiments with and without logarithmic compression of the scattering coefficients (see Section \ref{sec:logcomp}).
 
@@
 The standard configuration of $39$ coefficients coupled with an average-energy measure performs best in preliminary tests, so we use this in the following.
-The coefficients are averaged using $250\,\mathrm{ms}$ long non-overlapping windows so that each window represents structures of similar scale to the scattering coefficients.
-
-%.
-The MFCC vectors within each texture window are then averaged in time.
-As the \emph{ob} approaches involve a clustering step, we need to ensure that both MFCCs and scattering coefficients have a comparable number of vectors per recording.
-Thus, $t$ is set to $250\,\mathrm{ms}$, resulting in $120$ MFCCs texture windows per scene.
-
-%\ja{We've tested the impact of the logarithm here, but not of the feature selection or standardization.
-Should we do this as well?
-For the feature selection, we did say earlier that we would test its impact.}
-
-%Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval.
-\ml{What does this mean?}
+We average the coefficients using $250\,\mathrm{ms}$ long non-overlapping windows so that each window represents structures of scales close to that of scattering coefficients.
 
@@
 The evaluation is performed on the private part of the DCASE 2013 dataset.
-The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
-These closest neighbors are determined by a particular distance, which is one of \emph{RbQ-c}, \emph{RbQ-a}, or \emph{RbQ-w}.
-\vl{Avoid passive voice.}
+As a metric, we use the precision at rank $k$ ($p@k$).
+This number is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
+We determine these neighbors using one of the proposed similarity measures \emph{RbQ-c}, \emph{RbQ-a}, or \emph{RbQ-w}.
+We compute $p@k$ for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
+Note that $p@1$ is equal to the classification accuracy obtained by the nearest-neighbor classifier in a leave-one-out cross-validation setting.
 
-The metric $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
-Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item.
-The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene.
-The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \textit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization (EM) algorithm.
-Gaussian models are compared using Monte-Carlo sampling approximation.
-In order to ensure convergence of the EM algorithm when considering the scattering features as input, the dimension of the feature vector is reduced from 1367 to 30 by projecting the features over the 30 principal components computed over the whole dataset.
-\vl{Too much passive voice, too many instances of "use", "used".}
+The \emph{RbQ} measures are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene.
+The distance on this average feature vector is then used to determine $p@k$.
+For the BOF approach of Aucouturier et al.
+\cite{aucouturier2007bag}, $10$-component GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird}.
+The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation.
+To ensure convergence of the EM algorithm for the scattering features, we reduce their dimension from $1367$ to $30$ by projecting the features onto the top $30$ principal components of the dataset.
 
-The scaling parameter $q$ of the RBF kernels (see Eq.
-\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.%, being the number of scenes for \emph{early} and the number of clusters for RbQ approaches.
-\ml{What does this mean?}
-The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and $200$ replicates. For each method, the numbers of clusters is set to $8$.
-
-%\ja{I'm not sure what this means. Do we fix $m$ and $N$ here for all experiments or do we optimize them over the test data each time? If so, this will bias the results, no?}
+The scaling parameter $q$ of the RBF kernels (see Eq. \ref{eq:kc}) is set to $10\%$ of the number of data points to cluster. For each method, the numbers of clusters is set to $8$.
 

### 2018-02-17 paper/structureScene.tex
Edit results section

@@
 
-%The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation.
-Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering and MFCCs features.
-
-This section presents evaluation results for the acoustic scene similarity retrieval task.
-The $p@k$ for different settings are shown in Figure \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics.
-This allows us to also evaluate the recall capabilities of the system under evaluation.
+Results for the acoustic scene similarity retrieval task demonstrate the superiority of the logarithmically compressed scattering features compared to MFCCs.
+Combining these with the \emph{RbQ} cluster model, we obtain significant improvements over traditional BOF and summary statistic measures.
 
@@
 
-As can be seen on Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when considering MFCCs as features.
-Its performance decreases compared to the early approach when considering the scattering features.
-Extensive analysis of those results is out of the scope of this paper, but we can assume that this decrease is to to the dimensionality reduction operated on the scattering features before GMM training in order to ensure the convergence of the EM algorithm.
+As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied to MFCCs.
+In contrast, its performance decreases compared to the early approach when applied to scattering features.
+Extensive analysis of those results is out of the scope of this paper, but this likely results from the dimension reduction applied prior to estimating the GMMs in the BOF model.
 
-The early approach being simpler in terms of implementation and runtime complexity, this method is retained as baseline for the remaining.
+The early approach being simpler in terms of implementation and runtime complexity, we retain this method as baseline for the remainder of the experiments.
 
@@
 \vl{The x-axis label is wrong. precision should be replaced by "rank $k$". The y-axis should be "precision ($p@k$)". It would be good to have the y-axis range from $0\%$ to $100\%$.
-The lines should be thicker and follow a color palette of complementary colors (blue vs. orange). The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC. }}
+The lines should be thicker and follow a color palette of complementary colors (blue vs.
+orange).
+The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC.
+} \ja{Should we not have $k$ stricly smaller than $10$ here?}}
 \label{fig:ASS_-1}
@@
 
-As can be seen on Figure \ref{fig:ASS_0}, the logarithmic compression is beneficial for precision performance irrespective of the similarity considered.
+Figure \ref{fig:ASS_0} shows how logarithmic compression is beneficial irrespective of the similarity model used.
 The gain is more pronounced when considering alternative similarities.
+\ja{What's ``alternative'' similarities?}
 
@@
 For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (\emph{RbQ}) to improve the similarity measures between the scenes.
-However, it is worth noticing that \emph{RbQ-a} performs worse or comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
+However, it is worth noting that \emph{RbQ-a} performs comparably to or worse than \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
 This result is in line with the findings of \cite{lagrange:hal-01082501}.
 To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects.
+
 Furthermore, it appears that \emph{RbQ-c} is better able to characterize the classes compared to \emph{RbQ-w}.
-This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution.
+Although not the only way of incorporating the number of frames associated to each centroid, the earth mover's distance is a rather natural of doing so.
+Its worse performance therefore suggests that including this information may not always be desirable.
 Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
 On the contrary, two similar environments may share a lot of similar sound sources with only a few sources discriminating between them.
-\ml{This is a bit of a stretch, since it may just be that we used the wrong distance to take the weighting into account.
-Perhaps the non-normalized EMD is not appropriate here?}
 
-\ja{needs to be discussed}
-
-Considering $p@5$ as our metric (as in \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), the use of the log-scattering transform instead of MFCCs increases the performance from $0.31$ to $0.49$ and the use of the relevance-based quantization approach using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
-\vl{Too many instances of ``use'', ``use'', ``using''.
-It is worth reminding what the baselines are.}
-
-%The same observations are made for the MFCCs for $k\leq5$.
-However for a rank $k$ greater than $5$, all the \emph{RbQ} approaches perform equivalently.
-As such, one can conclude that the use of MFCCs allows us to retrieve closely related scenes, but their lack of discriminative capabilities making the cluster selection/weighting step useless for not so closely related scenes of the same class.
-
-%This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.
-\ja{What is the ``at some point'' referring to?
-In other words, why do we expect this to happen for larger $k$ but not smaller?}
-
-%However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}.
-The latter even outperforms \emph{ob-a} for $k\geq5$.
-From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class.
-This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.
+With $p@5$ as our metric (cf.
+\cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), we see that replacing MFCCs by the logarithmically compressed scattering transform increases performance from $0.31$ to $0.49$.
+In addition, the relevance-based quantization using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
+\vl{It is worth reminding what the baselines are.}
 

### 2018-02-17 paper/structureScene.tex
Edit conclusion

@@
 
-This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales.
-Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature.
-In order to extend the selectivity potential of late integration using SVMs to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval.
+This paper presents a new approach for modeling acoustic scenes based on scattering transforms at small scales and cluster-based representations at large scales.
+Compared to traditional BOF and summary statistics models, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature.
+To adequately capture such distinct events, we develop a cluster-based model and validate it using experiments on acoustic scene similarity retrieval.
 For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
 These outcomes shall be studied further in future work by considering larger databases and emerging tasks in computational bioacoustics \cite{wimmer2013sampling}.
-\vl{The verb utilize should be avoided.}
 

### 2018-02-19 paper/structureScene.tex
add editors requirements

--- 2018-02-17 paper/structureScene.tex
@@
 
-\author{Vincent Lostanlen\thanks{This work is supported by the ERC InvariantClass grant 320959.}, Gr\'egoire Lafay, Joakim And\'en, and Mathieu Lagrange}
+\author{Vincent Lostanlen         \and
+		Gr\'egoire Lafay  \and Joakim And\'en  \and Mathieu Lagrange.}
+
+\institute{Vincent Lostanlen \at
+              \email{vincent.lostanlen@nyu.edu}
+\and
+			Gr\'egoire Lafay \at
+              \email{lafaygregoire@gmail.com}
+\and
+			Joakim And\'en \at
+              \email{janden@flatironinstitute.org}           \and
+			Mathieu Lagrange \at
+              \email{mathieu.lagrange@cnrs.fr}
+}
+
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
 
@@
 The emerging field of computational ecoacoustics aims at retrieving high-level information from acoustic scenes recorded by some network of sensors.
-These networks gather large amounts of data requiring analysis.
-To decide which parts to inspect manually, we need tools that automatically mine the data, identifying recurring patterns and isolated events.
-This requires a similarity measure for acoustic scenes that does not impose strong assumptions on the data.
+Large amounts of unlabeled data are gathered using such networks that needs to be analyzed.
+In order to identify which parts require human inspection, one needs tools that can mine data and identify structures such as recurring patterns or isolated events.
+To do so, one needs to be able to measure the level of similarity among the data without strong assumptions.
 
-The state of the art in audio similarity measurement is the ``bag-of-frames'' approach, which models a recording using summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
-They successfully characterize static scenes with little variability in auditory content, but cannot accurately capture scenes with a few salient events superimposed over static background.
-To overcome this issue, we propose a two-scale representation which describes a recording using clusters of scattering coefficients.
-The scattering coefficients capture short-scale structure, while the cluster model captures longer time scales, allowing for more accurate characterization of sparse events.
-Evaluation within the acoustic scene similarity framework demonstrates the superiority of the proposed approach.
+State of the art approaches for computing such a similarity is the ``bag-of-frames'' approach, i.e., modeling the audio signal using the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
+Although they successfully characterize static scenes with little variability in auditory content, they are unable to accurately discriminate scenes with few salient events superimposed over a textured background.
+To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure that better handle the required modeling of sparse events.
+Experiments performed using the acoustic scene similarity framework demonstrates the superiority of the proposed approach.
 \keywords{
@@
 In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
-Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for to signal processing community.
+Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.
 
-These questions are explored in an emerging field of computational bioacoustics, known as ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
+This work falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
+%Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
+%Consequently, few well-designed datasets are readily available for evaluation purposes.
 An important problem in this field is that manually analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling}.
-Some sort of pre-screening is therefore required to reduce the need for human expert listening and annotation.
+Some sort of pre-screening is therefore needed to reduce the need for human expert listening and annotation.
 
-The most straightforward approach is to specify a closed set of sound classes, such as sounds of animals expected to live near the acoustic sensors.
-We then train models for these classes which are used to automatically annotate our recordings \cite{}.
-\ja{What's the reference here?} A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class.
-This allows the scientist to drastically reduce the amount of information requiring manual processing.
+The most straightforward approach is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}.
+A given time interval, for example a single day, is then represented by the number of events within the predefined typology that have been detected in that period of time.
+This approach allows the scientist to drastically reduce the amount of information that needs to be processed.
 However, this approach has two drawbacks.
-First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be trusted.
-Second, and more importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.
+The first is the reliance to trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be always be trusted.
+More importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where the quantities of interest have yet to be defined.
 
-To identify which parts need human inspection, we need tools that can detect both recurring patterns and sparsely distributed events.
-Identifying recurring patterns allows the user to focus on certain time points for manual annotation, while detection of more rare structures enables discovery of unforeseen phenomena.
+In order to identify which parts needs human inspection, one needs tools that can mine data and annotate to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors.
+The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.
 
-With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset.
-This requires us to represent an audio recording in a way that captures its distinctive qualities.
-This has previously been attempted using bag-of-frames approaches \cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
-Unfortunately, this only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterized by a few distinct sound events.
-Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neuroscience \cite{nelken2004processing} suggest that this more closely matches human acoustic perception.
-We believe that the failure to model such distinct events is one of the reasons why this approach is insufficient for moderate-sized evaluation datasets \cite{lagrange:hal-01082501}.
+With this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset.
+This requires the audio fragments to be represented in some way that captures their distinctive qualities.
+This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics.
+Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented with scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others.
+Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely human acoustic perception of acoustic scenes.
+We believe that the failure to model such distinct events is one of the reason why this approach is found to be insufficient when considering evaluation datasets of typical sizes \cite{lagrange:hal-01082501}.
 
-Solving the acoustic similarity retrieval first requires the ability to capture meaningful signal structure at small time scales.
-This is often achieved using mel-frequency cepstral coefficients (MFCCs).
-Originally developed for speech processing \cite{davis-mermelstein}, MFCCs have recently found wider use in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
-A richer representation, the scattering transform, has enjoyed significant success in various audio \cite{Anden2014} and biomedical \cite{chudacek} signal classification tasks.
-Its structure is that of a convolutional neural network \cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}, but with fixed filters.
-Specifically, it alternates convolutions with wavelet filters and pointwise nonlinearities to ensure time-shift invariance and time-warping stability \cite{Mallat2012}.
+%To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes.
+This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}.
+In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.
 
-For our task, one advantage of the scattering transform is not requiring a training step, allowing for a wider range of applications compared to learned features.
-Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unserpervised setting.
+Addressing this task first requires the ability to capture meaningful aspects within the signal at small time-scales.
+In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
+A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
+A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014,chudacek}.
+This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}.
 
-In this work, we propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an cluster model.
-This unsupervised model quantizes the scattering coefficients into a given number of clusters.
-These clusters to define a set of distances for acoustic similarity retrieval.
-Evaluating this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
+For the task at at hand, scattering features have the advantage of not requiring training and thus potentially lead to wider range of application than trained features.
+Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined, thus leading to an unserpervised setting.
+
+%The next question is the choice of a classifier.
+Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM) \cite{cristianini2000introduction}, to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision.
+This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes.
+The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it almost completely.
+Combining scattering transforms for low-level representation with late-integrated SVMs for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.
+
+%we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
+We also propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters.
+We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval.
+Testing this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
 
@@
 Section \ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task.
-Results are reported in Section \ref{sec:results}.
+Results are subsequently reported in Section \ref{sec:results}.
 
@@
 
-Computational bioacoustics refers to the numerical investigation of sound production, dispersion and reception in animals, including humans.
-A recent paradigm processes the audio stream in a holistic way over large time scales, without assuming that a single species is present throughout the recording.
-Automated systems within this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010} and migration patterns \cite{Obrist2010}.
-They also mark time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.
+%One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify species properties of vocalizations, for example to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
 
-A closely related field is that of urban sound environment analysis.
-In this context, a popular approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al. \cite{aucouturier2007bag}.
-It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.
-Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
-This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background \cite{nelken2013}.
-Such events are not well-characterized by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
-The latter approach should therefore prove more fruitful in measuring auditory scene similarity.
+Computational bioacoustics usually refers to the numerical investigation of sound production, dispersion and reception in animals (including humans).
+A recent paradigm  processes the audio stream in a holistic way over large time-scales, without assuming that a single species is present throughout the recording.
+Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as marking time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.
+
+%The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
+In the closely related field of urban sound environment analysis, a popular approach is the bag of frames (BOF) introduced in this context by Aucouturier et al.
+\cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
+Recently, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
+This contrasts with the typical morphology of acoustic scenes, which we believe is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events are superimposed upon a stationary acoustic background.
+As those distinct and sparse events are often highly relevant for the task at hand, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on summary statistics of short-term features.
 
 This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}.
-Studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Instead, cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} seem to rely upon higher-level cognitive attributes.
-These typically include the identities of the sound sources which consitute the scene.
+For example, studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene.
 It has been shown that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes.
-For example, in urban areas the presence of birds is likely to be heard in parks and are therefore strong pleasantness indicators.
+For example, the presence of birds is very likely to be heard in parks in urban areas and so are strong pleasantness indicators.
 Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments \cite{ricciardi2015sound,lavandier2006contribution}.
@@
 
-From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF models which treat each observation separately and do not capture their temporal structure.
-A method that takes this knowledge into account could therefore have potential for great impact in acoustic scene modeling, given a rich enough representation of these distinct events.
+From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation the same and do not incorporate temporal structure between features.
+A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes, provided that the representation of the representation of those distinct events is rich enough.
+
+%Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of acoustic scene similarity retrieval, much of the work dedicated to acoustic scene modeling is applied to the classification task.
+%An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.
+
+%The methods that performed the best all follow a similar structure, where low-level feature representations, such as MFCCs, are calculated over short windows and fed separately to a classifier, such as a support vector machine (SVM) \cite{cortes-vapnik,cristianini2000introduction}, for training.
+During the classification stage, a given recording is similarly processed into short windows represented by feature vectors which are classified individually by the classifier.
+The individial class assignments are then combined through a majority vote which assigns a class to the whole recording.
+This last step is known as late integration and allows the classifier to work on individual, shorter-scale sound events instead of trying to classify the recording as a whole.
+By doing this, the classifier is able to identify and exploit the distinct events that characterize the scene instead of trying to model its average structure.
+This is in agreement with the psychoacoustic results related to the perception of auditory scenes presented above.
+
+%\ml{These two last paragraphs are a bit out of place here, since we don't do classification in the paper. Perhaps there is some way to cut them down a bit?}
+
+%\gl{Othe this r paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory}
+
+%\subsection{Datasets}
+
+ % The latter is also considered in this paper as a validation of the proposed approach. \ja{What does this last sentence mean?}
+
+%In this paper, we shall demonstrate that the use of features based on the wavelet scattering transform together with object-based representation of the acoustic scene allows us to achieve state-of-the-art \ja{?} performance in both the ASSR and ASC paradigms.
 
@@
 
-Local invariance to time-shifting and stability to time-warping are necessary when representing acoustic scenes for similarity measurement.
-The scattering transform is designed to satisfy these properties while retaining high discriminative power.
-It is computed by applying auditory and modulation wavelet filter banks alternated with complex modulus nonlinearities.
+Scattering transforms are time-shift invariant representations of audio signals computed by applying auditory and modulation wavelet filter banks alternated with complex modulus nonlinearities.
+This section first explains the importance of local invariance to time-shifting and stability to time-warping in the representation of acoustic scenes, then describes how the scattering transform is designed to satisfy these properties while retaining high discriminative power.
 
 \subsection{Invariance and stability in audio signals}
-The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval.
-Indeed, recordings may be shifted locally in time without affecting similarity to other recordings.
-To discard this superfluous source of variability, signals are first mapped into a time-shift invariant feature space.
-These features are then used to calculate similarities.
-Since the features ensure invariance, it does not have to be learned when constructing the desired similarity measure.
+The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
+Indeed, recordings may be shifted locally without affecting their similarity to other recordings.
+To discard this superfluous source of variability, signals are first mapped into a time-shifting invariant feature space.
+These features are then used to calculate similarities or to train a classifier.
+Since invariance is ensured by the features, it does not have to be learned during this last model construction step.
 
@@
 Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events.
-These make up an important part of intra-class variability among natural sounds, so representations must be robust with respect to such transformations.
+These make up an important part of intra-class variability among natural sounds, so it is important for representation to be robust to such transformations.
 
@@
 $\boldsymbol{x}(t)$ is obtained by convolution with all wavelet filters.
-Applying a pointwise complex modulus the transform yields
+Applying pointwise complex modulus the transform yields
 the wavelet scalogram
@@
 
-To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\boldsymbol{\phi}(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
+To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\phi(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
 \begin{equation}
-\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \boldsymbol{\phi}(t),
+\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \phi(t),
 \end{equation}
@@
 At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
-Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into the time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered from $\boldsymbol{x_1}(t,\gamma_1)$ by a second wavelet transform and another complex modulus.
+Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into the time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform and another complex modulus.
 
@@
 In the following, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Lastly, the low-pass filter $\boldsymbol{\phi}(t)$ is applied to $\boldsymbol{x_2}(t, \gamma_1, \gamma_2)$ to guarantee local invariance to time-shifting, which yields the second-order scattering coefficients
+Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}(t, \gamma_1, \gamma_2)$ to guarantee local invariance to time-shifting, which yields the second-order scattering coefficients
 \begin{equation}
 \mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
-(\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \boldsymbol{\phi})(t).
+(\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \phi)(t).
 \end{equation}
 
-The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ denotes either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
+The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
 While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
@@
 Given a set of $d$-dimensional feature vectors $X_u = \{x_1^u, \ldots, x_L^u\}$, extracted from the scene $s_u$, where $u=\lbrace 1,2,\ldots,U\rbrace$, we would like to partition $X_u$ into a set $C_u = \{c^u_1, \ldots, c^u_M\}$ of $M$ clusters.
-This partition is obtained by minimizing the variance of each cluster and known as a $k$-means clustering \cite{lloyd}.
+The partitioning is done by minimizing squared error between the empirical mean, or centroid, of each cluster and the vectors belonging to it.
+This is known as $k$-means clustering \cite{lloyd} and has found widespread use for a variety of applications.
 Each scene $s_u$ is then described by a set of clusters $C_u$.
-Note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
+One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
 Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroids.
@@
 \begin{itemize}
-\item Relevance-based Quantization closest similarity (\emph{RbQ-c}): the similarity between two scenes $s_u$ and $s_v$ is equal to the largest similarity between their centroids
+\item Relevance-based Quantization closest similarity (\emph{RbQ-c}): the similarity between two scenes $s_u$ and $s_v$ is equal to the largest similarity between their centroids, that is
 \begin{equation}
@@
 \end{equation}
-\item Relevance-based Quantization average similarity  (\emph{RbQ-a}): the similarity between two scenes $s_u$ and $s_v$ is equal to the average of their centroid similarities
+\item Relevance-based Quantization average similarity  (\emph{RbQ-a}): the similarity between two scenes $s_u$ and $s_v$ is equal to the average of their centroid similarities, that is
 \begin{equation}
@@
 The weight $w_m^u$ for the $m$th centroid $\mu_m^u$ is the number of frames belonging to a particular cluster.
-The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance $\widehat{\EMD}$ introduced by \cite{pele2008linear}.
-The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied by the a dissimilarity measure between the histogram bins.
-Here, that measure is given by $1-K_{mn}^{uv}$.
+The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance ($\widehat{\EMD}$) introduced by \cite{pele2008linear}.
+The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied with the ``ground distance'' between the histogram bins.
+%In our case, the histogram counts are the clusters weights $w_m^u$ defined over the bins formed by the centroids $\mu_m^u$, so the ground distance is the centroid distance.
+\ml{From what we have below, the distance is not actually the distance between centroids, though.
+It's given by the kernel $K_{mn}^{uv}$.}
+
+%\ml{Do we need the details that follow? Is it not enough to refer to \cite{pele2009fast}?}
+%To compute the $\widehat{\EMD}$, we use the implementation proposed in \cite{pele2009fast}.
+Given two signatures $p_u$ and $p_v$, the $\widehat{\EMD}$ is computed by solving the following linear program:
+% \begin{equation}
+% \begin{split}
+% \widehat{\EMD}(p_u,p_v) &=\left( \min\limits_{\lbrace f_{nm}\rbrace} \sum\limits_{n,m} f_{nm}D_{nm}^{uv} \right) \\
+% &+ \left|\sum\limits_{n} w_n^u - \sum\limits_{m} w_m^v  \right| \max\limits_{n,m}\lbrace  D_{nm}^{uv}\rbrace.
+% \end{split}
+% \end{equation}
+%
+% \begin{equation*}
+% \mathrm{s.t.} \quad f_{nm}\geq0 \quad \sum\limits_{m} f_{nm} \leq w_n^u \quad \sum\limits_{n} f_{nm} \leq w_m^v
+% \end{equation*}
+%
+% \begin{equation*}
+% \sum\limits_{n,m}f_{nm} = \min\left( \sum\limits_{n} w_n^u ,\sum\limits_{m} w_m^v \right)
+% \end{equation*}
+% %\ja{Can we provide some interpretation for the different terms and constraints here?}
+% where $\lbrace f_{nm} \rbrace$ is the flow between the cluster weights $w_n^u$ and $w_m^v$, that is, the amount transported from the $n^{\textrm{th}}$ bin to ``supply the demand'' of the $m^\textrm{th}$ bin.
+We denote by $D^{uv}$ the ground distance, a matrix containing the pairwise distances between the centroids sets $\mu^u$ and $\mu^v$ which is computed from $K$:
+% \begin{equation*}
+% D_{mn}^{uv}=1-K_{mn}^{uv}.
+% \end{equation*}
+% \ml{Note that this is not actually a distance.}
+
+% To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernel $K^s$ is computed: %\cite{chapelle1999support,jing2003support}
+% \begin{equation}
+% \label{eq:ks}
+% K_{uv}^s = \exp\left( - \dfrac{\widehat{\EMD}(p_u,p_v)}{A} \right) \\
+% \end{equation}
+% with $A$ a scaling parameter, set to the mean value of the $\widehat{\EMD}$ between all the scenes.
+The resulting kernel $K^s$ is known as an $\EMD$ kernel, and it should be noted that there is no guarantee that this kernel is positive definite.
+\ml{Is this kernel used anywhere?
+For similarity, do we not just use the EMD distance directly?}
 
@@
 
-To evaluate the representations introduced in the previous section, we apply it to the acoustic scene similarity retrieval task.
+To evaluate the representations introduced in the previous section, we apply it to the acoustic scene similarity retrieval task on the DCASE 2013 dataset.
 Results demonstrate the improved performance of the relevance-based quantization of scattering coefficients compared to baseline methods using summary statistics of MFCCs.
-The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}}
+The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}} % \ja{What is ``the one''?}
+
+% , and $\alpha$ is a parameter which trades off between \ja{?}
 
@@
 
-The experiments in this paper are carried out on the DCASE 2013 dataset \cite{7100934}.
-This dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
+%For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}. Larger datasets exists: the
+%Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
+The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} dataset.
+This dataset consists of two parts, namely a public and a private subset, each made up of $100$ acoustic scene recordings, sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
 The dataset is evenly divided into $10$ acoustic scence classes.
@@
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+As a result, the DCASE 2013 dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+%In addition, it is still a challenging dataset, with a state-of-the-art system based using handcrafted features as input to SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.
+
+%Recent advances using hierarchical classifier architectures obtain between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.
+On the Rouen dataset, the state-of-the-art is currently $95\%$ \cite{bisot2016acoustic}.
+%, but here we conduct our experiments on the public part of the dataset. \ja{Any more information on this dataset?}
+
+%In this paper, we are interested in evaluating different algorithmic designs leading to different scene models, we therefore resort to straightforward ranking performance measures for the ASSR task.
+\ja{It's not clear to me how the first part implies the second part here} Similarly, we have picked SVMs as our classifier for the ASC task, since these are a standard tool in the community and allows for us to focus our analysis on the earlier processing stages.
+Moreover, since MFCCs are well-established features which perform well for these task, we shall include them baseline features in all experiments described in this paper.
 
@@
 
-We perform our experiments using both scattering coefficients and MFCCs.
-For the scattering transform, each $30$-second scene is described by $128$ vectors of dimension $1367$ computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$.
-Here we discard $3$ seconds from the beginning and end of the scene to avoid boundary artifacts.
-We also conduct experiments with and without logarithmic compression of the scattering coefficients (see Section \ref{sec:logcomp}).
+Experiments are carried out using scattering coefficients as well as baseline mel-frequency cepstral coefficients (MFCCs) as features.
+For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$.
+In this case, $3$ seconds are discarded at the beginning and end of the scene to avoid boundary artifacts.
+Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}).
 
@@
 The standard configuration of $39$ coefficients coupled with an average-energy measure performs best in preliminary tests, so we use this in the following.
-We average the coefficients using $250\,\mathrm{ms}$ long non-overlapping windows so that each window represents structures of scales close to that of scattering coefficients.
+The coefficients are averaged using $250\,\mathrm{ms}$ long non-overlapping windows so that each window represents structures of similar scale to the scattering coefficients.
+
+%.
+The MFCC vectors within each texture window are then averaged in time.
+As the \emph{ob} approaches involve a clustering step, we need to ensure that both MFCCs and scattering coefficients have a comparable number of vectors per recording.
+Thus, $t$ is set to $250\,\mathrm{ms}$, resulting in $120$ MFCCs texture windows per scene.
+
+%\ja{We've tested the impact of the logarithm here, but not of the feature selection or standardization.
+Should we do this as well?
+For the feature selection, we did say earlier that we would test its impact.}
+
+%Having validated the underlying assumptions for the cluster-based auditory scene representation, we evaluate its performance for the task of acoustic scene similarity retrieval.
+\ml{What does this mean?}
 
@@
 The evaluation is performed on the private part of the DCASE 2013 dataset.
-As a metric, we use the precision at rank $k$ ($p@k$).
-This number is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
-We determine these neighbors using one of the proposed similarity measures \emph{RbQ-c}, \emph{RbQ-a}, or \emph{RbQ-w}.
-We compute $p@k$ for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
-Note that $p@1$ is equal to the classification accuracy obtained by the nearest-neighbor classifier in a leave-one-out cross-validation setting.
+The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
+These closest neighbors are determined by a particular distance, which is one of \emph{RbQ-c}, \emph{RbQ-a}, or \emph{RbQ-w}.
 
-The \emph{RbQ} measures are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene.
-The distance on this average feature vector is then used to determine $p@k$.
-For the BOF approach of Aucouturier et al.
-\cite{aucouturier2007bag}, $10$-component GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird}.
-The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation.
-To ensure convergence of the EM algorithm for the scattering features, we reduce their dimension from $1367$ to $30$ by projecting the features onto the top $30$ principal components of the dataset.
+The metric $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
+Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooses the label of the closest neighbor for a given item.
+The \emph{RbQ} approaches are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene.
+The parameters of the BOF approach are the one used in \cite{aucouturier2007bag}, \textit{i.e.} 10 Gaussians per scene estimated using the Expectation Maximization (EM) algorithm.
+Gaussian models are compared using Monte-Carlo sampling approaximation.
+In order to ensure convergence of the EM algorithm when considering the scattering features as input, the dimension of the feature vector is reduced from 1367 to 30 by projecting the features over the 30 principal components computed over the whole dataset.
 
-The scaling parameter $q$ of the RBF kernels (see Eq. \ref{eq:kc}) is set to $10\%$ of the number of data points to cluster. For each method, the numbers of clusters is set to $8$.
+The scaling parameter $q$ of the RBF kernels (see Eq.
+\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.%, being the number of scenes for \emph{early} and the number of clusters for RbQ approaches.
+\ml{What does this mean?}
+The clustering for the \emph{RbQ} approaches is done using plain $k$-means with random initialization and $200$ replicates. For each method, the numbers of clusters is set to $8$.
+
+%\ja{I'm not sure what this means. Do we fix $m$ and $N$ here for all experiments or do we optimize them over the test data each time? If so, this will bias the results, no?}
 
@@
 \includegraphics[width=\columnwidth]{figures/baselines}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
-\vl{The x-axis label is wrong. precision should be replaced by "rank $k$". The y-axis should be "precision ($p@k$)". It would be good to have the y-axis range from $0\%$ to $100\%$.
-The lines should be thicker and follow a color palette of complementary colors (blue vs. orange). The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC. }}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
 \label{fig:ASS_-1}
@@
 \ml{This is a bit of a stretch, since it may just be that we used the wrong distance to take the weighting into account.
-Perhaps the non-normalized EMD is not appropriate here?}
+Perhaps the non-normalize d EMD is not appropriate here?}
 
@@
 Considering $p@5$ as our metric (as in \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), the use of the log-scattering transform instead of MFCCs increases the performance from $0.31$ to $0.49$ and the use of the relevance-based quantization approach using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
-\vl{Too many instances of ``use'', ``use'', ``using''.
-It is worth reminding what the baselines are.}
 
@@
 These outcomes shall be studied further in future work by considering larger databases and emerging tasks in computational bioacoustics \cite{wimmer2013sampling}.
-\vl{The verb utilize should be avoided.}
+
+\section*{Availability of data and materials}
+The dataset supporting the conclusions of this article is available in the dcase2013 repository,  \sloppy \url{http://c4dm.eecs.qmul.ac.uk/sceneseventschallenge/description.html}.
+The software supporting the conclusions of this article is available in
+\begin{itemize}
+\item Project name: article2017EstimationAmbiance
+\item \sloppy Project home page: \url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}
+\item Programming language: Matlab
+\item License: GNU GPL
+\item Any restrictions to use by non-academics: license needed
+\end{itemize}
+
+\section*{Competing interests}
+The authors declare that they have no competing interests.
+
+\section*{Funding}
+This study is co-funded by the ANR under project reference ANR-16-CE22-0012.
+
+\section*{Authors' contributions}
+GL and VL carried out the numerical experiments and drafted the manuscript. VL, GL, JA and ML participated in the design of the study and helped to draft the manuscript.
+All authors read and approved the final manuscript.
 

### 2018-02-19 experiments/unsupervised/taun3clustering.m

@@ -11,7 +11,7 @@ function [config, store, obs] = taun3clustering(config, setting, data)
 % Date: 17-Dec-2016
 
 % Set behavior for debug mode
-if nargin==0, unsupervised('do', 3, 'mask', {2 0 2 3 }); return; else store=[]; obs=[]; end
+if nargin==0, unsupervised('do', 3, 'mask', {2 0 2 1 0 1 1 0 2}); return; else store=[]; obs=[]; end
 
 %% seed
 
@@ -66,6 +66,12 @@ switch setting.integration
         end
         
     case 'early'
+         if setting.pca
+           % perform pca
+          coeff = pca(data.features');
+          data.features = data.features'*coeff;
+          data.features = data.features(:, 1:30)';
+        end
         
         store.centroid=cell2mat(arrayfun(@(x) mean(data.features(:,data.indSample==x),2),unique(data.indSample),'UniformOutput',false));
         store.indSample=unique(data.indSample);

### 2018-02-19 experiments/unsupervised/taunReport.m

@@ -8,6 +8,9 @@ function config = taunReport(config)
 
 if nargin==0, unsupervised('report', 'r', 'reportName', 'ml'); return; end
 
+config = expExpose(config, 'p', 'mask', {2 0 2 [1 3] 0 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1);
+return
+
 config = expExpose(config, 'p', 'mask', {2 0 2 [1 3] 0 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1, 'save', 'baselines');
 config = expExpose(config, 'p', 'mask', {2 0 2 [1 2] 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1, 'save', 'overall');
 config = expExpose(config, 'p', 'mask', {2 2 0 [1 2] 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1, 'save', 'log');

### 2018-02-19 paper/structureScene.tex

--- 2018-02-17 paper/structureScene.tex
@@
 
+\section*{Availability of data and materials}
+The dataset supporting the conclusions of this article is available in the dcase2013 repository,  \sloppy \url{http://c4dm.eecs.qmul.ac.uk/sceneseventschallenge/description.html}.
+The software supporting the conclusions of this article is available in
+\begin{itemize}
+\item Project name: article2017EstimationAmbiance
+\item \sloppy Project home page: \url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}
+\item Programming language: Matlab
+\item License: GNU GPL
+\item Any restrictions to use by non-academics: license needed
+\end{itemize}
+
+\section*{Competing interests}
+The authors declare that they have no competing interests.
+
+\section*{Funding}
+This study is co-funded by the ANR under project reference ANR-16-CE22-0012.
+
+\section*{Authors' contributions}
+GL and VL carried out the numerical experiments and drafted the manuscript. VL, GL, JA and ML participated in the design of the study and helped to draft the manuscript.
+All authors read and approved the final manuscript.
+
 %\bibliographystyle{spmpsci}

### 2018-02-21 paper/structureScene.tex
restrucutre experiments
new baseline figure

@@
 For the BOF approach of Aucouturier et al.
-\cite{aucouturier2007bag}, $10$-component GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird}.
+\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird}.
 The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation.
 To ensure convergence of the EM algorithm for the scattering features, we reduce their dimension from $1367$ to $30$ by projecting the features onto the top $30$ principal components of the dataset.
+the number of Gaussians is optimized for each type of features by grid search in the range $[2, 20]$.
+Best $p@5$ is reached with 8 and 4  Gaussians, respectively for MFCCs and scattering features.
+Recommended number of Gaussians for MFCCs given in \cite{aucouturier2007bag} is 10.
 
@@
 He received the B.S.
-degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, France, and the B.S.
+degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, FranArthritis National Research Foundatioce, and the B.S.
 degree in Musicology from the Sorbonne University, Paris, France, in 2011.

### 2018-02-21 experiments/unsupervised/taun3clustering.m

@@ -11,7 +11,7 @@ function [config, store, obs] = taun3clustering(config, setting, data)
 % Date: 17-Dec-2016
 
 % Set behavior for debug mode
-if nargin==0, unsupervised('do', 3, 'mask', {2 0 2 1 0 1 1 0 2}); return; else store=[]; obs=[]; end
+if nargin==0, unsupervised('do', 3, 'mask', {2 0 2 3 0 1 1 0 2}); return; else store=[]; obs=[]; end
 
 %% seed
 
@@ -78,8 +78,10 @@ switch setting.integration
         store.centroidClass=store.class;
         case 'gmm'
         expRandomSeed();
+        setting.nbg=8;
         if strcmp(setting.features(1:4), 'scat') 
            % perform pca
+           setting.nbg=4;
           coeff = pca(data.features');
           data.features = data.features'*coeff;
           data.features = data.features(:, 1:30)';
@@ -92,7 +94,7 @@ switch setting.integration
                 if strcmp(setting.dataset, '2016') && i==33
                     dn = dn + randn(size(dn))*1;
                 end
-                gmmi = createGmm(dn, 10);
+                gmmi = createGmm(dn, setting.nbg);
                 if ~(any(isnan(gmmi.priors)) && any(isnan(gmmi.covars(:))))
                     break;
                 end
@@ -101,7 +103,7 @@ switch setting.integration
                 fprintf(2, 'unable to learn gmm for sample %d, performing kmeans\n', i);
                 
                 nbcoef=size(dn,2);
-                gmmi = gmm(nbcoef, 10, 'diag');
+                gmmi = gmm(nbcoef, setting.nbg, 'diag');
                 % options = foptions;
                 options(14) = 200;	% Just use 20 iterations of k-means in initialisation
                 options(1)  = -1;

### 2018-02-21 experiments/unsupervised/taunReport.m

@@ -8,8 +8,6 @@ function config = taunReport(config)
 
 if nargin==0, unsupervised('report', 'r', 'reportName', 'ml'); return; end
 
-config = expExpose(config, 'p', 'mask', {2 0 2 [1 3] 0 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1);
-return
 
 config = expExpose(config, 'p', 'mask', {2 0 2 [1 3] 0 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1, 'save', 'baselines');
 config = expExpose(config, 'p', 'mask', {2 0 2 [1 2] 1 1}, 'expand', 'precision', 'obs', 'p', 'percent', 1, 'save', 'overall');

### 2018-02-21 paper/structureScene.tex
TODO added

@@
 
-As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied to MFCCs.
-In contrast, its performance decreases compared to the early approach when applied to scattering features.
-Extensive analysis of those results is out of the scope of this paper, but this likely results from the dimension reduction applied prior to estimating the GMMs in the BOF model.
+As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied to the two types of features.
+
+TODO ML comment new figure
 

### 2018-02-21 paper/structureScene.tex
update log compression discussion

@@
 The most straightforward approach is to specify a closed set of sound classes, such as sounds of animals expected to live near the acoustic sensors.
-We then train models for these classes which are used to automatically annotate our recordings \cite{}.
-\ja{What's the reference here?} A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class.
+Computational models are trainedthen  for these classes which are used to automatically annotate recordings \cite{quteprints96243}.
+A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class.
 This allows the scientist to drastically reduce the amount of information requiring manual processing.
@@
 For the BOF approach of Aucouturier et al.
-\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird}.
+\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird, moon1996expectation}.
 The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation.
@@
 
-As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied to the two types of features.
-
-TODO ML comment new figure
+As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied both to the MFCCs and the scattering features.
 
@@
 The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC.
-} \ja{Should we not have $k$ stricly smaller than $10$ here?}}
+} \ja{Should we not have $k$ stricly smaller than $10$ here?} \ml{very true, we should stop at k=9 for all figures.
+We can see it reaches a plateau.
+mfcc, early pca should be discarded.}}
 \label{fig:ASS_-1}
@@
 
-Figure \ref{fig:ASS_0} shows how logarithmic compression is beneficial irrespective of the similarity model used.
-The gain is more pronounced when considering alternative similarities.
+Figure \ref{fig:ASS_0} shows that logarithmic compression of the scattering features is beneficial.
+For clarity sake, data is shown  for the \emph{early} approach only, but a equivalent gain is achieved for the relevance-based quantization approaches.
+%The gain is more pronounced when considering alternative similarities.
 \ja{What's ``alternative'' similarities?}
@@
 \includegraphics[width=\columnwidth]{figures/log}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.\ml{show only early}}
 \label{fig:ASS_0}
@@
 \includegraphics[width=\columnwidth]{figures/overall}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.\ml{keep only early approaches, and adapt the text refering to it then.}}
 \label{fig:ASS_1}

### 2018-02-22 paper/lostanlenRelevanceScattering.tex
imporved figures

--- 2018-02-21 paper/structureScene.tex
@@
 \includegraphics[width=\columnwidth]{figures/baselines}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
-\vl{The x-axis label is wrong. precision should be replaced by "rank $k$". The y-axis should be "precision ($p@k$)". It would be good to have the y-axis range from $0\%$ to $100\%$.
-The lines should be thicker and follow a color palette of complementary colors (blue vs.
-orange).
-The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC.
-} \ja{Should we not have $k$ stricly smaller than $10$ here?} \ml{very true, we should stop at k=9 for all figures.
-We can see it reaches a plateau.
-mfcc, early pca should be discarded.}}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
 \label{fig:ASS_-1}
@@
 \includegraphics[width=\columnwidth]{figures/log}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.\ml{show only early}}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
 \label{fig:ASS_0}
@@
 \includegraphics[width=\columnwidth]{figures/overall}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.\ml{keep only early approaches, and adapt the text refering to it then.}}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression.}
 \label{fig:ASS_1}

### 2018-02-22 paper/figures/mat2csv.m

@@ -1,10 +1,22 @@
 
-files = dir('./*.mat')
+files = dir('./*.mat');
 
 for k=1:length(files)
     clear data
-   load(files(k).name) 
-   if (exist('data', 'var'))
-     csvwrite([files(k).name(1:end-3), 'csv'], data.meanData); 
-   end
+    load(files(k).name)
+    if (exist('data', 'var'))
+        csvwrite([files(k).name(1:end-3), 'csv'], data.meanData);
+    end
+    
+    h=openfig([files(k).name(1:end-3) 'fig']);
+    set(h, 'visible', 'on')
+    ylabel('p@k (%)');
+    legend('-DynamicLegend');
+    lines = findobj(gcf,'Type','Line');
+    for i = 1:numel(lines)
+        lines(i).LineWidth = 2.0;
+    end
+    axis([1 9 10 100]);
+    disp('edit');
+    saveas(gcf, [files(k).name(1:end-3) 'png'])
 end

### 2018-02-22 paper/lostanlenRelevanceScattering.tex
Spell check & remove American spelling

@@
 The state of the art in audio similarity measurement is the ``bag-of-frames'' approach, which models a recording using summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
-They successfully characterize static scenes with little variability in auditory content, but cannot accurately capture scenes with a few salient events superimposed over static background.
+They successfully characterise static scenes with little variability in auditory content, but cannot accurately capture scenes with a few salient events superimposed over static background.
 To overcome this issue, we propose a two-scale representation which describes a recording using clusters of scattering coefficients.
@@
 The most straightforward approach is to specify a closed set of sound classes, such as sounds of animals expected to live near the acoustic sensors.
-Computational models are trainedthen  for these classes which are used to automatically annotate recordings \cite{quteprints96243}.
+Computational models are trained then  for these classes which are used to automatically annotate recordings \cite{quteprints96243}.
 A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class.
@@
 This has previously been attempted using bag-of-frames approaches \cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
-Unfortunately, this only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterized by a few distinct sound events.
+Unfortunately, this only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events.
 Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neuroscience \cite{nelken2004processing} suggest that this more closely matches human acoustic perception.
@@
 For our task, one advantage of the scattering transform is not requiring a training step, allowing for a wider range of applications compared to learned features.
-Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unserpervised setting.
+Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unsupervised setting.
 
@@
 This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background \cite{nelken2013}.
-Such events are not well-characterized by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
+Such events are not well-characterised by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
 The latter approach should therefore prove more fruitful in measuring auditory scene similarity.
@@
 This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}.
-Studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}.
+Studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterise an acoustic scene \cite{guyot2005urban,kang2006urban}.
 Instead, cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} seem to rely upon higher-level cognitive attributes.
-These typically include the identities of the sound sources which consitute the scene.
+These typically include the identities of the sound sources which constitute the scene.
 It has been shown that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes.
@@
 Let $\boldsymbol{\psi}(t)$ a complex-valued analytic bandpass filter of
-center frequency $\xi_1$ and bandwidth $\xi_1/Q_1$, where $Q_1$ is the quality factor of the filter.
+central frequency $\xi_1$ and bandwidth $\xi_1/Q_1$, where $Q_1$ is the quality factor of the filter.
 A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
@@
 For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
-has a center frequency of $2^{-\gamma_1/Q_1}\xi_1$ and a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$ resulting in the same quality factor $Q_1$ as $\boldsymbol{\psi}$.
+has a central frequency of $2^{-\gamma_1/Q_1}\xi_1$ and a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$ resulting in the same quality factor $Q_1$ as $\boldsymbol{\psi}$.
 In the following, we set $\xi_1$ to $20~\mathrm{kHz}$, $J_1$ to $10$, and the quality factor $Q_1$, which is also the number of wavelets per octave, to $8$.
@@
 which is derived from the short-term Fourier transform (STFT) by averaging the frequency
-axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q_1}\xi_1$.
+axis into constant-Q subbands of central frequencies $2^{-\gamma_1/Q_1}\xi_1$.
 Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
@@
 We define second-order wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ in the same way as the first-order wavelets, but with parameters $\xi_2$, $J_2$, and $Q_2$.
-Consequently, they have center frequencies $2^{-\gamma_2/Q_2}\xi_2$ for $\gamma_2$ taking values between $0$ and $(J_2 Q_2 - 1)$.
+Consequently, they have central frequencies $2^{-\gamma_2/Q_2}\xi_2$ for $\gamma_2$ taking values between $0$ and $(J_2 Q_2 - 1)$.
 While this abuses notation slightly, the identity of the wavelets should be clear from context.
@@
 \label{fig:histograms}
-Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
+Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a central acoustic frequency of $302\,\mathrm{Hz}$,
 (a) before and (b) after logarithmic compression.}
@@
 We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone--a source of spurious variability in scene classification.
-Logarithmic compression linearizes this dependency, facilitating the construction of powerful invariants at the classifier stage.
+Logarithmic compression linearise this dependency, facilitating the construction of powerful invariants at the classifier stage.
 
@@
 This dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
-The dataset is evenly divided into $10$ acoustic scence classes.
+The dataset is evenly divided into $10$ acoustic scene classes.
 The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
@@
 
-Furthermore, it appears that \emph{RbQ-c} is better able to characterize the classes compared to \emph{RbQ-w}.
+Furthermore, it appears that \emph{RbQ-c} is better able to characterise the classes compared to \emph{RbQ-w}.
 Although not the only way of incorporating the number of frames associated to each centroid, the earth mover's distance is a rather natural of doing so.

### 2018-06-27 paper/revision/responseToReviewers.tex
addition of response

\documentclass[10pt]{article}
\usepackage{url}

\title{Reply to reviewers concerning submission JASM-D-18-00019 "Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio"}

\begin{document}

\maketitle

As a preamble, we would like to thank the reviewers for their comments and suggestions. Following these comments, we made several changes to the article, which are summarized here. The next sections list our answers to each of the reviewer’s comments, with references to the revised manuscript (page, column, and paragraph) where appropriate.

\section{Answer to Reviewer 1}

\begin{enumerate}

\item \emph{I understand that the authors cannot run such a test but I suggest that they only refer to ecoacoustics in the discussion by mentionning that this could be tested on other audio scenes as those recorded in natural environments.}

$\rightarrow$

\item \emph{The dataset is too quickly described. I had to read ref\#36 to understand how it was built and made of. It would be nice to have an example of a labelled recording either as a spectrogram or scalogram to estimate the complexity of the scenes recorded. The event classes should also be detailed.}

$\rightarrow$

\item \emph{It is a pity that the methods are not directly compared with the different strategies reported in ref\#36.}

$\rightarrow$

\item \emph{.P1L36: "superiority" and other superlative seems too strong. I would say that thet the proposed approach performed better on the dataset tested.}

$\rightarrow$

\item \emph{.P2L29: "cannot be trusted" seems too strong, any system is proned to errors. These errors need to be estimated so that we can estimate how much the system is reliable, but I would not roughly say that it cannot be trusted.}

$\rightarrow$

\item \emph{.P10L44: the title of this subsection does not seem to be totally appropriate, I would rename it "Evaluation and algorithm"}

$\rightarrow$

\item \emph{.P11L37: the comments about figure 3 (and others) is rather quick with no try of quantification of the differences/similarities between the p@k curves.}

$\rightarrow$

\item \emph{.P12L37: Delete the comment addressed to one of the authors}

$\rightarrow$

\item \emph{.P13L16: Using French for a project name might not be a good idea for reaching a large audience}

$\rightarrow$

\end{enumerate}

\section{Answer to Reviewer 2}

\begin{enumerate}

\item \emph{Only two applications are considered in the literature review, bioacoustics and urban sound environment, but they allow to present a problem that also exists in other application areas.}

$\rightarrow$

\item \emph{The authors present their technique as opposed to the BOF approach, that is considered "state of the art". BOF maybe a popular technique but the fact that gets rid of the temporal structure makes it useless in many tasks. In fact, in DCASE 2013 BOF was the chosen baseline system but no participant proposed that technique. Certainly, BOF allows the authors to make stronger the point about describing an acoustic scene from a few distinct events, but perhaps they should avoid to claim BOF as "the" state of the art in this problem.}

$\rightarrow$

\item \emph{It would be interesting to see the contribution to the results of the Gammatone wavelet in comparison to Morlet or another more classical wavelet.}

$\rightarrow$

\item \emph{There are several parameters that must be set in the experimentation. The reader may want to know if there is any specific parameter that is critical. What about M?}

$\rightarrow$

\item \emph{According to Section 6.2, the number of scattering features is much higher than MFCC features; 1367 instead of 40 for each vector, and there is a similar number of vectors per time unit. This would mean a much higher computational load. However, in the experiments, a projection is applied to reduce them to 30. Is it PCA? The convergence of the EM  algorithm is invoked, but the computational load may be an added reason.}

$\rightarrow$

\item \emph{I do not see the authors have mentioned in Section 7 the following observation (Fig.3): for MFCC the relevance-based quantization improves the performance wrt BOF.}

$\rightarrow$

\item \emph{Pag. 2, line 48: what "this" refers to?}

$\rightarrow$

\item \emph{Pag. 2, line 49: "More closely matches... than...}

$\rightarrow$

\item \emph{Pag. 3, line 22: an cluster -> a cluster}

$\rightarrow$

\item \emph{Pag. 3, line 23: Unconsistent sentence "These clusters to define..."}

$\rightarrow$

\item \emph{Equation (5): Notation is different from previous equations}

$\rightarrow$

\item \emph{Pag. 7, line 9: S1 should be S2}

$\rightarrow$

\item \emph{Pag. 9, line 48: suppress "the"}

$\rightarrow$

\end{enumerate}

\end{document}

### 2018-06-27 paper/lostanlenRelevanceScattering.tex
use endfloat
Address reviewer points 2.7-2.13

@@
 \usepackage[cmex10]{amsmath}
+
+\usepackage[pagewise]{lineno}
+\linenumbers
 
@@
 \fnbelowfloat
-
+\usepackage{endfloat}
 \usepackage{booktabs}
@@
 With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset.
-This requires us to represent an audio recording in a way that captures its distinctive qualities.
-This has previously been attempted using bag-of-frames approaches \cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
-Unfortunately, this only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events.
-Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neuroscience \cite{nelken2004processing} suggest that this more closely matches human acoustic perception.
-We believe that the failure to model such distinct events is one of the reasons why this approach is insufficient for moderate-sized evaluation datasets \cite{lagrange:hal-01082501}.
+To construct such an algorithm, we are required to represent an audio recording in a way that captures its distinctive qualities.
+A popular choice of representation is the bag of frames \cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
+Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events.
+Furthermore, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neuroscience \cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events.
+We believe that the failure to model such distinct events is one of the reasons why the bag-of-frames representation is insufficient for moderate-sized evaluation datasets \cite{lagrange:hal-01082501}.
 
@@
 
-In this work, we propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an cluster model.
+In this work, we propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by a cluster model.
 This unsupervised model quantizes the scattering coefficients into a given number of clusters.
-These clusters to define a set of distances for acoustic similarity retrieval.
+These clusters are then used to define a set of distances for acoustic similarity retrieval.
 Evaluating this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
@@
 \mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
-(\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \boldsymbol{\phi})(t).
+\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \boldsymbol{\phi}(t).
 \end{equation}
 
-The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ denotes either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
+The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ denotes either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
 While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
@@
 The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance $\widehat{\EMD}$ introduced by \cite{pele2008linear}.
-The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied by the a dissimilarity measure between the histogram bins.
+The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied by a dissimilarity measure between the histogram bins.
 Here, that measure is given by $1-K_{mn}^{uv}$.
@@
 \includegraphics[width=\columnwidth]{figures/baselines}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for several baseline approaches, as a function of the rank $k$.}
 \label{fig:ASS_-1}

### 2018-06-28 paper/revision/responseToReviewers.tex
Address reviewer points 2.7-2.13
Assign corrections for reviewers
Fix rendering of reviewer comment
Address some simple comments by R1
Address relation to DCASE 2013 systems

@@
 
-$\rightarrow$
+$\rightarrow$ ML
 
@@
 
-$\rightarrow$
+$\rightarrow$ VL
 
@@
 $\rightarrow$
+Since the tasks are different, it is not straightforward to compare the proposed methods with the approaches of ref\#36.
+In the former, we propose a method for acoustic scene similarity retrieval, while the latter discusses the problem of scene classification.
+In similarity retrieval, we would like to find other recordings in the dataset that are as similar to the original as possible (which here is defined as being in the same class).
+In scene classification, the task is to assign a class to each recording.
+In this sense, similarity retrieval is a more general problem.
+For our experiments, we calculate the p@k curves for each propsed method, allowing us to analyze the distribution of recordings most similar to a particular query recording.
+The classification setting is reproduced for $k = 1$, with the p@1 measure corresponding to the leave-one-out classification accuracy of nearest neighbors.
+As such, the classification task is a special case of the similarity retrieval setting considered here.
+We have modified the manuscript to make this distinction more clear.
 
@@
 
-$\rightarrow$
+$\rightarrow$ JA
 
@@
 $\rightarrow$
+This has been fixed in the revised manuscript.
 
@@
 
-$\rightarrow$
+$\rightarrow$ ML - Experiment with variable number $M$ for best approach.
 
@@
 $\rightarrow$
+This has been clarified in the revised manuscript.
 
@@
 $\rightarrow$
+This has been clarified in the revised manuscript.
 
-\item \emph{Pag. 3, line 22: an cluster -> a cluster}
+\item \emph{Pag. 3, line 22: an cluster -\textgreater{} a cluster}
 
 $\rightarrow$
+This has been clarified in the revised manuscript.
 
@@
 $\rightarrow$
+The notation has been fixed for consistency with the other equations.
 

### 2018-06-28 paper/lostanlenRelevanceScattering.tex
Address some simple comments by R1
Address relation to DCASE 2013 systems

@@
 The experiments in this paper are carried out on the DCASE 2013 dataset \cite{7100934}.
-This dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
+Although the dataset was constructed for the task of acoustic scene classification, where the goal is to correctly assign the class of a given recording, we can use the same recordings and class labels for the task of acoustic scene similarity retrieval.
+The dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
 The dataset is evenly divided into $10$ acoustic scene classes.
@@
 In addition, the relevance-based quantization using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
-\vl{It is worth reminding what the baselines are.}
 
@@
 \begin{itemize}
-\item Project name: article2017EstimationAmbiance
 \item \sloppy Project home page: \url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}

### 2018-06-28 paper/revision/responseToReviewers.tex
Address relation to BOF and state of the art

@@
 
-$\rightarrow$ JA
+$\rightarrow$
+As discussed above in the response to the first reviewer, the application in the DCASE 2013 paper, classification, is different from the one proposed in our work, similarity retrieval.
+For classification, the BOF approach is indeed not the state of the art.
+However, in acoustic scene similarity retrieval, we are not aware of any systems which outperform the BOF.
+As such, it is indeed the state of the art for the similarity retrieval task.
+The manuscript has been revised to clarify this point.
 

### 2018-06-28 paper/lostanlenRelevanceScattering.tex
Address relation to BOF and state of the art
More analysis of results in Figure 3

@@
 It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.
+While BOF has largely been superseded by more sophisticated methods for the task of acoustic scene classification \cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
 Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
@@
 
-As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied both to the MFCCs and the scattering features.
+As seen in Figure \ref{fig:ASS_-1}, scattering features significantly outperform the baseline MFCCs for both BOF and early integration schemes.
+This is expected, as the scattering transform extends the MFCCs by including supplementary amplitude modulation information \cite{Anden2014}.
+We also note that applying PCA reduction to the scattering transform has little effect for the early integration scheme.
+In the context of summary statistics, $30$ dimensions are sufficient to discriminate between auditory scenes.
+
+Comparing the BOF to early integration, both approaches perform similarly for MFCCs and PCA-reduced scattering features alike.
+This is in line with previous results on BOF \cite{lagrange:hal-01082501}, where it is found to perform similarly as similarity retrieval on features averaged over the entire recording.
 

### 2018-06-28 paper/revision/responseToReviewers.tex
More analysis of results in Figure 3
Address comments about MFCCs and RbQ vs. BOF
Respond to remark about PCA

@@
 
-$\rightarrow$ JA
+$\rightarrow$
+This section has been expanded in the revised manuscript to provide more interpretation of the results.
 
@@
 
-$\rightarrow$ JA
+$\rightarrow$
+As described in Section 6.3, the scattering features are indeed projected onto the top $30$ principal components for use in the BOF model.
+This is to mitigate the problem of running EM in such a high-dimensional space.
+For the other methods, it may reduce the computational load, but potentially at the cost of sacrificing certain information that may be leveraged by those other methods.
+For this reason, we have chosen to run them on the full, $1367$-dimensional, feature vectors.
 
@@
 
-$\rightarrow$ JA
+$\rightarrow$
+The manuscript has been revised to include this observation.
 

### 2018-06-29 paper/revision/responseToReviewers.tex
adressing comments

@@
 
-\item \emph{I understand that the authors cannot run such a test but I suggest that they only refer to ecoacoustics in the discussion by mentionning that this could be tested on other audio scenes as those recorded in natural environments.}
+\item \emph{I suggest that they only refer to ecoacoustics in the discussion by mentionning that this could be tested on other audio scenes as those recorded in natural environments.}
 
-$\rightarrow$ ML
+$\rightarrow$ The abstract introduction, background section and conclusion have been rewritten to widen the scope of application of the proposed method and suggest the ecoacoustic as potential area of interest for further research.
 
@@
 In this sense, similarity retrieval is a more general problem.
-For our experiments, we calculate the p@k curves for each propsed method, allowing us to analyze the distribution of recordings most similar to a particular query recording.
+For our experiments, we calculate the p@k curves for each proposed method, allowing us to analyze the distribution of recordings most similar to a particular query recording.
 The classification setting is reproduced for $k = 1$, with the p@1 measure corresponding to the leave-one-out classification accuracy of nearest neighbors.
@@
 We have modified the manuscript to make this distinction more clear.
+
+It is also worth considering that the BOF model is used differently in scene retrieval (our use case) and scene classification (ref\#36).
+In scene retrieval, given two scenes (a and b) modeled using GMMs (A and B), the similarity between two scenes is the similarity between A and B.
+In scene classification, each class of sound scenes (park, boulevard) are modeled using a GMM (P, B).
+Then, for an unknown scene $u$, the likelihood of $u$ given P and B are computed.
+The scene $u$ is then labeled P if the likelihood of $u$ given P is greater than the likelihood of $u$ given B.
+
+Even if the technical aspects are quite close, the way the algorithm is designed is different as the tasks they have to solve are differents.
+A synthetic description of the matter is added to Section 2.
 
@@
 
-$\rightarrow$ ML
+$\rightarrow$ We reformulated in a more abstract way the task we attempt to solve, and give potential areas of applications in the introduction and conlusion sections.
 
@@
 
-\item \emph{There are several parameters that must be set in the experimentation. The reader may want to know if there is any specific parameter that is critical. What about M?}
+\item \emph{There are several parameters that must be set in the experimentation. The reader may want to know if there is any specific parameter that is critical.}
 
-$\rightarrow$ ML - Experiment with variable number $M$ for best approach.
+$\rightarrow$ Considering the features design, we set the parameters to widely used values that have shown to be robust.
+We also ran a sensitivity analysis for the parameters that controls the proposed algorithm: scaling parameter $q$ and the number of clusters $M$.
+None of them are found to be critical providing that they are set in a reasonable range.
+It is worth noting that the scaling method is specifically designed for reducing sensitivity compared to the direct RBF kernel formulation.
+
+\item \emph{What about M?}
+
+$\rightarrow$ A discussion about the setting of the parameter M is added at the end of Section 6.3.
 

### 2018-06-29 paper/lostanlenRelevanceScattering.tex
Address comments about MFCCs and RbQ vs. BOF
Remove line numbers
adressing comments

--- 2018-06-28 paper/lostanlenRelevanceScattering.tex
@@
 \usepackage[cmex10]{amsmath}
-
-\usepackage[pagewise]{lineno}
-\linenumbers
 
@@
 \begin{abstract}
-The emerging field of computational ecoacoustics aims at retrieving high-level information from acoustic scenes recorded by some network of sensors.
+The emerging field of computational acoustic monitoring aims at retrieving high-level information from acoustic scenes recorded by some network of sensors.
 These networks gather large amounts of data requiring analysis.
-To decide which parts to inspect manually, we need tools that automatically mine the data, identifying recurring patterns and isolated events.
+To decide which parts to inspect further, we need tools that automatically mine the data, identifying recurring patterns and isolated events.
 This requires a similarity measure for acoustic scenes that does not impose strong assumptions on the data.
@@
 The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
-In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
+In order to measure the effect of human activity and climate change on animal biodiversity~\cite{sueur2015ecoacoustics}, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world~\cite{warren2006urban, NessSST13, stowell13b}.
+In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas~\cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic~\cite{gloaguen}.
 Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for to signal processing community.
 
-These questions are explored in an emerging field of computational bioacoustics, known as ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
-An important problem in this field is that manually analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling}.
+%These questions are explored in an emerging field of computational bioacoustics, known as ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science .
+%~\cite{wimmer2013sampling}
+
+An important problem is that manually analysing the recorded data to identify the quantities of interest is very costly.
 Some sort of pre-screening is therefore required to reduce the need for human expert listening and annotation.
 
-The most straightforward approach is to specify a closed set of sound classes, such as sounds of animals expected to live near the acoustic sensors.
-Computational models are trained then  for these classes which are used to automatically annotate recordings \cite{quteprints96243}.
+The most straightforward approach is to specify a closed set of sound classes, such as sounds classes expected to appear near the acoustic sensors.
+Computational models are then trained  for these classes which are used to automatically annotate recordings~\cite{quteprints96243}.
 A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class.
@@
 To construct such an algorithm, we are required to represent an audio recording in a way that captures its distinctive qualities.
-A popular choice of representation is the bag of frames \cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
+A popular choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
 Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events.
-Furthermore, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neuroscience \cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events.
-We believe that the failure to model such distinct events is one of the reasons why the bag-of-frames representation is insufficient for moderate-sized evaluation datasets \cite{lagrange:hal-01082501}.
+Furthermore, experiments in cognitive psychology~\cite{dubois2006cognitive} and cognitive neuroscience~\cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events.
+We believe that the failure to model such distinct events is one of the reasons why the bag-of-frames representation is insufficient~\cite{lagrange:hal-01082501}.
 
@@
 This is often achieved using mel-frequency cepstral coefficients (MFCCs).
-Originally developed for speech processing \cite{davis-mermelstein}, MFCCs have recently found wider use in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
-A richer representation, the scattering transform, has enjoyed significant success in various audio \cite{Anden2014} and biomedical \cite{chudacek} signal classification tasks.
-Its structure is that of a convolutional neural network \cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}, but with fixed filters.
-Specifically, it alternates convolutions with wavelet filters and pointwise nonlinearities to ensure time-shift invariance and time-warping stability \cite{Mallat2012}.
+Originally developed for speech processing~\cite{davis-mermelstein}, MFCCs have recently found wider use in music information retrieval~\cite{logan} and environmental audio processing~\cite{aucouturier2007bag}.
+A richer representation, the scattering transform, has enjoyed significant success in various audio~\cite{Anden2014} and biomedical~\cite{chudacek} signal classification tasks.
+Its structure is that of a convolutional neural network~\cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}, but with fixed filters.
+Specifically, it alternates convolutions with wavelet filters and pointwise nonlinearities to ensure time-shift invariance and time-warping stability~\cite{Mallat2012}.
 
@@
 
-Motivations of the proposed approach and a brief review of the state of the art in acoustic scene modeling are given in Section \ref{sec:soa}.
-We describe the scattering transform in Section \ref{sec:scattering}, discuss feature post-processing in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}.
-Section \ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task.
-Results are reported in Section \ref{sec:results}.
+Motivations of the proposed approach and a brief review of the state of the art in acoustic scene modeling are given in Section~\ref{sec:soa}.
+We describe the scattering transform in Section~\ref{sec:scattering}, discuss feature post-processing in Section~\ref{sec:design} and propose a cluster-based scene description in Section~\ref{sec:object}.
+Section~\ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task.
+Results are reported in Section~\ref{sec:results}.
 
@@
 
-Computational bioacoustics refers to the numerical investigation of sound production, dispersion and reception in animals, including humans.
+%Computational bioacoustics refers to the numerical investigation of sound production, dispersion and reception in animals, including humans.
 A recent paradigm processes the audio stream in a holistic way over large time scales, without assuming that a single species is present throughout the recording.
-Automated systems within this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010} and migration patterns \cite{Obrist2010}.
-They also mark time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.
+Automated systems within this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices~\cite{Bardeli2010} and migration patterns~\cite{Obrist2010}.
+They also mark time intervals of particular interest for detailed human inspection~\cite{rosenstock2002landbird}.
 
-A closely related field is that of urban sound environment analysis.
-In this context, a popular approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al. \cite{aucouturier2007bag}.
+Caracterization of the similarity between audio recordings be they at the scale of the minute, the hour, the day or larger is of interest for many applications areas involving acoustic monitoring such as urban sound environment analysis and ecoacoustics.
+In this context, a popular approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
 It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.
-Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
-This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background \cite{nelken2013}.
+
+It is worth mentionning that this task typically falls into a unsupervised paradigm where no prior knowldge is used to model a given scene.
+For each scene $s$, a model is compute $M_s$ is computed.
+The similarity among the scene $s_1$ and $s_2$ is computed as the similarity, see Section~\ref{sec:experiments} for further details.
+The BOF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}.
+In this case, each class of scenes of a given typology, say \{park, boulevard, square \}, is modeled by a GMM trained on scenes taken from a training set.
+In order to predict the class of a given scene $s$, the likelihood each model given the scene are computed.
+The scene is then labeled park if the likelihood of the GMM trained on park scenes is higher than the other likelihoods.
+
+While BOF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
+Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
+This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background~\cite{nelken2013}.
 Such events are not well-characterised by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
@@
 
-This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}.
-Studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterise an acoustic scene \cite{guyot2005urban,kang2006urban}.
-Instead, cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} seem to rely upon higher-level cognitive attributes.
+This statement has some support in auditory psychology as well as sound synthesis based on summary statistics~\cite{mcdermott2013summary}.
+Studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterise an acoustic scene~\cite{guyot2005urban,kang2006urban}.
+Instead, cognitive processes such as sound environment quality perception~\cite{dubois2006cognitive} or loudness judgment~\cite{kuwano_memory_2003} seem to rely upon higher-level cognitive attributes.
 These typically include the identities of the sound sources which constitute the scene.
@@
 For example, in urban areas the presence of birds is likely to be heard in parks and are therefore strong pleasantness indicators.
-Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments \cite{ricciardi2015sound,lavandier2006contribution}.
+Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments~\cite{ricciardi2015sound,lavandier2006contribution}.
 Although the complete set of events occurring within a given auditory stream may not be discernable even to human expects, research has shown that a small set of events (so-called markers) suffice to reliably predict many high-level attributes.
@@
 This results in the wavelet filters covering the whole range of human hearing, from $20~\mathrm{Hz}$ to $20~\mathrm{kHz}$.
-Setting $Q_1 = 8$ results in filters whose bandwidth approximates an equivalent rectangular bandwidth (ERB) scale \cite{Fastl2007}.
+Setting $Q_1 = 8$ results in filters whose bandwidth approximates an equivalent rectangular bandwidth (ERB) scale~\cite{Fastl2007}.
 
@@
 While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
-Indeed, higher-order scattering coefficients have been shown to contain reduced energy and are therefore of limited use \cite{irene}.
+Indeed, higher-order scattering coefficients have been shown to contain reduced energy and are therefore of limited use~\cite{irene}.
 
@@
 $\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
-wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
+wavelets with one vanishing moment~\cite{Venkitaraman2014}, and are shown in Figure~\ref{fig:gammatones}.
 In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
-Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
+Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking~\cite{Fastl2007}.
 The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
@@
 It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
-As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
+As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds~\cite{Smith2006}.
 This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.
@@
 
-Many algorithms in pattern recognition, including nearest neighbor classifiers and SVMs, tend to work best when all features follow a standard normal distribution across all training instances \cite{Hsu2003}.
+Many algorithms in pattern recognition, including nearest neighbor classifiers and SVMs, tend to work best when all features follow a standard normal distribution across all training instances~\cite{Hsu2003}.
 Yet the distribution of the scattering coefficients is skewed towards larger values.
@@
 In particular, we find that the logarithm performs particularly well in this respect.
-Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
+Figure~\ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
 
@@
 Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where $\gamma$ and $n$ denote feature and sample indices, respectively.
-Many algorithms operate better on features which have zero mean and unit variance to avoid mismatch in numeric ranges \cite{Hsu2003}.
+Many algorithms operate better on features which have zero mean and unit variance to avoid mismatch in numeric ranges~\cite{Hsu2003}.
 To standardise $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
@@
 
-%As part of the aforementioned research areas and applications, the emerging field of \emph{Acoustic Scene Analysis} (also called \emph{Sound Scene Analysis}) \cite{Stowell15} aims to develop approaches and systems for the automatic analysis of environmental sounds and soundscapes (originating both from urban or nature environments).
-While research methodologies in related fields such as Automatic Speech Recognition (ASR) \cite{Rabiner93} and Music Information Retrieval (MIR) \cite{Muller07} are now well established, research addressing Acoustic Scene Analysis remains relatively young.
+%As part of the aforementioned research areas and applications, the emerging field of \emph{Acoustic Scene Analysis} (also called \emph{Sound Scene Analysis})~\cite{Stowell15} aims to develop approaches and systems for the automatic analysis of environmental sounds and soundscapes (originating both from urban or nature environments).
+While research methodologies in related fields such as Automatic Speech Recognition (ASR)~\cite{Rabiner93} and Music Information Retrieval (MIR)~\cite{Muller07} are now well established, research addressing Acoustic Scene Analysis remains relatively young.
 
-%From a data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.
+%From a data processing point of view, a holistic scheme, such as the bag-of-frame approach~\cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions~\cite{lagrange:hal-01082501}.
 
-As discussed in Section \ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
-While this can be addressed in the supervised case using late integration of discriminative classifiers \cite{Anden2014}, this is not directly feasible in the unsupervised case.
-As the detection of events is still an open problem \cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
+As discussed in Section~\ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
+While this can be addressed in the supervised case using late integration of discriminative classifiers~\cite{Anden2014}, this is not directly feasible in the unsupervised case.
+As the detection of events is still an open problem~\cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
 
 Given a set of $d$-dimensional feature vectors $X_u = \{x_1^u, \ldots, x_L^u\}$, extracted from the scene $s_u$, where $u=\lbrace 1,2,\ldots,U\rbrace$, we would like to partition $X_u$ into a set $C_u = \{c^u_1, \ldots, c^u_M\}$ of $M$ clusters.
-This partition is obtained by minimizing the variance of each cluster and known as a $k$-means clustering \cite{lloyd}.
+This partition is obtained by minimizing the variance of each cluster and known as a $k$-means clustering~\cite{lloyd}.
 Each scene $s_u$ is then described by a set of clusters $C_u$.
-Note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
+Note that this quantization approach differs from unsupervised learning schemes such as the ones studied in~\cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
 Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroids.
 
-The similarity between the scene centroids $\mu_m^u$ over the entire dataset, is computed using a radial basis function (RBF) kernel $K$ combined with a local scaling method \cite{selfTuneManor2004}:
+The similarity between the scene centroids $\mu_m^u$ over the entire dataset, is computed using a radial basis function (RBF) kernel $K$ combined with a local scaling method~\cite{selfTuneManor2004}:
 \begin{equation}
@@
 The weight $w_m^u$ for the $m$th centroid $\mu_m^u$ is the number of frames belonging to a particular cluster.
-The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance $\widehat{\EMD}$ introduced by \cite{pele2008linear}.
+The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance $\widehat{\EMD}$ introduced by~\cite{pele2008linear}.
 The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied by a dissimilarity measure between the histogram bins.
@@
 
-The experiments in this paper are carried out on the DCASE 2013 dataset \cite{7100934}.
+The experiments in this paper are carried out on the DCASE 2013 dataset~\cite{7100934}.
 Although the dataset was constructed for the task of acoustic scene classification, where the goal is to correctly assign the class of a given recording, we can use the same recordings and class labels for the task of acoustic scene similarity retrieval.
@@
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
+As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices~\cite{lagrange:hal-01082501}.
 
@@
 Here we discard $3$ seconds from the beginning and end of the scene to avoid boundary artifacts.
-We also conduct experiments with and without logarithmic compression of the scattering coefficients (see Section \ref{sec:logcomp}).
+We also conduct experiments with and without logarithmic compression of the scattering coefficients (see Section~\ref{sec:logcomp}).
 
@@
 The distance on this average feature vector is then used to determine $p@k$.
-For the BOF approach of Aucouturier et al.
-\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird, moon1996expectation}.
+For the BOF approach of Aucouturier et al.~\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm~\cite{dempster-laird, moon1996expectation}.
 The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation.
@@
 Best $p@5$ is reached with 8 and 4  Gaussians, respectively for MFCCs and scattering features.
-Recommended number of Gaussians for MFCCs given in \cite{aucouturier2007bag} is 10.
+Recommended number of Gaussians for MFCCs given in~\cite{aucouturier2007bag} is 10.
 
-The scaling parameter $q$ of the RBF kernels (see Eq. \ref{eq:kc}) is set to $10\%$ of the number of data points to cluster. For each method, the numbers of clusters is set to $8$.
+The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.
+As the number of Gaussians for the BOF approach, the numbers of clusters $M$ controls the level of abstraction.
+For each method, the parameter $M$ is set to $8$.
+It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds).
 
@@
 Results for the acoustic scene similarity retrieval task demonstrate the superiority of the logarithmically compressed scattering features compared to MFCCs.
-Combining these with the \emph{RbQ} cluster model, we obtain significant improvements over traditional BOF and summary statistic measures.
+Combining these with the \emph{RbQ} cluster model, improvements are obtained over traditional BOF and summary statistic measures.
 
@@
 
-As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied both to the MFCCs and the scattering features.
+As seen in Figure~\ref{fig:ASS_-1}, scattering features significantly outperform the baseline MFCCs for both BOF and early integration schemes.
+This is expected, as the scattering transform extends the MFCCs by including supplementary amplitude modulation information~\cite{Anden2014}.
+We also note that applying PCA reduction to the scattering transform has little effect for the early integration scheme.
+In the context of summary statistics, $30$ dimensions are sufficient to discriminate between auditory scenes.
+
+Comparing the BOF to early integration, both approaches perform similarly for MFCCs and PCA-reduced scattering features alike.
+This is in line with previous results on BOF~\cite{lagrange:hal-01082501}, where it is found to perform similarly as similarity retrieval on features averaged over the entire recording.
 
@@
 
-Figure \ref{fig:ASS_0} shows that logarithmic compression of the scattering features is beneficial.
+Figure~\ref{fig:ASS_0} shows that logarithmic compression of the scattering features is beneficial.
 For clarity sake, data is shown  for the \emph{early} approach only, but a equivalent gain is achieved for the relevance-based quantization approaches.
@@
 However, it is worth noting that \emph{RbQ-a} performs comparably to or worse than \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
-This result is in line with the findings of \cite{lagrange:hal-01082501}.
+This result is in line with the findings of~\cite{lagrange:hal-01082501}.
 To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects.
+The same behavior is observed for MFCCs, with \emph{RbQ-c} and \emph{RbQ-w} outperforming \emph{early}, which is equivalent to the state-of-the-art BOF model, as seen previously.
 
@@
 
-With $p@5$ as our metric (cf.
-\cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), we see that replacing MFCCs by the logarithmically compressed scattering transform increases performance from $0.31$ to $0.49$.
+With $p@5$ as our metric (cf.~\cite{aucouturier2007bag} and~\cite{lagrange:hal-01082501}), we see that replacing MFCCs by the logarithmically compressed scattering transform increases performance from $0.31$ to $0.49$.
 In addition, the relevance-based quantization using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
@@
 For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
-These outcomes shall be studied further in future work by considering larger databases and emerging tasks in computational bioacoustics \cite{wimmer2013sampling}.
+These outcomes shall be studied further in future work by considering larger databases and emerging tasks in ecoacoustics~\cite{sueur2015ecoacoustics, wimmer2013sampling}.
 

### 2018-07-06 paper/lostanlenRelevanceScattering.tex
remove the word "superiority"

@@
 The scattering coefficients capture short-scale structure, while the cluster model captures longer time scales, allowing for more accurate characterization of sparse events.
-Evaluation within the acoustic scene similarity framework demonstrates the superiority of the proposed approach.
+Evaluation within the acoustic scene similarity framework demonstrates the interest of the proposed approach.
 \keywords{
@@
 
-Results for the acoustic scene similarity retrieval task demonstrate the superiority of the logarithmically compressed scattering features compared to MFCCs.
+Results for the acoustic scene similarity retrieval task demonstrate that logarithmically compressed scattering features outerform MFCCs.
 Combining these with the \emph{RbQ} cluster model, improvements are obtained over traditional BOF and summary statistic measures.

### 2018-07-06 paper/revision/responseToReviewers.tex
remove the word "superiority"
remove "cannot be trusted"

@@
 \documentclass[10pt]{article}
+\usepackage{amssymb}
 \usepackage{url}
@@
 
-\item \emph{.P1L36: "superiority" and other superlative seems too strong. I would say that thet the proposed approach performed better on the dataset tested.}
+\item \emph{.P1L36: "superiority" and other superlative seems too strong. I would say that the proposed approach performed better on the dataset tested.}
 
-$\rightarrow$ VL
+$\rightarrow$ VL $\checkmark$
 
-\item \emph{.P2L29: "cannot be trusted" seems too strong, any system is proned to errors.
+\item \emph{.P2L29: "cannot be trusted" seems too strong, any system is prone to errors.
 These errors need to be estimated so that we can estimate how much the system is reliable, but I would not roughly say that it cannot be trusted.}

### 2018-07-06 paper/lostanlenRelevanceScattering.tex
remove "cannot be trusted"
rename "Algorithm" subsection
e.g. -> \eg{}

@@
 Some sort of pre-screening is therefore required to reduce the need for human expert listening and annotation.
-
-The most straightforward approach is to specify a closed set of sound classes, such as sounds classes expected to appear near the acoustic sensors.
+To this aim, the most straightforward approach is to specify a closed set of sound classes, such as sounds classes expected to appear near the acoustic sensors.
 Computational models are then trained  for these classes which are used to automatically annotate recordings~\cite{quteprints96243}.
-A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class.
+A given time interval (\eg{} a single day) is then represented by the number of events detected during that interval for each class.
 This allows the scientist to drastically reduce the amount of information requiring manual processing.
 However, this approach has two drawbacks.
-First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be trusted.
+First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors.
 Second, and more importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.
@@
 The BOF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}.
-In this case, each class of scenes of a given typology, say \{park, boulevard, square \}, is modeled by a GMM trained on scenes taken from a training set.
+In this case, each class of scenes of a given typology, say \{\emph{park, boulevard, square }\}, is modeled by a GMM trained on scenes taken from a training set.
 In order to predict the class of a given scene $s$, the likelihood each model given the scene are computed.
@@
 
-\subsection{Algorithm}
+\subsection{Evaluation and algorithm}
 

### 2018-07-06 paper/lostanlenRelevanceScattering.tex
a priori -> prior

@@
 First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors.
-Second, and more importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.
+Second, and more importantly, it is based on prior knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.
 

### 2018-07-06 paper/lostanlenRelevanceScattering.tex
brush up intro

@@
 However, this approach has two drawbacks.
-First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors.
-Second, and more importantly, it is based on prior knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.
+First, it relies on trained models whose prediction on unseen data -- \ie{} sensors outside of the training set -- is prone to errors.
+Secondly, and more importantly, it is based on prior knowledge and thus cannot be considered for exploratory analysis, in which quantities of interest have yet to be defined.
 
-To identify which parts need human inspection, we need tools that can detect both recurring patterns and sparsely distributed events.
+To identify which parts need human inspection, one needs tools that are able to detect both recurring patterns and sparsely distributed events.
 Identifying recurring patterns allows the user to focus on certain time points for manual annotation, while detection of more rare structures enables discovery of unforeseen phenomena.
@@
 To construct such an algorithm, we are required to represent an audio recording in a way that captures its distinctive qualities.
-A popular choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
+A widespread choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
 Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events.
@@
 Caracterization of the similarity between audio recordings be they at the scale of the minute, the hour, the day or larger is of interest for many applications areas involving acoustic monitoring such as urban sound environment analysis and ecoacoustics.
-In this context, a popular approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
+In this context, a classical approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
 It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.

### 2018-07-06 paper/lostanlenRelevanceScattering.tex
cite mallat 2012 cpam

@@
 The wavelet scattering transform, described below, has both of these desired properties: invariance to time-shifting and stability to time-warping.
-The stability condition can be formulated as a Lipschitz continuity property, which guarantees that the feature transforms of $\boldsymbol{x}(t)$ and $\boldsymbol{x_\tau}(t)$ are close together if $|\tau'(t)|$ is bounded by a small constant.
+The stability condition can be formulated as a Lipschitz continuity property, which guarantees that the feature transforms of $\boldsymbol{x}(t)$ and $\boldsymbol{x_\tau}(t)$ are close together if $|\tau'(t)|$ is bounded by a small constant \cite{mallat2012cpam}.
 

### 2018-07-06 paper/lostanlenRelevanceScattering.tex
reduce whitespace in integral

@@
 \subsection{Wavelet scalogram}
-Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int_{-\infty}^{+\infty} x(t) \exp(- \mathrm{i} 2\pi \omega t) \; \mathrm{d}t$.
+Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int_{-\infty}^{+\infty} x(t) \exp(- \mathrm{i} 2\pi \omega t) \, \mathrm{d}t$.
 Let $\boldsymbol{\psi}(t)$ a complex-valued analytic bandpass filter of

### 2018-07-06 paper/lostanlenRelevanceScattering.tex
masker tone -> masking tone

@@
 The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
-Likewise, because critical bands are skewed towards higher frequencies, a masker tone has to be louder if it is above the stimulus in frequency rather than below.
+Likewise, because critical bands are skewed towards higher frequencies, a masking tone has to be louder if it is above the stimulus in frequency rather than below.
 

### 2018-07-09 paper/lostanlenRelevanceScattering.tex
cite sainath 2015 interspeech
cite the Gammatone benchmark in lostanlen2017phd

@@
 Likewise, because critical bands are skewed towards higher frequencies, a masking tone has to be louder if it is above the stimulus in frequency rather than below.
-
 It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
-As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds~\cite{Smith2006}.
+As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds~\cite{Smith2006,sainath2015interspeech}.
+In addition, Gammatone wavelets have proven to outperform Morlet wavelets on a benchmark of supervised musical instrument classification from scattering coefficients \cite{lostanlen2017phd}.
 This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.

### 2018-07-10 paper/lostanlenRelevanceScattering.tex
appendix

@@
 \begin{itemize}
-\item \sloppy Project home page: \url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}
+\item \sloppy Companion website: \url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}
 \item Programming language: Matlab
@@
 
+\section{Supplementary material}
+
+Time reversal of a real signal $\boldsymbol{x}(t)$ is equivalent
+to the complex conjugation of its Fourier transform $\boldsymbol{\widehat{x}}(\omega$).
+As a consequence, the Fourier transform modulus $\vert\boldsymbol{\widehat{x}}(\omega)\vert$
+is not only invariant to translation, but also invariant to time reversal.
+Yet, although invariance to translation is needed for classification,
+invariance to time reversal is an undesirable property.
+
+The situation is different in the time-frequency domain. If the wavelet
+$\boldsymbol{\psi}(t)$ has a symmetric envelope, that is, if there
+exists a phase shift $\varphi\in[0;2\pi[$ such that $\boldsymbol{\psi}(-t)=\mathrm{e}^{\mathrm{i}\varphi}\boldsymbol{\psi}(t)$,
+the wavelet transform is covariant with respect to time reversal:
+\begin{eqnarray*}
+\mathbf{U}_{1}(\mathcal{W}_{\boldsymbol{\tau}}\boldsymbol{x})(t,\gamma) & = & \left|\mathrm{e}^{\mathrm{i}\varphi}\int_{-\infty}^{+\infty}\boldsymbol{x}(-t^{\prime})\boldsymbol{\psi}_{2^{\gamma},T(2^{\gamma})}(t^{\prime}-t)\;\mathrm{d}t^{\prime}\right|\\
+ & = & \left|\mathrm{e}^{\mathrm{i}\varphi}\right|\times\left|\int_{-\infty}^{+\infty}\boldsymbol{x}(t^{\prime})\boldsymbol{\psi}_{2^{\gamma},T(2^{\gamma})}(-t-t^{\prime})\;\mathrm{d}t^{\prime}\right|\\
+ & = & \mathbf{U}_{1}\boldsymbol{x}(t,\gamma).
+\end{eqnarray*}
+Gabor wavelets have a symmetric envelope, as they satisfy $\boldsymbol{\psi}(-t)=-\boldsymbol{\psi}(t)$,
+i.e. $\varphi=\pi$. This is due to the fact that Gaussians have bell-shaped
+symmetric envelopes. Therefore, the Gabor wavelet transform commutes
+with time reversal: 
+\[
+\mathbf{U}_{1}(\mathcal{W}_{\boldsymbol{\tau}}\boldsymbol{x})(t,\gamma)=\mathbf{U}_{1}\boldsymbol{x}(-t,\gamma).
+\]
+A simple way to break this undesirable invariance is to choose $\boldsymbol{\psi}(t)$
+as an asymmetric wavelet instead of a Gabor symmetric wavelet.
+
+The complex-valued Gammatone wavelet is a modification of the real-valued
+Gammatone auditory filter, originated in auditory physiology. The
+Gammatone auditory filter of dimensionless frequency $1$ is defined
+as a gamma distribution of order $N\in\mathbb{N}^{*}$ and bandwidth
+$\sigma$ modulated by a sine wave, that is, 
+\[
+t^{N-1}\exp(-2\pi\sigma t)\cos(2\pi t).
+\]
+For a fixed $\sigma$, the integer $N$ controls the relative shape
+of the envelope, becoming less skewed as $N$ increases. Psychoacoustical
+experiments have shown that, for $N=4$, the Gammatone function provides
+a valid approximation of the basilar membrane response in the mammalian
+cochlea \cite{Flanagan1960,Patterson1976,Lyon2010}. In particular,
+it is asymmetric both in the time domain and in the Fourier domain,
+which allows to reproduce the asymmetry of temporal masking as well
+as the asymmetry of spectral masking \cite{Fastl2007}. It is thus
+used in computational models for auditory physiology \cite{Pressnitzer2005}.
+However, it does not comply with the Grossman-Morlet admissibility
+condition, because it has a non-negligible average. In addition, because
+the Gammatone auditory filter takes real values in the time domain,
+its Fourier transform satisfies Hermitian symmetry, which implies
+that it does not belong to the space $H^{2}$ of analytic functions.
+More generally, there are no real-valued functions in $H^{2}$ \cite{Grossmann1984}.
+
+With the aim of building a pseudo-analytic admissible Gammatone wavelet,
+\cite{Venkitaraman2014} have modified the definition of the Gammatone
+auditory filter, by replacing the real-valued sine wave $\cos(2\pi t)$
+by its analytic part $\exp(2\pi\mathrm{\mathrm{i}}t)$ and by taking
+the first derivative of the gamma distribution, thus ensuring null
+mean. The definition of the Gammatone wavelet becomes
+\[
+\boldsymbol{\psi}(t)=\left(2\pi(\mathrm{i}-\sigma)t^{N-1}+(N-1)t^{N-2}\right)\exp(-2\pi\sigma t)\exp(2\pi\mathrm{i}t)
+\]
+in the time domain, and
+\[
+\boldsymbol{\widehat{\psi}}(\omega)=\dfrac{\mathrm{i}\omega\times(N-1)!}{(\sigma+\mathrm{i}(\omega-\sigma))^{N}}
+\]
+in the Fourier domain. Besides its biological plausibility, the Gammatone
+wavelet enjoys a near-optimal time-frequency localization with respect
+to the Heisenberg uncertainty principle. Furthermore, this time-frequency
+localization tends to optimality as $N$ approaches infinity, because
+the limit $N\rightarrow+\infty$ yields a Gabor wavelet \cite{Cohen1995}.
+Last but not least, the Gammatone wavelet transform of finite order
+$N$ is causal, as opposed to the Morlet wavelet transform, which
+makes it better suited to real-time applications. From an evolutionary
+point of view, it has been argued that the Gammatone reaches a practical
+compromise between time-frequency localization and causality constraints.
+\cite{Venkitaraman2014} did not provide a formula for deducing $\sigma$
+from the specification of a quality factor $Q$.
+In this appendix, we provide a rationale for choosing the topmost center frequency $\xi$ of a Gammatone wavelet filter bank in a discrete-time setting.
+Then, we relate the bandwidth parameter $\sigma$ to the choice of a quality factor Q.
+
+In order to preserve energy and allow for perfect reconstruction,
+the Gammatone wavelet filter bank must satisfy the inequalities
+\[
+1-\varepsilon\leq\vert\boldsymbol{\hat{\phi}}(\omega)\vert+\sum_{\gamma}\vert\boldsymbol{\hat{\psi}}(2^{\gamma}\omega)\vert+\vert\boldsymbol{\hat{\psi}}(-2^{\gamma}\omega)\vert\leq1
+\]
+for all frequencies $\omega$, where $\varepsilon$ is a small margin
+\cite{Anden2014}. Satisfying the equation above near the Nyquist
+frequency $\omega=\pi$ can be achieved by placing the log-frequency
+$\log_{2}\xi$ of the first (topmost) wavelet in between the log-frequency
+$\log_{2}(\xi\times2^{-1/Q})$ of the second wavelet and the log-frequency
+$\log_{2}(2\pi-\xi)$ of the mirror of the first wavelet. We obtain
+the equation 
+\[
+\log_{2}\xi-\log_{2}(\xi\times2^{-1/Q})=\log_{2}(2\pi-\xi)-\log_{2}\xi,
+\]
+
+of which we deduce the identity
+\[
+\xi=\dfrac{2\pi}{1+2^{1/Q}}.
+\]
+For $Q=1$, this yields a center frequency of $\xi=\frac{2\pi}{3}$.
+For greater values of $Q$, the center frequency $\xi$ tends towards
+$\pi$.
+
+\subsubsection*{Bandwidth parameter}
+
+The quality factor $Q$ of the Gammatone wavelet is defined as the
+ratio between the center frequency $\xi$ of the wavelet $\boldsymbol{\hat{\psi}}(\omega)$
+and its bandwidth $B$ in the Fourier domain. This bandwidth is given
+by the difference between the two solutions $\omega$ of the following
+equation:
+
+\[
+\dfrac{\vert\boldsymbol{\hat{\psi}}(\omega)\vert}{\vert\boldsymbol{\hat{\psi}}(\xi)\vert}=\dfrac{\omega}{\xi}\times\left(1+\dfrac{\left(\omega-\xi\right)^{2}}{\sigma^{2}}\right)^{-N/2}=r,
+\]
+where the magnitude cutoff $r$ is most
+often set to $\sqrt{\frac{1}{2}}$. Let $\Delta\!\omega=\omega-\xi$.
+Raising the above equation to the power $N/2$ yields the following:
+\[
+\left(1+\dfrac{\Delta\!\omega}{\omega_{c}}\right)^{N/2}=r\times\left(1+\frac{\Delta\!\omega^{2}}{\alpha^{2}}\right).
+\]
+Since $\Delta\!\omega\ll\xi$, we may approximate the left-hand side
+with a first-order Taylor expansion. This leads to a quadratic equation
+of the variable $\Delta\negmedspace\omega$:
+\[
+\dfrac{r^{2/N}}{\sigma^{2}}\times\Delta\!\omega^{2}-\dfrac{2}{N\xi}\times\Delta\!\omega+\left(r^{2/N}-1\right)=0.
+\]
+The discriminant of the above equation is:
+\[
+D=4\times\left(\dfrac{1}{N^{2}\xi^{2}}+\dfrac{r^{2/N}\left(1-r^{2/N}\right)}{\sigma^{2}}\right),
+\]
+which is a positive number as long as $r<1$. The bandwidth $B$ of
+$\boldsymbol{\hat{\psi}}$ is given by the difference between the
+two solutions of the quadratic equation, that is:
+\[
+B=\dfrac{2\sigma^{2}}{r^{2/N}}\times\sqrt{\dfrac{1}{N^{2}\xi^{2}}+\dfrac{r^{2/N}\left(1-r^{2/N}\right)}{\sigma^{2}}}.
+\]
+Now, let us express the parameter $\alpha$ as a function of some
+required bandwidth $B$ at some cutoff threshold $r$. After having
+raised the above to its square and rearranged the terms, we obtain
+another quadratic equation, yet of the variable $\mbox{\ensuremath{\alpha}}^{2}$:
+\[
+\dfrac{4}{r^{4/N}N^{2}\xi^{2}}\sigma^{4}+\dfrac{4\times\left(1-r^{2/N}\right)}{r^{2/N}}\sigma^{2}-B^{2}=0
+\]
+We multiply the equation by $\dfrac{r^{2/N}}{4\times\left(1-r^{2/N}\right)}\neq0:$
+\[
+\dfrac{1}{r^{2/N}\left(1-r^{2/N}\right)N^{2}\xi^{2}}\sigma^{4}+\sigma^{2}-\dfrac{r^{2/N}B^{2}}{4\times\left(1-r^{2/N}\right)}=0
+\]
+This leads to defining $\sigma^{2}$ as the unique positive root of
+the above polynomial:
+\[
+\sigma^{2}=\dfrac{r^{2/N}\left(1-r^{2/N}\right)N^{2}\xi^{2}}{2}\times\left(\sqrt{1+\dfrac{B^{2}}{\left(1-r^{2/N}\right)^{2}N^{2}\xi^{2}}}-1\right).
+\]
+If the filter bank has to be approximately orthogonal, we typically
+set $B$ to $B=\left(1-2^{-1/Q}\right)\times\xi$.
+
+We conclude with the following closed form for $\alpha$:
+\[
+\alpha=K_{N}\times\sqrt{\dfrac{\sqrt{1+h_{N}\left(Q\right)^{2}}-1}{2}}\times\xi,
+\]
+where 
+\[
+K_{N}=r^{1/N}N\sqrt{1-r^{2/N}}\mbox{ and }h_{N}\left(Q\right)=\dfrac{1-2^{-1/Q}}{N\times\left(1-r^{2/N}\right)}.
+\]
+
 % that's all folks

### 2018-07-17 paper/lostanlenRelevanceScattering.tex
covariant -> equivariant

@@
 exists a phase shift $\varphi\in[0;2\pi[$ such that $\boldsymbol{\psi}(-t)=\mathrm{e}^{\mathrm{i}\varphi}\boldsymbol{\psi}(t)$,
-the wavelet transform is covariant with respect to time reversal:
+the wavelet transform is equivariant with respect to time reversal:
 \begin{eqnarray*}

### 2018-07-17 paper/lostanlenRelevanceScattering.tex
shorten supplementary material

--- 2018-07-10 paper/lostanlenRelevanceScattering.tex
@@
 invariance to time reversal is an undesirable property.
-
-The situation is different in the time-frequency domain. If the wavelet
-$\boldsymbol{\psi}(t)$ has a symmetric envelope, that is, if there
-exists a phase shift $\varphi\in[0;2\pi[$ such that $\boldsymbol{\psi}(-t)=\mathrm{e}^{\mathrm{i}\varphi}\boldsymbol{\psi}(t)$,
-the wavelet transform is covariant with respect to time reversal:
-\begin{eqnarray*}
-\mathbf{U}_{1}(\mathcal{W}_{\boldsymbol{\tau}}\boldsymbol{x})(t,\gamma) & = & \left|\mathrm{e}^{\mathrm{i}\varphi}\int_{-\infty}^{+\infty}\boldsymbol{x}(-t^{\prime})\boldsymbol{\psi}_{2^{\gamma},T(2^{\gamma})}(t^{\prime}-t)\;\mathrm{d}t^{\prime}\right|\\
- & = & \left|\mathrm{e}^{\mathrm{i}\varphi}\right|\times\left|\int_{-\infty}^{+\infty}\boldsymbol{x}(t^{\prime})\boldsymbol{\psi}_{2^{\gamma},T(2^{\gamma})}(-t-t^{\prime})\;\mathrm{d}t^{\prime}\right|\\
- & = & \mathbf{U}_{1}\boldsymbol{x}(t,\gamma).
-\end{eqnarray*}
-Gabor wavelets have a symmetric envelope, as they satisfy $\boldsymbol{\psi}(-t)=-\boldsymbol{\psi}(t)$,
-i.e. $\varphi=\pi$. This is due to the fact that Gaussians have bell-shaped
-symmetric envelopes. Therefore, the Gabor wavelet transform commutes
-with time reversal: 
-\[
-\mathbf{U}_{1}(\mathcal{W}_{\boldsymbol{\tau}}\boldsymbol{x})(t,\gamma)=\mathbf{U}_{1}\boldsymbol{x}(-t,\gamma).
-\]
-A simple way to break this undesirable invariance is to choose $\boldsymbol{\psi}(t)$
+A simple way to break invariance to time reversal is to choose $\boldsymbol{\psi}(t)$
 as an asymmetric wavelet instead of a Gabor symmetric wavelet.

### 2018-07-17 paper/lostanlenRelevanceScattering.tex
suppress negative whitespace in \Delta\omega
\boldsymbol{\psi} Fig 1

@@
 \label{fig:gammatones}
-Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
+Gammatone wavelets $\boldsymbol{\psi}(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
 Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
@@
 where the magnitude cutoff $r$ is most
-often set to $\sqrt{\frac{1}{2}}$. Let $\Delta\!\omega=\omega-\xi$.
+often set to $\sqrt{\frac{1}{2}}$. Let $\Delta\omega=\omega-\xi$.
 Raising the above equation to the power $N/2$ yields the following:
 \[
-\left(1+\dfrac{\Delta\!\omega}{\omega_{c}}\right)^{N/2}=r\times\left(1+\frac{\Delta\!\omega^{2}}{\alpha^{2}}\right).
+\left(1+\dfrac{\Delta\omega}{\omega_{c}}\right)^{N/2}=r\times\left(1+\frac{\Delta\omega^{2}}{\alpha^{2}}\right).
 \]
-Since $\Delta\!\omega\ll\xi$, we may approximate the left-hand side
+Since $\Delta\omega\ll\xi$, we may approximate the left-hand side
 with a first-order Taylor expansion. This leads to a quadratic equation
@@
 \[
-\dfrac{r^{2/N}}{\sigma^{2}}\times\Delta\!\omega^{2}-\dfrac{2}{N\xi}\times\Delta\!\omega+\left(r^{2/N}-1\right)=0.
+\dfrac{r^{2/N}}{\sigma^{2}}\times\Delta\omega^{2}-\dfrac{2}{N\xi}\times\Delta\omega+\left(r^{2/N}-1\right)=0.
 \]

### 2018-07-17 paper/lostanlenRelevanceScattering.tex
rename sainath2015 in bibtex file

@@
 The wavelet scattering transform, described below, has both of these desired properties: invariance to time-shifting and stability to time-warping.
-The stability condition can be formulated as a Lipschitz continuity property, which guarantees that the feature transforms of $\boldsymbol{x}(t)$ and $\boldsymbol{x_\tau}(t)$ are close together if $|\tau'(t)|$ is bounded by a small constant \cite{mallat2012cpam}.
+The stability condition can be formulated as a Lipschitz continuity property, which guarantees that the feature transforms of $\boldsymbol{x}(t)$ and $\boldsymbol{x_\tau}(t)$ are close together if $|\tau'(t)|$ is bounded by a small constant \cite{Mallat2012}.
 

### 2018-07-17 paper/lostanlenRelevanceScattering.tex
re-organize supplementary material

@@
 
-\section{Supplementary material}
+\section{Supplementary material: }
 
+The introductory paper on Gammatone wavelet \cite{Venkitaraman2014} does not provide a formula for deducing $\sigma$ from the specification of a quality factor $Q$.
+
+In this appendix, we provide a rationale for choosing the topmost center frequency $\xi$ of a Gammatone wavelet filter bank in a discrete-time setting.
+Then, we relate the bandwidth parameter $\sigma$ to the choice of a quality factor Q.
+
+\subsubsection*{Motivation}
 Time reversal of a real signal $\boldsymbol{x}(t)$ is equivalent
@@
 
+\subsubsection*{Related work}
+
 With the aim of building a pseudo-analytic admissible Gammatone wavelet,
@@
 compromise between time-frequency localization and causality constraints.
-\cite{Venkitaraman2014} did not provide a formula for deducing $\sigma$
-from the specification of a quality factor $Q$.
-In this appendix, we provide a rationale for choosing the topmost center frequency $\xi$ of a Gammatone wavelet filter bank in a discrete-time setting.
-Then, we relate the bandwidth parameter $\sigma$ to the choice of a quality factor Q.
+
+\subsubsection*{Center frequency parameter}
 

### 2018-07-30 paper/lostanlenRelevanceScattering.tex
whitespace

@@
 The BOF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}.
-In this case, each class of scenes of a given typology, say \{\emph{park, boulevard, square }\}, is modeled by a GMM trained on scenes taken from a training set.
+In this case, each class of scenes of a given typology, say \{\emph{park, boulevard, square}\}, is modeled by a GMM trained on scenes taken from a training set.
 In order to predict the class of a given scene $s$, the likelihood each model given the scene are computed.

### 2018-07-30 paper/lostanlenRelevanceScattering.tex
update intro of Supp material

@@
 
-\section{Supplementary material: }
+\section{Supplementary material}
 
-The introductory paper on Gammatone wavelet \cite{Venkitaraman2014} does not provide a formula for deducing $\sigma$ from the specification of a quality factor $Q$.
-
+The asymmetric amplitude profile of the Gammatone function makes it suitable to model temporal masking in auditory filterbanks.
+Yet, the introductory paper on Gammatone wavelets \cite{Venkitaraman2014} does not provide a formula for deducing $\sigma$ from the specification of a quality factor $Q$.
 In this appendix, we provide a rationale for choosing the topmost center frequency $\xi$ of a Gammatone wavelet filter bank in a discrete-time setting.

### 2018-07-30 paper/lostanlenRelevanceScattering.tex
cite Fastl 2007 in supplementary material
cite Venkitaraman in supplementary material

@@
 
-The asymmetric amplitude profile of the Gammatone function makes it suitable to model temporal masking in auditory filterbanks.
+The asymmetric amplitude profile of Gammatones makes them suitable to model temporal masking in auditory filterbanks~\cite{Fastl2007}.
 Yet, the introductory paper on Gammatone wavelets \cite{Venkitaraman2014} does not provide a formula for deducing $\sigma$ from the specification of a quality factor $Q$.
@@
 point of view, it has been argued that the Gammatone reaches a practical
-compromise between time-frequency localization and causality constraints.
+compromise between time-frequency localization and causality constraints \cite{Venkitaraman2014}.
 

### 2018-07-30 paper/lostanlenRelevanceScattering.tex
capitalization: BOF -> BoF

@@
 Caracterization of the similarity between audio recordings be they at the scale of the minute, the hour, the day or larger is of interest for many applications areas involving acoustic monitoring such as urban sound environment analysis and ecoacoustics.
-In this context, a classical approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
+In this context, a classical approach is the bag of frames (BoF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
 It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.
@@
 The similarity among the scene $s_1$ and $s_2$ is computed as the similarity, see Section~\ref{sec:experiments} for further details.
-The BOF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}.
+The BoF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}.
 In this case, each class of scenes of a given typology, say \{\emph{park, boulevard, square}\}, is modeled by a GMM trained on scenes taken from a training set.
@@
 
-While BOF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
+While BoF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
 Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
@@
 
-From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF models which treat each observation separately and do not capture their temporal structure.
+From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BoF models which treat each observation separately and do not capture their temporal structure.
 A method that takes this knowledge into account could therefore have potential for great impact in acoustic scene modeling, given a rich enough representation of these distinct events.
@@
 The distance on this average feature vector is then used to determine $p@k$.
-For the BOF approach of Aucouturier et al.~\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm~\cite{dempster-laird, moon1996expectation}.
+For the BoF approach of Aucouturier et al.~\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm~\cite{dempster-laird, moon1996expectation}.
 The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation.
@@
 The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.
-As the number of Gaussians for the BOF approach, the numbers of clusters $M$ controls the level of abstraction.
+As the number of Gaussians for the BoF approach, the numbers of clusters $M$ controls the level of abstraction.
 For each method, the parameter $M$ is set to $8$.
@@
 Results for the acoustic scene similarity retrieval task demonstrate that logarithmically compressed scattering features outerform MFCCs.
-Combining these with the \emph{RbQ} cluster model, improvements are obtained over traditional BOF and summary statistic measures.
+Combining these with the \emph{RbQ} cluster model, improvements are obtained over traditional BoF and summary statistic measures.
 
@@
 
-As seen in Figure~\ref{fig:ASS_-1}, scattering features significantly outperform the baseline MFCCs for both BOF and early integration schemes.
+As seen in Figure~\ref{fig:ASS_-1}, scattering features significantly outperform the baseline MFCCs for both BoF and early integration schemes.
 This is expected, as the scattering transform extends the MFCCs by including supplementary amplitude modulation information~\cite{Anden2014}.
@@
 
-Comparing the BOF to early integration, both approaches perform similarly for MFCCs and PCA-reduced scattering features alike.
-This is in line with previous results on BOF~\cite{lagrange:hal-01082501}, where it is found to perform similarly as similarity retrieval on features averaged over the entire recording.
+Comparing the BoF to early integration, both approaches perform similarly for MFCCs and PCA-reduced scattering features alike.
+This is in line with previous results on BoF~\cite{lagrange:hal-01082501}, where it is found to perform similarly as similarity retrieval on features averaged over the entire recording.
 
@@
 To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects.
-The same behavior is observed for MFCCs, with \emph{RbQ-c} and \emph{RbQ-w} outperforming \emph{early}, which is equivalent to the state-of-the-art BOF model, as seen previously.
+The same behavior is observed for MFCCs, with \emph{RbQ-c} and \emph{RbQ-w} outperforming \emph{early}, which is equivalent to the state-of-the-art BoF model, as seen previously.
 
@@
 This paper presents a new approach for modeling acoustic scenes based on scattering transforms at small scales and cluster-based representations at large scales.
-Compared to traditional BOF and summary statistics models, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature.
+Compared to traditional BoF and summary statistics models, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature.
 To adequately capture such distinct events, we develop a cluster-based model and validate it using experiments on acoustic scene similarity retrieval.
-For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
+For this task, we show significant improvements over the traditional BoF and summary statistics models based on both standard MFCCs and scattering features.
 These outcomes shall be studied further in future work by considering larger databases and emerging tasks in ecoacoustics~\cite{sueur2015ecoacoustics, wimmer2013sampling}.

### 2018-07-30 paper/lostanlenRelevanceScattering.tex
improve description of Eq 6

@@
 \end{equation}
-Here, $\mu_{m,q}^u$ and $\mu_{n,q}^v$ are the $q^{\textrm{th}}$ nearest neighbors to the centroids $\mu_m^u$ and $\mu_n^v$, respectively, and $\Vert \cdot \Vert$ denotes the Euclidean norm.
+Here, $\mu_{m,q}^u$ and $\mu_{n,q}^v$ are the $q^{\textrm{th}}$ nearest neighbors to the centroids $\mu_m^u$ and $\mu_n^v$, respectively, and the double bars $\Vert \cdot \Vert$ denotes the Euclidean norm.
 

### 2018-07-30 paper/lostanlenRelevanceScattering.tex
list DCASE classes

@@
 The BoF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}.
-In this case, each class of scenes of a given typology, say \{\emph{park, boulevard, square}\}, is modeled by a GMM trained on scenes taken from a training set.
+In this case, each class of scenes from a given typology, say \{\emph{park, boulevard, square}\}, is modeled by a GMM trained on scenes taken from a training set.
 In order to predict the class of a given scene $s$, the likelihood each model given the scene are computed.
-The scene is then labeled park if the likelihood of the GMM trained on park scenes is higher than the other likelihoods.
+The scene is then labeled park if the likelihood of the GMM trained on park scenes is higher than all other likelihoods.
 
@@
 The dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
-The dataset is evenly divided into $10$ acoustic scene classes.
+The dataset is evenly divided into $10$ acoustic scene classes: \emph{bus}, \emph{busy street}, \emph{office}, \emph{open air market}, \emph{park}, \emph{quiet street}, \emph{restaurant}, \emph{supermarket}, \emph{tube}, and \emph{tube station}.
 The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
-No
-systematic variations in the recordings covaried with scene
+No systematic variations in the recordings covaried with scene
 type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.

### 2018-07-30 paper/lostanlenRelevanceScattering.tex
rewrite one paragraph in dataset section

@@
 The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
-No systematic variations in the recordings covaried with scene
-type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
+In order to avoid any correlation between recording conditions and label distribution, all recordings were carried out under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
 As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices~\cite{lagrange:hal-01082501}.

### 2018-08-12 paper/lostanlenRelevanceScattering.tex
Add sensitivity analysis figure and subsection

@@
 It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds).
+Note that this is the only free parameter in the proposed method.
+However, except for $\emph{RbQ-a}$, the results are not very sensitive to the choice of $M$, as long as it is large enough to characterize the seen.
+For a numerical demonstration of this, please see the end of the next section.
 
@@
 \label{fig:ASS_1}
+\end{center}
+\end{figure}
+
+\subsubsection*{Sensitivity to number of clusters $M$}
+
+We now study the sensitivity of the precision at rank $5$ ($p@5$) with respect to the number of clusters $M$.
+The results are shown in Figure \ref{fig:sensitivityM}.
+
+For a small number of clusters ($M = 1$ or $M = 2$), all methods perform worse, since not enough discriminative sound objects are extracted from the recording.
+This corresponds a summary statistics model.
+For $M = 4$, most methods perform well, since this allows for better characterization of various signal structures in the scenes.
+As the number of clusters increases, the \emph{RbQ-a} method performs significantly worse for both scattering features and MFCCs since any distinct objects are averaged out by clusters representing the background.
+The \emph{RbQ-w} and \emph{RbQ-c} methods do better in this regard, as they are better able to emphasize the clusters that discriminate well between scenes.
+
+Using \emph{RbQ-c}, therefore, we are not very sensitive to the choice of $M$, as long as it is large enough to allow for separation of the discriminative sound objects from the background.
+This motivates our choice of $M = 8$ for the previous experiments in this section.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/sensitivityM}
+\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precision at rank $5$ ($p@5$) obtained for different methods as a function of the number of clusters $M$.}
+\label{fig:sensitivityM}
 \end{center}

### 2018-08-13 paper/revision/responseToReviewers.tex
rename "Algorithm" subsection
check Gammatone item in to-do list
updarte response to reviewer

@@
 
-$\rightarrow$ VL
+$\rightarrow$ VL $\checkmark$
 
@@
 
-$\rightarrow$ A discussion about the setting of the parameter M is added at the end of Section 6.3.
+$\rightarrow$ The number of clusters $M$ is indeed an important parameter.
+The higher $M$ is, the more complex the model and the computation are.
+Setting $M$ to 4 roughly states that there may be at most 4 different types of events in the given chunk of audio, here 30 sec.
+The analysis we added to the manuscript at the end of Section 7.
 

### 2018-08-13 paper/lostanlenRelevanceScattering.tex
updarte response to reviewer

@@
 As the number of Gaussians for the BoF approach, the numbers of clusters $M$ controls the level of abstraction.
-For each method, the parameter $M$ is set to $8$.
-It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds).
+For each method, unless otherwise stated, the parameter $M$ is set to $8$.
+It thus allows $8$ different types of observations to be modeled, which seems reasonable given the duration of the scene (30 seconds).
 Note that this is the only free parameter in the proposed method.
@@
 For a small number of clusters ($M = 1$ or $M = 2$), all methods perform worse, since not enough discriminative sound objects are extracted from the recording.
-This corresponds a summary statistics model.
+Please note that setting $M = 1$ is equivalent to the \emph{early} approach as this corresponds a summary statistics model.
 For $M = 4$, most methods perform well, since this allows for better characterization of various signal structures in the scenes.
-As the number of clusters increases, the \emph{RbQ-a} method performs significantly worse for both scattering features and MFCCs since any distinct objects are averaged out by clusters representing the background.
+As the number of clusters increases, the \emph{RbQ-a} method performs worse for both scattering features and MFCCs since any distinct objects are averaged out by clusters representing the background.
 The \emph{RbQ-w} and \emph{RbQ-c} methods do better in this regard, as they are better able to emphasize the clusters that discriminate well between scenes.

### 2018-08-13 paper/lostanlenRelevanceScattering.tex
typos fixes

@@
 In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas~\cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic~\cite{gloaguen}.
-Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for to signal processing community.
+Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for signal processing community.
 
@@
 A widespread choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
-Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events.
+Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events sparsely distributed over time.
 Furthermore, experiments in cognitive psychology~\cite{dubois2006cognitive} and cognitive neuroscience~\cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events.
@@
 
-For our task, one advantage of the scattering transform is not requiring a training step, allowing for a wider range of applications compared to learned features.
+For our task, one advantage of the scattering transform is that it does not require a training step, allowing for a wider range of applications compared to learned features.
 Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unsupervised setting.
@@
 While BoF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
-Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
+Though, this representation was recently shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
 This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background~\cite{nelken2013}.
@@
 
-Many algorithms in pattern recognition, including nearest neighbor classifiers and SVMs, tend to work best when all features follow a standard normal distribution across all training instances~\cite{Hsu2003}.
+Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances~\cite{Hsu2003}.
 Yet the distribution of the scattering coefficients is skewed towards larger values.
@@
 
-The experiments in this paper are carried out on the DCASE 2013 dataset~\cite{7100934}.
+The experiments in this paper are carried out on the publicly available DCASE 2013 dataset~\cite{7100934}.
 Although the dataset was constructed for the task of acoustic scene classification, where the goal is to correctly assign the class of a given recording, we can use the same recordings and class labels for the task of acoustic scene similarity retrieval.
@@
 However, except for $\emph{RbQ-a}$, the results are not very sensitive to the choice of $M$, as long as it is large enough to characterize the seen.
-For a numerical demonstration of this, please see the end of the next section.
+A numerical demonstration is provided at the end of the next section.
 

### 2018-08-17 paper/lostanlenRelevanceScattering.tex
add park_scalogram, include it in paper

--- 2018-08-12 paper/lostanlenRelevanceScattering.tex
@@
 As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices~\cite{lagrange:hal-01082501}.
+As an illustration, Figure \ref{fig:scalogram} represents the wavelet scalogram of one recording within the DCASE 2013, labeled as \emph{park}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/park_scalogram.png}
+\caption{
+\label{fig:scalogram}
+Wavelet scalogram $\boldsymbol{x}_1(t,\gamma_1)$ of the audio recording \emph{park04} in the DCASE 2013 dataset.
+We observe that this acoustic scene is a mixture of transient events (chirping birds, footsteps) and stationary texture (flowing water).}
+\end{center}
+\end{figure}
 
@@
 It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds).
-Note that this is the only free parameter in the proposed method.
-However, except for $\emph{RbQ-a}$, the results are not very sensitive to the choice of $M$, as long as it is large enough to characterize the seen.
-For a numerical demonstration of this, please see the end of the next section.
 
@@
 \label{fig:ASS_1}
-\end{center}
-\end{figure}
-
-\subsubsection*{Sensitivity to number of clusters $M$}
-
-We now study the sensitivity of the precision at rank $5$ ($p@5$) with respect to the number of clusters $M$.
-The results are shown in Figure \ref{fig:sensitivityM}.
-
-For a small number of clusters ($M = 1$ or $M = 2$), all methods perform worse, since not enough discriminative sound objects are extracted from the recording.
-This corresponds a summary statistics model.
-For $M = 4$, most methods perform well, since this allows for better characterization of various signal structures in the scenes.
-As the number of clusters increases, the \emph{RbQ-a} method performs significantly worse for both scattering features and MFCCs since any distinct objects are averaged out by clusters representing the background.
-The \emph{RbQ-w} and \emph{RbQ-c} methods do better in this regard, as they are better able to emphasize the clusters that discriminate well between scenes.
-
-Using \emph{RbQ-c}, therefore, we are not very sensitive to the choice of $M$, as long as it is large enough to allow for separation of the discriminative sound objects from the background.
-This motivates our choice of $M = 8$ for the previous experiments in this section.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/sensitivityM}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precision at rank $5$ ($p@5$) obtained for different methods as a function of the number of clusters $M$.}
-\label{fig:sensitivityM}
 \end{center}

### 2018-08-17 figs/spectrogram/park_scalogram.m

@@ -1,0 +1,53 @@
+
+
+%%
+opts{1}.time.size = N;
+opts{1}.time.T = 2^10;
+opts{1}.time.max_scale = 8*opts{1}.time.T;
+opts{1}.time.nFilters_per_octave = 24;
+opts{1}.time.gamma_bounds = ...
+    [1+opts{1}.time.nFilters_per_octave*1 ...
+    opts{1}.time.nFilters_per_octave*10];
+opts{1}.time.max_Q = 12;
+opts{1}.time.is_chunked = false;
+
+archs = sc_setup(opts);
+
+
+%%
+addpath(genpath('~/scattering.m'));
+data_dir = '~/datasets/scenes_stereo_testset';
+wav_id = 4;
+wav_name = ['park', sprintf('%02d', wav_id), '.wav'];
+wav_path = [data_dir, '/', wav_name];
+disp(wav_path);
+
+
+[y, sample_rate] = audioread(wav_path);
+y = 0.5 * (y(:, 1) + y(:, 2));
+N = 2^20;
+y = y(1:2^20);
+
+S = sc_propagate(y, archs);
+
+%%
+scalogram = S{1+1}.data.';
+
+x_duration = opts{1}.time.size / sample_rate;
+freq_hz = [50, 100, 200, 500, 1000, 2000, 5000, 10000, 20000];
+xi = archs{1}.banks{1}.spec.mother_xi * sample_rate;
+freq_semitones = 1 + ...
+    round(log2(xi ./ freq_hz) * opts{1}.time.nFilters_per_octave);
+freq_semitones = sort(unique(freq_semitones));
+
+imagesc([0.0, x_duration], opts{1}.time.gamma_bounds, ...
+    100 * flipud(log1p(1e1*scalogram)));
+colormap rev_magma;
+set(gca, 'YDir', 'normal');
+set(gca(), 'Xtick', [0:5:20.0]);
+set(gca(), 'Ytick', sort(opts{1}.time.gamma_bounds(2) - freq_semitones));
+set(gca(), 'YTickLabel', freq_hz/1000);
+xlabel('Time (s)');
+
+export_fig park_scalogram.png -m8 -transparent
+ylabel('Frequency (kHz)');

### 2018-08-17 figs/spectrogram/table_predictions.m

@@ -1,0 +1,111 @@
+system_strs = { ...
+    'cu1_feMf_ne5_prLm_reMofa_sc25_st1', ...
+    'co1_feSc_me1_ne5_prLm_reMofa_sc1000_st1'
+};
+
+queries = { ...
+    'Violin/ordinario/Vn-ord-G4-mf-4c', ...
+    'Trumpet-C/flatterzunge/TpC-flatt-G4-mf'};
+n_queries = length(queries);
+
+data_dir = '~/Downloads/SOL_0.9_HQ';
+
+addpath(genpath('~/scattering.m'));
+clear opts;
+opts{1}.time.size = 65536;
+opts{1}.time.T = 2^10;
+opts{1}.time.max_scale = 2*opts{1}.time.T;
+opts{1}.time.nFilters_per_octave = 48;
+opts{1}.time.gamma_bounds = ...
+    [1+opts{1}.time.nFilters_per_octave*1 ...
+    opts{1}.time.nFilters_per_octave*7];
+opts{1}.time.max_Q = opts{1}.time.nFilters_per_octave / 2;
+opts{1}.time.is_chunked = false;
+
+archs = sc_setup(opts);
+
+
+n_systems = length(system_strs);
+systems = cell(1, n_systems);
+clc();
+
+for q_id = 1:n_queries
+    query_str = queries{q_id};
+    
+    for system_id = 1:n_systems
+        system_str = system_strs{system_id};
+        disp(system_str)
+        disp([' Q: ', query_str]);
+        disp(' ');
+
+        systems{system_id} = load(system_str);
+        system = systems{system_id};
+        
+        query_id = ...
+            find(cellfun(@(x) strcmp(x, query_str), system.d.file));
+        neighbor_ids = system.d.nn(query_id,:);
+        
+        for i = 0:length(neighbor_ids)
+            if i > 0
+                neighbor_id = neighbor_ids(i);
+                file_name = system.d.file{neighbor_id};
+                disp([' ', num2str(i),': ', file_name]);
+            else
+                file_name = query_str;
+            end  
+            
+            neighbor_split = strsplit(file_name, '/');
+            instrument_str = neighbor_split{1};
+            
+            switch instrument_str
+                case 'Trumpet-C'
+                    family_str = 'Trumpets';
+                case 'Trumpet-C-straight'
+                    family_str = 'Trumpets';
+                case 'Violin'
+                    family_str = 'Strings';
+                case 'Violin-sordina'
+                    family_str = 'Strings';
[…]

### 2018-08-17 paper/revision/responseToReviewers.tex
add park_scalogram, include it in paper
reply to reviewers regarding scalogram
respond to reviewers re: gammatones

@@
 
-$\rightarrow$ VL
+$\rightarrow$ We have listed the classes from the DCASE taxonomy in full in Section 6.1. In addition, we have included the scalogram of a labeled recording (\emph{park04}} as Figure 3.
 
@@
 
-$\rightarrow$ VL $\checkmark$
+$\rightarrow$ VL $\checkmark$ We have not re-run experiments with Morlet wavelets instead of Gammatone wavelets, for two reasons.
+First, Morlet wavelets do not fit the whole narrative of the paper as well as Gammatone wavelets: while the arguments in favor of Morlet wavelets arise from harmonic analysis (Heisenberg time-frequency uncertainty tradeoff), those in favor of Gammatone wavelets arise from auditory neurophysiology.
+Because we work in the context of computational auditory scene analysis, \emph{a fortiori} in an unsupervised setting, we believe that it is preferable to borrow from a well-studied auditory model (Gammatone) rather than from a general-purpose operator (Morlet).
+The second reason why we did not conduct a systematic benchmark between Morlet and Gammatone is that, in previous experiments involving supervised classification of musical instruments with scattering coefficients, Gammatone wavelets appeared to outperform Morlet wavelets by a small margin, \ie{} from $80\%$ and $79\%$ test accuracy respectively.
+In all likelihood, the effect of the specific choice of wavelet is also small in the case of unsupervised similarity retrieval between auditory scenes, and possibly not statistically significant given the limited sample size of the DCASE 2013 dataset.
+We have expanded Section 3.4 to highlight these considerations.
+In addition, we have included a mathematical description of the Gammatone wavelet filter bank employed in this study as supplementary material.
+Indeed, the design of our Gammatone wavelet is slightly different than in the reference papers of Venkitaraman and Adiga, because, unlike them, we need to account for the influence of both the quality factor $Q$ and the polynomial degree parameter $N$.
+It is our hope that the proof of the formula linking $Q$ and $N$ to the parameter $\alpha$ in the Gammatone function will be beneficial to the signal processing community at large.
 

### 2018-08-17 paper/revision/responseToReviewers.tex
fix compilation issues

@@
 
-$\rightarrow$ We have listed the classes from the DCASE taxonomy in full in Section 6.1. In addition, we have included the scalogram of a labeled recording (\emph{park04}} as Figure 3.
+$\rightarrow$ We have listed the classes from the DCASE taxonomy in full in Section 6.1. In addition, we have included the scalogram of a labeled recording (\emph{park04}) as Figure 3.
 
@@
 Because we work in the context of computational auditory scene analysis, \emph{a fortiori} in an unsupervised setting, we believe that it is preferable to borrow from a well-studied auditory model (Gammatone) rather than from a general-purpose operator (Morlet).
-The second reason why we did not conduct a systematic benchmark between Morlet and Gammatone is that, in previous experiments involving supervised classification of musical instruments with scattering coefficients, Gammatone wavelets appeared to outperform Morlet wavelets by a small margin, \ie{} from $80\%$ and $79\%$ test accuracy respectively.
+The second reason why we did not conduct a systematic benchmark between Morlet and Gammatone is that, in previous experiments involving supervised classification of musical instruments with scattering coefficients, Gammatone wavelets appeared to outperform Morlet wavelets by a small margin, from $80\%$ and $79\%$ test accuracy respectively.
 In all likelihood, the effect of the specific choice of wavelet is also small in the case of unsupervised similarity retrieval between auditory scenes, and possibly not statistically significant given the limited sample size of the DCASE 2013 dataset.

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
sup material

--- 2018-08-17 paper/lostanlenRelevanceScattering.tex
@@
 %
-  \usepackage[pdftex]{graphicx}
+%\usepackage[pdftex]{graphicx}
 %
@@
 
-\usepackage[backend=bibtex,style=numeric-comp,sorting=none,firstinits=true]{biblatex}
-\bibliography{biblio}
+%\usepackage[backend=bibtex,style=numeric-comp,sorting=none,firstinits=true]{biblatex}
+%%\bibliography{biblio}
 %\renewcommand{\baselinestretch}{2}
 
+\graphicspath{{./}{figures/}}
+
 \begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-% Do not put math or special symbols in the title.
-%\title{An evaluation framework for event detection using a morphological model of acoustic scenes}
-\title{Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio}
-\titlerunning{Relevance-based Quantization of Scattering Features}
-
-\author{Vincent Lostanlen         \and
-		Gr\'egoire Lafay  \and Joakim And\'en  \and Mathieu Lagrange.}
-
-\institute{Vincent Lostanlen \at
-              \email{vincent.lostanlen@nyu.edu}
-\and
-			Gr\'egoire Lafay \at
-              \email{lafaygregoire@gmail.com}
-\and
-			Joakim And\'en \at
-              \email{janden@flatironinstitute.org}           \and
-			Mathieu Lagrange \at
-              \email{mathieu.lagrange@cnrs.fr}
-}
-
-\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
-
-% make the title area
-\maketitle
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
-\begin{abstract}
-The emerging field of computational acoustic monitoring aims at retrieving high-level information from acoustic scenes recorded by some network of sensors.
-These networks gather large amounts of data requiring analysis.
-To decide which parts to inspect further, we need tools that automatically mine the data, identifying recurring patterns and isolated events.
-This requires a similarity measure for acoustic scenes that does not impose strong assumptions on the data.
-
-The state of the art in audio similarity measurement is the ``bag-of-frames'' approach, which models a recording using summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
-They successfully characterise static scenes with little variability in auditory content, but cannot accurately capture scenes with a few salient events superimposed over static background.
-To overcome this issue, we propose a two-scale representation which describes a recording using clusters of scattering coefficients.
-The scattering coefficients capture short-scale structure, while the cluster model captures longer time scales, allowing for more accurate characterization of sparse events.
-Evaluation within the acoustic scene similarity framework demonstrates the interest of the proposed approach.
-\keywords{
-unsupervised learning \and data mining \and acoustic signal processing \and wavelet transforms \and audio databases \and content-based retrieval \and nearest neighbor searches \and acoustic sensors \and environmental sensors.}
-% \PACS{PACS code1 \and PACS code2 \and more}
-\end{abstract}
-
-\section{Introduction}
-
-The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
-In order to measure the effect of human activity and climate change on animal biodiversity~\cite{sueur2015ecoacoustics}, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world~\cite{warren2006urban, NessSST13, stowell13b}.
-In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas~\cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic~\cite{gloaguen}.
-Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for to signal processing community.
-
-%These questions are explored in an emerging field of computational bioacoustics, known as ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science .
-%~\cite{wimmer2013sampling}
-
-An important problem is that manually analysing the recorded data to identify the quantities of interest is very costly.
-Some sort of pre-screening is therefore required to reduce the need for human expert listening and annotation.
-To this aim, the most straightforward approach is to specify a closed set of sound classes, such as sounds classes expected to appear near the acoustic sensors.
-Computational models are then trained  for these classes which are used to automatically annotate recordings~\cite{quteprints96243}.
-A given time interval (\eg{} a single day) is then represented by the number of events detected during that interval for each class.
-This allows the scientist to drastically reduce the amount of information requiring manual processing.
-However, this approach has two drawbacks.
-First, it relies on trained models whose prediction on unseen data -- \ie{} sensors outside of the training set -- is prone to errors.
-Secondly, and more importantly, it is based on prior knowledge and thus cannot be considered for exploratory analysis, in which quantities of interest have yet to be defined.
-
-To identify which parts need human inspection, one needs tools that are able to detect both recurring patterns and sparsely distributed events.
-Identifying recurring patterns allows the user to focus on certain time points for manual annotation, while detection of more rare structures enables discovery of unforeseen phenomena.
-
-With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset.
-To construct such an algorithm, we are required to represent an audio recording in a way that captures its distinctive qualities.
-A widespread choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features.
-Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events.
-Furthermore, experiments in cognitive psychology~\cite{dubois2006cognitive} and cognitive neuroscience~\cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events.
-We believe that the failure to model such distinct events is one of the reasons why the bag-of-frames representation is insufficient~\cite{lagrange:hal-01082501}.
-
-Solving the acoustic similarity retrieval first requires the ability to capture meaningful signal structure at small time scales.
-This is often achieved using mel-frequency cepstral coefficients (MFCCs).
-Originally developed for speech processing~\cite{davis-mermelstein}, MFCCs have recently found wider use in music information retrieval~\cite{logan} and environmental audio processing~\cite{aucouturier2007bag}.
-A richer representation, the scattering transform, has enjoyed significant success in various audio~\cite{Anden2014} and biomedical~\cite{chudacek} signal classification tasks.
-Its structure is that of a convolutional neural network~\cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}, but with fixed filters.
-Specifically, it alternates convolutions with wavelet filters and pointwise nonlinearities to ensure time-shift invariance and time-warping stability~\cite{Mallat2012}.
-
-For our task, one advantage of the scattering transform is not requiring a training step, allowing for a wider range of applications compared to learned features.
-Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unsupervised setting.
-
-In this work, we propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by a cluster model.
-This unsupervised model quantizes the scattering coefficients into a given number of clusters.
-These clusters are then used to define a set of distances for acoustic similarity retrieval.
-Evaluating this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.
-
-Motivations of the proposed approach and a brief review of the state of the art in acoustic scene modeling are given in Section~\ref{sec:soa}.
-We describe the scattering transform in Section~\ref{sec:scattering}, discuss feature post-processing in Section~\ref{sec:design} and propose a cluster-based scene description in Section~\ref{sec:object}.
-Section~\ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task.
-Results are reported in Section~\ref{sec:results}.
-
-\section{Background} \label{sec:soa}
-
-%Computational bioacoustics refers to the numerical investigation of sound production, dispersion and reception in animals, including humans.
-A recent paradigm processes the audio stream in a holistic way over large time scales, without assuming that a single species is present throughout the recording.
-Automated systems within this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices~\cite{Bardeli2010} and migration patterns~\cite{Obrist2010}.
-They also mark time intervals of particular interest for detailed human inspection~\cite{rosenstock2002landbird}.
-
-Caracterization of the similarity between audio recordings be they at the scale of the minute, the hour, the day or larger is of interest for many applications areas involving acoustic monitoring such as urban sound environment analysis and ecoacoustics.
-In this context, a classical approach is the bag of frames (BoF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
-It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.
-
-It is worth mentionning that this task typically falls into a unsupervised paradigm where no prior knowldge is used to model a given scene.
-For each scene $s$, a model is compute $M_s$ is computed.
-The similarity among the scene $s_1$ and $s_2$ is computed as the similarity, see Section~\ref{sec:experiments} for further details.
-The BoF approach is also widely used in a supervised fashion for solving a classification task~\cite{7100934}.
-In this case, each class of scenes from a given typology, say \{\emph{park, boulevard, square}\}, is modeled by a GMM trained on scenes taken from a training set.
-In order to predict the class of a given scene $s$, the likelihood each model given the scene are computed.
-The scene is then labeled park if the likelihood of the GMM trained on park scenes is higher than all other likelihoods.
-
-While BoF has largely been superseded by more sophisticated methods for the task of acoustic scene classification~\cite{7100934}, it remains the best-performing model for acoustic scene similarity retrieval.
-Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of datasets~\cite{lagrange:hal-01082501}.
-This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background~\cite{nelken2013}.
-Such events are not well-characterised by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
-The latter approach should therefore prove more fruitful in measuring auditory scene similarity.
-
-This statement has some support in auditory psychology as well as sound synthesis based on summary statistics~\cite{mcdermott2013summary}.
-Studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterise an acoustic scene~\cite{guyot2005urban,kang2006urban}.
-Instead, cognitive processes such as sound environment quality perception~\cite{dubois2006cognitive} or loudness judgment~\cite{kuwano_memory_2003} seem to rely upon higher-level cognitive attributes.
-These typically include the identities of the sound sources which constitute the scene.
-It has been shown that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes.
-For example, in urban areas the presence of birds is likely to be heard in parks and are therefore strong pleasantness indicators.
-Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments~\cite{ricciardi2015sound,lavandier2006contribution}.
-Although the complete set of events occurring within a given auditory stream may not be discernable even to human expects, research has shown that a small set of events (so-called markers) suffice to reliably predict many high-level attributes.
-
-From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BoF models which treat each observation separately and do not capture their temporal structure.
-A method that takes this knowledge into account could therefore have potential for great impact in acoustic scene modeling, given a rich enough representation of these distinct events.
-
-\section{Wavelet scattering \label{sec:scattering}}
-
-Local invariance to time-shifting and stability to time-warping are necessary when representing acoustic scenes for similarity measurement.
-The scattering transform is designed to satisfy these properties while retaining high discriminative power.
-It is computed by applying auditory and modulation wavelet filter banks alternated with complex modulus nonlinearities.
-
-\subsection{Invariance and stability in audio signals}
-The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval.
-Indeed, recordings may be shifted locally in time without affecting similarity to other recordings.
-To discard this superfluous source of variability, signals are first mapped into a time-shift invariant feature space.
-These features are then used to calculate similarities.
-Since the features ensure invariance, it does not have to be learned when constructing the desired similarity measure.
-
-Formally, given a signal $\boldsymbol{x}(t)$, we would like its translation $\boldsymbol{x_c}(t) = \boldsymbol{x}(t-c)$ to be mapped to the same feature vector provided that $|c| \ll T$ for some maximum duration $T$ that specifies the extent of the time-shifting invariance.
-We can also define more complicated transformations by letting $c$ vary with $t$.
-In this case, we have $\boldsymbol{x_\tau}(t) = \boldsymbol{x}(t-\tau(t))$ for some function $\tau$, which performs a time-warping of $\boldsymbol{x}(t)$ to obtain $\boldsymbol{x_\tau}(t)$.
-Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events.
-These make up an important part of intra-class variability among natural sounds, so representations must be robust with respect to such transformations.
-
-The wavelet scattering transform, described below, has both of these desired properties: invariance to time-shifting and stability to time-warping.
-The stability condition can be formulated as a Lipschitz continuity property, which guarantees that the feature transforms of $\boldsymbol{x}(t)$ and $\boldsymbol{x_\tau}(t)$ are close together if $|\tau'(t)|$ is bounded by a small constant \cite{Mallat2012}.
-
-\subsection{Wavelet scalogram}
-Our convention for the Fourier transform of a continuous-time signal $\boldsymbol{x}(t)$ is $\boldsymbol{\hat{x}}(\omega) = \int_{-\infty}^{+\infty} x(t) \exp(- \mathrm{i} 2\pi \omega t) \, \mathrm{d}t$.
-Let $\boldsymbol{\psi}(t)$ a complex-valued analytic bandpass filter of
-central frequency $\xi_1$ and bandwidth $\xi_1/Q_1$, where $Q_1$ is the quality factor of the filter.
-A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
-according to a geometric sequence of scales $2^{\gamma_1/Q_1}$, obtaining
-\begin{equation}
-\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(2^{-\gamma_1/Q_1} t)\mbox{.}
-\end{equation}
-The variable $\gamma_1$ is a scale (an inverse log-frequency) taking integer values between $0$ and $(J_1 Q_1 - 1)$, where $J_1$ is the number of octaves spanned by the filter bank.
-For each $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
-has a central frequency of $2^{-\gamma_1/Q_1}\xi_1$ and a bandwidth of $2^{-\gamma_1/Q_1}\xi_1/Q_1$ resulting in the same quality factor $Q_1$ as $\boldsymbol{\psi}$.
-In the following, we set $\xi_1$ to $20~\mathrm{kHz}$, $J_1$ to $10$, and the quality factor $Q_1$, which is also the number of wavelets per octave, to $8$.
-This results in the wavelet filters covering the whole range of human hearing, from $20~\mathrm{Hz}$ to $20~\mathrm{kHz}$.
-Setting $Q_1 = 8$ results in filters whose bandwidth approximates an equivalent rectangular bandwidth (ERB) scale~\cite{Fastl2007}.
-
-The wavelet transform of an audio signal
-$\boldsymbol{x}(t)$ is obtained by convolution with all wavelet filters.
-Applying a pointwise complex modulus the transform yields
-the wavelet scalogram
-\begin{equation}
-\boldsymbol{x_1}(t, \gamma_1)
-= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t)\mbox{.}
-\end{equation}
-The scalogram bears resemblance to the constant-Q transform (CQT),
-which is derived from the short-term Fourier transform (STFT) by averaging the frequency
-axis into constant-Q subbands of central frequencies $2^{-\gamma_1/Q_1}\xi_1$.
-Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
-However, contrary to the CQT, the scalogram reaches a better time-frequency localization across the whole
-frequency range, whereas the temporal resolution of the traditional CQT is fixed by the support of the STFT analyzing window. %\cite{Brown1992}
-Therefore, the scalogram has a better temporal localization at high
-frequencies than the CQT, at the expense of a greater computational cost
-since the inverse fast Fourier transform routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
-However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.
-
-To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\boldsymbol{\phi}(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
-\begin{equation}
-\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \boldsymbol{\phi}(t),
-\end{equation}
-which is known as the set of first-order scattering coefficients.
-They capture the average spectral envelope of $\boldsymbol{x}(t)$ over scales of duration $T$ and where the spectral resolution varying with constant Q.
-In this way, they are closely related to the mel-frequency spectrogram and related features, such as MFCCs.
-
-\subsection{Extracting modulations with second-order scattering}
-In auditory scenes, short-time amplitude modulations may be caused by a variety of rapid mechanical interactions, including collision, friction, turbulent flow, and so on.
-At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
-Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into the time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered from $\boldsymbol{x_1}(t,\gamma_1)$ by a second wavelet transform and another complex modulus.
-
-We define second-order wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ in the same way as the first-order wavelets, but with parameters $\xi_2$, $J_2$, and $Q_2$.
-Consequently, they have central frequencies $2^{-\gamma_2/Q_2}\xi_2$ for $\gamma_2$ taking values between $0$ and $(J_2 Q_2 - 1)$.
-While this abuses notation slightly, the identity of the wavelets should be clear from context.
-The amplitude modulation spectrum resulting from a wavelet modulus decomposition using these second-order wavelets is then
-\begin{equation}
-\boldsymbol{x_2}(t,\gamma_1,\gamma_2) =
-\vert \boldsymbol{x_1} \ast \boldsymbol{\psi_{\gamma_2}} \vert(t,\gamma_1).
-\end{equation}
-In the following, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$.
-Lastly, the low-pass filter $\boldsymbol{\phi}(t)$ is applied to $\boldsymbol{x_2}(t, \gamma_1, \gamma_2)$ to guarantee local invariance to time-shifting, which yields the second-order scattering coefficients
-\begin{equation}
-\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
-\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \boldsymbol{\phi}(t).
-\end{equation}
-
-The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ denotes either $\gamma_1$ or $(\gamma_1,\gamma_2)$.
-While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
-Indeed, higher-order scattering coefficients have been shown to contain reduced energy and are therefore of limited use~\cite{irene}.
-
-\subsection{Gammatone wavelets}
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/gammatones}
-\caption{
-\label{fig:gammatones}
-Gammatone wavelets $\boldsymbol{\psi}(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
-Oscillations (red, blue) are the real and imaginary parts. The envelope (yellow) is the complex modulus.}
-\end{center}
-\end{figure}
-Wavelets
-$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
-wavelets with one vanishing moment~\cite{Venkitaraman2014}, and are shown in Figure~\ref{fig:gammatones}.
-In the context of auditory scene analysis, the asymmetric envelopes of Gammatone wavelets are more biologically plausible than the symmetric, Gaussian envelopes of the more widely used Morlet wavelets.
-Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking~\cite{Fastl2007}.
-The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
-Likewise, because critical bands are skewed towards higher frequencies, a masking tone has to be louder if it is above the stimulus in frequency rather than below.
-It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
-As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds~\cite{Smith2006,sainath2015interspeech}.
-In addition, Gammatone wavelets have proven to outperform Morlet wavelets on a benchmark of supervised musical instrument classification from scattering coefficients \cite{lostanlen2017phd}.
-This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.
-
-\section{Feature design}
-\label{sec:design}
-
-Before constructing models for similarity estimation, it is beneficial to process scattering coefficients to improve invariance, normality, and generalization power.
-In this section, we review two transformations which achieve these properties: logarithmic compression and standardisation.
-
-\subsection{Logarithmic compression}
-\label{sec:logcomp}
-
-Many algorithms in pattern recognition, including nearest neighbor classifiers and SVMs, tend to work best when all features follow a standard normal distribution across all training instances~\cite{Hsu2003}.
-Yet the distribution of the scattering coefficients is skewed towards larger values.
-We can reduce this skewness by applying a pointwise concave transformation to all coefficients.
-In particular, we find that the logarithm performs particularly well in this respect.
-Figure~\ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/compression}
-\caption{
-\label{fig:histograms}
-Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a central acoustic frequency of $302\,\mathrm{Hz}$,
-(a) before and (b) after logarithmic compression.}
-\label{fig:compression}
-\end{center}
-\end{figure}
-
-Taking the logarithm of a magnitude spectrum is ubiquitous in audio signal processing.
-Indeed, it is corroborated by the Weber-Fechner law in psychoacoustics, which states that the sensation of loudness is roughly proportional to the logarithm of the acoustic pressure.
-We must also recall that the measured amplitude of sound sources often decays polynomially with the distance to the microphone--a source of spurious variability in scene classification.
-Logarithmic compression linearise this dependency, facilitating the construction of powerful invariants at the classifier stage.
-
-\subsection{Standardisation}
-\label{sec:stand}
-
-Let $\mathbf{S}\boldsymbol{x}(\gamma,n)$ be a dataset, where $\gamma$ and $n$ denote feature and sample indices, respectively.
-Many algorithms operate better on features which have zero mean and unit variance to avoid mismatch in numeric ranges~\cite{Hsu2003}.
-To standardise $\mathbf{S}\boldsymbol{x}(\gamma,n)$, we subtract the sample mean vector $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ from $\mathbf{S}\boldsymbol{x}(\gamma,n)$ and divide the result by the sample standard deviation vector $\sigma[\mathbf{S}\boldsymbol{x}] (\gamma)$.
-The vectors $\mu[\mathbf{S}\boldsymbol{x}(\gamma)]$ and $\sigma[\mathbf{S}\boldsymbol{x}](\gamma)$ are estimated from the entire dataset.
-
-\section{Acoustic scene similarity retrieval}
-\label{sec:object}
-
-%As part of the aforementioned research areas and applications, the emerging field of \emph{Acoustic Scene Analysis} (also called \emph{Sound Scene Analysis})~\cite{Stowell15} aims to develop approaches and systems for the automatic analysis of environmental sounds and soundscapes (originating both from urban or nature environments).
-While research methodologies in related fields such as Automatic Speech Recognition (ASR)~\cite{Rabiner93} and Music Information Retrieval (MIR)~\cite{Muller07} are now well established, research addressing Acoustic Scene Analysis remains relatively young.
-
-%From a data processing point of view, a holistic scheme, such as the bag-of-frame approach~\cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions~\cite{lagrange:hal-01082501}.
-
-As discussed in Section~\ref{sec:soa}, results in sound perception suggest the appropriateness of source-driven representations of auditory scenes for predicting high-level properties.
-While this can be addressed in the supervised case using late integration of discriminative classifiers~\cite{Anden2014}, this is not directly feasible in the unsupervised case.
-As the detection of events is still an open problem~\cite{7100934}, we consider in this paper a generic quantization scheme in order to identify and represent time intervals of the scene that are coherent, thus likely to be dominated by a given source of interest.
-
-Given a set of $d$-dimensional feature vectors $X_u = \{x_1^u, \ldots, x_L^u\}$, extracted from the scene $s_u$, where $u=\lbrace 1,2,\ldots,U\rbrace$, we would like to partition $X_u$ into a set $C_u = \{c^u_1, \ldots, c^u_M\}$ of $M$ clusters.
-This partition is obtained by minimizing the variance of each cluster and known as a $k$-means clustering~\cite{lloyd}.
-Each scene $s_u$ is then described by a set of clusters $C_u$.
-Note that this quantization approach differs from unsupervised learning schemes such as the ones studied in~\cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset.
-Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroids.
-
-The similarity between the scene centroids $\mu_m^u$ over the entire dataset, is computed using a radial basis function (RBF) kernel $K$ combined with a local scaling method~\cite{selfTuneManor2004}:
-\begin{equation}
-\label{eq:kc}
-K_{mn}^{uv} = \exp\left( - \dfrac{\Vert \mu_m^u - \mu_n^v \Vert^2}{\Vert \mu_m^u - \mu_{m,q}^u \Vert \Vert \mu_n^v - \mu_{n,q}^v \Vert} \right).
-\end{equation}
-Here, $\mu_{m,q}^u$ and $\mu_{n,q}^v$ are the $q^{\textrm{th}}$ nearest neighbors to the centroids $\mu_m^u$ and $\mu_n^v$, respectively, and the double bars $\Vert \cdot \Vert$ denotes the Euclidean norm.
-
-To compute the similarity between two scenes, we consider several centroid-based similarity metrics:
-\begin{itemize}
-\item Relevance-based Quantization closest similarity (\emph{RbQ-c}): the similarity between two scenes $s_u$ and $s_v$ is equal to the largest similarity between their centroids
-\begin{equation}
-	\max_{m,n} K_{mn}^{uv},
-\end{equation}
-\item Relevance-based Quantization average similarity  (\emph{RbQ-a}): the similarity between two scenes $s_u$ and $s_v$ is equal to the average of their centroid similarities
-\begin{equation}
-	\frac{1}{M^2} \sum_{m,n} K_{mn}^{uv}
-\end{equation}
-and,
-\item Relevance-based Quantization weighted similarity  (\emph{RbQ-w}): the similarity between two scenes is computed using a variant of the earth mover's distance applied to the set of centroids each weighted by the number of frames assigned to its cluster.
-\end{itemize}
-
-For \emph{RbQ-w}, each scene is represented by a signature
-$$
-p_u = \lbrace(\mu_1^u,w_1^u),(\mu_2^u,w_2^u),\ldots,(\mu_M^u,w_M^u)\rbrace,
-$$
-where each of the $M$ centroids $\mu_1^u, \ldots, \mu_M^u$ are paired with corresponding weights $w_1^u, \ldots, w_M^u$.
-The weight $w_m^u$ for the $m$th centroid $\mu_m^u$ is the number of frames belonging to a particular cluster.
-The similarity between scenes is then given by a cross-bin histogram distance known as the non-normalized earth mover's distance $\widehat{\EMD}$ introduced by~\cite{pele2008linear}.
-The $\widehat{\EMD}$ computes the distance between two histograms by finding the minimal cost for transforming one histogram into the other, where cost is measured by the number of transported histogram counts multiplied by a dissimilarity measure between the histogram bins.
-Here, that measure is given by $1-K_{mn}^{uv}$.
-
-\section{Experiments}
-\label{sec:experiments}
-
-To evaluate the representations introduced in the previous section, we apply it to the acoustic scene similarity retrieval task.
-Results demonstrate the improved performance of the relevance-based quantization of scattering coefficients compared to baseline methods using summary statistics of MFCCs.
-The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}}
-
-\subsection{Dataset}
-
-The experiments in this paper are carried out on the DCASE 2013 dataset~\cite{7100934}.
-Although the dataset was constructed for the task of acoustic scene classification, where the goal is to correctly assign the class of a given recording, we can use the same recordings and class labels for the task of acoustic scene similarity retrieval.
-The dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration.
-The dataset is evenly divided into $10$ acoustic scene classes: \emph{bus}, \emph{busy street}, \emph{office}, \emph{open air market}, \emph{park}, \emph{quiet street}, \emph{restaurant}, \emph{supermarket}, \emph{tube}, and \emph{tube station}.
-The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
-In order to avoid any correlation between recording conditions and label distribution, all recordings were carried out under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type.
-As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices~\cite{lagrange:hal-01082501}.
-As an illustration, Figure \ref{fig:scalogram} represents the wavelet scalogram of one recording within the DCASE 2013, labeled as \emph{park}.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/park_scalogram.png}
-\caption{
-\label{fig:scalogram}
-Wavelet scalogram $\boldsymbol{x}_1(t,\gamma_1)$ of the audio recording \emph{park04} in the DCASE 2013 dataset.
-We observe that this acoustic scene is a mixture of transient events (chirping birds, footsteps) and stationary texture (flowing water).}
-\end{center}
-\end{figure}
-
-\subsection{Feature design}
-
-We perform our experiments using both scattering coefficients and MFCCs.
-For the scattering transform, each $30$-second scene is described by $128$ vectors of dimension $1367$ computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$.
-Here we discard $3$ seconds from the beginning and end of the scene to avoid boundary artifacts.
-We also conduct experiments with and without logarithmic compression of the scattering coefficients (see Section~\ref{sec:logcomp}).
-
-MFCCs are computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$ with full frequency range.
-The standard configuration of $39$ coefficients coupled with an average-energy measure performs best in preliminary tests, so we use this in the following.
-We average the coefficients using $250\,\mathrm{ms}$ long non-overlapping windows so that each window represents structures of scales close to that of scattering coefficients.
-
-\subsection{Evaluation and algorithm}
-
-The evaluation is performed on the private part of the DCASE 2013 dataset.
-As a metric, we use the precision at rank $k$ ($p@k$).
-This number is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items.
-We determine these neighbors using one of the proposed similarity measures \emph{RbQ-c}, \emph{RbQ-a}, or \emph{RbQ-w}.
-We compute $p@k$ for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items.
-Note that $p@1$ is equal to the classification accuracy obtained by the nearest-neighbor classifier in a leave-one-out cross-validation setting.
-
-The \emph{RbQ} measures are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene.
-The distance on this average feature vector is then used to determine $p@k$.
-For the BoF approach of Aucouturier et al.~\cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm~\cite{dempster-laird, moon1996expectation}.
-The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation.
-To ensure convergence of the EM algorithm for the scattering features, we reduce their dimension from $1367$ to $30$ by projecting the features onto the top $30$ principal components of the dataset.
-the number of Gaussians is optimized for each type of features by grid search in the range $[2, 20]$.
-Best $p@5$ is reached with 8 and 4  Gaussians, respectively for MFCCs and scattering features.
-Recommended number of Gaussians for MFCCs given in~\cite{aucouturier2007bag} is 10.
-
-The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.
-As the number of Gaussians for the BoF approach, the numbers of clusters $M$ controls the level of abstraction.
-For each method, the parameter $M$ is set to $8$.
-It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds).
-
-\section{Results \label{sec:results}}
-
-Results for the acoustic scene similarity retrieval task demonstrate that logarithmically compressed scattering features outerform MFCCs.
-Combining these with the \emph{RbQ} cluster model, improvements are obtained over traditional BoF and summary statistic measures.
-
-\subsubsection*{Baselines}
-
-As seen in Figure~\ref{fig:ASS_-1}, scattering features significantly outperform the baseline MFCCs for both BoF and early integration schemes.
-This is expected, as the scattering transform extends the MFCCs by including supplementary amplitude modulation information~\cite{Anden2014}.
-We also note that applying PCA reduction to the scattering transform has little effect for the early integration scheme.
-In the context of summary statistics, $30$ dimensions are sufficient to discriminate between auditory scenes.
-
-Comparing the BoF to early integration, both approaches perform similarly for MFCCs and PCA-reduced scattering features alike.
-This is in line with previous results on BoF~\cite{lagrange:hal-01082501}, where it is found to perform similarly as similarity retrieval on features averaged over the entire recording.
-
-The early approach being simpler in terms of implementation and runtime complexity, we retain this method as baseline for the remainder of the experiments.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/baselines}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for several baseline approaches, as a function of the rank $k$.}
-\label{fig:ASS_-1}
-\end{center}
-\end{figure}
-
-\subsubsection*{Logarithmic compression}
-
-Figure~\ref{fig:ASS_0} shows that logarithmic compression of the scattering features is beneficial.
-For clarity sake, data is shown  for the \emph{early} approach only, but a equivalent gain is achieved for the relevance-based quantization approaches.
-%The gain is more pronounced when considering alternative similarities.
-\ja{What's ``alternative'' similarities?}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/log}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
-\label{fig:ASS_0}
-\end{center}
-\end{figure}
-
-\subsubsection*{MFCC \vs scattering transform}
-
-Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{RbQ-c} approach.
-Overall, log-compressed scattering coefficients systematically outperform MFCCs.
-This is to be expected since the scattering coefficients capture larger-scale modulations, as opposed to MFCCs which only describe the short-time spectral envelope.
-
-\subsubsection*{Relevance-based quantization \vs early integration}
-
-For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (\emph{RbQ}) to improve the similarity measures between the scenes.
-However, it is worth noting that \emph{RbQ-a} performs comparably to or worse than \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids.
-This result is in line with the findings of~\cite{lagrange:hal-01082501}.
-To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects.
-The same behavior is observed for MFCCs, with \emph{RbQ-c} and \emph{RbQ-w} outperforming \emph{early}, which is equivalent to the state-of-the-art BoF model, as seen previously.
-
-Furthermore, it appears that \emph{RbQ-c} is better able to characterise the classes compared to \emph{RbQ-w}.
-Although not the only way of incorporating the number of frames associated to each centroid, the earth mover's distance is a rather natural of doing so.
-Its worse performance therefore suggests that including this information may not always be desirable.
-Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames.
-On the contrary, two similar environments may share a lot of similar sound sources with only a few sources discriminating between them.
-
-With $p@5$ as our metric (cf.~\cite{aucouturier2007bag} and~\cite{lagrange:hal-01082501}), we see that replacing MFCCs by the logarithmically compressed scattering transform increases performance from $0.31$ to $0.49$.
-In addition, the relevance-based quantization using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/overall}
-\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression.}
-\label{fig:ASS_1}
-\end{center}
-\end{figure}
-
-\section{Conclusion}
-
-This paper presents a new approach for modeling acoustic scenes based on scattering transforms at small scales and cluster-based representations at large scales.
-Compared to traditional BoF and summary statistics models, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature.
-To adequately capture such distinct events, we develop a cluster-based model and validate it using experiments on acoustic scene similarity retrieval.
-For this task, we show significant improvements over the traditional BoF and summary statistics models based on both standard MFCCs and scattering features.
-These outcomes shall be studied further in future work by considering larger databases and emerging tasks in ecoacoustics~\cite{sueur2015ecoacoustics, wimmer2013sampling}.
-
-\section*{Availability of data and materials}
-The dataset supporting the conclusions of this article is available in the dcase2013 repository,  \sloppy \url{http://c4dm.eecs.qmul.ac.uk/sceneseventschallenge/description.html}.
-The software supporting the conclusions of this article is available in
-\begin{itemize}
-\item \sloppy Companion website: \url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}
-\item Programming language: Matlab
-\item License: GNU GPL
-\item Any restrictions to use by non-academics: license needed
-\end{itemize}
-
-\section*{Competing interests}
-The authors declare that they have no competing interests.
-
-\section*{Funding}
-This study is co-funded by the ANR under project reference ANR-16-CE22-0012.
-
-\section*{Authors' contributions}
-GL and VL carried out the numerical experiments and drafted the manuscript. VL, GL, JA and ML participated in the design of the study and helped to draft the manuscript.
-All authors read and approved the final manuscript.
-
-%\bibliographystyle{spmpsci}
-%\bibliography{biblio}
-\printbibliography
-
-% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/lostanlen}}]{Vincent Lostanlen} was born in 1992.
-He received an engineering degree from TELECOM ParisTech in 2013 and his M.S.
-degree in acoustics, signal processing, and musical informatics (ATIAM) from the Universit\'{e} Pierre et Marie Curie (UPMC) and the Ircam in Paris, France, in 2013.
-Since 2013, he is a Ph.D.
-student at \'{E}cole normale sup\'{e}rieure in Paris, working under the supervision of St\'{e}phane Mallat.
-His research focuses on defining convolutional operators in the time-frequency domain with applications in audio classification.
-% \end{IEEEbiography}
-%
-% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/lafay}}]{Gr\'egoire Lafay} was born in 1990.
-He received the B.S.
-degree in Acoustic from the University Pierre and Marie Curie (UPMC), Paris, FranArthritis National Research Foundatioce, and the B.S.
-degree in Musicology from the Sorbonne University, Paris, France, in 2011.
-He received his M.S.
-degree in acoustics, signal processing, and musical informatics (ATIAM) from the UPMC and the Ircam in Paris, France, in 2013.
-Since 2013, he is a Ph.D.
-student at IRCCyN, Nantes, France.
-His research interests include acoustic scene similarity and classification as well as acoustic scene perception.
-% \end{IEEEbiography}
-%
-%
-% \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{figures/anden}}]{Joakim And\'en}
-% received the M.Sc.
-degree in mathematics from the Universit\'e Pierre et Marie Curie, Paris, France, in 2010 and the Ph.D.
-degree in applied mathematics from Ecole Polytechnique, Palaiseau, France, in 2014.
-He is currently a postdoctoral researcher with the Program in Applied and Computational mathematics at Princeton University, Princeton, USA.
-His research interests include signal processing, machine learning, and statistical data analysis.
-He is a member of the IEEE.
-%
-% \end{IEEEbiography}
-%
-% \begin{IEEEbiography}[{\includegraphics[width=1.25in,height=1.25in,clip,keepaspectratio]{figures/lagrange}}]{Mathieu Lagrange} is a CNRS research scientist at IRCCyN, a French laboratory dedicated to cybernetics.
-He obtained his Ph.D.
-in computer science at the University of Bordeaux in 2004, and visited several institutions in Canada (University of Victoria, McGill University) and in France (Orange Labs, TELECOM ParisTech, Ircam).
-His research focuses on machine listening algorithms applied to the analysis of musical and environmental audio.
-% \end{IEEEbiography}
-
-% biography section
-%
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-%\begin{IEEEbiography}{Mathieu Lagrange}
-%Biography text here.
-%\end{IEEEbiography}
-
-% if you will not have a photo at all:
-%\begin{IEEEbiographynophoto}{John Doe}
-%Biography text here.
-%\end{IEEEbiographynophoto}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-%\begin{IEEEbiographynophoto}{Jane Doe}
-%Biography text here.
-%\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
 
@@
 % that's all folks
+\begin{thebibliography}{10}
+\providecommand{\url}[1]{{#1}}
+\providecommand{\urlprefix}{URL }
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
+  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
+  \urlstyle{rm}\Url}\fi
+
+\bibitem{Anden2014}
+And{\'e}n, J., Mallat, S.: Deep scattering spectrum.
+\newblock IEEE Transactions on Signal Processing \textbf{62}(16), 4114--4128
+  (2014)
+
+\bibitem{Cohen1995}
+Cohen, L.: Time-frequency analysis.
+\newblock Prentice Hall (1995)
+
+\bibitem{Flanagan1960}
+Flanagan, J.: Models for approximating basilar membrane displacement.
+\newblock Journal of the Acoustical Society of America \textbf{32}(7), 937--937
+  (1960)
+
+\bibitem{Grossmann1984}
+Grossmann, A., Morlet, J.: {Decomposition of Hardy functions into square
+  integrable wavelets of constant shape}.
+\newblock SIAM Journal on Mathematical Analysis \textbf{15}(4), 723--736 (1984)
+
+\bibitem{Lyon2010}
+Lyon, R., Katsiamis, A., Drakakis, E.: History and future of auditory filter
+  models.
+\newblock In: Proceedings of the IEEE International Symposium on Circuits and
+  Systems (ISCS) (2010)
+
+\bibitem{Patterson1976}
+Patterson, R.: Auditory filter shapes derived with noise stimuli.
+\newblock Journal of the Acoustical Society of America \textbf{59}(3), 640--654
+  (1976)
+
+\bibitem{Pressnitzer2005}
+Pressnitzer, D., Gnansia, D.: Real-time auditory models.
+\newblock In: Proceedings of the International Computer Music Conference (ICMC)
+  (2005)
+
+\bibitem{Venkitaraman2014}
+Venkitaraman, A., Adiga, A., Seelamantula, C.S.: {Auditory-motivated Gammatone
+  wavelet transform}.
+\newblock Signal Processing \textbf{94}, 608--619 (2014)
+
+\bibitem{Fastl2007}
+Zwicker, E., Fastl, H.: Psychoacoustics: Facts and models, vol.~22.
+\newblock Springer Science \& Business Media (2013)
+
+\end{thebibliography}
+
 \end{document}

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
\alpha -> \sigma and \omega_c -> \xi in appendix

@@
 \[
-\left(1+\dfrac{\Delta\omega}{\omega_{c}}\right)^{N/2}=r\times\left(1+\frac{\Delta\omega^{2}}{\alpha^{2}}\right).
+\left(1+\dfrac{\Delta\omega}{\xi}\right)^{N/2}=r\times\left(1+\frac{\Delta\omega^{2}}{\sigma^{2}}\right).
 \]
@@
 \]
-Now, let us express the parameter $\alpha$ as a function of some
+Now, let us express the parameter $\sigma$ as a function of some
 required bandwidth $B$ at some cutoff threshold $r$. After having
 raised the above to its square and rearranged the terms, we obtain
-another quadratic equation, yet of the variable $\mbox{\ensuremath{\alpha}}^{2}$:
+another quadratic equation, yet of the variable $\mbox{\ensuremath{\sigma}}^{2}$:
 \[
@@
 
-We conclude with the following closed form for $\alpha$:
+We conclude with the following closed form for $\sigma$:
 \[
-\alpha=K_{N}\times\sqrt{\dfrac{\sqrt{1+h_{N}\left(Q\right)^{2}}-1}{2}}\times\xi,
+\sigma=K_{N}\times\sqrt{\dfrac{\sqrt{1+h_{N}\left(Q\right)^{2}}-1}{2}}\times\xi,
 \]

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
Gammatone wavelets have center freq \xi, not 1
bugfix Fourier transform of Gammatone wavelet
state that \xi and \sigma are expressed in radians
1 -> \xi in text

@@
 Then, we relate the bandwidth parameter $\sigma$ to the choice of a quality factor Q.
+We express both $\xi$ and $\sigma$ in radians, so that converting this value to Hertz would incur a multiplication by $2\pi/f_s$ where $f_s$ is the sampling frequency of the audio signal at hand.
 
@@
 Gammatone auditory filter, originated in auditory physiology. The
-Gammatone auditory filter of dimensionless frequency $1$ is defined
+Gammatone auditory filter of center frequency $\xi$ is defined
 as a gamma distribution of order $N\in\mathbb{N}^{*}$ and bandwidth
@@
 \[
-t^{N-1}\exp(-2\pi\sigma t)\cos(2\pi t).
+t^{N-1}\exp(-2\pi\sigma t)\cos(2\pi \xi t).
 \]
@@
 \[
-\boldsymbol{\psi}(t)=\left(2\pi(\mathrm{i}-\sigma)t^{N-1}+(N-1)t^{N-2}\right)\exp(-2\pi\sigma t)\exp(2\pi\mathrm{i}t)
+\boldsymbol{\psi}(t)=\left(2\pi(\mathrm{i}-\sigma)t^{N-1}+(N-1)t^{N-2}\right)\exp(-2\pi\sigma t)\exp(2\pi\mathrm{i}\xi t)
 \]
@@
 \[
-\boldsymbol{\widehat{\psi}}(\omega)=\dfrac{\mathrm{i}\omega\times(N-1)!}{(\sigma+\mathrm{i}(\omega-\sigma))^{N}}
+\boldsymbol{\widehat{\psi}}(\omega)=\dfrac{\mathrm{i}\omega\times(N-1)!}{(\sigma+\mathrm{i}(\omega-\xi))^{N}}
 \]

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
1 -> \xi in text, bis

@@
 \cite{Venkitaraman2014} have modified the definition of the Gammatone
-auditory filter, by replacing the real-valued sine wave $\cos(2\pi t)$
-by its analytic part $\exp(2\pi\mathrm{\mathrm{i}}t)$ and by taking
+auditory filter, by replacing the real-valued sine wave $\cos(2\pi \xi t)$
+by its analytic part $\exp(2\pi\mathrm{\mathrm{i}}\xi t)$ and by taking
 the first derivative of the gamma distribution, thus ensuring null

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
define sampling frequency \xi_{\mathrm{s}}

@@
 Then, we relate the bandwidth parameter $\sigma$ to the choice of a quality factor Q.
-We express both $\xi$ and $\sigma$ in radians, so that converting this value to Hertz would incur a multiplication by $2\pi/f_s$ where $f_s$ is the sampling frequency of the audio signal at hand.
 
@@
 \]
-for all frequencies $\omega$, where $\varepsilon$ is a small margin
-\cite{Anden2014}. Satisfying the equation above near the Nyquist
-frequency $\omega=\pi$ can be achieved by placing the log-frequency
+for all frequencies $\omega$ in the audible range, where $\varepsilon$ is a small margin
+\cite{Anden2014}. 
+Let $\xi_{\mathrm{s}}$ be the sampling rate of the audio signal $\boldsymbol{x}(t)$ at hand.
+Satisfying the equation above near the Nyquist
+frequency $\omega=\frac{1}{2}\xi_{\mathrm{s}}$ can be achieved by placing the log-frequency
 $\log_{2}\xi$ of the first (topmost) wavelet in between the log-frequency
 $\log_{2}(\xi\times2^{-1/Q})$ of the second wavelet and the log-frequency
-$\log_{2}(2\pi-\xi)$ of the mirror of the first wavelet. We obtain
+$\log_{2}(\xi_{\mathrm{s}}-\xi)$ of the mirror of the first wavelet. We obtain
 the equation

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
proofread up to page 2

@@
 \[
-\log_{2}\xi-\log_{2}(\xi\times2^{-1/Q})=\log_{2}(2\pi-\xi)-\log_{2}\xi,
+\log_{2}\xi-\log_{2}(\xi\times2^{-1/Q})=\log_{2}(\tfrac{1}{2}\xi_{\mathrm{s}}-\xi)-\log_{2}\xi,
 \]
 
-of which we deduce the identity
+of which we deduce the normalized identity
 \[
-\xi=\dfrac{2\pi}{1+2^{1/Q}}.
+\dfrac{\xi}{\xi_\mathrm{s}}=\dfrac{1}{1+2^{1/Q}}.
 \]
-For $Q=1$, this yields a center frequency of $\xi=\frac{2\pi}{3}$.
-For greater values of $Q$, the center frequency $\xi$ tends towards
-$\pi$.
+For $Q=1$, this yields a center frequency of $\xi=\frac{1}{3}\xi_{\mathrm{s}}$.
+For greater values of $Q$, the center frequency $\xi$ tends towards the Nyquist frequency $\frac{1}{2}\xi_{\mathrm{s}}$.
 

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
proofread most of appendix

@@
 \[
-\dfrac{\vert\boldsymbol{\hat{\psi}}(\omega)\vert}{\vert\boldsymbol{\hat{\psi}}(\xi)\vert}=\dfrac{\omega}{\xi}\times\left(1+\dfrac{\left(\omega-\xi\right)^{2}}{\sigma^{2}}\right)^{-N/2}=r,
+\dfrac{\vert\boldsymbol{\hat{\psi}}(\omega)\vert}{\vert\boldsymbol{\hat{\psi}}(\xi)\vert}
+=
+\dfrac{\omega}{\xi}\times
+\left(1+\dfrac{\left(\omega-\xi\right)^{2}}{\sigma^{2}}\right)^{-N/2}=r,
 \]
-where the magnitude cutoff $r$ is most
-often set to $\sqrt{\frac{1}{2}}$. Let $\Delta\omega=\omega-\xi$.
-Raising the above equation to the power $N/2$ yields the following:
+where the magnitude cutoff $r<1$ is most
+often set to $\sqrt{\frac{1}{2}}$. Let $\mathrm{\Delta}\omega=\omega-\xi$.
+Since $\mathrm{\Delta}\omega\ll\xi$, we may linearize the exponentiation to $(-\tfrac{N}{2})$ by the first-order Taylor expansion
 \[
-\left(1+\dfrac{\Delta\omega}{\xi}\right)^{N/2}=r\times\left(1+\frac{\Delta\omega^{2}}{\sigma^{2}}\right).
+\left(1 + \frac{(\mathrm{\Delta}\omega)^2}{\sigma^2}\right)^{\frac{N}{2}} 
+\approx
+1 + \frac{N}{2} \left(\frac{\mathrm{\Delta}\omega}{\sigma}\right)^2.
 \]
-Since $\Delta\omega\ll\xi$, we may approximate the left-hand side
-with a first-order Taylor expansion. This leads to a quadratic equation
-of the variable $\Delta\negmedspace\omega$:
+Furthermore, we may rewrite the ratio $\frac{\omega}{\xi}$ as $\frac{\xi + \mathrm{\Delta}\omega}{\xi} = (1 + \frac{\mathrm{\Delta}{\omega}}{\xi})$.
+This leads to a quadratic equation
+of the variable $\mathrm{\Delta}\omega$:
 \[
-\dfrac{r^{2/N}}{\sigma^{2}}\times\Delta\omega^{2}-\dfrac{2}{N\xi}\times\Delta\omega+\left(r^{2/N}-1\right)=0.
+1 + \dfrac{\mathrm{\Delta}\omega}{\xi}
+=
+r
+\times
+\left(1 + \dfrac{N}{2} \left(\dfrac{\mathrm{\Delta}\omega}{\sigma}\right)^2 \right),
 \]
-The discriminant of the above equation is:
+which rewrites as
 \[
-D=4\times\left(\dfrac{1}{N^{2}\xi^{2}}+\dfrac{r^{2/N}\left(1-r^{2/N}\right)}{\sigma^{2}}\right),
+(\mathrm{\Delta}\omega)^2
+- \dfrac{2\sigma^2}{Nr\xi} (\mathrm{\Delta}\omega)
+- \dfrac{2\sigma^2}{N} \dfrac{1-r}{r} = 0.
 \]
-which is a positive number as long as $r<1$. The bandwidth $B$ of
-$\boldsymbol{\hat{\psi}}$ is given by the difference between the
-two solutions of the quadratic equation, that is:
+
+The discriminant of the above equation is
 \[
-B=\dfrac{2\sigma^{2}}{r^{2/N}}\times\sqrt{\dfrac{1}{N^{2}\xi^{2}}+\dfrac{r^{2/N}\left(1-r^{2/N}\right)}{\sigma^{2}}}.
+D=\dfrac{4\sigma^2}{Nr}
+\left(
+\dfrac{\sigma^2}{Nr\xi^2}
++
+2(1-r)
+\right),
+\]
+which is a positive number because $r<1$.
+The bandwidth $B$ of $\boldsymbol{\hat{\psi}}$ is given by the difference between the two solutions of the quadratic equation, that is:
+\[
+B = 
+\sqrt{D} =
+\dfrac{2\sigma}{\sqrt{N r}}
+\sqrt{\dfrac{\sigma^2}{N r \xi^2} + 2 (1-r)}
+.
 \]
@@
 \[
-\dfrac{4}{r^{4/N}N^{2}\xi^{2}}\sigma^{4}+\dfrac{4\times\left(1-r^{2/N}\right)}{r^{2/N}}\sigma^{2}-B^{2}=0
+\dfrac{4}{N^2 r^2 \xi^2} (\sigma^2)^2 + \dfrac{8 (1-r)}{N r} \sigma^2 - B^2 = 0.
 \]
-We multiply the equation by $\dfrac{r^{2/N}}{4\times\left(1-r^{2/N}\right)}\neq0:$
+We multiply the equation by $\tfrac{1}{8} N^2 r^2 \xi^2 \neq 0$, thus yielding
 \[
-\dfrac{1}{r^{2/N}\left(1-r^{2/N}\right)N^{2}\xi^{2}}\sigma^{4}+\sigma^{2}-\dfrac{r^{2/N}B^{2}}{4\times\left(1-r^{2/N}\right)}=0
+\dfrac{1}{2} (\sigma^2)^2
++
+N r (1-r) \xi^2 \sigma^2
+-
+\dfrac{N^2 r^2 \xi^2 B^2}{8}
+= 0.
 \]
-This leads to defining $\sigma^{2}$ as the unique positive root of
-the above polynomial:
+The discriminant of the above equation is equal to:
 \[
-\sigma^{2}=\dfrac{r^{2/N}\left(1-r^{2/N}\right)N^{2}\xi^{2}}{2}\times\left(\sqrt{1+\dfrac{B^{2}}{\left(1-r^{2/N}\right)^{2}N^{2}\xi^{2}}}-1\right).
+D^\prime = N^2 r^2 \xi^2 \left(\xi^2 (1-r)^2 + \tfrac{1}{4}B^2\right).
+\]
+Because this discriminant is strictly positive, the equation has two roots.
+Yet, only the smallest of the two satisfies $\sigma^2 < \xi^2$.
+Therefore, we have
+\[
+\sigma^{2}=
+N r (1-r)
+\left(
+\sqrt{1 + \left(\dfrac{B}{2\xi (1-r)}\right)^2}
+- 1\right) \xi^2 
+.
 \]
@@
 set $B$ to $B=\left(1-2^{-1/Q}\right)\times\xi$.
-
 We conclude with the following closed form for $\sigma$:
-\[
-\sigma=K_{N}\times\sqrt{\dfrac{\sqrt{1+h_{N}\left(Q\right)^{2}}-1}{2}}\times\xi,
-\]
-where
-\[
-K_{N}=r^{1/N}N\sqrt{1-r^{2/N}}\mbox{ and }h_{N}\left(Q\right)=\dfrac{1-2^{-1/Q}}{N\times\left(1-r^{2/N}\right)}.
-\]
 

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
proofread last equation of appendix

@@
 If the filter bank has to be approximately orthogonal, we typically
-set $B$ to $B=\left(1-2^{-1/Q}\right)\times\xi$.
-We conclude with the following closed form for $\sigma$:
+set $B$ to
+\[
+B = 
+\left(\sqrt[Q]{2} - \dfrac{1}{\sqrt[Q]{2}} \right)
+\times \xi.
+\]
+We conclude with the following normalized closed form for $\sigma$:
+\[
+\dfrac{\sigma}{\xi} = \sqrt{N r (1-r)}
+\left(
+\sqrt{
+1 +
+\dfrac{\left(\sqrt[Q]{2} - \dfrac{1}{\sqrt[Q]{2}} \right)^2}{
+4 (1-r)^2
+}}
+- 1\right).
+\]
 

### 2018-09-12 paper/proof/13636_2018_138_MOESM1.tex
incorporate joakim's remarks to supp material

@@
 
-\section{Supplementary material}
+\section*{Supplementary material}
 
@@
 In this appendix, we provide a rationale for choosing the topmost center frequency $\xi$ of a Gammatone wavelet filter bank in a discrete-time setting.
-Then, we relate the bandwidth parameter $\sigma$ to the choice of a quality factor Q.
+We therefore relate the bandwidth parameter $\sigma$ to the choice of a quality factor Q.
 
@@
 The complex-valued Gammatone wavelet is a modification of the real-valued
-Gammatone auditory filter, originated in auditory physiology. The
+Gammatone auditory filter, which originated in auditory physiology. The
 Gammatone auditory filter of center frequency $\xi$ is defined
@@
 \cite{Venkitaraman2014} have modified the definition of the Gammatone
-auditory filter, by replacing the real-valued sine wave $\cos(2\pi \xi t)$
+auditory filter by replacing the real-valued sine wave $\cos(2\pi \xi t)$
 by its analytic part $\exp(2\pi\mathrm{\mathrm{i}}\xi t)$ and by taking
@@
 \]
-
-of which we deduce the normalized identity
+from which we deduce the normalized identity
 \[
@@
 raised the above to its square and rearranged the terms, we obtain
-another quadratic equation, yet of the variable $\mbox{\ensuremath{\sigma}}^{2}$:
+another quadratic equation, but now of the variable $\mbox{\ensuremath{\sigma}}^{2}$:
 \[
